### 2018-11-08 R/transect-analysis.Rmd
Organize data, start analysis notebook.

---
title: "Upper Ouémé Forest Fungi -- Soil Transect Analysis"
author: "Brendan Furneaux"
date: '`r paste0("October 10, 2018; Last modified ", strftime(file.mtime(knitr::current_input()), "%B %d, %Y"))`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries.
```{r}
library(magrittr)
library(tidyverse)
library(readxl)
```

Define the required files.
```{r}
base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
data.dir <- file.path(base.dir, "data")
soils.dir <- file.path(data.dir, "soils")
soils.file <- file.path(soils.dir, "transectsoil_corrected.csv")

lab.dir <- file.path(data.dir, "lab_setup")
plate.file <- file.path(lab.dir, "Brendan soil.xlsx")

pacbio.dir <- file.path(data.dir, "PacBio")
pacbio.otu.file <- file.path(pacbio.dir, "OTU_table.txt")

iontorrent.dir <- file.path(data.dir, "IonTorrent")
iontorrent.otu.file <- file.path(iontorrent.dir, "OTU_,matrix.txt")
```

Load the soil chemistry data
```{r}
soils <- read_tsv(soils.file, locale = locale(decimal_mark = ",")) %>%
  extract("Echantillon", into = c("Plot", "Sample", "Date"),
          regex = "(V1P[123])T([0-9]+)-([0-9]{4}-[0-9]{1,2}-[0-9]{1,2})") %>%
  mutate_at("Date", as.Date)
```

Load the data for the sample placements in the plates
```{r}
plate <- read_xlsx(plate.file, "Concentration samples", skip = 1, col_names = TRUE) %>%
  rename(Comment = X__1, Well = row) %>%
  select (-X__2)

```

Load the IonTorrent OTU table
```{r}
ion.otu <- read_tsv(iontorrent.otu.file) %>%
  column_to_rownames("OTU") %>%
  as.matrix
```

```{r}
pb.otu <- read_tsv(pacbio.otu.file)

pb.otumatrix <- pb.otu %>%
  select(Group:H7) %>%
  column_to_rownames("Group") %>%
  as.matrix
```

### 2018-12-18 R/filter-check.Rmd
Working on new demultiplexing and dada2 clustering, use Makefile

---
title: "Filter check"
output: beamer_presentation
---

```{r setup}
source("install_packages.R")

library(magrittr)
library(tidyverse)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)

if (interactive()) {
  base.dir <- getwd() %>%
    str_extract(".*oueme-fungi-transect")
  data.dir <- file.path(base.dir, "data")
  lab.dir <- file.path(data.dir, "lab_setup")
  gits7.file <- file.path(lab.dir, "Hectors tag primer plates.xlsx")
  its1.lr5.file <- file.path(lab.dir, "Brendan soil2.xlsx")
  tag.files <- list.files(file.path(lab.dir, "tags"), ".+\\.fasta", full.names = TRUE)
  pacbio.dir <- file.path(data.dir, "PacBio")
  pacbio.fastq.files <- list.files(pacbio.dir, "\\.fastq\\.gz", full.names = TRUE)
  in.name <- "pb_500_001"
  in.fastq <- file.path(pacbio.dir, "pb_500_001.reads_of_insert.fastq.gz")
  in.fwd <- str_replace(in.fastq, fixed(".fastq.gz"), ".gits7.blast")
  in.rev <- str_replace(in.fastq, fixed(".fastq.gz"), ".primer.blast")
  out.fastq <- str_replace(in.fastq, fixed(".fastq.gz"), ".demux.fastq.gz")
  out.groups <- str_replace(in.fastq, fixed(".fastq.gz"), ".groups")
  platekey <- file.path(lab.dir, "gITS7_platekey.csv")
  plate <- "short_002"
}
```

```{r}
fastq <- readFastq(out.fastq)
qual <- tibble(
  eexp = 10^(-1 * as(fastq@quality, "matrix")/10) %>%
  rowSums(na.rm = TRUE),
  length = width(fastq),
  erate = eexp/length)

ggplot(qual, aes(x = erate)) +
  stat_ecdf(geom = "step") +
  scale_x_log10(name = "Expected number of errors per base") +
  scale_y_continuous(name = "Fraction passing")

ggplot(qual, aes(x = eexp)) +
  stat_ecdf(geom = "step") +
  scale_x_log10(name = "Expected number of errors") +
  scale_y_continuous(name = "Fraction passing")

ggplot(qual, aes(x = length)) +
  stat_ecdf(geom = "step") +
  scale_x_log10(name = "Length (bp)") +
  scale_y_continuous(name = "Fraction Greater than length")

```

```{r}
fastq@sread[width(fastq)>3000] %>% as.character
```

```{r}
library(dada2)
out2.fastq <- str_replace(out.fastq, fixed(".fastq"), ".trim.fastq")
filterAndTrim(out.fastq, out2.fastq, maxLen = 2999)
derep <- derepFastq(out2.fastq)

err <- learnErrors(derep, errorEstimationFunction = dada2:::PacBioErrfun, multithread = TRUE)
dd <- dada(derep, err = err, BAND_SIZE=32, multithread = TRUE)
```

### 2018-12-18 R/transect-analysis.Rmd
Working on new demultiplexing and dada2 clustering, use Makefile

@@
 date: '`r paste0("October 10, 2018; Last modified ", strftime(file.mtime(knitr::current_input()), "%B %d, %Y"))`'
-output: html_document
+output:
+  beamer_presentation:
+    keep_tex: TRUE
+  bookdown::pdf_document2:
+    keep_tex: TRUE
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-Load libraries.
-```{r}
+knitr::opts_chunk$set(echo = FALSE, fig.width = 5, fig.height = 5)
 library(magrittr)
@@
 library(readxl)
+library(rwantshue)
+colorscheme <- iwanthue(seed=12, force_init = TRUE)
+doctype <- knitr::opts_knit$get("rmarkdown.pandoc.to")
+
+if (! is.null(doctype) && doctype == "beamer") {
+  knitr::opts_chunk$set(message = FALSE, warning = FALSE, out.width = "100%", fig.height = 4, fig.width = 8)
+}
 ```
 
-Define the required files.
-```{r}
+# Doctype
+
+Doctype is `r doctype`.
+
+```{r include = FALSE}
 base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
@@
 lab.dir <- file.path(data.dir, "lab_setup")
-plate.file <- file.path(lab.dir, "Brendan soil.xlsx")
+plate.file <- file.path(lab.dir, "Brendan_soil.xlsx")
+tags.file <- file.path(lab.dir, "Hectors_tag_primer_plates.xlsx")
 
 pacbio.dir <- file.path(data.dir, "PacBio")
-pacbio.otu.file <- file.path(pacbio.dir, "OTU_table.txt")
+pacbio.otu.file <- file.path(pacbio.dir, "PACBIO_OTU_table.xlsx")
+pacbio.funguild.file <- file.path(pacbio.dir, "PacBio_FUNGuild_in.guilds.web.txt")
+pacbio.group.files <- list.files(pacbio.dir, "*.groups")
 
 iontorrent.dir <- file.path(data.dir, "IonTorrent")
-iontorrent.otu.file <- file.path(iontorrent.dir, "OTU_,matrix.txt")
+iontorrent.otu.file <- file.path(iontorrent.dir, "higherLvl", "OTU.txt")
+iontorrent.classification.file <- file.path(iontorrent.dir, "hiera_BLAST.txt")
+iontorrent.funguild.file <- file.path(iontorrent.dir, "IonTorrent_FUNGuild.guilds.txt")
+iontorrent.repseq.file <- file.path(iontorrent.dir, "Otus_rep_seqs.fa")
 ```
 
-Load the soil chemistry data
 ```{r}
+# Load the soil chemistry data
 soils <- read_tsv(soils.file, locale = locale(decimal_mark = ",")) %>%
@@
 
-Load the data for the sample placements in the plates
 ```{r}
+# Load the data for the sample placements in the plates
+
+true_tags <- read_xlsx(plate.file, "Taggar fITS7", skip = 1) %>%
+  map(list(1:3, 5:7, 9:11), ~ select(.y, !!.x), .) %>%
+  map_dfr(set_names, c("Well", "No", "Tag"))
+
+wrong_tags <- read_xlsx(tags.file, skip = 1) %>%
+  select(Well = LOCATION, oligoname) %>%
+  mutate(Tag = str_replace(oligoname, "gITS7mod-", "") %>% as.numeric)
+
+well_key <- wrong_tags %>%
+  left_join(select(true_tags, Tag, True_Well = Well))
+
 plate <- read_xlsx(plate.file, "Concentration samples", skip = 1, col_names = TRUE) %>%
   rename(Comment = X__1, Well = row) %>%
-  select (-X__2)
+  select (-X__2) %>%
+  # Sample number consists of site, location along transect, and possible qualifier
+  tidyr::extract("Sample", c("Site", "X", "Qual"), "(V1P[13]|T[12])[ -][ST](\\d+)(\\D*)$") %>%
+  mutate_at("X", as.integer) %>%
+  # Sites were recorded differently in the two years
+  mutate(Site = case_when(Site == "V1P1" ~ "Ang",
+                          Site == "V1P3" ~ "Gan",
+                          Site == "T1" ~ "Gan",
+                          Site == "T2" ~ "Ang",
+                          TRUE ~ NA_character_)) %>%
+  # For some site/years, coordinates were measured from the edge of the plot rather than numbered from 1--25 
+  group_by(Year, Site) %>%
+  mutate(X = X - if (!any(is.na(X)) && max(X) > 25) 12 else 0) %>%
+  ungroup %>%
+  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)", remove = FALSE) %>%
+  mutate_at("Row", factor, levels = LETTERS) %>%
+  mutate_at(c("Column", "Row"), as.integer) %>%
+  mutate(sample_type = factor(Comment, levels = c("Blank", "Pos", "Pos. Kontroll")) %>%
+           fct_collapse(Blank = "Blank", Pos = c("Pos", "Pos. Kontroll")) %>%
+           fct_explicit_na("Sample"))
+
+plate %>% group_by(Year, Site) %>%
+  summarize_at("X", c("min", "max"))
+```
+
+```{r}
+# Load the IonTorrent OTU table
+ion.otu <- read_tsv(iontorrent.otu.file) %>%
+  mutate(n_reads = rowSums(select_if(., is_numeric))) %>%
+  left_join(read_tsv(iontorrent.classification.file)) %>%
+  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
+  left_join(read_tsv(iontorrent.funguild.file) %>% rename(OTU = Group)) %>%
+  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
+  rename(Kingdom = Domain) %>%
+  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "?") %>%
+  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "unidentified") %>%
+
+  # make "unident." as informative as possible
+  dplyr::mutate_at("Kingdom", replace_na, "unident.") %>%
+  dplyr::mutate(Phylum = ifelse(is.na(Phylum),
+                                ifelse(is.na(Kingdom),
+                                       NA_character_,
+                                       paste0("unident. ", Kingdom)),
+                                Phylum),
+                Class = ifelse(is.na(Class),
+                                ifelse(is.na(Phylum),
+                                       NA_character_,
+                                       paste0("unident. ", Phylum)),
+                                Class),
+                Order = ifelse(is.na(Order),
+                                ifelse(is.na(Class),
+                                       NA_character_,
+                                       paste0("unident. ", Class)),
+                                Order),
+                Family = ifelse(is.na(Family),
+                                ifelse(is.na(Order),
+                                       NA_character_,
+                                       paste0("unident. ", Order)),
+                                Family),
+                Genus = ifelse(is.na(Genus),
+                                ifelse(is.na(Family),
+                                       NA_character_,
+                                       paste0("unident. ", Family)),
+                                Genus),
+                Species = ifelse(is.na(Species),
+                                ifelse(is.na(Genus),
+                                       NA_character_,
+                                       paste0(Genus, " sp.")),
+                                Species)) %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            str_replace,
+            "(unident\\. )*unident(ified|\\.)",
+            "ZZZZunident.") %>%
+
+  # Sort taxonomically. Put the "unident." at the end
+  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            str_replace,
+            "ZZZZ",
+            "") %>%
+
+  # Add the representative sequence for each OTU and calculate the length
+  left_join(
+    seqinr::read.fasta(file = iontorrent.repseq.file, as.string = TRUE) %>%
+    { tibble(OTU = names(.),
+             RepSeq = unlist(.),
+             SeqLen = str_length(RepSeq)) }) %>%
+
+  # Remove the positive control
+  filter(is.na(Species) | Species != "Agaricus bisporus") %>%
+
+  # Keep the data in this order.
+  mutate_at(dplyr::vars(Kingdom:Species),
+            ~ factor(.) %>% fct_inorder)
+
+ion.platemap <- read_tsv(iontorrent.otu.file) %>%
+  column_to_rownames("OTU") %>%
+  as.matrix %>%
+  colSums %>%
+  tibble(Well = names(.), n_reads = .) %>%
+  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE) %>%
+  full_join(plate %>%
+              select(Plate, Well, sample_type) %>%
+              full_join(
+                expand.grid(Plate = 1:2,
+                            Row = LETTERS[1:8],
+                            Column = 1:12) %>%
+                  tidyr::unite(Well, Row, Column, sep = ""))) %>%
+  mutate(Plate = "IonTorrent") %>%
+  mutate_at("sample_type", fct_explicit_na, "None") %>%
+  group_by_if(~ !is.factor(.)) %>%
+  summarize(sample_type = paste(as.character(sort(unique(sample_type))), collapse = "/")) %>%
+  ungroup %>%
+  mutate(f_reads = n_reads / max(n_reads)) %>%
+  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
+  mutate_at("Row", factor, levels = LETTERS) %>%
+  mutate_at(c("Column", "Row"), as.integer)
+```
+
+# Demultiplexing
+
+\tiny
+
+We used a forward primer which included the forward IonTorrent sequencing adapter and an index, which distinguishes the well the sample was place in in the plate.
+
+\center
+
+```{tikz out.width = "60%"}
+\begin{tikzpicture}[line width=7]
+
+\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
+\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
+\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
+\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
+\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
+\draw[black] (start) -- ++(1,0) coordinate (start);}
+\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}
+\newcommand{\plate}{\draw[orange!40!yellow] (start) -- node[midway, above]{plate index}  ++(1, 0) coordinate (start);}
+
+\newcommand{\radapt}{\draw[pink] (start) -- node[midway, below]{rev adapter}  ++(3, 0) coordinate (start);}
+
+\newcommand{\fprimer}{\coordinate (pstart) at (start);
+\fadapt
+\well
+\gITS
+\coordinate (end) at (start);
+\path (end) ++(0,0.75) coordinate (end);
+\draw[line width = 2, dashed, |-|] (pstart) ++(0, 0.75) -- node[midway, above]{fwd primer}  (end);}
+
+\newcommand{\rprimer}{\coordinate (pstart) at (start);
+\ITS
+\coordinate (end) at (start);
+\path (end) ++(0,0.751) coordinate (end);
+\draw[line width = 2, dashed, |-|] (pstart) ++(0,0.75) -- node[midway, above]{rev primer}  (end);}
+
+\coordinate (start) at (0,0);
+\fprimer
+\seq
+\rprimer
+\end{tikzpicture}
+```
+
+# Demultiplexing
+
+\tiny
+
+Because we had more than 96 samples, we used two plates.  The same indexes were used for both plates.
+
+During IonTorrent library prep, sequencing adapters were ligated on to the amplicons.
+The forward adapters included index barcodes to distinguish the two plates.
+The orientation of the adapters was presumably random with respect to our amplicons.
+
+\center
+
+```{tikz out.width = "60%"}
+\begin{tikzpicture}[line width=7]
+
+\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
+\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
+\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
+\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
+\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
+\draw[black] (start) -- ++(1,0) coordinate (start);}
+\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}
+\newcommand{\plate}{\draw[orange!40!yellow] (start) -- node[midway, above]{plate index}  ++(1, 0) coordinate (start);}
+
+\newcommand{\radapt}{\draw[pink] (start) -- node[midway, below]{rev adapter}  ++(3, 0) coordinate (start);}
+
+\newcommand{\fprimer}{\coordinate (pstart) at (start);
+\fadapt
+\well
+\gITS
+\coordinate (end) at (start);
+\path (end) ++(0,0.75) coordinate (end);
+\draw[line width = 2, dashed, |-|] (pstart) ++(0, 0.75) -- node[midway, above]{fwd primer}  (end);}
+
+\newcommand{\rprimer}{\coordinate (pstart) at (start);
+\ITS
+\coordinate (end) at (start);
+\path (end) ++(0,0.751) coordinate (end);
+\draw[line width = 2, dashed, |-|] (pstart) ++(0,0.75) -- node[midway, above]{rev primer}  (end);}
+
+\coordinate (start) at (0,0);
+\fprimer
+\seq
+\rprimer
+
+\coordinate (start) at (-4, -4);
+\fadapt\plate\fprimer\seq\rprimer\radapt
+
+\coordinate (start) at (-3, -6.5);
+\radapt\fprimer\seq\rprimer\plate\fadapt
+\end{tikzpicture}
+```
+
+# Demultiplexing
+
+\tiny
+
+Prior to sequencing, IonTorrent uses surface-bound amplification to cover beads with many copies of the same sequence.  The forward adapter binds to the surface.
+
+Our samples all contained two copies of the forward adapter.
+Which one was used in each amplification step is random, but over the course of the entire amplification, the ones which result in shorter amplicons should be favored.
+
+\center
+
+```{tikz out.width = "60%"}
+\begin{tikzpicture}[line width=7]
+
+\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
+\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
+\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
+\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
+\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
+\draw[black] (start) -- ++(1,0) coordinate (start);}
+\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}
+\newcommand{\plate}{\draw[orange!40!yellow] (start) -- node[midway, above]{plate index}  ++(1, 0) coordinate (start);}
+
+\newcommand{\radapt}{\draw[pink] (start) -- node[midway, below]{rev adapter}  ++(3, 0) coordinate (start);}
+
+\newcommand{\fprimer}{\coordinate (pstart) at (start);
+\fadapt
+\well
+\gITS
+\coordinate (end) at (start);
+\path (end) ++(0,0.75) coordinate (end);
+\draw[line width = 2, dashed, |-|] (pstart) ++(0, 0.75) -- node[midway, above]{fwd primer}  (end);}
+
+\newcommand{\rprimer}{\coordinate (pstart) at (start);
+\ITS
+\coordinate (end) at (start);
+\path (end) ++(0,0.751) coordinate (end);
+\draw[line width = 2, dashed, |-|] (pstart) ++(0,0.75) -- node[midway, above]{rev primer}  (end);}
+
+\coordinate (start) at (0,0);
+\fprimer
+\seq
+\rprimer
+
+\coordinate (start) at (-4, -4);
+\fadapt\plate\fprimer\seq\rprimer\radapt
+
+\coordinate (start) at (-3, -6.5);
+\radapt\fprimer\seq\rprimer\plate\fadapt
+
+\coordinate (start) at (0, -10.5);
+\fprimer\seq\rprimer\radapt
+
+\coordinate (start) at (0, -13);
+\fprimer\seq\rprimer\plate\fadapt
+\end{tikzpicture}
+```
+
+# Plates
+```{r platemap}
+pacbio.group.files %>%
+  file.path(pacbio.dir, .) %>%
+  map_dfr(read_csv) %>%
+  group_by(group) %>%
+  summarize(n_reads = n()) %>%
+  ungroup %>%
+  tidyr::extract(group, c("Plate", "Well"), "(\\w+_\\d+)_([A-H]\\d+)") %>%
+  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
+  mutate_at("Row", factor, levels = LETTERS) %>%
+  mutate_at(c("Column", "Row"), as.integer) %>%
+  group_by(Plate) %>%
+  mutate(f_reads = n_reads / max(n_reads)) %>%
+  ggplot(aes(Column, Row, size = f_reads)) + 
+  facet_wrap(facets = "Plate", ncol = 3) + 
+  # geom_raster(aes(fill = sample_type)) +
+  geom_point(aes(alpha = ifelse(f_reads > 0, 1, 0))) +
+  # scale_fill_discrete(name = "Type") +
+  scale_size(name = "Normalized\nReads", range = c(0.1, 3), trans = "log1p") +
+  scale_alpha(guide = FALSE) +
+  ggplot2::coord_equal() +
+  scale_y_reverse(breaks = 1:8, labels = LETTERS[1:8]) +
+  scale_x_continuous(breaks = 1:12, limits = c(0.5, 12.5)) +
+  # scale_fill_brewer(name = "", type = "qual", palette = 3) +
+  theme(legend.box = "vertical", legend.position = "bottom", legend.key.size = unit(4, "mm"))
+
+pb.otu <- read_excel(pacbio.otu.file)
+
+pb.otumatrix <- pb.otu %>%
+  column_to_rownames("Group") %>%
+  select(A7_483_1:H7_483_1) %>%
+  as.matrix %>%
+  t %>%
+  as.tibble(rownames = "Sample") %>%
+  tidyr::extract("Sample", c("Well", "Plate"), regex = "([A-H]1?[0-9])_483_([12])", remove = FALSE) %>%
+  mutate_at("Plate", as.integer) %>%
+  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE)
+
+pb.otumatrix %>%
+  full_join(plate %>% select(Plate, Well, sample_type)) %>%
+  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
+  mutate_at("Row", factor, levels = LETTERS) %>%
+  mutate_at(c("Column", "Row"), as.integer) %>%
+  mutate_at("Plate", paste, "PacBio") %>%
+  mutate(n_reads = rowSums(select(., starts_with("Otu")))) %>%
+  full_join(expand(., Row, Column, Plate)) %>%
+  mutate(n_reads = ifelse(is.na(n_reads), 0, n_reads)) %>%
+  group_by(Plate) %>%
+  mutate(f_reads = n_reads / max(n_reads),
+         sample_type = ifelse(is.na(sample_type), "None", as.character(sample_type))) %>%
+  ungroup %>%
+  select(Row, Column, Plate, n_reads, f_reads, sample_type) %>%
+  bind_rows(ion.platemap) %>%
+  mutate_at("sample_type", factor, levels = c("Sample", "Pos", "Blank", "Pos/Sample", "Blank/Sample", "Sample/None", "Pos/None", "Blank/None", "None")) %>%
+
+  ggplot(aes(Column, Row, size = f_reads)) + 
+  facet_wrap(facets = "Plate", ncol = 3) + 
+  geom_raster(aes(fill = sample_type)) +
+  geom_point(aes(alpha = ifelse(f_reads > 0, 1, 0))) +
+  scale_fill_discrete(name = "Type") +
+  scale_size(name = "Normalized\nReads", range = c(0.1, 3), trans = "log1p") +
+  scale_alpha(guide = FALSE) +
+  ggplot2::coord_equal() +
+  scale_y_reverse(breaks = 1:8, labels = LETTERS[1:8]) +
+  scale_x_continuous(breaks = 1:12, limits = c(0.5, 12.5)) +
+  scale_fill_brewer(name = "", type = "qual", palette = 3) +
+  theme(legend.box = "vertical", legend.position = "bottom", legend.key.size = unit(4, "mm"))
+```
+
+```{r}
+# Parse out the identification and classification
+pb.otu %<>% select(-(A1_483_1:`ITS1-4_500_1`)) %>%
+  #parse the name
+  dplyr::bind_cols(
+    stringr::str_split(.$X__2, stringr::fixed("|"), n = 5, simplify = TRUE) %>%
+      tibble::as_data_frame() %>%
+      set_names(c("Subject.Name", "Accession", "Species.Hypothesis",
+                  "Ref.Type", "Classification"))) %>%
+  tidyr::extract("Classification",
+                 c("Kingdom", "Phylum", "Class", "Order",
+                                 "Family", "Genus", "Species"),
+                 "k__(.+)_p__(.+)_c__(.+)_o__(.+)_f__(.+)_g__(.+)_s__(.+)") %>%
+  dplyr::select(-X__2) %>%
+
+  # calculate the length of each sequence
+  dplyr::mutate(SeqLen = str_length(RepSeq)) %>%
+
+  #change "unident." to NA
+  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "unident.") %>%
+  # make "unident." as informative as possible
+  dplyr::mutate_at("Kingdom", replace_na, "unident.") %>%
+  dplyr::mutate(Phylum = ifelse(is.na(Phylum),
+                                ifelse(is.na(Kingdom),
+                                       NA_character_,
+                                       paste0("unident. ", Kingdom)),
+                                Phylum),
+                Class = ifelse(is.na(Class),
+                                ifelse(is.na(Phylum),
+                                       NA_character_,
+                                       paste0("unident. ", Phylum)),
+                                Class),
+                Order = ifelse(is.na(Order),
+                                ifelse(is.na(Class),
+                                       NA_character_,
+                                       paste0("unident. ", Class)),
+                                Order),
+                Family = ifelse(is.na(Family),
+                                ifelse(is.na(Order),
+                                       NA_character_,
+                                       paste0("unident. ", Order)),
+                                Family),
+                Genus = ifelse(is.na(Genus),
+                                ifelse(is.na(Family),
+                                       NA_character_,
+                                       paste0("unident. ", Family)),
+                                Genus),
+                Species = ifelse(is.na(Species),
+                                ifelse(is.na(Genus),
+                                       NA_character_,
+                                       paste0(Genus, " sp.")),
+                                Species)) %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            str_replace,
+            "(unident\\. )*unident(ified|\\.)",
+            "ZZZZunident.") %>%
+
+  # Sort taxonomically. Put the "unident." at the end
+  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            str_replace,
+            "ZZZZ",
+            "") %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            ~ factor(.) %>% fct_inorder) %>%
+
+  # Remove the positive control
+  filter(is.na(Species) | Species != "Agaricus_bisporus") %>%
+
+  # Add guild assignments
+  left_join(read_tsv(pacbio.funguild.file))
 
@@
 
-Load the IonTorrent OTU table
 ```{r}
-ion.otu <- read_tsv(iontorrent.otu.file) %>%
-  column_to_rownames("OTU") %>%
-  as.matrix
+# Parse the total sequence count
+pb.otu %<>%
+  rename(n_reads = `OTU_ID=seqs_no`) %>%
+  mutate_at("n_reads", str_replace, "Otu\\d+=", "") %>%
+  mutate_at("n_reads", as.integer)
 ```
@@
 ```{r}
-pb.otu <- read_tsv(pacbio.otu.file)
+all.otu <- bind_rows(
+  mutate(pb.otu, tech = "PacBio"),
+  mutate(ion.otu, tech = "IonTorrent")
+) %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            str_replace,
+            "(unident\\. )*unident(ified|\\.)",
+            "ZZZZunident.") %>%
+  # Sort taxonomically. Put the "unident." at the end
+  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            str_replace,
+            "ZZZZ",
+            "") %>%
+  mutate_at(dplyr::vars(Kingdom:Species),
+            ~ factor(.) %>% fct_inorder)
+```
 
-pb.otumatrix <- pb.otu %>%
-  select(Group:H7) %>%
-  column_to_rownames("Group") %>%
-  as.matrix
+```{r}
+rand_palette <- partial(colorscheme$hex,
+                        color_space = hcl_presets$colorblind_friendly,
+                        force_mode = FALSE)
 ```
+
+```{r}
+summarize_taxonomy <- function(data, band_rank, group_rank, ...,
+                               band_prop = 0.02, group_prop = 0.05, min_reads = 2,
+                               print.table = if (is.null(get0("doctype"))) TRUE else
+                                 if (doctype == "beamer") FALSE else
+                                   TRUE) {
+  band_rank <- ensym(band_rank)
+  group_rank <- ensym(group_rank)
+  classification <- tidyselect::vars_select(names(data), Kingdom:!!band_rank)
+  data <- filter(data, Kingdom != "", n_reads >= min_reads, ...)
+  data <- mutate_at(data, classification, fct_drop)
+
+  # count the reads and OTUs for taxa at each rank
+
+  if ("tech" %in% names(data)) data <- group_by(data, tech)
+  for(rank in classification) {
+    data <- group_by_at(data, rank, .add = TRUE)
+    data <-
+      mutate(data,
+             !!(glue::glue("{rank}.otus")) := n(),
+             !!(glue::glue("{rank}.reads")) := sum(n_reads))
+  }
+  class_counts <- glue::glue("{classification}.reads")
+  data <- group_by_at(data, class_counts, .add = TRUE)
+  data <- summarize(data,
+                    `# OTUs` = n(),
+                    `# reads` = sum(n_reads))
+  while(is_grouped_df(data)) data <- ungroup(data)
+
+  # Order each rank by number of reads, with most reads first.
+  data <- arrange_at(data, class_counts, .funs = desc)
+  data <- mutate_at(data, classification, fct_inorder)
+
+  # Print the data
+  if (print.table)
+    print(knitr::kable(select(data, one_of(classification), `# OTUs`, `# reads`),
+                       booktabs = TRUE))
+
+  # Group taxa with a small fraction of the total OTUs and reads (on all techs) into "Other"
+  if ("tech" %in% names(data)) data <- group_by(data, tech)
+
+  data <- mutate(data,
+                 total_OTUs = sum(`# OTUs`),
+                 total_reads = sum(`# reads`))
+  data <- group_by(data, !!group_rank, add = TRUE)
+  data <- mutate(data,
+                 group_OTUs = sum(`# OTUs`),
+                 group_reads = sum(`# reads`))
+  data <- group_by(data, !!band_rank, add = TRUE)
+  data <- mutate(data,
+                 band_OTUs = sum(`# OTUs`),
+                 band_reads = sum(`# reads`))
+  while(is_grouped_df(data)) data <- ungroup(data)
+
+  data <- mutate_at(data, deparse(group_rank), as.character)
+  data <- group_by(data, !!group_rank)
+  data <- mutate(data,
+                 group_other = !any(group_OTUs >= group_prop * total_OTUs,
+                                     group_reads >= group_prop * total_reads))
+  data <- ungroup(data)
+  data <- mutate(data, !!group_rank := ifelse(group_other, "other", !!group_rank),
+                 !!band_rank := as.character(!!band_rank))
+
+  data <- group_by_at(data, classification)
+  data <- mutate(data,
+                 band_other = !any(band_OTUs >= band_prop * total_OTUs,
+                                   band_reads >= band_prop * total_reads))
+  data <- ungroup(data)
+  data <- mutate(data, !!band_rank := ifelse(band_other, paste("other", !!group_rank), !!band_rank))
+  data <- mutate_at(data, deparse(band_rank), str_replace, "other( other)+", "other")
+
+  # data <- mutate_at(data, deparse(band_rank), fct_lump, w = data$`# OTUs`,
+  #                   prop = band_prop, other_level = "other")
+  # data <- mutate_at(data, deparse(group_rank), fct_lump, w = data$`# OTUs`,
+  #                   prop = group_prop, other_level = "other")
+  data <- mutate_at(data, classification, . %>% factor %>% fct_inorder)
+
+  pal <- rand_palette(n = n_distinct(data[[deparse(band_rank)]]))
+  data <- tidyr::gather(data, key = "count", value = "value", "# OTUs", "# reads")
+
+  p1 <- ggplot(data, aes(x = !!group_rank, y = value, fill = !!band_rank)) +
+    facet_wrap(facets = vars(count, tech), nrow = 2, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
+    geom_col(position = "stack") +
+    scale_x_discrete(deparse(group_rank), expand = c(0, 0)) +
+    scale_fill_manual(values = pal) +
+    theme(axis.text.x = element_text(angle = -20, vjust = 1, hjust = 0),
+          legend.position = "bottom", legend.text = element_text(size = 8),
+          legend.key.height = unit(5, "mm"))
+}
+```
+
+# High level taxonomic ID
+```{r, results='asis', fig.cap="OTU richness and read count for Phyla and Kingdoms, based on ITS2 metabarcoding"}
+#debugonce(summarize_taxonomy)
+summarize_taxonomy(all.otu, Phylum, Kingdom) + theme(legend.position = "right")
+```
+
+# Class-level Taxon ID
+```{r, results = 'asis', fig.cap="OTU richness and read count for Classes and Phyla, based on ITS2 metabarcoding"}
+summarize_taxonomy(all.otu, Class, Phylum) + theme(legend.position = "right")
+```
+
+# Order-level taxon ID
+
+```{r, results = 'asis', fig.cap="OTU richness and read count for Orders and Phyla, based on ITS2 metabarcoding"}
+summarize_taxonomy(all.otu, Order, Phylum) + theme(legend.position = "right")
+```
+
+# Genus-level taxon ID: Basidiomycota
+
+```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera in Basidiomycota, based on ITS2 metabarcoding"}
+summarize_taxonomy(all.otu, Genus, Order, Phylum == "Basidiomycota") + theme(legend.position = "right")
+```
+
+# Genus-level taxon ID: ECM
+
+```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera, probable ECM species only, based on FUNGuild assignment of OTUs from ITS2 metabarcoding"}
+summarize_taxonomy(all.otu, Genus, Order, str_detect(Guild, "Ectomycorrhizal")) + theme(legend.position = "right")
+```
+
+# Guild assignment
+```{r}
+all.otu %>%
+  rename(`# reads` = n_reads) %>%
+  mutate(`# OTUs` = 1) %>%
+  mutate_at("Guild", str_replace, "^-$", "unknown") %>%
+  mutate(Guild = fct_lump(Guild, prop = 0.02, w = `# reads`)) %>%
+  tidyr::gather(key = "count", value = "value", "# OTUs", "# reads") %>%
+  ggplot(aes(x = Guild, y = value, fill = Guild)) +
+    facet_wrap(facets = vars(count, tech), nrow = 2, scales = "free_y") +
+    geom_col(position = "stack") +
+    scale_x_discrete("Guild", expand = c(0, 0)) +
+    scale_fill_discrete( guide = "none") +
+    theme(axis.text.x = element_text(angle = -20, vjust = 1, hjust = 0), legend.position = "bottom")
+```
+
+# Genus-level ID: Saprotrophs
+```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera in Basidiomycota, based on short-read PacBio data"}
+summarize_taxonomy(all.otu, Genus, Order, str_detect(Guild, "Undefined Saprotroph")) + theme(legend.position = "right")
+```
+
+# Sequence length distribution
+
+```{r fig.width = 5, fig.height = 3}
+ ggplot(all.otu, aes(SeqLen, ..density.., weight = n_reads, group = tech, color = tech)) +
+  geom_freqpoly() +
+  # facet_wrap(facets = vars(tech)) +
+  scale_y_log10(name = "fraction of reads (log scale)") +
+  scale_x_continuous(name = "Sequence length") +
+  guides(color = guide_legend(direction = "horizontal", title = NULL)) +
+  theme(legend.position = "bottom")
+```
+
+# Taxon ID: short PacBio sequences
+```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera, representative sequence < 100bp"}
+summarize_taxonomy(all.otu, Genus, Order, SeqLen < 100) + theme(legend.position = "right")
+```
+
+# Taxon ID: long PacBio sequences
+\tiny
+```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera, probable ECM species only, based on FUNGuild assignment of OTUs from ITS2 metabarcoding"}
+all.otu %>% filter(SeqLen > 350) %>%
+  select(Group, n_reads, SeqLen, Order:Species) %>%
+  knitr::kable(booktabs = TRUE)
+```

### 2018-12-18 writing/demultiplex_issues.tex
Working on new demultiplexing and dada2 clustering, use Makefile

\begin{tikzpicture}[line width=5]

\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
\draw[black] (start) -- ++(1,0) coordinate (start);}
\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}

\coordinate (start) at (3,3);
\fadapt
\well
\gITS
\seq
\ITS4

\draw[line width = 5, red] (0,0) -- node[midway, below]{fwd adapter} (3,0);
\draw[line width = 5, blue] ++(0,0) -- node[midway,above]{well index} ++(1,0);
\draw[line width = 5, black!30!green] (4,0) -- node[midway, below]{gITS7} (5.5,0);
\draw[line width = 5] (5.5,0) -- (6.5,0);
\draw[line width = 5, dotted](6.5,0) -- node[midway, below]{sequence} (9,0);
\draw[line width = 5] (9,0) -- (10,0);
\draw[line width = 5, cyan] (10,0) -- node[midway, below]{ITS4} (11.5,0);
%\draw[line width = 5, purple] (11.5,0) -- node[midway, below]{rev adapter} (14.5,0);
\draw (0,0.5) -- (0,1) -- node[midway, above]{fwd primer}(5.5,1) -- (5.5, 0.5);
\draw (10,0.5) -- (10,1) -- node[midway, above]{rev primer}(11.5,1) -- (11.5, 0.5);

\draw[line width = 5, red] (-4,-3) -> node[midway, below]{fwd adapter} (-1,-3);
\draw[line width = 5, black!20!yellow] (-1,-3) -- node[midway,above]{plate index}(0,-3);
\draw[line width = 5, red] (0,-3) -- node[midway, below]{fwd adapter} (3,-3);
\draw[line width = 5, blue] (3,-3) -- node[midway,above]{well index}(4,-3);
\draw[line width = 5, black!30!green] (4,-3) -- node[midway, below]{gITS7} (5.5,-3);
\draw[line width = 5] (5.5,-3) -- (6.5,-3);
\draw[line width = 5, dotted](6.5,-3) -- node[midway, below]{sequence} (9,-3);
\draw[line width = 5] (9,-3) -- (10,-3);
\draw[line width = 5, cyan] (10,-3) -- node[midway, below]{ITS4} (11.5,-3);
\draw[line width = 5, orange] (11.5,-3) -> node[midway, below]{rev adapter} (14.5,-3);
%\draw[line width = 5, purple] (11.5,0) -- node[midway, below]{rev adapter} (14.5,0);
\draw (0,-2.25) -- (0,-2) -- node[midway, above]{fwd primer}(5.5,-2) -- (5.5, -2.25);
\draw (10,-2.25) -- (10,-2) -- node[midway, above]{rev primer}(11.5,-2) -- (11.5, -2.25);

\draw[line width = 5, red] (-4,-3 - 3) -> node[midway, below]{fwd adapter} (-1,-3- 3);
\draw[line width = 5, black!20!yellow] (-1,-3- 3) -- node[midway,above]{plate index}(0,-3- 3);
\draw[line width = 5, red] (11.5-0,-3- 3) -- node[midway, below]{fwd adapter} (11.5-3,-3- 3);
\draw[line width = 5, blue] (11.5-3,-3- 3) -- node[midway,above]{well index}(11.5-4,-3- 3);
\draw[line width = 5, black!30!green] (11.5-4,-3- 3) -- node[midway, below]{gITS7} (11.5-5.5,-3- 3);
\draw[line width = 5] (11.5-5.5,-3- 3) -- (11.5-6.5,-3- 3);
\draw[line width = 5, dotted](11.5-6.5,-3- 3) -- node[midway, below]{sequence} (11.5-9,-3- 3);
\draw[line width = 5] (11.5-9,-3- 3) -- (11.5-10,-3- 3);
\draw[line width = 5, cyan] (11.5-10,-3- 3) -- node[midway, below]{ITS4} (11.5-11.5,-3- 3);
\draw[line width = 5, orange] (11.5,-3- 3) -> node[midway, below]{rev adapter} (14.5,-3- 3);
%\draw[line width = 5, purple] (11.5,0) -- node[midway, below]{rev adapter} (14.5,0);
\draw (11.5-0,-2.25- 3) -- (11.5-0,-2- 3) -- node[midway, above]{fwd primer}(11.5-5.5,-2- 3) -- (11.5-5.5, -2.25- 3);
\draw (11.5-10,-2.25- 3) -- (11.5-10,-2- 3) -- node[midway, above]{rev primer}(11.5-11.5,-2- 3) -- (11.5-11.5, -2.25- 3);
\end{tikzpicture}

### 2018-12-18 R/LSU_align.R

@@ -1,0 +1,60 @@
+library(magrittr)
+library(tidyverse)
+library(Biostrings)
+
+if (interactive()) {
+  base.dir <- getwd() %>%
+    str_extract(".*oueme-fungi-transect")
+  data.dir <- file.path(base.dir, "data")
+  lab.dir <- file.path(data.dir, "lab_setup")
+  ref.dir <- file.path(base.dir, "reference")
+  lsuref <- file.path(ref.dir, "SILVA_132_LSURef_Fungi.fasta")
+  lsuout <- file.path(ref.dir, "SILVA_132_LSURef_Fungi.align.fasta")
+}
+
+lsu <- readRNAStringSet(lsuref)
+bases <- str_subset(alphabet(lsu), "\\w")
+sublsu <- subseq(lsu, 149001, width = 1000, allow.nonnarrowing = TRUE) %>%
+  as.matrix()
+lsumatrix <-
+  seq(1, min(width(lsu[1]), width(lsu[1]) %/% 1000 * 1000 + 1), by = 1000) %>%
+  map(~ { cat(., ": ")
+    subseq(lsu, ., width = min(1000, width(lsu[1]) - .)) %>%
+        as.matrix %>%
+        magrittr::extract(.,,plyr::alply(., .margins = 2, unique) %>%
+            map(is_in, bases) %>%
+            map_lgl(any) %T>%
+              {cat(sum(.), "\n")},
+          drop = FALSE)}) %>%
+  keep(~ length(.) > 0) %>%
+  do.call(cbind, .)
+
+lsumatrix %>%
+  apply(1, paste, collapse = "") %>%
+  RNAStringSet %>%
+  writeXStringSet(filepath = lsuout)
+
+lsuprofile <- 
+  apply(lsumatrix, 2, . %>% table %>% vegan::diversity() %>% exp)
+lsupresence <-
+  apply(lsumatrix, 2, . %>% is_in(bases) %>% sum) /
+  nrow(lsumatrix)
+  
+
+pllsumatrix %<>% keep(~ length(.) > 0)
+lsumatrix %<>% do.call(cbind, .)
+
+length(lsumatrix[[1]])
+subseq(lsu, 67001, width = 1000) %>%
+  as.matrix %>%
+  plyr::alply(2, unique) %>%
+  map(is_in, bases) %>%
+  map_lgl(any)
[…]

### 2018-12-18 R/demultiplex_one.R

@@ -1,0 +1,31 @@
+source("install_packages.R")
+
+library(magrittr)
+library(tidyverse)
+library(ShortRead)
+
+# for scripted use, these are specified in the makefile
+if (interactive()) {
+  base.dir <- getwd() %>%
+    str_extract(".*oueme-fungi-transect")
+  data.dir <- file.path(base.dir, "data")
+  lab.dir <- file.path(data.dir, "lab_setup")
+  in.fastq <- file.path(pacbio.dir, "pb_483_001.reads_of_insert.demux.fastq.gz")
+  out.fastq <- str_replace(in.fastq, fixed(".fastq.gz"), ".demux.fastq.gz")
+  in.groups <- str_replace(in.fastq, fixed(".demux.fastq.gz"), ".groups")
+  platekey <- file.path(lab.dir, "gITS7_platekey.csv")
+  dset <- "short"
+  plate <- paste0(dset, "_001")
+  demux.dir <- file.path(data.dir, "demultiplex")
+  plate.dir <- file.path(demux.dir, dset)
+  target <- file.path(plate.dir, paste0(plate, "_A1.fasta.gz"))
+}
+
+the_group <- str_replace(target, fixed(".fasta.gz"), "") %>%
+  basename
+fastq <- readFastq(in.fastq)
+groups <- read_csv(in.groups) %>%
+  filter(group == the_group)
+if (!dir.exists(dirname(target))) dir.create(dirname(target), recursive = TRUE)
+fastq %>% magrittr::extract(as.character(.@id) %in% groups$qseqid) %>%
+  writeFastq(file = target)

### 2018-12-18 R/install_packages.R

@@ -1,0 +1,39 @@
+if (!exists("force.install")) force.install <- FALSE
+
+maybe.install <- function(package, force.install = FALSE) {
+  if (!requireNamespace(package, quietly = TRUE) || force.install)
+    install.packages(package)
+}
+
+maybe.install.Bioc <- function(package, force.install = FALSE, version = "3.8") {
+  if (!requireNamespace(package, quietly = TRUE) || force.install)
+    BiocManager::install(package, version = version)
+}
+
+maybe.install.Git <- function(package, force.install = FALSE) {
+  if (!requireNamespace(package, quietly = TRUE) || force.install)
+    devtools::install_github(package)
+}
+
+
+lapply(c("magrittr",
+         "tidyverse",
+         "glue",
+         "BiocManager",
+         "readxl",
+         "plyr",
+         "knitr",
+         "devtools",
+         "seqinr"),
+       maybe.install,
+       force.install = force.install)
+
+lapply(c("ShortRead",
+         "Biostrings"),
+       maybe.install.Bioc,
+       force.install = force.install)
+
+lapply(c("benjjneb/dada2",
+         "hoesler/rwantshue"),
+       maybe.install.Git,
+       force.install = force.install)

### 2018-12-18 R/plate_check.R

@@ -1,0 +1,39 @@
+library(magrittr)
+library(tidyverse)
+library(readxl)
+library(tidyxl)
+
+if (interactive()) {
+  base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
+  data.dir <- file.path(base.dir, "data")
+  lab.dir <- file.path(data.dir, "lab_setup")
+  plate.file <- file.path(lab.dir, "Brendan_soil2.xlsx")
+  out.file <- file.path(lab.dir, "short.platekey.csv")
+}
+
+xlsx_cells(plate.file, "PCR-Setup") %>%
+  filter(data_type == "character")
+  
+
+plate <- read_xlsx(plate.file, "Concentration samples", skip = 2, col_names = TRUE) %>%
+  rename(Comment = X__1, Well = row) %>%
+  select (-X__2) %>%
+  # Sample number consists of site, location along transect, and possible qualifier
+  tidyr::extract("Sample", c("Site", "X", "Qual"), "(V1P[13]|T[12])[ -][ST](\\d+)(\\D*)$") %>%
+  mutate_at("X", as.integer) %>%
+  # Sites were recorded differently in the two years
+  mutate(Site = case_when(Site == "V1P1" ~ "Ang",
+                          Site == "V1P3" ~ "Gan",
+                          Site == "T1" ~ "Gan",
+                          Site == "T2" ~ "Ang",
+                          TRUE ~ NA_character_)) %>%
+  # For some site/years, coordinates were measured from the edge of the plot rather than numbered from 1--25 
+  group_by(Year, Site) %>%
+  mutate(X = X - if (!any(is.na(X)) && max(X) > 25) 12 else 0) %>%
+  ungroup %>%
+  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)", remove = FALSE) %>%
+  mutate_at("Row", factor, levels = LETTERS) %>%
+  mutate_at(c("Column", "Row"), as.integer) %>%
+  mutate(sample_type = factor(Comment, levels = c("Blank", "Pos", "Pos. Kontroll")) %>%
+           fct_collapse(Blank = "Blank", Pos = c("Pos", "Pos. Kontroll")) %>%
+           fct_explicit_na("Sample"))

### 2018-12-18 R/tags.extract.R

@@ -1,0 +1,91 @@
+library(magrittr)
+library(plyr)
+library(tidyverse)
+library(readxl)
+library(seqinr)
+
+# for scripted use, these are specified in the makefile
+if (interactive()) {
+  base.dir <- getwd() %>%
+    str_extract(".*oueme-fungi-transect")
+  data.dir <- file.path(base.dir, "data")
+  lab.dir <- file.path(data.dir, "lab_setup")
+  gits7.file <- file.path(lab.dir, "Hectors_tag_primer_plates.xlsx")
+  its1.lr5.file <- file.path(lab.dir, "Brendan_soil2.xlsx")
+  tags.dir <- file.path(lab.dir, "tags")
+  which.tags = c("gits7", "its1", "lr5", "its5")
+} else {
+  which.tags = basename(sys.frame(1)$ofile) %>%
+    str_extract("(gits7|its1|lr5|its4)")
+}
+
+# given a DNA sequence including IUPAC ambiguous bases, give all possible realizations
+all_ambiguous_bases <- function(s) {
+  if (str_length(s) == 0) return(s)
+  base <-
+    revalue(substring(s, 1, 1),
+            c(R = "GA",
+              Y = "TC",
+              S = "GC",
+              W = "AT",
+              K = "GT",
+              M = "AC",
+              D = "GTA",
+              H = "TAC",
+              B = "GTC",
+              V = "GAC",
+              N = "GATC"),
+            FALSE)
+  base <- strsplit(base, "") %>% unlist
+  if (str_length(s) == 1) return(base)
+  s <- substring(s, 2)
+  s <- expand.grid(base, all_ambiguous_bases(s), stringsAsFactors = FALSE)
+  do.call(paste0, s)
+}
+
+
+tags = list()
+
+# read the gITS7 tags
+if ("gits7" %in% which.tags) {
+  tags$gits7 <- read_xlsx(gits7.file, skip = 1) %>%
+    select(name = oligoname, object = sequence) %>%
+    filter(str_detect(name, "gITS7mod"))
+}
+# read the (non-tagged) primers
+if ("its4" %in% which.tags) {
+  tags$its4 <- read_xlsx(gits7.file, skip = 1) %>%
+    select(name = oligoname, object = sequence) %>%
+    filter(str_detect(name, "ITS4"), !str_detect(name, "-IT"))
+}
+# read the ITS1 tags
+if ("its1" %in% which.tags) {
[…]

### 2019-01-04 R/filter-check.Rmd
Demultiplex and quality filter.  Uses parallel processing in demultiplex_all.R, but the parallel BLAST doesn't seem to work.

@@
   lab.dir <- file.path(data.dir, "lab_setup")
-  gits7.file <- file.path(lab.dir, "Hectors tag primer plates.xlsx")
-  its1.lr5.file <- file.path(lab.dir, "Brendan soil2.xlsx")
-  tag.files <- list.files(file.path(lab.dir, "tags"), ".+\\.fasta", full.names = TRUE)
-  pacbio.dir <- file.path(data.dir, "PacBio")
-  pacbio.fastq.files <- list.files(pacbio.dir, "\\.fastq\\.gz", full.names = TRUE)
-  in.name <- "pb_500_001"
-  in.fastq <- file.path(pacbio.dir, "pb_500_001.reads_of_insert.fastq.gz")
-  in.fwd <- str_replace(in.fastq, fixed(".fastq.gz"), ".gits7.blast")
-  in.rev <- str_replace(in.fastq, fixed(".fastq.gz"), ".primer.blast")
-  out.fastq <- str_replace(in.fastq, fixed(".fastq.gz"), ".demux.fastq.gz")
-  out.groups <- str_replace(in.fastq, fixed(".fastq.gz"), ".groups")
-  platekey <- file.path(lab.dir, "gITS7_platekey.csv")
-  plate <- "short_002"
+  seq.dir <- file.path(base.dir, "raw_data")
+  dataset.file <- file.path(lab.dir, "datasets.csv")
+  datasets <- read_csv(dataset.file) %>%
+    mutate(demux.dir = file.path(seq.dir, Dataset, Seq.Run, "demultiplex"),
+           files = map(demux.dir, list.files, pattern = "*.fastq.gz",
+                       full.names = TRUE)) %>%
+    unnest() %>%
+    mutate(filesize = file.size(files)) %>%
+    filter(filesize > 0) %>%
+    mutate(fastq = map(files, readFastq),
+           eexp = map(fastq, ~ 10^(-1 * as(.@quality, "matrix")/10) %>%
+                        rowSums(na.rm = TRUE)),
+           length = map(fastq, width)) %>%
+    select(-fastq) %>%
+    unnest %>%
+    mutate(erate = eexp/length)
 }
@@
 ```{r}
-fastq <- readFastq(out.fastq)
-qual <- tibble(
-  eexp = 10^(-1 * as(fastq@quality, "matrix")/10) %>%
-  rowSums(na.rm = TRUE),
-  length = width(fastq),
-  erate = eexp/length)
-
-ggplot(qual, aes(x = erate)) +
+ggplot(datasets, aes(x = erate, group = Dataset, color = Dataset)) +
   stat_ecdf(geom = "step") +
@@
 
-ggplot(qual, aes(x = eexp)) +
+ggplot(datasets, aes(x = eexp, group = Dataset, color = Dataset)) +
   stat_ecdf(geom = "step") +
@@
 
-ggplot(qual, aes(x = length)) +
+ggplot(datasets, aes(x = length, group = Dataset, color = Dataset)) +
   stat_ecdf(geom = "step") +

### 2019-01-04 R/demultiplex_all.R

@@ -2,6 +2,7 @@ source("install_packages.R")
 
 library(magrittr)
 library(tidyverse)
+library(multidplyr)
 library(readxl)
 library(seqinr)
 library(glue)
@@ -34,33 +35,47 @@ if (interactive()) {
   dataset <- "short-ion"
   seq.run <- "is_057"
   fullstem <- file.path(seq.dir, dataset, seq.run)
-  in.fastq <- file.path(fullstem, "rawdata", "bc-subset", "IonXpress_001_rawlib.basecaller.fastq.gz")
+  in.fastq <- file.path(fullstem, "rawdata", "no_bc-subset", "rawlib.basecaller.fastq.gz")
+  demux.dir <- file.path(fullstem, "demultiplex")
   
   gits7.file <- file.path(lab.dir, "Hectors tag primer plates.xlsx")
   its1.lr5.file <- file.path(lab.dir, "Brendan soil2.xlsx")
   tag.files <- list.files(file.path(lab.dir, "tags"), ".+\\.fasta", full.names = TRUE)
   
-  in.fwd <- str_replace(in.fastq, fixed(".fastq.gz"), ".gits7_ion.blast")
-  in.rev <- str_replace(in.fastq, fixed(".fastq.gz"), ".its4.blast")
+  blastdb.fwd <- file.path(lab.dir, "tags", "gits7_ion")
+  blastdb.rev <- file.path(lab.dir, "tags", "its4")
+  
+  # in.fwd <- str_replace(in.fastq, fixed(".fastq.gz"), ".gits7_ion.blast")
+  # in.rev <- str_replace(in.fastq, fixed(".fastq.gz"), ".its4.blast")
   # in.mid <- str_replace(in.fastq, fixed(".fastq.gz"), ".its4.blast")
-  out.fastq <- str_replace(in.fastq, fixed(".fastq.gz"), ".demux.fastq.gz")
-  out.groups <- str_replace(in.fastq, fixed(".fastq.gz"), ".groups")
-  platekey <- file.path(lab.dir, "gITS7_platekey.csv")
-  plate <- "short_001"
+  #out.fastq <- str_replace(in.fastq, fixed(".fastq.gz"), ".demux.fastq.gz")
+  #out.groups <- str_replace(in.fastq, fixed(".fastq.gz"), ".groups")
+  platekey.file <- file.path(lab.dir, "gITS7_platekey.csv")
+  plate <- "short_ion_001"
[…]

### 2019-01-04 R/gits7_ion.extract.R

@@ -1,0 +1,1 @@
+tags.extract.R

### 2019-01-04 R/quality_trim.R

@@ -1,0 +1,79 @@
+source("install_packages.R", echo = TRUE)
+
+library(magrittr)
+library(tidyverse)
+library(ShortRead)
+
+# for scripted use, these are specified in the makefile
+if (interactive()) {
+  base.dir <- getwd() %>%
+    str_extract(".*oueme-fungi-transect")
+  data.dir <- file.path(base.dir, "data")
+  lab.dir <- file.path(data.dir, "lab_setup")
+  seq.dir <- file.path(base.dir, "raw_data")
+  
+  # choose a dataset and run for testing.
+  dataset <- "short-ion"
+  seq.run <- "is_057"
+  fullstem <- file.path(seq.dir, dataset, seq.run)
+  demux.dir <- file.path(fullstem, "demultiplex")
+  
+  # for interactive testing; choose the file in the dataset with the most reads.
+  in.file <- list.files(path = demux.dir, pattern = "*.fastq.gz", full.names = TRUE)
+  in.file <- in.file[!grepl(pattern = ".trim.", in.file, fixed = TRUE)]
+  in.file <- in.file[1]
+  
+  target <- sub(pattern = ".fastq.gz", replacement = ".trim.fastq.gz",
+                  x = in.file, fixed = TRUE)
+  target <- sub(pattern = "demultiplex", replacement = "trim", x = target,
+                  fixed = TRUE)
+  in.file <- "/home/brendan/Documents/Uppsala/Projects/oueme-fungi-transect/raw_data/short-pacbio/pb_483/demultiplex/short-pacbio-002_H1.fastq.gz"
+  target <- "/home/brendan/Documents/Uppsala/Projects/oueme-fungi-transect/raw_data/short-pacbio/pb_483/trim/short-pacbio-002_H1.trim.fastq.gz"
+  dataset.file <- file.path(lab.dir, "datasets.csv")
+} else {
+  cat("prereqs: ", prereqs, "\n")
+  in.file <- prereqs[grep(pattern = ".fastq.gz", 
+                  prereqs)]
+  cat("in.file: ", in.file, "\n")
+}
+
+# read the dataset definitions
+datasets <- read_csv(dataset.file)
+# figure out which dataset this file belongs to
+dataset <- str_extract(in.file, datasets$Dataset) %>% na.omit()
+cat("dataset: ", dataset, "\n")
+seq.run <- str_extract(in.file, datasets$Seq.Run) %>% na.omit()
[…]

### 2019-01-09 R/demultiplex_one.R
Convert from ad-hoc package management to packrat

@@ -1,42 +1,0 @@
-source("install_packages.R")
-
-library(magrittr)
-library(tidyverse)
-library(ShortRead)
-
-# for scripted use, these are specified in the makefile
-if (interactive()) {
-  base.dir <- getwd() %>%
-    str_extract(".*oueme-fungi-transect")
-  data.dir <- file.path(base.dir, "data")
-  lab.dir <- file.path(data.dir, "lab_setup")
-  in.fastq <- file.path(pacbio.dir, "pb_483_001.reads_of_insert.demux.fastq.gz")
-  out.fastq <- str_replace(in.fastq, fixed(".fastq.gz"), ".demux.fastq.gz")
-  in.groups <- str_replace(in.fastq, fixed(".demux.fastq.gz"), ".groups")
-  platekey <- file.path(lab.dir, "gITS7_platekey.csv")
-  dset <- "short"
-  plate <- paste0(dset, "_001")
-  demux.dir <- file.path(data.dir, "demultiplex")
-  plate.dir <- file.path(demux.dir, dset)
-  target <- file.path(plate.dir, paste0(plate, "_A1.fasta.gz"))
-} else {
-  in.groups <- str_subset(prereqs, "\\.groups")
-  in.fastq <- str_subset(prereqs, "\\.fastq\\.gz")
-}
-stopifnot(file.exists(in.groups),
-          file.exists(in.fastq))
-
-the_group <- str_replace(target, fixed(".fasta.gz"), "") %>%
-  basename
-groups <- read_csv(in.groups) %>%
-  filter(group == the_group)
-if (!dir.exists(dirname(target))) dir.create(dirname(target), recursive = TRUE)
-if (file.exists(target)) {
-  file.remove(target)
-}
-file.create(target)
-fastq <- FastqStreamer(in.fastq)
-while (length(fq <- yield(fastq))) {
-  fq %>% magrittr::extract(as.character(.@id) %in% groups$qseqid) %>%
-    writeFastq(file = target, mode = "a")
-}

### 2019-01-14 R/filter-check.Rmd
Convert from ad-hoc package management to packrat
Break long pacbio reads into ITS1, ITS2 and LSU before dada, because the long reads are all singletons.  Save output of dada and derep into the same file.

@@
 ```{r setup}
-source("install_packages.R")
+# source("install_packages.R")
 
@@
 ```{r}
+library("scales")
+reverselog_trans <- function(base = exp(1)) {
+    trans <- function(x) -log(x, base)
+    inv <- function(x) base^(-x)
+    trans_new(paste0("reverselog-", format(base)), trans, inv, 
+              log_breaks(base = base), 
+              domain = c(1e-100, Inf))
+}
+
 ggplot(datasets, aes(x = erate, group = Dataset, color = Dataset)) +
-  stat_ecdf(geom = "step") +
-  scale_x_log10(name = "Expected number of errors per base") +
+  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+  scale_x_continuous(name = "Expected number of errors per base",
+                     trans = reverselog_trans(10)) +
   scale_y_continuous(name = "Fraction passing")
@@
 ggplot(datasets, aes(x = eexp, group = Dataset, color = Dataset)) +
-  stat_ecdf(geom = "step") +
-  scale_x_log10(name = "Expected number of errors") +
+  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+  scale_x_continuous(name = "Expected number of errors",
+                     trans = reverselog_trans(10)) +
   scale_y_continuous(name = "Fraction passing")
@@
 ggplot(datasets, aes(x = length, group = Dataset, color = Dataset)) +
-  stat_ecdf(geom = "step") +
+  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
   scale_x_log10(name = "Length (bp)") +
   scale_y_continuous(name = "Fraction Greater than length")
+
+```
+```{r}
+datasets %>%
+  group_by(Dataset) %>%
+  do(out = dada2::plotQualityProfile(.$files, aggregate = TRUE) %>%
+       print) %>%
+  invisible
 

### 2019-01-14 R/dada.R

@@ -18,9 +18,10 @@ if (interactive()) {
   trim.dir <- file.path(fullstem, "trim")
   in.files <- list.files(path = trim.dir,
                         pattern = "\\.trim\\.fastq\\.gz$",
-                        full.names = TRUE)
+                        full.names = TRUE) 
   
   target <- paste0(file.path(data.dir, dataset), "_", seq.run, ".RDS")
+  dataset.file <- file.path(lab.dir, "datasets.csv")
 } else {
   cat("prereqs: ", prereqs, "\n")
   in.files <- list.files(path = prereqs,
@@ -44,6 +45,9 @@ datasets %<>% filter(Dataset == dataset,
 err.fun <- ifelse(datasets$Tech == "PacBio",
                   PacBioErrfun,
                   loessErrfun)
+if (!is.na(datasets$DadaOpt)) {
+  do.call(setDadaOpt, parse(text = datasets$DadaOpt))
+}
 
 # from here, based on sample script at http://benjjneb.github.io/dada2/bigdata.html
 # File parsing
@@ -54,14 +58,7 @@ set.seed(100)
 err <- learnErrors(in.files, nbases = 1e8, multithread=TRUE, randomize=TRUE,
                    errorEstimationFunction = err.fun,
                    verbose = TRUE)
-# Infer sequence variants
-dds <- vector("list", length(sample.names))
-names(dds) <- sample.names
-#for(sam in sample.names) {
-#  cat("Processing:", sam, "\n")
-  derep <- derepFastq(in.files)#[[sam]])
-  dds <- dada(derep, err=err, multithread=TRUE, pool = TRUE)
-#}
-# Construct sequence table and write to disk
-seqtab <- makeSequenceTable(dds)
-saveRDS(dds, target) # CHANGE ME to where you want sequence table saved
+derep <- derepFastq(in.files)#[[sam]])
+asv <- dada(derep, err=err, multithread=TRUE, pool = TRUE)
+
+save(derep, asv, file = target)

### 2019-01-14 R/extract_regions.R

@@ -1,0 +1,46 @@
+library(magrittr)
+library(tidyverse)
+library(glue)
+library(ShortRead)
+
+if (interactive()) {
+  base.dir <- getwd() %>%
+    str_extract(".*oueme-fungi-transect")
+  data.dir <- file.path(base.dir, "data")
+  ref.dir <- file.path(base.dir, "reference")
+  lab.dir <- file.path(data.dir, "lab_setup")
+  seq.dir <- file.path(base.dir, "raw_data")
+  
+  # choose a dataset and run for testing.
+  dataset <- "long-pacbio"
+  seq.run <- "pb_500"
+  plate <- "002"
+  well <- "D6"
+  position.file <- glue("{file.path(seq.dir, dataset, seq.run, 'trim', dataset)}-{plate}_{well}.trim.positions.txt")
+  seq.file <-  glue("{file.path(seq.dir, dataset, seq.run, 'trim', dataset)}-{plate}_{well}.trim.fastq.gz")
+  stem <- str_replace(position.file, fixed(".positions.txt"), "")
+} else {
+  position.file <- str_extract(prereqs, fixed(".positions.txt"))
+  seq.file <- str_extract(prereqs, fixed(".fasta.gz"))
+}
+
+
+pos <- read_tsv(position.file, col_names = c("seq", "length", "SSU", "ITS1",
+                                             "5_8S", "ITS2", "LSU", "comment")) %>%
+  gather(key = "region", value = "pos", SSU:LSU) %>%
+  mutate_at("pos", str_extract, pattern = "\\d+-\\d+") %>%
+  tidyr::extract(pos, into = c("start", "end"), regex = "(\\d+)-(\\d+)") %>%
+  mutate_at(vars(start:end), as.integer)
+
+for (r in c("ITS1", "ITS2", "LSU")) {
+  out.file <- glue("{stem}.{r}.fastq.gz")
+  if (file.exists(out.file)) file.remove(out.file)
+  fastq <- ShortReadQ()
+  writeFastq(fastq, out.file)
+  p <- filter(pos, region == r)
+  fastq <- readFastq(seq.file)
+  fastq <- fastq[match(p$seq, fastq@id)] %>%
+    narrow(start = p$start, end = p$end)
+  writeFastq(fastq, out.file, mode = "a")
+  dada2::plotQualityProfile(out.file) %>% print
+}

### 2019-01-14 R/make_demux.R

@@ -1,6 +1,6 @@
 # Generate a makefile to 
 
-source("install_packages.R", echo = TRUE)
+#source("install_packages.R", echo = TRUE)
 
 library(magrittr)
 library(tidyverse)
@@ -40,7 +40,13 @@ cat("#############################",
 
 # find potential source files
 datasets <- read_csv(dataset.file) %>%
-  mutate(dada.file = glue("$(DATADIR)/{Dataset}_{Seq.Run}.asv.RDS"))
+  mutate(Regions = map(Regions, ~parse(text = .)) %>%
+           map(eval)) %>%
+  unnest(Regions) %>%
+  mutate(Regions = as.character(Regions) %>%
+           paste0(".", .) %>%
+           str_replace(fixed("_NULL"), ""),
+         dada.file = glue("$(DATADIR)/{Dataset}_{Seq.Run}{Regions}.asv.Rdata"))
 
 datasets %>%
   with(paste0("dada : ", paste0(dada.file, collapse = ' \\\n       '))) %>%
@@ -94,11 +100,10 @@ datasets %>%
             "\t$(R)",
             "\tmv -f $@.prestamp $@",
             .sep = "\n") %>%
+  unique %>%
   glue_collapse(sep = "\n\n") %>%
   str_replace_all(fixed("|\n"), "\\\n") %>%
   cat("\n", sep = "", file = target, append = TRUE)
-
-
 
 # List of groups which need to be extracted.
 datasets %<>%
@@ -110,6 +115,7 @@ datasets %<>%
 datasets %>% glue_data("{OutFile} : {demux.dir}/.{Plate}-{shard} ;",
                        .sep = "\n",
                        .trim = FALSE) %>%
+  unique %>%
   glue_collapse(sep = "\n\n") %>%
   cat("\n", ., file = target, append = TRUE, sep = "")
 
@@ -118,22 +124,30 @@ datasets %>%
   mutate(demux.file = str_replace(OutFile, fixed(glue("-{shard}")), "")) %>%
   group_by(dada.file, demux.file) %>%
   summarize(shard.file = paste0(OutFile, collapse = " ")) %>%
-  mutate(TrimFile = str_replace(demux.file,
-                                fixed("demultiplex"),
-                                "trim"),
-         TrimFile = str_replace(TrimFile,
-                                fixed(".fastq.gz"),
-                                ".trim.fastq.gz")) %>%
   glue_data("demultiplex : {demux.file}",
[…]

### 2019-01-15 R/filter-check.Rmd

@@
   dataset.file <- file.path(lab.dir, "datasets.csv")
-  datasets <- read_csv(dataset.file) %>%
+  datasets <- read_csv(dataset.file)
+
+  reads <- datasets %>%
     mutate(demux.dir = file.path(seq.dir, Dataset, Seq.Run, "demultiplex"),
@@
 
-ggplot(datasets, aes(x = erate, group = Dataset, color = Dataset)) +
+ggplot(reads, aes(x = erate, group = Dataset, color = Dataset)) +
   geom_step(aes(y = 1- ..y..), stat = "ecdf") +
@@
 
-ggplot(datasets, aes(x = eexp, group = Dataset, color = Dataset)) +
+ggplot(reads, aes(x = eexp, group = Dataset, color = Dataset)) +
   geom_step(aes(y = 1- ..y..), stat = "ecdf") +
@@
 
-ggplot(datasets, aes(x = length, group = Dataset, color = Dataset)) +
+ggplot(reads, aes(x = length, group = Dataset, color = Dataset)) +
   geom_step(aes(y = 1- ..y..), stat = "ecdf") +
@@
 ```{r}
-datasets %>%
-  group_by(Dataset) %>%
-  do(out = dada2::plotQualityProfile(.$files, aggregate = TRUE) %>%
-       print) %>%
-  invisible
+counts <- read_csv(file.path(data.dir, "fastq.counts"), col_names = c("file", "reads"))
+counts %<>% mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique)),
+                   seq.run = map(file, str_extract, datasets$Seq.Run) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique)),
+                   well = str_extract(file, "[A-H]\\d\\d?"),
+                   shard = str_extract(file, "\\bx[a-z][a-z]\\b"),
+                   is.shard = !is.na(shard),
+                   region = str_extract(file, "(ITS[12]|LSU)"),
+                   is.trim = str_detect(file, "trim"),
+                   is.demux = is.na(region) & !is.trim & !is.na(well),
+                   type = ifelse(is.trim, "trim",
+                                 ifelse(!is.na(region), "ITSx",
+                                        ifelse(is.demux, "demux",
+                                               "original"))),
+                   plate = str_extract(file, "00\\d")) %>%
+  group_by(dataset, type, region, is.shard) %>%
+  summarize(reads = sum(reads)) %>%
+  ungroup() %>%
+  select(-is.shard) %>%
+  unique() %>%
+  spread(key = "type", value = "reads") %>%
+  select(dataset, original, demux, region, ITSx, trim) %>%
+  filter(!is.na(dataset))
 
+demux.counts <- read_lines(file.path(data.dir, "demux.counts")) %>%
+  str_match("([^:]+): (\\d+) sequences (.+)") %>%
+  magrittr::extract(,-1) %>%
+  set_colnames(c("file", "reads", "type")) %>%
+  as_tibble() %>%
+  mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique)),
+                   seq.run = map(file, str_extract, datasets$Seq.Run) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique))) %>%
+  group_by(dataset, seq.run, type) %>%
+  summarize(reads = sum(as.integer(reads))) %>%
+  ungroup() %>%
+  mutate(type = recode(type,
+                       "with at least one tag match. " = "any.tag",
+                       "match at least 90% of tag length. " = "tag.cover",
+                       "with at most 3 tag mismatches. " = "tag.score",
+                       "with no more than one distinct tag in each direction. " = "identifiable",
+                       "with tags in both directions. " = "both.tags",
+                       "with tags appearing in correct order. " = "tag.order")) %>%
+  spread(key = "type", value = "reads") %>%
+  select(dataset, any.tag, tag.cover, tag.score, identifiable, both.tags, tag.order) %>%
+  left_join(select(counts, dataset, original, demux) %>% filter(complete.cases(.))) %>%
+  select(dataset, original, any.tag:tag.order, demux) %>%
+  left_join(select(counts, dataset, region, ITSx, trim))
+
+short.pacbio <- new.env()
+load(file.path(data.dir, "short-pacbio_pb_483.asv.Rdata"), envir = short.pacbio)
+remove("derep", envir = short.pacbio)
+assign("seq.tab", dada2::makeSequenceTable(short.pacbio$asv), envir = short.pacbio)
+
+long.pacbio.ITS1 <- new.env()
+load(file.path(data.dir, "long-pacbio_pb_500.ITS1.asv.Rdata"), envir = long.pacbio.ITS1)
+remove("derep", envir = long.pacbio.ITS1)
+assign("seq.tab", dada2::makeSequenceTable(long.pacbio.ITS1$asv), envir = long.pacbio.ITS1)
+
+long.pacbio.ITS2 <- new.env()
+load(file.path(data.dir, "long-pacbio_pb_500.ITS2.asv.Rdata"), envir = long.pacbio.ITS2)
+remove("derep", envir = long.pacbio.ITS2)
+assign("seq.tab", dada2::makeSequenceTable(long.pacbio.ITS2$asv), envir = long.pacbio.ITS2)
+
+long.pacbio.LSU <- new.env()
+load(file.path(data.dir, "long-pacbio_pb_500.LSU.asv.Rdata"), envir = long.pacbio.LSU)
+remove("derep", envir = long.pacbio.LSU)
+assign("seq.tab", dada2::makeSequenceTable(long.pacbio.LSU$asv), envir = long.pacbio.LSU)
+
+short.ion <- new.env()
+load(file.path(data.dir, "short-ion_is_057.asv.Rdata"), envir = short.ion)
+remove("derep", envir = short.ion)
+assign("seq.tab", dada2::makeSequenceTable(short.ion$asv), envir = short.ion)
+
+dadacounts <- tribble( ~ dataset, ~region, ~assigned, ~asv,
+                       "short-pacbio", NA, sum(short.pacbio$seq.tab), ncol(short.pacbio$seq.tab),
+                       "long-pacbio", "ITS1", sum(long.pacbio.ITS1$seq.tab), ncol(long.pacbio.ITS1$seq.tab),
+                       "long-pacbio", "ITS2", sum(long.pacbio.ITS2$seq.tab), ncol(long.pacbio.ITS2$seq.tab),
+                       "long-pacbio", "LSU", sum(long.pacbio.LSU$seq.tab), ncol(long.pacbio.LSU$seq.tab),
+                       "short-ion", NA, sum(short.ion$seq.tab), ncol(short.ion$seq.tab))
 ```
-
-```{r}
-fastq@sread[width(fastq)>3000] %>% as.character
-```
-
-```{r}
-library(dada2)
-out2.fastq <- str_replace(out.fastq, fixed(".fastq"), ".trim.fastq")
-filterAndTrim(out.fastq, out2.fastq, maxLen = 2999)
-derep <- derepFastq(out2.fastq)
-
-err <- learnErrors(derep, errorEstimationFunction = dada2:::PacBioErrfun, multithread = TRUE)
-dd <- dada(derep, err = err, BAND_SIZE=32, multithread = TRUE)
-```

### 2019-01-16 R/filter-check.Rmd
Reduce length requirement for long pacbio quality trimming (because these will now be applied to ITSx products)

@@
 
-if (interactive()) {
+#if (interactive()) {
   base.dir <- getwd() %>%
@@
     mutate(erate = eexp/length)
-}
+#}
 ```
@@
   select(dataset, original, demux, region, ITSx, trim) %>%
-  filter(!is.na(dataset))
+  filter(!is.na(dataset)) %>%
+  group_by(dataset) %>%
+  mutate(original = na.omit(original),
+         demux = na.omit(demux)) %>%
+  filter(all(is.na(region)) | !is.na(region))
 
@@
 
+d
 short.pacbio <- new.env()
@@
                        "short-ion", NA, sum(short.ion$seq.tab), ncol(short.ion$seq.tab))
+
+demux.counts %>% left_join(dadacounts) %>%
+  filter(!is.na(assigned))
 ```

### 2019-01-25 R/filter-check.Rmd
Check where sequences are being lost in the pipeline.

--- 2019-01-14 R/filter-check.Rmd
@@
 library(ggplot2)
+library("scales")
 
-if (interactive()) {
+#if (interactive()) {
   base.dir <- getwd() %>%
@@
 ```{r}
-library("scales")
 reverselog_trans <- function(base = exp(1)) {
@@
 ```{r}
-datasets %>%
-  group_by(Dataset) %>%
-  do(out = dada2::plotQualityProfile(.$files, aggregate = TRUE) %>%
-       print) %>%
-  invisible
+counts <- read_csv(file.path(data.dir, "fastq.counts"), col_names = c("file", "reads"))
+counts %<>% mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique)),
+                   seq.run = map(file, str_extract, datasets$Seq.Run) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique)),
+                   well = str_extract(file, "[A-H]\\d\\d?"),
+                   shard = str_extract(file, "\\bx[a-z][a-z]\\b"),
+                   is.shard = !is.na(shard),
+                   region = str_extract(file, "(ITS[12]|LSU)"),
+                   is.trim = str_detect(file, "trim"),
+                   is.demux = is.na(region) & !is.trim & !is.na(well),
+                   type = ifelse(is.trim, "trim",
+                                 ifelse(!is.na(region), "ITSx",
+                                        ifelse(is.demux, "demux",
+                                               "original"))),
+                   plate = str_extract(file, "00\\d")) %>%
+  group_by(dataset, type, region, is.shard) %>%
+  summarize(reads = sum(reads)) %>%
+  ungroup() %>%
+  select(-is.shard) %>%
+  unique() %>%
+  spread(key = "type", value = "reads") %>%
+  select(dataset, original, demux, region, ITSx, trim) %>%
+  filter(!is.na(dataset)) %>%
+  group_by(dataset) %>%
+  mutate(original = na.omit(original),
+         demux = na.omit(demux)) %>%
+  filter(all(is.na(region)) | !is.na(region)) %>%
+  ungroup()
 
+demux.counts <- read_lines(file.path(data.dir, "demux.counts")) %>%
+  str_match("([^:]+): (\\d+) sequences ([^(]+)( \\(([FR]): (\\d+), ([FR]): (\\d+)\\))?\\.") %>%
+  magrittr::extract(,c(2,3,4,6,7,8, 9)) %>%
+  set_colnames(c("file", "reads", "type", "d1", "n1", "d2", "n2")) %>%
+  as_tibble() %>%
+  mutate_at(c("reads", "n1", "n2"), as.integer) %>%
+  mutate(fwd = case_when(!is.na(d1) & d1 == "F" ~ n1,
+                         !is.na(d2) & d2 == "F" ~ n2,
+                         TRUE ~ NA_integer_),
+         rev = case_when(!is.na(d1) & d1 == "R" ~ n1,
+                         !is.na(d2) & d2 == "R" ~ n2,
+                         TRUE ~ NA_integer_)) %>%
+  select(-(d1:n2)) %>%
+  mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique)),
+                   seq.run = map(file, str_extract, datasets$Seq.Run) %>%
+                     map_chr(purrr::compose(dplyr::first, sort, unique))) %>%
+  group_by(dataset, seq.run, type) %>%
+  summarize_at(c("reads", "fwd", "rev"), sum) %>%
+  ungroup() %>%
+  mutate(type = recode(type,
+                       "with at least one tag match" = "any.tag",
+                       "match at least 90% of tag length" = "tag.cover",
+                       "with at most 3 tag mismatches" = "tag.score",
+                       "with no more than one distinct tag in each direction" = "identifiable",
+                       "with tags in both directions" = "both.tags",
+                       "with tags appearing in correct order" = "tag.order")) %>%
+  gather(key = "dir", value = "reads", reads, fwd, rev) %>%
+  mutate_at("dir", str_replace, "^reads$", "") %>%
+  tidyr::unite("type", type, dir, sep = ".") %>%
+  spread(key = "type", value = "reads") %>%
+  set_names(str_replace(names(.), "\\.$", "")) %>%
+  keep(~!all(is.na(.))) %>%
+  select(dataset, starts_with("any.tag"), starts_with("tag.cover"),
+         starts_with("tag.score"), starts_with("identifiable"), both.tags, tag.order) %>%
+  left_join(select(counts, dataset, original, demux) %>%
+              filter(complete.cases(.)) %>%
+              unique()) %>%
+  select(dataset, original, any.tag:tag.order, demux) %>%
+  left_join(select(counts, dataset, region, ITSx, trim))
+
+dada.counts <-
+  demux.counts %>%
+  left_join(select(datasets, dataset = Dataset, Seq.Run)) %>%
+  mutate(file = glue("{data.dir}/{dataset}_{Seq.Run}.dada.seqtable.rds"),
+         file = ifelse(is.na(region),
+                       file,
+                       str_replace(file, "\\.dada", glue(".{region}.dada"))),
+         seq.tab = map(file, readRDS),
+         assigned = map_int(seq.tab, sum),
+         n_ASV = map_int(seq.tab, ncol)) %>%
+  select(-data, -seq.tab, -file)
+
+dada.counts
 ```
-
-```{r}
-fastq@sread[width(fastq)>3000] %>% as.character
-```
-
-```{r}
-library(dada2)
-out2.fastq <- str_replace(out.fastq, fixed(".fastq"), ".trim.fastq")
-filterAndTrim(out.fastq, out2.fastq, maxLen = 2999)
-derep <- derepFastq(out2.fastq)
-
-err <- learnErrors(derep, errorEstimationFunction = dada2:::PacBioErrfun, multithread = TRUE)
-dd <- dada(derep, err = err, BAND_SIZE=32, multithread = TRUE)
-```

### 2019-01-25 R/transect-analysis.Rmd
Check where sequences are being lost in the pipeline.

@@
 library(readxl)
-library(rwantshue)
+packrat::extlib("rwantshue")
 colorscheme <- iwanthue(seed=12, force_init = TRUE)

### 2019-01-26 writing/transect_paper.Rmd
Template for manuscript

---
title: "A Tale of Two Transects"
subtitle: "Also Two Sequencing Technologies, and Two Amplicons"
author: 
  - Brendan Furneaux^1^,
    Roël Houdanon^2^,
    Mohammad Bahram^1^,
    Anna Rosling^3^,
    Nourou S. Yorou^2^,
    Martin Ryberg^1^
  - ^1^Program in Systematic Biology,
    Department of Organismal Biology,
    Uppsala University,
    Uppsala, Sweden
  - ^2^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
    LEB,
    University of Parakou,
    Parakou, Benin
  - ^3^Program in Evolutionary Biology,
    Department of Ecology and Genetics,
    Uppsala University,
    Uppsala, Sweden
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography:
  - "all.bib"
  - "R.bib"
csl: fungal-ecology.csl
panflute-filters: figurefilter
panflute-path: '../../python'
output:
  bookdown::pdf_document2:
    fig_crop: no
    keep_tex: yes
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
  bookdown::word_document2: 
    keep_md: yes
---

```{r setup, include=FALSE}
library("knitr")
library("tikzDevice")
knitr::opts_chunk$set(dev="tikz",
                      # dev.args = list(pointsize=32),
                      echo=FALSE,
                      results='asis',
                      fig.show='asis',
                      external=TRUE)
options(tikzDefaultEngine = "luatex",
        tikzLualatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
                                "\\usepackage[version=4]{mhchem}\n",
                                "\\usepackage{siunitx}\n",
                                "\\DeclareSIUnit\\year{a}\n"))

base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
if (opts_chunk$get("dev") == "tikz") {
  # styles$Expression <- styles$LaTeX.Expression
  # styles$Name <- styles$LaTeX.Name
  # styles$Short.Name <- styles$LaTeX.Short.Name
  # styles$Unit <- paste0("\\si{", styles$LaTeX.Unit, "}")
}
write_bib(c("base", "vegan", "permute"), file = "R.bib", tweak = TRUE)
```
# Introduction

# Methods

# Results

# Discussion

# Bibliography {-}

<div id="refs"></div>

### 2019-01-30 R/filter-check.Rmd
Look at correspondence between PacBio and Ion short reads.

@@
 
+#}
+```
+```{r}
 reads <- datasets %>%
@@
     mutate(erate = eexp/length)
-#}
 ```
@@
          assigned = map_int(seq.tab, sum),
-         n_ASV = map_int(seq.tab, ncol)) %>%
-  select(-data, -seq.tab, -file)
+         n_ASV = map_int(seq.tab, ncol),
+         file = str_replace(file, "\\.dada\\.seqtable\\.rds",
+                            ".dada.nochim.rds"),
+         nochim = map(file, readRDS),
+         nochim.reads = map_int(nochim, sum),
+         nochim.ASV = map_int(nochim, ncol)) %>%
+  select(-seq.tab, -file, -nochim)
 
@@
 ```
+```{r}
+s.pb.seqs <- readRDS(glue("{data.dir}/short-pacbio_pb_483.dada.nochim.rds")) %>%
+  colSums %>%
+  tibble(seq = names(.),
+         pb.reads = .)
+s.ion.seqs <- readRDS(glue("{data.dir}/short-ion_is_057.dada.nochim.rds")) %>%
+  colSums %>%
+  tibble(seq = names(.),
+         ion.reads = .)
+
+(crosstab <- full_join(s.pb.seqs, s.ion.seqs, by = "seq") %>%
+  mutate(len = str_length(seq)) %>%
+  replace_na(list(pb.reads = 0, ion.reads = 0))) %>%
+  ggplot(aes(x = ion.reads, y = pb.reads, color = len)) +
+  geom_point() +
+  scale_y_log10() +
+  scale_x_log10()
+```
+
+IonTorrent reads which match a PacBio ASV: `r filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
+IonTorrent reads which do NOT match PacBio: `r filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
+PacBio reads which match an IonTorrent ASV: `r filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
+PacBio reads which do NOT match IonTorrent: `r filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`
+
+```{r}
+crosstab %>% arrange(pb.reads, ion.reads) %>%
+  write.csv(file.path(data.dir, "pb_vs_ion_ASVs.csv"))
+```
+
+```{r}
+demux.counts %>%
+  select(dataset, region) %>%
+  left_join(select(datasets, dataset = Dataset, Seq.Run)) %>%
+  mutate(file = glue("{data.dir}/{dataset}_{Seq.Run}.dada.taxonomy.rds"),
+         file = ifelse(is.na(region),
+                       file,
+                       str_replace(file, "\\.dada", glue(".{region}.dada")))) %>%
+  filter(file.exists(file)) %>%
+  mutate(tax = map(file, readRDS)) %>%
+  unnest %>%
+  filter(str_detect(Taxonomy, "Inocybe")) %>%
+  mutate(nreads = rowSums(select_if(., is.integer), na.rm = TRUE)) %>%
+  arrange(desc(nreads)) %>%
+  select(dataset, region, nreads, Taxonomy, seq) %>%
+  mutate(name = paste(dataset, region, substr(map_chr(seq, digest::digest, algo = "sha1"), 1, 8), sep = "_")) %$%
+  set_names(seq, name) %>%
+  DNAStringSet(use.names = TRUE) %>%
+  writeFasta(file.path(data.dir, "inocybe.fasta"))
+
+`short-ion_is_057.dada.nochim` %>%
+  colnames() %>%
+  set_names(paste0("seq", seq_along(.))) %>%
+  DNAStringSet(use.names = TRUE) %>%
+  writeFasta(file.path(data.dir, "short-ion.fasta"))
+```

### 2019-02-07 R/qual-check.Rmd

---
title: "Read quality check"
output: beamer_presentation
---

```{r setup, include = FALSE}
library(magrittr)
library(tidyverse)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")
library(drake)
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
datasets <- readd(datasets)
# xdata <- readd(itsxtrim_.its1.lr5.)
# pos <- readd(positions_pb500002A1f)
reads <- readd(qstats_join) %>%
      tidyr::extract(col = "file",
                     into = c("Seq.Run", "Plate", "Well", "Direction", "Region"),
                     regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)([fr]?)-([:alnum:]+)\\.trim\\.fastq\\.gz") %>%
  left_join(datasets)
```

```{r}
reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, 
              log_breaks(base = base), 
              domain = c(1e-100, Inf))
}
```

## Minimum quality score
```{r}
ggplot(reads, aes(x = minq,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Minimum quality score",
                     # trans = reverselog_trans(10),
                     # limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

## Expected error rate
```{r}

ggplot(reads, aes(x = erate,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors per base",
                     trans = reverselog_trans(10),
                     limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

## Expected number of errors
```{r}
ggplot(reads, aes(x = eexp,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors",
                     trans = reverselog_trans(10),
                     limits = c(NA, 0.1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

## Error-free probability

```{r}
ggplot(reads, aes(x = p.noerr,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expectation of being error-free",
                     # trans = log_trans(10),
                     limits = c(0, 1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")
```

## Read length

```{r}
ggplot(reads, aes(x = length,
                  linetype = Region,
                  group = paste(Dataset, Region),
                  color = Dataset)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_log10(name = "Length (bp)") +
  scale_y_continuous(name = "Fraction Greater than length")

```

## Expected error-free reads
```{r}
reads %>%
  group_by(Dataset, Plate, Region) %>%
  summarize(`Expected Error-Free` = sum(p.noerr),
            Total = n()) %>%
  knitr::kable()
```

```{r eval = FALSE}
library(dada2)
dada2::plotQualityProfile(unique(filter(reads, Tech == "PacBio")$files))
```

```{r eval = FALSE}
counts <- read_csv(file.path(data.dir, "fastq.counts"), col_names = c("file", "reads"))
counts %<>% mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
                     map_chr(purrr::compose(dplyr::first, sort, unique)),
                   seq.run = map(file, str_extract, datasets$Seq.Run) %>%
                     map_chr(purrr::compose(dplyr::first, sort, unique)),
                   well = str_extract(file, "[A-H]\\d\\d?"),
                   shard = str_extract(file, "\\bx[a-z][a-z]\\b"),
                   is.shard = !is.na(shard),
                   region = str_extract(file, "(ITS[12]|LSU)"),
                   is.trim = str_detect(file, "trim"),
                   is.demux = is.na(region) & !is.trim & !is.na(well),
                   type = ifelse(is.trim, "trim",
                                 ifelse(!is.na(region), "ITSx",
                                        ifelse(is.demux, "demux",
                                               "original"))),
                   plate = str_extract(file, "00\\d") %>% replace_na("001")) %>%
  group_by(dataset, type, plate, region, is.shard) %>%
  summarize(reads = sum(reads)) %>%
  ungroup() %>%
  select(-is.shard) %>%
  unique() %>%
  spread(key = "type", value = "reads") %>%
  select(dataset, plate, original, demux, region, ITSx, trim) %>%
  filter(!is.na(dataset)) %>%
  group_by(dataset) %>%
  mutate(original = na.omit(original),
         demux = na.omit(demux)) %>%
  filter(all(is.na(region)) | !is.na(region)) %>%
  ungroup()

demux.counts <- read_lines(file.path(data.dir, "demux.counts")) %>%
  str_match("([^:]+): (\\d+) sequences ([^(]+)( \\(([FR]): (\\d+), ([FR]): (\\d+)\\))?\\.") %>%
  magrittr::extract(,c(2,3,4,6,7,8, 9)) %>%
  set_colnames(c("file", "reads", "type", "d1", "n1", "d2", "n2")) %>%
  as_tibble() %>%
  mutate_at(c("reads", "n1", "n2"), as.integer) %>%
  mutate(fwd = case_when(!is.na(d1) & d1 == "F" ~ n1,
                         !is.na(d2) & d2 == "F" ~ n2,
                         TRUE ~ NA_integer_),
         rev = case_when(!is.na(d1) & d1 == "R" ~ n1,
                         !is.na(d2) & d2 == "R" ~ n2,
                         TRUE ~ NA_integer_)) %>%
  select(-(d1:n2)) %>%
  mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
           map_chr(purrr::compose(dplyr::first, sort, unique)),
         seq.run = map(file, str_extract, datasets$Seq.Run) %>%
           map_chr(purrr::compose(dplyr::first, sort, unique)),
         seq.plate = map(file, str_extract, paste0(datasets$Dataset, "-[0-9]{3}")) %>%
           map_chr(purrr::compose(dplyr::first, sort, unique))) %>%
  group_by(dataset, seq.run, seq.plate,type) %>%
  summarize_at(c("reads", "fwd", "rev"), sum) %>%
  ungroup() %>%
  mutate(type = recode(type,
                       "with at least one tag match" = "any.tag",
                       "match at least 90% of tag length" = "tag.cover",
                       "with at most 3 tag mismatches" = "tag.score",
                       "with no more than one distinct tag in each direction" = "identifiable",
                       "with tags in both directions" = "both.tags",
                       "with tags appearing in correct order" = "tag.order")) %>%
  gather(key = "dir", value = "reads", reads, fwd, rev) %>%
  mutate_at("dir", str_replace, "^reads$", "") %>%
  tidyr::unite("type", type, dir, sep = ".") %>%
  spread(key = "type", value = "reads") %>%
  set_names(str_replace(names(.), "\\.$", "")) %>%
  keep(~!all(is.na(.))) %>%
  select(dataset, starts_with("any.tag"), starts_with("tag.cover"),
         starts_with("tag.score"), starts_with("identifiable"), both.tags, tag.order) %>%
  left_join(select(counts, dataset, original, demux) %>%
              filter(complete.cases(.)) %>%
              unique()) %>%
  select(dataset, original, any.tag:tag.order, demux) %>%
  left_join(select(counts, dataset, region, ITSx, trim))

dada.counts <-
  demux.counts %>%
  left_join(select(datasets, dataset = Dataset, Seq.Run)) %>%
  mutate(file = glue("{data.dir}/{dataset}_{Seq.Run}.dada.seqtable.rds"),
         file = ifelse(is.na(region),
                       file,
                       str_replace(file, "\\.dada", glue(".{region}.dada"))),
         seq.tab = map(file, readRDS),
         assigned = map_int(seq.tab, sum),
         n_ASV = map_int(seq.tab, ncol),
         file = str_replace(file, "\\.dada\\.seqtable\\.rds",
                            ".dada.nochim.rds"),
         nochim = map(file, readRDS),
         nochim.reads = map_int(nochim, sum),
         nochim.ASV = map_int(nochim, ncol)) %>%
  select(-seq.tab, -file, -nochim)

dada.counts
```
```{r eval = FALSE}
s.pb.seqs <- readRDS(glue("{data.dir}/short-pacbio_pb_483.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         pb.reads = .)
s.ion.seqs <- readRDS(glue("{data.dir}/short-ion_is_057.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         ion.reads = .)

(crosstab <- full_join(s.pb.seqs, s.ion.seqs, by = "seq") %>%
  mutate(len = str_length(seq)) %>%
  replace_na(list(pb.reads = 0, ion.reads = 0))) %>%
  ggplot(aes(x = ion.reads, y = pb.reads, color = len)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10()
```

IonTorrent reads which match a PacBio ASV: ` filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
IonTorrent reads which do NOT match PacBio: ` filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
PacBio reads which match an IonTorrent ASV: ` filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
PacBio reads which do NOT match IonTorrent: ` filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`

```{r eval = FALSE}
crosstab %>% arrange(pb.reads, ion.reads) %>%
  write.csv(file.path(data.dir, "pb_vs_ion_ASVs.csv"))
```

```{r eval = FALSE}
demux.counts %>%
  select(dataset, region) %>%
  left_join(select(datasets, dataset = Dataset, Seq.Run)) %>%
  mutate(file = glue("{data.dir}/{dataset}_{Seq.Run}.dada.taxonomy.rds"),
         file = ifelse(is.na(region),
                       file,
                       str_replace(file, "\\.dada", glue(".{region}.dada")))) %>%
  filter(file.exists(file)) %>%
  mutate(tax = map(file, readRDS)) %>%
  unnest %>%
  filter(str_detect(Taxonomy, "Inocybe")) %>%
  mutate(nreads = rowSums(select_if(., is.integer), na.rm = TRUE)) %>%
  arrange(desc(nreads)) %>%
  select(dataset, region, nreads, Taxonomy, seq) %>%
  mutate(name = paste(dataset, region, substr(map_chr(seq, digest::digest, algo = "sha1"), 1, 8), sep = "_")) %$%
  set_names(seq, name) %>%
  DNAStringSet(use.names = TRUE) %>%
  writeFasta(file.path(data.dir, "inocybe.fasta"))

`short-ion_is_057.dada.nochim` %>%
  colnames() %>%
  set_names(paste0("seq", seq_along(.))) %>%
  DNAStringSet(use.names = TRUE) %>%
  writeFasta(file.path(data.dir, "short-ion.fasta"))
```

### 2019-02-07 R/transect-analysis.Rmd
update analysis for new data...

--- 2018-12-18 R/transect-analysis.Rmd
@@
 library(readxl)
-library(rwantshue)
+library(here)
+packrat::extlib("rwantshue")
 colorscheme <- iwanthue(seed=12, force_init = TRUE)
@@
 ```{r include = FALSE}
-base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
-data.dir <- file.path(base.dir, "data")
+data.dir <- here("data")
 soils.dir <- file.path(data.dir, "soils")
@@
 plate <- read_xlsx(plate.file, "Concentration samples", skip = 1, col_names = TRUE) %>%
-  rename(Comment = X__1, Well = row) %>%
-  select (-X__2) %>%
+  dplyr::rename(Comment = "..9", Well = row) %>%
+  select (-"..12") %>%
   # Sample number consists of site, location along transect, and possible qualifier

### 2019-02-11 R/filter-check.Rmd
Incorporate dada workflow through guild assignment into drake; use new .data argument in drake transforms.

@@
   lab.dir <- file.path(data.dir, "lab_setup")
-  seq.dir <- file.path(base.dir, "raw_data")
+  seq.dir <- file.path(base.dir, "sequences")
+  rawfastq.dir <- file.path(seq.dir, "rawfastq")
+  demux.dir <- file.path(seq.dir, "demux")
+  trim.dir <- file.path(seq.dir, "trim")
   dataset.file <- file.path(lab.dir, "datasets.csv")
@@
 ```
+
 ```{r}
+
 reads <- datasets %>%
-    mutate(demux.dir = file.path(seq.dir, Dataset, Seq.Run, "demultiplex"),
-           files = map(demux.dir, list.files, pattern = "*.fastq.gz",
-                       full.names = TRUE)) %>%
+  mutate(trim.files = map(paste0(Seq.Run, ".*\\.fastq\\.gz"), list.files, path = trim.dir,
+                          full.names = TRUE),
+         old.demux.files = map(paste0(Dataset, ".*\\.fastq\\.gz"), list.files,
+                               path = demux.dir, full.names = TRUE),
+         files = pmap(list(trim.files, old.demux.files), c)) %>%
+  select(-trim.files, -old.demux.files) %>%
   unnest() %>%
-    mutate(filesize = file.size(files)) %>%
-    filter(filesize > 0) %>%
+  filter(!str_detect(files, "-x[a-z]{2}\\.fastq\\.gz")) %>%
+  mutate(filesize = file.size(files),
+         Base.Calling = ifelse(str_detect(basename(files), Seq.Run),
+                               "arrow", "default")) %>%
+  filter(filesize > 0,
+         Base.Calling == "default" | Tech == "PacBio") %>%
   mutate(fastq = map(files, readFastq),
+         # fastq = map(fastq, ~magrittr::extract(., width(.) > 0)),
          eexp = map(fastq, ~ 10^(-1 * as(.@quality, "matrix")/10) %>%
                       rowSums(na.rm = TRUE)),
+         p.noerr = map(fastq, ~log1p(- 10^(-1 * as(.@quality, "matrix")/10)) %>%
+                         rowSums(na.rm = TRUE) %>%
+                         exp),
          length = map(fastq, width)) %>%
@@
 
-ggplot(reads, aes(x = erate, group = Dataset, color = Dataset)) +
+ggplot(reads, aes(x = erate,
+                  # group = paste(Base.Calling, Dataset),
+                  color = Dataset,
+                  linetype = Base.Calling)) +
   geom_step(aes(y = 1- ..y..), stat = "ecdf") +
   scale_x_continuous(name = "Expected number of errors per base",
-                     trans = reverselog_trans(10)) +
+                     trans = reverselog_trans(10),
+                     limits = c(NA, 1e-5),
+                     oob = squish) +
   scale_y_continuous(name = "Fraction passing")
 
-ggplot(reads, aes(x = eexp, group = Dataset, color = Dataset)) +
+ggplot(reads, aes(x = eexp,
+                  group = paste(Base.Calling, Dataset),
+                  color = Dataset,
+                  linetype = Base.Calling)) +
   geom_step(aes(y = 1- ..y..), stat = "ecdf") +
   scale_x_continuous(name = "Expected number of errors",
-                     trans = reverselog_trans(10)) +
+                     trans = reverselog_trans(10),
+                     limits = c(NA, 0.1),
+                     oob = squish) +
   scale_y_continuous(name = "Fraction passing")
 
-ggplot(reads, aes(x = length, group = Dataset, color = Dataset)) +
+ggplot(reads, aes(x = p.noerr,
+                  group = paste(Base.Calling, Dataset),
+                  color = Dataset,
+                  linetype = Base.Calling)) +
+  geom_step(aes(y = 1 - ..y..), stat = "ecdf") +
+  scale_x_continuous(name = "Expectation of being error-free",
+                     # trans = log_trans(10),
+                     limits = c(0, 1),
+                     oob = squish) +
+  scale_y_continuous(name = "Fraction passing")
+
+ggplot(reads, aes(x = length,
+                  group = paste(Base.Calling, Dataset),
+                  color = Dataset,
+                  linetype = Base.Calling)) +
   geom_step(aes(y = 1- ..y..), stat = "ecdf") +
@@
 
+reads %>%
+  group_by(Base.Calling, Dataset) %>%
+  summarize(`Expected Error-Free` = sum(p.noerr),
+            Total = n()) %>%
+  knitr::kable()
 ```
+
+```{r}
+library(dada2)
+dada2::plotQualityProfile(unique(filter(reads, Tech == "PacBio")$files))
+```
+
 ```{r}
@@
                                                "original"))),
-                   plate = str_extract(file, "00\\d")) %>%
-  group_by(dataset, type, region, is.shard) %>%
+                   plate = str_extract(file, "00\\d") %>% replace_na("001")) %>%
+  group_by(dataset, type, plate, region, is.shard) %>%
   summarize(reads = sum(reads)) %>%
@@
   spread(key = "type", value = "reads") %>%
-  select(dataset, original, demux, region, ITSx, trim) %>%
+  select(dataset, plate, original, demux, region, ITSx, trim) %>%
   filter(!is.na(dataset)) %>%
@@
          seq.run = map(file, str_extract, datasets$Seq.Run) %>%
+           map_chr(purrr::compose(dplyr::first, sort, unique)),
+         seq.plate = map(file, str_extract, paste0(datasets$Dataset, "-[0-9]{3}")) %>%
            map_chr(purrr::compose(dplyr::first, sort, unique))) %>%
-  group_by(dataset, seq.run, type) %>%
+  group_by(dataset, seq.run, seq.plate,type) %>%
   summarize_at(c("reads", "fwd", "rev"), sum) %>%

### 2019-02-11 R/assign_taxonomy.R

@@ -1,72 +1,0 @@
-library(magrittr)
-library(tidyverse)
-library(glue)
-library(assertthat)
-library(dada2)
-
-if (interactive()) {
-  base.dir <- getwd() %>%
-    str_extract(".*oueme-fungi-transect")
-  data.dir <- file.path(base.dir, "data")
-  ref.dir <- file.path(base.dir, "reference")
-  
-  # choose a dataset and run for testing.
-  dataset <- "long-pacbio"
-  seq.run <- "pb_500"
-  region = ".LSU"
-  in.file <- glue("{file.path(data.dir, dataset)}_{seq.run}{region}.dada.seqtable.rds")
-  target.rds <- str_replace(in.file, fixed(".seqtable.rds"), ".taxonomy.rds")
-  target.csv <- str_replace(in.file, fixed(".seqtable.rds"), ".taxonomy.csv")
-  reference <- file.path(ref.dir, "lsu_ref.fasta.gz")
-} else {
-  # Get the necessary information from environmental variables (set by make)
-  # or from stdin.
-  targets <- Sys.getenv("TARGETLIST") %>% str_split(" ") %>% unlist
-  assert_that(length(targets) > 0)
-  target.rds <- str_subset(targets, ".*\\.rds")
-  target.csv <- str_subset(targets, ".*\\.csv")
-  con <- file("stdin")
-  open(con, blocking = TRUE)
-  prereqs <- readLines(con)
-  close(con)
-  prereqs <- str_split(prereqs, " ") %>% unlist
-  in.file <- str_subset(prereqs, fixed(".nochim.rds"))
-  reference <- str_subset(prereqs, fixed(".fasta.gz"))
-}
-
-assert_that(file.exists(in.file),
-            is.readable(in.file),
-            file.exists(reference),
-            is.readable(reference),
-            is.dir(dirname(target.rds)),
-            is.dir(dirname(target.csv)),
-            is.writeable(dirname(target.rds)),
-            is.writeable(dirname(target.csv)))
-seq.table <- readRDS(in.file)
-
-tax <- seq.table %>%
-  colnames %>%
-  assignTaxonomy(reference, multithread = TRUE) %>%
-  as_tibble(rownames = "seq") %>%
-  # remove taxon rank prefixed from Unite reference
-  mutate_at(vars(-seq), str_replace, "^[kpcofgs]__", "") %>%
-  mutate(
-    # add Species to RDP reference
[…]

### 2019-02-11 R/extract_regions.R

@@ -10,9 +10,10 @@ extract_region <- function(infile, outfile, region, positions) {
   assert_that(dir.exists(dirname(outfile)))
   
   # if the region is "full", then we don't need to cut anything.
-  if (identical(region, "full")) {
-    file.copy(infile, outfile, overwrite = TRUE)
-    return()
+  if (region %in% c("full", "long", "short")) {
+    return(file.copy(from = infile, 
+              to = outfile,
+              overwrite = TRUE))
   }
   
   # make sure the file exists even if we don't have anything to write.

### 2019-02-11 R/transect-analysis.Rmd
Restructure to fit standard Snakemake project structure with ./scripts and ./config directories, part 1

@@
   summarize_at("X", c("min", "max"))
-```
-
-```{r}
-# Load the IonTorrent OTU table
-ion.otu <- read_tsv(iontorrent.otu.file) %>%
-  mutate(n_reads = rowSums(select_if(., is_numeric))) %>%
-  left_join(read_tsv(iontorrent.classification.file)) %>%
-  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
-  left_join(read_tsv(iontorrent.funguild.file) %>% rename(OTU = Group)) %>%
-  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
-  rename(Kingdom = Domain) %>%
-  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "?") %>%
-  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "unidentified") %>%
-
-  # make "unident." as informative as possible
-  dplyr::mutate_at("Kingdom", replace_na, "unident.") %>%
-  dplyr::mutate(Phylum = ifelse(is.na(Phylum),
-                                ifelse(is.na(Kingdom),
-                                       NA_character_,
-                                       paste0("unident. ", Kingdom)),
-                                Phylum),
-                Class = ifelse(is.na(Class),
-                                ifelse(is.na(Phylum),
-                                       NA_character_,
-                                       paste0("unident. ", Phylum)),
-                                Class),
-                Order = ifelse(is.na(Order),
-                                ifelse(is.na(Class),
-                                       NA_character_,
-                                       paste0("unident. ", Class)),
-                                Order),
-                Family = ifelse(is.na(Family),
-                                ifelse(is.na(Order),
-                                       NA_character_,
-                                       paste0("unident. ", Order)),
-                                Family),
-                Genus = ifelse(is.na(Genus),
-                                ifelse(is.na(Family),
-                                       NA_character_,
-                                       paste0("unident. ", Family)),
-                                Genus),
-                Species = ifelse(is.na(Species),
-                                ifelse(is.na(Genus),
-                                       NA_character_,
-                                       paste0(Genus, " sp.")),
-                                Species)) %>%
-  mutate_at(dplyr::vars(Kingdom:Species),
-            str_replace,
-            "(unident\\. )*unident(ified|\\.)",
-            "ZZZZunident.") %>%
-
-  # Sort taxonomically. Put the "unident." at the end
-  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
-  mutate_at(dplyr::vars(Kingdom:Species),
-            str_replace,
-            "ZZZZ",
-            "") %>%
-
-  # Add the representative sequence for each OTU and calculate the length
-  left_join(
-    seqinr::read.fasta(file = iontorrent.repseq.file, as.string = TRUE) %>%
-    { tibble(OTU = names(.),
-             RepSeq = unlist(.),
-             SeqLen = str_length(RepSeq)) }) %>%
-
-  # Remove the positive control
-  filter(is.na(Species) | Species != "Agaricus bisporus") %>%
-
-  # Keep the data in this order.
-  mutate_at(dplyr::vars(Kingdom:Species),
-            ~ factor(.) %>% fct_inorder)
-
-ion.platemap <- read_tsv(iontorrent.otu.file) %>%
-  column_to_rownames("OTU") %>%
-  as.matrix %>%
-  colSums %>%
-  tibble(Well = names(.), n_reads = .) %>%
-  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE) %>%
-  full_join(plate %>%
-              select(Plate, Well, sample_type) %>%
-              full_join(
-                expand.grid(Plate = 1:2,
-                            Row = LETTERS[1:8],
-                            Column = 1:12) %>%
-                  tidyr::unite(Well, Row, Column, sep = ""))) %>%
-  mutate(Plate = "IonTorrent") %>%
-  mutate_at("sample_type", fct_explicit_na, "None") %>%
-  group_by_if(~ !is.factor(.)) %>%
-  summarize(sample_type = paste(as.character(sort(unique(sample_type))), collapse = "/")) %>%
-  ungroup %>%
-  mutate(f_reads = n_reads / max(n_reads)) %>%
-  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
-  mutate_at("Row", factor, levels = LETTERS) %>%
-  mutate_at(c("Column", "Row"), as.integer)
 ```
@@
 
-pb.otumatrix <- pb.otu %>%
+pb.otumatrix <- readRDS(file.path(data.dir, "short-pacbio_pb_483.dada.seqtable.rds"))
+
+  pb.otu %>%
   column_to_rownames("Group") %>%

### 2019-02-11 R/demultiplex_all.R

@@ -50,7 +50,7 @@ if (interactive()) {
   base.dir <- getwd() %>%
     str_extract(".*oueme-fungi-transect")
   data.dir <- file.path(base.dir, "data")
-  lab.dir <- file.path(data.dir, "lab_setup")
+  lab.dir <- here("config")
   seq.dir <- file.path(base.dir, "raw_data")
   dataset <- "short-ion"
   seq.run <- "is_057"

### 2019-02-11 R/drake.R

@@ -16,9 +16,9 @@ library(rITSx)
 
 if (interactive()) {
   base.dir <- here()
-  r.dir <- here("R")
+  r.dir <- here("scripts")
   data.dir <- here("data")
-  lab.dir <- here("data", "lab_setup")
+  lab.dir <- here("config")
   seq.dir <- here("sequences")
   trim.dir <- file.path(seq.dir, "trim")
   region.dir <- file.path(seq.dir, "regions")

### 2019-02-11 R/make_demux.R

@@ -11,7 +11,7 @@ if (interactive()) {
   seq.dir <- here("sequences")
   demux.dir <- file.path(seq.dir, "demux")
   data.dir <- file.path(base.dir, "data")
-  lab.dir <- file.path(data.dir, "lab_setup")
+  lab.dir <- here("config")
   dataset.file <- file.path(lab.dir, "datasets.csv")
   target <- file.path(base.dir, "demux.make")
   splits <- expand.grid(letters, letters) %>%

### 2019-02-11 R/plate_check.R

@@ -6,7 +6,7 @@ library(tidyxl)
 if (interactive()) {
   base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
   data.dir <- file.path(base.dir, "data")
-  lab.dir <- file.path(data.dir, "lab_setup")
+  lab.dir <- here("config")
   plate.file <- file.path(lab.dir, "Brendan_soil2.xlsx")
   out.file <- file.path(lab.dir, "short.platekey.csv")
 }

### 2019-02-11 R/tags.extract.R

@@ -9,7 +9,7 @@ library(assertthat)
 # for scripted use, these are specified in the makefile
 if (interactive()) {
   data.dir <- here("data")
-  lab.dir <- file.path(data.dir, "lab_setup")
+  lab.dir <- here("config")
   gits7.file <- file.path(lab.dir, "Hectors_tag_primer_plates.xlsx")
   its1.lr5.file <- file.path(lab.dir, "Brendan_soil2.xlsx")
   tags.dir <- file.path(lab.dir, "tags")

### 2019-02-11 scripts/qual-check.Rmd
REstructure to Snakemake standard, part 2 (actual restructuring)

--- 2019-02-07 R/qual-check.Rmd
@@
 ```{r eval = FALSE}
-counts <- read_csv(file.path(data.dir, "fastq.counts"), col_names = c("file", "reads"))
-counts %<>% mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
-                     map_chr(purrr::compose(dplyr::first, sort, unique)),
-                   seq.run = map(file, str_extract, datasets$Seq.Run) %>%
-                     map_chr(purrr::compose(dplyr::first, sort, unique)),
-                   well = str_extract(file, "[A-H]\\d\\d?"),
-                   shard = str_extract(file, "\\bx[a-z][a-z]\\b"),
-                   is.shard = !is.na(shard),
-                   region = str_extract(file, "(ITS[12]|LSU)"),
-                   is.trim = str_detect(file, "trim"),
-                   is.demux = is.na(region) & !is.trim & !is.na(well),
-                   type = ifelse(is.trim, "trim",
-                                 ifelse(!is.na(region), "ITSx",
-                                        ifelse(is.demux, "demux",
-                                               "original"))),
-                   plate = str_extract(file, "00\\d") %>% replace_na("001")) %>%
-  group_by(dataset, type, plate, region, is.shard) %>%
+counts <- readd(seq_counts) %>% 
+  tidyr::extract(col = "file",
+                     into = c("Seq.Run", "Plate", "Well","Region","Type"),
+                     regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)[fr]?-?([:alnum:]*)\\.([:alnum:]+).+") %>%
+  mutate(Type = ifelse(Type == "trim" & str_length(Region) > 0,
+                       "itsx",
+                       Type)) %>%
+  group_by(Seq.Run, Region, Type) %>%
   summarize(reads = sum(reads)) %>%
-  ungroup() %>%
-  select(-is.shard) %>%
-  unique() %>%
-  spread(key = "type", value = "reads") %>%
-  select(dataset, plate, original, demux, region, ITSx, trim) %>%
-  filter(!is.na(dataset)) %>%
-  group_by(dataset) %>%
-  mutate(original = na.omit(original),
-         demux = na.omit(demux)) %>%
-  filter(all(is.na(region)) | !is.na(region)) %>%
+  spread(key = "Type", value = "reads") %>%
+  select(Seq.Run, trim, Region, itsx, qfilt) %>%
+  group_by(Seq.Run) %>%
+  mutate(trim = na.omit(trim)) %>%
+  filter(all(is.na(itsx)) | !is.na(itsx)) %>%
   ungroup()

### 2019-02-11 scripts/make_cutadapt_barcodes.R

@@ -1,0 +1,27 @@
+library(magrittr)
+library(tidyverse)
+library(Biostrings)
+library(here)
+
+if (interactive()) {
+  data.dir <- here("data")
+  lab.dir <- here("config")
+  seq.dir <- here("sequences")
+  dataset.file <- file.path(lab.dir, "datasets.csv")
+} else {
+  dataset.file <- Sys.getenv("DATASET")
+  target <- Sys.getenv("TARGETLIST")
+  
+  # get the prereq files on stdin
+  # there are too many to pass as an environmental variable!
+  con <- file("stdin")
+  open(con, blocking = TRUE)
+  prereqs <- readLines(con)
+  close(con)
+  prereqs <- str_split(prereqs, " ") %>% unlist
+  cat("prereqs: ", prereqs, "\n")
+  
+  in.files <- str_subset(prereqs,
+                         pattern = "\\.trim\\.fastq\\.gz$")
+  cat("in.files: ", in.files, "\n")
+}

### 2019-02-15 writing/qual-check.Rmd
move Rmd files to writing/ and ignore the directory on Packrat; fix report knitting on UPPMAX

--- 2019-02-11 scripts/qual-check.Rmd
@@
 # pos <- readd(positions_pb500002A1f)
-reads <- readd(qstats_join) %>%
-      tidyr::extract(col = "file",
-                     into = c("Seq.Run", "Plate", "Well", "Direction", "Region"),
-                     regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)([fr]?)-([:alnum:]+)\\.trim\\.fastq\\.gz") %>%
-  left_join(datasets)
+reads <- readd(qstats_join)
 ```

### 2019-02-15 build.sh

@@ -31,7 +31,8 @@
 rm -f demux.make &&
 
 module load R/3.5.0 gcc bioinfo-tools blast/2.7.1+ samtools/1.9 ITSx/1.1-beta \
-            gnuparallel/20180822 SMRT/5.0.1 cutadapt/1.16 Fastx/0.0.14 &&
+            gnuparallel/20180822 SMRT/5.0.1 cutadapt/1.16 Fastx/0.0.14 \
+            texlive/2018 &&
 
 # convert PacBio RSII files to .bam format
 # these are single-threaded options

### 2019-02-15 scripts/drake.R

@@ -28,6 +28,7 @@ if (interactive()) {
   region.dir <- file.path(seq.dir, "regions")
   filter.dir <- file.path(seq.dir, "filter")
   ref.dir <- here("reference")
+  rmd.dir <- here("writing")
   out.dir <- here("output")
   in.files <- list.files(path = trim.dir,
                          pattern = "\\.trim\\.fastq\\.gz$",
@@ -45,6 +46,7 @@ if (interactive()) {
   region.dir <- Sys.getenv("REGIONDIR")
   filter.dir <- Sys.getenv("FILTERDIR")
   ref.dir <- Sys.getenv("REF_ROOT")
+  rmd.dir <- Sys.getenv("RMDDIR")
   out.dir <- Sys.getenv("OUTDIR")
   # get the prereq files on stdin
   # there are too many to pass as an environmental variable!
@@ -329,14 +331,20 @@ plan <- drake_plan(
                     .id = RID)),
   # join all the quality stats into one data.frame
   qstats_join = target(
-    bind_rows(qstats),
+    bind_rows(qstats) %>%
+      tidyr::extract(
+        col = "file",
+        into = c("Seq.Run", "Plate", "Well", "Direction", "Region"),
+        regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)([fr]?)-([:alnum:]+)\\.trim\\.fastq\\.gz") %>%
+      left_join(datasets),
     transform = combine(qstats)),
   # knit a report about the quality stats.
   qstats_knit = {
-    if (!dir.exists(out.dir)) dir.create(out.dir, recursive = TRUE)
+    if (!dir.exists(!!out.dir)) dir.create(!!out.dir, recursive = TRUE)
     rmarkdown::render(
-      knitr_in(!!file.path(r.dir, "qual-check.Rmd")),
-      output_file = file_out(!!file.path(out.dir, "qual-check.pdf")))},
+      knitr_in(!!file.path(rmd.dir, "qual-check.Rmd")),
+      output_file = file_out(!!file.path(out.dir, "qual-check.pdf")),
+      output_dir = !!out.dir)},
   trace = TRUE
 )
 

### 2019-02-21 writing/qual-check.Rmd
Set up for clustermq, require patched version of dada2 which allows to specify fastq quality format, add comments, include full ITS in region extraction

@@
 library(drake)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, device = "tikz")
 ```
@@
 # pos <- readd(positions_pb500002A1f)
-reads <- readd(qstats_join)
+reads <- readd(qstats_join) %>%
+  mutate_at("Region", factor,
+            levels = c("ITS2", "ITS1", "short", "ITS", "LSU", "long")) %>%
+  mutate_at("Dataset", factor,
+            levels = c("short-ion", "short-pacbio", "long-pacbio")) %>%
+  group_by(Dataset, Plate, Region)
+reads_subset <- reads %>%
+  do(if (nrow(.) > 10000) sample_n(., 10000) else .)
+summarize(reads, n = n())
 ```
@@
 ```{r}
-ggplot(reads, aes(x = minq,
-                  linetype = Region,
+ggplot(reads_subset, aes(x = minq,
+                  linetype = Dataset,
                   group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+                  color = Region)) +
+  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
   scale_x_continuous(name = "Minimum quality score",
@@
                      oob = squish) +
-  scale_y_continuous(name = "Fraction passing")
+  scale_y_continuous(name = "Fraction passing") +
+  scale_color_viridis_d(option = "C") +
+  scale_linetype_discrete() +
+  theme_bw()
 ```
@@
 
-ggplot(reads, aes(x = erate,
-                  linetype = Region,
+ggplot(reads_subset, aes(x = erate,
+                  linetype = Dataset,
                   group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+                  color = Region)) +
+  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
   scale_x_continuous(name = "Expected number of errors per base",
@@
 ```{r}
-ggplot(reads, aes(x = eexp,
-                  linetype = Region,
+ggplot(reads_subset, aes(x = eexp,
+                  linetype = Dataset,
                   group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+                  color = Region)) +
+  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
   scale_x_continuous(name = "Expected number of errors",
@@
 ```{r}
-ggplot(reads, aes(x = p.noerr,
-                  linetype = Region,
+ggplot(reads_subset, aes(x = p.noerr,
+                  linetype = Dataset,
                   group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1 - ..y..), stat = "ecdf") +
+                  color = Region)) +
+  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
   scale_x_continuous(name = "Expectation of being error-free",
@@
 ```{r}
-ggplot(reads, aes(x = length,
-                  linetype = Region,
+ggplot(reads_subset, aes(x = length,
+                  linetype = Dataset,
                   group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
-  scale_x_log10(name = "Length (bp)") +
-  scale_y_continuous(name = "Fraction Greater than length")
+                  color = Region)) +
+  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
+  scale_x_log10(name = "Length (bp)",
+                limits = c(50, NA),
+                oob = squish) +
+  scale_y_continuous(name = "Fraction Greater than length") +
+  scale_color_viridis_d(option = "C") +
+  scale_linetype_discrete() +
+  theme_bw()
+```
 
+## Read length density
+```{r}
+ggplot(reads_subset, aes(x = length,
+                  linetype = Dataset,
+                  group = paste(Dataset, Region),
+                  color = Region)) +
+  geom_line(stat = "density") +
+  scale_x_log10(name = "Length (bp)",
+                limits = c(50, NA),
+                oob = squish) +
+  scale_y_continuous(name = "count") +
+  scale_color_viridis_d(option = "C") +
+  scale_linetype_discrete() +
+  theme_bw()
 ```
@@
 ```
+
+## ASV correspondences
+
 ```{r eval = FALSE}
@@
 
-IonTorrent reads which match a PacBio ASV: ` filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
-IonTorrent reads which do NOT match PacBio: ` filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
-PacBio reads which match an IonTorrent ASV: ` filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
-PacBio reads which do NOT match IonTorrent: ` filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`
+- IonTorrent reads which match a PacBio ASV: ` filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
+- IonTorrent reads which do NOT match PacBio: ` filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
+- PacBio reads which match an IonTorrent ASV: ` filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
+- PacBio reads which do NOT match IonTorrent: ` filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`
 

### 2019-02-21 scripts/plate_check.R

@@ -36,7 +36,7 @@ read_platemap <- function(filename, sheet, skip = 2) {
     # get rid of empty columns
     discard(~all(is.na(.))) %>%
     # make nice names
-    rename(Comment = V1,
+    dplyr::rename(Comment = V1,
            Well = row,
            DNA.Conc = "Conc. (ng/µl)",
            DNA.Tot = DNA,
@@ -70,5 +70,5 @@ read_platemap <- function(filename, sheet, skip = 2) {
                     levels = c("Blank", "Pos", "Pos. Kontroll")) %>%
              fct_collapse(Blank = "Blank", Pos = c("Pos", "Pos. Kontroll")) %>%
              fct_explicit_na("Sample")) %>%
-    mutate_at(Plate, formatC, width = 3, flag = "0")
+    mutate_at("Plate", formatC, width = 3, flag = "0")
 }

### 2019-02-23 build.sh
switch back to using UPPMAX module system for texlive because the conda version doesn't seem to work.

@@ -30,6 +30,8 @@
 # always redo this makefile, in case something (number of cores!) changes.
 rm -f demux.make &&
 
+## texlive from conda isn't functional
+module load texlive &&
 conda activate oueme1 &&
 
 # convert PacBio RSII files to .bam format

### 2019-02-23 writing/qual-check.Rmd
switch back to using UPPMAX module system for texlive because the conda version doesn't seem to work.
Use global theming

@@
 knitr::opts_chunk$set(echo = FALSE, device = "tikz")
+theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
 ```
@@
   scale_color_viridis_d(option = "C") +
-  scale_linetype_discrete() +
-  theme_bw()
+  scale_linetype_discrete()
 ```

### 2019-03-18 writing/transect_paper.Rmd
Start of methods outline

@@
 
+## Sampling
+
+## Soil Chemistry
+
+## DNA extraction, PCR, and sequencing
+
+  - lysis in field Zymo Research Xpedition kit
+  - short amplicon : gITS7 -- ITS4 (multiplexing index on gITS7)
+  - long amplicon : ITS1 -- LR5 (multiplexing index on both ends)
+  - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5
+  - long amplicon : PacBio RS II
+
+## Bioinformatics
+
+  - Two workflows for short reads:
+    - Mohammad's workflow in Pipecraft
+    - Brendan's workflow using Snakemake/Drake
+
+  - Brendan's pipeline:
+      - Ion Torrent
+          - used demultiplexing from SciLife (standard IonXpress)
+          - primers removed with `cutadapt` [@martin2011]
+      - PacBio
+          - CCS basecalls re-done with `ccs` version ___ [@pacificbiosciences2019]
+          - demultiplexed and primers removed with `cutadapt` [@martin2011]
+      - ITS1, ITS2, ITS, LSU extracted using ITSx [@Nilsson2010, @bengtsson-palme2013]
+      - extracted regions independently quality filtered by max expected error using `dada2::filterAndTrim()`
+      - dereplicated using `dada2::derepFastq()`
+      - denoised with `dada2::dada()`
+      - de novo chimera removal using `dada2::removeBimeraDenovo()`
+      - taxonomy assigned using `dada2::assignTaxonomy()` with Unite database (ITS, ITS1, ITS2, long, short)[@nilsson2019a], RDP (LSU, long)[@cole2014], and Silva (LSU, long) databases
+## Statistics
+
 # Results

### 2019-04-26 writing/group_presentation.Rnw
Presentation for Anna's group

\documentclass[10]{beamer}
\include{pres_preamble}

\title{Turnover scale of soil fungal community composition}
\subtitle{Presentation for Anna's Group Meeting}
\author{Brendan Furneaux}
\date{26 April, 2019}

\begin{document}

\frame{\titlepage}

<<setup, include=FALSE>>=
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
library(drake)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggtree)
library(castor)
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
loadd(bc_corr, bc_corr2, unif_corr, unif_corr2, qstats_join, datasets, longtree,
      taxon_ITS2_unite)
reads <- qstats_join %>%
  mutate_at("Region", factor,
            levels = c("ITS2", "ITS1", "short", "ITS", "LSU", "long")) %>%
  mutate_at("Dataset", factor,
            levels = c("short-ion", "short-pacbio", "long-pacbio")) %>%
  group_by(Dataset, Plate, Region)
reads_subset <- reads %>%
  do(if (nrow(.) > 10000) sample_n(., 10000) else .)
reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, 
              log_breaks(base = base), 
              domain = c(1e-100, Inf))
}
@

%\begin{frame}{Outline}
%\tableofcontents
%\end{frame}

\begin{frame}{Introduction}

\begin{itemize}
 \item Question: What is the scale of spatial (and temporal) autocorrelation in soil fungal community composition?
  \begin{itemize}
    \item Specifically in a tropical ECM forest.
  \end{itemize}
 \item Metabarcoding method comparison
\end{itemize}
\end{frame}

\begin{frame}{Upper Ouémé Forest Reserve}
\framesubtitle{Permanent study plots}
\vskip1.5em
\begincols
  \begincol{0.55\linewidth}
    \begin{figure}
      \begin{overprint}
      \onslide<1>
        \includegraphics{soudanian_benin}
      \onslide<2>
        \includegraphics{soudanian_benin_oueme}
      \onslide<3->
        \includegraphics{soudanian_benin_oueme_village}
      \end{overprint}
      \caption{Location of study sites in Bènin.}
    \end{figure}
  \endcol

  \begincol{0.45\linewidth}
  \begin{itemize}
    \tightlist
    \item<1-|alert@1> West Soudanian woodlands in central Bènin
    \item<1-|alert@1> 10° N Latitude
    \item<3-|alert@3> Two sites located near the villages Angaradebou and Gando
    \item<4-|alert@4> Dominated by \emph{Isoberlinia doka} (Caesalpinoideae; ECM host)
  \end{itemize}
  \endcol
\endcols
\end{frame}

\begin{frame}{Sampling}
  \begin{itemize}
    \item<1-|alert@1> \SI{25}{\m} transects.
    \item<1-|alert@1> Soil samples taken every 1 m.
    \item<2-|alert@2> $\SI{5}{\cm}\times\SI{5}{\cm}\times\SI{5}{\cm}$ sample, homogenize in plastic bag, then subsample \SI{50}{\mg}
    \item<3-|alert@3> Lysis \& preservation in field
    \item<4-|alert@4> Same locations in two years -- 2015 and 2016 (25 points × 2 years × 2 sites = 100 samples)
  \end{itemize}
\end{frame}

\begin{frame}{Sequencing}
  \framesubtitle{Which region?}
  \small
  \begincols
  \begincol{0.5\linewidth}
    \begin{block}{ITS2 (gITS7-ITS4)}
        \begin{itemize}
          \tightlist
          \item Internally transcribed spacer between 5.8S and 26/28S (LSU) subunits of ribosomal RNA
          \item Usually 250--350 bp.
          \item ITS --- standard DNA barcode for fungi.
          \item Relatively good species-level resolution, and good database support.
          \item Cannot be aligned across larger taxonomic groups.
          \item Groups with variant (or absent) 5.8S not sequenced.
        \end{itemize}
      \end{block}
    \endcol
    \begincol{0.5\linewidth}
      \begin{block}{Full ITS + partial LSU (ITS1-LR5)}
        \begin{itemize}
          \tightlist
          \item Full ITS plus the 5' end of LSU; 1500-2000 bp.
          \item ITS -- resolution to separate closely related taxa.
          \item LSU -- can be aligned across fungi.
          \item Primers are more universal than gITS7.
          \item Long amplicon with alternating variable and conserved regions --- susceptible to chimera formation and breakage.
        \end{itemize}
      \end{block}
    \endcol
  \endcols
\end{frame}

\begin{frame}{Sequencing}
  \framesubtitle{Which platform?}
  \begincols
  \begincol{0.5\linewidth}
    \begin{block}{IonTorrent}
      \begin{itemize}
        \tightlist
        \item Low cost per bp.
        \item Smaller minimum job than Illumina.
        \item Needs less DNA than PacBio.
        \item Bias against amplicons greater than 400 bp.
         \item Cannot sequence long ITS+LSU amplicon.
       \end{itemize}
     \end{block}
  \endcol
  \begincol{0.5\linewidth}
     \begin{block}{PacBio}
        \begin{itemize}
          \tightlist
          \item Higher cost per bp $\implies$ fewer samples or less sequencing depth for same budget.
          \item Requires more DNA.
          \item Good accuracy requires multiple circular reads.
        \end{itemize}
      \end{block}
  \endcol
  \endcols
  \centering

\end{frame}

\begin{frame}{Sequencing}
  \begin{block}{The plan}
  Three sequencing runs:
  \begin{itemize}
    \item Ion Proton S5, short amplicon
    \begin{itemize}
      \item Pool all samples on one chip
    \end{itemize}
    \item PacBio RSII, short amplicon
    \begin{itemize}
      \item Same amplicon library as IonTorrent.
      \item Two sequencing libraries, two SMRT cells per library.
    \end{itemize}
    \item PacBio RSII, long amplicon
    \begin{itemize}
      \item Two sequencing libraries, four SMRT cells per library.
    \end{itemize}
  \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Sequencing}
  \begin{block}{What happened}
  Three sequencing runs:
  \begin{itemize}
    \item \sout{IonTorrent, short amplicon}
    \begin{itemize}
      \item Accidentally used primers with IonTorrent adapters built in.
      \item Lost some of the multiplexing barcodes.
      \item Can't demultiplex the samples --- can only look at pool of all reads.
    \end{itemize}
    \item PacBio, short amplicon
    \begin{itemize}
      \item OK
    \end{itemize}
    \item PacBio, long amplicon
    \begin{itemize}
      \item OK
    \end{itemize}
  \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Bioinformatics}
  \begin{itemize}
    \item Redo RSII circular consensus basecalling using new algorithm (Arrow).
    \item Demultiplex and trim primers.
    \item Cut out regions (ITS1, ITS2, full ITS, LSU) using ITSx.
    \item Quality filter.
  \end{itemize}
\end{frame}

\begin{frame}{Expected error rate}
\framesubtitle{After \texttt{Arrow}}

<<errorrate, fig.width = 12>>=
ggplot(reads_subset, aes(x = erate,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors per base",
                     trans = reverselog_trans(10),
                     limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete()
@
\end{frame}

\begin{frame}{Bioinformatics}
\framesubtitle{continued...}
  \begin{itemize}
    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
      \begin{itemize}
        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
        \item Works on the smaller regions.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Expected total number of errors}
\framesubtitle{After \texttt{Arrow}}

<<totalerror, fig.width = 12>>=
ggplot(reads_subset, aes(x = eexp,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors",
                     trans = reverselog_trans(10),
                     limits = c(NA, 0.1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete()
@
\end{frame}

\begin{frame}{Bioinformatics}
\framesubtitle{continued...}
  \begin{itemize}
    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
      \begin{itemize}
        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
        \item Works on the smaller regions.
      \end{itemize}
    \item For each ITS2 ASV, align the whole long read and find the consensus.
      \begin{itemize}
        \item There were a lot of chimeras (15--25\%).
        \item Remove "outliers" from the alignments using \texttt{odseq}.
        \item Remaining chimeras are too few to change the consensus.
      \end{itemize}
    \item Align with \texttt{DECIPHER}
      \begin{itemize}
        \item Progressive aligner with inferred RNA secondary structure.
      \end{itemize}
    \item \texttt{RAxML} tree
  \end{itemize}
\end{frame}

\begin{frame}{Taxonomy assignment}
Both \texttt{dada2} and \texttt{DECIPHER} can do taxonomy assignment.
\begincols
\begincol{0.5\framewidth}
\begin{block}{dada2}
  \begin{itemize}
    \item Naïve Bayesian classifier.
    \item Uses a lot of memory.
    \item Needs every rank to be filled (\emph{incertae sedis}, "unidentified").
    \item Seems to put thins in places that don't match the tree.
  \end{itemize}
\end{block}
\endcol

\begincol{0.5\framewidth}
\begin{block}{DECIPHER}
  \begin{itemize}
    \item Some kind of machine learning algorithm.
    \item Uses less memory.
    \item Allows uneven classification depths.
    \item Very low bootstraps, especially for ITS.
  \end{itemize}
\end{block}
\endcol
\endcols

UNITE for ITS, RDP for LSU...  But there is no large database that has both for the same specimens!

I'm still working on it...
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Kingdoms}
<<kingtree, fig.width = 12, message=FALSE>>=
seqhash <- function(seq, len = 8) UseMethod("seqhash")
seqhash.character <- function(seq, len = 8) {
  h <- purrr::map_chr(seq, digest::digest)
  stringr::str_sub(h, end = len)
}
seqhash.XStringSet <- function(seq, len = 8) {
  seqhash.character(as.character(seq))
}
treetax <- taxon_ITS2_unite %>%
  mutate_at("seq", seqhash) %>%
  select(-Taxonomy, -nreads) 
outgroups <- treetax %>% filter(!is.na(Kingdom), Kingdom %in% c("Alveolata", "Chromista", "Plantae"), seq %in% longtree$tip.label)
p <- ggtree(tr = longtree)
longtree <- reroot(longtree, MRCA(p, outgroups$seq))

ggtree(groupOTU(longtree, split(treetax$seq, treetax$Kingdom))) +
  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Phyla}
<<phytree, fig.width = 12, message=FALSE>>=
ggtree(groupOTU(longtree, split(treetax$seq, treetax$Phylum))) +
  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Families --- Basidiomycota}
<<basidtree, fig.width = 12, message=FALSE>>=
basids <- treetax %>% filter(!is.na(Class),
                             Class %in% c("Geminibasidiomycetes",
                                         "Tremellomycetes",
                                         "Exobasidiomycetes",
                                         "Microbotryomycetes"),
                             seq %in% longtree$tip.label)

basidtree <- phylobase::phylo4(longtree) %>%
  phylobase::subset(mrca = basids$seq) %>%
  as("phylo")

basids2 <- treetax %>% filter(seq %in% basidtree$tip.label)

ggtree(groupOTU(basidtree, split(basids2$seq, basids2$Order)), layout = "circular") +
  aes(color = group) +
  scale_color_brewer(type = "qual", palette = "Paired", limits = unique(na.omit(basids2$Order))) +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Correlogram}
<<correlogram, fig.width = 12>>=
bind_rows(as.data.frame(bc_corr$mantel.res) %>%
            mutate(tdist = "same", metric = "bray"),
          as.data.frame(bc_corr2$mantel.res) %>%
            mutate(tdist = "diff", metric = "bray"),
          as.data.frame(unif_corr$mantel.res) %>%
            mutate(tdist = "same", metric = "unifrac"),
          as.data.frame(unif_corr2$mantel.res) %>%
            mutate(tdist = "diff", metric = "unifrac")) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Spatial Distance correlation") +
  scale_color_discrete(name = "Year") +
  xlab("distance (m)") +
  facet_wrap("metric") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
@
\end{frame}

\end{document}

### 2019-06-05 writing/coverage.Rmd

---
title: "Coverage tests"
author: "Brendan Furneaux"
date: "2/22/2019"
output: 
  beamer_presentation: 
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      external = FALSE,
                      device = "tikz")
library(magrittr)
library(drake)
library(ggplot2)
library(tidyverse)
theme_set(theme_bw(base_size = 24))

bigmaps <- drake::cached() %>%
  str_subset("dada_map_pb_500") %>%
  map(readd, character_only = TRUE)

pm <- readd(platemap) %>%
  mutate_at("Primer.Pair", tolower) %>%
  mutate_at(c("Year", "Site", "Qual", "Plate", "Well", "Primer.Pair"), factor)

st <- readd(big_seq_table_ITS2) %>%
  set_colnames(map_chr(colnames(.), digest::digest) %>% substr(1,8)) %>%
  as_tibble(rownames = "file") %>%
  tidyr::extract(
    col = "file",
    into = c("Seq.Run", "Plate", "Well", "Direction", "Region"),
    regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)([fr]?)-([:alnum:]+)\\.qfilt\\.fastq\\.gz") %>%
  group_by(Seq.Run, Plate, Well, Region) %>%
  summarize_if(is.numeric, sum) %>%
  ungroup() %>%
  left_join(readd(datasets)) %>%
  mutate(Primer.Pair = paste(str_replace(Forward, "_tag.*$", ""),
                             str_replace(Reverse, "_tag.*$", ""),
                             sep = "_")) %>%
  select(Seq.Run, Plate, Well, Region, Primer.Pair, everything()) %>%
  mutate_at(c("Seq.Run", "Region", "Primer.Pair"), factor) %>%
  mutate_at("Well", factor, levels = levels(pm$Well)) %>%
  mutate_at("Plate", factor, levels = levels(pm$Plate)) %>%
  gather(key = "ASV", value = "Reads",
         -one_of(c("Seq.Run", "Plate", "Well", "Region", "Primer.Pair", "Dataset", "Tech", "Runs", "Forward", "Reverse", "Regions", "PlateKey", "Homopolymer.Gap.Penalty", "Band.Size", "Pool"))) %>%
  mutate_at("Reads", as.integer) %>%
  filter(Reads > 0L)
```

# Accumulation curves

```{r coveragesample, results='asis'}
right_join(pm, st) %>%
    filter(!is.na(Year), Qual == "") %>%
    group_by(Site, X, Year, Seq.Run, ASV) %>%
    summarize(Reads = 1) %>%
    group_by(Site, Year, Seq.Run, ASV) %>%
    summarize(Reads = sum(Reads)) %>%
    spread(key = ASV, value = Reads, fill = 0L) %>%
    group_by(Site, Year) %>%
    group_walk(~ {cat("\n\n## Coverage by # samples | ",
                      as.character(.y$Site),
                      as.character(.y$Year), "\n\n")
      column_to_rownames(.x, "Seq.Run") %>%
                 as.matrix() %>%
                 t %>%
                 rbind(25, .) %>%
                 iNEXT::iNEXT(datatype = "incidence_freq") %>%
                 {iNEXT::ggiNEXT(.) +
                     xlab("Number of samples") +
                     ylab("ASV diversity") +
                     scale_fill_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_color_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_shape_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE)} %>%
                 plot})
```

```{r coveragereads, results='asis'}
right_join(pm, st) %>%
    filter(!is.na(Year), Qual == "") %>%
    group_by(Site, Year, Seq.Run, ASV) %>%
    summarize(Reads = sum(Reads)) %>%
    spread(key = ASV, value = Reads, fill = 0L) %>%
    group_by(Site, Year) %>%
    group_walk(~ {cat("\n\n## Coverage by # reads |",
                      as.character(.y$Site),
                      as.character(.y$Year), "\n\n")
      column_to_rownames(.x, "Seq.Run") %>%
                 as.matrix() %>%
                 t %>%
                 iNEXT::iNEXT(datatype = "abundance") %>%
                 {iNEXT::ggiNEXT(.) +
                     xlab("Number of reads") +
                     ylab("ASV diversity") +
                     scale_fill_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_color_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_shape_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE)} %>%
                 plot})
```
```{r coveragesamplereads, results='asis'}
d <- right_join(pm, st) %>%
    filter(!is.na(Year)) %>%
    group_by(Site, Year, X, Qual, Seq.Run, ASV) %>%
    summarize(Reads = sum(Reads)) %>%
    spread(key = ASV, value = Reads, fill = 0L) %>%
    group_by(Site, X)
cats <- ungroup(d) %>%
  select(Year, Seq.Run, Qual) %>%
  unique() %>%
  glue::glue_data("{Year}_{Seq.Run}_{Qual}")

group_walk(d, 
           ~ {cat("\n\n## Coverage by # reads |",
                  as.character(.y$Site),
                  "sample",
                  as.character(.y$X),
                  "\n\n")
             unite(.x, "Seq.Run", Year, Seq.Run, Qual) %>%
               column_to_rownames("Seq.Run") %>%
               as.matrix() %>%
               t %>%
               iNEXT::iNEXT(datatype = "abundance") %>%
               {iNEXT::ggiNEXT(.) +
                   xlab("Number of reads") +
                   ylab("ASV diversity") +
                   scale_fill_discrete(limits = cats,
                                       drop = FALSE) +
                   scale_color_discrete(limits = cats,
                                        drop = FALSE) +
                   scale_shape_discrete(limits = cats,
                                        drop = FALSE)} %>%
               plot})
```

# Resampling

## Read counts per site×sample×year

```{r}
right_join(pm, st) %>%
  filter(startsWith(as.character(Seq.Run), "pb"),
         !is.na(Year)) %>%
  group_by(Year, Site, X, Qual, Dataset) %>%
  summarize(Ntot = sum(Reads)) %>%
  ggplot(aes(x = Ntot, group = Dataset, color = Dataset)) +
  geom_line(mapping = aes(y = 1 - ..y..), stat = "ecdf") +
  xlab("Number of reads") +
  ylab("Fraction of samples")
```

## Read counts per site×sample

```{r}
right_join(pm, st) %>%
  filter(startsWith(as.character(Seq.Run), "pb"),
         !is.na(Year)) %>%
  group_by(Site, X, Qual, Dataset) %>%
  summarize(Ntot = sum(Reads)) %>%
  ggplot(aes(x = Ntot, group = Dataset, color = Dataset)) +
  geom_line(mapping = aes(y = 1 - ..y..), stat = "ecdf") +
  xlab("Number of reads") +
  ylab("Fraction of samples")
```

## New vs. expired buffer

 - in 2016, there was not enough unexpired field lysis buffer for all samples
 - sampled with expired buffer at all locations
 - sampled with unexpired buffer at approx. 1/3 of all locations.

## Reads: new vs. expired buffer

```{r}
right_join(pm, st) %>%
  filter(!is.na(Year), Qual != "*") %>%
  group_by(Dataset, Year, X, Site) %>%
  filter("" %in% Qual & "X" %in% Qual) %>%
  group_by(Dataset, Year, X, Site, Qual) %>%
  summarize(Reads = sum(Reads)) %>%
  mutate_at("Qual", forcats::fct_recode, Expired = "", New = "X") %>%
  spread(key = Qual, value = Reads, fill = 0L) %>%
  ggplot(aes(x = Expired, y = New, color = Dataset)) +
  geom_point() +
  geom_abline(slope = 1, linetype = 2, alpha = 0.5) +
  coord_fixed() +
  xlab("Expired buffer") +
  ylab("New buffer")
```

## Reads: new vs expired buffer

```{r}
right_join(pm, st) %>%
  filter(!is.na(Year), Qual != "*") %>%
  group_by(Dataset, Year, X, Site) %>%
  filter("" %in% Qual & "X" %in% Qual) %>%
  group_by(Dataset, Year, X, Site, Qual) %>%
  summarize(Reads = sum(Reads)) %>%
  mutate_at("Qual", forcats::fct_recode, Expired = "", New = "X") %>%
  spread(key = Qual, value = Reads, fill = 0L) %>%
  filter(New > 20, Expired > 20) %>%
  mutate(Ratio = New / Expired) %>%
  ggplot(aes(y = Ratio, x = Dataset)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, alpha = 0.5, linetype = 2)
```

### 2019-06-05 writing/group_presentation.Rnw

@@
 loadd(bc_corr, bc_corr2, unif_corr, unif_corr2, qstats_join, datasets, longtree,
-      taxon_ITS2_unite)
+      taxon_LSUcons_rdp)
 reads <- qstats_join %>%
@@
 }
-treetax <- taxon_ITS2_unite %>%
+treetax <- taxon_LSUcons_rdp %>%
   mutate_at("seq", seqhash) %>%

### 2019-06-05 writing/transect_paper.Rmd

--- 2019-01-26 writing/transect_paper.Rmd
@@
 
+## Sampling
+
+  - Ouémé Forest reserve
+  - Two locations in _Isoberlinia doka_ forest, approx. 30km apart
+  - transect: 25m, sampled at interval of 1m
+  - Sampled in May 2015 and June 2016
+  - remove coarse debris (typically very little) and take ~5cm×5cm×5cm sample
+  - homogenize by hand in ziplock bag
+  - take $\approx$ 50mg, lysis in field with Terralyzer
+  - remainder of sample dried at 40°C with electric dryer within 24-48 hours
+
+## Soil Chemistry
+
+  - Need to get methods from Nourou
+
+## DNA extraction, PCR, and sequencing
+
+  - lysis in field Zymo Research Xpedition kit
+  - short amplicon : gITS7 -- ITS4 (multiplexing index on gITS7)
+  - long amplicon : ITS1 -- LR5 (multiplexing index on both ends)
+  - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5
+  - long amplicon : PacBio RS II
+
+## Bioinformatics
+
+  - Two workflows for short reads:
+    - Mohammad's workflow in Pipecraft
+    - Brendan's workflow using Snakemake/Drake
+
+  - Mohammad's pipeline:
+
+  - Brendan's pipeline:
+      - Ion Torrent
+          - used demultiplexing from SciLife (standard IonXpress)
+          - primers removed with `cutadapt` [@martin2011]
+      - PacBio
+          - CCS basecalls re-done with `ccs` version ___ [@pacificbiosciences2019]
+          - demultiplexed and primers removed with `cutadapt` [@martin2011]
+      - ITS1, ITS2, ITS, LSU extracted using ITSx [@Nilsson2010; @bengtsson-palme2013]
+      - extracted regions independently quality filtered by max expected error using `dada2::filterAndTrim()`
+      - dereplicated using `dada2::derepFastq()`
+      - denoised with `dada2::dada()`
+      - de novo chimera removal using `dada2::removeBimeraDenovo()`
+      - taxonomy assigned using IDTAXA [@murali2018]
+          - UNITE version 8 [@nilsson2019a]
+              - extracted ITS1 and ITS2 using ITSx [@Nilsson2010; @bengtsson-palme2013]
+              - trained three classifiers: full ITS, ITS1, ITS2.
+          - RDP fungal LSU release 11.5 [@cole2014]
+              - chose only sequences containing conserved primer sites for ITS4 (3' side of H11; CAUAUnAnUnAGnSSAGGA) and LR5 (on H37; CGAAnUUUCnCUCAGGAUAGCnG).
+              - ambiguities based on [@cannone2002; @tedersoo2015]
+              - Trim all bases after LR5 because they are not present in our amplicons.
+          - SILVA LSU release 132 [@quast2013; @glockner2017; @yilmaz2014]
+              - filter out sequences annotates as prokaryotes, then treat as RDP.
+          - all databases
+              - select medioid sequence for each annotated species, based on minimum 5-mer distance to other sequences
+              - for sequences not annotated as species (e.g. "unidentified Basidiomycota"), randomly choose up to 20 per taxon
+              - Use up to 5 iterations of IDTAXA training, removing all "problem" sequences on each iteration, unless last member of annotated taxon.
+              - "underannotated" sequences (e.g., "unidentified Basidiomycota" confidently placed as Agaricomycetes) not protected from removal if last sequence annotated as such.
+      - ecological guild assigned using FUNGuild [@nguyen2016funguild]
+      - consensus long read and LSU region constructed for each ITS2 ASV:
+          - align reads using `DECIPHER` with RNA secondary structure calculation [@wright2015;@wright2016]
+          - remove outliers (i.e. chimeras) with `odseq` [@jehl2015]
+          - remove gap columns (>50%)
+          - take consensus (>50%) at non-gaps
+          - remove sequences with more than 3 ambiguous bases (bad consensus)
+      - initial alignment of long amplicons with DECIPHER [@wright2015;@wright2016]
+      - alignment/tree for long amplicons built iteratively with PASTA[@mirarab2014; @mirarab2014a]. Fastree tree, mafft align, OPAL merge, RaxML final tree
+
+## Statistics
+
 # Results

### 2019-06-10 writing/tech_compare.Rmd
RMD file to compare ASVs for each sequencing technology

---
title: "Read quality check"
output:
  beamer_presentation:
    latex_engine: xelatex
---

```{r setup, include = FALSE}
library(magrittr)
library(tidyverse)
library(here)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")
library(drake)
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
```

# Read number from PacBio vs. IonTorrent ASV-OTUs (ITS2)

```{r}
otu <- read_tsv(here("data/clusters/ITS2.table"),
                col_types = "ciii") %>% select(is_057, pb_500) %>%
  filter(rowSums(.) > 0)
otu %>%
  mutate(is_057 = ifelse(is_057 == 0, 0.6, is_057),
         pb_500 = ifelse(pb_500 == 0, 0.6, pb_500)) %>%
  ggplot(aes(x = is_057, y = pb_500)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Ion Torrent Reads") +
  ylab("PacBio Reads")
```

# IonTorrent ITS2 ASV-OTUs: coverage by PacBio long reads
```{r}
otu %>%
  arrange(desc(is_057)) %>%
  filter(is_057 > 0) %>%
  ggplot(aes(y = is_057, x = seq_along(is_057), fill = pb_500 > 0)) +
  geom_col(width = 1) +
  xlab("OTUs") +
  ylab("reads") +
  scale_fill_discrete(name = "PacBio")
```
`r otu %>% filter(is_057 > 0) %>% {formatC(sum(.$pb_500 > 0) / length(.$is_057)*100, digits = 1, flag = "0", format = "f")}`% of IonTorrent OTUs also include at least one PacBio ASV.

# PacBio ITS2 ASV-OTUs: coverage by Ion Torrent
```{r}
otu %>%
  arrange(desc(pb_500)) %>%
  filter(pb_500 > 0) %>%
  ggplot(aes(y = pb_500, x = seq_along(pb_500), fill = is_057 > 0)) +
  geom_col(width = 1) +
  xlab("OTUs") +
  ylab("reads") +
  scale_fill_discrete(name = "Ion Torrent")
```
`r otu %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio OTUs also include at least one IonTorrent ASV

### 2019-06-10 writing/tech_compare.Rmd
Add ASV comparison, fractions by read, secondary axis with community fraction

@@
 
+# Read number from PacBio vs. IonTorrent ASVs (ITS2)
+```{r}
+loadd(big_seq_table_ITS2)
+asv_table <- big_seq_table_ITS2 %>%
+  tibble::as_tibble(rownames = "filename") %>%
+  tidyr::gather(key = "seq", value = "reads", -1) %>%
+  dplyr::filter(reads >= 1) %>%
+  tidyr::extract(col = "filename",
+                 into = c("seq.run", "plate", "well", "dir", "region"),
+                 regex = "([a-z]+_\\d+)_(\\d+)-([A-H]1?[0-9])([fr]?)-([:alnum:]+).+") %>%
+  dplyr::filter(seq.run %in% c("is_057", "pb_500")) %>%
+  dplyr::group_by(seq.run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq.run, value = reads, fill = 0)
+
+asv_table %>% 
+  mutate(is_057 = ifelse(is_057 == 0, 0.6, is_057),
+                    pb_500 = ifelse(pb_500 == 0, 0.6, pb_500)) %>%
+  ggplot(aes(x = is_057, y = pb_500)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  xlab("Ion Torrent Reads") +
+  ylab("PacBio Reads")
+
+```
+
 # Read number from PacBio vs. IonTorrent ASV-OTUs (ITS2)
@@
 
+# IonTorrent ITS2 ASVs: coverage by PacBio long reads
+```{r}
+asv_table %>%
+  arrange(desc(is_057)) %>%
+  filter(is_057 > 0) %>%
+  ggplot(aes(y = is_057, x = seq_along(is_057), fill = pb_500 > 0)) +
+  geom_col(width = 1) +
+  xlab("ASVs") +
+  ylab("reads") +
+  scale_fill_discrete(name = "PacBio") +
+  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
+```
+`r asv_table %>% filter(is_057 > 0) %>% {formatC(sum(.$pb_500 > 0) / length(.$is_057)*100, digits = 1, flag = "0", format = "f")}`% of IonTorrent ASVs (`r asv_table %>% {formatC(sum(.$is_057[.$pb_500 > 0]) / sum(.$is_057) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one PacBio read.
+
 # IonTorrent ITS2 ASV-OTUs: coverage by PacBio long reads
@@
   ylab("reads") +
-  scale_fill_discrete(name = "PacBio")
+  scale_fill_discrete(name = "PacBio") +
+  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
 ```
-`r otu %>% filter(is_057 > 0) %>% {formatC(sum(.$pb_500 > 0) / length(.$is_057)*100, digits = 1, flag = "0", format = "f")}`% of IonTorrent OTUs also include at least one PacBio ASV.
+`r otu %>% filter(is_057 > 0) %>% {formatC(sum(.$pb_500 > 0) / length(.$is_057)*100, digits = 1, flag = "0", format = "f")}`% of IonTorrent OTUs (`r otu %>% {formatC(sum(.$is_057[.$pb_500 > 0]) / sum(.$is_057) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one PacBio ASV.
+
+# PacBio ITS2 ASVs: coverage by Ion Torrent
+```{r}
+asv_table %>%
+  arrange(desc(pb_500)) %>%
+  filter(pb_500 > 0) %>%
+  ggplot(aes(y = pb_500, x = seq_along(pb_500), fill = is_057 > 0)) +
+  geom_col(width = 1) +
+  xlab("ASVs") +
+  ylab("reads") +
+  scale_fill_discrete(name = "Ion Torrent") +
+  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
+```
+`r asv_table %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio ASVs (`r asv_table %>% {formatC(sum(.$pb_500[.$is_057 > 0]) / sum(.$pb_500) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one IonTorrent read
 
@@
   ylab("reads") +
-  scale_fill_discrete(name = "Ion Torrent")
+  scale_fill_discrete(name = "Ion Torrent") +
+  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
 ```
-`r otu %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio OTUs also include at least one IonTorrent ASV
+`r otu %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio OTUs (`r otu %>% {formatC(sum(.$pb_500[.$is_057 > 0]) / sum(.$pb_500) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one IonTorrent ASV

### 2019-06-10 writing/tech_compare.Rmd
Change titles so everything fits on the page

@@
 
-# IonTorrent ITS2 ASVs: coverage by PacBio long reads
+# IonTorrent ITS2 ASVs: coverage by long PacBio
 ```{r}
@@
 
-# IonTorrent ITS2 ASV-OTUs: coverage by PacBio long reads
+# IonTorrent ITS2 ASV-OTUs: coverage by long PacBio
 ```{r}

### 2019-06-10 writing/tech_compare.Rmd
improve graphs, add sample coverage analysis

@@
 library(drake)
+
+source(here("scripts/map_LSU.R"))
 knitr::opts_chunk$set(echo = FALSE, device = "tikz")
@@
 asv_table %>% 
-  mutate(is_057 = ifelse(is_057 == 0, 0.6, is_057),
-                    pb_500 = ifelse(pb_500 == 0, 0.6, pb_500)) %>%
+  filter(is_057 > 0, pb_500 > 0) %>%
   ggplot(aes(x = is_057, y = pb_500)) +
-  geom_point() +
+  geom_point(alpha = 0.2) +
   scale_x_log10() +
@@
   xlab("Ion Torrent Reads") +
-  ylab("PacBio Reads")
+  ylab("PacBio Reads") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = asv_table %>% filter(is_057 == 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = asv_table %>% filter(pb_500 == 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5)
 
@@
 otu %>%
-  mutate(is_057 = ifelse(is_057 == 0, 0.6, is_057),
-         pb_500 = ifelse(pb_500 == 0, 0.6, pb_500)) %>%
+  filter(is_057 > 0, pb_500 > 0) %>%
   ggplot(aes(x = is_057, y = pb_500)) +
-  geom_point() +
+  geom_point(alpha = 0.2) +
   scale_x_log10() +
@@
   xlab("Ion Torrent Reads") +
-  ylab("PacBio Reads")
+  ylab("PacBio Reads") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = otu %>% filter(is_057 == 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = otu %>% filter(pb_500 == 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5)
 ```
@@
 ```
-`r otu %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio OTUs (`r otu %>% {formatC(sum(.$pb_500[.$is_057 > 0]) / sum(.$pb_500) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one IonTorrent ASV
+`r otu %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio OTUs (`r otu %>% {formatC(sum(.$pb_500[.$is_057 > 0]) / sum(.$pb_500) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one IonTorrent ASV.
+
+# ASV sample coverage
+```{r}
+asv_samps <- big_seq_table_ITS2 %>%
+  tibble::as_tibble(rownames = "filename") %>%
+  tidyr::gather(key = "seq", value = "reads", -1) %>%
+  tidyr::extract(col = "filename",
+                 into = c("seq.run", "plate", "well", "dir", "region"),
+                 regex = "([a-z]+_\\d+)_(\\d+)-([A-H]1?[0-9])([fr]?)-([:alnum:]+).+") %>%
+  dplyr::filter(reads >= 1, seq.run %in% c("pb_500", "is_057")) %>%
+  dplyr::group_by(seq.run, well, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  dplyr::ungroup()
+
+asv_samp_plot <- asv_samps %>%
+  dplyr::group_by(seq.run, seq) %>%
+  dplyr::summarize(sample = n_distinct(well)) %>%
+  tidyr::spread(key = seq.run, value = sample, fill = 0)
+
+asv_samp_plot %>%
+  filter(is_057 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_500)) +
+  geom_point(alpha = 0.2) +
+  scale_x_log10() +
+  scale_y_log10() +
+  xlab("Ion Torrent wells") +
+  ylab("PacBio wells") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = asv_samp_plot %>% filter(is_057 == 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = asv_samp_plot %>% filter(pb_500 == 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500)) -
+                    mean(log10(is_057)), slope = 1),
+              linetype = 2L, alpha = 0.5)
+
+```
+
+# OTU Sample coverage
+
+```{r}
+otumap = read_tsv(here("data/clusters/ITS2.uc"),
+                  col_names = c("Type", "ID", "Length", "Identity", "Strand",
+                                "X1", "X2", "Align", "Query", "Target"),
+                  col_types = "cccccccccc") %>%
+  dplyr::select(otu = ID, seq = Query) %>%
+  dplyr::mutate_at("seq", stringr::str_remove_all, ";.+") %>%
+  unique()
+
+otu_samps <- asv_samps %>%
+  dplyr::mutate_at("seq", seqhash) %>%
+  dplyr::left_join(otumap, by = "seq")
+
+otu_samp_plot <- otu_samps %>%
+  dplyr::group_by(seq.run, otu) %>%
+  dplyr::summarize(sample = n_distinct(well)) %>%
+  tidyr::spread(key = seq.run, value = sample, fill = 0)
+
+otu_samp_plot %>%
+  filter(is_057 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_500)) +
+  geom_point(alpha = 0.2) +
+  scale_x_log10() +
+  scale_y_log10() +
+  xlab("Ion Torrent wells") +
+  ylab("PacBio wells") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = otu_samp_plot %>% filter(is_057 == 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = otu_samp_plot %>% filter(pb_500 == 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500)) -
+                    mean(log10(is_057)), slope = 1),
+              linetype = 2L, alpha = 0.5)
+```

### 2019-06-10 writing/tech_compare.Rmd
Slide on costs for future plans

@@
 ---
-title: "Read quality check"
+title: "Comparison of ASVs and OTUs between PacBio and Ion Torrent"
+author: "Brendan Furneaux"
+date: "10 June 2019"
 output:
@@
 ```
+
+# Costs going forward
+
+ - SciLife still has the short amplicon libraries for Ion Torrent
+ - Alternative 1: So we can resequence them on two Ion Torent S5 530 chips and demultiplex
+   - Cost (my best guess) $2 \text{ runs} \times 14000 \frac{\text{sek}}{\text{run}} = 28000 \text{ sek}$
+ - Alternative 2: Pool these two libraries (total ~150 samples) with ITS2 library from spatial paper (same primers, ~100 samples), sequence on one Illumina MiSeq 2×300bp lane (~18M read pairs)
+   - Compare total community recovered from PacBio, Ion Torrent, and Illumina on the transect samples
+   - Cost (best guess) $3 \text{ libraries} \times 1000 \frac{\text{sek}}{\text{ library}} + 1 \text{ run} \times 17000 \frac{\text{sek}}{\text{run}} = 20000 \text{ sek}$
+ - Alternative 3: Forget the libraries from the transect paper and sequence a long amplicon library for the spatial paper on Sequel
+    - We currently have 18 pools of 5 samples each. But not demultiplexed yet!
+    - Cost - ask Anna about Peter's sequencing cost.

### 2019-06-10 writing/tech_compare.Rmd
switch to log-scale community graphs

@@
   scale_fill_discrete(name = "PacBio") +
-  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
+  scale_y_continuous(trans = log10_trans(), sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
 ```
@@
   scale_fill_discrete(name = "PacBio") +
-  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
+  scale_y_continuous(trans = log10_trans(),
+                     sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
 ```
@@
   scale_fill_discrete(name = "Ion Torrent") +
-  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
+  scale_y_continuous(trans = log10_trans(),
+                     sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
 ```

### 2019-06-10 writing/tech_compare.Rmd
Move guide line to have 0 intercept

@@
   stat_smooth(method = "loess", formula = y ~ x) +
-  geom_abline(aes(intercept = mean(log10(pb_500)) -
-                    mean(log10(is_057)), slope = 1),
+  geom_abline(aes(intercept = 0, slope = 1),
               linetype = 2L, alpha = 0.5)

### 2019-08-11 writing/MSA2019_presentation.Rnw
copy presentation for Anna's group as base for MSA presentation.

--- 2019-06-05 writing/group_presentation.Rnw
@@
 
-\title{Turnover scale of soil fungal community composition}
+\title{Comparison of long and short amplicon metabarcoding to determination of fungal species diversity and spatial turnover in an ectomycorrhizal West African woodland}
 \subtitle{Presentation for Anna's Group Meeting}
-\author{Brendan Furneaux}
-\date{26 April, 2019}
+\author{Brendan Furneaux, Mohammad Bahram, Nourou Yorou, Anna Rosling, Martin Ryberg}
+\date{13 August, 2019}
 
@@
 @
+
+# Metabarcoding
+
+  - Cloning + Sanger -> 454 -> Illumina
+  - For fungi, usually ITS2
+      - short
+      - intraspecific variation in most groups
+      - good database support (Unite 8.0 - 817 130 fungal sequences)
+      - ITS is "official barcode for fungi" \cite{schoch2012}
+      - fungal specific primers
+  - problems
+      - not alignable
+      - cannot place unknown sequences
+      - variant 5.8S
+      - variable length
+
+  - LSU (26S/28S)
+      - alternating highly conserved domains and variable regions
+      - widely used for phylogenies
+      - universal primers
+      - less variable length
+      - more conserved than ITS
+      - not as good database support (RDP 11 fungal LSU - 125 525, Silva 132 LSU 907 382 of which )
+  - SSU (18S)
+      - alternating highly conserved and variable regions
+      - more conserved than LSU
+      - used for barcoding prokaryotes and "protists"
+      - good database support (Silva SSU), but not as good for fungi
 

### 2019-08-12 writing/MSA2019_presentation.Rnw
add author affiliations

--- 2019-06-05 writing/group_presentation.Rnw
@@
 
-\title{Turnover scale of soil fungal community composition}
-\subtitle{Presentation for Anna's Group Meeting}
-\author{Brendan Furneaux}
-\date{26 April, 2019}
+\title{Comparison of long and short amplicon metabarcoding for determination of fungal species diversity and spatial turnover in an ectomycorrhizal West African woodland}
+\author{\mbox{\textbf{Brendan Furneaux}\inst{1}} \and \mbox{Mohammad Bahram\inst{2}} \and \mbox{Nourou Yorou\inst{3}} \and \mbox{Anna Rosling\inst{1}} \and \mbox{Martin Ryberg\inst{1}}}
+\institute{\inst{1}Uppsala University \and \inst{2}Swedish University of Agricultural Sciences \and \inst{3}University of Parakou}
+\date{13 August, 2019}
 
@@
 @
+
+% # Metabarcoding
+% 
+%   - Cloning + Sanger -> 454 -> Illumina
+%   - For fungi, usually ITS2
+%       - short
+%       - intraspecific variation in most groups
+%       - good database support (Unite 8.0 - 817 130 fungal sequences)
+%       - ITS is "official barcode for fungi" \cite{schoch2012}
+%       - fungal specific primers
+%   - problems
+%       - not alignable
+%       - cannot place unknown sequences
+%       - variant 5.8S
+%       - variable length
+%   
+%   - LSU (26S/28S)
+%       - alternating highly conserved domains and variable regions
+%       - widely used for phylogenies
+%       - universal primers
+%       - less variable length
+%       - more conserved than ITS
+%       - not as good database support (RDP 11 fungal LSU - 125 525, Silva 132 LSU 907 382 of which )
+%   - SSU (18S)
+%       - alternating highly conserved and variable regions
+%       - more conserved than LSU
+%       - used for barcoding prokaryotes and "protists"
+%       - good database support (Silva SSU), but not as good for fungi
 

### 2019-08-12 writing/MSA2019_presentation.Rnw
add logo, comment out unneeded slides

@@
 \date{13 August, 2019}
+\logo{\includegraphics[width=2.5cm]{Uppsala_University_logo}}
 
@@
 
-\begin{frame}{Sequencing}
-  \framesubtitle{Which platform?}
-  \begincols
-  \begincol{0.5\linewidth}
-    \begin{block}{IonTorrent}
-      \begin{itemize}
-        \tightlist
-        \item Low cost per bp.
-        \item Smaller minimum job than Illumina.
-        \item Needs less DNA than PacBio.
-        \item Bias against amplicons greater than 400 bp.
-         \item Cannot sequence long ITS+LSU amplicon.
-       \end{itemize}
-     \end{block}
-  \endcol
-  \begincol{0.5\linewidth}
-     \begin{block}{PacBio}
-        \begin{itemize}
-          \tightlist
-          \item Higher cost per bp $\implies$ fewer samples or less sequencing depth for same budget.
-          \item Requires more DNA.
-          \item Good accuracy requires multiple circular reads.
-        \end{itemize}
-      \end{block}
-  \endcol
-  \endcols
-  \centering
-
-\end{frame}
-
-\begin{frame}{Sequencing}
-  \begin{block}{The plan}
-  Three sequencing runs:
-  \begin{itemize}
-    \item Ion Proton S5, short amplicon
-    \begin{itemize}
-      \item Pool all samples on one chip
-    \end{itemize}
-    \item PacBio RSII, short amplicon
-    \begin{itemize}
-      \item Same amplicon library as IonTorrent.
-      \item Two sequencing libraries, two SMRT cells per library.
-    \end{itemize}
-    \item PacBio RSII, long amplicon
-    \begin{itemize}
-      \item Two sequencing libraries, four SMRT cells per library.
-    \end{itemize}
-  \end{itemize}
-  \end{block}
-\end{frame}
-
-\begin{frame}{Sequencing}
-  \begin{block}{What happened}
-  Three sequencing runs:
-  \begin{itemize}
-    \item \sout{IonTorrent, short amplicon}
-    \begin{itemize}
-      \item Accidentally used primers with IonTorrent adapters built in.
-      \item Lost some of the multiplexing barcodes.
-      \item Can't demultiplex the samples --- can only look at pool of all reads.
-    \end{itemize}
-    \item PacBio, short amplicon
-    \begin{itemize}
-      \item OK
-    \end{itemize}
-    \item PacBio, long amplicon
-    \begin{itemize}
-      \item OK
-    \end{itemize}
-  \end{itemize}
-  \end{block}
-\end{frame}
+% \begin{frame}{Sequencing}
+%   \framesubtitle{Which platform?}
+%   \begincols
+%   \begincol{0.5\linewidth}
+%     \begin{block}{IonTorrent}
+%       \begin{itemize}
+%         \tightlist
+%         \item Low cost per bp.
+%         \item Smaller minimum job than Illumina.
+%         \item Needs less DNA than PacBio.
+%         \item Bias against amplicons greater than 400 bp.
+%          \item Cannot sequence long ITS+LSU amplicon.
+%        \end{itemize}
+%      \end{block}
+%   \endcol
+%   \begincol{0.5\linewidth}
+%      \begin{block}{PacBio}
+%         \begin{itemize}
+%           \tightlist
+%           \item Higher cost per bp $\implies$ fewer samples or less sequencing depth for same budget.
+%           \item Requires more DNA.
+%           \item Good accuracy requires multiple circular reads.
+%         \end{itemize}
+%       \end{block}
+%   \endcol
+%   \endcols
+%   \centering
+% 
+% \end{frame}
+% 
+% \begin{frame}{Sequencing}
+%   \begin{block}{The plan}
+%   Three sequencing runs:
+%   \begin{itemize}
+%     \item Ion Proton S5, short amplicon
+%     \begin{itemize}
+%       \item Pool all samples on one chip
+%     \end{itemize}
+%     \item PacBio RSII, short amplicon
+%     \begin{itemize}
+%       \item Same amplicon library as IonTorrent.
+%       \item Two sequencing libraries, two SMRT cells per library.
+%     \end{itemize}
+%     \item PacBio RSII, long amplicon
+%     \begin{itemize}
+%       \item Two sequencing libraries, four SMRT cells per library.
+%     \end{itemize}
+%   \end{itemize}
+%   \end{block}
+% \end{frame}
+% 
+% \begin{frame}{Sequencing}
+%   \begin{block}{What happened}
+%   Three sequencing runs:
+%   \begin{itemize}
+%     \item \sout{IonTorrent, short amplicon}
+%     \begin{itemize}
+%       \item Accidentally used primers with IonTorrent adapters built in.
+%       \item Lost some of the multiplexing barcodes.
+%       \item Can't demultiplex the samples --- can only look at pool of all reads.
+%     \end{itemize}
+%     \item PacBio, short amplicon
+%     \begin{itemize}
+%       \item OK
+%     \end{itemize}
+%     \item PacBio, long amplicon
+%     \begin{itemize}
+%       \item OK
+%     \end{itemize}
+%   \end{itemize}
+%   \end{block}
+% \end{frame}
 
@@
 
-<<errorrate, fig.width = 12>>=
-ggplot(reads_subset, aes(x = erate,
-                  linetype = Dataset,
-                  group = paste(Dataset, Region),
-                  color = Region)) +
-  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
-  scale_x_continuous(name = "Expected number of errors per base",
-                     trans = reverselog_trans(10),
-                     limits = c(NA, 1e-5),
-                     oob = squish) +
-  scale_y_continuous(name = "Fraction passing") +
-  scale_color_viridis_d(option = "C") +
-  scale_linetype_discrete()
-@
+% <<errorrate, fig.width = 12>>=
+% ggplot(reads_subset, aes(x = erate,
+%                   linetype = Dataset,
+%                   group = paste(Dataset, Region),
+%                   color = Region)) +
+%   geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
+%   scale_x_continuous(name = "Expected number of errors per base",
+%                      trans = reverselog_trans(10),
+%                      limits = c(NA, 1e-5),
+%                      oob = squish) +
+%   scale_y_continuous(name = "Fraction passing") +
+%   scale_color_viridis_d(option = "C") +
+%   scale_linetype_discrete()
+% @
 \end{frame}
@@
 
-<<totalerror, fig.width = 12>>=
-ggplot(reads_subset, aes(x = eexp,
-                  linetype = Dataset,
-                  group = paste(Dataset, Region),
-                  color = Region)) +
-  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
-  scale_x_continuous(name = "Expected number of errors",
-                     trans = reverselog_trans(10),
-                     limits = c(NA, 0.1),
-                     oob = squish) +
-  scale_y_continuous(name = "Fraction passing") +
-  scale_color_viridis_d(option = "C") +
-  scale_linetype_discrete()
-@
+% <<totalerror, fig.width = 12>>=
+% ggplot(reads_subset, aes(x = eexp,
+%                   linetype = Dataset,
+%                   group = paste(Dataset, Region),
+%                   color = Region)) +
+%   geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
+%   scale_x_continuous(name = "Expected number of errors",
+%                      trans = reverselog_trans(10),
+%                      limits = c(NA, 0.1),
+%                      oob = squish) +
+%   scale_y_continuous(name = "Fraction passing") +
+%   scale_color_viridis_d(option = "C") +
+%   scale_linetype_discrete()
+% @
 \end{frame}

### 2019-08-12 writing/MSA2019_presentation.Rnw
quality and length plots

@@
 
-\title{Comparison of long and short amplicon metabarcoding for determination of fungal species diversity and spatial turnover in an ectomycorrhizal West African woodland}
+\title{Comparison of long and short amplicon metabarcoding for determination of fungal species diversity and spatial turnover}
+\subtitle{in an ectomycorrhizal West African woodland}
 \author{\mbox{\textbf{Brendan Furneaux}\inst{1}} \and \mbox{Mohammad Bahram\inst{2}} \and \mbox{Nourou Yorou\inst{3}} \and \mbox{Anna Rosling\inst{1}} \and \mbox{Martin Ryberg\inst{1}}}
@@
 
+\logo{}
+
 <<setup, include=FALSE>>=
-knitr::opts_chunk$set(echo = FALSE, device = "tikz")
+knitr::opts_chunk$set(echo = FALSE, device = "tikz", cache = TRUE)
 library(drake)
@@
   \begin{itemize}
-    \item Redo RSII circular consensus basecalling using new algorithm (Arrow).
+    \item Redo RSII circular consensus basecalling using Sequel algorithm (Arrow).
     \item Demultiplex and trim primers.
     \item Cut out regions (ITS1, ITS2, full ITS, LSU) using ITSx.
-    \item Quality filter.
+    \item Quality filter - up to 3 expected errors.
   \end{itemize}
@@
 
-% <<errorrate, fig.width = 12>>=
-% ggplot(reads_subset, aes(x = erate,
-%                   linetype = Dataset,
-%                   group = paste(Dataset, Region),
-%                   color = Region)) +
-%   geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
-%   scale_x_continuous(name = "Expected number of errors per base",
-%                      trans = reverselog_trans(10),
-%                      limits = c(NA, 1e-5),
-%                      oob = squish) +
-%   scale_y_continuous(name = "Fraction passing") +
-%   scale_color_viridis_d(option = "C") +
-%   scale_linetype_discrete()
-% @
+<<errorrate, fig.width = 12, cache.check = reads_subset>>=
+filter(reads_subset, Tech == "PacBio") %>%
+  ggplot(aes(x = eexp,
+             group = paste(Dataset, Region),
+             color = Dataset)) +
+  geom_line(aes(y = 1 - ..y..), stat = "ecdf", size = 1) +
+  scale_x_continuous(name = "Expected number of errors",
+                     trans = reverselog_trans(10),
+                     limits = c(NA, 1e-5),
+                     oob = squish) +
+  facet_wrap(~Region, ncol = 2, strip.position = "right") +
+  scale_y_continuous(name = "Fraction passing") +
+  # scale_color_viridis_d(option = "C") +
+  geom_vline(aes(xintercept = 3), linetype = 2)
+@
 \end{frame}
@@
 
-% <<totalerror, fig.width = 12>>=
-% ggplot(reads_subset, aes(x = eexp,
-%                   linetype = Dataset,
-%                   group = paste(Dataset, Region),
-%                   color = Region)) +
-%   geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
-%   scale_x_continuous(name = "Expected number of errors",
-%                      trans = reverselog_trans(10),
-%                      limits = c(NA, 0.1),
-%                      oob = squish) +
-%   scale_y_continuous(name = "Fraction passing") +
-%   scale_color_viridis_d(option = "C") +
-%   scale_linetype_discrete()
-% @
+<<length, fig.width = 12, cache.check = reads_subset>>=
+filter(reads_subset, Tech == "PacBio", eexp <= 3) %>%
+  ggplot(aes(x = length,
+             group = paste(Dataset, Region),
+             color = Dataset, fill = Dataset)) +
+  geom_area(stat = "density", size = 1, alpha = 0.5) +
+  scale_x_log10(name = "Length (bp)",
+                limits = c(50, NA),
+                oob = squish) +
+  facet_wrap(~Region, ncol = 2, strip.position = "right", scales = "free_y") +
+  scale_y_continuous(name = "Fraction passing") +
+  # scale_color_viridis_d(option = "C") +
+  geom_vline(aes(xintercept = 3), linetype = 2)
+@
 \end{frame}
@@
   geom_point(size = 3) +
-  geom_line(alpha = 0.3) +
+  geom_line(alpha = 0.3, size = 1.5) +
   geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +

### 2019-08-13 writing/MSA2019_presentation.Rnw
Add rDNA and amplicon images to presentation

@@
 @
+
+\begin{frame}{Metabarcoding}
+\framesubtitle{rDNA regions}
+\includegraphics[width=\linewidth]{rDNA}
+
+\begin{tabular}{p{0.3\linewidht}p{0.3\linewidth}p{0.3\linewidth}}
+SSU & ITS & LSU \\ \midrule
+highly conserved w/variable regions & variable (5.8S conserved) & conserved w/variable regions \\
+used for prokaryotes and all eukaryotes & official barcode for fungi & mostly used for fungi, early phylogenetics \\
+\end{tabular}
+
+\end{frame}
 
@@
   \small
+  \includegraphics[width=\linewidth]{rDNA_amplicons}
+
   \begincols

### 2019-08-13 writing/MSA2019_presentation.Rnw
database comparisons

--- 2019-08-12 writing/MSA2019_presentation.Rnw
@@
 @
+
+\begin{frame}{Metabarcoding}
+\framesubtitle{rDNA regions}
+\includegraphics[width=\linewidth]{rDNA}
+
+\vfill
+
+\tiny
+\begin{tabular}{p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}}
+SSU & ITS & LSU \\ \midrule
+highly conserved w/variable regions & variable (5.8S conserved) & conserved w/variable regions \\[1en] 
+used for prokaryotes and all eukaryotes & official barcode for fungi & mostly used for fungi, early phylogenetics \\[1en]
+\end{tabular}
+
+\vfill
+
+\begincols
+  \begincol{0.3\linewidth}
+    \begin{block}{Silva 132}
+      2M seqs\\
+      700k "Ref NR 99" (clusters)\\
+      16k fungi, 60k other eukaryotes, 625k prokaryotes
+    \end{block}
+  \endcol
+  \begincol{0.3\linewidth}
+    \begin{block}{Unite 8.0}
+      800k seqs \\
+      120k "species hypotheses" (clusters)\\
+      70k fungi, 40k plants, 10k other euk \\
+      some sequences include LSU
+    \end{block}
+  \endcol
+  \begincol{0.3\linewidth}
+    \begin{block}{RDP 11 fungal LSU}
+      11k seqs\\
+      8k fungi, 3k other euk
+    \end{block}
+  \endcol
+\endcols
+
+\end{frame}
 

### 2019-08-13 writing/MSA2019_presentation.Rnw
add pictures

--- 2019-08-12 writing/MSA2019_presentation.Rnw
@@
 
-% # Metabarcoding
-% 
-%   - Cloning + Sanger -> 454 -> Illumina
-%   - For fungi, usually ITS2
-%       - short
-%       - intraspecific variation in most groups
-%       - good database support (Unite 8.0 - 817 130 fungal sequences)
-%       - ITS is "official barcode for fungi" \cite{schoch2012}
-%       - fungal specific primers
-%   - problems
-%       - not alignable
-%       - cannot place unknown sequences
-%       - variant 5.8S
-%       - variable length
-%   
-%   - LSU (26S/28S)
-%       - alternating highly conserved domains and variable regions
-%       - widely used for phylogenies
-%       - universal primers
-%       - less variable length
-%       - more conserved than ITS
-%       - not as good database support (RDP 11 fungal LSU - 125 525, Silva 132 LSU 907 382 of which )
-%   - SSU (18S)
-%       - alternating highly conserved and variable regions
-%       - more conserved than LSU
-%       - used for barcoding prokaryotes and "protists"
-%       - good database support (Silva SSU), but not as good for fungi
+\setbeamersize{text margin left=5mm,text margin right=5mm}
+
+\begin{frame}{rDNA Regions and Databases}
+\framesubtitle{rDNA regions}
+\vskip -1ex
+
+\includegraphics[width=\linewidth]{rDNA}
+
+\vfill
+
+\begin{overprint}
+\onslide<1>
+\small
+\begin{tabulary}{\linewidth}{r|L|L|L}
+& SSU (16S/18S) & ITS & LSU (23S/26S/28S) \\ \midrule
+Variation &
+  Highly conserved w/variable regions &
+  Variable (5.8S conserved) &
+  Conserved w/variable regions \\ \noalign{\smallskip}
+Uses & 
+  Prokaryota, broad Eukaryota &
+  Official barcode for Fungi &
+  Phylogenetics \\ \noalign{\smallskip}
+Alignable &
+  Yes &
+  No &
+  Yes \\ \noalign{\smallskip}
+Databases &
+  Silva, RDP (Prokaryotes only) &
+  Unite, Warcup &
+  Silva, RDP (Fungi) \\ 
+\end{tabulary}
+
+\onslide<2>
+\begin{columns}[t]
+  \begincol{0.3\linewidth}
+    \begin{block}{Silva 132 (2017)}
+      \small
+      \textbf{SSU Ref:}
+      \begin{itemize}
+        \footnotesize
+        \tightlist
+        \item 2M seqs
+        \item 700k "Ref NR 99" (clusters)
+        \item 16k Fungi, 60k other Eukaryota, 625k Prokaryota
+      \end{itemize}
+      \textbf{LSU Ref:}
+      \begin{itemize}
+        \footnotesize
+        \tightlist
+        \item 200k seqs
+        \item 5k Fungi, 24k other Eukaryota, 170k Prokaryota
+      \end{itemize}
+    \end{block}
+  \endcol
+  \begincol{0.3\linewidth}
+    \begin{block}{Unite 8.0 (2019)}
+    \small
+    \textbf{ITS:}
+      \begin{itemize}
+        \footnotesize
+        \tightlist
+        \item 800k seqs
+        \item 120k "species hypotheses" (clusters)
+        \item 70k Fungi, 40k plants, 10k other Eukaryota
+        \itemsome sequences include LSU
+      \end{itemize}
+    \end{block}
+  \endcol
+  \begincol{0.3\linewidth}
+    \small
+    \begin{block}{RDP 11 (2016)}
+      \small
+      \textbf{Fungal LSU training set:}\\
+      \begin{itemize}
+        \footnotesize
+        \tightlist
+        \item 11k seqs
+        \item 8k Fungi, 3k other Eukaryota
+        \item annotated to genus
+      \end{itemize}
+      \textbf{SSU dataset:}
+      \begin{itemize}
+        \footnotesize
+        \tightlist
+        \item Prokaryotes only
+      \end{itemize}
+      \textbf{Warcup Fungal ITS training set:}
+      \begin{itemize}
+        \footnotesize
+        \tightlist
+        \item 18k seqs
+        \item all Fungi
+      \end{itemize}
+    \end{block}
+  \endcol
+\end{columns}
+\center The three databases don't use the same taxonomy!
+\end{overprint}
+\end{frame}
 
@@
 
-\begin{frame}{Upper Ouémé Forest Reserve}
-\framesubtitle{Permanent study plots}
+\begin{frame}{Study site}
 \vskip1.5em
@@
         \includegraphics{soudanian_benin_oueme}
-      \onslide<3->
-        \includegraphics{soudanian_benin_oueme_village}
       \end{overprint}
@@
     \tightlist
-    \item<1-|alert@1> West Soudanian woodlands in central Bènin
-    \item<1-|alert@1> 10° N Latitude
-    \item<3-|alert@3> Two sites located near the villages Angaradebou and Gando
-    \item<4-|alert@4> Dominated by \emph{Isoberlinia doka} (Caesalpinoideae; ECM host)
+    \item West Soudanian woodlands in central Bènin
+    \item 10° N Latitude
+    \item Two sites 30km apart
+    \item Dominated by \emph{Isoberlinia doka} (Caesalpinoideae; ECM host)
+    \item Many AM trees and shrubs, plus three more ECM.
+    \item Fire \& termites
   \end{itemize}
@@
 \endcols
+\end{frame}
+
+\begin{frame}{Soudanian Woodlands}
+  \begin{overprint}
+    \onslide<1>
+      \includegraphics{Foret_Oueme_wide}
+    \onslide<2>
+      \includegraphics{woodland}
+  \end{overprint}
 \end{frame}
@@
           \tightlist
-          \item Full ITS plus the 5' end of LSU; 1500-2000 bp.
+          \item Full ITS plus the 5' end of LSU; 1000-2000 bp.
           \item ITS -- resolution to separate closely related taxa.
-          \item LSU -- can be aligned across fungi.
+          \item LSU -- can be aligned across Fungi.
           \item Primers are more universal than gITS7.

### 2019-08-21 writing/MSA2019_presentation.Rnw
final changes before MSA presentation

@@
 \logo{\includegraphics[width=2.5cm]{Uppsala_University_logo}}
+\setbeamersize{text margin left=5mm,text margin right=5mm}
 
@@
 theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
-loadd(bc_corr, bc_corr2, unif_corr, unif_corr2, qstats_join, datasets, longtree,
-      taxon_LSUcons_rdp)
+loadd(
+  bc_corr_long,
+  bc_corr_long2,
+  bc_corr_short,
+  bc_corr_short2,
+  # unif_corr,
+  # unif_corr2,
+  qstats_join,
+  datasets,
+  # longtree#,
+  # taxon_LSUcons_rdp
+  )
 reads <- qstats_join %>%
-  mutate_at("Region", factor,
+  mutate_at("region", factor,
             levels = c("ITS2", "ITS1", "short", "ITS", "LSU", "long")) %>%
-  mutate_at("Dataset", factor,
+  mutate_at("dataset", factor,
             levels = c("short-ion", "short-pacbio", "long-pacbio")) %>%
-  group_by(Dataset, Plate, Region)
+  group_by(dataset, plate, region)
 reads_subset <- reads %>%
@@
 
-\setbeamersize{text margin left=5mm,text margin right=5mm}
+\begin{frame}{Outline}
+\Large
+\begin{itemize}
+  \item Long amplicons for metabarcoding
+  \item Study site \& design
+  \item (early) Results
+\end{itemize}
+\end{frame}
 
-\begin{frame}{rDNA Regions and Databases}
-\framesubtitle{rDNA regions}
+\begin{frame}{rDNA Regions}
 \vskip -1ex
@@
 
-\vfill
+\vskip 2ex
 
@@
         \item 70k Fungi, 40k plants, 10k other Eukaryota
-        \itemsome sequences include LSU
+        \item some sequences include LSU
       \end{itemize}
@@
       \small
-      \textbf{Fungal LSU training set:}\\
+      \textbf{LSU training set:}\\
       \begin{itemize}
@@
         \tightlist
-        \item 18k seqs
-        \item all Fungi
+        \item 18k seqs, all Fungi
       \end{itemize}
@@
 
-\begin{frame}{Introduction}
+\begin{frame}{Why long amplicons?}
 
 \begin{itemize}
- \item Question: What is the scale of spatial (and temporal) autocorrelation in soil fungal community composition?
+ \item Combine ITS (as barcode) with LSU/SSU (phylogenetic inference)
+ \item Long read technologies are more tolerant of length variation
  \begin{itemize}
-    \item Specifically in a tropical ECM forest.
+    \item PCR length bias still an issue!
  \end{itemize}
- \item Metabarcoding method comparison
+ \item Very conserved primer sites reduce primer bias.
+ \item (Tedersoo et al., 2018)
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Why not long amplicons?}
+\begin{itemize}
+  \item Read quality
+    \begin{itemize}
+      \item PacBio circular consensus gives good quality up to a few kb
+      \item Oxford Nanopore is still problematic
+    \end{itemize}
+  \item Chimera formation in PCR.
+    \begin{itemize}
+      \item Alternation of conserved and variable regions is a recipe for chimera formation.
+    \end{itemize}
+  \item Cost -- for the same budget, Illumina gives many more samples and/or greater depth.
+  \item (Kennedy et al. 2018)
+
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Does it matter?}
+\begin{itemize}
+  \item Do a simple ecological study using different methods, see if the results change.
+  \item Measuring spatial turnover in soil fungal community.
 \end{itemize}
@@
 \begin{frame}{Study site}
-\vskip1.5em
+\vskip 1.5em
 \begincols
@@
   \begin{itemize}
-    \item<1-|alert@1> \SI{25}{\m} transects.
-    \item<1-|alert@1> Soil samples taken every 1 m.
-    \item<2-|alert@2> $\SI{5}{\cm}\times\SI{5}{\cm}\times\SI{5}{\cm}$ sample, homogenize in plastic bag, then subsample \SI{50}{\mg}
-    \item<3-|alert@3> Lysis \& preservation in field
-    \item<4-|alert@4> Same locations in two years -- 2015 and 2016 (25 points × 2 years × 2 sites = 100 samples)
+    \item \SI{25}{\m} transects.
+    \item Soil samples taken every 1 m.
+    \item $\SI{5}{\cm}\times\SI{5}{\cm}\times\SI{5}{\cm}$ sample, homogenize in plastic bag, then subsample \SI{50}{\mg}
+    \item Lysis \& preservation in field
+    \item Same locations in two years -- 2015 and 2016 (25 points × 2 years × 2 sites = 100 samples) 
   \end{itemize}
@@
 \begin{frame}{Sequencing}
-  \framesubtitle{Which region?}
-  \small
+  \framesubtitle{Which amplicon?}
   \includegraphics[width=\linewidth]{rDNA_amplicons}
-
-  \begincols
+  \vfill
+  \begin{columns}[t]
   \begincol{0.5\linewidth}
-    \begin{block}{ITS2 (gITS7-ITS4)}
+    \begin{block}{short amplicon}
       \begin{itemize}
         \tightlist
-          \item Internally transcribed spacer between 5.8S and 26/28S (LSU) subunits of ribosomal RNA
-          \item Usually 250--350 bp.
-          \item ITS --- standard DNA barcode for fungi.
-          \item Relatively good species-level resolution, and good database support.
-          \item Cannot be aligned across larger taxonomic groups.
-          \item Groups with variant (or absent) 5.8S not sequenced.
+        \item ITS2
+        \item primers gITS7 -- ITS4
+        \item 250--350 bp
+        \item Sequence same amplicon library with Illumina MiSeq and PacBio RSII
+        \item<2-> Don't have MiSeq data yet!
        \end{itemize}
@@
   \begincol{0.5\linewidth}
-      \begin{block}{Full ITS + partial LSU (ITS1-LR5)}
+     \begin{block}{long amplicon}
         \begin{itemize}
           \tightlist
-          \item Full ITS plus the 5' end of LSU; 1000-2000 bp.
-          \item ITS -- resolution to separate closely related taxa.
-          \item LSU -- can be aligned across Fungi.
-          \item Primers are more universal than gITS7.
-          \item Long amplicon with alternating variable and conserved regions --- susceptible to chimera formation and breakage.
+          \item Full ITS + partial LSU
+          \item Primers ITS1 -- LR5
+          \item 1200 -- 1800 bp
+          \item PacBio RSII
         \end{itemize}
@@
   \endcol
-  \endcols
+\end{columns}
 \end{frame}
-
-% \begin{frame}{Sequencing}
-%   \framesubtitle{Which platform?}
-%   \begincols
-%   \begincol{0.5\linewidth}
-%     \begin{block}{IonTorrent}
-%       \begin{itemize}
-%         \tightlist
-%         \item Low cost per bp.
-%         \item Smaller minimum job than Illumina.
-%         \item Needs less DNA than PacBio.
-%         \item Bias against amplicons greater than 400 bp.
-%          \item Cannot sequence long ITS+LSU amplicon.
-%        \end{itemize}
-%      \end{block}
-%   \endcol
-%   \begincol{0.5\linewidth}
-%      \begin{block}{PacBio}
-%         \begin{itemize}
-%           \tightlist
-%           \item Higher cost per bp $\implies$ fewer samples or less sequencing depth for same budget.
-%           \item Requires more DNA.
-%           \item Good accuracy requires multiple circular reads.
-%         \end{itemize}
-%       \end{block}
-%   \endcol
-%   \endcols
-%   \centering
-% 
-% \end{frame}
 % 
@@
     \item Redo RSII circular consensus basecalling using Sequel algorithm (Arrow).
-    \item Demultiplex and trim primers.
     \item Cut out regions (ITS1, ITS2, full ITS, LSU) using ITSx.
@@
 \begin{frame}{Expected error rate}
-\framesubtitle{After \texttt{Arrow}}
+\framesubtitle{After \texttt{Arrow} and \texttt{ITSx}}
 
 <<errorrate, fig.width = 12, cache.check = reads_subset>>=
-filter(reads_subset, Tech == "PacBio") %>%
+filter(reads_subset, tech == "PacBio") %>%
   ggplot(aes(x = eexp,
-             group = paste(Dataset, Region),
-             color = Dataset)) +
+             group = paste(dataset, region),
+             color = dataset)) +
   geom_line(aes(y = 1 - ..y..), stat = "ecdf", size = 1) +
@@
                      oob = squish) +
-  facet_wrap(~Region, ncol = 2, strip.position = "right") +
+  facet_wrap(~region, ncol = 2, strip.position = "right") +
   scale_y_continuous(name = "Fraction passing") +
@@
   \begin{itemize}
-    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
-      \begin{itemize}
-        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
-        \item Works on the smaller regions.
-      \end{itemize}
+    \item Find ITS2 amplicon sequence variants (ASVs) with \texttt{dada2}.
+    \item Generate LSU consensus for each ASV
   \end{itemize}
@@
 
-\begin{frame}{Expected total number of errors}
-\framesubtitle{After \texttt{Arrow}}
+\begin{frame}{Amplicon and individual region lengths}
+\framesubtitle{After \texttt{Arrow} and \texttt{ITSx}}
 
 <<length, fig.width = 12, cache.check = reads_subset>>=
-filter(reads_subset, Tech == "PacBio", eexp <= 3) %>%
+filter(reads_subset, tech == "PacBio", eexp <= 3) %>%
   ggplot(aes(x = length,
-             group = paste(Dataset, Region),
-             color = Dataset, fill = Dataset)) +
+             group = paste(dataset, region),
+             color = dataset, fill = dataset)) +
   geom_area(stat = "density", size = 1, alpha = 0.5) +
@@
                 oob = squish) +
-  facet_wrap(~Region, ncol = 2, strip.position = "right", scales = "free_y") +
+  facet_wrap(~region, ncol = 2, strip.position = "right", scales = "free_y") +
   scale_y_continuous(name = "Fraction passing") +
@@
 
-\begin{frame}{Bioinformatics}
-\framesubtitle{continued...}
+\begin{frame}{Long vs. short amplicons}
+\framesubtitle{ASV read counts}
+\begin{overprint}
+\onslide<1>
+Total ASVs:
 \begin{itemize}
-    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
-      \begin{itemize}
-        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
-        \item Works on the smaller regions.
+  \item Long amplicon: 646 ASVs
+  \item Short amplicon: 531 ASVs
 \end{itemize}
-    \item For each ITS2 ASV, align the whole long read and find the consensus.
-      \begin{itemize}
-        \item There were a lot of chimeras (15--25\%).
-        \item Remove "outliers" from the alignments using \texttt{odseq}.
-        \item Remaining chimeras are too few to change the consensus.
-      \end{itemize}
-    \item Align with \texttt{DECIPHER}
-      \begin{itemize}
-        \item Progressive aligner with inferred RNA secondary structure.
-      \end{itemize}
-    \item \texttt{RAxML} tree
-  \end{itemize}
+
+\onslide<2>
+<<readplot, fig.width = 12, message=FALSE>>=
+loadd(big_seq_table_ITS2)
+asv_table <- big_seq_table_ITS2 %>%
+  tibble::as_tibble(rownames = "filename") %>%
+  tidyr::gather(key = "seq", value = "reads", -1) %>%
+  dplyr::filter(reads >= 1) %>%
+  tidyr::extract(col = "filename",
+                 into = c("seq_run", "plate", "well", "dir", "region"),
+                 regex = "([a-z]+_\\d+)_(\\d+)-([A-H]1?[0-9])([fr]?)-([:alnum:]+).+") %>%
+  dplyr::filter(seq_run %in% c("pb_483", "pb_500")) %>%
+  dplyr::group_by(seq_run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq_run, value = reads, fill = 0) %>%
+  dplyr::rename(short_pacbio = pb_483, long_pacbio = pb_500)
+
+asv_table %>% 
+  filter(short_pacbio > 0, long_pacbio > 0) %>%
+  ggplot(aes(x = short_pacbio, y = long_pacbio)) +
+  geom_point(alpha = 0.2) +
+  scale_x_log10() +
+  scale_y_log10() +
+  xlab("Short amplicon reads") +
+  ylab("Long amplicon reads") +
+  geom_rug(aes(y = long_pacbio, x = 1), sides = "l", alpha = 0.2,
+           data = asv_table %>% filter(short_pacbio == 0)) +
+  geom_rug(aes(y = short_pacbio), sides = "b", alpha = 0.2,
+           data = asv_table %>% filter(long_pacbio == 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(long_pacbio[long_pacbio >= 1])) -
+                    mean(log10(short_pacbio[short_pacbio >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5)
+@
+\end{overprint}
 \end{frame}
 
-\begin{frame}{Taxonomy assignment}
-Both \texttt{dada2} and \texttt{DECIPHER} can do taxonomy assignment.
-\begincols
-\begincol{0.5\framewidth}
-\begin{block}{dada2}
-  \begin{itemize}
-    \item Naïve Bayesian classifier.
-    \item Uses a lot of memory.
-    \item Needs every rank to be filled (\emph{incertae sedis}, "unidentified").
-    \item Seems to put thins in places that don't match the tree.
-  \end{itemize}
-\end{block}
-\endcol
-
-\begincol{0.5\framewidth}
-\begin{block}{DECIPHER}
-  \begin{itemize}
-    \item Some kind of machine learning algorithm.
-    \item Uses less memory.
-    \item Allows uneven classification depths.
-    \item Very low bootstraps, especially for ITS.
-  \end{itemize}
-\end{block}
-\endcol
-\endcols
-
-UNITE for ITS, RDP for LSU...  But there is no large database that has both for the same specimens!
-
-I'm still working on it...
-\end{frame}
-
-\begin{frame}{Tree}
-\framesubtitle{Kingdoms}
-<<kingtree, fig.width = 12, message=FALSE>>=
-seqhash <- function(seq, len = 8) UseMethod("seqhash")
-seqhash.character <- function(seq, len = 8) {
-  h <- purrr::map_chr(seq, digest::digest)
-  stringr::str_sub(h, end = len)
-}
-seqhash.XStringSet <- function(seq, len = 8) {
-  seqhash.character(as.character(seq))
-}
-treetax <- taxon_LSUcons_rdp %>%
-  mutate_at("seq", seqhash) %>%
-  select(-Taxonomy, -nreads) 
-outgroups <- treetax %>% filter(!is.na(Kingdom), Kingdom %in% c("Alveolata", "Chromista", "Plantae"), seq %in% longtree$tip.label)
-p <- ggtree(tr = longtree)
-longtree <- reroot(longtree, MRCA(p, outgroups$seq))
-
-ggtree(groupOTU(longtree, split(treetax$seq, treetax$Kingdom))) +
-  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
-  theme(legend.position = NULL)
-@
-\end{frame}
-
-\begin{frame}{Tree}
-\framesubtitle{Phyla}
-<<phytree, fig.width = 12, message=FALSE>>=
-ggtree(groupOTU(longtree, split(treetax$seq, treetax$Phylum))) +
-  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
-  theme(legend.position = NULL)
-@
-\end{frame}
-
-\begin{frame}{Tree}
-\framesubtitle{Families --- Basidiomycota}
-<<basidtree, fig.width = 12, message=FALSE>>=
-basids <- treetax %>% filter(!is.na(Class),
-                             Class %in% c("Geminibasidiomycetes",
-                                         "Tremellomycetes",
-                                         "Exobasidiomycetes",
-                                         "Microbotryomycetes"),
-                             seq %in% longtree$tip.label)
-
-basidtree <- phylobase::phylo4(longtree) %>%
-  phylobase::subset(mrca = basids$seq) %>%
-  as("phylo")
-
-basids2 <- treetax %>% filter(seq %in% basidtree$tip.label)
-
-ggtree(groupOTU(basidtree, split(basids2$seq, basids2$Order)), layout = "circular") +
-  aes(color = group) +
-  scale_color_brewer(type = "qual", palette = "Paired", limits = unique(na.omit(basids2$Order))) +
-  theme(legend.position = NULL)
-@
-\end{frame}
+% \begin{frame}{Tree}
+% \framesubtitle{Kingdoms}
+% <<kingtree, fig.width = 12, message=FALSE>>=
+% seqhash <- function(seq, len = 8) UseMethod("seqhash")
+% seqhash.character <- function(seq, len = 8) {
+%   h <- purrr::map_chr(seq, digest::digest)
+%   stringr::str_sub(h, end = len)
+% }
+% seqhash.XStringSet <- function(seq, len = 8) {
+%   seqhash.character(as.character(seq))
+% }
+% treetax <- taxon_LSUcons_rdp %>%
+%   mutate_at("seq", seqhash) %>%
+%   select(-Taxonomy, -nreads) 
+% outgroups <- treetax %>% filter(!is.na(Kingdom), Kingdom %in% c("Alveolata", "Chromista", "Plantae"), seq %in% longtree$tip.label)
+% p <- ggtree(tr = longtree)
+% longtree <- reroot(longtree, MRCA(p, outgroups$seq))
+% 
+% ggtree(groupOTU(longtree, split(treetax$seq, treetax$Kingdom))) +
+%   aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
+%   theme(legend.position = NULL)
+% @
+% \end{frame}
+% 
+% \begin{frame}{Tree}
+% \framesubtitle{Phyla}
+% <<phytree, fig.width = 12, message=FALSE>>=
+% ggtree(groupOTU(longtree, split(treetax$seq, treetax$Phylum))) +
+%   aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
+%   theme(legend.position = NULL)
+% @
+% \end{frame}
+% 
+% \begin{frame}{Tree}
+% \framesubtitle{Families --- Basidiomycota}
+% <<basidtree, fig.width = 12, message=FALSE>>=
+% basids <- treetax %>% filter(!is.na(Class),
+%                              Class %in% c("Geminibasidiomycetes",
+%                                          "Tremellomycetes",
+%                                          "Exobasidiomycetes",
+%                                          "Microbotryomycetes"),
+%                              seq %in% longtree$tip.label)
+% 
+% basidtree <- phylobase::phylo4(longtree) %>%
+%   phylobase::subset(mrca = basids$seq) %>%
+%   as("phylo")
+% 
+% basids2 <- treetax %>% filter(seq %in% basidtree$tip.label)
+% 
+% ggtree(groupOTU(basidtree, split(basids2$seq, basids2$Order)), layout = "circular") +
+%   aes(color = group) +
+%   scale_color_brewer(type = "qual", palette = "Paired", limits = unique(na.omit(basids2$Order))) +
+%   theme(legend.position = NULL)
+% @
+% \end{frame}
 
 \begin{frame}{Correlogram}
+
 <<correlogram, fig.width = 12>>=
-bind_rows(as.data.frame(bc_corr$mantel.res) %>%
-            mutate(tdist = "same", metric = "bray"),
-          as.data.frame(bc_corr2$mantel.res) %>%
-            mutate(tdist = "diff", metric = "bray"),
-          as.data.frame(unif_corr$mantel.res) %>%
-            mutate(tdist = "same", metric = "unifrac"),
-          as.data.frame(unif_corr2$mantel.res) %>%
-            mutate(tdist = "diff", metric = "unifrac")) %>%
+bind_rows(as.data.frame(bc_corr_long$mantel.res) %>%
+            mutate(tdist = "same", seq_run = "long"),
+          as.data.frame(bc_corr_long2$mantel.res) %>%
+            mutate(tdist = "diff", seq_run = "long"),
+          as.data.frame(bc_corr_short$mantel.res) %>%
+            mutate(tdist = "same", seq_run = "short"),
+          as.data.frame(bc_corr_short2$mantel.res) %>%
+            mutate(tdist = "diff", seq_run = "short")) %>%
+  filter(complete.cases(.)) %>%
   mutate(Significant = `Pr(corrected)` < 0.05) %>%
@@
   xlab("distance (m)") +
-  facet_wrap("metric") +
+  facet_wrap("seq_run") +
   scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
@@
 
+\begin{frame}{Acknowledgements}
+\begin{columns}[t]
+  \begin{column}{0.5\linewidth}
+    \textbf{Uppsala University}\\
+    Anneli Svanholm\\
+    Bobby Sulistyo\\
+    Ylva Strid\\
+    Kerri Kluting
+  \end{column}
+  \begin{column}{0.5\linewidth}
+
+    \textbf{University of Parakou}\\
+    Roel Houdanon\\
+    Evans Codjia\\
+    Souleymane Boni\\
+    Hyppolite Aignon\\
+    Sylvestre Badou\\
+    Gerard Laourou\\
+    Aziz Boukary\\
+    Olivia Fayedi
+  \end{column}
+\end{columns}
+\includegraphics[width=0.5\linewidth]{formas}
+\end{frame}
+
 \end{document}

### 2019-08-21 scripts/drake-06-pretaxonomy.R

@@ -16,7 +16,8 @@ setup_log("pretaxonomy")
 
 #### pre-taxonomy ####
 # single-threaded targets after dada2, before taxonomy.
-if (target %in% od || !all(file.exists(outputs))) {
+if (target %in% od #|| !all(file.exists(outputs))
+    ) {
   cat("\n Making pre-taxonomy targets (loop)...\n")
   tictoc::tic()
   dconfig <- drake::drake_config(plan,

### 2019-08-21 scripts/drake.R

@@ -19,6 +19,7 @@ if (interactive()) {
   in_files <- list.files(path = trim_dir,
                          pattern = "\\.trim\\.fastq\\.gz$",
                          full.names = TRUE, recursive = TRUE)
+  in_files <- in_files[!grepl("is_057", in_files)]
   dataset_file <- file.path(lab_dir, "datasets.csv")
   regions_file <- file.path(lab_dir, "regions.csv")
   platemap_file <- file.path(lab_dir, "Brendan_soil2.xlsx")
@@ -551,7 +552,7 @@ plan <- drake_plan(
   # Replace raw reads which were mapped to a DADA2 ASV with the ASV,
   # and put them all in one tibble.  We'll only do this for pacbio long reads.
   combined = target(
-    tzara::combine_bigmaps(purrr::imap_dfr(c(dada_map), ~mutate(.x, name = .y)),
+    combine_bigmaps(dplyr::bind_rows(dada_map),
                            dplyr::bind_rows(raw)),
     transform = combine(dada_map, raw, .by = seq_run_ID, .tag_in = step)
   ),
@@ -561,6 +562,8 @@ plan <- drake_plan(
   # with at least 3 sequences.
   conseq = 
     combined_pb_500 %>%
+    tidyr::extract(name, c("seq_run", "plate", "well", "direction", "region"), "([pi][sb]_\\d{3})_(\\d{3})-([A-H]\\d{1,2})([fr])-(.+)\\.qfilt\\.fastq\\.gz") %>%
+    tidyr::spread(key = "region", value = "seq") %>%
     dplyr::group_by(ITS2) %>%
     dplyr::filter(!is.na(ITS2), dplyr::n() >= 3) %>%
     dplyr::summarize(nreads = dplyr::n(),
@@ -644,34 +647,51 @@ plan <- drake_plan(
       seqhash(ITS2)
     lt},
   
-  physeq = assemble_physeq(platemap, datasets, relabel_seqtable(big_seq_table_ITS2), longtree),
+  physeq = assemble_physeq(platemap, datasets, relabel_seqtable(big_seq_table_ITS2)),#, longtree),
   
   longphyseq = physeq %>%
-    phyloseq::prune_samples(samples = phyloseq::sample_data(.)$Dataset == "long-pacbio"
+    phyloseq::prune_samples(samples = phyloseq::sample_data(.)$dataset == "long-pacbio"
                             & phyloseq::sample_data(.)$sample_type == "Sample"
[…]

### 2019-08-21 scripts/mantel.R

@@ -4,21 +4,21 @@
 # convert colnames from the full sequence to an 8-character hash.
 # also pools forward and reverse reads from same well.
 
-relabel_seqtable <- function(seqtable, hashlen = 8) {
+relabel_seqtable <- function(seqtable) {
   seqtable %>%
     # set the new column names
-    magrittr::set_colnames(seqhash(colnames(.), len = hashlen)) %>%
+    magrittr::set_colnames(tzara::seqhash(colnames(.))) %>%
     # convert to a tibble for easier name column operations
     tibble::as_tibble(rownames = "file") %>%
     # parse the filename
     tidyr::extract(
       col = "file",
-      into = c("Seq.Run", "Plate", "Well", "Direction", "Region"),
+      into = c("seq_run", "plate", "well", "direction", "region"),
       regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)([fr]?)-([:alnum:]+)\\.qfilt\\.fastq\\.gz") %>%
     # create the ID
-    tidyr::unite("ID", Seq.Run, Plate, Well, sep = "") %>%
+    tidyr::unite("ID", seq_run, plate, well, sep = "") %>%
     # remove unnecessary columns
-    dplyr::select(-Direction, -Region) %>%
+    dplyr::select(-direction, -region) %>%
     # pool all reads from the same well
     dplyr::group_by(ID) %>%
     dplyr::summarise_all(sum) %>%
@@ -30,23 +30,23 @@ relabel_seqtable <- function(seqtable, hashlen = 8) {
 
 assemble_physeq <- function(platemap, datasets, seqtable, tree) {
   samp <- platemap %>%
-    dplyr::mutate_at("Primer.Pair", tolower) %>%
+    dplyr::mutate_at("primer_pair", tolower) %>%
     dplyr::left_join(
       datasets %>%
-        dplyr::select(Dataset, Seq.Run, Tech, Forward, Reverse) %>%
+        dplyr::select(dataset, seq_run, tech, forward, reverse) %>%
         dplyr::mutate(
-          Primer.Pair = paste(stringr::str_replace(Forward, "_tag.*$", ""),
-                              stringr::str_replace(Reverse, "_tag.*$", ""),
+          primer_pair = paste(stringr::str_replace(forward, "_tag.*$", ""),
+                              stringr::str_replace(reverse, "_tag.*$", ""),
[…]

### 2019-08-21 scripts/plate_check.R

@@ -14,62 +14,64 @@ read_platemap <- function(filename, sheet, skip = 2) {
                             col_names = TRUE, .name_repair = "minimal")
   # the data is not "tidy"; we have groups of columns for the different
   # primer pairs.  Find the cells which define the primer pairs.
-  primer.pair <- cells %>%
+  primer_pair <- cells %>%
     dplyr::filter(row < header_line,
            stringr::str_detect(character, "ITS")) %>%
     # parse out the primer names
     dplyr::select(col, character) %>%
-    tidyr::extract(character, c("Forward", "Reverse"),
+    tidyr::extract(character, c("forward", "reverse"),
                    regex = "^([:alnum:]+) ?and ?([:alnum:]+).*") %>%
-    tidyr::unite(Primer.Pair, Forward, Reverse, remove = FALSE) %>%
+    tidyr::unite(primer_pair, forward, reverse, remove = FALSE) %>%
     # find the columns they refer to
-    dplyr::mutate(col.end = dplyr::lead(col) - 1,
-                  col.end = tidyr::replace_na(col.end, ncol(raw))) %>%
+    dplyr::mutate(col_end = dplyr::lead(col) - 1,
+                  col_end = tidyr::replace_na(col_end, ncol(raw))) %>%
     dplyr::rowwise()
   # Get the data which applies to all primer pairs.
-  out_all <- raw[1:(min(primer.pair$col) - 1)]
+  out_all <- raw[1:(min(primer_pair$col) - 1)]
   # tidy the whole thing
-  dplyr::do(primer.pair,
+  dplyr::do(primer_pair,
      dplyr::bind_cols(out_all,
-               Primer.Pair = rep(.$Primer.Pair, nrow(raw)),
-               raw[(.$col):(.$col.end)])) %>%
+               primer_pair = rep(.$primer_pair, nrow(raw)),
+               raw[(.$col):(.$col_end)])) %>%
     # get rid of empty columns
     purrr::discard(~all(is.na(.))) %>%
     # make nice names
-    dplyr::rename(Comment = V1,
-           Well = row,
-           DNA.Conc = "Conc. (ng/µl)",
-           DNA.Tot = DNA,
-           PCR.Conc = "ng/µl",
-           PCR.Tot = "Total ng (39 µl)",
-           Pool.Vol = "µl Product") %>%
+    dplyr::rename(comment = V1,
+           well = row,
+           dna_conc = "Conc. (ng/µl)",
+           dna_tot = DNA,
+           pcr_conc = "ng/µl",
[…]

### 2019-10-09 writing/LSUx.Rmd
LSUx and reconstruct.  Intended to run up to consensus.

---
title: Comparison of DADA2 performance on LSUx domains (including ITS and 5.8S)
  rDNA
author: "Brendan Furneaux"
date: "October 2, 2019"
output:
  beamer_presentation:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
library(ggplot2)
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
library(drake)
library(magrittr)
targets <- drake_cache()$list()
expand.grid(plate = c("001", "002"),
            region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2",
                       "LSU3", "D3", "LSU4", "ITS", "LSU", "32S", "long"))

table <- tibble::tibble(name = targets[startsWith(targets, "nochim_pb_500")]) %>%
  tidyr::extract(col = name, into = c("plate", "region"),
                 regex = "nochim_pb_500_(00[12])_(.+)",
                 remove = FALSE) %>%
  dplyr::mutate_at("region", factor, levels = c("ITS1", "5_8S", "ITS2", "LSU1",
                                                "D1", "LSU2", "D2", "LSU3",
                                                "D3", "LSU4", "ITS", "LSU",
                                                "32S", "long")) %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(nochim = lapply(name, readd, character_only = TRUE)) %>%
  dplyr::summarize(nochim = list(dada2::mergeSequenceTables(tables = nochim)),
                   reads = sum(nochim),
                   ASVs = ncol(nochim),
                   length_min = min(nchar(colnames(nochim))),
                   length_q1 = quantile(rep(nchar(colnames(nochim)),
                                            colSums(nochim)), 0.25),
                   length_med = median(rep(nchar(colnames(nochim)),
                                           colSums(nochim))),
                   length_mean = mean(rep(nchar(colnames(nochim)),
                                          colSums(nochim))),
                   length_q3 = quantile(rep(nchar(colnames(nochim)),
                                            colSums(nochim)), 0.75),
                   length_max = max(nchar(colnames(nochim))))
```

## Reads vs. ASVs

```{r}
ggplot(table, aes(reads, ASVs, color = length_med, label = region)) +
  geom_text() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  xlab("mapped reads")
```

## Length distributions
```{r}
ggplot(table, aes(x = length_med, y = reads, label = region)) +
  geom_text() +
  xlab("length (bp)")
```

## Length distributions - Variable regions
```{r}
lengths_table <- table %>%
  mutate(
    nreads = map(nochim, colSums),
    length = map(nochim, ~ nchar(colnames(.)))
  ) %>%
  select(region, length, nreads) %>%
  unnest() %>%
  group_by(region, length) %>%
  summarize(nreads = sum(nreads))

lengths_table %>%
  filter(region %in% c("ITS1", "ITS2", "D1", "D2", "D3")) %>%
  dplyr::slice(rep(1:nrow(.), .$nreads)) %>%
  ggplot(aes(length)) +
  geom_density(bw = 2) +
  facet_wrap(~region,ncol = 2, strip.position = "right") +
  scale_y_continuous(breaks = NULL, labels = NULL)
```

## Length distributions -- conserved regions

```{r}
lengths_table %>%
  filter(region %in% c("5_8S", "LSU1", "LSU2", "LSU3", "LSU4")) %>%
  dplyr::slice(rep(1:nrow(.), .$nreads)) %>%
  ggplot(aes(length)) +
  geom_density(bw = 1) +
  facet_wrap(~region,ncol = 2, strip.position = "right") +
  scale_y_continuous(breaks = NULL, labels = NULL) +
  scale_x_continuous(limits = c(0, 300))
```

## Length distributions -- combined regions

```{r}
lengths_table %>%
  filter(region %in% c("ITS", "LSU", "32S", "long")) %>%
  dplyr::slice(rep(1:nrow(.), .$nreads)) %>%
  ggplot(aes(length)) +
  geom_density(bw = 10) +
  facet_wrap(~region,ncol = 2, strip.position = "right") +
  scale_y_continuous(breaks = NULL, labels = NULL) +
  scale_x_continuous(limits = c(0, 2000))
```

### 2019-10-09 writing/filter-check.Rmd
LSUx and reconstruct.  Intended to run up to consensus.

---
title: "Filter check"
output: beamer_presentation
---

```{r setup}
# source("install_packages.R")

library(magrittr)
library(tidyverse)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")

#if (interactive()) {
  base.dir <- getwd() %>%
    str_extract(".*oueme-fungi-transect")
  data.dir <- file.path(base.dir, "data")
  lab.dir <- file.path(base.dir, "config")
  seq.dir <- file.path(base.dir, "sequences")
  rawfastq.dir <- file.path(seq.dir, "rawfastq")
  demux.dir <- file.path(seq.dir, "demux")
  trim.dir <- file.path(seq.dir, "trim")
  dataset.file <- file.path(lab.dir, "datasets.csv")
  datasets <- read_csv(dataset.file)

#}
```

```{r}

reads <- datasets %>%
  mutate(
    trim.files = map(
      paste0(seq_run, ".*\\.fastq\\.gz"),
      list.files,
      path = trim.dir,
      full.names = TRUE
    ),
    old.demux.files = map(
      paste0(dataset, ".*\\.fastq\\.gz"),
      list.files,
      path = demux.dir,
      full.names = TRUE
    ),
    files = pmap(
      list(trim.files, old.demux.files),
      c
    )
  ) %>%
  select(-trim.files, -old.demux.files) %>%
  unnest() %>%
  filter(!str_detect(files, "-x[a-z]{2}\\.fastq\\.gz")) %>%
  mutate(filesize = file.size(files),
         base_calling = ifelse(str_detect(basename(files), seq_run),
                               "arrow", "default")) %>%
  filter(filesize > 0,
         base_calling == "default" | tech == "PacBio") %>%
  mutate(fastq = map(files, readFastq),
         # fastq = map(fastq, ~magrittr::extract(., width(.) > 0)),
         eexp = map(fastq, ~ 10^(-1 * as(.@quality, "matrix")/10) %>%
                      rowSums(na.rm = TRUE)),
         p.noerr = map(fastq, ~log1p(- 10^(-1 * as(.@quality, "matrix")/10)) %>%
                         rowSums(na.rm = TRUE) %>%
                         exp),
         length = map(fastq, width)) %>%
  select(-fastq) %>%
  unnest() %>%
  mutate(erate = eexp/length)
```

```{r}
reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, 
              log_breaks(base = base), 
              domain = c(1e-100, Inf))
}

ggplot(reads, aes(x = erate,
                  # group = paste(base_calling, dataset),
                  color = dataset,
                  linetype = base_calling)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors per base",
                     trans = reverselog_trans(10),
                     limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")

ggplot(reads, aes(x = eexp,
                  group = paste(base_calling, dataset),
                  color = dataset,
                  linetype = base_calling)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors",
                     trans = reverselog_trans(10),
                     limits = c(NA, 0.1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")

ggplot(reads, aes(x = p.noerr,
                  group = paste(base_calling, dataset),
                  color = dataset,
                  linetype = base_calling)) +
  geom_step(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expectation of being error-free",
                     # trans = log_trans(10),
                     limits = c(0, 1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing")

ggplot(reads, aes(x = length,
                  group = paste(base_calling, dataset),
                  color = dataset,
                  linetype = base_calling)) +
  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
  scale_x_log10(name = "Length (bp)") +
  scale_y_continuous(name = "Fraction Greater than length")

reads %>%
  group_by(base_calling, dataset) %>%
  summarize(`Expected Error-Free` = sum(p.noerr),
            Total = n()) %>%
  knitr::kable()
```

```{r}
library(dada2)
dada2::plotQualityProfile(unique(filter(reads, Tech == "PacBio")$files))
```

```{r}
counts <- read_csv(file.path(data.dir, "fastq.counts"), col_names = c("file", "reads"))
counts %<>% mutate(dataset = map(file, str_extract, datasets$dataset) %>%
                     map_chr(purrr::compose(dplyr::first, sort, unique)),
                   seq.run = map(file, str_extract, datasets$seq_run) %>%
                     map_chr(purrr::compose(dplyr::first, sort, unique)),
                   well = str_extract(file, "[A-H]\\d\\d?"),
                   shard = str_extract(file, "\\bx[a-z][a-z]\\b"),
                   is.shard = !is.na(shard),
                   region = str_extract(file, "(ITS[12]|LSU)"),
                   is.trim = str_detect(file, "trim"),
                   is.demux = is.na(region) & !is.trim & !is.na(well),
                   type = ifelse(is.trim, "trim",
                                 ifelse(!is.na(region), "ITSx",
                                        ifelse(is.demux, "demux",
                                               "original"))),
                   plate = str_extract(file, "00\\d") %>% replace_na("001")) %>%
  group_by(dataset, type, plate, region, is.shard) %>%
  summarize(reads = sum(reads)) %>%
  ungroup() %>%
  select(-is.shard) %>%
  unique() %>%
  spread(key = "type", value = "reads") %>%
  select(dataset, plate, original, demux, region, ITSx, trim) %>%
  filter(!is.na(dataset)) %>%
  group_by(dataset) %>%
  mutate(original = na.omit(original),
         demux = na.omit(demux)) %>%
  filter(all(is.na(region)) | !is.na(region)) %>%
  ungroup()

demux.counts <- read_lines(file.path(data.dir, "demux.counts")) %>%
  str_match("([^:]+): (\\d+) sequences ([^(]+)( \\(([FR]): (\\d+), ([FR]): (\\d+)\\))?\\.") %>%
  magrittr::extract(,c(2,3,4,6,7,8, 9)) %>%
  set_colnames(c("file", "reads", "type", "d1", "n1", "d2", "n2")) %>%
  as_tibble() %>%
  mutate_at(c("reads", "n1", "n2"), as.integer) %>%
  mutate(fwd = case_when(!is.na(d1) & d1 == "F" ~ n1,
                         !is.na(d2) & d2 == "F" ~ n2,
                         TRUE ~ NA_integer_),
         rev = case_when(!is.na(d1) & d1 == "R" ~ n1,
                         !is.na(d2) & d2 == "R" ~ n2,
                         TRUE ~ NA_integer_)) %>%
  select(-(d1:n2)) %>%
  mutate(dataset = map(file, str_extract, datasets$dataset) %>%
           map_chr(purrr::compose(dplyr::first, sort, unique)),
         seq.run = map(file, str_extract, datasets$seq_run) %>%
           map_chr(purrr::compose(dplyr::first, sort, unique)),
         seq.plate = map(file, str_extract, paste0(datasets$dataset, "-[0-9]{3}")) %>%
           map_chr(purrr::compose(dplyr::first, sort, unique))) %>%
  group_by(dataset, seq.run, seq.plate,type) %>%
  summarize_at(c("reads", "fwd", "rev"), sum) %>%
  ungroup() %>%
  mutate(type = recode(type,
                       "with at least one tag match" = "any.tag",
                       "match at least 90% of tag length" = "tag.cover",
                       "with at most 3 tag mismatches" = "tag.score",
                       "with no more than one distinct tag in each direction" = "identifiable",
                       "with tags in both directions" = "both.tags",
                       "with tags appearing in correct order" = "tag.order")) %>%
  gather(key = "dir", value = "reads", reads, fwd, rev) %>%
  mutate_at("dir", str_replace, "^reads$", "") %>%
  tidyr::unite("type", type, dir, sep = ".") %>%
  spread(key = "type", value = "reads") %>%
  set_names(str_replace(names(.), "\\.$", "")) %>%
  keep(~!all(is.na(.))) %>%
  select(dataset, starts_with("any.tag"), starts_with("tag.cover"),
         starts_with("tag.score"), starts_with("identifiable"), both.tags, tag.order) %>%
  left_join(select(counts, dataset, original, demux) %>%
              filter(complete.cases(.)) %>%
              unique()) %>%
  select(dataset, original, any.tag:tag.order, demux) %>%
  left_join(select(counts, dataset, region, ITSx, trim))

dada.counts <-
  demux.counts %>%
  left_join(select(datasets, dataset = dataset, seq_run)) %>%
  mutate(file = glue("{data.dir}/{dataset}_{seq_run}.dada.seqtable.rds"),
         file = ifelse(is.na(region),
                       file,
                       str_replace(file, "\\.dada", glue(".{region}.dada"))),
         seq.tab = map(file, readRDS),
         assigned = map_int(seq.tab, sum),
         n_ASV = map_int(seq.tab, ncol),
         file = str_replace(file, "\\.dada\\.seqtable\\.rds",
                            ".dada.nochim.rds"),
         nochim = map(file, readRDS),
         nochim.reads = map_int(nochim, sum),
         nochim.ASV = map_int(nochim, ncol)) %>%
  select(-seq.tab, -file, -nochim)

dada.counts
```
```{r}
s.pb.seqs <- readRDS(glue("{data.dir}/short-pacbio_pb_483.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         pb.reads = .)
s.ion.seqs <- readRDS(glue("{data.dir}/short-ion_is_057.dada.nochim.rds")) %>%
  colSums %>%
  tibble(seq = names(.),
         ion.reads = .)

(crosstab <- full_join(s.pb.seqs, s.ion.seqs, by = "seq") %>%
  mutate(len = str_length(seq)) %>%
  replace_na(list(pb.reads = 0, ion.reads = 0))) %>%
  ggplot(aes(x = ion.reads, y = pb.reads, color = len)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10()
```

IonTorrent reads which match a PacBio ASV: `r filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
IonTorrent reads which do NOT match PacBio: `r filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
PacBio reads which match an IonTorrent ASV: `r filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
PacBio reads which do NOT match IonTorrent: `r filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`

```{r}
crosstab %>% arrange(pb.reads, ion.reads) %>%
  write.csv(file.path(data.dir, "pb_vs_ion_ASVs.csv"))
```

```{r}
demux.counts %>%
  select(dataset, region) %>%
  left_join(select(datasets, dataset = dataset, seq_run)) %>%
  mutate(file = glue("{data.dir}/{dataset}_{seq_run}.dada.taxonomy.rds"),
         file = ifelse(is.na(region),
                       file,
                       str_replace(file, "\\.dada", glue(".{region}.dada")))) %>%
  filter(file.exists(file)) %>%
  mutate(tax = map(file, readRDS)) %>%
  unnest %>%
  filter(str_detect(Taxonomy, "Inocybe")) %>%
  mutate(nreads = rowSums(select_if(., is.integer), na.rm = TRUE)) %>%
  arrange(desc(nreads)) %>%
  select(dataset, region, nreads, Taxonomy, seq) %>%
  mutate(name = paste(dataset, region, substr(map_chr(seq, digest::digest, algo = "sha1"), 1, 8), sep = "_")) %$%
  set_names(seq, name) %>%
  DNAStringSet(use.names = TRUE) %>%
  writeFasta(file.path(data.dir, "inocybe.fasta"))

`short-ion_is_057.dada.nochim` %>%
  colnames() %>%
  set_names(paste0("seq", seq_along(.))) %>%
  DNAStringSet(use.names = TRUE) %>%
  writeFasta(file.path(data.dir, "short-ion.fasta"))
```

### 2019-10-09 scripts/drake-03-preDADA.R

@@ -13,7 +13,10 @@ setup_log("predada")
 #### pre-DADA2 ####
 # single-threaded targets after itsx
 # for local runs, ITSx targets will also run here.
-predada_targets <- stringr::str_subset(od, "^derep2_")
+predada_targets <- c(
+  purrr::keep(od, startsWith, "derep2_"),
+  purrr::keep(od, startsWith, "err_")
+)
 if (length(predada_targets)) {
   flog.info("Making pre-dada targets (loop)...")
   tictoc::tic()

### 2019-10-09 scripts/drake.R

@@ -168,21 +168,25 @@ itsx_meta <- datasets %>%
   mutate_at(c("join_derep", "join_derep_map", "itsxtrim", "lsux_pos", "search5_8S"), syms)
 
 #### predada_meta ####
-# predada_meta has one row region per input file
-# it is used in targets regions, filter, and derep2
+# predada_meta has one row per region per well
+# it is used in targets regionx, filter, and derep2
 # The index is well_ID + region
 # plate_ID is defined because it will be the index for dada_meta
 flog.info("Making predada_meta.")
 predada_meta <- itsx_meta %>%
+  group_by_at(c("seq_run", "regions", "plate", "well")) %>%
+  summarize(trim_file = list(c(trim_file)),
+            positions = list(c(positions))) %>%
+  ungroup() %>%
   mutate_at("regions", str_split, ",") %>%
-  unnest(region = regions, .preserve = ends_with("_ID")) %>%
+  unnest(region = regions, .preserve = c("trim_file", "positions")) %>%
   left_join(regions, by = "region") %>%
-  mutate(well_ID = glue("{seq_run}{plate}{well}{direction}"),
+  mutate(well_ID = glue("{seq_run}{plate}{well}"),
          plate_ID = glue("{seq_run}_{plate}"),
          range_ID = glue("{region}{range}"),
-         region_file = glue("{seq_run}_{plate}-{well}{direction}-{region}.trim.fastq.gz"),
-         filter_file = glue("{seq_run}_{plate}-{well}{direction}-{region}{range}.qfilt.fastq.gz")) %>%
-  mutate_at("positions", syms)
+         region_file = glue("{seq_run}_{plate}-{well}-{region}.trim.fastq.gz"),
+         filter_file = glue("{seq_run}_{plate}-{well}-{region}{range}.qfilt.fastq.gz")) %>%
+  mutate_at("positions", lapply, syms)
 
 if (interactive()) {
   #use a subset
@@ -201,18 +205,19 @@ if (interactive()) {
 }
 
 #### dada_meta ####
-# dada_meta has one row per region per plate
+# dada_meta has one row per region per sequencing run
 # it is used for targets err, dada, dadamap, seq_table, and nochim
-# The index is plate_ID + region (plate_ID)
+# The index is seq_run + region
 flog.info("Making dada_meta.")
 dada_meta <- predada_meta %>%
-  select(seq_run, plate, region, range, plate_ID, range_ID, seq_run) %>%
+  select(seq_run, region, range, range_ID) %>%
[…]

### 2019-10-09 scripts/scratch.R

@@ -770,3 +770,337 @@ ggtree(groupOTU(longtree, split(treetax$seq, treetax$Kingdom))) + aes(color = gr
 theme(legend.position = NULL)
 p + aes(color = Kingdom) 
 
+reconstruct_ITS <- 
+  readd(dada_map_pb_500_001_ITS1) %>%
+  select(seq.id, ITS1 = dada.seq) %>%
+  inner_join(
+    readd(dada_map_pb_500_001_5_8S) %>%
+               select(seq.id, `5_8S` = dada.seq),
+             by = "seq.id") %>%
+  inner_join(
+    readd(dada_map_pb_500_001_ITS2) %>%
+               select(seq.id, ITS2 = dada.seq),
+             by = "seq.id") %>%
+  full_join(
+    readd(dada_map_pb_500_001_ITS) %>%
+      select(seq.id, ITS = dada.seq),
+    by = "seq.id") %>%
+  mutate(ITS = coalesce(ITS, paste0(ITS1, `5_8S`, ITS2))) %>%
+  select(seq.id, ITS)
+
+is_bimera_ITS <- reconstruct_ITS %>%
+  group_by(ITS) %>%
+  summarize(abundance = n()) %>%
+  ungroup() %>%
+  dplyr::rename(sequence = ITS) %>%
+  dada2::isBimeraDenovo()
+
+reconstruct_LSU <- 
+  readd(dada_map_pb_500_001_LSU1) %>%
+  select(seq.id, LSU1 = dada.seq) %>%
+  inner_join(
+    readd(dada_map_pb_500_001_D1) %>%
+      select(seq.id, D1 = dada.seq),
+    by = "seq.id") %>%
+  inner_join(
+    readd(dada_map_pb_500_001_LSU2) %>%
+      select(seq.id, LSU2 = dada.seq),
+    by = "seq.id") %>%
+  inner_join(
+    readd(dada_map_pb_500_001_D2) %>%
+      select(seq.id, D2 = dada.seq),
+    by = "seq.id") %>%
+  inner_join(
+    readd(dada_map_pb_500_001_LSU3) %>%
+      select(seq.id, LSU3 = dada.seq),
+    by = "seq.id") %>%
+  inner_join(
+    readd(dada_map_pb_500_001_D3) %>%
+      select(seq.id, D3 = dada.seq),
+    by = "seq.id") %>%
+  inner_join(
+    readd(dada_map_pb_500_001_LSU4) %>%
[…]

### 2019-10-09 scripts/taxonomy_scratch.R

@@ -1,0 +1,495 @@
+library(drake)
+library(magrittr)
+loadd(conseq)
+conseq %<>% dplyr::filter(!is.na(long)) %>%
+  dplyr::mutate(label = tzara::seqhash(long))
+lsu_consensus <- dplyr::filter(conseq, !is.na(LSU)) %$%
+  set_names(LSU, label)
+its_consensus <- dplyr::filter(conseq, !is.na(ITS)) %$%
+  set_names(chartr("U", "T", ITS), label)
+long_consensus <- conseq %$%
+  set_names(long, label)
+idtaxaenv <- new.env()
+load("reference/rdp_train.idtaxa.Rdata", envir = idtaxaenv)
+rdp_train_idtaxa <- idtaxaenv$trainingSet
+remove("idtaxaenv")
+tax_idtaxa <- DECIPHER::IdTaxa(Biostrings::RNAStringSet(lsu_consensus),
+                               trainingSet = rdp_train_idtaxa,
+                               strand = "top",
+                               threshold = 40,
+                               processors = 3)
+tictoc::tic()
+tax_sintax <- sintax(seq = chartr("U", "T", lsu_consensus),
+                     db = "reference/rdp_train.LSU.sintax.fasta.gz",
+                     multithread = 3)
+tictoc::toc()
+tictoc::tic()
+tax_dada2 <- dada2::assignTaxonomy(chartr("U", "T", lsu_consensus),
+                                   refFasta = "reference/rdp_train.LSU.dada2.fasta.gz",
+                                   tryRC = FALSE,
+                                   outputBootstraps = TRUE,
+                                   verbose = TRUE,
+                                   multithread = 3)
+tictoc::toc()
+
+idtaxaenv <- new.env()
+load("reference/warcup.idtaxa.Rdata", envir = idtaxaenv)
+warcup_idtaxa <- idtaxaenv$trainingSet
+tax_its_idtaxa <- DECIPHER::IdTaxa(Biostrings::DNAStringSet(its_consensus),
+                                   trainingSet = warcup_idtaxa,
+                                   strand = "top",
+                                   threshold = 40,
+                                   processors = 3)
+tictoc::tic()
+tax_its_sintax <- sintax(seq = its_consensus,
+                         db = "reference/warcup.ITS.sintax.fasta.gz",
+                         multithread = 3)
+tictoc::toc()
+tictoc::tic()
+tax_its_dada2 <- dada2::assignTaxonomy(its_consensus,
[…]

### 2019-10-10 writing/transect_paper.Rmd
work on methods section

@@
           - demultiplexed and primers removed with `cutadapt` [@martin2011]
-      - ITS1, ITS2, ITS, LSU extracted using ITSx [@Nilsson2010; @bengtsson-palme2013]
+      - Illumina MiSeq
+          - pending
+      - Region extraction with LSUx (available as R package on github and conda)
+          - 5.8S located with Rfam [@kalvari2018] covariance model (CM) using `cmsearch` from Infernal 1.1.2 [@nawrocki2013]
+          - read trimmed at 5' end of 5.8S; remaining read aligned to modified 5.8S + LSU (=32S) CM from Ribosomal Data Project (RDP) [@glockner2017, @cole2014] with annotations for different LSU domains (definitions from @michot1984 and @raue1988), using `cmalign` from Infernal 1.1.2
+          - 5.8S, ITS2, D1, D2, D3, and conserved LSU domains (called LSU1, LSU2, LSU3, LSU4) extracted from alignment using annotations; ITS1 is everything before 5.8S.
       - extracted regions independently quality filtered by max expected error using `dada2::filterAndTrim()`
@@
       - de novo chimera removal using `dada2::removeBimeraDenovo()`
+      - full length amplicons reconstructed from ASVs of extracted domains (r package tzara -- available on github and conda)
+          - when a read is mapped to ASVs for all regions, they are concatenated.
+          - bimera detection on each trio of variable-conserved-variable domains (e.g., ITS1-5.8S-ITS2, ITS2-LSU1-D1, etc.); reads which are detected at bimera for any trio are discarded.
+          - For each ITS2 ASV, check each other domain:
+              - If some reads are mapped to ASVs, but some are not, assign the unassigned to the closest ASV as long as distance < 10. This is less stringent than DADA2.
+              - If no reads are mapped to ASVs, calculate consensus sequence:
+                  - align reads using `DECIPHER` with RNA secondary structure calculation [@wright2015;@wright2016]
+                  - remove outliers (i.e. chimeras) with `odseq` [@jehl2015]
+                  - remove gap columns (>50%)
+                  - take consensus (>50%) at non-gaps
+                  - remove sequences with more than 3 ambiguous bases (bad consensus)
+          - repeat process using each other domain as reference, in decreasing order of diversity.
+
       - taxonomy assigned using IDTAXA [@murali2018]
@@
       - consensus long read and LSU region constructed for each ITS2 ASV:
-          - align reads using `DECIPHER` with RNA secondary structure calculation [@wright2015;@wright2016]
-          - remove outliers (i.e. chimeras) with `odseq` [@jehl2015]
-          - remove gap columns (>50%)
-          - take consensus (>50%) at non-gaps
-          - remove sequences with more than 3 ambiguous bases (bad consensus)
+
       - initial alignment of long amplicons with DECIPHER [@wright2015;@wright2016]

### 2019-11-07 writing/transect_paper.Rmd
methods bullet points

@@
 
-      - taxonomy assigned using IDTAXA [@murali2018]
+      - taxonomy assignment
+          - three methods: 
+              - IDTAXA [@murali2018]
+              - SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016]
+              - RDP Naïve Bayesian Classifier (NBC) [@wang2007] as implemented in DADA2
+          - three databases:
               - UNITE version 8 [@nilsson2019a]
                   - extracted ITS1 and ITS2 using ITSx [@Nilsson2010; @bengtsson-palme2013]
-              - trained three classifiers: full ITS, ITS1, ITS2.
-          - RDP fungal LSU release 11.5 [@cole2014]
+                  - trained three classifiers: full (for ITS, 5_8S, full amplicons), ITS1, ITS2.
+              - RDP fungal LSU training set 11.5 [@cole2014]
                   - chose only sequences containing conserved primer sites for ITS4 (3' side of H11; CAUAUnAnUnAGnSSAGGA) and LR5 (on H37; CGAAnUUUCnCUCAGGAUAGCnG).
@@
                   - Trim all bases after LR5 because they are not present in our amplicons.
-          - SILVA LSU release 132 [@quast2013; @glockner2017; @yilmaz2014]
-              - filter out sequences annotates as prokaryotes, then treat as RDP.
-          - all databases
-              - select medioid sequence for each annotated species, based on minimum 5-mer distance to other sequences
-              - for sequences not annotated as species (e.g. "unidentified Basidiomycota"), randomly choose up to 20 per taxon
-              - Use up to 5 iterations of IDTAXA training, removing all "problem" sequences on each iteration, unless last member of annotated taxon.
-              - "underannotated" sequences (e.g., "unidentified Basidiomycota" confidently placed as Agaricomycetes) not protected from removal if last sequence annotated as such.
+                  - used to identify long amplicon, 32S, LSU
+                  - taxonomy for non-fungi looked up from NCBI based on accession numbers
+              - RDP Warcup fungal ITS training set [@wang2007]
+                  - ITS1 and ITS2 extracted as for Unite;
+          - taxonomic classification in the three datasets were standardized to @tedersoo2017c
+              - full classification to genus level available in supplementary material
+              - apparently used by Unite for nonfungi; @tedersoo2018 used by Unite for fungi
+              - groups monophyletic
+              - SINTAX and NBC require a ranked system, as opposed to more accepted classifications like @adl2019
       - ecological guild assigned using FUNGuild [@nguyen2016funguild]
-      - consensus long read and LSU region constructed for each ITS2 ASV:
+      - alignment
+          - preliminary alignment of 32S to RDP covariance model with `cmalign` from Infernal (aligns only conserved regions)
+          - concatenate ITS1 to beginning of alignment
+          - create guide tree using alignment of only conserved, gap-free columns in preliminary alignment, using UPGMA in DECIPHER [@wright2016]
+          - realign nonconserved sites using MLocARNA (progressive simultaneous alignment and folding) from LocARNA 2.0.0RC8 [@will2007; @will2012; @will2013].
 
-      - initial alignment of long amplicons with DECIPHER [@wright2015;@wright2016]
       - alignment/tree for long amplicons built iteratively with PASTA[@mirarab2014; @mirarab2014a]. Fastree tree, mafft align, OPAL merge, RaxML final tree

### 2019-11-07 writing/transect_paper.Rmd
sampling methods

@@
 
+Sampling was conducted at two sites approximately 30 km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+At each site, 25 soil samples were collected at intervals of 1 m.
+For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a knife blade.
+Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
+Approximately 50 mg total of soil was collected from two locations in the homogenized soil and placed in Xpedition lysis solution (Zymo Research) and lysed in the field using a handheld bead-beater (Terralyzer; Zymo Research).
+The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis.
+
   - Ouémé Forest reserve
@@
           - bimera detection on each trio of variable-conserved-variable domains (e.g., ITS1-5.8S-ITS2, ITS2-LSU1-D1, etc.); reads which are detected at bimera for any trio are discarded.
-          - For each ITS2 ASV, check each other domain:
-              - If some reads are mapped to ASVs, but some are not, assign the unassigned to the closest ASV as long as distance < 10. This is less stringent than DADA2.
-              - If no reads are mapped to ASVs, calculate consensus sequence:
-                  - align reads using `DECIPHER` with RNA secondary structure calculation [@wright2015;@wright2016]
-                  - remove outliers (i.e. chimeras) with `odseq` [@jehl2015]
-                  - remove gap columns (>50%)
-                  - take consensus (>50%) at non-gaps
-                  - remove sequences with more than 3 ambiguous bases (bad consensus)
-          - repeat process using each other domain as reference, in decreasing order of diversity.
 
@@
               - RDP fungal LSU training set 11.5 [@cole2014]
-                  - chose only sequences containing conserved primer sites for ITS4 (3' side of H11; CAUAUnAnUnAGnSSAGGA) and LR5 (on H37; CGAAnUUUCnCUCAGGAUAGCnG).
-                  - ambiguities based on [@cannone2002; @tedersoo2015]
-                  - Trim all bases after LR5 because they are not present in our amplicons.
-                  - used to identify long amplicon, 32S, LSU
+                  - used to identify long amplicon, LSU
                   - taxonomy for non-fungi looked up from NCBI based on accession numbers
@@
       - ecological guild assigned using FUNGuild [@nguyen2016funguild]
-      - alignment
+      - alignment as RNA in DECIPHER (iterative progressive alignment with RNS secondary structure prediction) [@wright2015]
+      - tree using RAxML 8.2 [@stamatakis2014]; GTR+GAMMA; rapid bootstrapping with MRE-IGN stopping criterion [@pattengale2010].
+
+<!--
           - preliminary alignment of 32S to RDP covariance model with `cmalign` from Infernal (aligns only conserved regions)
@@
           - create guide tree using alignment of only conserved, gap-free columns in preliminary alignment, using UPGMA in DECIPHER [@wright2016]
-          - realign nonconserved sites using MLocARNA (progressive simultaneous alignment and folding) from LocARNA 2.0.0RC8 [@will2007; @will2012; @will2013].
+          - realign nonconserved sites using MLocARNA (progressive simultaneous alignment and folding) from LocARNA 2.0.0RC8 [@will2007; @will2012; @will2013]. -->
 

### 2019-11-08 writing/transect_paper.Rmd
wrote a paragraph about LSU region extraction.

@@
 
-  - Ouémé Forest reserve
-  - Two locations in _Isoberlinia doka_ forest, approx. 30km apart
-  - transect: 25m, sampled at interval of 1m
-  - Sampled in May 2015 and June 2016
-  - remove coarse debris (typically very little) and take ~5cm×5cm×5cm sample
-  - homogenize by hand in ziplock bag
-  - take $\approx$ 50mg, lysis in field with Terralyzer
-  - remainder of sample dried at 40°C with electric dryer within 24-48 hours
-
 ## Soil Chemistry
@@
 
-  - lysis in field Zymo Research Xpedition kit
+After field lysis, DNA was extracted using the Zymo Research Soil/Fecal Prep kit **check name**.
+DNA for short amplicons was amplified using the primer pairs gITS7[@ihrmark2012]--ITS4[@white1990amplification].
   - short amplicon : gITS7 -- ITS4 (multiplexing index on gITS7)
+  - short PCR protocol:
   - long amplicon : ITS1 -- LR5 (multiplexing index on both ends)
+  - long PCR protocol:
   - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5
@@
       - PacBio
-          - CCS basecalls re-done with `ccs` version ___ [@pacificbiosciences2019]
+          - CCS basecalls re-done with `ccs` from SMRTtools version 5.0.1 [@pacificbiosciences2019]
           - demultiplexed and primers removed with `cutadapt` [@martin2011]
@@
           - pending
-      - Region extraction with LSUx (available as R package on github and conda)
-          - 5.8S located with Rfam [@kalvari2018] covariance model (CM) using `cmsearch` from Infernal 1.1.2 [@nawrocki2013]
-          - read trimmed at 5' end of 5.8S; remaining read aligned to modified 5.8S + LSU (=32S) CM from Ribosomal Data Project (RDP) [@glockner2017, @cole2014] with annotations for different LSU domains (definitions from @michot1984 and @raue1988), using `cmalign` from Infernal 1.1.2
-          - 5.8S, ITS2, D1, D2, D3, and conserved LSU domains (called LSU1, LSU2, LSU3, LSU4) extracted from alignment using annotations; ITS1 is everything before 5.8S.
+
+Raw reads were divided into regions/domains by matching to covariance models, which account for both nucleotide sequence and RNA secondary structure.
+First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013].
+This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013; data not shown].
+Each read was then aligned to a CM based on the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017, @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
+The seed alignment was truncated at the LR5 primer site, and the reference line was modified to annotate the 5.8S RNA, ITS2 region, and the divergent "D" domains of the 28S RNA as defined by @raue1988.
+The alignment for each read was then split into alternating more-conserved and less-conserved regions according to these annotations:
+ITS1 (for the entire read prior to the 5' end of 5.8S), 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
+where the four "LSU" regions represent the conserved regions of the 28S rRNA between the "D" regions.
+The process of extracting the regions is automated in the new R package `LSUx`, available on github and conda-forge (**but not yet**).
+
       - extracted regions independently quality filtered by max expected error using `dada2::filterAndTrim()`

### 2019-11-15 writing/transect_paper.Rmd
Write more methods.

@@
 
+Circular consensus sequence (CCS) basecalls were made using the `ccs` command from SMRTtools version 5.0.1  [@pacificbiosciences2019].
+The resulting sequences were demultiplexed and sequencing primers were removed using `cutadapt` version 1.18 [@martin2011].
+
 - Brendan's pipeline:
@@
           - used demultiplexing from SciLife (standard IonXpress)
-          - primers removed with `cutadapt` [@martin2011]
+          - primers removed with 
       - PacBio
-          - CCS basecalls re-done with `ccs` from SMRTtools version 5.0.1 [@pacificbiosciences2019]
-          - demultiplexed and primers removed with `cutadapt` [@martin2011]
       - Illumina MiSeq
@@
 
-      - extracted regions independently quality filtered by max expected error using `dada2::filterAndTrim()`
-      - dereplicated using `dada2::derepFastq()`
-      - denoised with `dada2::dada()`
+Each of the extracted regions was independently filtered for length (table **X**) and a maximum of three expected errors using the `filterAndTrim` function from DADA2 [@callahan2016, @callahan2018].
+Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2.
+The error model for DADA2 denoising was fit using only the 5.8S RNA region.
+Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons).
+Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
+
+The ITS2 region was used to identify full-length amplicons ASVs, due to the fact that it was the only full-length region/domain shared by both long and short amplicons.
+
       - de novo chimera removal using `dada2::removeBimeraDenovo()`

### 2019-11-19 writing/transect_paper.Rmd
write more methods

@@
 The error model for DADA2 denoising was fit using only the 5.8S RNA region.
-Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons).
+Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, **different sequencing technologies**).
 Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
 
-The ITS2 region was used to identify full-length amplicons ASVs, due to the fact that it was the only full-length region/domain shared by both long and short amplicons.
+The ITS2 region was used to identify full-length amplicon ASVs, due to the fact that it was the only full-length region/domain shared by both long and short amplicons.
+For each ITS2 ASV from the long amplicon data set, the denoised sequences for the other regions corresponding to the same sequencing reads were concatenated to form a set of full-length reads.
+For reads which were not assigned a denoised sequence for every region, the raw read for the region was used instead.
+Because ITS2 is one of the most variable regions in the amplified regions, reads with identical ITS2 regions are expected have extremely similar sequences in the other regions, unless they are chimeric.
+Thus, the concatenated regions for each ASV were aligned in R using the DECIPHER package [@wright2015], and outlier sequences (chimeras) were removed from each alignment using the odseq package [@jehl2015].
+The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence.
+Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
+The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available on github ([https://github.com/brendanf/tzara]), CRAN (not yet), and bioconda (not yet).
 
-      - de novo chimera removal using `dada2::removeBimeraDenovo()`
-      - full length amplicons reconstructed from ASVs of extracted domains (r package tzara -- available on github and conda)
-          - when a read is mapped to ASVs for all regions, they are concatenated.
-          - bimera detection on each trio of variable-conserved-variable domains (e.g., ITS1-5.8S-ITS2, ITS2-LSU1-D1, etc.); reads which are detected at bimera for any trio are discarded.
+# Taxonomy assignment
 
-      - taxonomy assignment
-          - three methods: 
-              - IDTAXA [@murali2018]
-              - SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016]
-              - RDP Naïve Bayesian Classifier (NBC) [@wang2007] as implemented in DADA2
-          - three databases:
-              - UNITE version 8 [@nilsson2019a]
-                  - extracted ITS1 and ITS2 using ITSx [@Nilsson2010; @bengtsson-palme2013]
-                  - trained three classifiers: full (for ITS, 5_8S, full amplicons), ITS1, ITS2.
-              - RDP fungal LSU training set 11.5 [@cole2014]
-                  - used to identify long amplicon, LSU
-                  - taxonomy for non-fungi looked up from NCBI based on accession numbers
-              - RDP Warcup fungal ITS training set [@wang2007]
-                  - ITS1 and ITS2 extracted as for Unite;
-          - taxonomic classification in the three datasets were standardized to @tedersoo2017c
-              - full classification to genus level available in supplementary material
-              - apparently used by Unite for nonfungi; @tedersoo2018 used by Unite for fungi
-              - groups monophyletic
-              - SINTAX and NBC require a ranked system, as opposed to more accepted classifications like @adl2019
+New taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@wang2007] were created, conforming to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
+In particular, the classification for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
+Although this system is not formally published, it seems to be in use for non-fungal eukaryotes in the Unite database, and it is a system with only purportedly monophyletic taxa and a uniform set of taxon ranks.
+These make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as @adl2019.
+FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).
+
+Taxonomic assignment was performed to genus level on the ITS and LSU regions using Unite/Warcup and RDP, respectively, as taxonomic references.
+For each region/reference combination, taxonomy was assigned using three algorithms:
+the RDP Naïve Bayesian Classifier (NBC) as implemented in DADA2;
+SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
+and IDTAXA [@murali2018].
+Each ASV thus was thus given nine taxonomic assignments (three region/reference combinations $\times$ three algorithms).
+
       - ecological guild assigned using FUNGuild [@nguyen2016funguild]

### 2019-11-20 writing/transect_paper.Rmd
writing methods

@@
 library("tikzDevice")
+tryCatch(
+  tinytex::tinytex_root(),
+  error = function(e) tinytex::install_tinytex()
+  )
 knitr::opts_chunk$set(dev="tikz",
@@
 
-## DNA extraction, PCR, and sequencing
+## DNA extraction, amplification, and sequencing
 
 After field lysis, DNA was extracted using the Zymo Research Soil/Fecal Prep kit **check name**.
-DNA for short amplicons was amplified using the primer pairs gITS7[@ihrmark2012]--ITS4[@white1990amplification].
-  - short amplicon : gITS7 -- ITS4 (multiplexing index on gITS7)
-  - short PCR protocol:
-  - long amplicon : ITS1 -- LR5 (multiplexing index on both ends)
-  - long PCR protocol:
+DNA was quantified using PicoGreen *according to the manufacturer's protocol*?
+
+Two different regions of rDNA were amplified (figure).
+The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and LSU rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] as the reverse primer.
+
+DNA for short amplicons was amplified using gITS7 [@ihrmark2012] as the forward primer and a equimolar mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] as the reverse primer (hereafter ITS4m).
+gITS7 primers were indexed for multiplexing (**supp info**).
+Amplification was performed by polymerase chain reaction (PCR) in 60µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.3 U *taq* polymerase (Dream *Taq*, Thermo Fisher) and 10--20 ng purified DNA in Dream *Taq* buffer.
+The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
+Each reaction was split into three aliquots and amplified separately to reduce the effect of PCR stochasticity (**ref**), and then pooled after amplification.
+
+```{tikz rDNA_amplicons, fig.cap = "rDNA amplicons", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
+
+```
+
+DNA for long amplicons was extracted using ITS1 [@white1990amplification] for the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+Both primers were indexed for multiplexing (**supp info**).
+PCR was performed in 60 µL reactions as for the short amplicons, but with 500 µM of each of the two primers.
+Reaction conditions were 10 min at 95°, 30 cycles of 45 s at 95°, 45 s at 59°, and 90 s at 72°, and finally 10 min at 72°.
+Each reaction was split into three aliquots as for short amplicons.
+
   - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5

### 2019-11-20 scripts/lsux.R

@@ -257,7 +257,7 @@ LSUx <- function(seq, cm_5.8S, cm_32S, glocal = TRUE, ITS1 = FALSE, cpu) {
   futile.logger::flog.info("Extracting LSU regions.", name = "LSUx")
   pos <- extract_LSU(aln = aln$alignment, rf = aln$RF)
   pos <- dplyr::mutate_at(pos, "seq_name", stringr::str_replace, "^[^|]*\\|", "")
-  pos <- dplyr::mutate_if(pos, is.integer, add, cms$seq_from - 1L)
+  pos <- dplyr::mutate_if(pos, is.integer, '+', cms$seq_from - 1L)
   if (isTRUE(ITS1)) {
     no_ITS1 <- pos$`5_8S_start` == 1 | is.na(pos$`5_8S_start`)
     if (all(no_ITS1)) {

### 2019-11-21 writing/transect_paper.Rmd
more methods

@@
 
-Two different regions of rDNA were amplified (figure).
-The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and LSU rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] as the reverse primer.
+Two different regions of the nuclear ribosomal DNA were amplified (figure \@ref(fig:rDNA)).
+The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) ribosomal RNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] (hereafter ITS4m) as the reverse primer.
+The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the ribosomal small subunit (SSU) RNA, the full ITS region including the 5.8S rRNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable domains.
 
-DNA for short amplicons was amplified using gITS7 [@ihrmark2012] as the forward primer and a equimolar mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] as the reverse primer (hereafter ITS4m).
-gITS7 primers were indexed for multiplexing (**supp info**).
-Amplification was performed by polymerase chain reaction (PCR) in 60µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.3 U *taq* polymerase (Dream *Taq*, Thermo Fisher) and 10--20 ng purified DNA in Dream *Taq* buffer.
+The gITS7 primers for the short amplicon were indexed for multiplexing (**supp info**).
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher) and 3--7 ng purified DNA in Dream *Taq* buffer.
 The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
-Each reaction was split into three aliquots and amplified separately to reduce the effect of PCR stochasticity (**ref**), and then pooled after amplification.
+Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity (**ref**), which were pooled after amplification.
 
-```{tikz rDNA_amplicons, fig.cap = "rDNA amplicons", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
+```{tikz rDNA, fig.cap = "rDNA amplicons", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
 
@@
 
-DNA for long amplicons was extracted using ITS1 [@white1990amplification] for the forward primer and LR5 [@vilgalys1990] as the reverse primer.
-Both primers were indexed for multiplexing (**supp info**).
-PCR was performed in 60 µL reactions as for the short amplicons, but with 500 µM of each of the two primers.
+Both primers for the long amplicon were indexed for multiplexing (**supp info**).
+PCR was performed as for the short amplicons, but with 500 µM of each of the two primers.
 Reaction conditions were 10 min at 95°, 30 cycles of 45 s at 95°, 45 s at 59°, and 90 s at 72°, and finally 10 min at 72°.
-Each reaction was split into three aliquots as for short amplicons.
+Each reaction was performed in three technical replicates as for short amplicons.
 
@@
 
-  - Mohammad's pipeline:
-
 Circular consensus sequence (CCS) basecalls were made using the `ccs` command from SMRTtools version 5.0.1  [@pacificbiosciences2019].
@@
 
-- Brendan's pipeline:
-      - Ion Torrent
-          - used demultiplexing from SciLife (standard IonXpress)
-          - primers removed with 
-      - PacBio
-      - Illumina MiSeq
-          - pending
-
-Raw reads were divided into regions/domains by matching to covariance models, which account for both nucleotide sequence and RNA secondary structure.
+Raw reads were divided into regions/domains by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
 First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013].
 This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013; data not shown].
-Each read was then aligned to a CM based on the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017, @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
-The seed alignment was truncated at the LR5 primer site, and the reference line was modified to annotate the 5.8S RNA, ITS2 region, and the divergent "D" domains of the 28S RNA as defined by @raue1988.
-The alignment for each read was then split into alternating more-conserved and less-conserved regions according to these annotations:
-ITS1 (for the entire read prior to the 5' end of 5.8S), 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
-where the four "LSU" regions represent the conserved regions of the 28S rRNA between the "D" regions.
-The process of extracting the regions is automated in the new R package `LSUx`, available on github and conda-forge (**but not yet**).
+Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017, @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
+The seed alignment modified by truncation at the LR5 primer site, and the reference line was modified to annotate the 5.8S RNA, ITS2 region, and the divergent "D" domains of the 28S RNA as defined by @raue1988.
+The CM alignment for each read was then split into alternating more-conserved and less-conserved regions according to these annotations:
+ITS1, 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
+where the four "LSU" regions represent the conserved regions of the 28S rRNA flanking the "D" regions.
+For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
+No attempt was made to remove the short fragment of the SSU from the 5' end of ITS1 in the long amplicons; this is too short to be reliably detected by a CM or the HMMs employed by ITSx.
+The process of extracting the regions is automated in the new R package `LSUx`, available on github, CRAN, and bioconda (**but not yet**).
 
-Each of the extracted regions was independently filtered for length (table **X**) and a maximum of three expected errors using the `filterAndTrim` function from DADA2 [@callahan2016, @callahan2018].
+Each of the extracted regions was independently filtered for length (table **X**) and a maximum of three expected errors.
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2.
@@
 Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
-The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available on github ([https://github.com/brendanf/tzara]), CRAN (not yet), and bioconda (not yet).
+A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
+The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available on github (https://github.com/brendanf/tzara), CRAN (not yet), and bioconda (not yet).
 
-# Taxonomy assignment
+### Taxonomy assignment
 
@@
 
+## Alignment and phylogenetic inference
+
+Long amplicon ASVs were aligned as RNA using DECIPHER [@wright2015].
+As attempt was made to refine this alignment using Pasta [@mirarab2014a], but the resulting maximum liklihood (ML) trees did not improve on the ML achieved using the original alignment from DECIPHER.
+Since ML is the optimization criterion used by Pasta, the original DECIPHER alignment was retained.
+A final ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
+
+The tree was rooted outside the fungi by using the most abundant ASV which was confidently assigned to the Chlorophyta by all 6 applicable taxonomic assignment methods.
+Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
+Inspection of the tree along with taxonomic assignments revealed that the genus *Tulasnella* (Basidiomycota), which is known to have accelerated evolution in the rDNA [@moncalvo2006], was nested within the Metazoa, so it was excluded from further analysis.
+The kingdom Fungi was then identified as the minimal clade containing all remaining ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
+ASVs falling outside this clade were excluded from further analysis.
+
       - ecological guild assigned using FUNGuild [@nguyen2016funguild]
@@
 
-      - alignment/tree for long amplicons built iteratively with PASTA[@mirarab2014; @mirarab2014a]. Fastree tree, mafft align, OPAL merge, RaxML final tree
-
 ## Statistics

### 2019-12-02 writing/transect_paper.Rmd
methods and intro outline

@@
 # Introduction
+Send this outline by Thursday Dec 5
+Meet Monday Dec 2 about
+
+## Introduction to ECM communities
+
+## Turnover scale in ECM forests
+New ecosystem type
+
+## Long reads vs short reads/phylogenetic methods
+Phylogenetic methods
 
@@
 
-  - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5
-  - long amplicon : PacBio RS II
+Amplicons were purified using SPRI beads [Vesterinen et al 2016] and quantified fluorometrically using PicoGreen dsDNA (Quant-iT) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan) according to the manufacturer's protocol.
+100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
+Each library was sequenced using a Pacific Biosciences RS II sequences.
+Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
 
@@
 Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017, @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
-The seed alignment modified by truncation at the LR5 primer site, and the reference line was modified to annotate the 5.8S RNA, ITS2 region, and the divergent "D" domains of the 28S RNA as defined by @raue1988.
-The CM alignment for each read was then split into alternating more-conserved and less-conserved regions according to these annotations:
+The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regionss:
 ITS1, 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
@@
 For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
-No attempt was made to remove the short fragment of the SSU from the 5' end of ITS1 in the long amplicons; this is too short to be reliably detected by a CM or the HMMs employed by ITSx.
+No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx.
 The process of extracting the regions is automated in the new R package `LSUx`, available on github, CRAN, and bioconda (**but not yet**).
@@
 Each of the extracted regions was independently filtered for length (table **X**) and a maximum of three expected errors.
-Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2.
+Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")`.
 The error model for DADA2 denoising was fit using only the 5.8S RNA region.
@@
 
-The ITS2 region was used to identify full-length amplicon ASVs, due to the fact that it was the only full-length region/domain shared by both long and short amplicons.
 For each ITS2 ASV from the long amplicon data set, the denoised sequences for the other regions corresponding to the same sequencing reads were concatenated to form a set of full-length reads.
@@
 New taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@wang2007] were created, conforming to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
-In particular, the classification for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
-Although this system is not formally published, it seems to be in use for non-fungal eukaryotes in the Unite database, and it is a system with only purportedly monophyletic taxa and a uniform set of taxon ranks.
+In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
+Although the latter system is not formally published, it seems to be in use for non-fungal eukaryotes in the Unite database, and it is a system with only purportedly monophyletic taxa and a uniform set of taxon ranks.
 These make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as @adl2019.
@@
 
-      - ecological guild assigned using FUNGuild [@nguyen2016funguild]
-      - alignment as RNA in DECIPHER (iterative progressive alignment with RNS secondary structure prediction) [@wright2015]
-      - tree using RAxML 8.2 [@stamatakis2014]; GTR+GAMMA; rapid bootstrapping with MRE-IGN stopping criterion [@pattengale2010].
-
 <!--
@@
 
+Ecological community dissimilarity matrices were calculated using methods "bray" (both long and short amplicons) and "unifrac" (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
+Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m,
+
 # Results

### 2019-12-05 writing/transect_paper.Rmd
Don't refine MAFFT alignment, this takes FOREVER (and is throttled to 8 cores)

@@
 # Introduction
-Send this outline by Thursday Dec 5
-Meet Monday Dec 2 about
 
@@
 ## Turnover scale in ECM forests
-New ecosystem type
+
+- Meta analysis @Bahram2013
+    -
 
@@
 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
-Each library was sequenced using a Pacific Biosciences RS II sequences.
+Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGS) at SciLifeLab in Uppsala, Sweden.
 Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
+
+Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGS.
 
@@
 
-Circular consensus sequence (CCS) basecalls were made using the `ccs` command from SMRTtools version 5.0.1  [@pacificbiosciences2019].
+Circular consensus sequence (CCS) basecalls for SMRT sequences were made using `ccs` version 3.4 [@pacificbiosciences2019] using the default settings.
 The resulting sequences were demultiplexed and sequencing primers were removed using `cutadapt` version 1.18 [@martin2011].
+Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible.
 
@@
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")`.
-The error model for DADA2 denoising was fit using only the 5.8S RNA region.
-Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, **different sequencing technologies**).
+The error model for DADA2 denoising was fit using the 5.8S RNA region for long amplicons, and using the entire read for short amplicons.
+Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
 Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
@@
 
-New taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@wang2007] were created, conforming to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
+Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@wang2007] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
 In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
-Although the latter system is not formally published, it seems to be in use for non-fungal eukaryotes in the Unite database, and it is a system with only purportedly monophyletic taxa and a uniform set of taxon ranks.
-These make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as @adl2019.
+Although the latter system is not formally published, it seems to be in use for non-fungal eukaryotes in the Unite database.
+Additionally, it is a system with only purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as @adl2019.
 FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).
@@
 and IDTAXA [@murali2018].
-Each ASV thus was thus given nine taxonomic assignments (three region/reference combinations $\times$ three algorithms).
+Each ASV thus was thus given up to nine taxonomic assignments (three region/reference combinations $\times$ three algorithms).
 
@@
 
-Long amplicon ASVs were aligned as RNA using DECIPHER [@wright2015].
-As attempt was made to refine this alignment using Pasta [@mirarab2014a], but the resulting maximum liklihood (ML) trees did not improve on the ML achieved using the original alignment from DECIPHER.
+Unique LSU regions from long amplicon ASVs were aligned as RNA using DECIPHER [@wright2015] with up to 10 iterations of progressive alignment and conserved secondary structure calculation and 10 refinement iterations.
+This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had type 1 introns after this position [@holst-jensen1999].
+Full length long amplicon ASVs (including ITS1, 5.8S, and ITS2 regions) were aligned and truncated in the same way.
+
+An attempt was made to refine the long amplicon alignment using Pasta [@mirarab2014a], but the resulting maximum likelihood (ML) trees did not improve on the ML achieved using the original alignment from DECIPHER.
 Since ML is the optimization criterion used by Pasta, the original DECIPHER alignment was retained.
-A final ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
+
+ML trees were produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
+The LSU tree was not constrained topologically, but the long amplicon tree was constrained by the result for the LSU tree.
+
+Short amplicon ASVs without a perfect match to one of the long amplicon ASVs were added to the long amplicon alignment using the `--add` option of MAFFT version 7.453 [@katoh2012] with a 10-mer based guide tree and two iterations.
+The short ASVs were then placed on the long amplicon tree using EPA-NG version 0.3.6 [@barbera2019].
+A "grafted" tree including the short amplicon ASVs in the most probable positions from the EPA-NG results was produced using GAPPA version 0.5.1 [@czech2019a].
+When more than one short amplicon ASV was placed in the same position on the tree, GAPPA by default groups all of these sequences together in a polytomy.
+The branch leading to each of these polytomies was deleted by setting its length to zero and then using the `d2multi` function from the `ape` package in R.
+The resulting tree was used as a guide tree to construct a final ML tree in RAxML, using the GTR+GAMMA model, in order to resolve relationships between different sequences assigned to the same branch.
+No bootstrapping was performed on the final tree.
 

### 2019-12-05 scripts/drake.R

@@ -684,7 +684,7 @@ plan <- drake_plan(
       y = .,
       add = "add",
       method = "10merpair",
-      maxiterate = "1000",
+      maxiterate = "2",
       thread = ignore(ncpus),
       quiet = FALSE,
       exec = Sys.which("mafft")

### 2019-12-05 writing/transect_paper.Rmd
Intro outline

@@
 
+- Ectomycorrhizas (ECM) are a symbiosis between fungi and plant roots.
+- x% of plants, y% of fungi
+- The plants are often ecologically and economically important trees (Pinaceae, Fagaceae, Dipterocarpaceae, Amherstiae)
+- The fungi provide nutrients and protection from pathogens, the plant provides sugar.
+- Many well-known mushrooms are produced by ECM (Amanita, Cantharellus, Boletus) but many ECM produce inconspicuous (Tomentella) or no (Cenococcum) fruit bodies.
+- Anyway the fruitbodies are ephemeral, so study of ECM communities requires looking at vegetative structures.
+- Not easy to culture, so we need molecular methods.
+
 ## Turnover scale in ECM forests
 
+- @lilleskov2004 - autocorrelation only <2.6 m at most sites.  Based on Sanger sequencing root tips in temperate forests.
+
 - Meta analysis @Bahram2013
-    -
+  - Includes data from Benin (same site? ask Nourou/Leho) but this is unpublished
+  - conclude that there is greater range of distance decay in tropical forests; because of less host availability? older forests?  *Russulaceae*?
+  - 65 m in Benin
+
+- @pickles2012a
+  - T-RFLP of root tips, temperate forest
+  - Autocorrelation < 3.4 m
+
+- Data from African wooded savannas - @tedersoo2011 but only southern band
+  - Irregular sampling with no samples nearer than 8m
+  - based on sanger sequencing root tips
+  - Significant Moran's I only in <10m class
 
 ## Long reads vs short reads/phylogenetic methods
-Phylogenetic methods
+
+- Clustering vs. phylogenetic methods
+    - Clustering-based OTU approach may both "clump" and "split" species.
+    - Distance-based clustering conflates intra-species variation and sequencing error
+    - Denoising (DADA2 [@callahan2017], Deblur [@amir2017], UNOISE2 [@edgar2016]) attempts to control for sequencing error while leaving intra-species variation.
+    - Phylogenetic community distance measures like UNIFRAC [@wong2016; look for earlier reference] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
+
+- Short reads
+    - 454 [@buee2009], Ion Torrent, Illumina [@schmidt2013].
+    - For fungi, frequently ITS1 or ITS2 [@schoch2012]
+    - Dependent on matching to database for taxonomic placement.
+    - Phylogenetic placement algorithms [@matsen2010; @berger2011] are not easy to apply because ITS is not in most LSU alignments (different situation than V4/V5 of 18S)
+    - Possible to place on taxonomy-based tree [@tedersoo2018]
+
+- Long reads
+    - Can include both LSU/SSU and ITS in same amplicon for better taxonomic placement [@Tedersoo2018]
+    - can build tree directly
+    - short reads from same location can be mapped to tree.
+    - Sequencing depth tradeoff [@Kennedy2018]
 
@@
 Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25 soil samples were collected at intervals of 1 m.
+At each site, 25 soil samples were collected at intervals of 1 m in May 2015.
+Additional samples were collected from one third of the sample locations (3-m spacing) one year later in June 2016.
 For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a knife blade.
@@
   - Need to get methods from Nourou
+  - Or just not use it?
 

### 2019-12-17 writing/ISP_update.Rmd
update tech comparison presentation and create ISP followup presentation

---
title: "ISP 1-month followup"
author: "Brendan Furneaux"
date: "12 Dec 2019"
output:
  beamer_presentation:
    latex_engine: xelatex
---

```{r setup, include = FALSE}
library(magrittr)
library(tidyverse)
library(here)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")
library(drake)

cache <- drake_cache(here(".light"))
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
```

# Read number from long PacBio vs. IonTorrent (ITS2)
```{r, message=FALSE}
loadd(big_seq_table_ITS2, cache = cache)
asv_table <- big_seq_table_ITS2 %>%
  tibble::as_tibble(rownames = "filename") %>%
  tidyr::gather(key = "seq", value = "reads", -1) %>%
  dplyr::filter(reads >= 1) %>%
  tidyr::extract(col = "filename",
                 into = c("seq.run", "plate", "well", "dir", "region"),
                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
  dplyr::group_by(seq.run, seq) %>%
  dplyr::summarize(reads = sum(reads)) %>%
  tidyr::spread(key = seq.run, value = reads, fill = 0)

otu <- read_tsv(here("data/clusters/ITS2.table")) %>%
  column_to_rownames("#OTU ID") %>%
  as.matrix() %>%
  t() %>%
  as_tibble(rownames = "sample") %>%
  tidyr::extract(
    col = "sample",
    into = c("seq_run", "plate", "well"),
    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
  ) %>%
  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
  dplyr::filter(reads >= 1) %>%
  dplyr::group_by(seq_run, seq) %>%
  dplyr::summarize(reads = sum(reads)) %>%
  tidyr::spread(key = seq_run, value = reads, fill = 0)

big_table <- bind_rows(
  mutate(asv_table, type = "ASV"),
  mutate(otu, type = "OTU")
)
big_table %>%
  filter(is_057 > 0, pb_500 > 0) %>%
  ggplot(aes(x = is_057, y = pb_500)) +
  geom_point(alpha = 0.2) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Ion Torrent Short Reads") +
  ylab("PacBio Long Reads") +
  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(is_057 == 0, pb_500 > 0)) +
  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
           data = big_table %>% filter(pb_500 == 0, is_057 > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
                    mean(log10(is_057[is_057 >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  facet_wrap(~type)
```

# Read number from short PacBio vs. IonTorrent (ITS2)
```{r}
big_table %>% 
  filter(is_057 > 0, pb_483 > 0) %>%
  ggplot(aes(x = is_057, y = pb_483)) +
  geom_point(alpha = 0.2) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Ion Torrent Short Reads") +
  ylab("PacBio Short Reads") +
  geom_rug(aes(y = pb_483, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(is_057 == 0, pb_483 > 0)) +
  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
           data = big_table %>% filter(pb_483 == 0, is_057 > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10(pb_483[pb_483 >= 1])) -
                    mean(log10(is_057[is_057 >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  facet_wrap(~type)

```

# Read number from long vs. short PacBio ASVs (ITS2)
```{r}
big_table %>% 
  filter(pb_483 > 0, pb_500 > 0) %>%
  ggplot(aes(x = pb_483, y = pb_500)) +
  geom_point(alpha = 0.2) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("PacBio Short Reads") +
  ylab("PacBio Long Reads") +
  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(pb_483 == 0, pb_500 > 0)) +
  geom_rug(aes(y = pb_483), sides = "b", alpha = 0.2,
           data = big_table %>% filter(pb_500 == 0, pb_483 > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
                    mean(log10(pb_483[pb_483 >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  facet_wrap(~type)
```

# Mantel Correlograms - long vs. short PacBio

```{r}
loadd(
  "correlog_bray_0_long.pacbio",                                 
  "correlog_bray_0_short.ion",
  "correlog_bray_0_short.pacbio",
  "correlog_bray_1_long.pacbio",
  "correlog_bray_1_short.ion",
  "correlog_bray_1_short.pacbio",
  "correlog_unifrac_0_long.pacbio",
  "correlog_unifrac_0_short.ion",
  "correlog_unifrac_0_short.pacbio",
  "correlog_unifrac_1_long.pacbio",
  "correlog_unifrac_1_short.ion",
  "correlog_unifrac_1_short.pacbio",
  cache = cache
)
bind_rows(as.data.frame(correlog_bray_0_long.pacbio$mantel.res) %>%
            mutate(tdist = "same", seq_run = "long", metric = "Bray-Curtis"),
          as.data.frame(correlog_bray_1_long.pacbio$mantel.res) %>%
            mutate(tdist = "diff", seq_run = "long", metric = "Bray-Curtis"),
          as.data.frame(correlog_bray_0_short.pacbio$mantel.res) %>%
            mutate(tdist = "same", seq_run = "short", metric = "Bray-Curtis"),
          as.data.frame(correlog_bray_1_short.pacbio$mantel.res) %>%
            mutate(tdist = "diff", seq_run = "short", metric = "Bray-Curtis"),
          as.data.frame(correlog_unifrac_0_long.pacbio$mantel.res) %>%
            mutate(tdist = "same", seq_run = "long", metric = "UNIFRAC"),
          as.data.frame(correlog_unifrac_1_long.pacbio$mantel.res) %>%
            mutate(tdist = "diff", seq_run = "long", metric = "UNIFRAC"),
          as.data.frame(correlog_unifrac_0_short.pacbio$mantel.res) %>%
            mutate(tdist = "same", seq_run = "short", metric = "UNIFRAC"),
          as.data.frame(correlog_unifrac_1_short.pacbio$mantel.res) %>%
            mutate(tdist = "diff", seq_run = "short", metric = "UNIFRAC")) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Spatial Distance correlation") +
  scale_color_discrete(name = "Year") +
  xlab("distance (m)") +
  facet_grid(seq_run ~ metric) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
```

# Progress since last time:

 - Sent samples for Illumina sequences (including repurification).
 - Run pipeline with Ion Torrent data (`r sum(asv_table$is_057 > 0)` ASVs, `r sum(otu$is_057 > 0)` 97% OTUs)
 - Map short reads (both Pac Bio and Ion Torrent) into long-read tree using EPA.
    - Resulting tree conflicts with taxonomic assignments (unlike long-read tree)
    - Probably because aligning the short reads isn't very accurate.
 - Write text for methods, outline for Introduction

### 2019-12-17 writing/tech_compare.Rmd
update tech comparison presentation and create ISP followup presentation

@@
 
+cache <- drake_cache(here(".light"))
 source(here("scripts/map_LSU.R"))
@@
 
-# Read number from PacBio vs. IonTorrent ASVs (ITS2)
+# Read number from long PacBio vs. IonTorrent (ITS2)
 ```{r}
-loadd(big_seq_table_ITS2)
+loadd(big_seq_table_ITS2, cache = cache)
 asv_table <- big_seq_table_ITS2 %>%
@@
                  into = c("seq.run", "plate", "well", "dir", "region"),
-                 regex = "([a-z]+_\\d+)_(\\d+)-([A-H]1?[0-9])([fr]?)-([:alnum:]+).+") %>%
-  dplyr::filter(seq.run %in% c("is_057", "pb_500")) %>%
+                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
   dplyr::group_by(seq.run, seq) %>%
@@
 
-asv_table %>% 
+otu <- read_tsv(here("data/clusters/ITS2.table")) %>%
+  column_to_rownames("#OTU ID") %>%
+  as.matrix() %>%
+  t() %>%
+  as_tibble(rownames = "sample") %>%
+  tidyr::extract(
+    col = "sample",
+    into = c("seq_run", "plate", "well"),
+    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
+  ) %>%
+  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
+  dplyr::filter(reads >= 1) %>%
+  dplyr::group_by(seq_run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq_run, value = reads, fill = 0)
+
+big_table <- bind_rows(
+  mutate(asv_table, type = "ASV"),
+  mutate(otu, type = "OTU")
+)
+big_table %>%
   filter(is_057 > 0, pb_500 > 0) %>%
@@
   scale_y_log10() +
-  xlab("Ion Torrent Reads") +
-  ylab("PacBio Reads") +
+  xlab("Ion Torrent Short Reads") +
+  ylab("PacBio Long Reads") +
   geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
-           data = asv_table %>% filter(is_057 == 0)) +
+           data = big_table %>% filter(is_057 == 0, pb_500 > 0)) +
   geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
-           data = asv_table %>% filter(pb_500 == 0)) +
+           data = big_table %>% filter(pb_500 == 0, is_057 > 0)) +
   stat_smooth(method = "loess", formula = y ~ x) +
@@
                     mean(log10(is_057[is_057 >= 1])), slope = 1),
-              linetype = 2L, alpha = 0.5)
+              linetype = 2L, alpha = 0.5) +
+  facet_wrap(~type)
+```
+
+# Read number from short PacBio vs. IonTorrent (ITS2)
+```{r}
+big_table %>% 
+  filter(is_057 > 0, pb_483 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_483)) +
+  geom_point(alpha = 0.2) +
+  scale_x_log10() +
+  scale_y_log10() +
+  xlab("Ion Torrent Short Reads") +
+  ylab("PacBio Short Reads") +
+  geom_rug(aes(y = pb_483, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(is_057 == 0, pb_483 > 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_483 == 0, is_057 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_483[pb_483 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  facet_wrap(~type)
 
@@
 
-# Read number from PacBio vs. IonTorrent ASV-OTUs (ITS2)
-
+# Read number from long vs. short PacBio ASVs (ITS2)
 ```{r}
-otu <- read_tsv(here("data/clusters/ITS2.table"),
-                col_types = "ciii") %>% select(is_057, pb_500) %>%
-  filter(rowSums(.) > 0)
-otu %>%
-  filter(is_057 > 0, pb_500 > 0) %>%
-  ggplot(aes(x = is_057, y = pb_500)) +
+big_table %>% 
+  filter(pb_483 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = pb_483, y = pb_500)) +
   geom_point(alpha = 0.2) +
@@
   scale_y_log10() +
-  xlab("Ion Torrent Reads") +
-  ylab("PacBio Reads") +
+  xlab("PacBio Short Reads") +
+  ylab("PacBio Long Reads") +
   geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
-           data = otu %>% filter(is_057 == 0)) +
-  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
-           data = otu %>% filter(pb_500 == 0)) +
+           data = big_table %>% filter(pb_483 == 0, pb_500 > 0)) +
+  geom_rug(aes(y = pb_483), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_500 == 0, pb_483 > 0)) +
   stat_smooth(method = "loess", formula = y ~ x) +
   geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
-                    mean(log10(is_057[is_057 >= 1])), slope = 1),
-              linetype = 2L, alpha = 0.5)
+                    mean(log10(pb_483[pb_483 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  facet_wrap(~type)
 ```

### 2019-12-19 writing/transect_paper.Rmd
start putting results figures in manuscript

@@
   - "R.bib"
-csl: fungal-ecology.csl
-panflute-filters: figurefilter
-panflute-path: '../../python'
+csl: molecular-ecology.csl
 output:
   bookdown::pdf_document2:
-    fig_crop: no
+    fig_crop: yes
     keep_tex: yes
@@
     keep_md: yes
+    fig_crop: yes
 ---
@@
 library("tikzDevice")
+library("ggplot2")
+library("tidyverse")
+library("drake")
+library("here")
 tryCatch(
@@
   )
-knitr::opts_chunk$set(dev="tikz",
+knitr::opts_chunk$set(
+  dev = "pdf",
   # dev.args = list(pointsize=32),
-                      echo=FALSE,
-                      results='asis',
-                      fig.show='asis',
-                      external=TRUE)
-options(tikzDefaultEngine = "luatex",
-        tikzLualatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
+  echo = FALSE,
+  # results = 'asis',
+  # fig.show = 'asis',
+  fig.width = 6,
+  fig.height = 4.5,
+  message = FALSE,
+  warning = FALSE,
+  external = TRUE
+)
+options(
+  tikzDefaultEngine = "luatex",
+  tikzXelatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
                           "\\usepackage[version=4]{mhchem}\n",
                           "\\usepackage{siunitx}\n",
-                                "\\DeclareSIUnit\\year{a}\n"))
+                          "\\DeclareSIUnit\\year{a}\n",
+                          "\\usepackage{fontspec}"),
+  tikzMetricsDictionary = here("writing/transect_paper-tikzDictionary")
+)
 
@@
 write_bib(c("base", "vegan", "permute"), file = "R.bib", tweak = TRUE)
+cache = drake_cache(here(".light"))
+light_plan <- readRDS(here("data/plan/drake_light.rds"))
+theme_set(theme_bw())
 ```
@@
     - Denoising (DADA2 [@callahan2017], Deblur [@amir2017], UNOISE2 [@edgar2016]) attempts to control for sequencing error while leaving intra-species variation.
-    - Phylogenetic community distance measures like UNIFRAC [@wong2016; look for earlier reference] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
+    - Phylogenetic community distance measures like UNIFRAC [@wong2016, look for earlier reference] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
 
@@
 
-Amplicons were purified using SPRI beads [Vesterinen et al 2016] and quantified fluorometrically using PicoGreen dsDNA (Quant-iT) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan) according to the manufacturer's protocol.
+Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
+
+Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically using PicoGreen dsDNA (Quant-iT) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan) according to the manufacturer's protocol.
 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
@@
 
-  - Two workflows for short reads:
-    - Mohammad's workflow in Pipecraft
-    - Brendan's workflow using Snakemake/Drake
-
 Circular consensus sequence (CCS) basecalls for SMRT sequences were made using `ccs` version 3.4 [@pacificbiosciences2019] using the default settings.
@@
 First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013].
-This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013; data not shown].
-Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017, @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
+This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].
+Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
 The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regionss:
@@
 
-Each of the extracted regions was independently filtered for length (table **X**) and a maximum of three expected errors.
+```{r region-length}
+datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
+regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
+regions %>%
+  select(Region = region, "Minimum length" = min_length, "Maximum length" = max_length) %>%
+  right_join(tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4"))) %>%
+  mutate_at("Region", str_replace_all, "_", ".") %>%
+  knitr::kable(
+    booktabs = TRUE,
+    caption = "Minimum and maximum allowed lengths for each extracted region."
+  )
+```
+
+Each of the extracted regions was independently filtered for length (Table \@ref(tab:region-length)) and a maximum of three expected errors.
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")`.
@@
 
-The tree was rooted outside the fungi by using the most abundant ASV which was confidently assigned to the Chlorophyta by all 6 applicable taxonomic assignment methods.
+The tree was rooted outside the Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
@@
 
-Ecological community dissimilarity matrices were calculated using methods "bray" (both long and short amplicons) and "unifrac" (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
-Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m,
+Ecological community dissimilarity matrices were calculated using the species based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
+Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m.
+Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
+
+Additionally, dissimilarities were plotted as a function of spatial distance up to 25 m and fit using an exponential variogram function using `gstat` version `r packageVersion("gstat")` using weighting .
+The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.
 
 # Results
+
+```{r read-counts}
+
+```
+
+- *Agaricus bisporus* from positive control was most abundant ASV in all sequencing runs
+  - only one ASV in both PacBio runs
+  - several ASVs in IonTorrent -- due to sequencing errors not controlled by DADA2 or intragenome variation only detected in deep sequencing?
+
+(ref:readcap) ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
+Dashed line represents a constant ratio of read numbers.
+The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading.
+
+```{r pbl-vs-its, fig.cap="Comparison between read number for PacBio RSII long reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+loadd(big_seq_table_ITS2, cache = cache)
+asv_table <- big_seq_table_ITS2 %>%
+  tibble::as_tibble(rownames = "filename") %>%
+  tidyr::gather(key = "seq", value = "reads", -1) %>%
+  dplyr::filter(reads >= 1) %>%
+  tidyr::extract(col = "filename",
+                 into = c("seq.run", "plate", "well", "dir", "region"),
+                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
+  dplyr::group_by(seq.run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq.run, value = reads, fill = 0)
+
+otu <- read_tsv(
+  here("data/clusters/ITS2.table"),
+  col_types = cols(
+    .default = col_integer(),
+    `#OTU ID` = col_character()
+  )
+) %>%
+  column_to_rownames("#OTU ID") %>%
+  as.matrix() %>%
+  t() %>%
+  as_tibble(rownames = "sample") %>%
+  tidyr::extract(
+    col = "sample",
+    into = c("seq_run", "plate", "well"),
+    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
+  ) %>%
+  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
+  dplyr::filter(reads >= 1) %>%
+  dplyr::group_by(seq_run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq_run, value = reads, fill = 0)
+
+big_table <- bind_rows(
+  mutate(asv_table, type = "ASV"),
+  mutate(otu, type = "OTU")
+) %>%
+  select(-seq)
+
+comparisons <- crossing(
+  x = factor(datasets$seq_run, ordered = TRUE),
+  y = factor(datasets$seq_run, ordered = TRUE),
+  type = c("ASV", "OTU")
+) %>%
+  filter(y > x) %>%
+  mutate_all(as.character) %>%
+  mutate(
+    cross = paste(x, y, sep = "-")
+  ) %>%
+  group_by(cross, type) %>%
+  group_modify(function(d, g) {
+    big_table %>%
+      filter(type == g$type) %>%
+      select(!!d$x, !!d$y) %>%
+      filter(.data[[d$x]] > 0 & .data[[d$y]] > 0) %>%
+      mutate_all(log) %>%
+      lm() %>%
+      broom::glance() %>%
+      bind_cols(d, .)
+    })
+
+compR2 <- function(x, y, type) {
+  format(filter(comparisons, x == !!x, y == !!y, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
+}
+
+big_table %>%
+  filter(is_057 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_500)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10(limits = c(1, 1e4)) +
+  coord_equal() +
+  xlab("Ion Torrent Short Reads") +
+  ylab("PacBio Long Reads") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(is_057 == 0, pb_500 > 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_500 == 0, is_057 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  facet_wrap(~type)
+```
+
+- Figure \@ref(fig:pbl-vs-its)
+  - ASVs/OTUs are correlated between Ion Torrent and PacBio; 
+  ASVs $R^2 = `r compR2("is_057", "pb_500", "ASV")`$; 
+  OTUs $R^2 = `r compR2("is_057", "pb_500", "OTU")`$
+  - "hook" due to sampling effects with rare species
+  - lots of ASVs/OTUs detected in only one tech
+
+- Figure \@ref(fig:pbs-vs-its)
+  - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
+  - $R^2 = `r compR2("is_057", "pb_483", "OTU")`$
+  - sequencing technologies are comparable, but ASVs not so much
+  - not such good match with ASVs -- suggests DADA2 doesn't perform equally
+
+- Figure \@ref(fig:pbl-vs-pbs)
+  - No "hook", because depth is similar
+  - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
+
+```{r pbs-vs-its, message=FALSE, fig.cap="Comparison between read number for PacBio RSII short reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+big_table %>% 
+  filter(is_057 > 0, pb_483 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_483)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10(limits = c(1, 1e4)) +
+  xlab("Ion Torrent Short Reads") +
+  ylab("PacBio Short Reads") +
+  coord_equal() +
+  geom_rug(aes(y = pb_483, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(is_057 == 0, pb_483 > 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_483 == 0, is_057 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_483[pb_483 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  facet_wrap(~type)
+
+```
+
+```{r pbl-vs-pbs, message=FALSE, fig.cap="Comparison between read number for PacBio RSII long and short amplicon reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+big_table %>% 
+  filter(pb_483 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = pb_483, y = pb_500)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_equal() +
+  xlab("PacBio Short Reads") +
+  ylab("PacBio Long Reads") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(pb_483 == 0, pb_500 > 0)) +
+  geom_rug(aes(y = pb_483), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_500 == 0, pb_483 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
+                    mean(log10(pb_483[pb_483 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  facet_wrap(~type)
+```
+
+(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (left two columnn) or weighted Unifrac (right column) dissimilarities and spatial distance, for short (left column) and long (right column) amplicons, and for the entire soil fungal community (top row) or only ECM taxa (bottom row).
+Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.
+
+```{r correlog-ecm, fig.cap="(ref:correlog-cap)"}
+
+light_plan %>%
+  filter(startsWith(target, "correlog")) %>%
+  select(target) %>%
+  separate(
+    target,
+    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
+    sep = "_",
+    remove = FALSE
+  ) %>%
+  mutate(
+    tdist = paste(tdist, "year"),
+    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
+    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
+  ) %>%
+  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
+  mutate_at("target", purrr::map, "mantel.res") %>%
+  mutate_at("target", purrr::map, as.data.frame) %>%
+  unnest(target) %>%
+  filter(complete.cases(.)) %>%
+  mutate(Significant = `Pr(corrected)` < 0.05) %>%
+  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
+  geom_point(size = 3) +
+  geom_line(alpha = 0.3, size = 1.5) +
+  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
+  # geom_line(aes(y = exp(predict(bc_fit)))) +
+  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
+  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
+  scale_color_discrete(name = "Time lag") +
+  xlab("Distance (m)") +
+  facet_grid(guild ~ metric + amplicon) +
+  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
+```
+
+(ref:variog) Spatial variogram fits for community dissimilarities.
+Each point represents one pair of samples taken at the same time.
+The blue line and shaded area represent a LOESS smooth of the data and its associated confidence interval, respectively.
+The red line is an exponential fit for the relationship between community dissimilarity and spatial distance.
+Horizontal dashed lines represent the values of the fit function at zero distance ("nugget") and infinite distance ("sill").
+The vertical dashed line is at the characteristic distance of the fit function ("range").
+
+```{r variog, warning = FALSE, message = FALSE}
+variog_data <- light_plan %>%
+  filter(startsWith(target, "variofit")) %>%
+  select(target, dist, dist_spatial) %>%
+  separate("target", c("step", "guild", "metric", "tech", "amplicon"), sep = "_", remove = FALSE) %>%
+  mutate(
+    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
+    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
+  ) %>%
+  mutate_at(c("target", "dist", "dist_spatial"), lapply, readd, character_only = TRUE, cache = cache) %>%
+  mutate_at(c("dist", "dist_spatial"), lapply, unclass) %>%
+  mutate(sills = purrr::map(target, ~ cumsum(.$psill)),
+         range = purrr::map_dbl(target, ~.$range[2]),
+         vgline = purrr::map(target, gstat::variogramLine, maxdist = 25, n = 100))
+
+variog_points <- unnest(variog_data, c("dist", "dist_spatial")) %>%
+  filter(dist_spatial <= 20)
+variog_empir <- variog_points %>%
+  group_by(guild, metric, tech, amplicon, dist_spatial) %>%
+  summarize_at("dist", mean)
+variog_sills <- unnest(variog_data, "sills")
+variog_fitline <- unnest(select(variog_data,-"dist"), "vgline")
+
+variog_points %>%
+  ggplot(aes(dist_spatial, dist)) +
+  geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
+  geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
+  geom_point(shape = 1, alpha = 0.2) +
+  geom_line(data = variog_empir, color = "blue", size = 1.5) +
+  geom_line(data = variog_fitline,
+            aes(dist, gamma), inherit.aes = FALSE, color = "red", size = 0.5) +
+  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
+  ylab("Community dissimilarity") +
+  facet_grid(guild ~ metric + amplicon)
+
+```
 

### 2019-12-20 writing/transect_paper.Rmd
Initial figures for results section, sent to Martin & Anna before Christmas.

--- 2019-12-05 writing/transect_paper.Rmd
@@
   - "R.bib"
-csl: fungal-ecology.csl
-panflute-filters: figurefilter
-panflute-path: '../../python'
+csl: molecular-ecology.csl
 output:
+  bookdown::word_document2: 
+    keep_md: yes
   bookdown::pdf_document2:
-    fig_crop: no
+    fig_crop: yes
     keep_tex: yes
@@
       in_header: preamble.tex
-  bookdown::word_document2: 
-    keep_md: yes
 ---
@@
 library("tikzDevice")
+library("ggplot2")
+library("magrittr")
+library("tidyverse")
+library("drake")
+library("here")
+library("phyloseq")
 tryCatch(
@@
 }
-write_bib(c("base", "vegan", "permute"), file = "R.bib", tweak = TRUE)
+write_bib(c("base", "phyloseq", "gstat"), file = "R.bib", tweak = TRUE)
+cache = drake_cache(here(".light"))
+light_plan <- readRDS(here("data/plan/drake_light.rds"))
+theme_set(theme_bw())
 ```
@@
     - can build tree directly
-    - short reads from same location can be mapped to tree.
+    - short reads from same location can potentially be mapped to tree. (but we didn't find that this works)
     - Sequencing depth tradeoff [@Kennedy2018]
@@
 
-Short amplicon ASVs without a perfect match to one of the long amplicon ASVs were added to the long amplicon alignment using the `--add` option of MAFFT version 7.453 [@katoh2012] with a 10-mer based guide tree and two iterations.
+*Short amplicon ASVs without a perfect match to one of the long amplicon ASVs were added to the long amplicon alignment using the `--add` option of MAFFT version 7.453 [@katoh2012] with a 10-mer based guide tree and two iterations.
 The short ASVs were then placed on the long amplicon tree using EPA-NG version 0.3.6 [@barbera2019].
@@
 The resulting tree was used as a guide tree to construct a final ML tree in RAxML, using the GTR+GAMMA model, in order to resolve relationships between different sequences assigned to the same branch.
-No bootstrapping was performed on the final tree.
+No bootstrapping was performed on the final tree.* -- this didn't work.
 
-The tree was rooted outside the fungi by using the most abundant ASV which was confidently assigned to the Chlorophyta by all 6 applicable taxonomic assignment methods.
+The tree was rooted outside the Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
@@
 ASVs falling outside this clade were excluded from further analysis.
+
+*Taxonomic assignment of ASVs which were represented in the long amplicon set was refined using the phylogenetic tree.
 
@@
 
-Ecological community dissimilarity matrices were calculated using methods "bray" (both long and short amplicons) and "unifrac" (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
-Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m,
+*Ecological community dissimilarity matrices were calculated using the species based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
+Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m.
+Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.*
+
+*Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation.
+, and fit using a joint exponential spatio-temporal model using the `fit.StVariogram` function from `gstat` version `r packageVersion("gstat")` [@R-gstat].
+The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.*
 
 # Results
+
+## Read counts
+```{r read-counts}
+
+```
+
+- How many reads passed quality control
+- how many became ASVs
+- how many ASVs in different regions
+
+```{r agaricus-check}
+agaricus <- readd(taxon_table, cache = cache) %>%
+  filter(rank == "genus", taxon == "Agaricus", n_diff == 1) %$%
+  unique(label)
+agaricus_reads <- readd(proto_physeq, cache = cache) %>%
+  phyloseq::subset_samples(sample_type == "Pos") %>%
+  prune_taxa(taxa = agaricus) %>%
+  prune_taxa(taxa = colSums(otu_table(.)) > 0) %>%
+  merge_samples("seq_run") %>%
+  otu_table()
+
+agaricus_order <- readd(proto_physeq, cache = cache) %>%
+  merge_samples("seq_run") %>%
+  otu_table() %>%
+  apply(MARGIN = 1, order, decreasing = TRUE) %>%
+  apply(MARGIN = 2, match, x = 1)
+```
+
+## Results from control
+
+- *Agaricus bisporus* from positive control was most abundant ASV in Ion and long PacBio runs, 5th highest in short PacBio
+  - only one ASV in both PacBio runs
+  - four ASVs in IonTorrent -- due to sequencing errors not controlled by DADA2 or intragenome variation only detected in deep sequencing?
+
+(ref:readcap) ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
+Dashed line represents a constant ratio of read numbers.
+The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading.
+$R^2$ value displayed is for log-transformed non-zero read numbers.
+
+## Comparison of read counts from different technologies
+
+```{r pbl-vs-its, fig.cap="Comparison between read number for PacBio RSII long reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+loadd(big_seq_table_ITS2, cache = cache)
+asv_table <- big_seq_table_ITS2 %>%
+  tibble::as_tibble(rownames = "filename") %>%
+  tidyr::gather(key = "seq", value = "reads", -1) %>%
+  dplyr::filter(reads >= 1) %>%
+  tidyr::extract(col = "filename",
+                 into = c("seq.run", "plate", "well", "dir", "region"),
+                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
+  dplyr::group_by(seq.run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq.run, value = reads, fill = 0)
+
+otu <- read_tsv(
+  here("data/clusters/ITS2.table"),
+  col_types = cols(
+    .default = col_integer(),
+    `#OTU ID` = col_character()
+  )
+) %>%
+  column_to_rownames("#OTU ID") %>%
+  as.matrix() %>%
+  t() %>%
+  as_tibble(rownames = "sample") %>%
+  tidyr::extract(
+    col = "sample",
+    into = c("seq_run", "plate", "well"),
+    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
+  ) %>%
+  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
+  dplyr::filter(reads >= 1) %>%
+  dplyr::group_by(seq_run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq_run, value = reads, fill = 0)
+
+big_table <- bind_rows(
+  mutate(asv_table, type = "ASV"),
+  mutate(otu, type = "OTU")
+) %>%
+  select(-seq)
+
+comparisons <- crossing(
+  x = factor(datasets$seq_run, ordered = TRUE),
+  y = factor(datasets$seq_run, ordered = TRUE),
+  type = c("ASV", "OTU")
+) %>%
+  filter(y > x) %>%
+  mutate_all(as.character) %>%
+  mutate(
+    cross = paste(x, y, sep = "-")
+  ) %>%
+  group_by(cross, type) %>%
+  group_modify(function(d, g) {
+    big_table %>%
+      filter(type == g$type) %>%
+      select(!!d$x, !!d$y) %>%
+      filter(.data[[d$x]] > 0 & .data[[d$y]] > 0) %>%
+      mutate_all(log) %>%
+      lm() %>%
+      broom::glance() %>%
+      bind_cols(d, .)
+    }) %>%
+  mutate(
+    label = paste("R^2 ==", formatC(r.squared, format = "f", digits = 2))
+  )
+
+compR2 <- function(x, y, type) {
+  format(filter(comparisons, x == !!x, y == !!y, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
+}
+
+big_table %>%
+  filter(is_057 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_500)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10(limits = c(1, 1e4)) +
+  coord_equal() +
+  xlab("Ion Torrent Short Reads") +
+  ylab("PacBio Long Reads") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(is_057 == 0, pb_500 > 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_500 == 0, is_057 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  geom_text(
+    aes(label = label),
+    data = filter(comparisons, x == "is_057", y == "pb_500"),
+    x = 1,
+    y = 3.5,
+    color = "black",
+    hjust = 0,
+    parse = TRUE,
+    inherit.aes = FALSE
+  ) +
+  facet_wrap(~type)
+```
+
+- Figure \@ref(fig:pbl-vs-its)
+  - ASVs/OTUs are correlated between Ion Torrent and PacBio; 
+  ASVs $R^2 = `r compR2("is_057", "pb_500", "ASV")`$; 
+  OTUs $R^2 = `r compR2("is_057", "pb_500", "OTU")`$
+  - "hook" at lower left due to sampling effects with rare species
+  - lots of ASVs/OTUs detected in only one tech
+
+- Figure \@ref(fig:pbs-vs-its)
+  - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
+  - $R^2 = `r compR2("is_057", "pb_483", "OTU")`$
+  - sequencing technologies yield comparable results (given read depth) with sufficient clustering
+  - not such good match with ASVs -- suggests DADA2 doesn't perform equally on different datasets
+
+- Figure \@ref(fig:pbl-vs-pbs)
+  - No "hook", because depth is similar
+  - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
+
+```{r pbs-vs-its, message=FALSE, fig.cap="Comparison between read number for PacBio RSII short reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+big_table %>% 
+  filter(is_057 > 0, pb_483 > 0) %>%
+  ggplot(aes(x = is_057, y = pb_483)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10(limits = c(1, 1e4)) +
+  xlab("Ion Torrent Short Reads") +
+  ylab("PacBio Short Reads") +
+  coord_equal() +
+  geom_rug(aes(y = pb_483, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(is_057 == 0, pb_483 > 0)) +
+  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_483 == 0, is_057 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_483[pb_483 >= 1])) -
+                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  geom_text(
+    aes(label = label),
+    data = filter(comparisons, x == "is_057", y == "pb_483"),
+    x = 1,
+    y = 3.5,
+    color = "black",
+    hjust = 0,
+    parse = TRUE,
+    inherit.aes = FALSE
+  ) +
+  facet_wrap(~type)
+
+```
+
+```{r pbl-vs-pbs, message=FALSE, fig.cap="Comparison between read number for PacBio RSII long and short amplicon reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+big_table %>% 
+  filter(pb_483 > 0, pb_500 > 0) %>%
+  ggplot(aes(x = pb_483, y = pb_500)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_equal() +
+  xlab("PacBio Short Reads") +
+  ylab("PacBio Long Reads") +
+  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(pb_483 == 0, pb_500 > 0)) +
+  geom_rug(aes(y = pb_483), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(pb_500 == 0, pb_483 > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
+                    mean(log10(pb_483[pb_483 >= 1])), slope = 1),
+              linetype = 2L, alpha = 0.5) +
+  geom_text(
+    aes(label = label),
+    data = filter(comparisons, x == "pb_483", y == "pb_500"),
+    x = 1,
+    y = 3.5,
+    color = "black",
+    hjust = 0,
+    parse = TRUE,
+    inherit.aes = FALSE
+  ) +
+  facet_wrap(~type)
+```
+
+## Distance decay
+
+(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (top two rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top row) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).
+Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.
+
+```{r correlog, fig.cap="(ref:correlog-cap)"}
+
+light_plan %>%
+  filter(startsWith(target, "correlog")) %>%
+  select(target) %>%
+  separate(
+    target,
+    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
+    sep = "_",
+    remove = FALSE
+  ) %>%
+  mutate(
+    tdist = paste(tdist, "year"),
+    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
+    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
+  ) %>%
+  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
+  mutate_at("target", purrr::map, "mantel.res") %>%
+  mutate_at("target", purrr::map, as.data.frame) %>%
+  unnest(target) %>%
+  filter(complete.cases(.)) %>%
+  mutate(Significant = `Pr(corrected)` < 0.05) %>%
+  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
+  geom_point(size = 3) +
+  geom_line(alpha = 0.3, size = 1.5) +
+  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
+  # geom_line(aes(y = exp(predict(bc_fit)))) +
+  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
+  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
+  scale_color_discrete(name = "Time lag") +
+  xlab("Distance (m)") +
+  facet_grid(metric + amplicon ~ guild, scales = "fixed") +
+  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
+```
+
+Figure \@ref(fig:correlog):
+
+  - Long amplicons found longer autocorrelation of Bray-Curtis for all fungi, but not ECM.  Is this due to some particular group?
+  - Long amplicon also detects some autocorrelation even after 1 year, absent in other tests.
+  - Unifrac has less significant autocorrelation.
+
+(ref:variog) Spatial variogram fits for community dissimilarities.
+Each point represents one pair of samples taken at the same time.
+Thick blue and red lines represent the empirical variogram (mean of community dissimilarities at each distance).
+Thin lines is an exponential fit for the relationship between community dissimilarity and spatial distance.
+Horizontal dashed lines represent the values of the fit function at zero distance ("nugget") and infinite distance ("sill").
+The vertical dashed line is at the characteristic distance of the fit function ("range").
+
+```{r variog, fig.cap = "(ref:variog)", include = FALSE}
+variog_data <- light_plan %>%
+  filter(step == "variofit") %>%
+  select(variofit, variog, guild, metric, tech, amplicon) %>%
+  mutate_at(c("guild", "metric", "tech", "amplicon"), str_replace_all, "\"", "") %>%
+  mutate(
+    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
+    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
+  ) %>%
+  mutate_at(c("variofit", "variog"), lapply, readd, character_only = TRUE, cache = cache) %>%
+  mutate(sills = purrr::map(variofit, ~ cumsum(.$psill)),
+         range = purrr::map_dbl(variofit, ~.$range[2]),
+         vgline = purrr::map(variofit, gstat::variogramLine, maxdist = 25, n = 100))
+
+variog_points <- unnest(variog_data, c("variog")) %>%
+  filter(dist <= 25)
+variog_empir <- variog_points %>%
+  group_by(guild, metric, tech, amplicon, dist) %>%
+  summarize_at("gamma", mean)
+variog_sills <- unnest(variog_data, "sills")
+variog_fitline <- unnest(variog_data, "vgline")
+
+variog_points %>%
+  ggplot(aes(dist, gamma)) +
+  geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
+  geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
+  geom_point(shape = 1, alpha = 0.2) +
+  geom_line(data = variog_empir, color = "blue", size = 1.5) +
+  geom_line(data = variog_fitline,
+            aes(dist, gamma), inherit.aes = FALSE, color = "red", size = 0.5) +
+  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
+  ylab("Community dissimilarity") +
+  facet_grid(guild ~ metric + amplicon)
+
+```
+
+(ref:spatio) Spatiotemporal variogram fits for community dissimilarities.
+The thick lines represent the empirical variogram (mean of community distances at each spatial distance) for samples taken at the same time (red) or separated by one year (blue).
+Thin lines are an exponential fit to the variograms.
+
+## Spatio-temporal variogram
+```{r spatio-temporal, fig.cap = "(ref:spatio)"}
+variog_data <- light_plan %>%
+  filter(step == "variofitST") %>%
+  select(variofitST, variogST, guild, metric, tech, amplicon) %>%
+  mutate_at(c("guild", "metric", "tech", "amplicon"), str_replace_all, "\"", "") %>%
+  mutate(
+    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
+    amplicon = factor(
+      amplicon,
+      levels = c("short", "long"),
+      labels = c("short amplicon", "long amplicon")
+    )
+  ) %>%
+  mutate_at(
+    c("variogST", "variofitST"),
+    lapply,
+    readd,
+    character_only = TRUE,
+    cache = cache)
+
+variog_points <- unnest(variog_data, c("variogST")) %>%
+  filter(dist <= 25) %>%
+  mutate_at("timelag", factor)
+
+variog_empir <- variog_points %>%
+  group_by(guild, metric, tech, amplicon, dist, timelag) %>%
+  summarize_at("gamma", mean)
+
+variog_confint <- variog_points %>%
+  group_by(guild, metric, tech, amplicon, dist, timelag) %>%
+  summarize(
+    lower = quantile(gamma, 0.025),
+    upper = quantile(gamma, 0.975)
+  )
+
+variog_fit <- variog_data %>%
+  mutate_at("variofitST", lapply, gstat:::vgmMetric, expand_grid(spacelag = seq(1, 25, 0.5), timelag = 0:1)) %>%
+  unnest(variofitST) %>%
+  rename(dist = spacelag) %>%
+  mutate_at("timelag", factor)
+
+variog_empir %>%
+  ggplot(aes(dist, gamma, color = timelag)) +
+  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
+  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
+  # geom_jitter(shape = 46, alpha = 0.5) +
+  geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(size = 1.5, alpha = 0.5) +
+  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
+  geom_line(data = filter(variog_fit, timelag == 1), size = 0.5, color = "darkcyan", alpha = 0.8) +
+  geom_line(data = filter(variog_fit, timelag == 0), size = 0.5, color = "red1", alpha = 0.8) +
+  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
+  scale_color_discrete() +
+  scale_y_continuous(name = "Community dissimilarity") +
+  facet_grid(metric + amplicon ~ guild, scales = "free")
+
+```
+
+(ref:variofit) Parameters for variogram fits shown in figure \@ref(fig:spatio-temporal).
+Range: characteristic range of exponential function; Anisotropy: effective spatial distance represented by one year; R/A: Range divided by anisotropy; the characteristic range of the exponential function in the time domain.
+
+```{r variofit}
+variog_data %>%
+  mutate_at("variofitST", lapply, purrr::compose(as.data.frame, as.list, gstat::extractPar)) %>%
+  unnest(variofitST) %>%
+  arrange(metric, amplicon, guild) %>%
+  select(Guild = guild, Amplicon = amplicon, Metric = metric,
+         Range = range, Anisotropy = anis) %>%
+  mutate("R/A" = Range/Anisotropy,
+         "R/A" = ifelse(`R/A` > 100, NA, `R/A`)) %>%
+  mutate_at(c("Range", "Anisotropy", "R/A"), formatC, digits = 1,
+            format = "f") %>%
+  mutate_at("Range", paste, "m") %>%
+  mutate_at("Anisotropy", paste, "m/a") %>%
+  mutate_at("R/A", paste, "a") %>%
+  mutate_at("R/A", str_replace, "NA a", "-") %>%
+  mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
+  kable(booktabs = TRUE, caption = "(ref:variofit)")
+```
+
+Figure \@ref(fig:spatio-temporal) and Table \@ref(tab:variofit):
+
+ - Fit looks good for long Pacbio w/Bray-Curtis.
+ - Unifrac shows unconvincing distance structure. (No surprise that the model doesn't fit well)
+ - Distance structure for short amplicon is really confusing.
+Why are the samples 25 m apart so similar?
+(Even after 1 year?) Could the samples be mixed up?
+Or is this just a statistical artifact (very few sample pairs have that separation).
+ - ECM has shorter range and less anisotropy for all metrics
+   - shorter range = more spatially variable
+   - less anisotropy = less temporally variable
 

### 2019-12-25 writing/transect_paper.Rmd
improve figures

@@
 output:
-  bookdown::word_document2: 
-    keep_md: yes
   bookdown::pdf_document2:
@@
       in_header: preamble.tex
+  bookdown::word_document2: 
+    keep_md: yes
 ---
@@
 ```{r region-length}
-datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
+datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc") %>%
+
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
@@
 
-(ref:readcap) ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
+(ref:readcap) Comparison between read numbers for different sequencing strategies, by ASV (left) and 97% OTU (right).
+ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
 Dashed line represents a constant ratio of read numbers.
@@
 
-```{r pbl-vs-its, fig.cap="Comparison between read number for PacBio RSII long reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
+```{r pb-vs-it, fig.cap="(ref:readcap)", fig.height=8}
 loadd(big_seq_table_ITS2, cache = cache)
@@
 
-comparisons <- crossing(
-  x = factor(datasets$seq_run, ordered = TRUE),
-  y = factor(datasets$seq_run, ordered = TRUE),
+multi_table <- crossing(
+  x_var = factor(datasets$seq_run, ordered = TRUE),
+  y_var = factor(datasets$seq_run, ordered = TRUE),
   type = c("ASV", "OTU")
 ) %>%
-  filter(y > x) %>%
+  filter(x_var < y_var) %>%
   mutate_all(as.character) %>%
-  mutate(
-    cross = paste(x, y, sep = "-")
-  ) %>%
-  group_by(cross, type) %>%
-  group_modify(function(d, g) {
+  group_by(x_var, y_var, type) %>%
+  group_modify(
+    function(d, g) {
       big_table %>%
         filter(type == g$type) %>%
-      select(!!d$x, !!d$y) %>%
-      filter(.data[[d$x]] > 0 & .data[[d$y]] > 0) %>%
-      mutate_all(log) %>%
-      lm() %>%
-      broom::glance() %>%
-      bind_cols(d, .)
-    }) %>%
+        select(x = !!g$x_var, y = !!g$y_var) %>%
+        filter(x > 0 | y > 0)
+    }
+  )
+
+comparisons <- multi_table %>%
+  group_modify(
+    function(d, g) {
+      filter_all(d, ~. > 0) %>%
+        mutate_all(log10) %>%
+      {bind_cols(
+      lm(y ~ x, .) %>%
+        broom::glance(.),
+      lm(y ~ offset(x), .) %>%
+        broom::tidy(.) %>%
+        select(term, estimate) %>%
+        pivot_wider(names_from = term, values_from = estimate)
+      )}
+    }
+  ) %>%
   mutate(
@@
   )
+
+# comparisons <- crossing(
+#   x_var = factor(datasets$seq_run, ordered = TRUE),
+#   y_var = factor(datasets$seq_run, ordered = TRUE),
+#   type = c("ASV", "OTU")
+# ) %>%
+#   filter(y_var < x_var) %>%
+#   mutate_all(as.character) %>%
+#   group_by(x_var, y_var, type) %>%
+#   group_modify(function(d, g) {
+#     big_table %>%
+#       filter(type == g$type) %>%
+#       select(x = !!g$x_var, y = !!g$y_var) %>%
+#       filter(.data[[g$x_var]] > 0 & .data[[g$y_var]] > 0) %>%
+#       mutate_all(log10) %>%
+#       lm() %>%
+#       {bind_rows(
+#       broom::glance(.),
+#       broom::tidy(.) %>%
+#         select(term, estimate) %>%
+#         pivot_wider(names_from = term, values_from = estimate)
+#       )} %>%
+#       bind_cols(d, .)
+#     }) %>%
+#   mutate(
+#     label = paste("R^2 ==", formatC(r.squared, format = "f", digits = 2))
+#   )
 
@@
 
-big_table %>%
-  filter(is_057 > 0, pb_500 > 0) %>%
-  ggplot(aes(x = is_057, y = pb_500)) +
+# p <- multi_table %>%
+#   filter(x > 0, y > 0, type == "ASV") %>%
+#   ggplot(aes(y = y, x = x)) +
+#   geom_point(alpha = 0.2, shape = 1) +
+#   scale_x_log10() +
+#   scale_y_log10(limits = c(1, 1e4)) +
+#   coord_equal() +
+#   xlab(NULL) +
+#   ylab(NULL) +
+#   geom_rug(aes(y = y), sides = "l", alpha = 0.2,
+#            data = multi_table %>% filter(x == 0, y > 0)) +
+#   geom_rug(aes(x = x), sides = "b", alpha = 0.2,
+#            data = multi_table %>% filter(y == 0, x > 0)) +
+#   facet_grid(y_var ~ x_var, switch = "both", space = "free") +
+#   stat_smooth(method = "loess", formula = y ~ x) +
+#   geom_abline(aes(intercept = `(Intercept)`, slope = 1),
+#               linetype = 2L, alpha = 0.5, data = comparisons) +
+#   geom_text(
+#     aes(label = label),
+#     data = comparisons,
+#     x = 1,
+#     y = 3.5,
+#     color = "black",
+#     hjust = 0,
+#     parse = TRUE,
+#     inherit.aes = FALSE
+#   ) +
+#     theme(strip.placement = "outside", strip.background = element_blank())
+# g <- ggplotGrob(p)
+# g$grobs[g$layout$name %in% c("panel-1-2")] <- NULL
+# g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2")),]
+# grid.newpage()
+# grid.draw(g)
+
+expand_seq_run <- function(labels) {
+  lapply(
+    labels,
+    )
+}
+
+read_comparison <- function(multi_table, comparisons, type) {
+  multi_table <- filter(multi_table, type == !!type)
+  comparisons <- filter(comparisons, type == !!type)
+  p <- multi_table %>%
+    filter(x > 0, y > 0) %>%
+    ggplot(aes(y = y, x = x)) +
     geom_point(alpha = 0.2, shape = 1) +
@@
     coord_equal() +
-  xlab("Ion Torrent Short Reads") +
-  ylab("PacBio Long Reads") +
-  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
-           data = big_table %>% filter(is_057 == 0, pb_500 > 0)) +
-  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
-           data = big_table %>% filter(pb_500 == 0, is_057 > 0)) +
+    xlab(NULL) +
+    ylab(NULL) +
+    geom_rug(aes(y = y), sides = "l", alpha = 0.2,
+             data = multi_table %>% filter(x == 0, y > 0)) +
+    geom_rug(aes(x = x), sides = "b", alpha = 0.2,
+             data = multi_table %>% filter(y == 0, x > 0)) +
+    facet_grid(y_var ~ x_var, switch = "both", space = "free", ) +
     stat_smooth(method = "loess", formula = y ~ x) +
-  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
-                    mean(log10(is_057[is_057 >= 1])), slope = 1),
+    geom_abline(aes(intercept = `(Intercept)`, slope = 1),
+                linetype = 2L, alpha = 0.5, data = comparisons) +
+    geom_text(
+      aes(label = label),
+      data = comparisons,
+      x = 0.5,
+      y = 3.5,
+      color = "black",
+      hjust = 0,
+      parse = TRUE,
+      inherit.aes = FALSE
+    ) +
+    theme(strip.placement = "outside", strip.background = element_blank())
+  g <- ggplotGrob(p)
+  g$grobs[g$layout$name %in% c("panel-1-2")] <- NULL
+  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2")),]
+  g
+}
+ggpubr::ggarrange(
+  read_comparison(multi_table, comparisons, type = "ASV"),
+  read_comparison(multi_table, comparisons, type = "OTU"),
+  ncol = 1,
+  labels = "auto",
+  label.x = 0.9
+)
+
+read_comparison <- function(big_table, comparisons, x, y, xlab, ylab) {
+  x_var <- enquo(x)
+  y_var <- enquo(y)
+  big_table %>%
+  filter(!!x_var > 0, !!y_var > 0) %>%
+  ggplot(aes(x = !!x_var, y = !!y_var)) +
+  geom_point(alpha = 0.2, shape = 1) +
+  scale_x_log10() +
+  scale_y_log10(limits = c(1, 1e4)) +
+  coord_equal() +
+  xlab(xlab) +
+  ylab(ylab) +
+  geom_rug(aes(y = !!y_var, x = 1), sides = "l", alpha = 0.2,
+           data = big_table %>% filter(!!x_var == 0, !!y_var > 0)) +
+  geom_rug(aes(y = !!x_var), sides = "b", alpha = 0.2,
+           data = big_table %>% filter(!!y_var == 0, !!x_var > 0)) +
+  stat_smooth(method = "loess", formula = y ~ x) +
+  geom_abline(aes(intercept = mean(log10((!!y_var)[(!!y_var) >= 1])) -
+                    mean(log10((!!x_var)[(!!x_var) >= 1])), slope = 1),
               linetype = 2L, alpha = 0.5) +
@@
     aes(label = label),
-    data = filter(comparisons, x == "is_057", y == "pb_500"),
+    data = filter(comparisons, .data$x == quo_name(x_var), .data$y == quo_name(y_var)),
     x = 1,
@@
   ) +
-  facet_wrap(~type)
+  facet_wrap(~type, strip.position = "right", scales = "free")
+}
+# ggpubr::ggarrange(
+#   ncol = 1,
+#   labels = "auto",
+#   read_comparison(
+#     big_table,comparisons,
+#     is_057, pb_500,
+#     "Ion Torrent Short Reads", "PacBio Long Reads"),
+#   read_comparison(
+#     big_table, comparisons,
+#     is_057, pb_483,
+#     "Ion Torrent Short Reads", "PacBio Short Reads"),
+#   read_comparison(
+#     big_table, comparisons,
+#     pb_483, pb_500,
+#     "PacBio Short Reads", "PacBio Long Reads")
+# )
 ```
 
-- Figure \@ref(fig:pbl-vs-its)
+- Figure \@ref(fig:pb-vs-it)a
   - ASVs/OTUs are correlated between Ion Torrent and PacBio; 
@@
 
-- Figure \@ref(fig:pbs-vs-its)
+- Figure \@ref(fig:pb-vs-it)b
   - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
@@
 
-- Figure \@ref(fig:pbl-vs-pbs)
+- Figure \@ref(fig:pb-vs-it)c
   - No "hook", because depth is similar
   - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
-
-```{r pbs-vs-its, message=FALSE, fig.cap="Comparison between read number for PacBio RSII short reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
-big_table %>% 
-  filter(is_057 > 0, pb_483 > 0) %>%
-  ggplot(aes(x = is_057, y = pb_483)) +
-  geom_point(alpha = 0.2, shape = 1) +
-  scale_x_log10() +
-  scale_y_log10(limits = c(1, 1e4)) +
-  xlab("Ion Torrent Short Reads") +
-  ylab("PacBio Short Reads") +
-  coord_equal() +
-  geom_rug(aes(y = pb_483, x = 1), sides = "l", alpha = 0.2,
-           data = big_table %>% filter(is_057 == 0, pb_483 > 0)) +
-  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
-           data = big_table %>% filter(pb_483 == 0, is_057 > 0)) +
-  stat_smooth(method = "loess", formula = y ~ x) +
-  geom_abline(aes(intercept = mean(log10(pb_483[pb_483 >= 1])) -
-                    mean(log10(is_057[is_057 >= 1])), slope = 1),
-              linetype = 2L, alpha = 0.5) +
-  geom_text(
-    aes(label = label),
-    data = filter(comparisons, x == "is_057", y == "pb_483"),
-    x = 1,
-    y = 3.5,
-    color = "black",
-    hjust = 0,
-    parse = TRUE,
-    inherit.aes = FALSE
-  ) +
-  facet_wrap(~type)
-
-```
-
-```{r pbl-vs-pbs, message=FALSE, fig.cap="Comparison between read number for PacBio RSII long and short amplicon reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
-big_table %>% 
-  filter(pb_483 > 0, pb_500 > 0) %>%
-  ggplot(aes(x = pb_483, y = pb_500)) +
-  geom_point(alpha = 0.2, shape = 1) +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_equal() +
-  xlab("PacBio Short Reads") +
-  ylab("PacBio Long Reads") +
-  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
-           data = big_table %>% filter(pb_483 == 0, pb_500 > 0)) +
-  geom_rug(aes(y = pb_483), sides = "b", alpha = 0.2,
-           data = big_table %>% filter(pb_500 == 0, pb_483 > 0)) +
-  stat_smooth(method = "loess", formula = y ~ x) +
-  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
-                    mean(log10(pb_483[pb_483 >= 1])), slope = 1),
-              linetype = 2L, alpha = 0.5) +
-  geom_text(
-    aes(label = label),
-    data = filter(comparisons, x == "pb_483", y == "pb_500"),
-    x = 1,
-    y = 3.5,
-    color = "black",
-    hjust = 0,
-    parse = TRUE,
-    inherit.aes = FALSE
-  ) +
-  facet_wrap(~type)
-```
 
@@
 variog_empir <- variog_points %>%
+  # mutate(bin = cut(dist, c(1:13 - 0.5, 14.5, 16.5, 19.5, 24.5))) %>%
   group_by(guild, metric, tech, amplicon, dist, timelag) %>%
-  summarize_at("gamma", mean)
+  summarize(
+    gamma = mean(gamma),
+    # dist = mean(dist),
+    n = n()
+  )
 
@@
 variog_fit <- variog_data %>%
-  mutate_at("variofitST", lapply, gstat:::vgmMetric, expand_grid(spacelag = seq(1, 25, 0.5), timelag = 0:1)) %>%
+  mutate_at("variofitST", lapply,
+            gstat:::vgmMetric,
+            expand_grid(spacelag = seq(1, 25, 0.5), timelag = 0:1)
+            ) %>%
   unnest(variofitST) %>%
@@
 variog_empir %>%
-  ggplot(aes(dist, gamma, color = timelag)) +
+  ggplot(aes(dist, gamma, color = timelag, size = n)) +
   # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
@@
   # geom_jitter(shape = 46, alpha = 0.5) +
-  geom_line(size = 1.5, alpha = 0.5) +
-  geom_point(size = 1.5, alpha = 0.5) +
+  # geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(alpha = 0.5) +
   # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +

### 2020-01-13 writing/transect_paper.Rmd
Changes from Anna's first edit

@@
 Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25 soil samples were collected at intervals of 1 m in May 2015.
+At each site, 25 soil samples were along a linear transect at intervals of 1 m in May 2015.
 Additional samples were collected from one third of the sample locations (3-m spacing) one year later in June 2016.
@@
 
-## Soil Chemistry
+<!-- ## Soil Chemistry
 
   - Need to get methods from Nourou
-  - Or just not use it?
+  - or just not use it?
+  -->
 
@@
 Two different regions of the nuclear ribosomal DNA were amplified (figure \@ref(fig:rDNA)).
-The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) ribosomal RNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] (hereafter ITS4m) as the reverse primer.
+The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) ribosomal RNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] (hereafter ITS4m) as the reverse primer.
 The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the ribosomal small subunit (SSU) RNA, the full ITS region including the 5.8S rRNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable domains.
@@
 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
-Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGS) at SciLifeLab in Uppsala, Sweden.
+Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
+for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752 37 UPPSALA).
 Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
 
-Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGS.
+Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGC.
 
@@
 and IDTAXA [@murali2018].
-Each ASV thus was thus given up to nine taxonomic assignments (three region/reference combinations $\times$ three algorithms).
+Each ASV thus was thus given up to nine preliminary taxonomic assignments (three region/reference combinations $\times$ three algorithms).
+
+Sequences were assigned as ECM if
 
@@
 
-*Ecological community dissimilarity matrices were calculated using the species based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
+Ecological community dissimilarity matrices were calculated using the species based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
 Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m.
-Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.*
+Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
 
-*Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation.
+Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation.
 , and fit using a joint exponential spatio-temporal model using the `fit.StVariogram` function from `gstat` version `r packageVersion("gstat")` [@R-gstat].
-The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.*
+The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.
 
@@
 
+# Acknowledgements
+
+The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
+Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden
+
 # Bibliography {-}

### 2020-01-16 writing/transect_paper.Rmd
Address some of Anna's comments

@@
 
-## Turnover scale in ECM forests
-
-- @lilleskov2004 - autocorrelation only <2.6 m at most sites.  Based on Sanger sequencing root tips in temperate forests.
-
-- Meta analysis @Bahram2013
-  - Includes data from Benin (same site? ask Nourou/Leho) but this is unpublished
-  - conclude that there is greater range of distance decay in tropical forests; because of less host availability? older forests?  *Russulaceae*?
-  - 65 m in Benin
-
-- @pickles2012a
-  - T-RFLP of root tips, temperate forest
-  - Autocorrelation < 3.4 m
-
-- Data from African wooded savannas - @tedersoo2011 but only southern band
-  - Irregular sampling with no samples nearer than 8m
-  - based on sanger sequencing root tips
-  - Significant Moran's I only in <10m class
-
 ## Long reads vs short reads/phylogenetic methods
@@
 
+## Turnover scale in ECM forests
+
+*why is turnover scale interesting*
+
+- @lilleskov2004 - autocorrelation only <2.6 m at most sites.  Based on Sanger sequencing root tips in temperate forests.
+
+- Meta analysis @Bahram2013
+  - Includes data from Benin (same site? ask Nourou/Leho) but this is unpublished
+  - conclude that there is greater range of distance decay in tropical forests; because of less host availability? older forests?  *Russulaceae*?
+  - 65 m in Benin
+
+- @pickles2012a
+  - T-RFLP of root tips, temperate forest
+  - Autocorrelation < 3.4 m
+
+- Data from African wooded savannas - @tedersoo2011 but only southern band
+  - Irregular sampling with no samples nearer than 8m
+  - based on sanger sequencing root tips
+  - Significant Moran's I only in <10m class
+
+*statement of question*
+
 # Methods
@@
 
-Sampling was conducted at two sites approximately 30 km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064
+°; Gan: N 9.75678° W 2.31058 °) approximately 30 km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
 Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 50 mg total of soil was collected from two locations in the homogenized soil and placed in Xpedition lysis solution (Zymo Research) and lysed in the field using a handheld bead-beater (Terralyzer; Zymo Research).
-The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis.
+Approximately 50 mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0 mL microtube containing 750 mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
+<!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->
 
@@
 
-After field lysis, DNA was extracted using the Zymo Research Soil/Fecal Prep kit **check name**.
-DNA was quantified using PicoGreen *according to the manufacturer's protocol*?
+After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
+DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
 
-Two different regions of the nuclear ribosomal DNA were amplified (figure \@ref(fig:rDNA)).
-The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) ribosomal RNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] (hereafter ITS4m) as the reverse primer.
-The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the ribosomal small subunit (SSU) RNA, the full ITS region including the 5.8S rRNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable domains.
+Two different fragments of the nuclear rDNA were amplified (figure \@ref(fig:rDNA)).
+The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] (hereafter ITS4m) as the reverse primer.
+The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the ribosomal small subunit (SSU) RNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable domains (figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 
 The gITS7 primers for the short amplicon were indexed for multiplexing (**supp info**).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher) and 3--7 ng purified DNA in Dream *Taq* buffer.
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7 ng purified DNA in Dream *Taq* buffer.
 The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
-Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity (**ref**), which were pooled after amplification.
+Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity, which were pooled after amplification.
 
-```{tikz rDNA, fig.cap = "rDNA amplicons", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
+(ref:rDNA) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines), as well as primer sites used in this study (red and blue arrows) and the first three variable regions in LSU (grey).)
+
+```{tikz rDNA, fig.cap = "(ref:rDNA)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
 
@@
 
-Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically using PicoGreen dsDNA (Quant-iT) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan) according to the manufacturer's protocol.
-100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
+Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
+An aliquot of 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
 Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
@@
 
+We attempted to denoise both long and short amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019, but only a very small fraction of the long amplicon reads were mapped to amplicon sequence variants (ASVs) at the end of this process.
+We believe that this was due to `DADA2`'s requirement that the seed sequence of each ASV not be a singleton read.
+For the combination of long reads, relatively high diversity, and relatively low sequencing depth in this study, this requirement was not met except by the most abundant sequences.
+We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
+
 Raw reads were divided into regions/domains by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
-First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013].
-This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].
-Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
-The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regionss:
+First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
+No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
+<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
+The fragment of each read beginning with the 5.8S rRNA was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
+The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regions:
 ITS1, 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
-where the four "LSU" regions represent the conserved regions of the 28S rRNA flanking the "D" regions.
+where LSU1-4 represent the conserved regions of the 28S rRNA flanking the variable D1-3 regions.
 For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
-No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx.
-The process of extracting the regions is automated in the new R package `LSUx`, available on github, CRAN, and bioconda (**but not yet**).
+The process of extracting the regions is automated in the new R package `LSUx`, available on github (https://github.com/brendanf/LSUx).
 
 ```{r region-length}
-datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc") %>%
+datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
 
@@
 regions %>%
-  select(Region = region, "Minimum length" = min_length, "Maximum length" = max_length) %>%
-  right_join(tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4"))) %>%
+  select(Region = region, "Min. length" = min_length, "Max. length" = max_length) %>%
+  right_join(
+    tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4")),
+    by = "Region"
+    ) %>%
   mutate_at("Region", str_replace_all, "_", ".") %>%
@@
 Each of the extracted regions was independently filtered for length (Table \@ref(tab:region-length)) and a maximum of three expected errors.
-Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")`.
-The error model for DADA2 denoising was fit using the 5.8S RNA region for long amplicons, and using the entire read for short amplicons.
+Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
+The error model for `DADA2` denoising was fit using the 5.8S RNA region for long amplicons, and using the entire read for short amplicons.
 Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
-Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
+Chimeras within each region were removed using `removeBimeraDenovoTable` from `DADA2`.
 
 For each ITS2 ASV from the long amplicon data set, the denoised sequences for the other regions corresponding to the same sequencing reads were concatenated to form a set of full-length reads.
-For reads which were not assigned a denoised sequence for every region, the raw read for the region was used instead.
-Because ITS2 is one of the most variable regions in the amplified regions, reads with identical ITS2 regions are expected have extremely similar sequences in the other regions, unless they are chimeric.
-Thus, the concatenated regions for each ASV were aligned in R using the DECIPHER package [@wright2015], and outlier sequences (chimeras) were removed from each alignment using the odseq package [@jehl2015].
+For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
+Because ITS2 is one of the most variable regions in the amplified regions, reads with identical ITS2 regions are expected have highly similar sequences in the other regions, unless the amplicon was chimeric.
+Thus, the concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015], and outlier sequences (interpreted as likely chimeras) were removed from each alignment using the `odseq` package [@jehl2015].
 The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence.
-Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
+Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed *(how many)*.
+The count and sample distribution of reads assigned to each ASV were calculated in order to form a sample x ASV community matrix.
 A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
-The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available on github (https://github.com/brendanf/tzara), CRAN (not yet), and bioconda (not yet).
+The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available on github (https://github.com/brendanf/tzara).
 
@@
 
-Taxonomic assignment was performed to genus level on the ITS and LSU regions using Unite/Warcup and RDP, respectively, as taxonomic references.
+Taxonomic assignment was performed to genus level separately on the ITS and LSU regions using Unite/Warcup and RDP, respectively, as taxonomic references.
 For each region/reference combination, taxonomy was assigned using three algorithms:
@@
 and IDTAXA [@murali2018].
-Each ASV thus was thus given up to nine preliminary taxonomic assignments (three region/reference combinations $\times$ three algorithms).
+Each ASV was thus given up to nine preliminary taxonomic assignments (three region/reference combinations $\times$ three algorithms).
 
-Sequences were assigned as ECM if
+Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
+All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
+
+## Clustering
+
+For comparison with clustering-based OTU methods, ASVs were clustered at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
 
@@
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
-Inspection of the tree along with taxonomic assignments revealed that the genus *Tulasnella* (Basidiomycota), which is known to have accelerated evolution in the rDNA [@moncalvo2006], was nested within the Metazoa, so it was excluded from further analysis.
+Inspection of the tree along with taxonomic assignments revealed that two ASVs assigned to the genus *Tulasnella* (Basidiomycota), which is known to have accelerated evolution in the rDNA [@moncalvo2006], were nested within the Metazoa, so they were excluded.
 The kingdom Fungi was then identified as the minimal clade containing all remaining ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
-ASVs falling outside this clade were excluded from further analysis.
+ASVs falling outside this clade were not included in the fungal community analysis.
 
-*Taxonomic assignment of ASVs which were represented in the long amplicon set was refined using the phylogenetic tree.
+Taxonomic assignment of ASVs which were represented in the long amplicon dataset was refined using the phylogenetic tree.
+*Describe this more.*
 
@@
 
-Ecological community dissimilarity matrices were calculated using the species based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
+Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
 Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m.
@@
 
-Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation.
-, and fit using a joint exponential spatio-temporal model using the `fit.StVariogram` function from `gstat` version `r packageVersion("gstat")` [@R-gstat].
+Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation, and fit using a joint exponential spatio-temporal model using the `fit.StVariogram` function from `gstat` version `r packageVersion("gstat")` [@R-gstat].
 The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.

### 2020-01-16 writing/transect_paper.Rmd
add conserved regions to rDNA figure, improve discussion of LSUx, mention unused samples.

@@
 At each site, 25 soil samples were along a linear transect at intervals of 1 m in May 2015.
-Additional samples were collected from one third of the sample locations (3-m spacing) one year later in June 2016.
+One third of the sample locations (3-m spacing) were resampled one year later in June 2016.
 For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a knife blade.
@@
 <!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->
+
+An additional sample was collected at every sampling location in 2016 using an older batch of lysis buffer for preservation.
+Sequencing results for these samples were found not to be comparable with results from the newer batch, so these samples were not included in spatial analyses.
+However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
 
@@
 Two different fragments of the nuclear rDNA were amplified (figure \@ref(fig:rDNA)).
-The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] (hereafter ITS4m) as the reverse primer.
-The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the ribosomal small subunit (SSU) RNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable domains (figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
+The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the small subunit (SSU) rDNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 
@@
 
-(ref:rDNA) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines), as well as primer sites used in this study (red and blue arrows) and the first three variable regions in LSU (grey).)
+(ref:rDNA) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).
+D1-3 (grey) represent the first three variable regions in LSU [@michot1984], while LSU1-4 (black) represent the conserved regions.
+Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
 
@@
 
-Raw reads were divided into regions/domains by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
-First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
+Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
+First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
 No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
 <!--This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
-The fragment of each read beginning with the 5.8S rRNA was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
-The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regions:
-ITS1, 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
-where LSU1-4 represent the conserved regions of the 28S rRNA flanking the variable D1-3 regions.
+The fragment of each read beginning with the 5.8S rRNA was then aligned to a specialized CM using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
+The CM was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
+The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in figure \@ref(fig:rDNA),
+where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
 For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
-The process of extracting the regions is automated in the new R package `LSUx`, available on github (https://github.com/brendanf/LSUx).
+Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available on github (https://github.com/brendanf/LSUx).
 
@@
 
-Each of the extracted regions was independently filtered for length (Table \@ref(tab:region-length)) and a maximum of three expected errors.
+Each of the extracted regions was independently filtered for length (table \@ref(tab:region-length)) and a maximum of three expected errors.
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
@@
 (ref:variofit) Parameters for variogram fits shown in figure \@ref(fig:spatio-temporal).
-Range: characteristic range of exponential function; Anisotropy: effective spatial distance represented by one year; R/A: Range divided by anisotropy; the characteristic range of the exponential function in the time domain.
+Range: range at which exponential function is at 95% of its maximum value; Anisotropy: effective spatial distance represented by one year; R/A: Range divided by anisotropy; the characteristic range of the exponential function in the time domain.
 
@@
          Range = range, Anisotropy = anis) %>%
-  mutate("R/A" = Range/Anisotropy,
+  mutate(Range = 3*Range,
+         "R/A" = Range/Anisotropy,
          "R/A" = ifelse(`R/A` > 100, NA, `R/A`)) %>%
@@
 
-Figure \@ref(fig:spatio-temporal) and Table \@ref(tab:variofit):
+Figure \@ref(fig:spatio-temporal) and table \@ref(tab:variofit):
 

### 2020-01-17 writing/transect_paper.Rmd
add supplementary figures for amplicon lengths

@@
   - Brendan Furneaux^1^,
-    Roël Houdanon^2^,
     Mohammad Bahram^1^,
@@
 theme_set(theme_bw())
+
+datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
+
+regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
 ```
@@
 
-Two different fragments of the nuclear rDNA were amplified (figure \@ref(fig:rDNA)).
+Two different fragments of the nuclear rDNA were amplified (Figure \@ref(fig:rDNA)).
 The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
-The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the small subunit (SSU) rDNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the small subunit (SSU) rDNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (Figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 
@@
 The CM was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
-The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in figure \@ref(fig:rDNA),
+The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure \@ref(fig:rDNA),
 where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
@@
 
-```{r region-length}
-datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
-
-regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
-regions %>%
-  select(Region = region, "Min. length" = min_length, "Max. length" = max_length) %>%
-  right_join(
-    tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4")),
-    by = "Region"
-    ) %>%
-  mutate_at("Region", str_replace_all, "_", ".") %>%
-  knitr::kable(
-    booktabs = TRUE,
-    caption = "Minimum and maximum allowed lengths for each extracted region."
-  )
-```
-
-Each of the extracted regions was independently filtered for length (table \@ref(tab:region-length)) and a maximum of three expected errors.
+Each of the extracted regions was independently filtered for length (Table \@ref(tab:region-length)) and a maximum of three expected errors.
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
@@
   - only one ASV in both PacBio runs
-  - four ASVs in IonTorrent -- due to sequencing errors not controlled by DADA2 or intragenome variation only detected in deep sequencing?
+  - four ASVs in IonTorrent -- probably due to sequencing errors not controlled by DADA2 (most abundant ASV had `r agaricus_reads["is_057", "39d80c19"][[1]]` reads, others had `r glue::glue_collapse(c(unclass(agaricus_reads["is_057", -1])), sep = ", ", last = ", and ")`)
 
@@
 
-(ref:variofit) Parameters for variogram fits shown in figure \@ref(fig:spatio-temporal).
+(ref:variofit) Parameters for variogram fits shown in Figure \@ref(fig:spatio-temporal).
 Range: range at which exponential function is at 95% of its maximum value; Anisotropy: effective spatial distance represented by one year; R/A: Range divided by anisotropy; the characteristic range of the exponential function in the time domain.
@@
 
-Figure \@ref(fig:spatio-temporal) and table \@ref(tab:variofit):
+Figure \@ref(fig:spatio-temporal) and Table \@ref(tab:variofit):
 
@@
 <div id="refs"></div>
+
+\processdelayedfloats
+
+# Supplementary Information {-}
+
+\beginsupplement
+
+```{r reads-table}
+reads_table <- cache$list() %>%
+  purrr::keep(startsWith, "big_seq_table_") %>%
+  purrr::map(readd, character_only = TRUE, cache = cache) %>%
+  purrr::map(tibble::as_tibble, rownames = "sample") %>%
+  purrr::map_dfr(tidyr::pivot_longer, -"sample", names_to = "seq", values_to = "reads") %>%
+  dplyr::filter(reads > 0) %>%
+  tidyr::extract(
+    "sample",
+    c("seq_run", "plate", "well", "region"),
+    regex = "([pi][bs]_\\d{3})_(\\d{3})_([A-H]1?\\d)_(.+)"
+  ) %>%
+  # dplyr::mutate_at(
+  #   "region",
+  #   factor,
+  #   levels = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3",
+  #              "D3", "LSU4", "ITS", "LSU", "32S", "long")) %>%
+  dplyr::group_by(seq_run, region, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  dplyr::mutate(length = nchar(seq)) %>%
+  dplyr::ungroup()
+
+region_table <- reads_table %>%
+  dplyr::group_by(seq_run, region) %>%
+  dplyr::summarize(
+    ASVs = dplyr::n_distinct(seq),
+    length_min = min(length),
+    length_q1 = reldist::wtd.quantile(length, 0.25, weight = reads),
+    length_med = reldist::wtd.quantile(length, 0.5, weight = reads),
+    length_mean = weighted.mean(length, w = reads),
+    length_q3 = reldist::wtd.quantile(length, 0.75, weight = reads),
+    length_max = max(length),
+    reads = sum(reads)
+  )
+```
+
+(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  *a*) Length distribution.  *b*) Empirical cumulative distribution function.
+
+```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
+full_dens <- reads_table %>%
+  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
+  dplyr::filter(region %in% c("long", "short")) %>%
+  dplyr::group_by(seq_run) %>%
+  dplyr::mutate(
+    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
+      factor(levels = c("Short", "Long")),
+    reads = reads / sum(reads)
+  ) %>%
+  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  stat_density(bw = 0.5, geom = "line", position = "identity") +
+  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
+  scale_y_continuous(name = "Density") +
+  scale_x_continuous(name = NULL) +
+  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")
+
+full_ecdf <- reads_table %>%
+  dplyr::filter(region  %in% c("long", "short")) %>%
+  dplyr::group_by(seq_run) %>%
+  dplyr::mutate(reads = reads / sum(reads)) %>%
+  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  stat_ecdf(geom = "line") +
+  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
+  scale_y_continuous(name = "Cumulative read fraction") +
+  scale_x_continuous(name = "Length (bp)") +
+  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))
+
+ggpubr::ggarrange(full_dens, full_ecdf, ncol = 1, labels = "auto")
+```
+
+```{r region-length}
+regions %>%
+  select(Region = region, "Min. length" = min_length, "Max. length" = max_length) %>%
+  right_join(
+    tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4")),
+    by = "Region"
+    ) %>%
+  mutate_at("Region", str_replace_all, "_", ".") %>%
+  knitr::kable(
+    booktabs = TRUE,
+    caption = "Minimum and maximum allowed lengths for each extracted region."
+  )
+```
+
+(ref:region-asv) Plot of total number of ASVs recovered vs. fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons.
+
+```{r regions, fig.cap="(ref:region-asv)"}
+region_table %>%
+  dplyr::filter(seq_run == "pb_500") %>%
+  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
+  geom_point() +
+  ggrepel::geom_text_repel() +
+  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
+  xlab("Mapped reads")
+```
+
+(ref:region-length) Length distribution for each region extracted from long amplicons.  Filtering limits are shown as dashed vertical lines.
+
+```{r region-lengths, fig.cap = "(ref:region-length)", fig.height = 6}
+baseregions <- c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3",
+               "LSU4")
+
+region_limits <- dplyr::filter(regions, region %in% baseregions) %>%
+  dplyr::mutate_at(
+    "region",
+    factor,
+    levels = baseregions
+    ) %>%
+  tidyr::pivot_longer(
+    cols = c("min_length", "max_length"),
+    names_to = "limit",
+    names_pattern = "(min|max)_length",
+    values_to = "length"
+  )
+
+reads_table %>%
+  dplyr::filter(region %in% baseregions,
+         seq_run == "pb_500") %>%
+  dplyr::mutate_at(
+    "region",
+    factor,
+    levels = baseregions
+    ) %>%
+  ggplot(aes(length, weight = reads)) +
+  stat_density(bw = 0.5, color = "#8da0cb", geom = "line") +
+  geom_vline(data = region_limits,
+             aes(xintercept = length),
+             linetype = "dashed", alpha = 0.5) +
+  facet_wrap(~region,ncol = 1, strip.position = "right", scales = "free_y") +
+  scale_y_continuous(breaks = NULL, labels = NULL, name = "Read density") +
+  scale_x_continuous(limits = c(0, 500), name = "Length (bp)")
+```
+
+(ref:ITS2-length) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies.
+*a*) Length distribution.
+*b*) Empirical cumulative distribution function.
+
+```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
+its2_dens <- reads_table %>%
+  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
+  dplyr::filter(region == "ITS2") %>%
+  dplyr::group_by(seq_run) %>%
+  dplyr::mutate(
+    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
+      factor(levels = c("Short", "Long")),
+    reads = reads / sum(reads)
+  ) %>%
+  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  stat_density(bw = 0.5, geom = "line", position = "identity") +
+  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
+  scale_y_continuous(name = "Density") +
+  scale_x_continuous(name = NULL) +
+  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")
+
+its2_ecdf <- reads_table %>%
+  dplyr::filter(region == "ITS2") %>%
+  dplyr::group_by(seq_run) %>%
+  dplyr::mutate(reads = reads / sum(reads)) %>%
+  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  stat_ecdf(geom = "line") +
+  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
+  scale_y_continuous(name = "Cumulative read fraction") +
+  scale_x_continuous(name = "ITS2 Length (bp)") +
+  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))
+
+ggpubr::ggarrange(its2_dens, its2_ecdf, ncol = 1, labels = "auto")
+```

### 2020-01-20 writing/transect_paper.Rmd
ensure phylotax doesn't assign a taxon to an unidentified outgroup; improve methods discussion of LSUx, tzara and phylotax

@@
 datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
+regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
 
-regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
+reads_table <- cache$list() %>%
+  purrr::keep(startsWith, "big_seq_table_") %>%
+  purrr::map(readd, character_only = TRUE, cache = cache) %>%
+  purrr::map(tibble::as_tibble, rownames = "sample") %>%
+  purrr::map_dfr(tidyr::pivot_longer, -"sample", names_to = "seq", values_to = "reads") %>%
+  dplyr::filter(reads > 0) %>%
+  tidyr::extract(
+    "sample",
+    c("seq_run", "plate", "well", "region"),
+    regex = "([pi][bs]_\\d{3})_(\\d{3})_([A-H]1?\\d)_(.+)"
+  ) %>%
+  # dplyr::mutate_at(
+  #   "region",
+  #   factor,
+  #   levels = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3",
+  #              "D3", "LSU4", "ITS", "LSU", "32S", "long")) %>%
+  dplyr::group_by(seq_run, region, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  dplyr::mutate(length = nchar(seq)) %>%
+  dplyr::ungroup()
+
+region_table <- reads_table %>%
+  dplyr::group_by(seq_run, region) %>%
+  dplyr::summarize(
+    ASVs = dplyr::n_distinct(seq),
+    length_min = min(length),
+    length_q1 = reldist::wtd.quantile(length, 0.25, weight = reads),
+    length_med = reldist::wtd.quantile(length, 0.5, weight = reads),
+    length_mean = weighted.mean(length, w = reads),
+    length_q3 = reldist::wtd.quantile(length, 0.75, weight = reads),
+    length_max = max(length),
+    reads = sum(reads)
+  )
 ```
+
 # Introduction
@@
 
-We attempted to denoise both long and short amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019, but only a very small fraction of the long amplicon reads were mapped to amplicon sequence variants (ASVs) at the end of this process.
-We believe that this was due to `DADA2`'s requirement that the seed sequence of each ASV not be a singleton read.
-For the combination of long reads, relatively high diversity, and relatively low sequencing depth in this study, this requirement was not met except by the most abundant sequences.
+### Denoising
+
+We attempted to denoise both long and short amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019
+However, only `r region_table$ASVs[region_table$region=="long"]` amplicon sequence variants (ASVs), were obtained.
+We believe that this was due to `DADA2`'s requirement that the seed sequence of each ASV be represented by at least two error-free reads.
+If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
+Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
+For the combination of long reads (median *X* bp after trimming) and moderate error rate (*X*% mean ccs quality score) in this study, the expected number of reads to achieve two error-free reads is `r round(2/(0.995)^1550)`.
+In comparison, the equivalent requirement for the short amplicon (median *X* bp after trimming, mean ccs quality score *Y*%) is only `r round(2/(0.999)^350)`
+Given the high diversity relative to sequencing depth in this study (*X* ASVs from PacBio short amplicons, *Y* total long amplicon reads), this requirement was not met for the long amplicons except by the most abundant sequences.
 We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
@@
 <!--This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
-The fragment of each read beginning with the 5.8S rRNA was then aligned to a specialized CM using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
-The CM was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
+A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
+A CM was generated from the alignment using `cmbuild` from Infernal.
+The fragment of each read beginning with the 5.8S rRNA was then aligned to the CM using `cmalign` from Infernal.
 The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure \@ref(fig:rDNA),
 where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
-For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
+For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
 Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available on github (https://github.com/brendanf/LSUx).
@@
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
-The error model for `DADA2` denoising was fit using the 5.8S RNA region for long amplicons, and using the entire read for short amplicons.
+The error model for `DADA2` denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
 Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
+For PacBio libraries, `DADA2` was run with complete pooling and a band size of 16.
+For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the `DADA2` FAQ (https://benjjneb.github.io/dada2/faq.html).
 Chimeras within each region were removed using `removeBimeraDenovoTable` from `DADA2`.
@@
 For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
-Because ITS2 is one of the most variable regions in the amplified regions, reads with identical ITS2 regions are expected have highly similar sequences in the other regions, unless the amplicon was chimeric.
-Thus, the concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015], and outlier sequences (interpreted as likely chimeras) were removed from each alignment using the `odseq` package [@jehl2015].
+Because ITS2 is the most variable of the amplified regions (Supplementary Figure \@ref(fig:regions)), reads with identical ITS2 regions are expected have highly similar sequences in the other regions, unless the amplicon was chimeric.
+The concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015].
+Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the `odseq` package [@jehl2015], using the default threshold of 0.025.
 The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence.
@@
 
-Taxonomic assignment of ASVs which were represented in the long amplicon dataset was refined using the phylogenetic tree.
-*Describe this more.*
+Taxonomic assignments of ASVs which were represented in the long amplicon dataset were refined using the phylogenetic tree.
+A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
+A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
+The result of this process was to give a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade where which was consistently given an assignment, as well as clarify the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.
 
@@
 ## Read counts
+
 ```{r read-counts}
@@
 
-```{r reads-table}
-reads_table <- cache$list() %>%
-  purrr::keep(startsWith, "big_seq_table_") %>%
-  purrr::map(readd, character_only = TRUE, cache = cache) %>%
-  purrr::map(tibble::as_tibble, rownames = "sample") %>%
-  purrr::map_dfr(tidyr::pivot_longer, -"sample", names_to = "seq", values_to = "reads") %>%
-  dplyr::filter(reads > 0) %>%
-  tidyr::extract(
-    "sample",
-    c("seq_run", "plate", "well", "region"),
-    regex = "([pi][bs]_\\d{3})_(\\d{3})_([A-H]1?\\d)_(.+)"
-  ) %>%
-  # dplyr::mutate_at(
-  #   "region",
-  #   factor,
-  #   levels = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3",
-  #              "D3", "LSU4", "ITS", "LSU", "32S", "long")) %>%
-  dplyr::group_by(seq_run, region, seq) %>%
-  dplyr::summarize(reads = sum(reads)) %>%
-  dplyr::mutate(length = nchar(seq)) %>%
-  dplyr::ungroup()
-
-region_table <- reads_table %>%
-  dplyr::group_by(seq_run, region) %>%
-  dplyr::summarize(
-    ASVs = dplyr::n_distinct(seq),
-    length_min = min(length),
-    length_q1 = reldist::wtd.quantile(length, 0.25, weight = reads),
-    length_med = reldist::wtd.quantile(length, 0.5, weight = reads),
-    length_mean = weighted.mean(length, w = reads),
-    length_q3 = reldist::wtd.quantile(length, 0.75, weight = reads),
-    length_max = max(length),
-    reads = sum(reads)
-  )
-```
-
 (ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  *a*) Length distribution.  *b*) Empirical cumulative distribution function.
@@
 
-(ref:region-asv) Plot of total number of ASVs recovered vs. fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons.
+(ref:region-asv) Plot of total number of ASVs recovered vs.
+fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
+Some reggions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU.
+For this figure, subregions were concatenated prior to generating ASVs with `DADA2`.
 

### 2020-01-20 scripts/taxonomy.R

@@ -738,9 +738,19 @@ relabel_tree <- function(tree, old, new, chimeras = character(0)) {
 
 # If all members of the clade are assigned uniquely to one taxon, then assign
 # that taxon
-# Otherwise, if there exists one clade that is at least one of the possibilities
+# Otherwise, if there exists one taxon that is at least one of the possibilities
 # for all of the members, then assign that one.
 clade_taxon <- function(tree, tax, node, rank) {
+  # Check the direct children
+  # If one of them is completely unidentified, then we don't want to assign a
+  # taxon here, because there's no way to know if the unassigned child is in the
+  # group or not.
+  children <- phangorn::Children(tree, node)
+  for (child in children) {
+    tips <- tree$tip.label[phangorn::Descendants(tree, child, type = "tips")[[1]]]
+    taxa <- dplyr::filter(tax, label %in% tips, rank == !!rank)
+    if (nrow(taxa) == 0) return(NA_character_)
+  }
   tips <- tree$tip.label[phangorn::Descendants(tree, node, type = "tips")[[1]]]
   taxa <- dplyr::filter(tax, label %in% tips, rank == !!rank) %>%
     dplyr::group_by(label)

### 2020-01-20 writing/transect_paper.Rmd

@@
 
-## Clustering
-
-For comparison with clustering-based OTU methods, ASVs were clustered at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
-
-## Alignment and phylogenetic inference
+### Clustering
+
+For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
+
+### Alignment and phylogenetic inference
 
@@
 
-Taxonomic assignments of ASVs which were represented in the long amplicon dataset were refined using the phylogenetic tree.
+Taxonomic assignments of ASVs which were represented in the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure \@ref(fig:phylotax)).
 A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
@@
 ## Read counts
+
+Long (short) amplicon sequencing with PacBio yielded *X* (*Y*) circular consensus reads, of which *A* (*B*) were successfully demultiplexed.
+The 5.8S region was detected in *C* (*D*) reads.
+Ion Torrent sequencing yielded *E* demultiplexed reads, with the 5.8S region detected in *F*.
 
@@
 
+Figure \@ref/fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs
+
 - Figure \@ref(fig:pb-vs-it)a
@@
 
+(ref:phylotax) Hypothetical example of phylogenetic refinement of taxonomic assignments.
+In the example, a clade includes  seven OTUs (A-F), which have been identified by two methods as belonging to taxa "Tax1" or "Tax2", or are unidentified ("unk"), as shown in the tip labels.
+No taxon is assigned at node 1, because one of the child branches (A) is completely unassigned.
+No taxon is assigned at node 2, because the assignments at C and F are inconsistent.
+Node 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
+Node 4 (and thus also node 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
+
+```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
+extree <-
+  tidytree::as.treedata(ape::read.tree(text = "(A:1,((B:1,C:1):1,((E:1,F:1):1,D:1):1):1);")) %>%
+  tidytree::full_join(
+    tibble(label = LETTERS[1:6],
+           tax1 = c("unk", "Tax1", "Tax2", "Tax2", "unk", "unk"),
+           tax2 = c("unk", "Tax2", "Tax2", "Tax1", "unk", "Tax1")),
+    by = "label"
+  )
+ggtree::ggtree(extree) +
+  ggtree::geom_tiplab(x = 4, align = 2) +
+  xlim(0, 6) +
+  ggtree::geom_tiplab(
+    aes(color = tax1, label = tax1),
+    x = 4.7,
+    hjust = 0.5
+  ) +
+  ggtree::geom_tiplab(
+    aes(color = tax2, label = tax2),
+    x = 5.7,
+    hjust = 0.5
+  ) +
+  ggtree::geom_tiplab(x = 5.2, label = "/", hjust = 0.5) +
+  ggtree::geom_nodelab(aes(label = node - 6), geom = "label") +
+  ggtree::geom_hilight(node = 10, fill = "steelblue", extendto = 4) +
+  ggtree::geom_hilight(node = 9, fill = "tomato", extendto = 4) +
+  scale_color_manual(values = c(Tax1 = "steelblue", Tax2 = "tomato", unk = "grey50")) +
+  scale_y_reverse()
+```
+
 (ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  *a*) Length distribution.  *b*) Empirical cumulative distribution function.

### 2020-01-22 writing/transect_paper.Rmd
add table of read and ASV counts through pipeline

@@
 
-datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
+datasets <- read_csv(here("config/datasets.csv"), col_types = "ccccicccccicc")
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
@@
   )
+loadd(big_seq_table_ITS2, cache = cache)
+asv_table <- big_seq_table_ITS2 %>%
+  tibble::as_tibble(rownames = "filename") %>%
+  tidyr::gather(key = "seq", value = "reads", -1) %>%
+  dplyr::filter(reads >= 1) %>%
+  tidyr::extract(col = "filename",
+                 into = c("seq.run", "plate", "well", "dir", "region"),
+                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
+  dplyr::group_by(seq.run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq.run, value = reads, fill = 0)
+
+otu <- read_tsv(
+  here("data/clusters/ITS2.table"),
+  col_types = cols(
+    .default = col_integer(),
+    `#OTU ID` = col_character()
+  )
+) %>%
+  column_to_rownames("#OTU ID") %>%
+  as.matrix() %>%
+  t() %>%
+  as_tibble(rownames = "sample") %>%
+  tidyr::extract(
+    col = "sample",
+    into = c("seq_run", "plate", "well"),
+    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
+  ) %>%
+  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
+  dplyr::filter(reads >= 1) %>%
+  dplyr::group_by(seq_run, seq) %>%
+  dplyr::summarize(reads = sum(reads)) %>%
+  tidyr::spread(key = seq_run, value = reads, fill = 0)
 ```
@@
 
-We attempted to denoise both long and short amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019
-However, only `r region_table$ASVs[region_table$region=="long"]` amplicon sequence variants (ASVs), were obtained.
-We believe that this was due to `DADA2`'s requirement that the seed sequence of each ASV be represented by at least two error-free reads.
+```{r length-qual}
+
+parse_qstat <- function(d) {
+  mutate_at(d, "file", basename) %>%
+  tidyr::extract(col = "file", into = c("seq_run", "plate", "well"),
+                 regex = "([pi][bs]_\\d{3})_(\\d{3})-?([A-H]1?\\d)?[rf]?[.].*")
+}
+
+demuxlength <- readd(qstats_length, cache = cache) %>%
+  parse_qstat() %>%
+  filter(is.na(step), !is.na(well)) %>%
+  group_by(seq_run) %>%
+  summarize(length = reldist::wtd.quantile(length, weight = nreads)) %>%
+  deframe()
+
+demuxqual <- readd(qstats_erate, cache = cache) %>%
+  parse_qstat() %>%
+  filter(is.na(step), !is.na(well)) %>%
+  group_by(seq_run) %>%
+  summarize(qual = round(weighted.mean(erate, w = nreads, na.rm = TRUE), 4)) %>%
+  deframe()
+
+readcounts <- 
+  readd(qstats_n, cache = cache) %>%
+  parse_qstat()
+
+rawcounts <- readcounts %>%
+  filter(is.na(step), is.na(well)) %>%
+  group_by(seq_run) %>%
+  summarize(nreads =  prettyNum(sum(nreads), big.mark = " ")) %>%
+  deframe()
+
+demuxcounts <- readcounts %>%
+  filter(is.na(step), !is.na(well)) %>%
+  group_by(seq_run) %>%
+  summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
+  deframe()
+
+regioncounts <- readcounts %>%
+  filter(step == "lsux") %>%
+  group_by(seq_run) %>%
+  summarize(nreads = prettyNum(
+    sum(nreads) %/% if (seq_run[1] == "pb_500") 14 else 4,
+    big.mark = " ")
+  ) %>%
+  deframe()
+```
+
+We attempted to denoise both long and short amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019.
+However, only `r region_table$ASVs[region_table$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r round(100*region_table$reads[region_table$region == "long"]/as.integer(gsub(" ", "", demuxcounts["pb_500"])), 1)`% of the trimmed reads.
+We believe that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
+The `DADA2` algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
 If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
 Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median *X* bp after trimming) and moderate error rate (*X*% mean ccs quality score) in this study, the expected number of reads to achieve two error-free reads is `r round(2/(0.995)^1550)`.
-In comparison, the equivalent requirement for the short amplicon (median *X* bp after trimming, mean ccs quality score *Y*%) is only `r round(2/(0.999)^350)`
-Given the high diversity relative to sequencing depth in this study (*X* ASVs from PacBio short amplicons, *Y* total long amplicon reads), this requirement was not met for the long amplicons except by the most abundant sequences.
+For the combination of long reads (median $L=`r demuxlength['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(demuxqual['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - demuxqual['pb_500'])^demuxlength['pb_500'])), big.mark = " ")`.
+Given the high diversity relative to sequencing depth in this study (`r region_table$ASVs[region_table$seq_run == "pb_483" & region_table$region == "short"]` ASVs based on PacBio short amplicons, `r demuxcounts["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
+In comparison, the equivalent requirement for the short amplicon ($L=`r demuxlength['pb_483']` \text{ bp}$, $\epsilon=`r demuxqual['pb_483']`$) is only `r round(2/(1 - demuxqual['pb_483'])^demuxlength['pb_483'], 1)` reads.
 We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
@@
 In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
-Although the latter system is not formally published, it seems to be in use for non-fungal eukaryotes in the Unite database.
-Additionally, it is a system with only purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as @adl2019.
+Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
+Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [ @adl2019], which utilizes hierarchical nameless ranks.
 FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).
@@
 
-An attempt was made to refine the long amplicon alignment using Pasta [@mirarab2014a], but the resulting maximum likelihood (ML) trees did not improve on the ML achieved using the original alignment from DECIPHER.
-Since ML is the optimization criterion used by Pasta, the original DECIPHER alignment was retained.
+<!--An attempt was made to refine the long amplicon alignment using Pasta [@mirarab2014a], but the resulting maximum likelihood (ML) trees did not improve on the ML achieved using the original alignment from DECIPHER.
+Since ML is the optimization criterion used by Pasta, the original DECIPHER alignment was retained.-->
 
@@
 
-Long (short) amplicon sequencing with PacBio yielded *X* (*Y*) circular consensus reads, of which *A* (*B*) were successfully demultiplexed.
-The 5.8S region was detected in *C* (*D*) reads.
-Ion Torrent sequencing yielded *E* demultiplexed reads, with the 5.8S region detected in *F*.
-
-```{r read-counts}
-
+Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
+The 5.8S region was detected in `r regioncounts["pb_500"]` (`r regioncounts["pb_483"]`) reads.
+Ion Torrent sequencing yielded `r demuxcounts["is_057"]` demultiplexed reads, with the 5.8S region detected in `r regioncounts["is_057"]`.
+
+(ref:readcounts) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages.
+Raw: raw reads as delivered by sequencing center;
+Trim: reads with primers and demultiplexing indices removed;
+LSUx: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
+ITS2: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
+short, ITS, LSU, long: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"song\" denote the full-length short and long amplicons, respectively.
+Short amplicons were mapped to full ITS, LSU, and long amplicon ASVs when
+
+```{r}
+bioinf_table <- bind_rows(
+  enframe(rawcounts, name = "seq_run", value = "reads") %>%
+    mutate(
+      reads = as.integer(gsub(" ", "", reads)),
+      step = "Raw"
+    ),
+  enframe(demuxcounts, name = "seq_run", value = "reads") %>%
+    mutate(
+      reads = as.integer(gsub(" ", "", reads)),
+      step = "Trim"
+    ),
+  enframe(regioncounts, name = "seq_run", value = "reads") %>%
+    mutate(
+      reads = as.integer(gsub(" ", "", reads)),
+      step = "LSUx"
+    ),
+  filter(region_table, region == "ITS2") %>%
+    select(seq_run, reads, step = region, ASVs),
+  pivot_longer(asv_table, -1, names_to = "seq_run", values_to = "reads") %>%
+    filter(reads > 0) %>%
+    mutate_at("seq", chartr, old = "T", new = "U") %>%
+    inner_join(
+      select(readd(allseqs, cache = cache), seq = ITS2, long, short, ITS, LSU) %>%
+        pivot_longer(-1, names_to = "region", values_to = "consensus") %>%
+        filter(!is.na(consensus)) %>%
+        select(-consensus) %>%
+        unique(),
+      by = "seq"
+    ) %>%
+    group_by(seq_run, region) %>%
+    summarize(reads = sum(reads, na.rm = TRUE), ASVs = n()) %>%
+    select(seq_run, reads, step = region, ASVs)
+) %>%
+  left_join(select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
+  mutate(
+    step = ifelse(tech == "Ion" & step == "Trim", "Raw", step),
+  ) %>%
+  select(tech, amplicon, step, reads, ASVs) %>%
+  pivot_longer(c("reads", "ASVs"), names_to = "type", values_to = "count") %>%
+  filter(!is.na(count)) %>%
+  mutate(
+    step = factor(step, levels = c("Raw", "Trim", "LSUx", "ITS2",
+                                   "short", "ITS", "LSU", "long")),
+    tech = factor(tech, levels = c("PacBio", "Ion", "Illumina")),
+    amplicon = factor(amplicon, levels = c("Long", "Short")),
+    type = factor(type, c("ASVs", "reads"))
+  ) %>%
+  arrange(tech, amplicon, type) %>%
+  pivot_wider(
+    id_cols = "step",
+    names_from = c("tech", "amplicon", "type"),
+    values_from = c("count")
+    ) %>%
+  arrange(step) %>%
+  column_to_rownames("step")
+
+options(knitr.kable.NA = '–')
+bioinf_header <- tibble(name = names(bioinf_table)) %>%
+  separate(name, into = c("tech", "amplicon", "type")) 
+bioinf_table %>%
+  set_names(gsub(".*_", "", names(.))) %>%
+  kable(
+    booktabs = TRUE,
+    caption = "(ref:readcounts)",
+    align = "rrrrrr",
+    format.args = list(big.mark = "\\,"),
+    linesep = c(rep("", 3), rep("\\addlinespace", 2), rep("", 4)),
+    escape = FALSE
+    ) %>%
+  kableExtra::add_header_above(
+    header = c(
+      " " = 1,
+      bioinf_header$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
+    )
+  ) %>%
+  kableExtra::add_header_above(
+    header = c(
+      " " = 1,
+      rle(bioinf_header$tech) %$% set_names(lengths, values)
+    )
+  )
 ```
@@
 ```{r pb-vs-it, fig.cap="(ref:readcap)", fig.height=8}
-loadd(big_seq_table_ITS2, cache = cache)
-asv_table <- big_seq_table_ITS2 %>%
-  tibble::as_tibble(rownames = "filename") %>%
-  tidyr::gather(key = "seq", value = "reads", -1) %>%
-  dplyr::filter(reads >= 1) %>%
-  tidyr::extract(col = "filename",
-                 into = c("seq.run", "plate", "well", "dir", "region"),
-                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
-  dplyr::group_by(seq.run, seq) %>%
-  dplyr::summarize(reads = sum(reads)) %>%
-  tidyr::spread(key = seq.run, value = reads, fill = 0)
-
-otu <- read_tsv(
-  here("data/clusters/ITS2.table"),
-  col_types = cols(
-    .default = col_integer(),
-    `#OTU ID` = col_character()
-  )
-) %>%
-  column_to_rownames("#OTU ID") %>%
-  as.matrix() %>%
-  t() %>%
-  as_tibble(rownames = "sample") %>%
-  tidyr::extract(
-    col = "sample",
-    into = c("seq_run", "plate", "well"),
-    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
-  ) %>%
-  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
-  dplyr::filter(reads >= 1) %>%
-  dplyr::group_by(seq_run, seq) %>%
-  dplyr::summarize(reads = sum(reads)) %>%
-  tidyr::spread(key = seq_run, value = reads, fill = 0)
-
 big_table <- bind_rows(
@@
 ) %>%
-  filter(x_var < y_var) %>%
+  filter(
+    x_var < y_var,
+    x_var %in% names(big_table),
+    y_var %in% names(big_table)
+  ) %>%
   mutate_all(as.character) %>%
@@
   mutate(Range = 3*Range,
-         "R/A" = Range/Anisotropy,
+         "Time Range" = Range/Anisotropy,
          "R/A" = ifelse(`R/A` > 100, NA, `R/A`)) %>%
-  mutate_at(c("Range", "Anisotropy", "R/A"), formatC, digits = 1,
+  mutate_at(c("Range", "Time Range"), formatC, digits = 1,
             format = "f") %>%
@@
 extree <-
-  tidytree::as.treedata(ape::read.tree(text = "(A:1,((B:1,C:1):1,((E:1,F:1):1,D:1):1):1);")) %>%
+  treeio::as.treedata(ape::read.tree(text = "(A:1,((B:1,C:1):1,((E:1,F:1):1,D:1):1):1);")) %>%
   tidytree::full_join(

### 2020-01-22 writing/transect_paper.Rmd
change import format for datasets to match new columns

@@
 
-datasets <- read_csv(here("config/datasets.csv"), col_types = "ccccicccccicc")
+datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccicc")
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
@@
 
-## Read counts
+## Sequencing
 
@@
 (ref:readcounts) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages.
-Raw: raw reads as delivered by sequencing center;
-Trim: reads with primers and demultiplexing indices removed;
-LSUx: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
-ITS2: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
-short, ITS, LSU, long: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"song\" denote the full-length short and long amplicons, respectively.
+*Raw*: raw reads as delivered by sequencing center;
+*Trim*: reads with primers and demultiplexing indices removed;
+*LSUx*: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
+*ITS2*: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
+*short*, *ITS*, *LSU*, *long*: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.
 Short amplicons were mapped to full ITS, LSU, and long amplicon ASVs when
@@
 (ref:variofit) Parameters for variogram fits shown in Figure \@ref(fig:spatio-temporal).
-Range: range at which exponential function is at 95% of its maximum value; Anisotropy: effective spatial distance represented by one year; R/A: Range divided by anisotropy; the characteristic range of the exponential function in the time domain.
+Range: range at which exponential function is at 95% of its maximum value; Anisotropy: effective spatial distance represented by one year; the characteristic range of the exponential function in the time domain.
 
@@
          "Time Range" = Range/Anisotropy,
-         "R/A" = ifelse(`R/A` > 100, NA, `R/A`)) %>%
+         "Time Range" = ifelse(`Time Range` > 100, NA, `Time Range`)) %>%
   mutate_at(c("Range", "Time Range"), formatC, digits = 1,

### 2020-01-22 scripts/drake_plan_heavy.R

@@ -44,7 +44,7 @@ setup_log("drakeplan")
 
 # load the dataset and region definitions
 flog.info("Loading datasets file.")
-datasets <- read_csv(config$dataset, col_types = "cccicccccicc")
+datasets <- read_csv(config$dataset, col_types = "cccccccicccccicc")
 flog.info("Loading regions file.\n")
 regions <- read_csv(config$regions, col_types = "cccciiiiic") %>%
   mutate_at("range", replace_na, "") %>%

### 2020-01-24 writing/transect_paper.Rmd
partial results text, Venn diagrams, taxonomy bar charts

@@
   tidyr::spread(key = seq_run, value = reads, fill = 0)
+
+k_or_M <- function(x, ...) {
+  case_when(
+    abs(x) > 1e6 ~ paste(format(x / 1e6, ...), "M"),
+    abs(x) > 1e3 ~ paste(format(x / 1e3, ...), "k"),
+    TRUE ~ paste(format(x, ...), " ")
+  )
+}
+
+percent <- function(x, ...) {
+  paste0(format(as.numeric(x) * 100), "%")
+}
+
+text_list <- function(x, ...) {
+  x = format(x, ...)
+  glue::glue_collapse(x, sep = ", ", last = ", and ")
+}
 ```
@@
 An aliquot of 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
-Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
+Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
 for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752 37 UPPSALA).
@@
 
-## Sequencing
-
+The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).'
 Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
@@
 Ion Torrent sequencing yielded `r demuxcounts["is_057"]` demultiplexed reads, with the 5.8S region detected in `r regioncounts["is_057"]`.
+
+### Region extraction
+
+The 5.8S region was located in `r text_list(bioin)
 
@@
 
-```{r}
+```{r bioinfo}
 bioinf_table <- bind_rows(
@@
     summarize(reads = sum(reads, na.rm = TRUE), ASVs = n()) %>%
-    select(seq_run, reads, step = region, ASVs)
+    select(seq_run, reads, step = region, ASVs),
+  # pivot_longer(otu, -1, names_to = "seq_run", values_to = "reads") %>%
+  #   filter(reads > 0) %>%
+  #   mutate_at("seq", chartr, old = "T", new = "U") %>%
+  #   inner_join(
+  #     select(readd(allseqs, cache = cache), seq = ITS2, long, short, ITS, LSU) %>%
+  #       pivot_longer(-1, names_to = "region", values_to = "consensus") %>%
+  #       filter(!is.na(consensus)) %>%
+  #       select(-consensus) %>%
+  #       unique(),
+  #     by = "seq"
+  #   ) %>%
+  #   group_by(seq_run, region) %>%
+  #   summarize(reads = sum(reads, na.rm = TRUE), OTUs = n()) %>%
+  #   select(seq_run, reads, step = region, OTUs)
 ) %>%
@@
                                    "short", "ITS", "LSU", "long")),
-    tech = factor(tech, levels = c("PacBio", "Ion", "Illumina")),
+    tech = factor(tech, levels = c("PacBio", "Ion Torrent", "Illumina")),
     amplicon = factor(amplicon, levels = c("Long", "Short")),
@@
 bioinf_header <- tibble(name = names(bioinf_table)) %>%
-  separate(name, into = c("tech", "amplicon", "type")) 
+  separate(name, into = c("tech", "amplicon", "type"), sep = "_") %>%
+  mutate(tech = paste(tech, plyr::mapvalues(tech, datasets$tech, datasets$machine, FALSE)))
 bioinf_table %>%
@@
     align = "rrrrrr",
-    format.args = list(big.mark = "\\,"),
+    format.args = list(big.mark = " "),
     linesep = c(rep("", 3), rep("\\addlinespace", 2), rep("", 4)),
@@
 ```
-
-- How many reads passed quality control
-- how many became ASVs
-- how many ASVs in different regions
 
@@
 
-## Results from control
-
-- *Agaricus bisporus* from positive control was most abundant ASV in Ion and long PacBio runs, 5th highest in short PacBio
-  - only one ASV in both PacBio runs
-  - four ASVs in IonTorrent -- probably due to sequencing errors not controlled by DADA2 (most abundant ASV had `r agaricus_reads["is_057", "39d80c19"][[1]]` reads, others had `r glue::glue_collapse(c(unclass(agaricus_reads["is_057", -1])), sep = ", ", last = ", and ")`)
+### Reproducibility of read counts from different technologies
+
+(ref:venn-table-asv) Correspondences between ASVs found by different sequencing strategies.
+Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets , including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+
+```{r venn-table-asv}
+venn_ASV <-
+  asv_table %>%
+  select(seq, pb_500, pb_483, is_057) %>%
+  mutate_if(is.numeric, list(found = ~. > 0)) %>%
+  group_by_at(vars(ends_with("_found"))) %>%
+  summarize_if(is.numeric, list(ASVs = ~sum(. > 0), reads = sum)) %>%
+  ungroup() %>%
+  mutate_at(vars(ends_with("_found")), as.integer) %>%
+  tidyr::unite(col = "set", 1:3, sep = "") %>%
+  pivot_longer(
+    -1,
+    names_to = c("seq_run", "what"),
+    values_to = "value",
+    names_pattern = "([[:alpha:]]+[-._]\\d+)_(.+)"
+  ) %>%
+  pivot_wider(names_from = "what", values_from = "value") %>%
+  group_by(seq_run) %>%
+  mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
+  mutate_at(
+    vars(ends_with("_frac")),
+    formatC,
+    digits = 2,
+    format = "f",
+    drop0trailing = FALSE
+  ) %>%
+  mutate_at(
+    c("reads", "ASVs"),
+    format
+  ) %>%
+  mutate(
+    ASVs_frac = if_else(grepl(" +0", ASVs), "", ASVs_frac),
+    ASVs = if_else(grepl(" +0", ASVs), "", ASVs),
+    reads_frac = if_else(grepl(" +0", reads), "", reads_frac),
+    reads = if_else(grepl(" +0", reads), "", reads)
+  ) %>%
+  select(set, seq_run, ASVs, ASVs_frac, reads, reads_frac) %>% {
+    left_join(
+      group_by(., set) %>%
+        summarize(ASVs = as.integer(first(ASVs[ASVs != ""]))),
+
+      select(., set, seq_run, ASV_frac = ASVs_frac, reads, frac = reads_frac) %>%
+      pivot_wider(
+        names_from = "seq_run",
+        values_from = c("ASV_frac", "reads", "frac")
+      )
+    )
+  } %>%
+  mutate(n = str_count(set, "1")) %>%
+  arrange(n, desc(set)) %>%
+  select(set, ASVs, ends_with("pb_500"), ends_with("pb_483"), ends_with("is_057")) %>%
+  column_to_rownames("set")
+
+venn_ASV %>%
+  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
+  set_names(gsub("_", " ", names(.))) %>%
+  kable(
+    booktabs = TRUE,
+    caption = "(ref:venn-table-asv)",
+    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
+    align = ("rcrrcrrcrr"),
+    row.names = FALSE
+  ) %>%
+  kableExtra::add_header_above(
+    header = c(
+      " " = 1,
+      names(venn_ASV)[-(1)] %>% 
+        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
+        rle() %$%
+        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
+      )
+  ) %>%
+  kableExtra::add_header_above(
+    header = c(
+      " " = 1,
+       names(venn_ASV)[-(1)] %>% 
+        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
+        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
+        rle() %$%
+        set_names(lengths, values)
+      )
+  )
+```
+
+(ref:venn-table-otu) Correspondences between 97% OTUs found by different sequencing strategies.
+
+```{r venn-table-otu}
+venn_OTU <-
+  otu %>%
+  select(seq, pb_500, pb_483, is_057) %>%
+  mutate_if(is.numeric, list(found = ~. > 0)) %>%
+  group_by_at(vars(ends_with("_found"))) %>%
+  summarize_if(is.numeric, list(OTUs = ~sum(. > 0), reads = sum)) %>%
+  ungroup() %>%
+  mutate_at(vars(ends_with("_found")), as.integer) %>%
+  tidyr::unite(col = "set", 1:3, sep = "") %>%
+  pivot_longer(
+    -1,
+    names_to = c("seq_run", "what"),
+    values_to = "value",
+    names_pattern = "([[:alpha:]]+[-._]\\d+)_(.+)"
+  ) %>%
+  pivot_wider(names_from = "what", values_from = "value") %>%
+  group_by(seq_run) %>%
+  mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
+  mutate_at(
+    vars(ends_with("_frac")),
+    formatC,
+    digits = 2,
+    format = "f",
+    drop0trailing = FALSE
+  ) %>%
+  mutate_at(
+    c("reads", "OTUs"),
+    format
+  ) %>%
+  mutate(
+    OTUs_frac = if_else(grepl(" +0", OTUs), "", OTUs_frac),
+    OTUs = if_else(grepl(" +0", OTUs), "", OTUs),
+    reads_frac = if_else(grepl(" +0", reads), "", reads_frac),
+    reads = if_else(grepl(" +0", reads), "", reads)
+  ) %>%
+  select(set, seq_run, OTUs, OTUs_frac, reads, reads_frac) %>%  {
+    left_join(
+      group_by(., set) %>%
+        summarize(OTUs = as.integer(first(OTUs[OTUs != ""]))),
+
+      select(., set, seq_run, OTU_frac = OTUs_frac, reads, frac = reads_frac) %>%
+      pivot_wider(
+        names_from = "seq_run",
+        values_from = c("OTU_frac", "reads", "frac")
+      )
+    )
+  } %>%
+  mutate(n = str_count(set, "1")) %>%
+  arrange(n, desc(set)) %>%
+  select(set, OTUs, ends_with("pb_500"), ends_with("pb_483"), ends_with("is_057")) %>%
+  column_to_rownames("set")
+
+venn_OTU %>%
+  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
+  set_names(gsub("_", " ", names(.))) %>%
+  kable(
+    booktabs = TRUE,
+    caption = "(ref:venn-table-otu)",
+    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
+    align = ("rcrrcrrcrr"),
+    row.names = FALSE
+  ) %>%
+  kableExtra::add_header_above(
+    header = c(
+      " " = 1,
+      names(venn_OTU)[-1] %>% 
+        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
+        rle() %$%
+        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
+      )
+  ) %>%
+  kableExtra::add_header_above(
+    header = c(
+     " " = 1,
+      names(venn_OTU)[-1] %>% 
+        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
+        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
+        rle() %$%
+        set_names(lengths, values)
+      )
+
+  )
+```
+
+(ref:venn) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different sequencing strategies.
+In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.
+
+```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
+vennplot_ASV <- venn_ASV %>%
+  rownames_to_column("set") %>%
+  mutate(set = map_chr(set,
+                       ~ strsplit(., "")[[1]] %>%
+                     as.integer() %>%
+                     as.logical() %>%
+                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
+                     paste(collapse = "&"))) %>%
+  mutate(
+    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057,
+                       sep = "/") %>%
+      gsub("/+", "/", .) %>%
+      gsub("(^/|/$)", "", .),
+    label = paste(ASVs, reads_frac, sep = "\n")
+  )
+
+eulerr::venn(deframe(select(vennplot_ASV, set, ASVs))) %>%
+  plot(
+    legend = list(
+      labels = c("PacBio Long", "PacBio Short", "Ion Torrent")
+    ),
+    quantities = vennplot_ASV$label,
+    main = "ASV"
+  )
+
+vennplot_OTU <- venn_OTU %>%
+  rownames_to_column("set") %>%
+  mutate(n = str_count(set, "1")) %>%
+  arrange(n, desc(set)) %>%
+  mutate(set = map_chr(set,
+                       ~ strsplit(., "")[[1]] %>%
+                     as.integer() %>%
+                     as.logical() %>%
+                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
+                     paste(collapse = "&"))) %>%
+  mutate(
+    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057,
+                       sep = "/") %>%
+      gsub("/+", "/", .) %>%
+      gsub("(^/|/$)", "", .),
+    label = paste(OTUs, reads_frac, sep = "\n")
+  )
+
+eulerr::venn(deframe(select(vennplot_OTU, set, OTUs))) %>%
+  plot(
+    legend = FALSE,
+    labels = FALSE,
+    quantities = vennplot_OTU$label,
+    main = "OTU"
+  )
+
+```
 
@@
 
-## Comparison of read counts from different technologies
-
 ```{r pb-vs-it, fig.cap="(ref:readcap)", fig.height=8}
@@
 ) %>%
-  select(-seq)
+  select(-seq) %>%
+  select(type, everything())
+
+names(big_table)[-1] <-
+  tibble(seq_run = names(big_table)[-1]) %>%
+  left_join(datasets, by = "seq_run") %>%
+  glue::glue_data("{tech} {machine} - {amplicon}")
 
 multi_table <- crossing(
-  x_var = factor(datasets$seq_run, ordered = TRUE),
-  y_var = factor(datasets$seq_run, ordered = TRUE),
+  x_var = factor(glue::glue_data(datasets, "{tech} {machine} - {amplicon}"), ordered = TRUE),
+  y_var = factor(glue::glue_data(datasets, "{tech} {machine} - {amplicon}"), ordered = TRUE),
   type = c("ASV", "OTU")
@@
 compR2 <- function(x, y, type) {
-  format(filter(comparisons, x == !!x, y == !!y, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
+  x <- plyr::mapvalues(
+    x,
+    datasets$seq_run,
+    paste(
+      datasets$tech,
+      datasets$machine,
+      "-",
+      datasets$amplicon),
+    FALSE
+  )
+  y <- plyr::mapvalues(
+    y,
+    datasets$seq_run,
+    paste(
+      datasets$tech,
+      datasets$machine,
+      "-",
+      datasets$amplicon),
+    FALSE
+  )
+  format(filter(comparisons, x == x_var, y == y_var, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
 }
@@
 
-Figure \@ref/fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs
-
-- Figure \@ref(fig:pb-vs-it)a
-  - ASVs/OTUs are correlated between Ion Torrent and PacBio; 
-  ASVs $R^2 = `r compR2("is_057", "pb_500", "ASV")`$; 
-  OTUs $R^2 = `r compR2("is_057", "pb_500", "OTU")`$
-  - "hook" at lower left due to sampling effects with rare species
-  - lots of ASVs/OTUs detected in only one tech
-
-- Figure \@ref(fig:pb-vs-it)b
-  - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
-  - $R^2 = `r compR2("is_057", "pb_483", "OTU")`$
-  - sequencing technologies yield comparable results (given read depth) with sufficient clustering
-  - not such good match with ASVs -- suggests DADA2 doesn't perform equally on different datasets
-
-- Figure \@ref(fig:pb-vs-it)c
-  - No "hook", because depth is similar
-  - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
+The number of ASVs and OTUs shared between the different sequencing strategies are shown in Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in figure \@ref(fig:venn).
+ASVs shared between all datasets represented `r text_list(percent(venn_ASV[7, c("frac_pb_500", "frac_pb_483", "frac_is_057")]))` of the reads for the long and short PacBio datasets and Ion Torrent dataset, respectively.
+These fractions increased to `r text_list(percent(venn_OTU[7, c("frac_pb_500", "frac_pb_483", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+In particular, >99% and `r percent(sum(as.numeric(venn_OTU[c("011", "111"), "frac_is_057"])))` of reads in the PacBio and Ion Torrent short-amplicon datasets belonged to OTUs shared between the two datasets.
+In contrast, `r percent(venn_ASV["100", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(venn_OTU["100", "frac_pb_500"])` after OTU clustering.
+
+Figure \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
+In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = `r compR2("is_057", "pb_500", "ASV")` \text{ to } `r compR2("is_057", "pb_483", "OTU")`$).
+Comparisons including the Ion Torrent dataset all display a "hook" at low read counts, which is probably due to statistical sampling error when comparing read counts between datasets with very different sequencing depth.
+The strongest correlation was for the short amplicon PacBio and Ion Torrent datasets after OTU clustering ($R^2 = `r compR2("is_057", "pb_483", "OTU")`$), which were sequenced from the same amplicon library.
+
+## Taxonomic assignment
+
+  - what fraction assigned at each level (or representative ranks; phylum, (order), genus)
+  - what fraction are fungi
+  - comparison of different region/database and algorithm
+  - breakdown of groups
+  - fraction ECM
+
+```{r}
+taxon_reads <- 
+  readd(proto_physeq, cache = cache) %>%
+  phyloseq::merge_samples(group = "seq_run") %>%
+  phyloseq::otu_table() %>%
+  t() %>%
+  {.@.Data} %>%
+  as_tibble(rownames = "label") %>%
+  mutate_if(is.numeric, ~./sum(.)) %>%
+  pivot_longer(-1, names_to = "seq_run", values_to = "reads") %>%
+  filter(reads > 0) %>%
+  left_join(select(datasets, seq_run, amplicon), by = "seq_run") %>%
+  expand_grid(
+    tibble(
+      reference = c("unite", "warcup", "rdp_train"),
+      region = c("ITS", "ITS", "LSU")
+    ),
+      method = c("dada2", "idtaxa", "sintax")
+  ) %>%
+  filter(
+    ifelse(amplicon == "Short", region == "ITS", TRUE),
+    ifelse(amplicon == "Long", region %in% c("ITS", "LSU"), TRUE)
+  ) %>%
+  left_join(
+    readd(taxon_table, cache = cache) %>%
+      mutate(region = ifelse(region == "short", "ITS", region)) %>%
+      mutate_at("taxon", na_if, "NA") %>%
+      select(label, method, reference, rank, taxon, region) %>%
+      pivot_wider(
+        names_from = "rank",
+        values_from = "taxon"
+      ),
+    by = c("label", "reference", "region", "method")
+  ) %>%
+  mutate(
+    kingdom = ifelse(!is.na(phylum) & reference == "warcup", "Fungi", kingdom)
+  )
+```
+
+(ref:taxa-kingdom) Taxonomic assignment of ASVs to kingdom.
+N/A denotes ASVs which could not be identified to kingdom level.
+Height of each bar represents the fraction of reads represented by that bar.
+
+```{r taxa-kingdom, fig.cap = "(ref:taxa-kingdom)"}
+taxon_reads %>%
+  group_by(reference, method, seq_run, kingdom) %>%
+  summarize(reads = sum(reads), ASVs = n()) %>%
+  ggplot(aes(x = method, y = reads, fill = kingdom)) +
+  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
+  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+(ref:ASV-kingdom) Taxonomic assignment of ASVs to kingdom.
+N/A denotes ASVs which could not be identified to kingdom level.
+Height of each bar represents the fraction of total ASVs represented by that bar.
+
+```{r ASV-kingdom, fig.cap = "(ref:ASV-kingdom)"}
+taxon_reads %>%
+  group_by(reference, method, seq_run, kingdom) %>%
+  summarize(reads = sum(reads), ASVs = n()) %>%
+  ggplot(aes(x = method, y = ASVs, fill = kingdom)) +
+  geom_bar(position = "fill", stat = "identity", color = "white", size = 0.2) +
+  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+(ref:taxa-phylum) Taxonomic assignment of ASVs to fungal phylum.
+Phyla representing less than 1% of total reads are grouped as "other".
+N/A denotes ASVs which were identified as fungi, but could not be identified to phylum level.
+Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
+Height of each bar represents the fraction of reads represented by that bar.
+
+```{r taxa-phylum, fig.cap = "(ref:taxa-phylum)"}
+taxon_reads %>%
+  filter(kingdom == "Fungi") %>%
+  group_by(reference, method, phylum, seq_run) %>%
+  summarize(reads = sum(reads), ASVs = n()) %>%
+  ungroup() %>%
+  mutate(
+    phylum = ifelse(reads < 0.01 & !is.na(phylum), " other", phylum)
+  ) %>%
+  group_by(reference, method, phylum, seq_run) %>%
+  summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
+  # complete(reference, method, phylum, fill = list(reads = 0, ASVs = 0)) %>%
+  ggplot(aes(x = method, y = reads, fill = phylum)) +
+  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
+  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
+  theme(axis.text.x = element_text(angle = 90))
+
+```
+
+(ref:asco-class) Taxonomic assignment of ASVs to ascomycete order.
+Orders representing less than 1% of total reads are grouped as "other".
+N/A denotes ASVs which were identified as Ascomycota, but could not be identified to order level.
+Reads which were not identified to phylum, or which were identified as a non-Ascomycota phylum, are not included.
+Height of each bar represents the fraction of total reads represented by that bar.
+
+```{r asco-class, fig.cap = "(ref:asco-class)"}
+taxon_reads %>%
+  filter(phylum == "Ascomycota") %>%
+  group_by(reference, method, order, seq_run) %>%
+  summarize(reads = sum(reads), ASVs = n()) %>%
+  ungroup() %>%
+  mutate(
+    order = ifelse(reads < 0.01 & !is.na(order), " other", order)
+  ) %>%
+  group_by(reference, method, order, seq_run) %>%
+  summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
+  ggplot(aes(x = method, y = reads, fill = order)) +
+  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
+  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
+  theme(axis.text.x = element_text(angle = 90))
+
+```
+
+(ref:basid-family) Taxonomic assignment of ASVs to basidiomycete family.
+Families representing less than 1% of total reads are grouped as "other".
+N/A denotes ASVs which were identified as Basidiomycota, but could not be identified to family level.
+Reads which were not identified to phylum, or which were identified as a non-Basidiomycota phylum, are not included.
+Height of each bar represents the fraction of total reads represented by that bar.
+
+```{r basid-family, fig.cap = "(ref:basid-family)"}
+taxon_reads %>%
+  filter(phylum == "Basidiomycota") %>%
+  group_by(reference, method, family, seq_run) %>%
+  summarize(reads = sum(reads), ASVs = n()) %>%
+  ungroup() %>%
+  mutate(
+    family = ifelse(reads < 0.01 & !is.na(family), " other", family)
+  ) %>%
+  group_by(reference, method, family, seq_run) %>%
+  summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
+  ggplot(aes(x = method, y = reads, fill = family)) +
+  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
+  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
+  theme(axis.text.x = element_text(angle = 90))
+
+```
+
+### Positive control
+
+*Agaricus bisporus*, the positive control, was represented by a single ASV in both long- and short-amplicon PacBio datasets.
+It was the most abundant ASV in the long-amplicon dataset, and the fifth most abundant ASV in the short-amplicon dataset.
+The same ASV was also the most abundant ASV in the Ion Torrent dataset, but three additional ASVs identified as *Agaricus bisporus* were also present in that dataset.
+The abundances of these ASVs (`r glue::glue_collapse(c(unclass(agaricus_reads["is_057", ])), sep = ", ", last = ", and ")`) was inconsistent with both heterozygosity (which would tend toward equal read counts for two variants) and intragenomic variation (which would tend toward ratios on the order of the number of rDNA regions in the genome, i.e.
+10--1000).
+This suggests that even after `DADA2` denoising, the Ion Torrent dataset still includes spurious ASVs due to PCR and/or sequencing errors.
 
@@
 
+## Spatio-temporal variogram
+
 (ref:spatio) Spatiotemporal variogram fits for community dissimilarities.
@@
 
-## Spatio-temporal variogram
 ```{r spatio-temporal, fig.cap = "(ref:spatio)"}
@@
   scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
-  scale_color_discrete() +
+  scale_color_discrete(name = "Time lag") +
   scale_y_continuous(name = "Community dissimilarity") +
@@
 
-(ref:variofit) Parameters for variogram fits shown in Figure \@ref(fig:spatio-temporal).
-Range: range at which exponential function is at 95% of its maximum value; Anisotropy: effective spatial distance represented by one year; the characteristic range of the exponential function in the time domain.
+(ref:variofit) Parameters for variogram fits shown in Figure \@ref(fig:spatio-temporal). Range: range at which exponential function is at 95% of its maximum value.
 
@@
   mutate_at("Range", paste, "m") %>%
-  mutate_at("Anisotropy", paste, "m/a") %>%
-  mutate_at("R/A", paste, "a") %>%
-  mutate_at("R/A", str_replace, "NA a", "-") %>%
+  mutate_at("Time Range", paste, "a") %>%
+  mutate_at("Time Range", str_replace, "NA a", "-") %>%
   mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
+  select(Guild, Amplicon, Metric, "Space Range" = Range, "Time Range") %>%
   kable(booktabs = TRUE, caption = "(ref:variofit)")
@@
 # Discussion
+
+## Comparison of tech
+
+  - length bias not clearly observed between PacBio and Ion Torrent short reads
+  - long reads had more very short ITS2 sequences -- who are these?
+  - lots of ASVs/OTUs detected in only one tech
+
+  - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
+  - sequencing technologies yield comparable results (given read depth) with sufficient clustering
+  - not such good match with ASVs -- suggests DADA2 doesn't perform equally on different datasets
+  - No "hook", because depth is similar
+  - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
 

### 2020-01-24 scripts/drake_plan_light.R

@@ -14,7 +14,9 @@ library(assertr)
 library(disk.frame)
 
 source(file.path(config$rdir, "variogram.R"))
-
+datasets <- read_csv(config$dataset, col_types = "cccccccicccccicc") %>%
+  mutate_at("amplicon", stringr::str_to_lower) %>%
+  filter(tech != "Illumina")
 physeq_meta <-
   tidyr::crossing(
     dplyr::select(datasets, "seq_run", "tech", "dataset", "amplicon"),
@@ -277,7 +279,7 @@ plan2 <- drake_plan(
   physeq = target(
     {
       physeq <- proto_physeq
-      if (amplicon == "Long") {
+      if (amplicon == "long") {
         phyloseq::phy_tree(physeq) <- fungi_tree_decipher_unconst_long
       }
       physeq <- physeq %>%
@@ -349,7 +351,7 @@ plan2 <- drake_plan(
   ),
   trace = TRUE
 ) %>%
-  dplyr::filter(ifelse(amplicon == '"Short"', metric != '"wunifrac"', TRUE) %|% TRUE)
+  dplyr::filter(ifelse(amplicon == '"short"', metric != '"wunifrac"', TRUE) %|% TRUE)
 
 saveRDS(plan2, "data/plan/drake_light.rds")
 

### 2020-01-29 writing/transect_paper.Rmd
move more data munging to drake plan, work on heat trees

@@
 csl: molecular-ecology.csl
+always_allow_html: true
 output:
@@
   )
+
 knitr::opts_chunk$set(
@@
 
-reads_table <- cache$list() %>%
-  purrr::keep(startsWith, "big_seq_table_") %>%
-  purrr::map(readd, character_only = TRUE, cache = cache) %>%
-  purrr::map(tibble::as_tibble, rownames = "sample") %>%
-  purrr::map_dfr(tidyr::pivot_longer, -"sample", names_to = "seq", values_to = "reads") %>%
-  dplyr::filter(reads > 0) %>%
-  tidyr::extract(
-    "sample",
-    c("seq_run", "plate", "well", "region"),
-    regex = "([pi][bs]_\\d{3})_(\\d{3})_([A-H]1?\\d)_(.+)"
-  ) %>%
-  # dplyr::mutate_at(
-  #   "region",
-  #   factor,
-  #   levels = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3",
-  #              "D3", "LSU4", "ITS", "LSU", "32S", "long")) %>%
-  dplyr::group_by(seq_run, region, seq) %>%
-  dplyr::summarize(reads = sum(reads)) %>%
-  dplyr::mutate(length = nchar(seq)) %>%
-  dplyr::ungroup()
-
-region_table <- reads_table %>%
-  dplyr::group_by(seq_run, region) %>%
-  dplyr::summarize(
-    ASVs = dplyr::n_distinct(seq),
-    length_min = min(length),
-    length_q1 = reldist::wtd.quantile(length, 0.25, weight = reads),
-    length_med = reldist::wtd.quantile(length, 0.5, weight = reads),
-    length_mean = weighted.mean(length, w = reads),
-    length_q3 = reldist::wtd.quantile(length, 0.75, weight = reads),
-    length_max = max(length),
-    reads = sum(reads)
-  )
-loadd(big_seq_table_ITS2, cache = cache)
-asv_table <- big_seq_table_ITS2 %>%
-  tibble::as_tibble(rownames = "filename") %>%
-  tidyr::gather(key = "seq", value = "reads", -1) %>%
-  dplyr::filter(reads >= 1) %>%
-  tidyr::extract(col = "filename",
-                 into = c("seq.run", "plate", "well", "dir", "region"),
-                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
-  dplyr::group_by(seq.run, seq) %>%
-  dplyr::summarize(reads = sum(reads)) %>%
-  tidyr::spread(key = seq.run, value = reads, fill = 0)
-
-otu <- read_tsv(
-  here("data/clusters/ITS2.table"),
-  col_types = cols(
-    .default = col_integer(),
-    `#OTU ID` = col_character()
-  )
-) %>%
-  column_to_rownames("#OTU ID") %>%
-  as.matrix() %>%
-  t() %>%
-  as_tibble(rownames = "sample") %>%
-  tidyr::extract(
-    col = "sample",
-    into = c("seq_run", "plate", "well"),
-    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
-  ) %>%
-  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
-  dplyr::filter(reads >= 1) %>%
-  dplyr::group_by(seq_run, seq) %>%
-  dplyr::summarize(reads = sum(reads)) %>%
-  tidyr::spread(key = seq_run, value = reads, fill = 0)
+loadd(
+  reads_table,
+  region_table,
+  asv_table,
+  otu_table,
+  demuxlength,
+  demuxqual,
+  readcounts,
+  rawcounts,
+  demuxcounts,
+  regioncounts,
+  cache = cache
+)
 
@@
 ### Denoising
-
-```{r length-qual}
-
-parse_qstat <- function(d) {
-  mutate_at(d, "file", basename) %>%
-  tidyr::extract(col = "file", into = c("seq_run", "plate", "well"),
-                 regex = "([pi][bs]_\\d{3})_(\\d{3})-?([A-H]1?\\d)?[rf]?[.].*")
-}
-
-demuxlength <- readd(qstats_length, cache = cache) %>%
-  parse_qstat() %>%
-  filter(is.na(step), !is.na(well)) %>%
-  group_by(seq_run) %>%
-  summarize(length = reldist::wtd.quantile(length, weight = nreads)) %>%
-  deframe()
-
-demuxqual <- readd(qstats_erate, cache = cache) %>%
-  parse_qstat() %>%
-  filter(is.na(step), !is.na(well)) %>%
-  group_by(seq_run) %>%
-  summarize(qual = round(weighted.mean(erate, w = nreads, na.rm = TRUE), 4)) %>%
-  deframe()
-
-readcounts <- 
-  readd(qstats_n, cache = cache) %>%
-  parse_qstat()
-
-rawcounts <- readcounts %>%
-  filter(is.na(step), is.na(well)) %>%
-  group_by(seq_run) %>%
-  summarize(nreads =  prettyNum(sum(nreads), big.mark = " ")) %>%
-  deframe()
-
-demuxcounts <- readcounts %>%
-  filter(is.na(step), !is.na(well)) %>%
-  group_by(seq_run) %>%
-  summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
-  deframe()
-
-regioncounts <- readcounts %>%
-  filter(step == "lsux") %>%
-  group_by(seq_run) %>%
-  summarize(nreads = prettyNum(
-    sum(nreads) %/% if (seq_run[1] == "pb_500") 14 else 4,
-    big.mark = " ")
-  ) %>%
-  deframe()
-```
 
@@
 For each region/reference combination, taxonomy was assigned using three algorithms:
-the RDP Naïve Bayesian Classifier (NBC) as implemented in DADA2;
+the RDP Naïve Bayesian Classifier (RDPC) as implemented in DADA2;
 SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
@@
 ```{r bioinfo}
-bioinf_table <- bind_rows(
-  enframe(rawcounts, name = "seq_run", value = "reads") %>%
-    mutate(
-      reads = as.integer(gsub(" ", "", reads)),
-      step = "Raw"
-    ),
-  enframe(demuxcounts, name = "seq_run", value = "reads") %>%
-    mutate(
-      reads = as.integer(gsub(" ", "", reads)),
-      step = "Trim"
-    ),
-  enframe(regioncounts, name = "seq_run", value = "reads") %>%
-    mutate(
-      reads = as.integer(gsub(" ", "", reads)),
-      step = "LSUx"
-    ),
-  filter(region_table, region == "ITS2") %>%
-    select(seq_run, reads, step = region, ASVs),
-  pivot_longer(asv_table, -1, names_to = "seq_run", values_to = "reads") %>%
-    filter(reads > 0) %>%
-    mutate_at("seq", chartr, old = "T", new = "U") %>%
-    inner_join(
-      select(readd(allseqs, cache = cache), seq = ITS2, long, short, ITS, LSU) %>%
-        pivot_longer(-1, names_to = "region", values_to = "consensus") %>%
-        filter(!is.na(consensus)) %>%
-        select(-consensus) %>%
-        unique(),
-      by = "seq"
-    ) %>%
-    group_by(seq_run, region) %>%
-    summarize(reads = sum(reads, na.rm = TRUE), ASVs = n()) %>%
-    select(seq_run, reads, step = region, ASVs),
-  # pivot_longer(otu, -1, names_to = "seq_run", values_to = "reads") %>%
-  #   filter(reads > 0) %>%
-  #   mutate_at("seq", chartr, old = "T", new = "U") %>%
-  #   inner_join(
-  #     select(readd(allseqs, cache = cache), seq = ITS2, long, short, ITS, LSU) %>%
-  #       pivot_longer(-1, names_to = "region", values_to = "consensus") %>%
-  #       filter(!is.na(consensus)) %>%
-  #       select(-consensus) %>%
-  #       unique(),
-  #     by = "seq"
-  #   ) %>%
-  #   group_by(seq_run, region) %>%
-  #   summarize(reads = sum(reads, na.rm = TRUE), OTUs = n()) %>%
-  #   select(seq_run, reads, step = region, OTUs)
-) %>%
-  left_join(select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
-  mutate(
-    step = ifelse(tech == "Ion" & step == "Trim", "Raw", step),
-  ) %>%
-  select(tech, amplicon, step, reads, ASVs) %>%
-  pivot_longer(c("reads", "ASVs"), names_to = "type", values_to = "count") %>%
-  filter(!is.na(count)) %>%
-  mutate(
-    step = factor(step, levels = c("Raw", "Trim", "LSUx", "ITS2",
-                                   "short", "ITS", "LSU", "long")),
-    tech = factor(tech, levels = c("PacBio", "Ion Torrent", "Illumina")),
-    amplicon = factor(amplicon, levels = c("Long", "Short")),
-    type = factor(type, c("ASVs", "reads"))
-  ) %>%
-  arrange(tech, amplicon, type) %>%
-  pivot_wider(
-    id_cols = "step",
-    names_from = c("tech", "amplicon", "type"),
-    values_from = c("count")
-    ) %>%
-  arrange(step) %>%
-  column_to_rownames("step")
-
+loadd(bioinf_table, cache = cache)
 options(knitr.kable.NA = '–')
@@
 ```{r venn-table-asv}
-venn_ASV <-
-  asv_table %>%
-  select(seq, pb_500, pb_483, is_057) %>%
-  mutate_if(is.numeric, list(found = ~. > 0)) %>%
-  group_by_at(vars(ends_with("_found"))) %>%
-  summarize_if(is.numeric, list(ASVs = ~sum(. > 0), reads = sum)) %>%
-  ungroup() %>%
-  mutate_at(vars(ends_with("_found")), as.integer) %>%
-  tidyr::unite(col = "set", 1:3, sep = "") %>%
-  pivot_longer(
-    -1,
-    names_to = c("seq_run", "what"),
-    values_to = "value",
-    names_pattern = "([[:alpha:]]+[-._]\\d+)_(.+)"
-  ) %>%
-  pivot_wider(names_from = "what", values_from = "value") %>%
-  group_by(seq_run) %>%
-  mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
-  mutate_at(
-    vars(ends_with("_frac")),
-    formatC,
-    digits = 2,
-    format = "f",
-    drop0trailing = FALSE
-  ) %>%
-  mutate_at(
-    c("reads", "ASVs"),
-    format
-  ) %>%
-  mutate(
-    ASVs_frac = if_else(grepl(" +0", ASVs), "", ASVs_frac),
-    ASVs = if_else(grepl(" +0", ASVs), "", ASVs),
-    reads_frac = if_else(grepl(" +0", reads), "", reads_frac),
-    reads = if_else(grepl(" +0", reads), "", reads)
-  ) %>%
-  select(set, seq_run, ASVs, ASVs_frac, reads, reads_frac) %>% {
-    left_join(
-      group_by(., set) %>%
-        summarize(ASVs = as.integer(first(ASVs[ASVs != ""]))),
-
-      select(., set, seq_run, ASV_frac = ASVs_frac, reads, frac = reads_frac) %>%
-      pivot_wider(
-        names_from = "seq_run",
-        values_from = c("ASV_frac", "reads", "frac")
-      )
-    )
-  } %>%
-  mutate(n = str_count(set, "1")) %>%
-  arrange(n, desc(set)) %>%
-  select(set, ASVs, ends_with("pb_500"), ends_with("pb_483"), ends_with("is_057")) %>%
-  column_to_rownames("set")
+loadd(venn_ASV, cache = cache)
 
@@
 ```{r venn-table-otu}
-venn_OTU <-
-  otu %>%
-  select(seq, pb_500, pb_483, is_057) %>%
-  mutate_if(is.numeric, list(found = ~. > 0)) %>%
-  group_by_at(vars(ends_with("_found"))) %>%
-  summarize_if(is.numeric, list(OTUs = ~sum(. > 0), reads = sum)) %>%
-  ungroup() %>%
-  mutate_at(vars(ends_with("_found")), as.integer) %>%
-  tidyr::unite(col = "set", 1:3, sep = "") %>%
-  pivot_longer(
-    -1,
-    names_to = c("seq_run", "what"),
-    values_to = "value",
-    names_pattern = "([[:alpha:]]+[-._]\\d+)_(.+)"
-  ) %>%
-  pivot_wider(names_from = "what", values_from = "value") %>%
-  group_by(seq_run) %>%
-  mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
-  mutate_at(
-    vars(ends_with("_frac")),
-    formatC,
-    digits = 2,
-    format = "f",
-    drop0trailing = FALSE
-  ) %>%
-  mutate_at(
-    c("reads", "OTUs"),
-    format
-  ) %>%
-  mutate(
-    OTUs_frac = if_else(grepl(" +0", OTUs), "", OTUs_frac),
-    OTUs = if_else(grepl(" +0", OTUs), "", OTUs),
-    reads_frac = if_else(grepl(" +0", reads), "", reads_frac),
-    reads = if_else(grepl(" +0", reads), "", reads)
-  ) %>%
-  select(set, seq_run, OTUs, OTUs_frac, reads, reads_frac) %>%  {
-    left_join(
-      group_by(., set) %>%
-        summarize(OTUs = as.integer(first(OTUs[OTUs != ""]))),
-
-      select(., set, seq_run, OTU_frac = OTUs_frac, reads, frac = reads_frac) %>%
-      pivot_wider(
-        names_from = "seq_run",
-        values_from = c("OTU_frac", "reads", "frac")
-      )
-    )
-  } %>%
-  mutate(n = str_count(set, "1")) %>%
-  arrange(n, desc(set)) %>%
-  select(set, OTUs, ends_with("pb_500"), ends_with("pb_483"), ends_with("is_057")) %>%
-  column_to_rownames("set")
+loadd(venn_OTU, cache = cache)
 
@@
 ```{r pb-vs-it, fig.cap="(ref:readcap)", fig.height=8}
-big_table <- bind_rows(
-  mutate(asv_table, type = "ASV"),
-  mutate(otu, type = "OTU")
-) %>%
-  select(-seq) %>%
-  select(type, everything())
-
-names(big_table)[-1] <-
-  tibble(seq_run = names(big_table)[-1]) %>%
-  left_join(datasets, by = "seq_run") %>%
-  glue::glue_data("{tech} {machine} - {amplicon}")
-
-multi_table <- crossing(
-  x_var = factor(glue::glue_data(datasets, "{tech} {machine} - {amplicon}"), ordered = TRUE),
-  y_var = factor(glue::glue_data(datasets, "{tech} {machine} - {amplicon}"), ordered = TRUE),
-  type = c("ASV", "OTU")
-) %>%
-  filter(
-    x_var < y_var,
-    x_var %in% names(big_table),
-    y_var %in% names(big_table)
-  ) %>%
-  mutate_all(as.character) %>%
-  group_by(x_var, y_var, type) %>%
-  group_modify(
-    function(d, g) {
-      big_table %>%
-        filter(type == g$type) %>%
-        select(x = !!g$x_var, y = !!g$y_var) %>%
-        filter(x > 0 | y > 0)
-    }
-  )
-
-comparisons <- multi_table %>%
-  group_modify(
-    function(d, g) {
-      filter_all(d, ~. > 0) %>%
-        mutate_all(log10) %>%
-      {bind_cols(
-      lm(y ~ x, .) %>%
-        broom::glance(.),
-      lm(y ~ offset(x), .) %>%
-        broom::tidy(.) %>%
-        select(term, estimate) %>%
-        pivot_wider(names_from = term, values_from = estimate)
-      )}
-    }
-  ) %>%
-  mutate(
-    label = paste("R^2 ==", formatC(r.squared, format = "f", digits = 2))
-  )
-
-# comparisons <- crossing(
-#   x_var = factor(datasets$seq_run, ordered = TRUE),
-#   y_var = factor(datasets$seq_run, ordered = TRUE),
-#   type = c("ASV", "OTU")
-# ) %>%
-#   filter(y_var < x_var) %>%
-#   mutate_all(as.character) %>%
-#   group_by(x_var, y_var, type) %>%
-#   group_modify(function(d, g) {
-#     big_table %>%
-#       filter(type == g$type) %>%
-#       select(x = !!g$x_var, y = !!g$y_var) %>%
-#       filter(.data[[g$x_var]] > 0 & .data[[g$y_var]] > 0) %>%
-#       mutate_all(log10) %>%
-#       lm() %>%
-#       {bind_rows(
-#       broom::glance(.),
-#       broom::tidy(.) %>%
-#         select(term, estimate) %>%
-#         pivot_wider(names_from = term, values_from = estimate)
-#       )} %>%
-#       bind_cols(d, .)
-#     }) %>%
-#   mutate(
-#     label = paste("R^2 ==", formatC(r.squared, format = "f", digits = 2))
-#   )
+loadd(
+  big_table,
+  multi_table,
+  comparisons,
+  cache = cache
+)
 
@@
   format(filter(comparisons, x == x_var, y == y_var, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
-}
-
-# p <- multi_table %>%
-#   filter(x > 0, y > 0, type == "ASV") %>%
-#   ggplot(aes(y = y, x = x)) +
-#   geom_point(alpha = 0.2, shape = 1) +
-#   scale_x_log10() +
-#   scale_y_log10(limits = c(1, 1e4)) +
-#   coord_equal() +
-#   xlab(NULL) +
-#   ylab(NULL) +
-#   geom_rug(aes(y = y), sides = "l", alpha = 0.2,
-#            data = multi_table %>% filter(x == 0, y > 0)) +
-#   geom_rug(aes(x = x), sides = "b", alpha = 0.2,
-#            data = multi_table %>% filter(y == 0, x > 0)) +
-#   facet_grid(y_var ~ x_var, switch = "both", space = "free") +
-#   stat_smooth(method = "loess", formula = y ~ x) +
-#   geom_abline(aes(intercept = `(Intercept)`, slope = 1),
-#               linetype = 2L, alpha = 0.5, data = comparisons) +
-#   geom_text(
-#     aes(label = label),
-#     data = comparisons,
-#     x = 1,
-#     y = 3.5,
-#     color = "black",
-#     hjust = 0,
-#     parse = TRUE,
-#     inherit.aes = FALSE
-#   ) +
-#     theme(strip.placement = "outside", strip.background = element_blank())
-# g <- ggplotGrob(p)
-# g$grobs[g$layout$name %in% c("panel-1-2")] <- NULL
-# g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2")),]
-# grid.newpage()
-# grid.draw(g)
-
-expand_seq_run <- function(labels) {
-  lapply(
-    labels,
-    )
 }
@@
 }
-# ggpubr::ggarrange(
-#   ncol = 1,
-#   labels = "auto",
-#   read_comparison(
-#     big_table,comparisons,
-#     is_057, pb_500,
-#     "Ion Torrent Short Reads", "PacBio Long Reads"),
-#   read_comparison(
-#     big_table, comparisons,
-#     is_057, pb_483,
-#     "Ion Torrent Short Reads", "PacBio Short Reads"),
-#   read_comparison(
-#     big_table, comparisons,
-#     pb_483, pb_500,
-#     "PacBio Short Reads", "PacBio Long Reads")
-# )
 ```
@@
 
+The fraction of reads and ASVs assigned to each taxonomic rank is shown in Figure \@ref(fig:tax-chart).
+For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs.
+A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
+Across sequencing technologies, amplicons, and reference databases, the RDPC algorithm assigned the greatest fraction of reads and ASVs, followed by SINTAX, and then IDTAXA, with the exception of lower taxonomic ranks (order for the Long amplicon, family and genus for both amplicons) using the warcup reference database, where SINTAX assigned the greatest fraction of reads, but not ASVs.
+
+(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, RDPC, SINTAX, IDTAXA).
+
+```{r tax-chart, fig.cap = "(ref:tax-chart)"}
+readd(taxon_reads, cache = cache) %>%
+  pivot_longer(names_to = "rank", cols = c("kingdom", "phylum", "class", "order", "family", "genus"), values_to = "taxon") %>%
+  group_by(amplicon, tech, Algorithm, reference, rank, ID = !is.na(taxon)) %>%
+  summarize(reads = sum(reads), ASVs = n()) %>%
+  group_by(amplicon, tech, Algorithm, reference, rank) %>%
+  mutate(ASVs = ASVs / sum(ASVs)) %>%
+  ungroup() %>%
+  filter(ID) %>%
+  mutate_at("rank", factor, levels = c("kingdom", "phylum", "class", "order", "family", "genus")) %>%
+  rename(Reads = reads) %>%
+  mutate_at("reference", as.character) %>%
+  mutate(
+    reference = map(reference, ~ if (. == "All") c("Unite", "Warcup", "RDP") else .)
+  ) %>%
+  unnest(reference) %>%
+  mutate_at("reference", factor, levels = c("Unite", "Warcup", "RDP")) %>%
+  pivot_longer(cols = c("Reads", "ASVs"), names_to = "type", values_to = "frac") %>%
+  ggplot(aes(x = rank, y = frac, ymax = frac, group = Algorithm, fill = Algorithm)) +
+  ggnomics::facet_nested(reference ~ type + tech + amplicon) +
+  # geom_ribbon(ymin = 0, alpha = 0.5) +
+  # geom_line(aes(color = Algorithm)) +
+  geom_col(position = "dodge") +
+  ylab("Fraction of reads assigned") +
+  xlab(NULL) +
+  theme(axis.text.x = element_text(vjust = 0.5, angle = 90),
+        legend.direction = "horizontal",
+        legend.position = "bottom")
+```
+
+Taxonomic composition of the sequenced soil community is summarized in Figure \@ref(fig:taxa).
+
   - what fraction assigned at each level (or representative ranks; phylum, (order), genus)
@@
 
-```{r}
-taxon_reads <- 
-  readd(proto_physeq, cache = cache) %>%
-  phyloseq::merge_samples(group = "seq_run") %>%
-  phyloseq::otu_table() %>%
-  t() %>%
-  {.@.Data} %>%
-  as_tibble(rownames = "label") %>%
-  mutate_if(is.numeric, ~./sum(.)) %>%
-  pivot_longer(-1, names_to = "seq_run", values_to = "reads") %>%
-  filter(reads > 0) %>%
-  left_join(select(datasets, seq_run, amplicon), by = "seq_run") %>%
-  expand_grid(
-    tibble(
-      reference = c("unite", "warcup", "rdp_train"),
-      region = c("ITS", "ITS", "LSU")
-    ),
-      method = c("dada2", "idtaxa", "sintax")
-  ) %>%
-  filter(
-    ifelse(amplicon == "Short", region == "ITS", TRUE),
-    ifelse(amplicon == "Long", region %in% c("ITS", "LSU"), TRUE)
-  ) %>%
-  left_join(
-    readd(taxon_table, cache = cache) %>%
-      mutate(region = ifelse(region == "short", "ITS", region)) %>%
-      mutate_at("taxon", na_if, "NA") %>%
-      select(label, method, reference, rank, taxon, region) %>%
-      pivot_wider(
-        names_from = "rank",
-        values_from = "taxon"
-      ),
-    by = c("label", "reference", "region", "method")
-  ) %>%
-  mutate(
-    kingdom = ifelse(!is.na(phylum) & reference == "warcup", "Fungi", kingdom)
-  )
-```
-
-(ref:taxa-kingdom) Taxonomic assignment of ASVs to kingdom.
-N/A denotes ASVs which could not be identified to kingdom level.
+(ref:taxa-kingdom) Taxonomic assignment of ASVs to kingdom by different reference databases and algorithms.
+"unidentified" denotes ASVs which could not be identified to kingdom level.
+"other" denotes kingdoms represented by less than 1% of reads in all datasets.
+The Warcup database did not include non-Fungi, so for that database, ASVs were considered to be identified to Fungi if they were identified to any phylum.
 Height of each bar represents the fraction of reads represented by that bar.
 
-```{r taxa-kingdom, fig.cap = "(ref:taxa-kingdom)"}
-taxon_reads %>%
-  group_by(reference, method, seq_run, kingdom) %>%
-  summarize(reads = sum(reads), ASVs = n()) %>%
-  ggplot(aes(x = method, y = reads, fill = kingdom)) +
+```{r taxon-plot}
+taxon_plot <- function(.data, rank, ..., y = reads, cutoff = NULL) {
+  rank <- enquo(rank)
+  y <- enquo(y)
+  ranks <- c("kingdom", "phylum", "class", "order", "family", "genus")
+  .data <- .data %>%
+    group_by(reference, Algorithm, tech, amplicon) %>%
+    mutate(ASVs = n()) %>%
+    ungroup() %>%
+    filter(...) %>%
+    arrange_at(ranks) %>%
+    mutate_at(ranks, ~ factor(., levels = c(NA, "other", discard(unique(.), is.na)))) %>%
+    group_by(reference, Algorithm, tech, amplicon, !!rank) %>%
+    summarize(reads = sum(reads), ASVs = n()/max(ASVs))
+
+  if (!is.null(cutoff)) {
+    .data <- group_by(.data, !!rank) %>%
+      group_map(
+
+        ~ if (all(pull(.x, !!y) < cutoff)) {
+          mutate(.x, !!rank := factor("other", levels = levels(!!rank)))
+        } else {
+          .x
+        },
+        keep = TRUE
+      ) %>%
+      bind_rows() %>%
+      group_by(reference, Algorithm, !!rank, tech, amplicon) %>%
+      summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
+      ungroup()
+  }
+
+  .data <- mutate(.data, !!rank := fct_drop(!!rank))
+  vals <- levels(pull(.data, !!rank))
+  # if ("other" %in% vals) vals <- c("other", vals) %>% magrittr::extract(!duplicated(.))
+  # if (any(is.na(vals))) vals <- c(NA, vals) %>% magrittr::extract(!duplicated(.))
+
+  rank_label <- as_label(rank)
+  if (rank_label == str_to_lower(rank_label)) rank_label <- str_to_title(rank_label)
+  y_label <- as_label(y)
+  if (y_label == str_to_lower(y_label)) y_label <- str_to_title(y_label)
+  ggplot(.data, aes(x = Algorithm, y = !!y, fill = !!rank)) +
     geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
-  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
-  theme(axis.text.x = element_text(angle = 90))
-```
-
-(ref:ASV-kingdom) Taxonomic assignment of ASVs to kingdom.
-N/A denotes ASVs which could not be identified to kingdom level.
+    ggnomics::facet_nested(
+      cols = vars(tech, amplicon, reference),
+      scales = "free_x",
+      space = "free_x",
+      nest_line = TRUE,
+      bleed = FALSE,
+      resect = unit(1, "mm")
+      ) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
+          strip.background = element_blank(),
+          panel.spacing = unit(3, "pt")) +
+    scale_fill_discrete(
+      breaks = vals,
+      labels = replace_na(as.character(vals), "unidentified"),
+      name = rank_label
+    ) +
+    ylab(paste("Fraction of", y_label))
+
+}
+
+```
+
+```{r taxa-kingdom, fig.cap = "(ref:taxa-kingdom)", include = FALSE}
+loadd(taxon_reads, cache = cache)
+taxon_plot(taxon_reads, kingdom, cutoff = 0.01)
+```
+
+(ref:ASV-kingdom) Taxonomic assignment of ASVs to kingdom by different reference databases and algorithms.
+"unidentified" denotes ASVs which could not be identified to kingdom level.
+"other" denotes kingdoms represented by less than 1% of ASVs in all datasets.
+The Warcup database did not include non-Fungi, so ASVs were considered to be identified to Fungi if they were identified to any phylum.
 Height of each bar represents the fraction of total ASVs represented by that bar.
 
-```{r ASV-kingdom, fig.cap = "(ref:ASV-kingdom)"}
-taxon_reads %>%
-  group_by(reference, method, seq_run, kingdom) %>%
-  summarize(reads = sum(reads), ASVs = n()) %>%
-  ggplot(aes(x = method, y = ASVs, fill = kingdom)) +
-  geom_bar(position = "fill", stat = "identity", color = "white", size = 0.2) +
-  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
-  theme(axis.text.x = element_text(angle = 90))
+```{r ASV-kingdom, fig.cap = "(ref:ASV-kingdom)", include = FALSE}
+taxon_plot(taxon_reads, kingdom, y = ASVs, cutoff = 0.01)
 ```
@@
 Phyla representing less than 1% of total reads are grouped as "other".
-N/A denotes ASVs which were identified as fungi, but could not be identified to phylum level.
+"unidentified" denotes ASVs which were identified as fungi, but could not be identified to phylum level.
 Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
@@
 
-```{r taxa-phylum, fig.cap = "(ref:taxa-phylum)"}
-taxon_reads %>%
-  filter(kingdom == "Fungi") %>%
-  group_by(reference, method, phylum, seq_run) %>%
-  summarize(reads = sum(reads), ASVs = n()) %>%
-  ungroup() %>%
-  mutate(
-    phylum = ifelse(reads < 0.01 & !is.na(phylum), " other", phylum)
-  ) %>%
-  group_by(reference, method, phylum, seq_run) %>%
-  summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
-  # complete(reference, method, phylum, fill = list(reads = 0, ASVs = 0)) %>%
-  ggplot(aes(x = method, y = reads, fill = phylum)) +
-  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
-  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
-  theme(axis.text.x = element_text(angle = 90))
-
+```{r taxa-phylum, fig.cap = "(ref:taxa-phylum)", include = FALSE}
+taxon_plot(taxon_reads, rank = phylum, cutoff = 0.01, kingdom == "Fungi")
 ```
@@
 Orders representing less than 1% of total reads are grouped as "other".
-N/A denotes ASVs which were identified as Ascomycota, but could not be identified to order level.
+"unidentified" denotes ASVs which were identified as Ascomycota, but could not be identified to order level.
 Reads which were not identified to phylum, or which were identified as a non-Ascomycota phylum, are not included.
@@
 
-```{r asco-class, fig.cap = "(ref:asco-class)"}
-taxon_reads %>%
-  filter(phylum == "Ascomycota") %>%
-  group_by(reference, method, order, seq_run) %>%
-  summarize(reads = sum(reads), ASVs = n()) %>%
-  ungroup() %>%
-  mutate(
-    order = ifelse(reads < 0.01 & !is.na(order), " other", order)
-  ) %>%
-  group_by(reference, method, order, seq_run) %>%
-  summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
-  ggplot(aes(x = method, y = reads, fill = order)) +
-  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
-  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
-  theme(axis.text.x = element_text(angle = 90))
-
+```{r asco-class, fig.cap = "(ref:asco-class)", include = FALSE}
+taxon_plot(taxon_reads, order, phylum == "Ascomycota", cutoff = 0.01)
 ```
@@
 Families representing less than 1% of total reads are grouped as "other".
-N/A denotes ASVs which were identified as Basidiomycota, but could not be identified to family level.
+"unidentified" denotes ASVs which were identified as Basidiomycota, but could not be identified to family level.
 Reads which were not identified to phylum, or which were identified as a non-Basidiomycota phylum, are not included.
@@
 
-```{r basid-family, fig.cap = "(ref:basid-family)"}
+```{r basid-family, fig.cap = "(ref:basid-family)", include = FALSE}
+taxon_plot(taxon_reads, family, phylum == "Basidiomycota", cutoff = 0.01)
+```
+
+(ref:ecm) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic identification by different reference databases and algorithms.
+Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
+"unidentified" denotes ASVs which were identified as Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
+Height of each bar represents the fraction of reads represented by that bar.
+
+```{r ECM, fig.cap = "(ref:ecm)"}
+  readd(taxon_reads, cache = cache) %>%
+  taxon_plot(ECM, kingdom == "Fungi")
+```
+
+(ref:ecm-families) Taxonomic composition of
+
+```{r ECM-families, fig.cap = "(ref:ecm-families)"}
+  readd(taxon_reads, cache = cache) %>%
+  taxon_plot(
+    family,
+    !is.na(as.character(ECM)),
+    ECM != "non-ECM",
+    kingdom == "Fungi",
+    cutoff = 0.01
+  )
+```
+
+```{r taxdata}
+loadd(taxdata, cache = cache)
+```
+
+(ref:taxa) Heat tree summarizing taxonomic composition of soil community.
+Color and size of nodes represent the fractional read abundance of all ASVs belonging to that taxonomic group.
+Color and thickness of branches represent the fractional ASV richness of that taxonomic group.
+Both measures are a mean across sequencing run, reference database, and taxonomic assignment algorithm.
+Diverse but relatively rare groups are thus shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as large nodes on relatively thin branches (e.g., Russulales).
+Groups which are not represented by at least 1% of reads or 1% of ASVs are not shown.
+
+```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
+set.seed(3)
+readd(taxdata, cache = cache) %>%
+  metacoder::heat_tree(
+    layout = "davidson-harel",
+    # initial_layout = "reingold-tilford",
+    node_size = rowMeans(.$data$tax_read[,-1]),
+    node_color = rowMeans(.$data$tax_read[,-1]),
+    node_label = replace_na(taxon_names, "unknown"),
+    node_size_range = c(0.005, .035),
+    node_size_axis_label = "Read abundance",
+    node_color_axis_label = "Read abundance",
+    node_label_size_range = c(0.01, 0.03),
+    edge_size = rowMeans(.$data$tax_asv[,-1]),
+    edge_size_trans = "linear",
+    edge_size_axis_label = "ASV count",
+    edge_size_range = c(0.0025, 0.03),
+    edge_color = rowMeans(.$data$tax_asv[,-1]),
+    edge_color_axis_label = "ASV count",
+    aspect_ratio = 3/2,
+    make_node_legend = FALSE,
+    make_edge_legend = FALSE,
+    output_file = here("temp/heattree.pdf")
+  )
+```
+
+(ref:heattree-length-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
+Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
+Values are log_2_ of the ratio of median values across all datasets (sequencing technology, assignment algorithm) using the Unite database for each amplicon.
+
+```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare)"}
+taxdata2 <- readd(taxdata, cache = cache)
+# taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
+# taxdata2$data$read_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = read_cols)
+# taxdata2$data$tax_asv <- metacoder::calc_taxon_abund(taxdata2, "asv_table", cols = asv_cols)
+# taxdata2$data$tax_read <- metacoder::calc_taxon_abund(taxdata2, "read_table", cols = read_cols)
+taxdata2$data$diff_read <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_read",
+  cols = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_Unite_.+"),
+  groups = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_Unite_.+") %>%
+    str_replace("reads_.+_(.+)_Unite_.+", "\\1")
+  )
+taxdata2$data$diff_asv <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_asv",
+  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
+  groups = names(taxdata2$data$tax_asv) %>%
+    keep(startsWith, "ASVs_") %>%
+    str_replace("ASVs_.+_(.+)_.+_.+", "\\1")
+)
+
+set.seed(3)
+taxdata2 %>%
+  metacoder::heat_tree(
+    layout = "davidson-harel",
+    initial_layout = "reingold-tilford",
+
+    node_size = rowMeans(.$data$tax_read[,-1]),
+    node_size_range = c(0.005, .035),
+    node_size_axis_label = "Read abundance",
+
+    node_color = .$data$diff_read$log2_median_ratio,
+    node_color_range = metacoder::diverging_palette(),
+    node_color_trans = "linear",
+    node_color_axis_label = "Read abundance ratio",
+    node_color_interval = c(-4.5, 4.5),
+
+    edge_size = rowMeans(.$data$tax_asv[,-1]),
+    edge_size_range = c(0.0025, 0.03),
+    edge_size_trans = "linear",
+    edge_size_axis_label = "ASV count",
+
+    edge_color = .$data$diff_asv$log2_median_ratio,
+    edge_color_range = metacoder::diverging_palette(),
+    edge_color_trans = "linear",
+    edge_color_axis_label = "ASV count ratio",
+    edge_color_interval = c(-3, 3),
+
+    node_label = ifelse(
+      pmax(abs(.$data$diff_read$log2_median_ratio),
+           abs(.$data$diff_asv$log2_median_ratio)) > 0.5,
+      taxon_names,
+      ""
+    ),
+    node_label_size_range = c(0.010, 0.03),
+    title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2))
+  )
+```
+
+(ref:heattree-db-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
+Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
+Values are log_2_ of the ratio of median values across all datasets (sequencing technology, reference database, assignment algorithm) for each amplicon.
+
+```{r heattree-db-compare, fig.cap = "(ref:heattree-db-compare)"}
+taxdata2 <- readd(taxdata, cache = cache)
+# taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
+# taxdata2$data$read_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = read_cols)
+# taxdata2$data$tax_asv <- metacoder::calc_taxon_abund(taxdata2, "asv_table", cols = asv_cols)
+# taxdata2$data$tax_read <- metacoder::calc_taxon_abund(taxdata2, "read_table", cols = read_cols)
+taxdata2$data$diff_read <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_read",
+  cols = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_PacBio_Long_(Unite|RDP)"),
+  groups = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_PacBio_Long_(Unite|RDP)") %>%
+    str_replace("reads_.+_.+_(.+)_.+", "\\1")
+  )
+taxdata2$data$diff_asv <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_asv",
+  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
+  groups = names(taxdata2$data$tax_asv) %>%
+    keep(startsWith, "ASVs_") %>%
+    str_replace("ASVs_.+_(.+)_.+_.+", "\\1")
+)
+
+set.seed(2)
+taxdata2 %>%
+  metacoder::heat_tree(
+    layout = "davidson-harel",
+    initial_layout = "reingold-tilford",
+
+    node_size = rowMeans(.$data$tax_read[,-1]),
+    node_size_range = c(0.005, .035),
+    node_size_axis_label = "Read abundance",
+
+    node_color = .$data$diff_read$log2_median_ratio,
+    node_color_range = metacoder::diverging_palette(),
+    node_color_trans = "linear",
+    node_color_axis_label = "Read abundance ratio",
+    node_color_interval = c(-4.5, 4.5),
+
+    edge_size = rowMeans(.$data$tax_asv[,-1]),
+    edge_size_range = c(0.0025, 0.03),
+    edge_size_trans = "linear",
+    edge_size_axis_label = "ASV count",
+
+    edge_color = .$data$diff_asv$log2_median_ratio,
+    edge_color_range = metacoder::diverging_palette(),
+    edge_color_trans = "linear",
+    edge_color_axis_label = "ASV count ratio",
+    edge_color_interval = c(-3, 3),
+
+    node_label = ifelse(
+      pmax(abs(.$data$diff_read$log2_median_ratio),
+           abs(.$data$diff_asv$log2_median_ratio)) > 0.5,
+      taxon_names,
+      ""
+    ),
+    node_label_size_range = c(0.010, 0.03),
+    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
+    aspect_ratio = 3/2
+  )
+```
+
+```{r ecm-heattree}
+taxdata2 <- readd(taxdata_ECM, cache = cache)
+
+val_cols <- names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_Unite_.+")
+
+taxdata2$filter_obs(
+  "tax_read",
+  rowSums(select_at(taxdata2$data$tax_read, val_cols)) > 0,
+  drop_taxa = TRUE,
+  drop_obs = TRUE,
+  supertaxa = TRUE,
+  reassign_obs = FALSE) %>%
+  invisible()
+
+taxdata2$data$diff_read <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_read",
+  cols = val_cols,
+  groups = val_cols %>%
+    str_replace("reads_.+_(.+)_Unite_.+", "\\1"),
+  func = function(abund1, abund2) {
+    list(read_ratio = log10(mean(abund1) / mean(abund2)))
+  }
+)
+taxdata2$data$diff_asv <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_asv",
+  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
+  groups = names(taxdata2$data$tax_asv) %>%
+    keep(startsWith, "ASVs_") %>%
+    str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
+  func = function(abund1, abund2) {
+    list(ASV_ratio = log10(mean(abund1) / mean(abund2)))
+  }
+)
+
+set.seed(3)
+taxdata2 %>%
+  metacoder::heat_tree(
+    layout = "davidson-harel",
+    initial_layout = "reingold-tilford",
+
+    node_size = rowMeans(.$data$tax_read[,-1]),
+    node_size_range = c(0.01, .035),
+    node_size_axis_label = "Read abundance",
+
+    node_color = read_ratio,
+    node_color_range = metacoder::diverging_palette(),
+    node_color_trans = "linear",
+    node_color_axis_label = "Read abundance ratio",
+    node_color_interval = c(-3, 3),
+
+    edge_size = rowMeans(.$data$tax_asv[,-1]),
+    edge_size_range = c(0.005, 0.03),
+    edge_size_trans = "linear",
+    edge_size_axis_label = "ASV count",
+
+    edge_color = ASV_ratio,
+    edge_color_range = metacoder::diverging_palette(),
+    edge_color_trans = "linear",
+    edge_color_axis_label = "ASV count ratio",
+    edge_color_interval = c(-0.75, 0.75),
+
+    node_label = taxon_names,
+    node_label_size_range = c(0.02, 0.03),
+    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
+    aspect_ratio = 3/2
+  )
+```
+
+```{r heattree-ecm, fig.cap = "Heat tree for all sequencing reads."}
+taxdata <- 
   taxon_reads %>%
-  filter(phylum == "Basidiomycota") %>%
-  group_by(reference, method, family, seq_run) %>%
-  summarize(reads = sum(reads), ASVs = n()) %>%
-  ungroup() %>%
-  mutate(
-    family = ifelse(reads < 0.01 & !is.na(family), " other", family)
-  ) %>%
-  group_by(reference, method, family, seq_run) %>%
-  summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
-  ggplot(aes(x = method, y = reads, fill = family)) +
-  geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
-  facet_wrap(~seq_run + reference, nrow = 1, scales = "free_x") +
-  theme(axis.text.x = element_text(angle = 90))
-
+  filter(!is.na(as.character(ECM)), ECM != "non-ECM") %>%
+  select(-region, -seq_run) %>%
+  pivot_wider(
+    names_from = c("tech", "amplicon", "reference", "Algorithm"),
+    values_from = "reads",
+    values_fill = list(reads = 0),
+    names_repair = "universal"
+    ) %>%
+  taxa::parse_tax_data(
+    class_cols = c("kingdom", "phylum", "class", "order", "family", "genus")
+  )
+
+samples <- names(which(map_lgl(taxdata$data$tax_data, is.numeric)))
+
+taxdata$data$asv_table <- metacoder::calc_obs_props(taxdata, "tax_data", cols = samples)
+taxdata$data$tax_table <- metacoder::calc_taxon_abund(taxdata, "asv_table", cols = samples) %>%
+  column_to_rownames("taxon_id")
+taxdata <- taxdata$filter_taxa(do.call(pmax, taxdata$data$tax_table[,-1]) > 0.01)
+taxdata$data$tax_table <- taxdata$data$tax_table[taxdata$taxon_ids(),] %>%
+  rownames_to_column("taxon_id")
+taxdata$data$diff_table <- metacoder::compare_groups(
+  taxdata,
+  data = "tax_table",
+  cols = names(taxdata$data$tax_table) %>% keep(endsWith, "Unite_RDPC"),
+  groups = names(taxdata$data$tax_table) %>%
+    keep(endsWith, "Unite_RDPC") %>%
+    str_replace("_Unite_RDPC", "")
+  )
+
+theme_replace(plot.background = element_blank(),
+              panel.border = element_blank(), panel.background = element_blank())
+ metacoder::heat_tree_matrix(
+  taxdata,
+  data = "diff_table",
+  node_size = rowMeans(taxdata$data$tax_table[,-1]),
+  node_label = replace_na(taxon_names, "unknown"),
+  node_size_axis_label = "Proportion of reads",
+  node_color = log2_median_ratio,
+  node_color_range = metacoder::diverging_palette(),
+  node_color_trans = "linear",
+  node_color_interval = c(-3, 3),
+  edge_color_interval = c(-3, 3),
+  key_size = 0.7,
+  seed = 2,
+  layout = "davidson-harel"
+)
 ```
@@
   - Unifrac has less significant autocorrelation.
-
-(ref:variog) Spatial variogram fits for community dissimilarities.
-Each point represents one pair of samples taken at the same time.
-Thick blue and red lines represent the empirical variogram (mean of community dissimilarities at each distance).
-Thin lines is an exponential fit for the relationship between community dissimilarity and spatial distance.
-Horizontal dashed lines represent the values of the fit function at zero distance ("nugget") and infinite distance ("sill").
-The vertical dashed line is at the characteristic distance of the fit function ("range").
-
-```{r variog, fig.cap = "(ref:variog)", include = FALSE}
-variog_data <- light_plan %>%
-  filter(step == "variofit") %>%
-  select(variofit, variog, guild, metric, tech, amplicon) %>%
-  mutate_at(c("guild", "metric", "tech", "amplicon"), str_replace_all, "\"", "") %>%
-  mutate(
-    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
-    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
-    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
-  ) %>%
-  mutate_at(c("variofit", "variog"), lapply, readd, character_only = TRUE, cache = cache) %>%
-  mutate(sills = purrr::map(variofit, ~ cumsum(.$psill)),
-         range = purrr::map_dbl(variofit, ~.$range[2]),
-         vgline = purrr::map(variofit, gstat::variogramLine, maxdist = 25, n = 100))
-
-variog_points <- unnest(variog_data, c("variog")) %>%
-  filter(dist <= 25)
-variog_empir <- variog_points %>%
-  group_by(guild, metric, tech, amplicon, dist) %>%
-  summarize_at("gamma", mean)
-variog_sills <- unnest(variog_data, "sills")
-variog_fitline <- unnest(variog_data, "vgline")
-
-variog_points %>%
-  ggplot(aes(dist, gamma)) +
-  geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
-  geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
-  geom_point(shape = 1, alpha = 0.2) +
-  geom_line(data = variog_empir, color = "blue", size = 1.5) +
-  geom_line(data = variog_fitline,
-            aes(dist, gamma), inherit.aes = FALSE, color = "red", size = 0.5) +
-  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
-  ylab("Community dissimilarity") +
-  facet_grid(guild ~ metric + amplicon)
-
-```
 

### 2020-02-01 writing/transect_paper.Rmd
merge duplicate taxa in heat trees, remove Angaradebou data due to low read counts, clean up taxonomy section and heat trees, add supplementary plots about read depth in different samples

@@
   ggplot(aes(x = rank, y = frac, ymax = frac, group = Algorithm, fill = Algorithm)) +
-  ggnomics::facet_nested(reference ~ type + tech + amplicon) +
+  ggnomics::facet_nested(
+    reference ~ type + tech + amplicon,
+    resect = unit(1, "mm"),
+    nest_line = TRUE,
+    bleed = FALSE
+  ) +
   # geom_ribbon(ymin = 0, alpha = 0.5) +
@@
         legend.direction = "horizontal",
-        legend.position = "bottom")
+        legend.position = "bottom",
+        strip.background = element_blank())
 ```
@@
     arrange_at(ranks) %>%
-    mutate_at(ranks, ~ factor(., levels = c(NA, "other", discard(unique(.), is.na)))) %>%
+    mutate_at(
+      ranks,
+      ~ factor(
+        .,
+        levels = c(NA, "other", discard(unique(.), is.na)),
+        exclude = NULL
+      )
+    ) %>%
     group_by(reference, Algorithm, tech, amplicon, !!rank) %>%
-    summarize(reads = sum(reads), ASVs = n()/max(ASVs))
+    summarize(reads = sum(reads), ASVs = n()/max(ASVs)) %>%
+    ungroup()
 
@@
 
+  rank_label <- as_label(rank)
+
   .data <- mutate(.data, !!rank := fct_drop(!!rank))
@@
 
-  rank_label <- as_label(rank)
   if (rank_label == str_to_lower(rank_label)) rank_label <- str_to_title(rank_label)
@@
 
-```{r ECM-families, fig.cap = "(ref:ecm-families)"}
+```{r ECM-families, fig.cap = "(ref:ecm-families)", include = FALSE}
   readd(taxon_reads, cache = cache) %>%
@@
 Color and thickness of branches represent the fractional ASV richness of that taxonomic group.
-Both measures are a mean across sequencing run, reference database, and taxonomic assignment algorithm.
-Diverse but relatively rare groups are thus shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as large nodes on relatively thin branches (e.g., Russulales).
-Groups which are not represented by at least 1% of reads or 1% of ASVs are not shown.
+Both measures are a mean across sequencing runs, reference databases, and taxonomic assignment algorithms.
+Diverse but relatively rare groups are thus shown as small, light-colored nodes on relatively thick, darker-green-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as large nodes on relatively thin branches (e.g., Russulales).
+Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "...".
+ASVs which could not be further assigned are labelled "?".
 
 ```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
-set.seed(3)
+set.seed(4)
 readd(taxdata, cache = cache) %>%
@@
     edge_color_axis_label = "ASV count",
-    aspect_ratio = 3/2,
+    # aspect_ratio = 3/2,
     make_node_legend = FALSE,
@@
 
-(ref:heattree-length-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
-Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
-Values are log_2_ of the ratio of median values across all datasets (sequencing technology, assignment algorithm) using the Unite database for each amplicon.
-
-```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare)"}
-taxdata2 <- readd(taxdata, cache = cache)
-# taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
-# taxdata2$data$read_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = read_cols)
-# taxdata2$data$tax_asv <- metacoder::calc_taxon_abund(taxdata2, "asv_table", cols = asv_cols)
-# taxdata2$data$tax_read <- metacoder::calc_taxon_abund(taxdata2, "read_table", cols = read_cols)
-taxdata2$data$diff_read <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_read",
-  cols = names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_.+_.+_Unite_.+"),
-  groups = names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_.+_.+_Unite_.+") %>%
-    str_replace("reads_.+_(.+)_Unite_.+", "\\1")
-  )
-taxdata2$data$diff_asv <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_asv",
-  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
-  groups = names(taxdata2$data$tax_asv) %>%
-    keep(startsWith, "ASVs_") %>%
-    str_replace("ASVs_.+_(.+)_.+_.+", "\\1")
-)
-
-set.seed(3)
-taxdata2 %>%
-  metacoder::heat_tree(
-    layout = "davidson-harel",
-    initial_layout = "reingold-tilford",
-
-    node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.005, .035),
-    node_size_axis_label = "Read abundance",
-
-    node_color = .$data$diff_read$log2_median_ratio,
-    node_color_range = metacoder::diverging_palette(),
-    node_color_trans = "linear",
-    node_color_axis_label = "Read abundance ratio",
-    node_color_interval = c(-4.5, 4.5),
-
-    edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.0025, 0.03),
-    edge_size_trans = "linear",
-    edge_size_axis_label = "ASV count",
-
-    edge_color = .$data$diff_asv$log2_median_ratio,
-    edge_color_range = metacoder::diverging_palette(),
-    edge_color_trans = "linear",
-    edge_color_axis_label = "ASV count ratio",
-    edge_color_interval = c(-3, 3),
-
-    node_label = ifelse(
-      pmax(abs(.$data$diff_read$log2_median_ratio),
-           abs(.$data$diff_asv$log2_median_ratio)) > 0.5,
-      taxon_names,
-      ""
-    ),
-    node_label_size_range = c(0.010, 0.03),
-    title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2))
-  )
-```
-
 (ref:heattree-db-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
@@
 
-```{r heattree-db-compare, fig.cap = "(ref:heattree-db-compare)"}
+```{r heattree-db-compare, fig.cap = "(ref:heattree-db-compare)", include = FALSE}
 taxdata2 <- readd(taxdata, cache = cache)
@@
     keep(str_detect, "reads_PacBio_Long_(Unite|RDP)") %>%
-    str_replace("reads_.+_.+_(.+)_.+", "\\1")
-  )
-taxdata2$data$diff_asv <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_asv",
-  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
-  groups = names(taxdata2$data$tax_asv) %>%
-    keep(startsWith, "ASVs_") %>%
-    str_replace("ASVs_.+_(.+)_.+_.+", "\\1")
-)
-
-set.seed(2)
-taxdata2 %>%
-  metacoder::heat_tree(
-    layout = "davidson-harel",
-    initial_layout = "reingold-tilford",
-
-    node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.005, .035),
-    node_size_axis_label = "Read abundance",
-
-    node_color = .$data$diff_read$log2_median_ratio,
-    node_color_range = metacoder::diverging_palette(),
-    node_color_trans = "linear",
-    node_color_axis_label = "Read abundance ratio",
-    node_color_interval = c(-4.5, 4.5),
-
-    edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.0025, 0.03),
-    edge_size_trans = "linear",
-    edge_size_axis_label = "ASV count",
-
-    edge_color = .$data$diff_asv$log2_median_ratio,
-    edge_color_range = metacoder::diverging_palette(),
-    edge_color_trans = "linear",
-    edge_color_axis_label = "ASV count ratio",
-    edge_color_interval = c(-3, 3),
-
-    node_label = ifelse(
-      pmax(abs(.$data$diff_read$log2_median_ratio),
-           abs(.$data$diff_asv$log2_median_ratio)) > 0.5,
-      taxon_names,
-      ""
-    ),
-    node_label_size_range = c(0.010, 0.03),
-    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
-    aspect_ratio = 3/2
-  )
-```
-
-```{r ecm-heattree}
-taxdata2 <- readd(taxdata_ECM, cache = cache)
-
-val_cols <- names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_.+_.+_Unite_.+")
-
-taxdata2$filter_obs(
-  "tax_read",
-  rowSums(select_at(taxdata2$data$tax_read, val_cols)) > 0,
-  drop_taxa = TRUE,
-  drop_obs = TRUE,
-  supertaxa = TRUE,
-  reassign_obs = FALSE) %>%
-  invisible()
-
-taxdata2$data$diff_read <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_read",
-  cols = val_cols,
-  groups = val_cols %>%
-    str_replace("reads_.+_(.+)_Unite_.+", "\\1"),
+    str_replace("reads_.+_.+_(.+)_.+", "\\1"),
   func = function(abund1, abund2) {
@@
   func = function(abund1, abund2) {
-    list(ASV_ratio = log10(mean(abund1) / mean(abund2)))
+    list(asv_ratio = log10(mean(abund1) / mean(abund2)))
   }
@@
 
-set.seed(3)
+taxdata2 <- taxdata2$filter_taxa(!is.nan(asv_ratio), !is.nan(read_ratio),
+                                 reassign_obs = FALSE)
+
+set.seed(2)
 taxdata2 %>%
@@
     node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.01, .035),
+    node_size_range = c(0.005, .035),
     node_size_axis_label = "Read abundance",
@@
     node_color_axis_label = "Read abundance ratio",
-    node_color_interval = c(-3, 3),
+    node_color_interval = c(-4.5, 4.5),
 
     edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.005, 0.03),
+    edge_size_range = c(0.0025, 0.03),
     edge_size_trans = "linear",
@@
 
-    edge_color = ASV_ratio,
+    edge_color = asv_ratio,
     edge_color_range = metacoder::diverging_palette(),
@@
     edge_color_axis_label = "ASV count ratio",
-    edge_color_interval = c(-0.75, 0.75),
-
-    node_label = taxon_names,
-    node_label_size_range = c(0.02, 0.03),
+    edge_color_interval = c(-3, 3),
+
+    node_label = ifelse(
+      pmax(abs(.$data$diff_read$read_ratio),
+           abs(.$data$diff_asv$asv_ratio)) > 0.5,
+      taxon_names,
+      ""
+    ),
+    node_label_size_range = c(0.010, 0.03),
     # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
     aspect_ratio = 3/2
-  )
-```
-
-```{r heattree-ecm, fig.cap = "Heat tree for all sequencing reads."}
-taxdata <- 
-  taxon_reads %>%
-  filter(!is.na(as.character(ECM)), ECM != "non-ECM") %>%
-  select(-region, -seq_run) %>%
-  pivot_wider(
-    names_from = c("tech", "amplicon", "reference", "Algorithm"),
-    values_from = "reads",
-    values_fill = list(reads = 0),
-    names_repair = "universal"
-    ) %>%
-  taxa::parse_tax_data(
-    class_cols = c("kingdom", "phylum", "class", "order", "family", "genus")
-  )
-
-samples <- names(which(map_lgl(taxdata$data$tax_data, is.numeric)))
-
-taxdata$data$asv_table <- metacoder::calc_obs_props(taxdata, "tax_data", cols = samples)
-taxdata$data$tax_table <- metacoder::calc_taxon_abund(taxdata, "asv_table", cols = samples) %>%
-  column_to_rownames("taxon_id")
-taxdata <- taxdata$filter_taxa(do.call(pmax, taxdata$data$tax_table[,-1]) > 0.01)
-taxdata$data$tax_table <- taxdata$data$tax_table[taxdata$taxon_ids(),] %>%
-  rownames_to_column("taxon_id")
-taxdata$data$diff_table <- metacoder::compare_groups(
-  taxdata,
-  data = "tax_table",
-  cols = names(taxdata$data$tax_table) %>% keep(endsWith, "Unite_RDPC"),
-  groups = names(taxdata$data$tax_table) %>%
-    keep(endsWith, "Unite_RDPC") %>%
-    str_replace("_Unite_RDPC", "")
-  )
-
-theme_replace(plot.background = element_blank(),
-              panel.border = element_blank(), panel.background = element_blank())
- metacoder::heat_tree_matrix(
-  taxdata,
-  data = "diff_table",
-  node_size = rowMeans(taxdata$data$tax_table[,-1]),
-  node_label = replace_na(taxon_names, "unknown"),
-  node_size_axis_label = "Proportion of reads",
-  node_color = log2_median_ratio,
-  node_color_range = metacoder::diverging_palette(),
-  node_color_trans = "linear",
-  node_color_interval = c(-3, 3),
-  edge_color_interval = c(-3, 3),
-  key_size = 0.7,
-  seed = 2,
-  layout = "davidson-harel"
   )
@@
 ```
+
+(ref:heattree-length-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
+Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
+Values are log_10_ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
+Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
+Refer to Figure \@ref(fig:taxa) for complete taxon labeling.
+
+```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare)"}
+taxdata2 <- readd(taxdata, cache = cache)
+# taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
+# taxdata2$data$read_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = read_cols)
+# taxdata2$data$tax_asv <- metacoder::calc_taxon_abund(taxdata2, "asv_table", cols = asv_cols)
+# taxdata2$data$tax_read <- metacoder::calc_taxon_abund(taxdata2, "read_table", cols = read_cols)
+taxdata2$data$diff_read <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_read",
+  cols = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_.+_.+"),
+  groups = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_.+_.+") %>%
+    str_replace("reads_.+_(.+)_.+_.+", "\\1"),
+  func = function(abund1, abund2) {
+    list(read_ratio = log10(mean(abund1) / mean(abund2)))
+  }
+  )
+taxdata2$data$diff_asv <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_asv",
+  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
+  groups = names(taxdata2$data$tax_asv) %>%
+    keep(startsWith, "ASVs_") %>%
+    str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
+  func = function(abund1, abund2) {
+    list(asv_ratio = log10(mean(abund1) / mean(abund2)))
+  }
+)
+
+set.seed(4)
+taxdata2 %>%
+  metacoder::heat_tree(
+    layout = "davidson-harel",
+    # initial_layout = "reingold-tilford",
+
+    node_size = rowMeans(.$data$tax_read[,-1]),
+    node_size_range = c(0.005, .035),
+    node_size_axis_label = "Read abundance",
+
+    node_color = read_ratio,
+    node_color_range = metacoder::diverging_palette(),
+    node_color_trans = "linear",
+    node_color_axis_label = "Read abundance ratio",
+    node_color_interval = c(-6, 6),
+
+    edge_size = rowMeans(.$data$tax_asv[,-1]),
+    edge_size_range = c(0.0025, 0.03),
+    edge_size_trans = "linear",
+    edge_size_axis_label = "ASV count",
+
+    edge_color = asv_ratio,
+    edge_color_range = metacoder::diverging_palette(),
+    edge_color_trans = "linear",
+    edge_color_axis_label = "ASV count ratio",
+    edge_color_interval = c(-3, 3),
+
+    node_label = ifelse(
+      pmax(abs(.$data$diff_read$read_ratio),
+           abs(.$data$diff_asv$asv_ratio)) > 0.5,
+      taxon_names,
+      ""
+    ),
+    # aspect_ratio = 3/2,
+    node_label_size_range = c(0.010, 0.03)
+    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2))
+  )
+```
+
+(ref:ecm-heattree) Variation in read abundance and diversity for different amplicons, for ECM taxa only.
+Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
+Values are log_10_ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
+
+```{r ecm-heattree, fig.cap = "(ref:ecm-heattree)"}
+taxdata2 <- readd(taxdata_ECM, cache = cache)
+
+val_cols <- names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_Unite_.+")
+
+taxdata2$filter_obs(
+  "tax_read",
+  rowSums(select_at(taxdata2$data$tax_read, val_cols)) > 0,
+  drop_taxa = TRUE,
+  drop_obs = TRUE,
+  supertaxa = TRUE,
+  reassign_obs = FALSE) %>%
+  invisible()
+
+taxdata2$data$diff_read <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_read",
+  cols = val_cols,
+  groups = val_cols %>%
+    str_replace("reads_.+_(.+)_Unite_.+", "\\1"),
+  func = function(abund1, abund2) {
+    list(read_ratio = log10(mean(abund1) / mean(abund2)))
+  }
+)
+taxdata2$data$diff_asv <- metacoder::compare_groups(
+  taxdata2,
+  data = "tax_asv",
+  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
+  groups = names(taxdata2$data$tax_asv) %>%
+    keep(startsWith, "ASVs_") %>%
+    str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
+  func = function(abund1, abund2) {
+    list(ASV_ratio = log10(mean(abund1) / mean(abund2)))
+  }
+)
+
+taxdata2 <- taxdata2$filter_taxa(!is.nan(ASV_ratio), !is.nan(read_ratio))
+
+set.seed(3)
+taxdata2 %>%
+  metacoder::heat_tree(
+    layout = "davidson-harel",
+    initial_layout = "reingold-tilford",
+
+    node_size = rowMeans(.$data$tax_read[,-1]),
+    node_size_range = c(0.01, .035),
+    node_size_axis_label = "Read abundance",
+
+    node_color = read_ratio,
+    node_color_range = metacoder::diverging_palette(),
+    node_color_trans = "linear",
+    node_color_axis_label = "Read abundance ratio",
+    node_color_interval = c(-3, 3),
+
+    edge_size = rowMeans(.$data$tax_asv[,-1]),
+    edge_size_range = c(0.005, 0.03),
+    edge_size_trans = "linear",
+    edge_size_axis_label = "ASV count",
+
+    edge_color = ASV_ratio,
+    edge_color_range = metacoder::diverging_palette(),
+    edge_color_trans = "linear",
+    edge_color_axis_label = "ASV count ratio",
+    edge_color_interval = c(-0.75, 0.75),
+
+    node_label = taxon_names,
+    node_label_size_range = c(0.02, 0.03),
+    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
+    aspect_ratio = 3/2
+  )
+```
+
+(ref:short-its) Kingdom-level taxonomic composition of ITS2 ASVs <= 140 bp.
+Vertical axis shows fraction of total reads in each dataset.
+
+```{r short-its, fig.cap = ""}
+readd(big_seq_table_ITS2, cache = cache) %>%
+  colSums() %>%
+  enframe(name = "ITS2", value = "nreads") %>%
+  mutate(length = nchar(ITS2)) %>%
+  filter(length <= 140) %>%
+  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
+  semi_join(readd(taxon_reads, cache = cache), ., by = c("label" = "ITS2")) %>%
+  # group_by(ITS2) %>%
+  # filter(!"Short" %in% amplicon) %>%
+  taxon_plot(kingdom, y = reads)
+```
+
+(ref:sample-depth) Sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016, after full bioinformatics pipeline.
+Ang 2015 samples are underrepresented, especially in the PacBio long amplicon dataset.
+Dotted horizontal line at proposed cutoff for inclusion in community analysis.
+"2016 X" is using the newer batch of extraction buffer, "2016" is using the older batch.
+
+```{r sample-depth, fig.cap = "(ref:sample-depth)"}
+readd(proto_physeq, cache = cache) %>% {
+  mutate(
+    sample_data(.),
+    reads = otu_table(.) %>% rowSums()
+  )
+} %>%
+  filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
+  complete(seq_run, tech, amplicon, site, x, nesting(year, qual)) %>%
+  mutate(
+    qual = fct_collapse(qual, " " = c("*", "")),
+    "Sample group" = paste(year, qual)
+  ) %>%
+  ggplot(aes(x = x, y = reads, group = `Sample group`, fill = `Sample group`)) +
+  geom_col(position = "dodge") +
+  ggnomics::facet_nested(
+    cols = vars(tech, amplicon, site),
+    rows = vars(`Sample group`),
+    nest_line = TRUE,
+    resect = unit(1, "mm")
+  ) +
+  scale_y_log10() +
+  xlab("x (m)") +
+  theme(
+    legend.position = "bottom",
+    strip.background = element_blank()
+  ) +
+  geom_hline(yintercept = 100, linetype = 2, alpha = 0.5)
+
+```
+
+(ref:plate-depth) Sequencing depth for each well in the sequencing plates (001, 002), after full bioinformatics pipeline.
+Darker shading represents greater sequencing depth, as a fraction of the total number of sequences from the sequencing run.
+The border color for each well where sequences were recovered indicates the sample group, i.e.
+site name (Ang or Gan), sampling year (2015 or 2016), and buffer batch (presence or absence of X; 2016 samples with X are the good buffer); Pos = positive control, Blank = PCR blank.
+Ion Torrent (is_057) samples for the two plates are mixed, shown here as plate 001.
+Note that many of the Ang2015 samples are missing or have extremely low number of reads for the pb_500 (long PacBio) sequencing run.
+
+```{r plate-depth, fig.cap = "(ref:plate-depth)"}
+readd(proto_physeq, cache = cache) %>% {
+  mutate(
+    sample_data(.),
+    reads = otu_table(.) %>% rowSums()
+  )
+} %>%
+  group_by(seq_run, plate) %>%
+  mutate(
+    reads = reads/sum(reads),
+    row = factor(LETTERS[row], levels = rev(LETTERS[1:8])),
+    qual = fct_collapse(qual, " " = c("*", "")),
+    "Sample group" = coalesce(
+      str_c(site, year, qual),
+      as.character(sample_type),
+      "Unused"
+      )
+  ) %>%
+  ggplot(aes(x = column, y = row, fill = reads, color = `Sample group`)) +
+  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
+  ggnomics::facet_nested(cols = vars(plate), rows = vars(tech, amplicon),
+                         resect = unit(1, "mm"), nest_line = TRUE) +
+  scale_color_discrete(name = "Sample group") +
+  scale_fill_gradient(low = "gray80", high = "black") +
+  coord_equal() +
+  scale_x_continuous(breaks = 1:9) +
+  theme(strip.background = element_blank())
+```
+
+(ref:raw-depth) Sequencing depth for each well in the sequencing plates (001, 002), after demultiplexing and primer trimming.
+Darker shading represents greater sequencing depth, as a fraction of the total number of sequences from the sequencing run.
+The border color for each well where sequences were recovered indicates the sample group, i.e.
+site name (Ang or Gan), sampling year (2015 or 2016), and buffer batch (presence or absence of X; 2016 samples with X are the good buffer); Pos = positive control, Blank = PCR blank, NA = unused well.
+Ion Torrent (is_057) samples for the two plates are mixed, shown here as plate 001.
+Note that many of the Ang2015 samples are missing or have extremely low number of reads for the pb_500 (long PacBio) sequencing run.
+
+```{r raw-depth, fig.cap = "(ref:raw-depth)"}
+readd(qstats_n, cache = cache) %>%
+  mutate(file = str_replace(file, ".*/", "") %>%
+           str_replace("[.].*", "")) %>%
+  tidyr::extract(col = "file", into = c("seq_run", "plate", "row", "col"),
+                 regex = "([[:alpha:]]{2}[-_]\\d+)_(\\d+)-([A-H])(\\d+)") %>%
+  filter(is.na(step), !is.na(seq_run)) %>%
+  mutate_at("col", as.integer) %>%
+  left_join(
+    readd(proto_physeq, cache = cache) %>%
+      sample_data() %>%
+      select(plate, row, column, site, year, qual, sample_type) %>%
+      mutate(row = LETTERS[row]) %>%
+      unique(),
+    by = c("plate", "row", "col" = "column")
+  ) %>%
+  left_join(select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
+  group_by(seq_run, plate) %>%
+  mutate(
+    nreads = nreads/sum(nreads),
+    row = factor(row, levels = rev(LETTERS[1:8])),
+    qual = fct_collapse(qual, " " = c("*", "")),
+    "Sample group" = coalesce(
+      str_c(site, year, qual),
+      as.character(sample_type)
+      )
+  ) %>%
+  ggplot(aes(x = col, y = row, fill = nreads, color = `Sample group`)) +
+  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
+  ggnomics::facet_nested(cols = vars(plate), rows = vars(tech, amplicon),
+                         resect = unit(1, "mm"), nest_line = TRUE) +
+  scale_fill_gradient(low = "gray80", high = "black") +
+  coord_equal() +
+  scale_color_discrete(name = "Sample group") +
+  scale_x_continuous(breaks = 1:9) +
+  theme(strip.background = element_blank())
+```
+
+(ref:pcr-concentration) Concentration of PCR products in ng/µL for the two plates of samples.
+
+```{r pcr-concentration, fig.cap = "(ref:pcr-conentration)"}
+tidyxl::xlsx_cells(here("config/Brendan_soil3.xlsx"), sheets = c("BSoil 1 Pico green", "BSoil 2 pico green", "BSoil 1 Picco Green ITS 1 LR5", "BSoil 2 Picco Green its1 LR5")) %>%
+  filter(between(col, 4, 12), between(row, 77, 84), numeric > 0) %>% 
+  mutate(row = LETTERS[row - 76], col = col - 3,
+         sheet = plyr::mapvalues(sheet, c("BSoil 1 Pico green", "BSoil 2 pico green", "BSoil 1 Picco Green ITS 1 LR5", "BSoil 2 Picco Green its1 LR5"), c("short_001", "short_002", "long_001", "long_002"))) %>%
+  select(sheet, row, col, Concentration = numeric) %>%
+  separate(sheet, into = c("amplicon", "plate")) %>%
+  left_join(
+    readd(proto_physeq, cache = cache) %>%
+      sample_data() %>%
+      select(plate, row, column, site, year, qual, sample_type) %>%
+      mutate(row = LETTERS[row]) %>%
+      unique(),
+    by = c("plate", "row", "col" = "column")
+  ) %>%
+  mutate(
+    row = factor(row, levels = rev(LETTERS[1:8])),
+    qual = fct_collapse(qual, " " = c("*", "")),
+    "Sample group" = coalesce(
+      str_c(site, year, qual),
+      as.character(sample_type)
+      )
+  ) %>%
+  ggplot(aes(x = col, y = row, fill = Concentration, color = `Sample group`)) +
+  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
+  ggnomics::facet_nested(cols = vars(plate), rows = vars(amplicon),
+                         resect = unit(1, "mm"), nest_line = TRUE) +
+  scale_fill_gradient(low = "gray80", high = "black") +
+  coord_equal() +
+  scale_color_discrete(name = "Sample group") +
+  scale_x_continuous(breaks = 1:9) +
+  theme(strip.background = element_blank())
+
+```
+
+(ref:dna-conc) DNA concentration in soil extracts.
+
+```{r dna-conc, fig.cap = "(ref:dna-conc)"}
+readxl::read_xls(here("config/Brendan soil samples DNA quantification.xls"),
+                 col_types = c("text", "text", "numeric")) %>%
+  filter(str_detect(Sample, "(T\\d-S\\d+|V\\dP\\d T\\d+)")) %>%
+  rename(year = Year) %>%
+  # Sample name is actually three pieces of information
+  tidyr::extract("Sample", c("site", "x", "qual"),
+                 "(V1P[13]|T[12])[ -][ST](\\d+)(\\D*)$") %>%
+  dplyr::mutate_at("x", as.integer) %>%
+  # Sites were recorded differently in the two years
+  dplyr::mutate(
+    year = factor(year),
+    site = dplyr::case_when(site == "V1P1" ~ "Ang",
+                                        site == "V1P3" ~ "Gan",
+                                        site == "T1" ~ "Gan",
+                                        site == "T2" ~ "Ang",
+                                        TRUE ~ NA_character_)) %>%
+  # For some site/years, coordinates were measured from the edge of
+  # the plot rather than numbered from 1--25 
+  dplyr::group_by(year, site) %>%
+  dplyr::mutate(x = x - if (!any(is.na(x)) && max(x) > 25) 12 else 0) %>%
+  dplyr::ungroup() %>%
+  left_join(
+    readd(proto_physeq, cache = cache) %>%
+      sample_data() %>%
+      select(plate, row, column, site, year, qual, sample_type, x) %>%
+      mutate(row = LETTERS[row]) %>%
+      unique(),
+    by = c("site", "year", "qual", "x")
+  ) %>%
+  mutate(
+    row = factor(row, levels = rev(LETTERS[1:8])),
+    qual = ifelse(qual == "*", "", qual),
+    "Sample group" = coalesce(
+      str_c(site, year, qual),
+      as.character(sample_type)
+      )
+  ) %>%
+  ggplot(aes(x = column, y = row, fill = `Conc. (ng/µl)`, color = `Sample group`)) +
+  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
+  ggnomics::facet_nested(cols = vars(plate),
+                         resect = unit(1, "mm"), nest_line = TRUE) +
+  scale_fill_gradient(low = "gray80", high = "black", na.value = "white") +
+  coord_equal() +
+  scale_color_discrete(name = "Sample group") +
+  scale_x_continuous(breaks = 1:9) +
+  theme(strip.background = element_blank())
+
+```
+
+```{r}
+physeq <- readd(proto_physeq, cache = cache) %>%
+  `sample_data<-`(inset2(
+    sample_data(.),
+    "qual",
+    value = fct_collapse(sample_data(.)$qual, "-" = c("", "*"))
+    )) %>%
+  merge_samples(group = glue::glue_data(
+    sample_data(.),
+    "{year}_{qual}_{tech}_{amplicon}"
+  ))
+
+ranks <- c("root", "kingdom", "phylum", "class", "order", "family", "genus")
+physeq <- readd(taxon_reads, cache = cache) %>%
+  mutate(root = "Root") %>%
+  group_by(label) %>%
+  arrange(amplicon, reference, Algorithm, .by_group = TRUE) %>%
+  summarize_at(ranks, first) %>%
+  left_join(
+    tibble(label = taxa_names(physeq)),
+    .,
+    by = "label"
+  ) %>% {
+    split(select(., -label), .$label)
+  } %>%
+  lapply(unlist) %>%
+  build_tax_table() %>%
+  `tax_table<-`(physeq, .) %>%
+  tax_glom(taxrank = "family", NArm = FALSE)
+```
+
+```{r}
+taxmap <- metacoder::parse_phyloseq(physeq)
+taxmap$data$otu_table <- metacoder::calc_obs_props(taxmap, "otu_table")
+taxmap$data$read_count <- metacoder::calc_taxon_abund(taxmap, "otu_table")
+taxmap$data$read_diff <- metacoder::compare_groups(taxmap, "read_count", cols = c("2016_-_PacBio_long", "2016_X_PacBio_long"), groups = c("Xpedition", "Lifeguard"))
+```
+
+```{r}
+taxmap$data$read_diff <- metacoder::compare_groups(taxmap, "read_count", cols = c("2016_-_PacBio_long", "2016_X_PacBio_long"), groups = c("Xpedition", "Lifeguard"))
+taxa::filter_taxa(
+  taxmap,
+  taxmap$data$read_count$`2016_-_PacBio_long` > 0 |
+    taxmap$data$read_count$`2016_X_PacBio_long` > 0,
+  drop_obs = TRUE,
+  reassign_obs = FALSE
+  ) %>%
+metacoder::heat_tree(
+  node_label = taxon_names,
+  node_color = log2_median_ratio,
+  node_size = rowMeans(.$data$read_count[-1]),
+  node_color_range = metacoder::diverging_palette(),
+  edge_size_interval = c(0, 0.3),
+  output_file = here("temp/buffer.pdf")
+)
+
+```
+
+```{r}
+taxmap$data$read_diff <- metacoder::compare_groups(
+  taxmap,
+  "read_count",
+  cols = c("2015_-_PacBio_long", "2016_X_PacBio_long", "2016_-_PacBio_long"),
+  groups = c("Xpedition 2015", "Xpedition 2016", "Lifeguard 2016")
+)
+taxa::filter_taxa(
+  taxmap,
+  taxmap$data$read_count$`2015_-_PacBio_long` > 0 |
+    taxmap$data$read_count$`2016_X_PacBio_long` > 0 |
+    taxmap$data$read_count$`2016_-_PacBio_long` > 0,
+  drop_obs = TRUE,
+  reassign_obs = FALSE
+  ) %>%
+metacoder::heat_tree_matrix(
+  "read_diff",
+  node_label = taxon_names,
+  node_color = log2_median_ratio,
+  node_size = rowMeans(.$data$read_count[-1]),
+  node_color_range = metacoder::diverging_palette(),
+  edge_size = rowMeans(.$data$read_count[-1]),
+  edge_size_range = c(0, 0.03),
+  edge_size_trans = "linear",
+  node_size_range = c(0, 0.03),
+  node_label_size_range = c(0.005, 0.03),
+  layout = "davidson-harel",
+  initial_layout = "reingold-tilford",
+  output_file = here("temp/compare_long.pdf")
+)
+
+```
+
+```{r}
+taxmap$data$read_diff <- metacoder::compare_groups(
+  taxmap,
+  "read_count",
+  cols = c("2015_-_PacBio_short", "2016_X_PacBio_short", "2016_-_PacBio_short"),
+  groups = c("Xpedition 2015", "Xpedition 2016", "Lifeguard 2016")
+)
+taxa::filter_taxa(
+  taxmap,
+  taxmap$data$read_count$`2015_-_PacBio_short` > 0 |
+    taxmap$data$read_count$`2016_X_PacBio_short` > 0 |
+    taxmap$data$read_count$`2016_-_PacBio_short` > 0,
+  drop_obs = TRUE,
+  reassign_obs = FALSE
+  ) %>%
+metacoder::heat_tree_matrix(
+  "read_diff",
+  node_label = taxon_names,
+  node_color = log2_median_ratio,
+  node_size = rowMeans(.$data$read_count[-1]),
+  node_color_range = metacoder::diverging_palette(),
+  edge_size = rowMeans(.$data$read_count[-1]),
+  edge_size_range = c(0.0002, 0.03),
+  edge_size_trans = "linear",
+  node_size_range = c(0.0002, 0.03),
+  node_label_size_range = c(0.005, 0.03),
+  layout = "davidson-harel",
+  initial_layout = "reingold-tilford",
+  output_file = here("temp/compare_short.pdf")
+)
+
+```
+
+```{r}
+pseq <- readd(proto_physeq, cache = cache) %>%
+  `sample_data<-`(inset2(
+    sample_data(.),
+    "qual",
+    value = fct_collapse(sample_data(.)$qual, "-" = c("", "*"))
+    ))
+
+sample_data(pseq) %>%
+  select(.tech = tech, .amplicon = amplicon, .site = site, .year = year, .qual = qual) %>%
+  filter(.tech != "Ion Torrent", !is.na(.site), !is.na(.year)) %>%
+  unique() %>%
+  rowwise() %>%
+  pmap_dfr(
+    function(.tech, .amplicon, .site, .year, .qual, phyloseq) {
+      selection <- with(sample_data(phyloseq), tech == .tech &
+          amplicon == .amplicon &
+          site == .site &
+          year == .year &
+          qual == .qual)
+      eval(substitute(
+        subset_samples(
+        phyloseq,
+        subset = selection
+      ),
+      list(selection = selection))) %>% 
+        prune_samples(samples = rowSums(otu_table(.)) > 0) %>%
+        prune_taxa(taxa = colSums(otu_table(.)) > 0) %>% {
+        full_join(
+        distance(., "bray") %>% broom::tidy() %>% rename(bray = distance),
+        sample_data(.) %>% rownames_to_column() %$% set_names(x, rowname) %>% dist() %>% broom::tidy()
+      )
+      } %>%
+        mutate(
+          tech = .tech,
+          amplicon = .amplicon,
+          site = .site,
+          year = .year,
+          qual = .qual
+        )
+    },
+    phyloseq = pseq
+  ) %>%
+  group_by(distance, tech, amplicon, site, year, qual) %>%
+  summarize(bray = mean(bray, na.rm = TRUE), n = n()) %>%
+  ggplot(aes(distance, bray, size = n)) +
+  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
+  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
+  # geom_jitter(shape = 46, alpha = 0.5) +
+  # geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(alpha = 0.5) +
+  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
+  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
+  scale_y_continuous(name = "Community dissimilarity") +
+  ggnomics::facet_nested(year + qual + site ~ amplicon)
+```

### 2020-02-01 scripts/drake_plan_light.R

@@ -15,6 +15,8 @@ library(disk.frame)
 
 source(file.path(config$rdir, "variogram.R"))
 source(file.path(config$rdir, "qstats.R"))
+source(file.path(config$rdir, "taxonomy.R"))
+source(file.path(config$rdir, "plate_check.R"))
 
 choosevars <- function(d, g, .data) {
   .data %>%
@@ -301,8 +303,9 @@ plan2 <- drake_plan(
       }
       physeq <- physeq %>%
         phyloseq::subset_samples(sample_type == "Sample") %>%
-        phyloseq::subset_samples(qual == "X" | year == "2015") %>%
-        phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 0 ) %>%
+        phyloseq::subset_samples(buffer == "Xpedition") %>%
+        phyloseq::subset_samples(site == "Gan") %>%
+        phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 100 ) %>%
         phyloseq::prune_samples(samples = phyloseq::sample_data(.)[["tech"]] == tech) %>%
         phyloseq::prune_samples(samples = phyloseq::sample_data(.)[["amplicon"]] == amplicon)
       
@@ -868,6 +871,30 @@ plan2 <- drake_plan(
         reassign_obs = FALSE
       )
       
+      retaxa_table <- tibble::tibble(
+        classification = out$classifications(),
+        id = out$taxon_ids()
+      ) %>%
+        group_by(classification) %>%
+        mutate(newid = first(id)) %>%
+        filter(id != newid)
+      for (n in names(out$data)) {
+        if ("taxon_id" %in% names(out$data[[n]]))
+        out$data[[n]]$taxon_id <- plyr::mapvalues(
+          out$data[[n]]$taxon_id,
+          retaxa_table$id,
+          retaxa_table$newid,
+          FALSE
+        )
+      }
+      for (n in c("asv_table", "read_table", "tax_asv", "tax_read")) {
+        out$data[[n]] <- out$data[[n]] %>%
+          group_by(taxon_id) %>%
+          summarize_all(sum)
+      }
+      out <- out$filter_taxa(!taxon_ids %in% retaxa_table$id, drop_obs = TRUE,
+                             reassign_obs = FALSE, reassign_taxa = FALSE)
+      
       out$data$tax_read$mean = rowMeans(out$data$tax_read[,-1])
[…]

### 2020-02-01 scripts/taxonomy.R

@@ -817,7 +817,7 @@ phylotax_ <- function(tree, taxa, node, ranks, e) {
           rank == r,
           taxon != !!taxon
         ) %>%
-        dplyr::select(-rank, -taxon, -n_tot, -n_diff, -n_reads, -confidence)
+        dplyr::select(-rank, -taxon, -n_tot, -n_diff, -n_reads, -confidence, -n_method, -n_reference)
       newAssign <- tibble::tibble(
         label = tips,
         rank = r,

### 2020-02-02 writing/transect_paper.Rmd
cleo

@@
 venn_ASV %>%
+  filter(!is.na(ASVs)) %>%
   set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
@@
   mutate(
-    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057,
+    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                        sep = "/") %>%
@@
 
-eulerr::venn(deframe(select(vennplot_ASV, set, ASVs))) %>%
+eulerr::venn(deframe(select(vennplot_ASV, set, ASVs)) %>% replace_na(0)) %>%
   plot(
     legend = list(
-      labels = c("PacBio Long", "PacBio Short", "Ion Torrent")
+      labels = c("PacBio Long", "PacBio Short", "Ion Torrent", "Illumina")
     ),
@@
     scale_x_log10() +
-    scale_y_log10(limits = c(1, 1e4)) +
+    scale_y_log10() +
     coord_equal() +
@@
              data = multi_table %>% filter(y == 0, x > 0)) +
-    facet_grid(y_var ~ x_var, switch = "both", space = "free", ) +
+    facet_grid(y_var ~ x_var, switch = "both", space = "free") +
     stat_smooth(method = "loess", formula = y ~ x) +
@@
       data = comparisons,
-      x = 0.5,
-      y = 3.5,
+      x = 0.2,
+      y = 5.5,
       color = "black",
@@
   g <- ggplotGrob(p)
-  g$grobs[g$layout$name %in% c("panel-1-2")] <- NULL
-  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2")),]
+  g$grobs[g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")] <- NULL
+  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")),]
   g
@@
 ```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
-set.seed(4)
+set.seed(2)
 readd(taxdata, cache = cache) %>%
@@
     node_label = replace_na(taxon_names, "unknown"),
-    node_size_range = c(0.005, .035),
+    node_size_range = c(0.002, .035),
     node_size_axis_label = "Read abundance",
@@
     edge_size_axis_label = "ASV count",
-    edge_size_range = c(0.0025, 0.03),
+    edge_size_range = c(0.001, 0.03),
     edge_color = rowMeans(.$data$tax_asv[,-1]),
@@
   xlab("Distance (m)") +
-  facet_grid(metric + amplicon ~ guild, scales = "fixed") +
+  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed") +
   scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
@@
   scale_y_continuous(name = "Community dissimilarity") +
-  facet_grid(metric + amplicon ~ guild, scales = "free")
+  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free")
 
@@
   unnest(variofitST) %>%
-  arrange(metric, amplicon, guild) %>%
-  select(Guild = guild, Amplicon = amplicon, Metric = metric,
+  arrange(metric, amplicon, tech, guild) %>%
+  select(Guild = guild, Amplicon = amplicon, Tech = tech, Metric = metric,
          Range = range, Anisotropy = anis) %>%
@@
   mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
-  select(Guild, Amplicon, Metric, "Space Range" = Range, "Time Range") %>%
+  select(Guild, Amplicon, Tech, Metric, "Space Range" = Range, "Time Range") %>%
   kable(booktabs = TRUE, caption = "(ref:variofit)")
@@
 ```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
-full_dens <- reads_table %>%
+full_dens <- 
+  readd(reads_table, cache = cache) %>%
   dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
@@
 
-full_ecdf <- reads_table %>%
+full_ecdf <- readd(reads_table, cache = cache) %>%
   dplyr::filter(region  %in% c("long", "short")) %>%
@@
   stat_ecdf(geom = "line") +
-  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
+  scale_color_brewer(type = "qual",
+                     labels = c("Short - Ion", "Short - PacBio", "Long - PacBio", "Short - Illumina"),
+                     name = NULL, palette = "Set2") +
   scale_y_continuous(name = "Cumulative read fraction") +
   scale_x_continuous(name = "Length (bp)") +
-  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))
+  theme(legend.position = c(1, 0.02), legend.justification = c(1, 0), legend.background = element_blank())
 
@@
 ```{r regions, fig.cap="(ref:region-asv)"}
-region_table %>%
+readd(region_table, cache = cache) %>%
   dplyr::filter(seq_run == "pb_500") %>%
@@
 
-reads_table %>%
+readd(reads_table, cache = cache) %>%
   dplyr::filter(region %in% baseregions,
@@
 ```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
-its2_dens <- reads_table %>%
+its2_dens <- readd(reads_table, cache = cache) %>%
   dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
@@
 
-its2_ecdf <- reads_table %>%
+its2_ecdf <- readd(reads_table, cache = cache) %>%
   dplyr::filter(region == "ITS2") %>%
@@
 
-set.seed(4)
+set.seed(2)
 taxdata2 %>%
@@
     node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.005, .035),
+    node_size_range = c(0.002, .035),
     node_size_axis_label = "Read abundance",
@@
     edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.0025, 0.03),
+    edge_size_range = c(0.001, 0.03),
     edge_size_trans = "linear",
@@
     node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.01, .035),
+    node_size_range = c(0.002, .035),
     node_size_axis_label = "Read abundance",
@@
     edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.005, 0.03),
+    edge_size_range = c(0.001, 0.03),
     edge_size_trans = "linear",
@@
   filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
-  complete(seq_run, tech, amplicon, site, x, nesting(year, qual)) %>%
+  complete(nesting(seq_run, tech, amplicon), site, x, nesting(year, qual)) %>%
   mutate(
@@
   scale_y_continuous(name = "Community dissimilarity") +
-  ggnomics::facet_nested(year + qual + site ~ amplicon)
-```
+  ggnomics::facet_nested(year + qual + site ~ tech + amplicon)
+```

### 2020-02-02 scripts/qstats.R

@@ -58,6 +58,6 @@ q_stats.character <- function(sreadq, ..., qualityType = "FastqQuality") {
 
 parse_qstat <- function(d) {
   mutate_at(d, "file", basename) %>%
-    tidyr::extract(col = "file", into = c("seq_run", "plate", "well"),
-                   regex = "([pi][bs]_\\d{3})_(\\d{3})-?([A-H]1?\\d)?[rf]?[.].*")
+    tidyr::extract(col = "file", into = c("seq_run", "plate", "well", "read"),
+                   regex = "([piS][bsH][-_]\\d{3,4})_(\\d{3})-?([A-H]1?\\d)?(_R[12])?[rf]?[.].*")
 }

### 2020-02-03 writing/transect_paper.Rmd
Clean up figures after addition of Illumina data.

@@
 
-An additional sample was collected at every sampling location in 2016 using an older batch of lysis buffer for preservation.
-Sequencing results for these samples were found not to be comparable with results from the newer batch, so these samples were not included in spatial analyses.
+An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
+Sequencing results for these samples were found not to be comparable with results preserved using the Xpedition^TM^ lysis buffer (Supplementary figure \@ref(fig:)), so these samples were not included in spatial analyses.
 However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
@@
 After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
+Samples preserved using 
 DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
@@
 
+linesep <- venn_ASV %>%
+  filter(!is.na(ASVs)) %>%
+  select(starts_with("reads")) %>%
+  mutate_all(nchar) %>%
+  as.matrix %>%
+  is_greater_than(0) %>%
+  rowSums %>%
+  rle() %$%
+  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
+  unlist() %>%
+  head(length(.) - 2) %>%
+  c("\\midrule")
+
 venn_ASV %>%
@@
     caption = "(ref:venn-table-asv)",
-    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
+    linesep = linesep,
     align = ("rcrrcrrcrr"),
@@
 
+linesep <- venn_OTU %>%
+  filter(!is.na(OTUs)) %>%
+  select(starts_with("reads")) %>%
+  mutate_all(nchar) %>%
+  as.matrix %>%
+  is_greater_than(0) %>%
+  rowSums %>%
+  rle() %$%
+  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
+  unlist() %>%
+  head(length(.) - 2) %>%
+  c("\\midrule")
+
 venn_OTU %>%
@@
     caption = "(ref:venn-table-otu)",
-    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
+    linesep = linesep,
     align = ("rcrrcrrcrr"),
@@
   rownames_to_column("set") %>%
+  filter(set != "Total") %>%
   mutate(set = map_chr(set,
@@
   rownames_to_column("set") %>%
-  mutate(n = str_count(set, "1")) %>%
-  arrange(n, desc(set)) %>%
+  filter(set != "Total") %>%
   mutate(set = map_chr(set,
@@
 
-eulerr::venn(deframe(select(vennplot_OTU, set, OTUs))) %>%
+eulerr::venn(deframe(select(vennplot_OTU, set, OTUs)) %>% replace_na(0)) %>%
   plot(
@@
     geom_point(alpha = 0.2, shape = 1) +
-    scale_x_log10() +
-    scale_y_log10() +
+    scale_x_log10(
+      breaks = c(1, 1000, 1000000),
+      labels = c("1", "1k", "1M"),
+      minor_breaks = c(10, 100, 10000, 100000),
+      limits = c(1, 1e6)
+    ) +
+    scale_y_log10(
+      breaks = c(1, 1000, 1000000),
+      labels = c("1", "1k", "1M"),
+      minor_breaks = c(10, 100, 10000, 100000),
+      limits = c(1, 1e6)
+    ) +
     coord_equal() +
@@
              data = multi_table %>% filter(y == 0, x > 0)) +
-    facet_grid(y_var ~ x_var, switch = "both", space = "free") +
+    ggnomics::facet_nested(y_amplicon + y_var ~ x_amplicon + x_var, switch = "both", space = "free", nest_line = TRUE) +
     stat_smooth(method = "loess", formula = y ~ x) +
@@
       data = comparisons,
-      x = 0.2,
-      y = 5.5,
+      x = 6,
+      y = 0.7,
       color = "black",
-      hjust = 0,
+      hjust = 1,
       parse = TRUE,
@@
 )
-
-read_comparison <- function(big_table, comparisons, x, y, xlab, ylab) {
-  x_var <- enquo(x)
-  y_var <- enquo(y)
-  big_table %>%
-  filter(!!x_var > 0, !!y_var > 0) %>%
-  ggplot(aes(x = !!x_var, y = !!y_var)) +
-  geom_point(alpha = 0.2, shape = 1) +
-  scale_x_log10() +
-  scale_y_log10(limits = c(1, 1e4)) +
-  coord_equal() +
-  xlab(xlab) +
-  ylab(ylab) +
-  geom_rug(aes(y = !!y_var, x = 1), sides = "l", alpha = 0.2,
-           data = big_table %>% filter(!!x_var == 0, !!y_var > 0)) +
-  geom_rug(aes(y = !!x_var), sides = "b", alpha = 0.2,
-           data = big_table %>% filter(!!y_var == 0, !!x_var > 0)) +
-  stat_smooth(method = "loess", formula = y ~ x) +
-  geom_abline(aes(intercept = mean(log10((!!y_var)[(!!y_var) >= 1])) -
-                    mean(log10((!!x_var)[(!!x_var) >= 1])), slope = 1),
-              linetype = 2L, alpha = 0.5) +
-  geom_text(
-    aes(label = label),
-    data = filter(comparisons, .data$x == quo_name(x_var), .data$y == quo_name(y_var)),
-    x = 1,
-    y = 3.5,
-    color = "black",
-    hjust = 0,
-    parse = TRUE,
-    inherit.aes = FALSE
-  ) +
-  facet_wrap(~type, strip.position = "right", scales = "free")
-}
 ```
@@
   ggnomics::facet_nested(
-    reference ~ type + tech + amplicon,
+    tech + amplicon ~ type + reference,
     resect = unit(1, "mm"),
@@
 set.seed(2)
+theme <- theme_update(panel.border = element_blank())
 readd(taxdata, cache = cache) %>%
@@
   )
+theme_set(theme)
 ```
@@
 set.seed(2)
+theme <- theme_update(panel.border = element_blank())
 taxdata2 %>%
@@
 
-(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (top two rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top row) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).
+(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).
 Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.
@@
     guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
-    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
-    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
+    amplicon = factor(amplicon, levels = c("Short", "Long"))
   ) %>%
@@
   xlab("Distance (m)") +
-  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed") +
-  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
+  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed", nest_line = TRUE) +
+  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
+  theme(strip.background = element_blank())
 ```
@@
   mutate(
-    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
-    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
+    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("All Fungi", "ECM")),
+    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
     amplicon = factor(
       amplicon,
-      levels = c("short", "long"),
-      labels = c("short amplicon", "long amplicon")
+      levels = c("Short", "Long")
     )
@@
   scale_color_discrete(name = "Time lag") +
+  scale_size_continuous(name = "Comparisons") +
   scale_y_continuous(name = "Community dissimilarity") +
-  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free")
+  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free", nest_line = TRUE) +
+  theme(strip.background = element_blank())
 
@@
 
-(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  *a*) Length distribution.  *b*) Empirical cumulative distribution function.
+(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.
+Length distribution (*a*, *c*) and empirical cumulative distribution function (*b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.
 
 ```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
-full_dens <- 
-  readd(reads_table, cache = cache) %>%
-  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
+full_dens <- readd(reads_table, cache = cache) %>%
+  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon)) %>%
   dplyr::filter(region %in% c("long", "short")) %>%
   dplyr::group_by(seq_run) %>%
+  dplyr::mutate_at("amplicon", factor, levels = c("Short", "Long")) %>%
+  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
   dplyr::mutate(
-    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
-      factor(levels = c("Short", "Long")),
     reads = reads / sum(reads)
-  ) %>%
-  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  )
+
+full_dens_short <- full_dens %>%
+  filter(amplicon == "Short") %>%
+  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
   stat_density(bw = 0.5, geom = "line", position = "identity") +
@@
   scale_y_continuous(name = "Density") +
-  scale_x_continuous(name = NULL) +
-  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")
+  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
+  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
+  theme(strip.background = element_blank())
+
+full_dens_long <- full_dens %>%
+  filter(amplicon == "Long") %>%
+  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
+  stat_density(bw = 0.5, geom = "line", position = "identity") +
+  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
+  scale_y_continuous(name = "Density") +
+  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
+  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
+  theme(strip.background = element_blank())
 
@@
   dplyr::filter(region  %in% c("long", "short")) %>%
+  left_join(select(datasets, seq_run, amplicon, tech, machine), by = "seq_run") %>%
+  mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
   dplyr::group_by(seq_run) %>%
-  dplyr::mutate(reads = reads / sum(reads)) %>%
-  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  dplyr::mutate(reads = reads / sum(reads))
+
+full_ecdf_short <- full_ecdf %>%
+  filter(amplicon == "Short") %>%
+  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
   stat_ecdf(geom = "line") +
   scale_color_brewer(type = "qual",
-                     labels = c("Short - Ion", "Short - PacBio", "Long - PacBio", "Short - Illumina"),
                      name = NULL, palette = "Set2") +
-  scale_y_continuous(name = "Cumulative read fraction") +
+  scale_y_continuous(name = "ECDF") +
+  scale_x_continuous(name = NULL) +
+  facet_grid( amplicon ~ ., scales = "free_x") +
+  theme(
+    legend.position = c(1, 0.02),
+    legend.justification = c(1, 0),
+    legend.background = element_blank(),
+    strip.background = element_blank()
+  )
+
+full_ecdf_long <- full_ecdf %>%
+  filter(amplicon == "Long") %>%
+  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
+  stat_ecdf(geom = "line") +
+  scale_color_brewer(type = "qual",
+                     name = NULL, palette = "Set2", direction = -1) +
+  scale_y_continuous(name = "ECDF") +
   scale_x_continuous(name = "Length (bp)") +
-  theme(legend.position = c(1, 0.02), legend.justification = c(1, 0), legend.background = element_blank())
-
-ggpubr::ggarrange(full_dens, full_ecdf, ncol = 1, labels = "auto")
+  facet_grid( amplicon ~ ., scales = "free_x") +
+  theme(
+    legend.position = c(1, 0.02),
+    legend.justification = c(1, 0),
+    legend.background = element_blank(),
+    strip.background = element_blank()
+  )
+
+ggpubr::ggarrange(
+  full_dens_short,
+  full_ecdf_short,
+  full_dens_long,
+  full_ecdf_long,
+  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
+
+)
 ```
@@
 ```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
-its2_dens <- readd(reads_table, cache = cache) %>%
-  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
+its2 <- readd(reads_table, cache = cache) %>%
+  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
+  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
   dplyr::filter(region == "ITS2") %>%
+  arrange(tech, amplicon) %>%
+  mutate(
+    group = paste(tech, "-", amplicon) %>% factor(levels = unique(.))
+  ) %>%
   dplyr::group_by(seq_run) %>%
   dplyr::mutate(
-    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
-      factor(levels = c("Short", "Long")),
     reads = reads / sum(reads)
-  ) %>%
-  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  )
+
+its2_dens <- its2 %>%
+  ggplot(aes(length, weight = reads, color = group, group = group)) +
   stat_density(bw = 0.5, geom = "line", position = "identity") +
@@
   scale_y_continuous(name = "Density") +
-  scale_x_continuous(name = NULL) +
-  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")
-
-its2_ecdf <- readd(reads_table, cache = cache) %>%
-  dplyr::filter(region == "ITS2") %>%
-  dplyr::group_by(seq_run) %>%
-  dplyr::mutate(reads = reads / sum(reads)) %>%
-  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
+  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
+  ggnomics::facet_nested(rows = vars(tech, amplicon), scales = "free_y",
+                         nest_line = TRUE) +
+  theme(strip.background = element_blank())
+
+its2_ecdf <- its2 %>%
+  mutate(tech = "", amplicon = "") %>%
+  ggplot(aes(length, weight = reads, color = group, group = group)) +
   stat_ecdf(geom = "line") +
-  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
-  scale_y_continuous(name = "Cumulative read fraction") +
+  scale_color_brewer(type = "qual", name = NULL, palette = "Set2") +
+  scale_y_continuous(name = "ECDF") +
   scale_x_continuous(name = "ITS2 Length (bp)") +
-  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))
-
-ggpubr::ggarrange(its2_dens, its2_ecdf, ncol = 1, labels = "auto")
+  ggnomics::facet_nested(rows = vars(tech, amplicon), scales = "free_y",
+                         nest_line = TRUE) +
+  theme(
+    strip.background = element_blank(),
+    legend.position = c(0.98, 0.02),
+    legend.justification = c(1, 0)
+  )
+
+ggpubr::ggarrange(its2_dens, its2_ecdf, ncol = 1, labels = "auto", heights = c(4, 2))
+```
+
+(ref:short-its) Kingdom-level taxonomic composition of ITS2 ASVs <= 140 bp.
+Vertical axis shows fraction of total reads in each dataset.
+
+```{r short-its, fig.cap = ""}
+readd(big_seq_table_ITS2, cache = cache) %>%
+  colSums() %>%
+  enframe(name = "ITS2", value = "nreads") %>%
+  mutate(length = nchar(ITS2)) %>%
+  filter(length <= 140) %>%
+  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
+  semi_join(readd(taxon_reads, cache = cache), ., by = c("label" = "ITS2")) %>%
+  # group_by(ITS2) %>%
+  # filter(!"Short" %in% amplicon) %>%
+  taxon_plot(kingdom, y = reads)
 ```
@@
 set.seed(2)
+oldtheme <- theme_update(panel.border = element_blank())
 taxdata2 %>%
@@
   )
+theme_set(oldtheme)
 ```
@@
 set.seed(3)
+oldtheme <- theme_update(panel.border = element_blank())
 taxdata2 %>%
@@
   )
-```
-
-(ref:short-its) Kingdom-level taxonomic composition of ITS2 ASVs <= 140 bp.
-Vertical axis shows fraction of total reads in each dataset.
-
-```{r short-its, fig.cap = ""}
-readd(big_seq_table_ITS2, cache = cache) %>%
-  colSums() %>%
-  enframe(name = "ITS2", value = "nreads") %>%
-  mutate(length = nchar(ITS2)) %>%
-  filter(length <= 140) %>%
-  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(readd(taxon_reads, cache = cache), ., by = c("label" = "ITS2")) %>%
-  # group_by(ITS2) %>%
-  # filter(!"Short" %in% amplicon) %>%
-  taxon_plot(kingdom, y = reads)
-```
-
-(ref:sample-depth) Sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016, after full bioinformatics pipeline.
-Ang 2015 samples are underrepresented, especially in the PacBio long amplicon dataset.
-Dotted horizontal line at proposed cutoff for inclusion in community analysis.
-"2016 X" is using the newer batch of extraction buffer, "2016" is using the older batch.
+theme_set(oldtheme)
+```
+
+(ref:sample-depth) DNA concentrations after extraction (leftmost column) and PCR (second and third columns), and sequencing depth (right three columns) along transects at the two sites Ang and Gan for the years 2015 and 2016.
+Sequencing reads which were filtered out during quality control are colored dark grey.
+Dotted horizontal line at 100 reads indicates cutoff for inclusion in community analysis.
+2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using LifeGuard.
 
 ```{r sample-depth, fig.cap = "(ref:sample-depth)"}
+reads_plot <- readd(qstats_n, cache = cache) %>%
+  mutate(file = str_replace(file, ".*/", "") %>%
+           str_replace("[.].*", "")) %>%
+  tidyr::extract(col = "file", into = c("seq_run", "plate", "row", "col", "read"),
+                 regex = "([[:alpha:]]{2}[-_]\\d+)_(\\d+)-([A-H])(\\d+)(_R[12])?[rf]?") %>%
+  filter(is.na(step), !is.na(seq_run), read != "_R2" | is.na(read)) %>%
+  group_by(seq_run, plate, row, col) %>%
+  summarize(raw_reads = sum(nreads)) %>%
+  ungroup() %>%
+  mutate_at("col", as.integer) %>%
+  left_join(
     readd(proto_physeq, cache = cache) %>% {
@@
     } %>%
-  filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
-  complete(nesting(seq_run, tech, amplicon), site, x, nesting(year, qual)) %>%
+      select(seq_run, plate, row, column, site, year, qual, sample_type, x, buffer, tech, amplicon, reads) %>%
+      mutate(row = LETTERS[row]) %>%
+      unique(),
+    by = c("seq_run", "plate", "row", "col" = "column")
+  ) %>%
+  filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x), !is.na(buffer)) %>%
+  complete(nesting(seq_run, tech, amplicon), site, x, nesting(year, buffer)) %>%
+  pivot_longer(cols = c("reads", "raw_reads"), names_to = "type", values_to = "reads") %>%
   mutate(
-    qual = fct_collapse(qual, " " = c("*", "")),
-    "Sample group" = paste(year, qual)
-  ) %>%
+    type = factor(type, levels = c("raw_reads", "reads")),
+    "Sample group" = paste(year, buffer) %>%
+      ifelse(type == "raw_reads", "Raw", .) %>%
+      fct_relevel("Raw")
+  ) %>%
+  arrange(`Sample group`) %>%
   ggplot(aes(x = x, y = reads, group = `Sample group`, fill = `Sample group`)) +
-  geom_col(position = "dodge") +
+  geom_col(position = "identity") +
   ggnomics::facet_nested(
-    cols = vars(tech, amplicon, site),
-    rows = vars(`Sample group`),
+    cols = vars(tech, amplicon),
+    rows = vars(year, buffer, site),
     nest_line = TRUE,
@@
   ) +
-  scale_y_log10() +
+  scale_fill_manual(
+    values = c(RColorBrewer::brewer.pal(3, "Set2") %>% set_names(c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition")), Raw = "gray20"),
+    breaks = c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition", "Raw"),
+    labels = c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition", "Filtered")) +
+  scale_y_log10(breaks = c(10, 1e3, 1e5),
+                labels = c("10", "1k", "100k"), position = "right") +
   xlab("x (m)") +
@@
     legend.position = "bottom",
-    strip.background = element_blank()
+    strip.background = element_blank(),
+    panel.spacing = unit(1, "mm"),
+    plot.margin = margin(l = 0.5, unit = "mm"),
+    axis.text.x = element_text(angle = 90, vjust = 0.5)
   ) +
@@
 
-```
-
-(ref:plate-depth) Sequencing depth for each well in the sequencing plates (001, 002), after full bioinformatics pipeline.
-Darker shading represents greater sequencing depth, as a fraction of the total number of sequences from the sequencing run.
-The border color for each well where sequences were recovered indicates the sample group, i.e.
-site name (Ang or Gan), sampling year (2015 or 2016), and buffer batch (presence or absence of X; 2016 samples with X are the good buffer); Pos = positive control, Blank = PCR blank.
-Ion Torrent (is_057) samples for the two plates are mixed, shown here as plate 001.
-Note that many of the Ang2015 samples are missing or have extremely low number of reads for the pb_500 (long PacBio) sequencing run.
-
-```{r plate-depth, fig.cap = "(ref:plate-depth)"}
+conc_plot <- readd(proto_physeq, cache = cache) %>% 
+        sample_data(.) %>%
+      select(plate, row, column, site, year, sample_type, x, buffer, amplicon, dna_conc, pcr_conc) %>%
+  arrange(plate, row, column, amplicon) %>% 
+  unique() %>% 
+  mutate_at("dna_conc", as.double) %>%
+  pivot_longer(cols = c("dna_conc", "pcr_conc"), names_to = "step", values_to = "conc", names_pattern = "([[:alpha:]]+)_conc") %>%
+  filter(!is.na(year)) %>%
+  mutate(
+    amplicon = ifelse(step == "dna", "", amplicon),
+    step = factor(step, levels = c("dna", "pcr"), labels = c("DNA", "PCR product")),
+    conc = 1000 * conc,
+    "Sample group" = paste(year, buffer)
+  ) %>%
+  unique() %>%
+  complete(nesting(amplicon, step), site, x, nesting(year, buffer, `Sample group`)) %>%
+  arrange(`Sample group`) %>%
+  ggplot(aes(x = x, y = conc, group = `Sample group`, fill = `Sample group`)) +
+  geom_col() +
+  ggnomics::facet_nested(
+    cols = vars(step, amplicon),
+    rows = vars(year, buffer, site),
+    nest_line = TRUE,
+    resect = unit(1, "mm"),
+    switch = "y"
+  ) +
+  scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Set2") %>% set_names(c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition")), guide = "none") +
+  scale_y_log10(
+    breaks = c(10, 1e3, 1e5),
+    labels = c("0.1", "1", "100"),
+    name = "DNA concentration (ng/µL)"
+  ) +
+  xlab("x (m)") +
+  theme(
+    legend.position = "bottom",
+    strip.background = element_blank(),
+    strip.placement = "outside",
+    panel.spacing = unit(1, "mm"),
+    plot.margin = margin(r = 0.5, unit = "mm"),
+    axis.text.x = element_text(angle = 90, vjust = 0.5)
+  )
+
+ggpubr::ggarrange(conc_plot, reads_plot, ncol = 2, common.legend = TRUE) +
+  theme(panel.spacing = unit(0, "mm"))
+```
+
+```{r plate-loading}
 readd(proto_physeq, cache = cache) %>% {
@@
 } %>%
-  group_by(seq_run, plate) %>%
-  mutate(
-    reads = reads/sum(reads),
-    row = factor(LETTERS[row], levels = rev(LETTERS[1:8])),
-    qual = fct_collapse(qual, " " = c("*", "")),
-    "Sample group" = coalesce(
-      str_c(site, year, qual),
-      as.character(sample_type),
-      "Unused"
-      )
-  ) %>%
-  ggplot(aes(x = column, y = row, fill = reads, color = `Sample group`)) +
-  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
-  ggnomics::facet_nested(cols = vars(plate), rows = vars(tech, amplicon),
-                         resect = unit(1, "mm"), nest_line = TRUE) +
-  scale_color_discrete(name = "Sample group") +
-  scale_fill_gradient(low = "gray80", high = "black") +
-  coord_equal() +
-  scale_x_continuous(breaks = 1:9) +
-  theme(strip.background = element_blank())
-```
-
-(ref:raw-depth) Sequencing depth for each well in the sequencing plates (001, 002), after demultiplexing and primer trimming.
-Darker shading represents greater sequencing depth, as a fraction of the total number of sequences from the sequencing run.
-The border color for each well where sequences were recovered indicates the sample group, i.e.
-site name (Ang or Gan), sampling year (2015 or 2016), and buffer batch (presence or absence of X; 2016 samples with X are the good buffer); Pos = positive control, Blank = PCR blank, NA = unused well.
-Ion Torrent (is_057) samples for the two plates are mixed, shown here as plate 001.
-Note that many of the Ang2015 samples are missing or have extremely low number of reads for the pb_500 (long PacBio) sequencing run.
-
-```{r raw-depth, fig.cap = "(ref:raw-depth)"}
-readd(qstats_n, cache = cache) %>%
-  mutate(file = str_replace(file, ".*/", "") %>%
-           str_replace("[.].*", "")) %>%
-  tidyr::extract(col = "file", into = c("seq_run", "plate", "row", "col"),
-                 regex = "([[:alpha:]]{2}[-_]\\d+)_(\\d+)-([A-H])(\\d+)") %>%
-  filter(is.na(step), !is.na(seq_run)) %>%
-  mutate_at("col", as.integer) %>%
-  left_join(
-    readd(proto_physeq, cache = cache) %>%
-      sample_data() %>%
-      select(plate, row, column, site, year, qual, sample_type) %>%
-      mutate(row = LETTERS[row]) %>%
-      unique(),
-    by = c("plate", "row", "col" = "column")
-  ) %>%
-  left_join(select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
-  group_by(seq_run, plate) %>%
-  mutate(
-    nreads = nreads/sum(nreads),
-    row = factor(row, levels = rev(LETTERS[1:8])),
-    qual = fct_collapse(qual, " " = c("*", "")),
-    "Sample group" = coalesce(
-      str_c(site, year, qual),
-      as.character(sample_type)
-      )
-  ) %>%
-  ggplot(aes(x = col, y = row, fill = nreads, color = `Sample group`)) +
-  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
-  ggnomics::facet_nested(cols = vars(plate), rows = vars(tech, amplicon),
-                         resect = unit(1, "mm"), nest_line = TRUE) +
-  scale_fill_gradient(low = "gray80", high = "black") +
-  coord_equal() +
-  scale_color_discrete(name = "Sample group") +
-  scale_x_continuous(breaks = 1:9) +
-  theme(strip.background = element_blank())
-```
-
-(ref:pcr-concentration) Concentration of PCR products in ng/µL for the two plates of samples.
-
-```{r pcr-concentration, fig.cap = "(ref:pcr-conentration)"}
-tidyxl::xlsx_cells(here("config/Brendan_soil3.xlsx"), sheets = c("BSoil 1 Pico green", "BSoil 2 pico green", "BSoil 1 Picco Green ITS 1 LR5", "BSoil 2 Picco Green its1 LR5")) %>%
-  filter(between(col, 4, 12), between(row, 77, 84), numeric > 0) %>% 
-  mutate(row = LETTERS[row - 76], col = col - 3,
-         sheet = plyr::mapvalues(sheet, c("BSoil 1 Pico green", "BSoil 2 pico green", "BSoil 1 Picco Green ITS 1 LR5", "BSoil 2 Picco Green its1 LR5"), c("short_001", "short_002", "long_001", "long_002"))) %>%
-  select(sheet, row, col, Concentration = numeric) %>%
-  separate(sheet, into = c("amplicon", "plate")) %>%
-  left_join(
-    readd(proto_physeq, cache = cache) %>%
-      sample_data() %>%
-      select(plate, row, column, site, year, qual, sample_type) %>%
-      mutate(row = LETTERS[row]) %>%
-      unique(),
-    by = c("plate", "row", "col" = "column")
-  ) %>%
-  mutate(
-    row = factor(row, levels = rev(LETTERS[1:8])),
-    qual = fct_collapse(qual, " " = c("*", "")),
-    "Sample group" = coalesce(
-      str_c(site, year, qual),
-      as.character(sample_type)
-      )
-  ) %>%
-  ggplot(aes(x = col, y = row, fill = Concentration, color = `Sample group`)) +
-  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
-  ggnomics::facet_nested(cols = vars(plate), rows = vars(amplicon),
-                         resect = unit(1, "mm"), nest_line = TRUE) +
-  scale_fill_gradient(low = "gray80", high = "black") +
-  coord_equal() +
-  scale_color_discrete(name = "Sample group") +
-  scale_x_continuous(breaks = 1:9) +
-  theme(strip.background = element_blank())
-
-```
-
-(ref:dna-conc) DNA concentration in soil extracts.
-
-```{r dna-conc, fig.cap = "(ref:dna-conc)"}
-readxl::read_xls(here("config/Brendan soil samples DNA quantification.xls"),
-                 col_types = c("text", "text", "numeric")) %>%
-  filter(str_detect(Sample, "(T\\d-S\\d+|V\\dP\\d T\\d+)")) %>%
-  rename(year = Year) %>%
-  # Sample name is actually three pieces of information
-  tidyr::extract("Sample", c("site", "x", "qual"),
-                 "(V1P[13]|T[12])[ -][ST](\\d+)(\\D*)$") %>%
-  dplyr::mutate_at("x", as.integer) %>%
-  # Sites were recorded differently in the two years
-  dplyr::mutate(
-    year = factor(year),
-    site = dplyr::case_when(site == "V1P1" ~ "Ang",
-                                        site == "V1P3" ~ "Gan",
-                                        site == "T1" ~ "Gan",
-                                        site == "T2" ~ "Ang",
-                                        TRUE ~ NA_character_)) %>%
-  # For some site/years, coordinates were measured from the edge of
-  # the plot rather than numbered from 1--25 
-  dplyr::group_by(year, site) %>%
-  dplyr::mutate(x = x - if (!any(is.na(x)) && max(x) > 25) 12 else 0) %>%
-  dplyr::ungroup() %>%
-  left_join(
-    readd(proto_physeq, cache = cache) %>%
-      sample_data() %>%
-      select(plate, row, column, site, year, qual, sample_type, x) %>%
-      mutate(row = LETTERS[row]) %>%
-      unique(),
-    by = c("site", "year", "qual", "x")
-  ) %>%
-  mutate(
-    row = factor(row, levels = rev(LETTERS[1:8])),
-    qual = ifelse(qual == "*", "", qual),
-    "Sample group" = coalesce(
-      str_c(site, year, qual),
-      as.character(sample_type)
-      )
-  ) %>%
-  ggplot(aes(x = column, y = row, fill = `Conc. (ng/µl)`, color = `Sample group`)) +
-  geom_tile(size = 0.4, height = 0.9, width = 0.9) +
-  ggnomics::facet_nested(cols = vars(plate),
-                         resect = unit(1, "mm"), nest_line = TRUE) +
-  scale_fill_gradient(low = "gray80", high = "black", na.value = "white") +
-  coord_equal() +
-  scale_color_discrete(name = "Sample group") +
-  scale_x_continuous(breaks = 1:9) +
-  theme(strip.background = element_blank())
-
+  filter(tech != "Ion Torrent", amplicon == "Short", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
+  select(plate, well, tech, reads, year, site, buffer) %>%
+  pivot_wider(names_from = "tech", values_from = "reads") %>%
+  ggplot(aes(x = PacBio, y = Illumina, color = year)) +
+  geom_point()
 ```
@@
     )) %>%
+  subset_samples(!is.na(site)) %>%
   merge_samples(group = glue::glue_data(
     sample_data(.),
-    "{year}_{qual}_{tech}_{amplicon}"
+    "{year}_{buffer}_{tech}_{amplicon}"
   ))
@@
 
-```{r}
+(ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
+
+```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)"}
 taxmap <- metacoder::parse_phyloseq(physeq)
@@
 taxmap$data$read_count <- metacoder::calc_taxon_abund(taxmap, "otu_table")
-taxmap$data$read_diff <- metacoder::compare_groups(taxmap, "read_count", cols = c("2016_-_PacBio_long", "2016_X_PacBio_long"), groups = c("Xpedition", "Lifeguard"))
-```
-
-```{r}
-taxmap$data$read_diff <- metacoder::compare_groups(taxmap, "read_count", cols = c("2016_-_PacBio_long", "2016_X_PacBio_long"), groups = c("Xpedition", "Lifeguard"))
-taxa::filter_taxa(
-  taxmap,
-  taxmap$data$read_count$`2016_-_PacBio_long` > 0 |
-    taxmap$data$read_count$`2016_X_PacBio_long` > 0,
-  drop_obs = TRUE,
-  reassign_obs = FALSE
-  ) %>%
-metacoder::heat_tree(
-  node_label = taxon_names,
-  node_color = log2_median_ratio,
-  node_size = rowMeans(.$data$read_count[-1]),
-  node_color_range = metacoder::diverging_palette(),
-  edge_size_interval = c(0, 0.3),
-  output_file = here("temp/buffer.pdf")
-)
-
-```
-
-```{r}
 taxmap$data$read_diff <- metacoder::compare_groups(
@@
   "read_count",
-  cols = c("2015_-_PacBio_long", "2016_X_PacBio_long", "2016_-_PacBio_long"),
-  groups = c("Xpedition 2015", "Xpedition 2016", "Lifeguard 2016")
-)
+  cols = names(taxmap$data$read_count) %>%
+    keep(endsWith, "Long"),
+  groups = names(taxmap$data$read_count) %>%
+    keep(endsWith, "Long") %>%
+    str_match("(2015|2016)_(Xpedition|LifeGuard)") %>% {
+      paste(.[,3], .[,2])
+    },
+  func = function(abund1, abund2) {
+    list(read_ratio = log10(mean(abund1) / mean(abund2)) %>% ifelse(is.nan(.), 0, .))
+  },
+  combinations = list(
+    c("Xpedition 2016", "Xpedition 2015"),
+    c("Xpedition 2016", "LifeGuard 2016"),
+    c("Xpedition 2015", "LifeGuard 2016")
+  )
+)
+oldtheme <- theme_update(panel.border = element_blank())
 taxa::filter_taxa(
   taxmap,
-  taxmap$data$read_count$`2015_-_PacBio_long` > 0 |
-    taxmap$data$read_count$`2016_X_PacBio_long` > 0 |
-    taxmap$data$read_count$`2016_-_PacBio_long` > 0,
+  taxmap$data$read_count$`2015_Xpedition_PacBio_Long` > 0 |
+    taxmap$data$read_count$`2016_Xpedition_PacBio_Long` > 0 |
+    taxmap$data$read_count$`2016_LifeGuard_PacBio_Long` > 0,
   drop_obs = TRUE,
@@
   node_label = taxon_names,
-  node_color = log2_median_ratio,
+  node_color = read_ratio,
   node_size = rowMeans(.$data$read_count[-1]),
+  node_size_axis_label = "Read fraction",
   node_color_range = metacoder::diverging_palette(),
+  node_color_axis_label = "Log10 Read Ratio",
+  node_color_interval = c(-3, 3),
+  node_color_trans = "linear",
   edge_size = rowMeans(.$data$read_count[-1]),
@@
   edge_size_trans = "linear",
+  edge_size_axis_label = "ASV fraction",
   node_size_range = c(0, 0.03),
@@
 )
-
-```
-
-```{r}
+theme_set(oldtheme)
+
+```
+
+(ref:buffer-compare-short) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
+
+```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
 taxmap$data$read_diff <- metacoder::compare_groups(
@@
   "read_count",
-  cols = c("2015_-_PacBio_short", "2016_X_PacBio_short", "2016_-_PacBio_short"),
-  groups = c("Xpedition 2015", "Xpedition 2016", "Lifeguard 2016")
+  cols = names(taxmap$data$read_count) %>%
+    keep(endsWith, "Short"),
+  groups = names(taxmap$data$read_count) %>%
+    keep(endsWith, "Short") %>%
+    str_match("(2015|2016)_(Xpedition|LifeGuard)") %>% {
+      paste(.[,3], .[,2])
+    },
+  func = function(abund1, abund2) {
+    list(read_ratio = log10(mean(abund1) / mean(abund2)) %>% ifelse(is.nan(.), 0, .))
+  },
+  combinations = list(
+    c("Xpedition 2016", "Xpedition 2015"),
+    c("Xpedition 2016", "LifeGuard 2016"),
+    c("Xpedition 2015", "LifeGuard 2016")
+  )
 )
@@
   taxmap,
-  taxmap$data$read_count$`2015_-_PacBio_short` > 0 |
-    taxmap$data$read_count$`2016_X_PacBio_short` > 0 |
-    taxmap$data$read_count$`2016_-_PacBio_short` > 0,
+  taxmap$data$read_count %>% select(ends_with("Short")) %>% rowSums() > 0,
   drop_obs = TRUE,
@@
   node_label = taxon_names,
-  node_color = log2_median_ratio,
+  node_color = read_ratio,
   node_size = rowMeans(.$data$read_count[-1]),
+  node_size_axis_label = "Read fraction",
   node_color_range = metacoder::diverging_palette(),
+  node_color_axis_label = "Log10 Read Ratio",
+  node_color_interval = c(-3, 3),
+  node_color_trans = "linear",
   edge_size = rowMeans(.$data$read_count[-1]),
-  edge_size_range = c(0.0002, 0.03),
+  edge_size_range = c(0, 0.03),
   edge_size_trans = "linear",
-  node_size_range = c(0.0002, 0.03),
+  edge_size_axis_label = "ASV fraction",
+  node_size_range = c(0, 0.03),
   node_label_size_range = c(0.005, 0.03),
@@
 ```
-
-```{r}
-pseq <- readd(proto_physeq, cache = cache) %>%
-  `sample_data<-`(inset2(
-    sample_data(.),
-    "qual",
-    value = fct_collapse(sample_data(.)$qual, "-" = c("", "*"))
-    ))
-
-sample_data(pseq) %>%
-  select(.tech = tech, .amplicon = amplicon, .site = site, .year = year, .qual = qual) %>%
-  filter(.tech != "Ion Torrent", !is.na(.site), !is.na(.year)) %>%
-  unique() %>%
-  rowwise() %>%
-  pmap_dfr(
-    function(.tech, .amplicon, .site, .year, .qual, phyloseq) {
-      selection <- with(sample_data(phyloseq), tech == .tech &
-          amplicon == .amplicon &
-          site == .site &
-          year == .year &
-          qual == .qual)
-      eval(substitute(
-        subset_samples(
-        phyloseq,
-        subset = selection
-      ),
-      list(selection = selection))) %>% 
-        prune_samples(samples = rowSums(otu_table(.)) > 0) %>%
-        prune_taxa(taxa = colSums(otu_table(.)) > 0) %>% {
-        full_join(
-        distance(., "bray") %>% broom::tidy() %>% rename(bray = distance),
-        sample_data(.) %>% rownames_to_column() %$% set_names(x, rowname) %>% dist() %>% broom::tidy()
-      )
-      } %>%
-        mutate(
-          tech = .tech,
-          amplicon = .amplicon,
-          site = .site,
-          year = .year,
-          qual = .qual
-        )
-    },
-    phyloseq = pseq
-  ) %>%
-  group_by(distance, tech, amplicon, site, year, qual) %>%
-  summarize(bray = mean(bray, na.rm = TRUE), n = n()) %>%
-  ggplot(aes(distance, bray, size = n)) +
-  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
-  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
-  # geom_jitter(shape = 46, alpha = 0.5) +
-  # geom_line(size = 1.5, alpha = 0.5) +
-  geom_point(alpha = 0.5) +
-  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
-  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
-  scale_y_continuous(name = "Community dissimilarity") +
-  ggnomics::facet_nested(year + qual + site ~ tech + amplicon)
-```

### 2020-02-04 writing/transect_paper.Rmd
improve ASV/OTU tables

@@
 
-k_or_M <- function(x, ...) {
+k_or_M <- function(x, ..., .sep = " ", .fun = format) {
   case_when(
-    abs(x) > 1e6 ~ paste(format(x / 1e6, ...), "M"),
-    abs(x) > 1e3 ~ paste(format(x / 1e3, ...), "k"),
-    TRUE ~ paste(format(x, ...), " ")
+    abs(x) > 1e6 ~ paste(.fun(x / 1e6, ...), "M", sep = .sep),
+    abs(x) > 1e3 ~ paste(.fun(x / 1e3, ...), "k", sep = .sep),
+    is.na(x) ~ NA_character_,
+    TRUE ~ paste(.fun(x, ...), " ", sep = .sep)
   )
@@
 After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
-Samples preserved using 
+Samples preserved using LifeGuard were first centrifuged at XX ×g for Y minutes, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
 DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
@@
 
-The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).'
+The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
 Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
@@
   c("\\midrule")
-
+empty_background = if (options("knitr.table.format") == "html") "#777777" else "gray"
 venn_ASV %>%
   filter(!is.na(ASVs)) %>%
+  mutate_all(~ifelse(
+    . == "",
+    kableExtra::cell_spec("      ", background = empty_background, background_as_tile = FALSE), .)) %>%
   set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
@@
     align = ("rcrrcrrcrr"),
-    row.names = FALSE
+    row.names = FALSE,
+    escape = FALSE
   ) %>%
@@
   c("\\midrule")
-
+venn_OTU %>% as.matrix() %>% equals("") %>% which()
 venn_OTU %>%
+  filter(!is.na(OTUs)) %>%
+  mutate_all(~ifelse(. == "", kableExtra::cell_spec("      ", background = empty_background, background_as_tile = FALSE), .)) %>%
   set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
@@
     linesep = linesep,
-    align = ("rcrrcrrcrr"),
-    row.names = FALSE
+    align = ("rcrrcrrcrrcrr"),
+    row.names = FALSE,
+    escape = FALSE
   ) %>%
@@
       )
-
-  )
+  ) %>%
+  kableExtra::landscape()
 ```

### 2020-02-04 scripts/drake_plan_light.R

@@ -33,6 +33,15 @@ choosevars <- function(d, g, .data) {
     )
 }
 
+k_or_M <- function(x, ..., .sep = " ", .function = format) {
+  case_when(
+    abs(x) > 1e6 ~ paste(.function(x / 1e6, ...), "M", sep = .sep),
+    abs(x) > 1e3 ~ paste(.function(x / 1e3, ...), "k", sep = .sep),
+    is.na(x) ~ NA_character_,
+    TRUE ~ paste(.function(x, ...), " ", sep = .sep)
+  )
+}
+
 datasets <- read_csv(config$dataset, col_types = "cccccccicccccicc")
 physeq_meta <-
   tidyr::crossing(
@@ -572,7 +581,7 @@ plan2 <- drake_plan(
     mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
     bind_rows(
       group_by(., seq_run) %>%
-        summarize_if(is.numeric, sum) %>%
+        summarize_if(is.numeric, sum, na.rm = TRUE) %>%
         mutate(set = "Total")
     ) %>%
     mutate_at(
@@ -583,19 +592,23 @@ plan2 <- drake_plan(
       drop0trailing = FALSE
     ) %>%
     mutate_at(
-      c("reads", "ASVs"),
-      format
+      "reads",
+      k_or_M,
+      .sep = "",
+      .function = formatC,
+      format = "fg",
+      digits = 2
     ) %>%
     mutate(
-      ASVs_frac = if_else(grepl(" +0", ASVs), "", ASVs_frac),
-      ASVs = if_else(grepl(" +0", ASVs), "", ASVs),
-      reads_frac = if_else(grepl(" +0", reads), "", reads_frac),
-      reads = if_else(grepl(" +0", reads), "", reads)
+      ASVs_frac = if_else(grepl("^ *[0.]+ *$", ASVs), "", ASVs_frac),
+      ASVs = na_if(ASVs, 0),
+      reads_frac = if_else(grepl("^ *[0.]+ *$", reads), "", reads_frac),
+      reads = if_else(grepl("^ *[0.]+ *$", reads), "", reads)
     ) %>%
     select(set, seq_run, ASVs, ASVs_frac, reads, reads_frac) %>% {
       left_join(
         group_by(., set) %>%
-          summarize(ASVs = as.integer(first(ASVs[ASVs != ""]))),
+          summarize(ASVs = max(ASVs, na.rm = TRUE) %>% ifelse(is.finite(.), ., NA)),
[…]

### 2020-02-04 writing/transect_paper.Rmd
find strict consensus taxonomy for short amplicon sequences

@@
 
-Taxonomic assignments of ASVs which were represented in the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure \@ref(fig:phylotax)).
+Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure \@ref(fig:phylotax)).
 A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
@@
 The result of this process was to give a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade where which was consistently given an assignment, as well as clarify the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.
+This refinement algorithm is referred to as "PHYLOTAX".
+
+ASVs from the short amplicon datasets were refined using only the final strict consensus step, i.e., an assignment at a given rank was accepted if there was no conflict between the different assignment algorithms at greater than 50% confidence.
+This refinement method is referred to as "Consensus".
+Additionally, a hybrid method, "PHYLOTAX+C", was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
 
@@
 
-(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, RDPC, SINTAX, IDTAXA).
+(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, PHYLOTAX+C, Consensus, RDPC, SINTAX, IDTAXA).
 
@@
   mutate_at("reference", as.character) %>%
+  mutate_at("reference", replace_na, "All") %>%
   mutate(
-    reference = map(reference, ~ if (. == "All") c("Unite", "Warcup", "RDP") else .)
+    reference = map2(reference, amplicon, ~ if (.x == "All") c("Unite", "Warcup", if (.y == "Long") "RDP" else NULL) else .)
   ) %>%
@@
   geom_col(position = "dodge") +
+  # scale_fill_brewer(type = "qual", palette = 2) +
   ylab("Fraction of reads assigned") +
@@
         legend.position = "bottom",
+        legend.
         strip.background = element_blank())
@@
     target,
-    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
+    into = c("step", "guild", "metric", "tech", "amplicon", "method", "tdist"),
     sep = "_",
@@
   xlab("Distance (m)") +
-  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed", nest_line = TRUE) +
+  ggnomics::facet_nested(metric + tech + amplicon + method ~ guild, scales = "fixed", nest_line = TRUE) +
   scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
@@
   filter(step == "variofitST") %>%
-  select(variofitST, variogST, guild, metric, tech, amplicon) %>%
-  mutate_at(c("guild", "metric", "tech", "amplicon"), str_replace_all, "\"", "") %>%
+  select(variofitST, variogST, guild, metric, tech, amplicon, algorithm) %>%
+  mutate_at(c("guild", "metric", "tech", "amplicon", "algorithm"), str_replace_all, "\"", "") %>%
   mutate(
@@
       levels = c("Short", "Long")
+    ),
+    algorithm = factor(
+      algorithm,
+      levels = c("Consensus", "PHYLOTAX", "PHYLOTAX+Cons"),
+      labels = c("Cons", "PHYLO", "PHYLO+C")
     )
@@
   # mutate(bin = cut(dist, c(1:13 - 0.5, 14.5, 16.5, 19.5, 24.5))) %>%
-  group_by(guild, metric, tech, amplicon, dist, timelag) %>%
+  group_by(guild, metric, tech, amplicon, algorithm, dist, timelag) %>%
   summarize(
@@
 variog_confint <- variog_points %>%
-  group_by(guild, metric, tech, amplicon, dist, timelag) %>%
+  group_by(guild, metric, tech, amplicon, algorithm, dist, timelag) %>%
   summarize(
@@
   scale_y_continuous(name = "Community dissimilarity") +
-  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free", nest_line = TRUE) +
+  ggnomics::facet_nested(metric + tech + amplicon + algorithm ~ guild, scales = "free", nest_line = TRUE) +
   theme(strip.background = element_blank())
@@
   arrange(metric, amplicon, tech, guild) %>%
-  select(Guild = guild, Amplicon = amplicon, Tech = tech, Metric = metric,
-         Range = range, Anisotropy = anis) %>%
+  select(Amplicon = amplicon, Tech = tech, Guild = guild, Algorithm = algorithm,
+         Metric = metric, Range = range, Anisotropy = anis) %>%
   mutate(Range = 3*Range,
@@
   mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
-  select(Guild, Amplicon, Tech, Metric, "Space Range" = Range, "Time Range") %>%
+  select(Amplicon, Tech, Guild, Algorithm, Metric, "Space Range" = Range, "Time Range") %>%
   kable(booktabs = TRUE, caption = "(ref:variofit)")
@@
 
-```{r short-its, fig.cap = ""}
+```{r short-its, fig.cap = "(ref:short-its)"}
 readd(big_seq_table_ITS2, cache = cache) %>%

### 2020-02-04 scripts/drake_plan_light.R

@@ -43,6 +43,23 @@ k_or_M <- function(x, ..., .sep = " ", .function = format) {
 }
 
 datasets <- read_csv(config$dataset, col_types = "cccccccicccccicc")
+
+# combo_meta <- tibble(
+#   group = c("fungi", "euk"),
+#   taxname = "short",
+#   reftree = c("fungi_tree_decipher_unconst_long", "tree_decipher_unconst_long"),
+#   conf_taxon = glue::glue("conf_taxon_short_{group}"),
+#   phylotaxon = glue::glue("phylotaxon_unconst_long_{group}")
+# ) %>%
+#   mutate_at(c("reftree", "conf_taxon", "phylotaxon"), syms)
+
+guilds_meta <- tibble(
+  consensus_taxa = c("phylotaxon_decipher_unconst_long_fungi", "conf_taxon_short_fungi", "best_taxon_short_fungi"),
+  amplicon = c("Long", "Short", "Short"),
+  algorithm = c("PHYLOTAX", "Consensus", "PHYLOTAX+Cons")
+) %>%
+  mutate_at("consensus_taxa", syms)
+
 physeq_meta <-
   tidyr::crossing(
     dplyr::select(datasets, "seq_run", "tech", "dataset", "amplicon"),
@@ -51,7 +68,14 @@ physeq_meta <-
   dplyr::filter(
     seq_run != "is_057", # couldn't be demultiplexed
   ) %>%
-  dplyr::select(-dataset)
+  dplyr::select(-dataset) %>%
+  left_join(select(guilds_meta, amplicon, algorithm), by = "amplicon") %>%
+  mutate(
+    ecm = glue::glue("ecm_{algorithm}"),
+    ecm2 = glue::glue("ecm2_{algorithm}"),
+    ecm3 = glue::glue("ecm3_{algorithm}")
+  ) %>%
+  mutate_at(vars(starts_with("ecm")), compose(syms, make.names))
 
 plan2 <- drake_plan(
   # targets which are imported from first plan
@@ -113,22 +137,31 @@ plan2 <- drake_plan(
   
   tree_decipher_LSU = target(
     raxml_decipher_LSU$bipartitions,
-    transform = map(outname = "decipher_LSU", group = "euk", .tag_in = step, .tag_out = c(euktree, tree))
+    transform = map(
+      treename = "decipher_LSU",
+      group = "euk",
+      .tag_in = step, .tag_out = c(euktree, tree), .id = FALSE)
   ),
   
[…]

### 2020-02-05 writing/transect_paper.Rmd
Write positive control sequences to a file
modify results on read counts correlations to include Illumina.

@@
 *short*, *ITS*, *LSU*, *long*: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.
-Short amplicons were mapped to full ITS, LSU, and long amplicon ASVs when
+Short amplicons were mapped to full ITS, LSU, and long amplicon ASVs when they had a perfect match in the ITS2 region.
 
@@
 
+readd(allseqs, cache = cache) %>% filter(hash %in% taxa_names(agaricus_reads)) %$%
+  set_names(short, hash) %>%
+  discard(is.na) %>%
+  Biostrings::RNAStringSet() %>%
+  Biostrings::writeXStringSet(here("temp/agaricus.fasta"))
+
 agaricus_order <- readd(proto_physeq, cache = cache) %>%
@@
 (ref:venn-table-asv) Correspondences between ASVs found by different sequencing strategies.
-Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets , including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
 
@@
 (ref:venn-table-otu) Correspondences between 97% OTUs found by different sequencing strategies.
+Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTU frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
 
@@
       gsub("(^/|/$)", "", .),
-    label = paste(ASVs, reads_frac, sep = "\n")
+    label = str_c(ASVs, reads_frac, sep = "\n") %>%
+      replace_na("-")
   )
@@
     quantities = vennplot_ASV$label,
-    main = "ASV"
+    main = "ASV",
+    gp = grid::gpar(cex = 0.8)
   )
@@
       gsub("(^/|/$)", "", .),
-    label = paste(OTUs, reads_frac, sep = "\n")
+    label = str_c(OTUs, reads_frac, sep = "\n") %>%
+      replace_na("-")
   )
@@
     quantities = vennplot_OTU$label,
-    main = "OTU"
+    main = "OTU",
+    gp = grid::gpar(cex = 0.6)
   )
@@
 compR2 <- function(x, y, type) {
+  xamp <- plyr::mapvalues(
+    x,
+    datasets$seq_run,
+    datasets$amplicon,
+    FALSE
+  )
+  yamp <- plyr::mapvalues(
+    y,
+    datasets$seq_run,
+    datasets$amplicon,
+    FALSE
+  )
   x <- plyr::mapvalues(
@@
       datasets$tech,
-      datasets$machine,
-      "-",
-      datasets$amplicon),
+      datasets$machine),
     FALSE
@@
       datasets$tech,
-      datasets$machine,
-      "-",
-      datasets$amplicon),
+      datasets$machine),
     FALSE
   )
-  format(filter(comparisons, x == x_var, y == y_var, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
+  formatC(
+    filter(
+      comparisons,
+      x == x_var,
+      xamp == x_amplicon,
+      y == y_var,
+      yamp == y_amplicon,
+      type == !!type
+    )$r.squared,
+    digits = 2,
+    format = "f")
 }
@@
 The number of ASVs and OTUs shared between the different sequencing strategies are shown in Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in figure \@ref(fig:venn).
-ASVs shared between all datasets represented `r text_list(percent(venn_ASV[7, c("frac_pb_500", "frac_pb_483", "frac_is_057")]))` of the reads for the long and short PacBio datasets and Ion Torrent dataset, respectively.
-These fractions increased to `r text_list(percent(venn_OTU[7, c("frac_pb_500", "frac_pb_483", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
-In particular, >99% and `r percent(sum(as.numeric(venn_OTU[c("011", "111"), "frac_is_057"])))` of reads in the PacBio and Ion Torrent short-amplicon datasets belonged to OTUs shared between the two datasets.
-In contrast, `r percent(venn_ASV["100", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(venn_OTU["100", "frac_pb_500"])` after OTU clustering.
+ASVs shared between all datasets represented `r text_list(percent(venn_ASV["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
+These fractions increased to `r text_list(percent(venn_OTU["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+In particular, `r text_list(percent(colSums(mutate_all(venn_OTU[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric))))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
+In contrast, `r percent(venn_ASV["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(venn_OTU["1000", "frac_pb_500"])` after OTU clustering.
 
 Figure \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
-In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = `r compR2("is_057", "pb_500", "ASV")` \text{ to } `r compR2("is_057", "pb_483", "OTU")`$).
-Comparisons including the Ion Torrent dataset all display a "hook" at low read counts, which is probably due to statistical sampling error when comparing read counts between datasets with very different sequencing depth.
-The strongest correlation was for the short amplicon PacBio and Ion Torrent datasets after OTU clustering ($R^2 = `r compR2("is_057", "pb_483", "OTU")`$), which were sequenced from the same amplicon library.
+In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = compR2("SH-2257", "is_057", "OTU")`$).
+Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
+(`r compR2("pb_483", "SH-2257", "ASV")` to `r compR2("pb_483", "SH-2257", "OTU")`,
+`r compR2("pb_483", "is_057", "ASV")` to `r compR2("pb_483", "is_057", "OTU")`, and
+`r compR2("SH-2257", "is_057", "ASV")` to `r compR2("SH-2257", "is_057", "OTU")`,
+for PacBio vs. Illumina, PacBio vs. Ion Torrent, and Illumina vs. Ion Torrent, respectively),
+but this was not, or only weakly, the case for correlations between the long amplicon library and short amplicon library
+(`r compR2("pb_500", "pb_483", "ASV")` to `r compR2("pb_500", "pb_483", "OTU")`,
+`r compR2("pb_500", "SH-2257", "ASV")` to `r compR2("pb_500", "SH-2257", "OTU")`, and
+`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs PacBio, Illumina, and Ion Torrent short reads, respectively).
 

### 2020-02-06 writing/transect_paper.Rmd
better variogram fits

@@
 percent <- function(x, ...) {
-  paste0(format(as.numeric(x) * 100), "%")
+  paste0(formatC(as.numeric(x) * 100, ...), "%")
 }
@@
 venn_ASV %>%
+  rownames_to_column(" ") %>%
   filter(!is.na(ASVs)) %>%
@@
     kableExtra::cell_spec("  ", background = empty_background, background_as_tile = FALSE), .)) %>%
+  mutate_at(" ", str_replace, "[01]+", "") %>%
   set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
@@
     header = c(
-      " " = 1,
+      " " = 2,
       names(venn_ASV)[-(1)] %>% 
@@
     header = c(
-      " " = 1,
+      " " = 2,
        names(venn_ASV)[-(1)] %>% 
@@
 venn_OTU %>%
+  rownames_to_column(" ") %>%
   filter(!is.na(OTUs)) %>%
-  mutate_all(~ifelse(. == "", kableExtra::cell_spec("      ", background = empty_background, background_as_tile = FALSE), .)) %>%
+  mutate_all(~ifelse(
+    . == "",
+    kableExtra::cell_spec(
+      "  ",
+      background = empty_background,
+      background_as_tile = FALSE
+    ),
+    .)) %>%
+  mutate_at(" ", str_replace, "[01]+", "") %>%
   set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
@@
     header = c(
-      " " = 1,
+      " " = 2,
       names(venn_OTU)[-1] %>% 
@@
     header = c(
-     " " = 1,
+     " " = 2,
       names(venn_OTU)[-1] %>% 
@@
 
-(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, PHYLOTAX+C, Consensus, RDPC, SINTAX, IDTAXA).
+(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
+Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
+They are plotted in each column to compare with results for the individual databases.
 
@@
         legend.position = "bottom",
-        legend.
         strip.background = element_blank())
-```
-
-Taxonomic composition of the sequenced soil community is summarized in Figure \@ref(fig:taxa).
-
-  - what fraction assigned at each level (or representative ranks; phylum, (order), genus)
-  - what fraction are fungi
-  - comparison of different region/database and algorithm
-  - breakdown of groups
-  - fraction ECM
-
-(ref:taxa-kingdom) Taxonomic assignment of ASVs to kingdom by different reference databases and algorithms.
-"unidentified" denotes ASVs which could not be identified to kingdom level.
-"other" denotes kingdoms represented by less than 1% of reads in all datasets.
-The Warcup database did not include non-Fungi, so for that database, ASVs were considered to be identified to Fungi if they were identified to any phylum.
-Height of each bar represents the fraction of reads represented by that bar.
-
-```{r taxon-plot}
-taxon_plot <- function(.data, rank, ..., y = reads, cutoff = NULL) {
-  rank <- enquo(rank)
-  y <- enquo(y)
-  ranks <- c("kingdom", "phylum", "class", "order", "family", "genus")
-  .data <- .data %>%
-    group_by(reference, Algorithm, tech, amplicon) %>%
-    mutate(ASVs = n()) %>%
-    ungroup() %>%
-    filter(...) %>%
-    arrange_at(ranks) %>%
-    mutate_at(
-      ranks,
-      ~ factor(
-        .,
-        levels = c(NA, "other", discard(unique(.), is.na)),
-        exclude = NULL
-      )
-    ) %>%
-    group_by(reference, Algorithm, tech, amplicon, !!rank) %>%
-    summarize(reads = sum(reads), ASVs = n()/max(ASVs)) %>%
-    ungroup()
-
-  if (!is.null(cutoff)) {
-    .data <- group_by(.data, !!rank) %>%
-      group_map(
-
-        ~ if (all(pull(.x, !!y) < cutoff)) {
-          mutate(.x, !!rank := factor("other", levels = levels(!!rank)))
-        } else {
-          .x
-        },
-        keep = TRUE
-      ) %>%
-      bind_rows() %>%
-      group_by(reference, Algorithm, !!rank, tech, amplicon) %>%
-      summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
-      ungroup()
-  }
-
-  rank_label <- as_label(rank)
-
-  .data <- mutate(.data, !!rank := fct_drop(!!rank))
-  vals <- levels(pull(.data, !!rank))
-  # if ("other" %in% vals) vals <- c("other", vals) %>% magrittr::extract(!duplicated(.))
-  # if (any(is.na(vals))) vals <- c(NA, vals) %>% magrittr::extract(!duplicated(.))
-
-  if (rank_label == str_to_lower(rank_label)) rank_label <- str_to_title(rank_label)
-  y_label <- as_label(y)
-  if (y_label == str_to_lower(y_label)) y_label <- str_to_title(y_label)
-  ggplot(.data, aes(x = Algorithm, y = !!y, fill = !!rank)) +
-    geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
-    ggnomics::facet_nested(
-      cols = vars(tech, amplicon, reference),
-      scales = "free_x",
-      space = "free_x",
-      nest_line = TRUE,
-      bleed = FALSE,
-      resect = unit(1, "mm")
-      ) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
-          strip.background = element_blank(),
-          panel.spacing = unit(3, "pt")) +
-    scale_fill_discrete(
-      breaks = vals,
-      labels = replace_na(as.character(vals), "unidentified"),
-      name = rank_label
-    ) +
-    ylab(paste("Fraction of", y_label))
-
-}
-
-```
-
-```{r taxa-kingdom, fig.cap = "(ref:taxa-kingdom)", include = FALSE}
-loadd(taxon_reads, cache = cache)
-taxon_plot(taxon_reads, kingdom, cutoff = 0.01)
-```
-
-(ref:ASV-kingdom) Taxonomic assignment of ASVs to kingdom by different reference databases and algorithms.
-"unidentified" denotes ASVs which could not be identified to kingdom level.
-"other" denotes kingdoms represented by less than 1% of ASVs in all datasets.
-The Warcup database did not include non-Fungi, so ASVs were considered to be identified to Fungi if they were identified to any phylum.
-Height of each bar represents the fraction of total ASVs represented by that bar.
-
-```{r ASV-kingdom, fig.cap = "(ref:ASV-kingdom)", include = FALSE}
-taxon_plot(taxon_reads, kingdom, y = ASVs, cutoff = 0.01)
-```
-
-(ref:taxa-phylum) Taxonomic assignment of ASVs to fungal phylum.
-Phyla representing less than 1% of total reads are grouped as "other".
-"unidentified" denotes ASVs which were identified as fungi, but could not be identified to phylum level.
-Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
-Height of each bar represents the fraction of reads represented by that bar.
-
-```{r taxa-phylum, fig.cap = "(ref:taxa-phylum)", include = FALSE}
-taxon_plot(taxon_reads, rank = phylum, cutoff = 0.01, kingdom == "Fungi")
-```
-
-(ref:asco-class) Taxonomic assignment of ASVs to ascomycete order.
-Orders representing less than 1% of total reads are grouped as "other".
-"unidentified" denotes ASVs which were identified as Ascomycota, but could not be identified to order level.
-Reads which were not identified to phylum, or which were identified as a non-Ascomycota phylum, are not included.
-Height of each bar represents the fraction of total reads represented by that bar.
-
-```{r asco-class, fig.cap = "(ref:asco-class)", include = FALSE}
-taxon_plot(taxon_reads, order, phylum == "Ascomycota", cutoff = 0.01)
-```
-
-(ref:basid-family) Taxonomic assignment of ASVs to basidiomycete family.
-Families representing less than 1% of total reads are grouped as "other".
-"unidentified" denotes ASVs which were identified as Basidiomycota, but could not be identified to family level.
-Reads which were not identified to phylum, or which were identified as a non-Basidiomycota phylum, are not included.
-Height of each bar represents the fraction of total reads represented by that bar.
-
-```{r basid-family, fig.cap = "(ref:basid-family)", include = FALSE}
-taxon_plot(taxon_reads, family, phylum == "Basidiomycota", cutoff = 0.01)
-```
-
-(ref:ecm) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic identification by different reference databases and algorithms.
-Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
-"unidentified" denotes ASVs which were identified as Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
-Height of each bar represents the fraction of reads represented by that bar.
-
-```{r ECM, fig.cap = "(ref:ecm)"}
-  readd(taxon_reads, cache = cache) %>%
-  taxon_plot(ECM, kingdom == "Fungi")
-```
-
-(ref:ecm-families) Taxonomic composition of
-
-```{r ECM-families, fig.cap = "(ref:ecm-families)", include = FALSE}
-  readd(taxon_reads, cache = cache) %>%
-  taxon_plot(
-    family,
-    !is.na(as.character(ECM)),
-    ECM != "non-ECM",
-    kingdom == "Fungi",
-    cutoff = 0.01
-  )
-```
-
-```{r taxdata}
-loadd(taxdata, cache = cache)
 ```
@@
 ```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
+fungi_read <- readd(taxdata, cache = cache) %>%
+  taxa::filter_taxa(
+    taxon_names == "Fungi",
+    reassign_obs = FALSE,
+    reassign_taxa = FALSE
+    ) %>%
+  extract2("data") %>%
+  extract2("tax_read") %>%
+  select(ends_with("PHYLOTAX")) %>%
+  set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
+  unlist()
+
+fungi_asv <- readd(taxdata, cache = cache) %>%
+  taxa::filter_taxa(
+    taxon_names == "Fungi",
+    reassign_obs = FALSE,
+    reassign_taxa = FALSE
+    ) %>%
+  extract2("data") %>%
+  extract2("tax_asv") %>%
+  select(ends_with("PHYLOTAX")) %>%
+  set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
+  unlist()
+
+ecm_counts <- readd(taxon_reads, cache = cache) %>%
+  filter(Algorithm == "PHYLOTAX") %>%
+  group_by(seq_run) %>%
+  mutate(ASVs = 1/n()) %>%
+  filter(ECM %in% c("Possible ECM", "Probable ECM", "Highly Probable ECM")) %>%
+  summarize_at(c("ASVs", "reads"), sum)
+
+ecm_read <- select(ecm_counts, seq_run, reads) %>% deframe()
+ecm_asv <- select(ecm_counts, seq_run, ASVs) %>% deframe()
+
 set.seed(2)
@@
 
+Taxonomic composition of the sequenced soil community is summarized in Figure \@ref(fig:taxa).
+According to the PHYLOTAX assignments, Fungi represented `r percent(fungi_read["PacBio_Long"], format = "f", digits = 1)` of the reads and `r percent(fungi_asv["PacBio_Long"], format = "f", digits = 1)` of the ASVs in the long amplicon library, compared to `r paste(percent(range(fungi_read[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads and `r paste(percent(range(fungi_asv[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs in the short amplicon library.
+Fungi categorized as ECM made up `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads and `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs in the long amplicon library, and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs in the short amplicon library.
+Assignment to kingdom Fungi and ECM status for different amplicons, sequencing technologies, reference databases, and algorithms are shown in Supplementary Figure \@ref(ref:ecm).
+
+```{r taxon-plot}
+taxon_plot <- function(.data, rank, ..., y = reads, cutoff = NULL) {
+  rank <- enquo(rank)
+  y <- enquo(y)
+  ranks <- c("kingdom", "phylum", "class", "order", "family", "genus")
+  .data <- .data %>%
+    group_by(reference, Algorithm, tech, amplicon) %>%
+    mutate(ASVs = n()) %>%
+    ungroup() %>%
+    filter(...) %>%
+    arrange_at(ranks) %>%
+    mutate_at(
+      ranks,
+      ~ factor(
+        .,
+        levels = c(NA, "other", discard(unique(.), is.na)),
+        exclude = NULL
+      )
+    ) %>%
+    group_by(reference, Algorithm, tech, amplicon, !!rank) %>%
+    summarize(reads = sum(reads), ASVs = n()/max(ASVs)) %>%
+    ungroup()
+
+  if (!is.null(cutoff)) {
+    .data <- group_by(.data, !!rank) %>%
+      group_map(
+
+        ~ if (all(pull(.x, !!y) < cutoff)) {
+          mutate(.x, !!rank := factor("other", levels = levels(!!rank)))
+        } else {
+          .x
+        },
+        keep = TRUE
+      ) %>%
+      bind_rows() %>%
+      group_by(reference, Algorithm, !!rank, tech, amplicon) %>%
+      summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
+      ungroup()
+  }
+
+  rank_label <- as_label(rank)
+
+  .data <- mutate(.data, !!rank := fct_drop(!!rank))
+  vals <- levels(pull(.data, !!rank))
+  # if ("other" %in% vals) vals <- c("other", vals) %>% magrittr::extract(!duplicated(.))
+  # if (any(is.na(vals))) vals <- c(NA, vals) %>% magrittr::extract(!duplicated(.))
+
+  if (rank_label == str_to_lower(rank_label)) rank_label <- str_to_title(rank_label)
+  y_label <- as_label(y)
+  if (y_label == str_to_lower(y_label)) y_label <- str_to_title(y_label)
+  .data <- .data %>%
+    mutate(tech = fct_relabel(tech, ~ paste(., plyr::mapvalues(., datasets$tech, datasets$machine, FALSE))
+    ))
+  ggplot(.data, aes(x = Algorithm, y = !!y, fill = !!rank)) +
+    geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
+    ggnomics::facet_nested(
+      cols = vars(tech, amplicon, reference),
+      scales = "free_x",
+      space = "free_x",
+      nest_line = TRUE,
+      bleed = FALSE,
+      resect = unit(1, "mm")
+      ) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
+          strip.background = element_blank(),
+          panel.spacing = unit(3, "pt")) +
+    scale_fill_discrete(
+      breaks = vals,
+      labels = replace_na(as.character(vals), "unidentified"),
+      name = rank_label
+    ) +
+    ylab(paste("Fraction of", y_label))
+
+}
+
+```
+
 (ref:heattree-db-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
@@
   ) %>%
+  filter(startsWith(method, "PHYLOTAX")) %>%
   mutate(
@@
   # geom_line(aes(y = exp(predict(bc_fit)))) +
-  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
+  scale_y_continuous(name = "Mantel correlation") +
   scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
@@
   xlab("Distance (m)") +
-  ggnomics::facet_nested(metric + tech + amplicon + method ~ guild, scales = "fixed", nest_line = TRUE) +
+  ggnomics::facet_nested(amplicon + tech + metric ~ guild, scales = "fixed", nest_line = TRUE) +
   scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
@@
 
-(ref:spatio) Spatiotemporal variogram fits for community dissimilarities.
-The thick lines represent the empirical variogram (mean of community distances at each spatial distance) for samples taken at the same time (red) or separated by one year (blue).
-Thin lines are an exponential fit to the variograms.
+(ref:spatio) Relationship between spatiotemporal distance and community dissimilarities.
+Circles represent community data from short (top two rows) and long (bottom two rows) amplicon libraries, sequenced by Illumina MiSeq (top row) or PacBio RS II (bottom three rows).
+Community dissimilarities are calculated using the Bray-Curtis dissimilarity for all datasets (top three rows) and using the weighted UNIFRAC dissimilarity for the long amplicon library, for which a phylogenetic tree could be constructed (bottom row).
+The left column represents the full fungal community, and the right column only sequences identified as ECM.
+The color of each circle represents the time lag between samples being compared (0 or 1 year), and the size represents the number of comparisons for that spatial distance and time lag.
+Lines are the best-fit lines for an exponential decay to max model.
+The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0 year timelag) or spatiotemporal (for the 1 year time lag) distance.
 
@@
 variog_data <- light_plan %>%
-  filter(step == "variofitST") %>%
-  select(variofitST, variogST, guild, metric, tech, amplicon, algorithm) %>%
+  filter(step == "correlog") %>%
+  select(correlog, timelag, guild, metric, tech, amplicon, algorithm) %>%
+  mutate_at("correlog", lapply, readd, character_only = TRUE, cache = cache) %>%
+  mutate_at("correlog", map, "mantel.res") %>%
+  mutate_at("correlog", map_lgl, ~ min(.x[,"Pr(corrected)"], na.rm = TRUE) < 0.05) %>%
+  pivot_wider(names_from = "timelag", values_from = "correlog", names_prefix = "mantel_") %>%
+  left_join(
+    light_plan %>%
+      filter(step == "variofit2") %>%
+      select(variofit2, guild, metric, tech, amplicon, algorithm),
+    by = c("guild", "metric", "tech", "amplicon", "algorithm")
+    ) %>%
+  left_join(
+    light_plan %>%
+      filter(step == "variofitST2") %>%
+      select(variofitST2, variogST, guild, metric, tech, amplicon, algorithm),
+    by = c("guild", "metric", "tech", "amplicon", "algorithm")
+    ) %>%
+  mutate(
+    variofit = ifelse(mantel_1, variofitST2, variofit2)
+    ) %>%
+  mutate_at(c("variofit", "variogST", "variofitST2", "variofit2"), map, readd, character_only = TRUE, cache = cache) %>%
+  select(mantel_0, mantel_1, variofitST2, variofit2, variofit, variogST, guild, metric, tech, amplicon, algorithm) %>%
   mutate_at(c("guild", "metric", "tech", "amplicon", "algorithm"), str_replace_all, "\"", "") %>%
@@
       levels = c("Consensus", "PHYLOTAX", "PHYLOTAX+Cons"),
-      labels = c("Cons", "PHYLO", "PHYLO+C")
-    )
-  ) %>%
-  mutate_at(
-    c("variogST", "variofitST"),
-    lapply,
-    readd,
-    character_only = TRUE,
-    cache = cache)
+      labels = c("Cons", "PHYLO", "PHYLO")
+    )
+  )
 
 variog_points <- unnest(variog_data, c("variogST")) %>%
+  mutate_at("timelag", replace_na, "0") %>%
   filter(dist <= 25) %>%
@@
 variog_fit <- variog_data %>%
-  mutate_at("variofitST", lapply,
-            gstat:::vgmMetric,
-            expand_grid(spacelag = seq(1, 25, 0.5), timelag = 0:1)
-            ) %>%
-  unnest(variofitST) %>%
-  rename(dist = spacelag) %>%
+  mutate_at("variofit", map,
+            ~ mutate(.y, gamma = predict(.x, newdata = .y)),
+            expand_grid(dist = seq(1, 25, 0.5), timelag = 0:1)
+            ) %>%
+  unnest(variofit) %>%
+  filter(mantel_1 | timelag == "0", mantel_0) %>%
   mutate_at("timelag", factor)
@@
 variog_empir %>%
+  filter(algorithm == "PHYLO") %>%
   ggplot(aes(dist, gamma, color = timelag, size = n)) +
+  ggnomics::facet_nested(amplicon + tech + metric ~ guild, scales = "free", nest_line = TRUE, space = "free_y") +
   # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
@@
   # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
-  geom_line(data = filter(variog_fit, timelag == 1), size = 0.5, color = "darkcyan", alpha = 0.8) +
-  geom_line(data = filter(variog_fit, timelag == 0), size = 0.5, color = "red1", alpha = 0.8) +
+  geom_line(data = filter(variog_fit, timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
+  geom_line(data = filter(variog_fit, timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
   scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
@@
   scale_y_continuous(name = "Community dissimilarity") +
-  ggnomics::facet_nested(metric + tech + amplicon + algorithm ~ guild, scales = "free", nest_line = TRUE) +
   theme(strip.background = element_blank())
@@
 
-```{r variofit}
+```{r variofit, results="asis"}
 variog_data %>%
-  mutate_at("variofitST", lapply, purrr::compose(as.data.frame, as.list, gstat::extractPar)) %>%
-  unnest(variofitST) %>%
-  arrange(metric, amplicon, tech, guild) %>%
-  select(Amplicon = amplicon, Tech = tech, Guild = guild, Algorithm = algorithm,
-         Metric = metric, Range = range, Anisotropy = anis) %>%
-  mutate(Range = 3*Range,
-         "Time Range" = Range/Anisotropy,
-         "Time Range" = ifelse(`Time Range` > 100, NA, `Time Range`)) %>%
-  mutate_at(c("Range", "Time Range"), formatC, digits = 1,
-            format = "f") %>%
-  mutate_at("Range", paste, "m") %>%
-  mutate_at("Time Range", paste, "a") %>%
-  mutate_at("Time Range", str_replace, "NA a", "-") %>%
-  mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
-  select(Amplicon, Tech, Guild, Algorithm, Metric, "Space Range" = Range, "Time Range") %>%
-  kable(booktabs = TRUE, caption = "(ref:variofit)")
+  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
+  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
+  mutate(params = map(variofit, broom::tidy)) %>%
+  unnest(params) %>%
+  bind_cols(
+  pmap_dfr(
+    .,
+    function(mantel, variofit, term, ...) {
+      if (mantel) {
+        tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% as.data.frame,
+                 error = function(e) tibble(conf.low = NA_real_, conf.high = NA_real_))
+      } else {
+        tibble(conf.low = NA_real_, conf.high = NA_real_)
+      }
+    })
+  ) %>%
+  filter(term %in% c("range", "timerange")) %>%
+  mutate(value = ifelse(
+    is.na(estimate) | is.na(conf.low) | is.na(conf.high),
+    "-",
+    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
+  )) %>%
+  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
+  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
+  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
+  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
 ```
@@
 theme_set(oldtheme)
+```
+
+(ref:ecm) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments.
+Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
+"unidentified" denotes ASVs which were assigned to kingdom Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
+Height of each bar represents the fraction of reads represented by that bar.
+Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; "All" denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.
+
+```{r ECM, fig.cap = "(ref:ecm)"}
+  readd(taxon_reads, cache = cache) %>%
+  mutate_at("reference", replace_na, "All") %>%
+  taxon_plot(ECM, kingdom == "Fungi") +
+  theme(legend.position = "bottom", legend.title = element_blank())
 ```

### 2020-02-06 scripts/variogram.R

@@ -20,6 +20,7 @@ as_variogram <- function(vario) {
       np = as.numeric(dplyr::n()),
       dist = mean(dist),
       gamma = mean(gamma),
+      var = var(gamma),
       dir.hor = 0,
       dir.ver = 0,
       id = "var1"

### 2020-02-07 writing/transect_paper.Rmd
variogram table

@@
 
-  - Long amplicons found longer autocorrelation of Bray-Curtis for all fungi, but not ECM.  Is this due to some particular group?
-  - Long amplicon also detects some autocorrelation even after 1 year, absent in other tests.
-  - Unifrac has less significant autocorrelation.
+  - Very similar results for different amplicons and technologies
+  - Unifrac shows very little distance structure.
 
@@
         if (mantel) {
-        tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% as.data.frame,
-                 error = function(e) tibble(conf.low = NA_real_, conf.high = NA_real_))
+          tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% inset2("term", term) %>% as.data.frame,
+                   error = function(e) tibble(term = term, conf.low = NA_real_, conf.high = NA_real_))
         } else {
-        tibble(conf.low = NA_real_, conf.high = NA_real_)
+          tibble(term = term, conf.low = NA_real_, conf.high = NA_real_)
         }
-    })
+      }) %>%
+      select(-term)
   ) %>%
@@
   )) %>%
+  filter(timelag == "0" | term == "timerange") %>%
   pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
@@
 
- - Fit looks good for long Pacbio w/Bray-Curtis.
+ - Spatial ranges are all consistent, given the wide confidence intervals.
+ - For most models the temporal aspect showed poor convergence.
  - Unifrac shows unconvincing distance structure. (No surprise that the model doesn't fit well)
- - Distance structure for short amplicon is really confusing.
-Why are the samples 25 m apart so similar?
-(Even after 1 year?) Could the samples be mixed up?
-Or is this just a statistical artifact (very few sample pairs have that separation).
- - ECM has shorter range and less anisotropy for all metrics
-   - shorter range = more spatially variable
-   - less anisotropy = less temporally variable
+ - similar results for whole community and ECM -- ECM dominate community.
 

### 2020-02-10 writing/transect_paper.Rmd
combine venn diagram calculations into a function

@@
 ```{r venn-table-asv}
-loadd(venn_ASV, cache = cache)
-
-linesep <- venn_ASV %>%
-  filter(!is.na(ASVs)) %>%
-  select(starts_with("reads")) %>%
-  mutate_all(nchar) %>%
-  as.matrix %>%
-  is_greater_than(0) %>%
-  rowSums %>%
-  rle() %$%
-  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
-  unlist() %>%
-  head(length(.) - 2) %>%
-  c("\\midrule")
-empty_background = if (options("knitr.table.format") == "html") "#777777" else "gray"
-venn_ASV %>%
-  rownames_to_column(" ") %>%
-  filter(!is.na(ASVs)) %>%
-  mutate_all(~ifelse(
-    . == "",
-    kableExtra::cell_spec("  ", background = empty_background, background_as_tile = FALSE), .)) %>%
-  mutate_at(" ", str_replace, "[01]+", "") %>%
-  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
-  set_names(gsub("_", " ", names(.))) %>%
-  kable(
-    booktabs = TRUE,
-    caption = "(ref:venn-table-asv)",
-    linesep = linesep,
-    align = ("rcrrcrrcrr"),
-    row.names = FALSE,
-    escape = FALSE
-  ) %>%
-  kableExtra::add_header_above(
-    header = c(
-      " " = 2,
-      names(venn_ASV)[-(1)] %>% 
-        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
-        rle() %$%
-        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
-      )
-  ) %>%
-  kableExtra::add_header_above(
-    header = c(
-      " " = 2,
-       names(venn_ASV)[-(1)] %>% 
-        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
-        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
-        rle() %$%
-        set_names(lengths, values)
-      )
-  )
+readd(venn_ASV, cache = cache) %>%
+venn_table(ASVs, "(ref:venn-table-asv)")
 ```
@@
 ```{r venn-table-otu}
-loadd(venn_OTU, cache = cache)
-
-linesep <- venn_OTU %>%
-  filter(!is.na(OTUs)) %>%
-  select(starts_with("reads")) %>%
-  mutate_all(nchar) %>%
-  as.matrix %>%
-  is_greater_than(0) %>%
-  rowSums %>%
-  rle() %$%
-  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
-  unlist() %>%
-  head(length(.) - 2) %>%
-  c("\\midrule")
-venn_OTU %>% as.matrix() %>% equals("") %>% which()
-venn_OTU %>%
-  rownames_to_column(" ") %>%
-  filter(!is.na(OTUs)) %>%
-  mutate_all(~ifelse(
-    . == "",
-    kableExtra::cell_spec(
-      "  ",
-      background = empty_background,
-      background_as_tile = FALSE
-    ),
-    .)) %>%
-  mutate_at(" ", str_replace, "[01]+", "") %>%
-  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
-  set_names(gsub("_", " ", names(.))) %>%
-  kable(
-    booktabs = TRUE,
-    caption = "(ref:venn-table-otu)",
-    linesep = linesep,
-    align = ("rcrrcrrcrrcrr"),
-    row.names = FALSE,
-    escape = FALSE
-  ) %>%
-  kableExtra::add_header_above(
-    header = c(
-      " " = 2,
-      names(venn_OTU)[-1] %>% 
-        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
-        rle() %$%
-        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
-      )
-  ) %>%
-  kableExtra::add_header_above(
-    header = c(
-     " " = 2,
-      names(venn_OTU)[-1] %>% 
-        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
-        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
-        rle() %$%
-        set_names(lengths, values)
-      )
-  ) %>%
-  kableExtra::landscape()
+readd(venn_OTU, cache = cache) %>%
+venn_table(OTUs, "(ref:venn-table-otu)")
 ```
@@
 ```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
-vennplot_ASV <- venn_ASV %>%
-  rownames_to_column("set") %>%
-  filter(set != "Total") %>%
-  mutate(set = map_chr(set,
-                       ~ strsplit(., "")[[1]] %>%
-                     as.integer() %>%
-                     as.logical() %>%
-                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
-                     paste(collapse = "&"))) %>%
-  mutate(
-    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
-                       sep = "/") %>%
-      gsub("/+", "/", .) %>%
-      gsub("(^/|/$)", "", .),
-    label = str_c(ASVs, reads_frac, sep = "\n") %>%
-      replace_na("-")
-  )
-
-eulerr::venn(deframe(select(vennplot_ASV, set, ASVs)) %>% replace_na(0)) %>%
+readd(vennplot_ASV, cache = cache) %>% {
   plot(
+    select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
     legend = list(
@@
     ),
-    quantities = vennplot_ASV$label,
-    main = "ASV",
-    gp = grid::gpar(cex = 0.8)
-  )
-
-vennplot_OTU <- venn_OTU %>%
-  rownames_to_column("set") %>%
-  filter(set != "Total") %>%
-  mutate(set = map_chr(set,
-                       ~ strsplit(., "")[[1]] %>%
-                     as.integer() %>%
-                     as.logical() %>%
-                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
-                     paste(collapse = "&"))) %>%
-  mutate(
-    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
-                       sep = "/") %>%
-      gsub("/+", "/", .) %>%
-      gsub("(^/|/$)", "", .),
-    label = str_c(OTUs, reads_frac, sep = "\n") %>%
-      replace_na("-")
-  )
-
-eulerr::venn(deframe(select(vennplot_OTU, set, OTUs)) %>% replace_na(0)) %>%
+    quantities = .$label,
+    main = "ASV"
+  )
+}
+
+readd(vennplot_OTU, cache = cache) %>% {
   plot(
+    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
     legend = FALSE,
     labels = FALSE,
-    quantities = vennplot_OTU$label,
-    main = "OTU",
-    gp = grid::gpar(cex = 0.6)
-  )
+    quantities = .$label,
+    main = "OTU"
+  )
+}
 

### 2020-02-10 scripts/drake_plan_light.R

@@ -20,6 +20,7 @@ source(file.path(config$rdir, "variogram.R"))
 source(file.path(config$rdir, "qstats.R"))
 source(file.path(config$rdir, "taxonomy.R"))
 source(file.path(config$rdir, "plate_check.R"))
+source(file.path(config$rdir, "output_functions.R"))
 
 choosevars <- function(d, g, .data) {
   .data %>%
@@ -702,138 +703,13 @@ plan2 <- drake_plan(
     arrange(step) %>%
     column_to_rownames("step"),
   
-  venn_ASV =
-    asv_table %>%
-    select(seq, pb_500, pb_483, `SH-2257`, is_057) %>%
-    mutate_if(is.numeric, list(found = ~. > 0)) %>%
-    group_by_at(vars(ends_with("_found"))) %>%
-    summarize_if(is.numeric, list(ASVs = ~sum(. > 0), reads = sum)) %>%
-    ungroup() %>%
-    mutate_at(vars(ends_with("_found")), as.integer) %>%
-    complete(pb_500_found, pb_483_found, `SH-2257_found`, is_057_found) %>%
-    mutate_if(is.integer, replace_na, 0L) %>%
-    mutate_if(is.double, replace_na, 0L) %>%
-    tidyr::unite(col = "set", 1:4, sep = "") %>%
-    filter(set != "0000") %>%
-    pivot_longer(
-      -1,
-      names_to = c("seq_run", "what"),
-      values_to = "value",
-      names_pattern = "([[:alpha:]]+[-._]\\d+)_(.+)"
-    ) %>%
-    pivot_wider(names_from = "what", values_from = "value") %>%
-    group_by(seq_run) %>%
-    mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
-    bind_rows(
-      group_by(., seq_run) %>%
-        summarize_if(is.numeric, sum, na.rm = TRUE) %>%
-        mutate(set = "Total")
-    ) %>%
-    mutate_at(
-      vars(ends_with("_frac")),
-      formatC,
-      digits = 2,
-      format = "f",
-      drop0trailing = FALSE
-    ) %>%
-    mutate_at(
-      "reads",
-      k_or_M,
-      .sep = "",
-      .function = formatC,
-      format = "fg",
-      digits = 2
-    ) %>%
-    mutate(
-      ASVs_frac = if_else(grepl("^ *[0.]+ *$", ASVs), "", ASVs_frac),
[…]

### 2020-02-10 writing/transect_paper.Rmd
Text for end of results

@@
 
-Figure \@ref(fig:correlog):
-
-  - Very similar results for different amplicons and technologies
-  - Unifrac shows very little distance structure.
+Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure \@ref(fig:correlog).
+Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 3m for the total fungal community, and 1–2m for the ECM community.
+In both cases, the greatest correlation magnitudes and distances are associated with Illumina sequencing of the short amplicon library, followed by PacBio sequencing of the long amplicon library, with PacBio sequencing of the short amplicon library detecting the least spatial structure.
+
+In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1m 
+distance class showing a significant correlation at $p < 0.05$.
+
+The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1 year and 0m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
+For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0m and 1 year was not significant, but there was a negative correlation at time lag of 1 year and a distance of 1m for both sequencing technologies.
+This negative correlation was significant in each short amplicon correlogram except for the ECM community sequenced by Illumina.
 
 ## Spatio-temporal variogram
+
+Variograms relating community dissimilarities and spatial distances are shown in Figure \@ref(fig:spatio).
+For the cases where the Mantel correlogram detected a significant positive correlation, the best-fit exponential curve is also displayed; for the long amplicon library, where there was also a significant positive correlation after a time lag of one year, the best fit curve for time lag is also included.
+The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in \@ref(tab:variofit).
+Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0m was included in the 95% confidence interval.
+Although the best fit spatial ranges vary between the different datasets by a factor of about 3, there is overlap of the 95% confidence intervals for all of the Bray-Curtis variograms across communities, amplicon libraries, and sequencing technologies.
 
@@
 
-```{r spatio-temporal, fig.cap = "(ref:spatio)"}
+```{r spatio, fig.cap = "(ref:spatio)"}
 variog_data <- light_plan %>%
@@
   mutate_at("correlog", map, "mantel.res") %>%
-  mutate_at("correlog", map_lgl, ~ min(.x[,"Pr(corrected)"], na.rm = TRUE) < 0.05) %>%
+  mutate_at("correlog", map_lgl, ~ any(.x[,"Pr(corrected)"] < 0.05 & .x[, "Mantel.cor"] > 0, na.rm = TRUE)) %>%
   pivot_wider(names_from = "timelag", values_from = "correlog", names_prefix = "mantel_") %>%
@@
   theme(strip.background = element_blank())
-
 ```
@@
   filter(term %in% c("range", "timerange")) %>%
+  mutate_at("conf.low", replace_na, 0) %>%
   mutate(value = ifelse(
-    is.na(estimate) | is.na(conf.low) | is.na(conf.high),
+    is.na(estimate) | is.na(conf.high),
     "-",
@@
 
-Figure \@ref(fig:spatio-temporal) and Table \@ref(tab:variofit):
-
- - Spatial ranges are all consistent, given the wide confidence intervals.
- - For most models the temporal aspect showed poor convergence.
- - Unifrac shows unconvincing distance structure. (No surprise that the model doesn't fit well)
- - similar results for whole community and ECM -- ECM dominate community.
-
 # Discussion
 
-## Comparison of tech
+## Comparison of technologies
 
   - length bias not clearly observed between PacBio and Ion Torrent short reads
-  - long reads had more very short ITS2 sequences -- who are these?
+  - long reads had more very short ITS2 sequences Supplementary Figure \@ref(fig:ITS2-length)) -- these are mostly Alveolata (Supplementary Figure \@ref(fig:short-its))
   - lots of ASVs/OTUs detected in only one tech
 
-  - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
+  - very few PacBio OTUs undetected by Ion torrent/Illumina from same library; only rare OTUS undetected by PacBio -- sequencing depth seems to be the biggest effect.
+
   - sequencing technologies yield comparable results (given read depth) with sufficient clustering
   - not such good match with ASVs -- suggests DADA2 doesn't perform equally on different datasets
-  - No "hook", because depth is similar
+
   - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
 
+## Use of Phylogeny to improve taxonomic identification
+
+  - Strict consensus usually increases the number of assigned sequences relative to any single method, except for family and genus level identifications.
+  - Different methods and databases bring mostly complementary, non-contradictory information.
+  - Unite (using full ITS) identifies a similar fraction of reads and ASVs as RDP using LSU.
+
+## Differences in taxonomic bias of different amplicons
+  - Long amplicon library includes more non-fungi, particularly Alveolates and Metazoans.
+  - Also more of a few isolated fungal genera -- Tremellodendron, Marcelleina, Tulasnella.
+  - Short amplicon recovers more Sordariomycetes, Saccharomycetes, Glomeromycota -- the latter two maybe due to sequencing depth.
+
+## Bias in ecological results using different amplicons and sequencing technologies
+  - UNIFRAC does not detect spatial structure within this relatively homogeneous community -- largest fraction of weighted branch length is between Ascomycota and Basidiomycota.
+  - choice of PCR primer (and possibly PCR replicate?) is more important for repeatability that choice of sequencing technology.
+
 # Acknowledgements
@@
 The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
-Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden
+Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden.
 

### 2020-02-10 writing/transect_paper.Rmd
Add Illumina info to methods

@@
 Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
-for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752 37 UPPSALA).
+for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752 37 UPPSALA, Sweden).
 Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
 
-Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGC.
+Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read lengh of 300 bp at the SNP&SEQ Technology Platform (Dept.
+of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751 44 UPPSALA, Sweden).
 
@@
 
-Circular consensus sequence (CCS) basecalls for SMRT sequences were made using `ccs` version 3.4 [@pacificbiosciences2019] using the default settings.
-The resulting sequences were demultiplexed and sequencing primers were removed using `cutadapt` version 1.18 [@martin2011].
+Circular consensus sequence (CCS) basecalls for PacBio sequences were made using `ccs` version 3.4 [@pacificbiosciences2019] using the default settings.
+The resulting sequences, as well as the paired-end Illumina sequences, were demultiplexed and sequencing primers were removed using `cutadapt` version 2.8 [@martin2011].
 Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible.
@@
 ### Denoising
+
+Forward and reverse Illumina reads were quality controlled (max 3 expected errors in either read) and denoised using `DADA2`.
+Forward and reverse reads were
 
@@
 *Raw*: raw reads as delivered by sequencing center;
-*Trim*: reads with primers and demultiplexing indices removed;
+*Trim*: reads with primers and demultiplexing barcodes removed;
 *LSUx*: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
@@
 ```{r bioinfo}
+#TODO add filtered reads (currently inflated numbers for non-Illumina)
+#TODO add Illumina raw reads
+#TODO asterisks for mappde ASVs
 loadd(bioinf_table, cache = cache)

### 2020-02-10 writing/transect_paper.Rmd
Examine positive control after better Illumina trimming.

@@
   )
-```
-
-```{r agaricus-check}
-agaricus <- readd(taxon_table, cache = cache) %>%
-  filter(rank == "genus", taxon == "Agaricus", n_diff == 1) %$%
-  unique(label)
-agaricus_reads <- readd(proto_physeq, cache = cache) %>%
-  phyloseq::subset_samples(sample_type == "Pos") %>%
-  prune_taxa(taxa = agaricus) %>%
-  prune_taxa(taxa = colSums(otu_table(.)) > 0) %>%
-  merge_samples("seq_run") %>%
-  otu_table()
-
-readd(allseqs, cache = cache) %>% filter(hash %in% taxa_names(agaricus_reads)) %$%
-  set_names(short, hash) %>%
-  discard(is.na) %>%
-  Biostrings::RNAStringSet() %>%
-  Biostrings::writeXStringSet(here("temp/agaricus.fasta"))
-
-agaricus_order <- readd(proto_physeq, cache = cache) %>%
-  merge_samples("seq_run") %>%
-  otu_table() %>%
-  apply(MARGIN = 1, order, decreasing = TRUE) %>%
-  apply(MARGIN = 2, match, x = 1)
 ```
@@
 
-*Agaricus bisporus*, the positive control, was represented by a single ASV in both long- and short-amplicon PacBio datasets.
-It was the most abundant ASV in the long-amplicon dataset, and the fifth most abundant ASV in the short-amplicon dataset.
-The same ASV was also the most abundant ASV in the Ion Torrent dataset, but three additional ASVs identified as *Agaricus bisporus* were also present in that dataset.
-The abundances of these ASVs (`r glue::glue_collapse(c(unclass(agaricus_reads["is_057", ])), sep = ", ", last = ", and ")`) was inconsistent with both heterozygosity (which would tend toward equal read counts for two variants) and intragenomic variation (which would tend toward ratios on the order of the number of rDNA regions in the genome, i.e.
-10--1000).
-This suggests that even after `DADA2` denoising, the Ion Torrent dataset still includes spurious ASVs due to PCR and/or sequencing errors.
+```{r agaricus-check}
+agaricus <- readd(taxon_table, cache = cache) %>%
+  filter(rank == "genus", taxon == "Agaricus", n_diff == 1) %$%
+  unique(label)
+agaricus_reads <- readd(proto_physeq, cache = cache) %>%
+  phyloseq::subset_samples(sample_type == "Pos" | (seq_run == "is_057" & well == "E7")) %>%
+  prune_taxa(taxa = agaricus) %>%
+  prune_taxa(taxa = colSums(otu_table(.)) > 0) %>%
+  otu_table()
+
+pos_control <- agaricus_reads %>% as.matrix() %>% colSums() %>% which.max() %>% names()
+
+pos_control_reads <- readd(proto_physeq, cache = cache) %>%
+  phyloseq::subset_samples(sample_type != "Pos" & !(seq_run == "is_057" & well == "E7")) %>%
+  prune_taxa(taxa = pos_control) %>%
+  otu_table() %>%
+  rowSums()
+all_reads <- readd(proto_physeq, cache = cache) %>%
+  phyloseq::subset_samples(sample_type != "Pos" & !(seq_run == "is_057" & well == "E7")) %>%
+  otu_table() %>%
+  as.matrix() %>%
+  rowSums()
+
+pos_control_data <-
+  readd(proto_physeq, cache = cache) %>%
+  phyloseq::subset_samples(sample_type != "Pos" & !(seq_run == "is_057" & well == "E7")) %>%
+  sample_data() %>%
+  as.data.frame() %>%
+  mutate(
+    pc_reads = pos_control_reads,
+    all_reads = all_reads,
+    pc_frac = pc_reads/all_reads
+) %>%
+  select(seq_run, well, plate, pc_frac, pc_reads, all_reads)
+
+readd(allseqs, cache = cache) %>% filter(hash %in% taxa_names(agaricus_reads)) %$%
+  set_names(short, hash) %>%
+  discard(is.na) %>%
+  Biostrings::RNAStringSet() %>%
+  Biostrings::writeXStringSet(here("temp/agaricus.fasta"))
+
+agaricus_order <- readd(proto_physeq, cache = cache) %>%
+  merge_samples("seq_run") %>%
+  otu_table() %>%
+  apply(MARGIN = 1, order, decreasing = TRUE) %>%
+  apply(MARGIN = 2, match, x = 1)
+```
+
+*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
+Additional ASVs belonging to the genus *Agaricus* were also detected in the same wells as the positive control in the Ion Torrent dataset, but these were sufficiently different that they probably represent additional *Agaricus* species from the field samples, and were found in the same wells due to incomplete demultiplexing.
+*A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair.
+The abundance of the second ASV was `r agaricus_reads %>% as.data.frame() %>% with(f2ab49e8/`39d80c19`) %>% keep(~.
+> 0) %>% percent(format = "f", digits = 2) %>% paste(collapse = " and ")` that of the primary ASV in the two Illumina positive controls.
+The consistency of this ratio suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
+
+The primary *A.
+bisporus* ASV was the most abundant ASV in the long-amplicon PacBio dataset and the Ion Torrent dataset, and the fifth most abundant ASV in the short-amplicon PacBio dataset, and the third most abundant ASV in the Illumina dataset.
+This ASV was also found in `r pos_control_data %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, and represented `r pos_control_data %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
 
@@
 ```{r correlog, fig.cap="(ref:correlog-cap)"}
-
 light_plan %>%

### 2020-02-11 writing/transect_paper.Rmd
move more code out of the paper into the local drake plan

@@
 write_bib(c("base", "phyloseq", "gstat"), file = "R.bib", tweak = TRUE)
-cache = drake_cache(here(".light"))
+
+# use the "light" plan cache instead of the default cache at .drake
+cache <- drake_cache(here(".light"))
+readd <- function(..., cache = drake_cache(here(".light"))) {
+  force(cache)
+  drake::readd(..., cache = cache)
+}
+loadd <- function(..., cache = drake_cache(here(".light"))) {
+  force(cache)
+  drake::loadd(..., cache = cache, envir = parent.frame())
+}
+
 light_plan <- readRDS(here("data/plan/drake_light.rds"))
@@
 
-k_or_M <- function(x, ..., .sep = " ", .fun = format) {
-  case_when(
-    abs(x) > 1e6 ~ paste(.fun(x / 1e6, ...), "M", sep = .sep),
-    abs(x) > 1e3 ~ paste(.fun(x / 1e3, ...), "k", sep = .sep),
-    is.na(x) ~ NA_character_,
-    TRUE ~ paste(.fun(x, ...), " ", sep = .sep)
-  )
-}
-
-percent <- function(x, ...) {
-  paste0(formatC(as.numeric(x) * 100, ...), "%")
-}
-
-text_list <- function(x, ...) {
-  x = format(x, ...)
-  glue::glue_collapse(x, sep = ", ", last = ", and ")
-}
+source(here(file.path("scripts", "output_functions.R")))
 ```
@@
 
-We attempted to denoise both long and short amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019.
-However, only `r region_table$ASVs[region_table$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r round(100*region_table$reads[region_table$region == "long"]/as.integer(gsub(" ", "", demuxcounts["pb_500"])), 1)`% of the trimmed reads.
+We attempted to denoise both long and short PacBio amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019.
+However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
 We believe that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
@@
 
-The 5.8S region was located in `r text_list(bioin)
-
 (ref:readcounts) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages.
@@
 #TODO add filtered reads (currently inflated numbers for non-Illumina)
-#TODO add Illumina raw reads
-#TODO asterisks for mappde ASVs
-loadd(bioinf_table, cache = cache)
+#TODO asterisks for mapped ASVs
+# loadd(bioinf_table)
 options(knitr.kable.NA = '–')
-bioinf_header <- tibble(name = names(bioinf_table)) %>%
-  separate(name, into = c("tech", "amplicon", "type"), sep = "_") %>%
-  mutate(tech = paste(tech, plyr::mapvalues(tech, datasets$tech, datasets$machine, FALSE)))
-bioinf_table %>%
+
+readd(bioinf_table) %>%
   set_names(gsub(".*_", "", names(.))) %>%
@@
       " " = 1,
-      bioinf_header$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
+      readd(bioinf_header)$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
     )
@@
       " " = 1,
-      rle(bioinf_header$tech) %$% set_names(lengths, values)
+      rle(readd(bioinf_header)$tech) %$% set_names(lengths, values)
     )
@@
 ```{r venn-table-asv}
-readd(venn_ASV, cache = cache) %>%
+readd(venn_ASV) %>%
 venn_table(ASVs, "(ref:venn-table-asv)")
@@
 ```{r venn-table-otu}
-readd(venn_OTU, cache = cache) %>%
+readd(venn_OTU) %>%
 venn_table(OTUs, "(ref:venn-table-otu)")
@@
 ```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
-readd(vennplot_ASV, cache = cache) %>% {
+readd(vennplot_ASV) %>% {
   plot(
@@
 
-readd(vennplot_OTU, cache = cache) %>% {
+readd(vennplot_OTU) %>% {
   plot(
@@
 
-(ref:readcap) Comparison between read numbers for different sequencing strategies, by ASV (left) and 97% OTU (right).
+(ref:readcap) Comparison between read numbers for different sequencing strategies, by ASV (top) and 97% OTU (bottom).
 ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
@@
 ```{r pb-vs-it, fig.cap="(ref:readcap)", fig.height=8}
-loadd(
-  big_table,
-  multi_table,
-  comparisons,
-  cache = cache
-)
-
-compR2 <- function(x, y, type) {
-  xamp <- plyr::mapvalues(
-    x,
-    datasets$seq_run,
-    datasets$amplicon,
-    FALSE
-  )
-  yamp <- plyr::mapvalues(
-    y,
-    datasets$seq_run,
-    datasets$amplicon,
-    FALSE
-  )
-  x <- plyr::mapvalues(
-    x,
-    datasets$seq_run,
-    paste(
-      datasets$tech,
-      datasets$machine),
-    FALSE
-  )
-  y <- plyr::mapvalues(
-    y,
-    datasets$seq_run,
-    paste(
-      datasets$tech,
-      datasets$machine),
-    FALSE
-  )
-  formatC(
-    filter(
-      comparisons,
-      x == x_var,
-      xamp == x_amplicon,
-      y == y_var,
-      yamp == y_amplicon,
-      type == !!type
-    )$r.squared,
-    digits = 2,
-    format = "f")
-}
-
-read_comparison <- function(multi_table, comparisons, type) {
-  multi_table <- filter(multi_table, type == !!type)
-  comparisons <- filter(comparisons, type == !!type)
-  p <- multi_table %>%
-    filter(x > 0, y > 0) %>%
-    ggplot(aes(y = y, x = x)) +
-    geom_point(alpha = 0.2, shape = 1) +
-    scale_x_log10(
-      breaks = c(1, 1000, 1000000),
-      labels = c("1", "1k", "1M"),
-      minor_breaks = c(10, 100, 10000, 100000),
-      limits = c(1, 1e6)
-    ) +
-    scale_y_log10(
-      breaks = c(1, 1000, 1000000),
-      labels = c("1", "1k", "1M"),
-      minor_breaks = c(10, 100, 10000, 100000),
-      limits = c(1, 1e6)
-    ) +
-    coord_equal() +
-    xlab(NULL) +
-    ylab(NULL) +
-    geom_rug(aes(y = y), sides = "l", alpha = 0.2,
-             data = multi_table %>% filter(x == 0, y > 0)) +
-    geom_rug(aes(x = x), sides = "b", alpha = 0.2,
-             data = multi_table %>% filter(y == 0, x > 0)) +
-    ggnomics::facet_nested(y_amplicon + y_var ~ x_amplicon + x_var, switch = "both", space = "free", nest_line = TRUE) +
-    stat_smooth(method = "loess", formula = y ~ x) +
-    geom_abline(aes(intercept = `(Intercept)`, slope = 1),
-                linetype = 2L, alpha = 0.5, data = comparisons) +
-    geom_text(
-      aes(label = label),
-      data = comparisons,
-      x = 6,
-      y = 0.7,
-      color = "black",
-      hjust = 1,
-      parse = TRUE,
-      inherit.aes = FALSE
-    ) +
-    theme(strip.placement = "outside", strip.background = element_blank())
-  g <- ggplotGrob(p)
-  g$grobs[g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")] <- NULL
-  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")),]
-  g
-}
 ggpubr::ggarrange(
-  read_comparison(multi_table, comparisons, type = "ASV"),
-  read_comparison(multi_table, comparisons, type = "OTU"),
+  read_comparison(
+    readd(multi_table),
+    readd(comparisons),
+    type = "ASV"
+  ),
+  read_comparison(
+    readd(multi_table),
+    readd(comparisons),
+    type = "OTU"
+  ),
   ncol = 1,
@@
 The number of ASVs and OTUs shared between the different sequencing strategies are shown in Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in figure \@ref(fig:venn).
-ASVs shared between all datasets represented `r text_list(percent(venn_ASV["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
-These fractions increased to `r text_list(percent(venn_OTU["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
-In particular, `r text_list(percent(colSums(mutate_all(venn_OTU[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric))))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
-In contrast, `r percent(venn_ASV["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(venn_OTU["1000", "frac_pb_500"])` after OTU clustering.
+ASVs shared between all datasets represented `r text_list(percent(readd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
+These fractions increased to `r text_list(percent(readd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+In particular, `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
+In contrast, `r percent(readd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
 
 Figure \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
-In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = compR2("SH-2257", "is_057", "OTU")`$).
+In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = `r compR2("SH-2257", "is_057", "OTU")`$).
 Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
@@
 ```{r tax-chart, fig.cap = "(ref:tax-chart)"}
-readd(taxon_reads, cache = cache) %>%
-  pivot_longer(names_to = "rank", cols = c("kingdom", "phylum", "class", "order", "family", "genus"), values_to = "taxon") %>%
-  group_by(amplicon, tech, Algorithm, reference, rank, ID = !is.na(taxon)) %>%
-  summarize(reads = sum(reads), ASVs = n()) %>%
-  group_by(amplicon, tech, Algorithm, reference, rank) %>%
-  mutate(ASVs = ASVs / sum(ASVs)) %>%
-  ungroup() %>%
-  filter(ID) %>%
-  mutate_at("rank", factor, levels = c("kingdom", "phylum", "class", "order", "family", "genus")) %>%
-  rename(Reads = reads) %>%
-  mutate_at("reference", as.character) %>%
-  mutate_at("reference", replace_na, "All") %>%
-  mutate(
-    reference = map2(reference, amplicon, ~ if (.x == "All") c("Unite", "Warcup", if (.y == "Long") "RDP" else NULL) else .)
-  ) %>%
-  unnest(reference) %>%
-  mutate_at("reference", factor, levels = c("Unite", "Warcup", "RDP")) %>%
-  pivot_longer(cols = c("Reads", "ASVs"), names_to = "type", values_to = "frac") %>%
-  ggplot(aes(x = rank, y = frac, ymax = frac, group = Algorithm, fill = Algorithm)) +
-  ggnomics::facet_nested(
-    tech + amplicon ~ type + reference,
-    resect = unit(1, "mm"),
-    nest_line = TRUE,
-    bleed = FALSE
-  ) +
-  # geom_ribbon(ymin = 0, alpha = 0.5) +
-  # geom_line(aes(color = Algorithm)) +
-  geom_col(position = "dodge") +
-  # scale_fill_brewer(type = "qual", palette = 2) +
-  ylab("Fraction of reads assigned") +
-  xlab(NULL) +
-  theme(axis.text.x = element_text(vjust = 0.5, angle = 90),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        strip.background = element_blank())
+readd(tax_chart)
 ```
@@
 Diverse but relatively rare groups are thus shown as small, light-colored nodes on relatively thick, darker-green-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as large nodes on relatively thin branches (e.g., Russulales).
-Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "...".
+Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
 ASVs which could not be further assigned are labelled "?".
@@
 ```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
-fungi_read <- readd(taxdata, cache = cache) %>%
+fungi_read <- readd(taxdata) %>%
   taxa::filter_taxa(
@@
 
-fungi_asv <- readd(taxdata, cache = cache) %>%
+fungi_asv <- readd(taxdata) %>%
   taxa::filter_taxa(
@@
 
-ecm_counts <- readd(taxon_reads, cache = cache) %>%
+ecm_counts <- readd(taxon_reads) %>%
   filter(Algorithm == "PHYLOTAX") %>%
@@
 theme <- theme_update(panel.border = element_blank())
-readd(taxdata, cache = cache) %>%
+readd(taxdata) %>%
   metacoder::heat_tree(
@@
 
-```{r taxon-plot}
-taxon_plot <- function(.data, rank, ..., y = reads, cutoff = NULL) {
-  rank <- enquo(rank)
-  y <- enquo(y)
-  ranks <- c("kingdom", "phylum", "class", "order", "family", "genus")
-  .data <- .data %>%
-    group_by(reference, Algorithm, tech, amplicon) %>%
-    mutate(ASVs = n()) %>%
-    ungroup() %>%
-    filter(...) %>%
-    arrange_at(ranks) %>%
-    mutate_at(
-      ranks,
-      ~ factor(
-        .,
-        levels = c(NA, "other", discard(unique(.), is.na)),
-        exclude = NULL
-      )
-    ) %>%
-    group_by(reference, Algorithm, tech, amplicon, !!rank) %>%
-    summarize(reads = sum(reads), ASVs = n()/max(ASVs)) %>%
-    ungroup()
-
-  if (!is.null(cutoff)) {
-    .data <- group_by(.data, !!rank) %>%
-      group_map(
-
-        ~ if (all(pull(.x, !!y) < cutoff)) {
-          mutate(.x, !!rank := factor("other", levels = levels(!!rank)))
-        } else {
-          .x
-        },
-        keep = TRUE
-      ) %>%
-      bind_rows() %>%
-      group_by(reference, Algorithm, !!rank, tech, amplicon) %>%
-      summarize(reads = sum(reads), ASVs = sum(ASVs)) %>%
-      ungroup()
-  }
-
-  rank_label <- as_label(rank)
-
-  .data <- mutate(.data, !!rank := fct_drop(!!rank))
-  vals <- levels(pull(.data, !!rank))
-  # if ("other" %in% vals) vals <- c("other", vals) %>% magrittr::extract(!duplicated(.))
-  # if (any(is.na(vals))) vals <- c(NA, vals) %>% magrittr::extract(!duplicated(.))
-
-  if (rank_label == str_to_lower(rank_label)) rank_label <- str_to_title(rank_label)
-  y_label <- as_label(y)
-  if (y_label == str_to_lower(y_label)) y_label <- str_to_title(y_label)
-  .data <- .data %>%
-    mutate(tech = fct_relabel(tech, ~ paste(., plyr::mapvalues(., datasets$tech, datasets$machine, FALSE))
-    ))
-  ggplot(.data, aes(x = Algorithm, y = !!y, fill = !!rank)) +
-    geom_bar(position = "stack", stat = "identity", color = "white", size = 0.2) +
-    ggnomics::facet_nested(
-      cols = vars(tech, amplicon, reference),
-      scales = "free_x",
-      space = "free_x",
-      nest_line = TRUE,
-      bleed = FALSE,
-      resect = unit(1, "mm")
-      ) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
-          strip.background = element_blank(),
-          panel.spacing = unit(3, "pt")) +
-    scale_fill_discrete(
-      breaks = vals,
-      labels = replace_na(as.character(vals), "unidentified"),
-      name = rank_label
-    ) +
-    ylab(paste("Fraction of", y_label))
-
-}
-
-```
-
 (ref:heattree-db-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
@@
 ```{r heattree-db-compare, fig.cap = "(ref:heattree-db-compare)", include = FALSE}
-taxdata2 <- readd(taxdata, cache = cache)
+taxdata2 <- readd(taxdata)
 # taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
@@
 
-```{r agaricus-check}
-agaricus <- readd(taxon_table, cache = cache) %>%
-  filter(rank == "genus", taxon == "Agaricus", n_diff == 1) %$%
-  unique(label)
-agaricus_reads <- readd(proto_physeq, cache = cache) %>%
-  phyloseq::subset_samples(sample_type == "Pos" | (seq_run == "is_057" & well == "E7")) %>%
-  prune_taxa(taxa = agaricus) %>%
-  prune_taxa(taxa = colSums(otu_table(.)) > 0) %>%
-  otu_table()
-
-pos_control <- agaricus_reads %>% as.matrix() %>% colSums() %>% which.max() %>% names()
-
-pos_control_reads <- readd(proto_physeq, cache = cache) %>%
-  phyloseq::subset_samples(sample_type != "Pos" & !(seq_run == "is_057" & well == "E7")) %>%
-  prune_taxa(taxa = pos_control) %>%
-  otu_table() %>%
-  rowSums()
-all_reads <- readd(proto_physeq, cache = cache) %>%
-  phyloseq::subset_samples(sample_type != "Pos" & !(seq_run == "is_057" & well == "E7")) %>%
-  otu_table() %>%
-  as.matrix() %>%
-  rowSums()
-
-pos_control_data <-
-  readd(proto_physeq, cache = cache) %>%
-  phyloseq::subset_samples(sample_type != "Pos" & !(seq_run == "is_057" & well == "E7")) %>%
-  sample_data() %>%
-  as.data.frame() %>%
-  mutate(
-    pc_reads = pos_control_reads,
-    all_reads = all_reads,
-    pc_frac = pc_reads/all_reads
-) %>%
-  select(seq_run, well, plate, pc_frac, pc_reads, all_reads)
-
-readd(allseqs, cache = cache) %>% filter(hash %in% taxa_names(agaricus_reads)) %$%
-  set_names(short, hash) %>%
-  discard(is.na) %>%
-  Biostrings::RNAStringSet() %>%
-  Biostrings::writeXStringSet(here("temp/agaricus.fasta"))
-
-agaricus_order <- readd(proto_physeq, cache = cache) %>%
-  merge_samples("seq_run") %>%
-  otu_table() %>%
-  apply(MARGIN = 1, order, decreasing = TRUE) %>%
-  apply(MARGIN = 2, match, x = 1)
-```
-
 *Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
@@
 *A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair.
-The abundance of the second ASV was `r agaricus_reads %>% as.data.frame() %>% with(f2ab49e8/`39d80c19`) %>% keep(~.
+The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
 > 0) %>% percent(format = "f", digits = 2) %>% paste(collapse = " and ")` that of the primary ASV in the two Illumina positive controls.
-The consistency of this ratio suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
+The consistency of this ratio across PCR replicates suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
 
@@
 bisporus* ASV was the most abundant ASV in the long-amplicon PacBio dataset and the Ion Torrent dataset, and the fifth most abundant ASV in the short-amplicon PacBio dataset, and the third most abundant ASV in the Illumina dataset.
-This ASV was also found in `r pos_control_data %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, and represented `r pos_control_data %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
+This ASV was also found in `r readd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, and represented `r readd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
 
@@
   ) %>%
-  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
+  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
   mutate_at("target", purrr::map, "mantel.res") %>%
@@
 ```{r spatio, fig.cap = "(ref:spatio)"}
-variog_data <- light_plan %>%
-  filter(step == "correlog") %>%
-  select(correlog, timelag, guild, metric, tech, amplicon, algorithm) %>%
-  mutate_at("correlog", lapply, readd, character_only = TRUE, cache = cache) %>%
-  mutate_at("correlog", map, "mantel.res") %>%
-  mutate_at("correlog", map_lgl, ~ any(.x[,"Pr(corrected)"] < 0.05 & .x[, "Mantel.cor"] > 0, na.rm = TRUE)) %>%
-  pivot_wider(names_from = "timelag", values_from = "correlog", names_prefix = "mantel_") %>%
-  left_join(
-    light_plan %>%
-      filter(step == "variofit2") %>%
-      select(variofit2, guild, metric, tech, amplicon, algorithm),
-    by = c("guild", "metric", "tech", "amplicon", "algorithm")
-    ) %>%
-  left_join(
-    light_plan %>%
-      filter(step == "variofitST2") %>%
-      select(variofitST2, variogST, guild, metric, tech, amplicon, algorithm),
-    by = c("guild", "metric", "tech", "amplicon", "algorithm")
-    ) %>%
-  mutate(
-    variofit = ifelse(mantel_1, variofitST2, variofit2)
-    ) %>%
-  mutate_at(c("variofit", "variogST", "variofitST2", "variofit2"), map, readd, character_only = TRUE, cache = cache) %>%
-  select(mantel_0, mantel_1, variofitST2, variofit2, variofit, variogST, guild, metric, tech, amplicon, algorithm) %>%
-  mutate_at(c("guild", "metric", "tech", "amplicon", "algorithm"), str_replace_all, "\"", "") %>%
-  mutate(
-    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("All Fungi", "ECM")),
-    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
-    amplicon = factor(
-      amplicon,
-      levels = c("Short", "Long")
-    ),
-    algorithm = factor(
-      algorithm,
-      levels = c("Consensus", "PHYLOTAX", "PHYLOTAX+Cons"),
-      labels = c("Cons", "PHYLO", "PHYLO")
-    )
-  )
-
-variog_points <- unnest(variog_data, c("variogST")) %>%
-  mutate_at("timelag", replace_na, "0") %>%
-  filter(dist <= 25) %>%
-  mutate_at("timelag", factor)
-
-variog_empir <- variog_points %>%
-  # mutate(bin = cut(dist, c(1:13 - 0.5, 14.5, 16.5, 19.5, 24.5))) %>%
-  group_by(guild, metric, tech, amplicon, algorithm, dist, timelag) %>%
-  summarize(
-    gamma = mean(gamma),
-    # dist = mean(dist),
-    n = n()
-  )
-
-variog_confint <- variog_points %>%
-  group_by(guild, metric, tech, amplicon, algorithm, dist, timelag) %>%
-  summarize(
-    lower = quantile(gamma, 0.025),
-    upper = quantile(gamma, 0.975)
-  )
-
-variog_fit <- variog_data %>%
-  mutate_at("variofit", map,
-            ~ mutate(.y, gamma = predict(.x, newdata = .y)),
-            expand_grid(dist = seq(1, 25, 0.5), timelag = 0:1)
-            ) %>%
-  unnest(variofit) %>%
-  filter(mantel_1 | timelag == "0", mantel_0) %>%
-  mutate_at("timelag", factor)
-
-variog_empir %>%
+
+readd(variog_empir) %>%
   filter(algorithm == "PHYLO") %>%
@@
   # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
-  geom_line(data = filter(variog_fit, timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
-  geom_line(data = filter(variog_fit, timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
+  geom_line(data = filter(readd(variog_fit), timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
+  geom_line(data = filter(readd(variog_fit), timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
   scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
@@
 ```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
-full_dens <- readd(reads_table, cache = cache) %>%
+full_dens <- readd(reads_table) %>%
   dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon)) %>%
@@
 
-full_ecdf <- readd(reads_table, cache = cache) %>%
+full_ecdf <- readd(reads_table) %>%
   dplyr::filter(region  %in% c("long", "short")) %>%
@@
 ```{r regions, fig.cap="(ref:region-asv)"}
-readd(region_table, cache = cache) %>%
+readd(region_table) %>%
   dplyr::filter(seq_run == "pb_500") %>%
@@
 
-readd(reads_table, cache = cache) %>%
+readd(reads_table) %>%
   dplyr::filter(region %in% baseregions,
@@
 ```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
-its2 <- readd(reads_table, cache = cache) %>%
+its2 <- readd(reads_table) %>%
   dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
@@
 ```{r short-its, fig.cap = "(ref:short-its)"}
-readd(big_seq_table_ITS2, cache = cache) %>%
+readd(big_seq_table_ITS2) %>%
   colSums() %>%
@@
   mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(readd(taxon_reads, cache = cache), ., by = c("label" = "ITS2")) %>%
+  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
   # group_by(ITS2) %>%
@@
 ```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare)"}
-taxdata2 <- readd(taxdata, cache = cache)
+taxdata2 <- readd(taxdata)
 # taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
@@
 ```{r ECM, fig.cap = "(ref:ecm)"}
-  readd(taxon_reads, cache = cache) %>%
+  readd(taxon_reads) %>%
   mutate_at("reference", replace_na, "All") %>%
@@
 ```{r ecm-heattree, fig.cap = "(ref:ecm-heattree)"}
-taxdata2 <- readd(taxdata_ECM, cache = cache)
+taxdata2 <- readd(taxdata_ECM)
 
@@
 ```{r sample-depth, fig.cap = "(ref:sample-depth)"}
-reads_plot <- readd(qstats_n, cache = cache) %>%
+reads_plot <- readd(qstats_n) %>%
   mutate(file = str_replace(file, ".*/", "") %>%
@@
   left_join(
-    readd(proto_physeq, cache = cache) %>% {
+    readd(proto_physeq) %>% {
       mutate(
@@
 
-conc_plot <- readd(proto_physeq, cache = cache) %>% 
+conc_plot <- readd(proto_physeq) %>% 
         sample_data(.) %>%
@@
 ```{r plate-loading}
-readd(proto_physeq, cache = cache) %>% {
+readd(proto_physeq) %>% {
   mutate(
@@
 ```{r}
-physeq <- readd(proto_physeq, cache = cache) %>%
+physeq <- readd(proto_physeq) %>%
   `sample_data<-`(inset2(
@@
 ranks <- c("root", "kingdom", "phylum", "class", "order", "family", "genus")
-physeq <- readd(taxon_reads, cache = cache) %>%
+physeq <- readd(taxon_reads) %>%
   mutate(root = "Root") %>%

### 2020-02-11 scripts/variogram.R

@@ -20,7 +20,7 @@ as_variogram <- function(vario) {
       np = as.numeric(dplyr::n()),
       dist = mean(dist),
       gamma = mean(gamma),
-      var = var(gamma),
+      # var = var(gamma),
       dir.hor = 0,
       dir.ver = 0,
       id = "var1"
@@ -59,7 +59,7 @@ as_variogramST <- function(vario) {
       np = as.numeric(dplyr::n()),
       dist = mean(dist),
       gamma = mean(gamma),
-      var = var(gamma),
+      # var = var(gamma),
       id = "var1"
     ) %>%
     dplyr::group_by(bin) %>%

### 2020-02-11 writing/transect_paper.Rmd
Include comments from Anna/Martin as HTML comments in text.

@@
 - Ectomycorrhizas (ECM) are a symbiosis between fungi and plant roots.
-- x% of plants, y% of fungi
+- x% of plants, y% of fungi <!-- Brundrett for plants and Ornella Comandini , Andrea C. Rinaldi, Thomas W. Kuyper for fungi (with some calculations) - Martin -->
 - The plants are often ecologically and economically important trees (Pinaceae, Fagaceae, Dipterocarpaceae, Amherstiae)
 - The fungi provide nutrients and protection from pathogens, the plant provides sugar.
+
+## section
+
 - Many well-known mushrooms are produced by ECM (Amanita, Cantharellus, Boletus) but many ECM produce inconspicuous (Tomentella) or no (Cenococcum) fruit bodies.
 - Anyway the fruitbodies are ephemeral, so study of ECM communities requires looking at vegetative structures.
-- Not easy to culture, so we need molecular methods.
+- Not easy to culture, so we need molecular methods. <!-- I think these two points should be very brief, like maybe one sentence for both. - Martin-->
 
@@
   - Clustering-based OTU approach may both "clump" and "split" species.
+<!-- It would be nice with a transition here from previous paragraph.
+Possible effects of splitting and lumping on turnover.
+Or what turnover may depend on.
+e.g.
+individual size, niche turnover.
+Martin-->
   - Distance-based clustering conflates intra-species variation and sequencing error
@@
 - Short reads
-    - 454 [@buee2009], Ion Torrent, Illumina [@schmidt2013].
+  - 454 [@buee2009] <!--I'm personally not very found of this reference because it’s the kind of sloppy and fast work that some people do to be first rather than doing a good job.
+You can use it if you like but I think it already has more reference than it deserves.-AR-->, Ion Torrent, Illumina [@schmidt2013].
   - For fungi, frequently ITS1 or ITS2 [@schoch2012]
@@
   - short reads from same location can potentially be mapped to tree. (but we didn't find that this works)
-    - Sequencing depth tradeoff [@Kennedy2018]
+  - Sequencing depth tradeoff [@Kennedy2018] <!--Maybe this should start the short/long read part.
+Or be part of a transition between with the connection of OTU/phylogenetic methods and sequencing strategy.
+- Martin-->
 
@@
 Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25 soil samples were along a linear transect at intervals of 1 m in May 2015.
+At each site, 25 soil samples were collected along a linear transect at intervals of 1 m in May 2015.
 One third of the sample locations (3-m spacing) were resampled one year later in June 2016.
-For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a knife blade.
+For each sample, any coarse organic debris <!--Do you have any definition?
+Wider than 5 mm?
+MR--> was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a <!-- sterilized?
+-MR--> knife blade.
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
@@
 After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
-Samples preserved using LifeGuard were first centrifuged at XX ×g for Y minutes, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
+Samples preserved using LifeGuard were first centrifuged at 10000×g for 1 minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
 DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
@@
 Two different fragments of the nuclear rDNA were amplified (Figure \@ref(fig:rDNA)).
-The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
-The long amplicon (approximately 1550 bp) included approximately 12 bp at the 3' end of the small subunit (SSU) rDNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (Figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
+The long amplicon (approximately 1500 bp) included approximately 12 bp at the 3' end of the small subunit (SSU) rDNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (Figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 
 The gITS7 primers for the short amplicon were indexed for multiplexing (**supp info**).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7 ng purified DNA in Dream *Taq* buffer.
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150µM ITS4m, 2mM McCl_2_, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7 ng purified DNA in Dream *Taq* buffer.
 The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
@@
 
+<!--If limited by space, this could be in supp, and then just described in text. -MR-->
 ```{tikz rDNA, fig.cap = "(ref:rDNA)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
@@
 An aliquot of 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
+The first library contained all samples from 2016 plus one PCR blank and one positive control (68 total samples), and the second containes all samples from 2015 plus one PCR blank and one positive control (42 total samples).
 Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
@@
 
-Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read lengh of 300 bp at the SNP&SEQ Technology Platform (Dept.
+Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300 bp at the SNP&SEQ Technology Platform (Dept.
 of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751 44 UPPSALA, Sweden).
+The Illumina library was pooled with samples for another project, with one half of the reads from one lane devoted to this project.
 
@@
 Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible.
+For Ion Torrent and PacBio, reads were discarded if they did not have the appropriate primers on both ends.
+Reads were searched in both directions, and reads where the primers were found in the reverse direction were reverse complemented before further analysis.
+For Illumina sequences, primers were also searched for and removed on the 3' ends of the reads, in case of readthrough with short amplicons.
+Read pairs were retained when PCR primers were detected at the 5' ends of both the forward and reverse read.
+Read pairs where the primers were found in reverse orientation were kept in seperate files, but were retained in their original orientation until after denoising.
 
@@
 
-Forward and reverse Illumina reads were quality controlled (max 3 expected errors in either read) and denoised using `DADA2`.
-Forward and reverse reads were
+Forward and reverse Illumina reads were quality controlled (truncation after first base with quality <= 10, and then max 3 expected errors in either read) and denoised using `DADA2`.
+Forward and reverse reads were denoised using seperate error models and then merged.
 
@@
 No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
-<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
+<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
 A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
 A CM was generated from the alignment using `cmbuild` from Infernal.
-The fragment of each read beginning with the 5.8S rRNA was then aligned to the CM using `cmalign` from Infernal.
+The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
 The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure \@ref(fig:rDNA),
@@
 and IDTAXA [@murali2018].
-Each ASV was thus given up to nine preliminary taxonomic assignments (three region/reference combinations $\times$ three algorithms).
+Each ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
 
@@
 
-Unique LSU regions from long amplicon ASVs were aligned as RNA using DECIPHER [@wright2015] with up to 10 iterations of progressive alignment and conserved secondary structure calculation and 10 refinement iterations.
-This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had type 1 introns after this position [@holst-jensen1999].
+Unique LSU regions from long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10 iterations of progressive alignment and conserved RNA secondary structure calculation and 10 refinement iterations.
+<!-- The LSU tree ends up not looking as nice as the tree which includes ITS (I don't know why, except that maybe 5.8S is more helpful than ITS is harmful?) so I'm using that.
+Maybe I don't need to mention the LSU-only tree.-->
+This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had type 1 introns
+<!-- Just curious, what are Type 1 introns -AR--> <!-- It has to do with the enzyme that cuts them out, I think, but I don't actually know my introns are Type 1, just that this is a location where Type 1 introns are known to occur.
+-BF-->
+after this position [@holst-jensen1999].
 Full length long amplicon ASVs (including ITS1, 5.8S, and ITS2 regions) were aligned and truncated in the same way.
-
-<!--An attempt was made to refine the long amplicon alignment using Pasta [@mirarab2014a], but the resulting maximum likelihood (ML) trees did not improve on the ML achieved using the original alignment from DECIPHER.
-Since ML is the optimization criterion used by Pasta, the original DECIPHER alignment was retained.-->
 
@@
 The LSU tree was not constrained topologically, but the long amplicon tree was constrained by the result for the LSU tree.
-
-*Short amplicon ASVs without a perfect match to one of the long amplicon ASVs were added to the long amplicon alignment using the `--add` option of MAFFT version 7.453 [@katoh2012] with a 10-mer based guide tree and two iterations.
-The short ASVs were then placed on the long amplicon tree using EPA-NG version 0.3.6 [@barbera2019].
-A "grafted" tree including the short amplicon ASVs in the most probable positions from the EPA-NG results was produced using GAPPA version 0.5.1 [@czech2019a].
-When more than one short amplicon ASV was placed in the same position on the tree, GAPPA by default groups all of these sequences together in a polytomy.
-The branch leading to each of these polytomies was deleted by setting its length to zero and then using the `d2multi` function from the `ape` package in R.
-The resulting tree was used as a guide tree to construct a final ML tree in RAxML, using the GTR+GAMMA model, in order to resolve relationships between different sequences assigned to the same branch.
-No bootstrapping was performed on the final tree.* -- this didn't work.
+*Trees and alignments are available in Treebase. (not yet)*
 
@@
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
-Inspection of the tree along with taxonomic assignments revealed that two ASVs assigned to the genus *Tulasnella* (Basidiomycota), which is known to have accelerated evolution in the rDNA [@moncalvo2006], were nested within the Metazoa, so they were excluded.
-The kingdom Fungi was then identified as the minimal clade containing all remaining ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
+The kingdom Fungi was identified as the minimal clade containing all ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
 ASVs falling outside this clade were not included in the fungal community analysis.
@@
 This refinement method is referred to as "Consensus".
-Additionally, a hybrid method, "PHYLOTAX+C", was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
-
-<!--
-          - preliminary alignment of 32S to RDP covariance model with `cmalign` from Infernal (aligns only conserved regions)
-          - concatenate ITS1 to beginning of alignment
-          - create guide tree using alignment of only conserved, gap-free columns in preliminary alignment, using UPGMA in DECIPHER [@wright2016]
-          - realign nonconserved sites using MLocARNA (progressive simultaneous alignment and folding) from LocARNA 2.0.0RC8 [@will2007; @will2012; @will2013]. -->
-
-## Statistics
+Additionally, a hybrid method, was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
+
+## Spatial analysis
 

### 2020-02-11 writing/transect_paper.Rmd
intro text

@@
 
-- Ectomycorrhizas (ECM) are a symbiosis between fungi and plant roots.
-- x% of plants, y% of fungi <!-- Brundrett for plants and Ornella Comandini , Andrea C. Rinaldi, Thomas W. Kuyper for fungi (with some calculations) - Martin -->
-- The plants are often ecologically and economically important trees (Pinaceae, Fagaceae, Dipterocarpaceae, Amherstiae)
-- The fungi provide nutrients and protection from pathogens, the plant provides sugar.
-
-## section
-
-- Many well-known mushrooms are produced by ECM (Amanita, Cantharellus, Boletus) but many ECM produce inconspicuous (Tomentella) or no (Cenococcum) fruit bodies.
-- Anyway the fruitbodies are ephemeral, so study of ECM communities requires looking at vegetative structures.
-- Not easy to culture, so we need molecular methods. <!-- I think these two points should be very brief, like maybe one sentence for both. - Martin-->
-
-## Long reads vs short reads/phylogenetic methods
-
-- Clustering vs. phylogenetic methods
-  - Clustering-based OTU approach may both "clump" and "split" species.
-<!-- It would be nice with a transition here from previous paragraph.
-Possible effects of splitting and lumping on turnover.
-Or what turnover may depend on.
-e.g.
-individual size, niche turnover.
-Martin-->
-  - Distance-based clustering conflates intra-species variation and sequencing error
-  - Denoising (DADA2 [@callahan2017], Deblur [@amir2017], UNOISE2 [@edgar2016]) attempts to control for sequencing error while leaving intra-species variation.
-  - Phylogenetic community distance measures like UNIFRAC [@wong2016, look for earlier reference] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
-
-- Short reads
-  - 454 [@buee2009] <!--I'm personally not very found of this reference because it’s the kind of sloppy and fast work that some people do to be first rather than doing a good job.
-You can use it if you like but I think it already has more reference than it deserves.-AR-->, Ion Torrent, Illumina [@schmidt2013].
-  - For fungi, frequently ITS1 or ITS2 [@schoch2012]
-  - Dependent on matching to database for taxonomic placement.
-  - Phylogenetic placement algorithms [@matsen2010; @berger2011] are not easy to apply because ITS is not in most LSU alignments (different situation than V4/V5 of 18S)
-  - Possible to place on taxonomy-based tree [@tedersoo2018]
-
-- Long reads
-  - Can include both LSU/SSU and ITS in same amplicon for better taxonomic placement [@Tedersoo2018]
-  - can build tree directly
-  - short reads from same location can potentially be mapped to tree. (but we didn't find that this works)
-  - Sequencing depth tradeoff [@Kennedy2018] <!--Maybe this should start the short/long read part.
-Or be part of a transition between with the connection of OTU/phylogenetic methods and sequencing strategy.
-- Martin-->
+Ectomycorrhizas (ECM) are a symbiosis between fungi and plant root tips.
+The fungi provide nutrients and protection from pathogens in exchange for sugards from the plant [@Smith2010].
+Approximately 8% of described fungal species [@ainsworth2008ainsworth;@Rinaldi2008] and 2% of land plant species [@Brundrett2017a] form ECM.
+The plants include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
+Amherstieae.
+
+Although many well-known mushrooms are produced by ECM fungi (e.g., *Amanita*, *Cantharellus*, *Boletus*), many ECM fungi instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
+Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures.
+Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures.
+The advent of high-thoughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils.
+
+## Metabarcoding methods
+
+As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
+Traditionally, fungal metabarcoding studies have targetted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
+The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis.
+If taxonomic identification is desired, it has usually been performed by database searches using BLAST (ref).
+This approach comes with some weaknesses.
+
+While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the large taxonomic scales typically present in environmental samples.
+Additionally the wide range of length variation of these regions may introduce bias in recovery of different taxa.
+Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].
+The recent advent of long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) introduce the possibility of sequencing longer amplicons which include both the ITS regions and the flanking, more conserved SSU and/or LSU regions [@Tedersoo2018].
+This has the potential to improve taxonomic placement of sequences without close database matches, and to allow the alignment of metabarcoding reads.
+However, long-read technologies are currently more expensive per read, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
+
+Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
+In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while leaving intra-species variation.
+The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent all true amplicon sequences from the sample.
+Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are not as dataset specific [@callahan2017].
+
+Assignment of taxonomic identities using BLAST requires *a priori* choice of threshholds for different taxonomic ranks.
+Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments [e.g.,@wang2007;@edgar2016a;@murali2018a].
+In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017].
+
+Phylogenetic information has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
+Because OTU clustering may both "clump" different species in a single OTU, and "split" a single species into multiple OTUS [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
+In contrast, phylogenetic community distance measures like UNIFRAC [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
+Phylogenetic placement algorithms [@matsen2010; @berger2011] are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
+Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018].
+However, long amplicon sequencing allows the inclusion of alignable regions for construction of more fully resolved phylogenetic trees directly from metabarcoding reads.
+
+We investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM Soudanian woodland in Benin.
 
@@
 
-*why is turnover scale interesting*
-
-- @lilleskov2004 - autocorrelation only <2.6 m at most sites.  Based on Sanger sequencing root tips in temperate forests.
-
-- Meta analysis @Bahram2013
-  - Includes data from Benin (same site? ask Nourou/Leho) but this is unpublished
-  - conclude that there is greater range of distance decay in tropical forests; because of less host availability? older forests?  *Russulaceae*?
-  - 65 m in Benin
-
-- @pickles2012a
-  - T-RFLP of root tips, temperate forest
-  - Autocorrelation < 3.4 m
-
-- Data from African wooded savannas - @tedersoo2011 but only southern band
-  - Irregular sampling with no samples nearer than 8m
-  - based on sanger sequencing root tips
-  - Significant Moran's I only in <10m class
-
-*statement of question*
+Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
+Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
+It varies between different ecosystems and taxonomic groups.
+Turnover scale is often measured by the range at which Moran's I statistic indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve community dissimilarity vs.
+distance [@legendre2012].
+
+Several studies have investigated the turnover scale in ECM forests, but so far they have been based on analysis of ECM root tips, rather than bulk soil.
+@lilleskov2004 found autocorrelation only at ranges <2.6 m at most sites in temperate forests, based on Sanger sequencing.
+Similarly, @pickles2012a found autocorrelation at distances <3.4 m, also in temperate forests, based on T-RFLP.
+Two studies have included data from tropical African woodlands and wooded savannahs.
+@tedersoo2011 found autocorrelation at ranges <10 m in Miombo woodland based on Sanger sequencing.
+However, this study was based on irregular sampling with a minimum spacing of 8 m, so this as the smallest distance class.
+@Bahram2013 performed a meta-analysis of turnover scale in ECM communities, including data from Benin, where they found turnover range of 65 m based on distance-decay curves.
+They concluded that the range of distance decay is greater in tropical forests than temperate forests.
+
+We investigated turnover range in both the total fungal community and ECM community by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
+We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RPP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs.
+weighted UNIFRAC).
+We also present new algorithms for dividing the rDNA into regions, combining denoising results from multiple regions, and incorporating phylogenetic information into taxonomic assignments.
 
@@
 
-Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064
-°; Gan: N 9.75678° W 2.31058 °) approximately 30 km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30 km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
 Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
@@
 
-## Distance decay
+## Spatial analysis
 

### 2020-02-12 writing/transect_paper.Rmd
move heattree code to light plan

@@
 ```{r ecm-heattree, fig.cap = "(ref:ecm-heattree)"}
-taxdata2 <- readd(taxdata_ECM)
-
-val_cols <- names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_.+_.+_Unite_.+")
-
-taxdata2$filter_obs(
-  "tax_read",
-  rowSums(select_at(taxdata2$data$tax_read, val_cols)) > 0,
-  drop_taxa = TRUE,
-  drop_obs = TRUE,
-  supertaxa = TRUE,
-  reassign_obs = FALSE) %>%
-  invisible()
-
-taxdata2$data$diff_read <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_read",
-  cols = val_cols,
-  groups = val_cols %>%
-    str_replace("reads_.+_(.+)_Unite_.+", "\\1"),
-  func = function(abund1, abund2) {
-    list(read_ratio = log10(mean(abund1) / mean(abund2)))
-  }
-)
-taxdata2$data$diff_asv <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_asv",
-  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
-  groups = names(taxdata2$data$tax_asv) %>%
-    keep(startsWith, "ASVs_") %>%
-    str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
-  func = function(abund1, abund2) {
-    list(ASV_ratio = log10(mean(abund1) / mean(abund2)))
-  }
-)
-
-taxdata2 <- taxdata2$filter_taxa(!is.nan(ASV_ratio), !is.nan(read_ratio))
-
-set.seed(3)
 oldtheme <- theme_update(panel.border = element_blank())
-taxdata2 %>%
-  metacoder::heat_tree(
-    layout = "davidson-harel",
-    initial_layout = "reingold-tilford",
-
-    node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.002, .035),
-    node_size_axis_label = "Read abundance",
-
-    node_color = read_ratio,
-    node_color_range = metacoder::diverging_palette(),
-    node_color_trans = "linear",
-    node_color_axis_label = "Read abundance ratio",
-    node_color_interval = c(-3, 3),
-
-    edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.001, 0.03),
-    edge_size_trans = "linear",
-    edge_size_axis_label = "ASV count",
-
-    edge_color = ASV_ratio,
-    edge_color_range = metacoder::diverging_palette(),
-    edge_color_trans = "linear",
-    edge_color_axis_label = "ASV count ratio",
-    edge_color_interval = c(-0.75, 0.75),
-
-    node_label = taxon_names,
-    node_label_size_range = c(0.02, 0.03),
-    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
-    aspect_ratio = 3/2
-  )
+readd(ecm_heattree)
 theme_set(oldtheme)

### 2020-02-12 scripts/drake_plan_light.R

@@ -1365,6 +1365,79 @@ plan2 <- drake_plan(
     )
     out
   },
+  
+  ecm_heattree = {
+    taxdata2 <- taxdata_ECM
+    
+    val_cols <- names(taxdata2$data$tax_read) %>%
+      keep(str_detect, "reads_.+_.+_Unite_.+")
+    
+    taxdata2$filter_obs(
+      "tax_read",
+      rowSums(select_at(taxdata2$data$tax_read, val_cols)) > 0,
+      drop_taxa = TRUE,
+      drop_obs = TRUE,
+      supertaxa = TRUE,
+      reassign_obs = FALSE) %>%
+      invisible()
+    
+    taxdata2$data$diff_read <- metacoder::compare_groups(
+      taxdata2,
+      data = "tax_read",
+      cols = val_cols,
+      groups = val_cols %>%
+        str_replace("reads_.+_(.+)_Unite_.+", "\\1"),
+      func = function(abund1, abund2) {
+        list(read_ratio = log10(mean(abund1) / mean(abund2)))
+      }
+    )
+    taxdata2$data$diff_asv <- metacoder::compare_groups(
+      taxdata2,
+      data = "tax_asv",
+      cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
+      groups = names(taxdata2$data$tax_asv) %>%
+        keep(startsWith, "ASVs_") %>%
+        str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
+      func = function(abund1, abund2) {
+        list(ASV_ratio = log10(mean(abund1) / mean(abund2)))
+      }
+    )
+    
+    taxdata2 <- taxdata2$filter_taxa(!is.nan(ASV_ratio), !is.nan(read_ratio))
+    
+    set.seed(3)
+    taxdata2 %>%
+      metacoder::heat_tree(
+        layout = "davidson-harel",
+        initial_layout = "reingold-tilford",
+        
+        node_size = rowMeans(.$data$tax_read[,-1]),
+        node_size_range = c(0.002, .035),
+        node_size_axis_label = "Read abundance",
+        
+        node_color = read_ratio,
+        node_color_range = metacoder::diverging_palette(),
+        node_color_trans = "linear",
+        node_color_axis_label = "Read abundance ratio",
+        node_color_interval = c(-3, 3),
+        
+        edge_size = rowMeans(.$data$tax_asv[,-1]),
+        edge_size_range = c(0.001, 0.03),
[…]

### 2020-02-12 writing/transect_paper.Rmd
clarify methods and results

@@
 
-loadd(
-  reads_table,
-  region_table,
-  asv_table,
-  otu_table,
-  demuxlength,
-  demuxqual,
-  readcounts,
-  rawcounts,
-  demuxcounts,
-  regioncounts,
-  cache = cache
-)
-
 source(here(file.path("scripts", "output_functions.R")))
@@
 The fungi provide nutrients and protection from pathogens in exchange for sugards from the plant [@Smith2010].
-Approximately 8% of described fungal species [@ainsworth2008ainsworth;@Rinaldi2008] and 2% of land plant species [@Brundrett2017a] form ECM.
+Approximately 8% of described fungal species [@ainsworth2008ainsworth;@Rinaldi2008] and 2% of land plant species [@brundrett2017a] form ECM.
 The plants include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
@@
 It varies between different ecosystems and taxonomic groups.
-Turnover scale is often measured by the range at which Moran's I statistic indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve community dissimilarity vs.
+Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve community dissimilarity vs.
 distance [@legendre2012].
@@
 Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median $L=`r demuxlength['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(demuxqual['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - demuxqual['pb_500'])^demuxlength['pb_500'])), big.mark = " ")`.
-Given the high diversity relative to sequencing depth in this study (`r region_table$ASVs[region_table$seq_run == "pb_483" & region_table$region == "short"]` ASVs based on PacBio short amplicons, `r demuxcounts["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
-In comparison, the equivalent requirement for the short amplicon ($L=`r demuxlength['pb_483']` \text{ bp}$, $\epsilon=`r demuxqual['pb_483']`$) is only `r round(2/(1 - demuxqual['pb_483'])^demuxlength['pb_483'], 1)` reads.
+For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = " ")`.
+Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r readd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
+In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
 We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
@@
 
-Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation, and fit using a joint exponential spatio-temporal model using the `fit.StVariogram` function from `gstat` version `r packageVersion("gstat")` [@R-gstat].
-The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.
+Additionally, empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by @Legendre2012 using the `nls` function in R.
+Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
+For datasets where the Mantel corellogram indicated spatial correlation between samples taken in seperate years, the model was re-fit with an additional term to represent temporal correlation:
+$$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
+where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity from replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill", and $a_t$ the equivalent temporal range.
+95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package in R.
 
@@
 The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
-Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
-The 5.8S region was detected in `r regioncounts["pb_500"]` (`r regioncounts["pb_483"]`) reads.
-Ion Torrent sequencing yielded `r demuxcounts["is_057"]` demultiplexed reads, with the 5.8S region detected in `r regioncounts["is_057"]`.
-
-### Region extraction
+Long (short) amplicon sequencing with PacBio yielded `r readd(rawcounts)["pb_500"]` (`r readd(rawcounts)["pb_483"]`) circular consensus reads, of which `r readd(demuxcounts)["pb_500"]` (`r readd(demuxcounts)["pb_483"]`) were successfully demultiplexed.
+The 5.8S region was detected in `r readd(regioncounts)["pb_500"]` (`r readd(regioncounts)["pb_483"]`) reads.
+Ion Torrent sequencing yielded `r readd(demuxcounts)["is_057"]` demultiplexed reads, with the 5.8S region detected in `r readd(regioncounts)["is_057"]`.
 
@@
 
+### Region extraction
+
+- Fraction where 5.8S was detected
+- Length/ASV count/ASV fraction of extracted regions.
+
 ### Reproducibility of read counts from different technologies
@@
 }
-
+```
+
+(ref:venn2) (*alternate to \@ref(fig:venn)*) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different sequencing technologies from the same short amplicon library.
+In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.
+
+```{r venn2, fig.cap = "(ref:venn2)", out.height = "30%", fig.show="hold"}
+venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+  vennplot_data(ASVs) %>% {
+  plot(
+     select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
+     legend = list(
+         labels = c("PacBio Short", "Illumina", "Ion Torrent")
+     ),
+     quantities = .$label,
+     main = "ASV",
+     fills = c("grey85", "lightblue", "lightcoral"))
+  }
+
+venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+  vennplot_data(OTUs) %>% {
+  plot(
+    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
+    legend = FALSE,
+    labels = FALSE,
+    quantities = .$label,
+    main = "OTU",
+     fills = c("grey85", "lightblue", "lightcoral")
+  )
+}
+```
+
+(ref:venn3) (*alternate to \@ref(fig:venn)*) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different amplicons.
+In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each amplicon are shown below.
+Long amplicons were sequenced on PacBio RS II, short amplicons were sequenced on PacBio RS II, Illumina MiSeq, and Ion Torrent Ion S5.
+ASV/OTU counts for short amplicons reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
+
+```{r venn3, fig.cap = "(ref:venn2)", out.height = "30%", fig.show="hold"}
+readd(asv_table) %>%
+  mutate_if(is.numeric, ~ . / sum(.)) %>%
+  mutate(long = pb_500,
+         short = rowMeans(.[,c("pb_483", "SH-2257", "is_057")])) %>%
+  select(seq, long, short) %>%
+venndata(ASVs, cols = c("long", "short")) %>%
+  vennplot_data(ASVs) %>% {
+  plot(
+     select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
+     legend = list(
+         labels = c("Long amplicon", "Short amplicon")
+     ),
+     quantities = .$label,
+     main = "ASV",
+     fills = c("white", "slateblue1"))
+  }
+
+readd(otu_table) %>%
+  mutate_if(is.numeric, ~ . / sum(.)) %>%
+  mutate(long = pb_500,
+         short = rowMeans(.[,c("pb_483", "SH-2257", "is_057")])) %>%
+  select(seq, long, short) %>%
+venndata(OTUs, cols = c("long", "short")) %>%
+  vennplot_data(OTUs) %>% {
+  plot(
+    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
+    legend = FALSE,
+    labels = FALSE,
+    quantities = .$label,
+    main = "OTU",
+     fills = c("white", "slateblue1")
+  )
+}
 ```
@@
 `r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs PacBio, Illumina, and Ion Torrent short reads, respectively).
+
+Almost all of the short amplicons from all three technologies were between 240 and 375 bp long (Supplementary Figure \@ref(fig:full-length)a).
+Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure \@ref(fig:full-length)b).
+In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of
 
@@
 ## Spatial analysis
+
+### Autocorrelation
 
@@
 ```{r variofit, results="asis"}
-variog_data %>%
+readd(variog_data) %>%
   select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
@@
 ```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
-extree <-
-  treeio::as.treedata(ape::read.tree(text = "(A:1,((B:1,C:1):1,((E:1,F:1):1,D:1):1):1);")) %>%
-  tidytree::full_join(
-    tibble(label = LETTERS[1:6],
-           tax1 = c("unk", "Tax1", "Tax2", "Tax2", "unk", "unk"),
-           tax2 = c("unk", "Tax2", "Tax2", "Tax1", "unk", "Tax1")),
-    by = "label"
-  )
-ggtree::ggtree(extree) +
-  ggtree::geom_tiplab(x = 4, align = 2) +
-  xlim(0, 6) +
-  ggtree::geom_tiplab(
-    aes(color = tax1, label = tax1),
-    x = 4.7,
-    hjust = 0.5
-  ) +
-  ggtree::geom_tiplab(
-    aes(color = tax2, label = tax2),
-    x = 5.7,
-    hjust = 0.5
-  ) +
-  ggtree::geom_tiplab(x = 5.2, label = "/", hjust = 0.5) +
-  ggtree::geom_nodelab(aes(label = node - 6), geom = "label") +
-  ggtree::geom_hilight(node = 10, fill = "steelblue", extendto = 4) +
-  ggtree::geom_hilight(node = 9, fill = "tomato", extendto = 4) +
-  scale_color_manual(values = c(Tax1 = "steelblue", Tax2 = "tomato", unk = "grey50")) +
-  scale_y_reverse()
-```
-
-(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.
-Length distribution (*a*, *c*) and empirical cumulative distribution function (*b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.
+readd(phylotax_fig)
+```
+
+(ref:full-length) Comparison of the length of denoised sequences from different length amplicons and sequencing technologies.
+Length distribution (*a*, *c*) and empirical cumulative distribution function (ECDF; *b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.
 
 ```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
-full_dens <- readd(reads_table) %>%
-  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon)) %>%
-  dplyr::filter(region %in% c("long", "short")) %>%
-  dplyr::group_by(seq_run) %>%
-  dplyr::mutate_at("amplicon", factor, levels = c("Short", "Long")) %>%
-  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
-  dplyr::mutate(
-    reads = reads / sum(reads)
-  )
-
-full_dens_short <- full_dens %>%
-  filter(amplicon == "Short") %>%
-  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
-  stat_density(bw = 0.5, geom = "line", position = "identity") +
-  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
-  scale_y_continuous(name = "Density") +
-  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
-  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
-  theme(strip.background = element_blank())
-
-full_dens_long <- full_dens %>%
-  filter(amplicon == "Long") %>%
-  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
-  stat_density(bw = 0.5, geom = "line", position = "identity") +
-  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
-  scale_y_continuous(name = "Density") +
-  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
-  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
-  theme(strip.background = element_blank())
-
-full_ecdf <- readd(reads_table) %>%
-  dplyr::filter(region  %in% c("long", "short")) %>%
-  left_join(select(datasets, seq_run, amplicon, tech, machine), by = "seq_run") %>%
-  mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
-  dplyr::group_by(seq_run) %>%
-  dplyr::mutate(reads = reads / sum(reads))
-
-full_ecdf_short <- full_ecdf %>%
-  filter(amplicon == "Short") %>%
-  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
-  stat_ecdf(geom = "line") +
-  scale_color_brewer(type = "qual",
-                     name = NULL, palette = "Set2") +
-  scale_y_continuous(name = "ECDF") +
-  scale_x_continuous(name = NULL) +
-  facet_grid( amplicon ~ ., scales = "free_x") +
-  theme(
-    legend.position = c(1, 0.02),
-    legend.justification = c(1, 0),
-    legend.background = element_blank(),
-    strip.background = element_blank()
-  )
-
-full_ecdf_long <- full_ecdf %>%
-  filter(amplicon == "Long") %>%
-  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
-  stat_ecdf(geom = "line") +
-  scale_color_brewer(type = "qual",
-                     name = NULL, palette = "Set2", direction = -1) +
-  scale_y_continuous(name = "ECDF") +
-  scale_x_continuous(name = "Length (bp)") +
-  facet_grid( amplicon ~ ., scales = "free_x") +
-  theme(
-    legend.position = c(1, 0.02),
-    legend.justification = c(1, 0),
-    legend.background = element_blank(),
-    strip.background = element_blank()
-  )
-
 ggpubr::ggarrange(
-  full_dens_short,
-  full_ecdf_short,
-  full_dens_long,
-  full_ecdf_long,
+  readd(full_dens_short),
+  readd(full_ecdf_short),
+  readd(full_dens_long),
+  readd(full_ecdf_long),
   ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
-
 )
@@
 fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
-Some reggions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU.
+Some regions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU.
 For this figure, subregions were concatenated prior to generating ASVs with `DADA2`.
@@
   dplyr::filter(seq_run == "pb_500") %>%
+  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", is.na(step), !is.na(well))$nreads)) %>%
+  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
   ggplot(aes(reads, ASVs, color = length_med, label = region)) +
@@
   scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
-  xlab("Mapped reads")
+  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
 ```
@@
 ```{r region-lengths, fig.cap = "(ref:region-length)", fig.height = 6}
-baseregions <- c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3",
-               "LSU4")
-
-region_limits <- dplyr::filter(regions, region %in% baseregions) %>%
-  dplyr::mutate_at(
-    "region",
-    factor,
-    levels = baseregions
-    ) %>%
-  tidyr::pivot_longer(
-    cols = c("min_length", "max_length"),
-    names_to = "limit",
-    names_pattern = "(min|max)_length",
-    values_to = "length"
-  )
-
-readd(reads_table) %>%
-  dplyr::filter(region %in% baseregions,
-         seq_run == "pb_500") %>%
-  dplyr::mutate_at(
-    "region",
-    factor,
-    levels = baseregions
-    ) %>%
-  ggplot(aes(length, weight = reads)) +
-  stat_density(bw = 0.5, color = "#8da0cb", geom = "line") +
-  geom_vline(data = region_limits,
-             aes(xintercept = length),
-             linetype = "dashed", alpha = 0.5) +
-  facet_wrap(~region,ncol = 1, strip.position = "right", scales = "free_y") +
-  scale_y_continuous(breaks = NULL, labels = NULL, name = "Read density") +
-  scale_x_continuous(limits = c(0, 500), name = "Length (bp)")
+readd(region_length_fig)
 ```
@@
 ```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
-its2 <- readd(reads_table) %>%
-  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
-  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
-  dplyr::filter(region == "ITS2") %>%
-  arrange(tech, amplicon) %>%
-  mutate(
-    group = paste(tech, "-", amplicon) %>% factor(levels = unique(.))
-  ) %>%
-  dplyr::group_by(seq_run) %>%
-  dplyr::mutate(
-    reads = reads / sum(reads)
-  )
-
-its2_dens <- its2 %>%
-  ggplot(aes(length, weight = reads, color = group, group = group)) +
-  stat_density(bw = 0.5, geom = "line", position = "identity") +
-  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
-  scale_y_continuous(name = "Density") +
-  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
-  ggnomics::facet_nested(rows = vars(tech, amplicon), scales = "free_y",
-                         nest_line = TRUE) +
-  theme(strip.background = element_blank())
-
-its2_ecdf <- its2 %>%
-  mutate(tech = "", amplicon = "") %>%
-  ggplot(aes(length, weight = reads, color = group, group = group)) +
-  stat_ecdf(geom = "line") +
-  scale_color_brewer(type = "qual", name = NULL, palette = "Set2") +
-  scale_y_continuous(name = "ECDF") +
-  scale_x_continuous(name = "ITS2 Length (bp)") +
-  ggnomics::facet_nested(rows = vars(tech, amplicon), scales = "free_y",
-                         nest_line = TRUE) +
-  theme(
-    strip.background = element_blank(),
-    legend.position = c(0.98, 0.02),
-    legend.justification = c(1, 0)
-  )
-
-ggpubr::ggarrange(its2_dens, its2_ecdf, ncol = 1, labels = "auto", heights = c(4, 2))
+ggpubr::ggarrange(
+  readd(its2_dens),
+  readd(its2_ecdf),
+  ncol = 1,
+  labels = "auto",
+  heights = c(4, 2)
+)
 ```
@@
   # filter(!"Short" %in% amplicon) %>%
-  taxon_plot(kingdom, y = reads)
+  taxon_plot(kingdom, y = reads, datasets = datasets)
 ```
@@
   mutate_at("reference", replace_na, "All") %>%
-  taxon_plot(ECM, kingdom == "Fungi") +
+  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
   theme(legend.position = "bottom", legend.title = element_blank())

### 2020-02-12 writing/transect_paper.Rmd
work on results, knit versions to send before we work together on the discussion. Also send more figure code to the plan.

@@
 output:
+  bookdown::word_document2: 
+    keep_md: yes
   bookdown::pdf_document2:
@@
 At each site, 25 soil samples were collected along a linear transect at intervals of 1 m in May 2015.
-One third of the sample locations (3-m spacing) were resampled one year later in June 2016.
+One third of the sample locations (3 m spacing) were resampled one year later in June 2016.
 For each sample, any coarse organic debris <!--Do you have any definition?
@@
 
-Forward and reverse Illumina reads were quality controlled (truncation after first base with quality <= 10, and then max 3 expected errors in either read) and denoised using `DADA2`.
-Forward and reverse reads were denoised using seperate error models and then merged.
-
-We attempted to denoise both long and short PacBio amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019.
+Forward and reverse Illumina reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
+Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using seperate error models and then merged.
+
+We attempted to denoise both long and short PacBio amplicons using `DADA2` according to the steps outlined in  the supplementary information in @callahan2019.
 However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
@@
 The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
-Long (short) amplicon sequencing with PacBio yielded `r readd(rawcounts)["pb_500"]` (`r readd(rawcounts)["pb_483"]`) circular consensus reads, of which `r readd(demuxcounts)["pb_500"]` (`r readd(demuxcounts)["pb_483"]`) were successfully demultiplexed.
-The 5.8S region was detected in `r readd(regioncounts)["pb_500"]` (`r readd(regioncounts)["pb_483"]`) reads.
-Ion Torrent sequencing yielded `r readd(demuxcounts)["is_057"]` demultiplexed reads, with the 5.8S region detected in `r readd(regioncounts)["is_057"]`.
+Long (short) amplicon sequencing with PacBio yielded `r readd(rawcounts)["pb_500"]` (`r readd(rawcounts)["pb_483"]`) circular consensus reads, of which `r readd(demuxcounts)["pb_500"]` (`r readd(demuxcounts)["pb_483"]`) were successfully trimmed and demultiplexed.
+Illumina sequencing yielded `r readd(rawcounts)["SH-2257"]` read pairs, of which `r readd(demuxcounts)["SH-2257"]` were successfully trimmed and demultiplexed, and `r readd(filtercounts_full)["SH-2257"]` passed quality filtering.
+Ion Torrent sequencing yielded `r readd(demuxcounts)["is_057"]` demultiplexed reads.
 
@@
 Figure \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
-In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = `r compR2("SH-2257", "is_057", "OTU")`$).
+In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = `r compR2("SH-2257", "is_057", "OTU")` to $).
 Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
@@
 
-## Spatio-temporal variogram
-
-Variograms relating community dissimilarities and spatial distances are shown in Figure \@ref(fig:spatio).
+Distance-decay curves relating community dissimilarities and spatial distances are shown in Figure \@ref(fig:spatio).
 For the cases where the Mantel correlogram detected a significant positive correlation, the best-fit exponential curve is also displayed; for the long amplicon library, where there was also a significant positive correlation after a time lag of one year, the best fit curve for time lag is also included.
 The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in \@ref(tab:variofit).
-Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0m was included in the 95% confidence interval.
+Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0 m was included in the 95% confidence interval.
 Although the best fit spatial ranges vary between the different datasets by a factor of about 3, there is overlap of the 95% confidence intervals for all of the Bray-Curtis variograms across communities, amplicon libraries, and sequencing technologies.

### 2020-02-12 scripts/drake_plan_light.R

@@ -1,3 +1,4 @@
+remove(list = ls())
 if (exists("snakemake")) {
   snakemake@source(".Rprofile", echo = FALSE)
   load(snakemake@input[["drakedata"]])
@@ -13,6 +14,7 @@ library(drake)
 library(assertr)
 library(disk.frame)
 library(ape)
+library(ggplot2)
 
 source(file.path(config$rdir, "dada.R"))
 source(file.path(config$rdir, "mantel.R"))
@@ -21,8 +23,6 @@ source(file.path(config$rdir, "qstats.R"))
 source(file.path(config$rdir, "taxonomy.R"))
 source(file.path(config$rdir, "plate_check.R"))
 source(file.path(config$rdir, "output_functions.R"))
-
-remove(readd, loadd)
 
 choosevars <- function(d, g, .data) {
   .data %>%
@@ -37,16 +37,8 @@ choosevars <- function(d, g, .data) {
     )
 }
 
-k_or_M <- function(x, ..., .sep = " ", .function = format) {
-  case_when(
-    abs(x) > 1e6 ~ paste(.function(x / 1e6, ...), "M", sep = .sep),
-    abs(x) > 1e3 ~ paste(.function(x / 1e3, ...), "k", sep = .sep),
-    is.na(x) ~ NA_character_,
-    TRUE ~ paste(.function(x, ...), " ", sep = .sep)
-  )
-}
-
 datasets <- read_csv(config$dataset, col_types = "cccccccicccccicc")
+regions <- read_csv(config$regions, col_types = "cccciiiiic")
 
 # combo_meta <- tibble(
 #   group = c("fungi", "euk"),
@@ -627,6 +619,20 @@ plan2 <- drake_plan(
     summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
     deframe(),
   
+  filtercounts_full = readcounts %>%
+    filter(step == "filter", is.na(region) | region %in% c("long", "short")) %>%
+    group_by(seq_run, region) %>%
+    summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
+    select(seq_run, nreads) %>%
+    deframe(), 
+  
+  filtercounts_ITS2 = readcounts %>%
+    filter(step == "filter", region == "ITS2") %>%
+    group_by(seq_run, region) %>%
+    summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
[…]

### 2020-02-13 writing/transect_paper.Rmd
Complete draft as sent to Anna

@@
 ---
-title: "A Tale of Two Transects"
-subtitle: "Also Two Sequencing Technologies, and Two Amplicons"
+title: "Long- and Short-Read Metabarcoding Give Similar Results for Spatial Turnover Range in an Ectomycorrhizal Savanna Woodland"
 author: 
@@
 
+# Abstract
+
+Ectomycorrhizal (ECM) communities are characterized by mutualistic symbiosis between certain fungi and plant roots.
+Spatial turnover rate is one important characteristic of ecological communities, which is relevant for biodiversity assessment and sampling design.
+We investigated turnover rate in an ECM fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400 bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600 bp) with PacBio RS II.
+Resulting turnover rates depended more on statistical method than sequencing technology.
+We found significant autocorrelation in distance classes <3 m using a Mantel correlogram, and a best-fit distance-decay curve range of 17--19 m, both shorter than previously reported for similar habitats.
+The use of long-amplicon sequencing enables construction of a phylogenetic tree from metabarcoding reads, which facilitates taxonomic identification of sequences.
+However, long reads present issues for denoising algorithms in diverse communities.
+We present a solution that splits the reads into shorter homologous regions prior to denoising, and then reconstructs the full denoised reads.
+In the choice between short and long amplicons, we suggest a hybrid approach using short amplicons for sampling breadth and depth, and long amplicons to characterize the local species pool for improved identification and phylogenetic analyses.
+
 # Introduction
 
-## Introduction to ECM communities
-
 Ectomycorrhizas (ECM) are a symbiosis between fungi and plant root tips.
-The fungi provide nutrients and protection from pathogens in exchange for sugards from the plant [@Smith2010].
-Approximately 8% of described fungal species [@ainsworth2008ainsworth;@Rinaldi2008] and 2% of land plant species [@brundrett2017a] form ECM.
-The plants include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
-Amherstieae.
+The fungi provide nutrients and protection from pathogens in exchange for sugars from the plant [@Smith2010].
+Approximately 8% of described fungal species are believed to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
+Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
+Amherstieae [@brundrett2017a].
 
 Although many well-known mushrooms are produced by ECM fungi (e.g., *Amanita*, *Cantharellus*, *Boletus*), many ECM fungi instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
-Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures.
+Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures [@horton2001].
 Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures.
-The advent of high-thoughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils.
-
-## Metabarcoding methods
+The advent of high-thoughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils [@Lindahl2013].
 
@@
 Traditionally, fungal metabarcoding studies have targetted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
-The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis.
-If taxonomic identification is desired, it has usually been performed by database searches using BLAST (ref).
-This approach comes with some weaknesses.
-
-While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the large taxonomic scales typically present in environmental samples.
+The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
+If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013].
+However, this traditional approach comes with some potential weaknesses.
+
+While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the large taxonomic scales typically present in environmental samples [@Lindahl2013].
 Additionally the wide range of length variation of these regions may introduce bias in recovery of different taxa.
@@
 The recent advent of long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) introduce the possibility of sequencing longer amplicons which include both the ITS regions and the flanking, more conserved SSU and/or LSU regions [@Tedersoo2018].
-This has the potential to improve taxonomic placement of sequences without close database matches, and to allow the alignment of metabarcoding reads.
+This has the potential to improve taxonomic placement of sequences that lack close database matches, and to allow the alignment of metabarcoding reads.
 However, long-read technologies are currently more expensive per read, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
@@
 In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while leaving intra-species variation.
-The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent all true amplicon sequences from the sample.
+The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
 Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are not as dataset specific [@callahan2017].
@@
 Phylogenetic information has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
-Because OTU clustering may both "clump" different species in a single OTU, and "split" a single species into multiple OTUS [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
+Because OTU clustering may both "clump" different species into a single OTU, and "split" a single species into multiple OTUs [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
 In contrast, phylogenetic community distance measures like UNIFRAC [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
-Phylogenetic placement algorithms [@matsen2010; @berger2011] are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
+Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
 Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018].
@@
 We investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM Soudanian woodland in Benin.
-
-## Turnover scale in ECM forests
 
@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples were found not to be comparable with results preserved using the Xpedition^TM^ lysis buffer (Supplementary figure \@ref(fig:)), so these samples were not included in spatial analyses.
+Sequencing results for these samples were found not to be comparable with results preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-compare-long) and \@ref(fig:buffer-compare-short)), so these samples were not included in spatial analyses.
 However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
@@
 
-Two different fragments of the nuclear rDNA were amplified (Figure \@ref(fig:rDNA)).
-The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
-The long amplicon (approximately 1500 bp) included approximately 12 bp at the 3' end of the small subunit (SSU) rDNA, the full ITS region including the 5.8S rDNA, and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (Figure \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
-
-The gITS7 primers for the short amplicon were indexed for multiplexing (**supp info**).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150µM ITS4m, 2mM McCl_2_, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7 ng purified DNA in Dream *Taq* buffer.
+Two different fragments of the nuclear rDNA were amplified (Figure\ \@ref(fig:rDNA)).
+The short amplicon (approximately 300 bp) targetted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
+The long amplicon (approximately 1500 bp) targetted the full ITS region including the 5.8S rDNA and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+
+The gITS7 primers for the short amplicon were indexed for multiplexing (**Supplementary File**).
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7 ng purified DNA in Dream *Taq* buffer.
 The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
@@
 
-Both primers for the long amplicon were indexed for multiplexing (**supp info**).
+Both primers for the long amplicon were indexed for multiplexing (**Supplementary File**).
 PCR was performed as for the short amplicons, but with 500 µM of each of the two primers.
@@
 An aliquot of 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
-The first library contained all samples from 2016 plus one PCR blank and one positive control (68 total samples), and the second containes all samples from 2015 plus one PCR blank and one positive control (42 total samples).
 Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
@@
 Forward and reverse Illumina reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
-Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using seperate error models and then merged.
+Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using seperate error models and pseudo-pooling, and then forward and reverse reads were merged.
 
@@
 The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
-The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure \@ref(fig:rDNA),
+The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure\ \@ref(fig:rDNA),
 where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
@@
 
-Each of the extracted regions was independently filtered for length (Table \@ref(tab:region-length)) and a maximum of three expected errors.
+Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors.
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
@@
 For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
-Because ITS2 is the most variable of the amplified regions (Supplementary Figure \@ref(fig:regions)), reads with identical ITS2 regions are expected have highly similar sequences in the other regions, unless the amplicon was chimeric.
+Because ITS2 is the most variable of the amplified regions (Supplementary Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected have highly similar sequences in the other regions, unless the amplicon was chimeric.
 The concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015].
@@
 The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence.
-Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed *(how many)*.
-The count and sample distribution of reads assigned to each ASV were calculated in order to form a sample x ASV community matrix.
+Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
+The count and sample distribution of reads assigned to each ASV were calculated in order to form a sample × ASV community matrix.
 A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
@@
 Maybe I don't need to mention the LSU-only tree.-->
-This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had type 1 introns
+This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns
 <!-- Just curious, what are Type 1 introns -AR--> <!-- It has to do with the enzyme that cuts them out, I think, but I don't actually know my introns are Type 1, just that this is a location where Type 1 introns are known to occur.
 -BF-->
-after this position [@holst-jensen1999].
+after this position, as also observed in several fungal species by @holst-jensen1999.
 Full length long amplicon ASVs (including ITS1, 5.8S, and ITS2 regions) were aligned and truncated in the same way.
@@
 
-Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure \@ref(fig:phylotax)).
+Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure\ \@ref(fig:phylotax)).
 A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
@@
 
-Additionally, empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by @Legendre2012 using the `nls` function in R.
+Additionally, empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R.
 Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
@@
 
-The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
+The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table\ \@ref(tab:bioinfo).
 Long (short) amplicon sequencing with PacBio yielded `r readd(rawcounts)["pb_500"]` (`r readd(rawcounts)["pb_483"]`) circular consensus reads, of which `r readd(demuxcounts)["pb_500"]` (`r readd(demuxcounts)["pb_483"]`) were successfully trimmed and demultiplexed.
@@
 Ion Torrent sequencing yielded `r readd(demuxcounts)["is_057"]` demultiplexed reads.
+`r readd(filtercounts_full)['pb_500']` (`r percent(readd(filtercounts_full)['pb_500']/readd(demuxcounts)['pb_500'], digits = 1, format = "f")`) of long amplicon PacBio reads passed length and quality filtering stages, compared with `r readd(filtercounts_full)['pb_483']` (`r percent(readd(filtercounts_full)['pb_483']/readd(demuxcounts)['pb_483'], digits = 1, format = "f")`) of short amplicon PacBio reads, `r readd(filtercounts_full)['SH-2257']` (`r percent(readd(filtercounts_full)['SH-2257']/readd(demuxcounts)['SH-2257'], digits = 1, format = "f")`) of Illumina read pairs, and `r readd(filtercounts_full)['is_057']` (`r percent(readd(filtercounts_full)['is_057']/readd(demuxcounts)['is_057'], digits = 1, format = "f")`) of Ion Torrent reads.
 
@@
 
-### Region extraction
-
-- Fraction where 5.8S was detected
-- Length/ASV count/ASV fraction of extracted regions.
-
-### Reproducibility of read counts from different technologies
-
-(ref:venn-table-asv) Correspondences between ASVs found by different sequencing strategies.
-Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
-
-```{r venn-table-asv}
-readd(venn_ASV) %>%
-venn_table(ASVs, "(ref:venn-table-asv)")
-```
-
-(ref:venn-table-otu) Correspondences between 97% OTUs found by different sequencing strategies.
-Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTU frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
-
-```{r venn-table-otu}
-readd(venn_OTU) %>%
-venn_table(OTUs, "(ref:venn-table-otu)")
-```
-
-(ref:venn) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different sequencing strategies.
+## Region extraction
+
+The 5.8S rDNA was located in `r readd(regioncounts)['pb_500']` long amplicon PacBio reads (`r percent(readd(regioncounts)['pb_500']/readd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total), `r readd(regioncounts)['pb_483']` short amplicon PacBio reads (`r percent(readd(regioncounts)['pb_483']/readd(demuxcounts)['pb_483'], digits = 1, format = "f")`), and `r readd(regioncounts)['is_057']` Ion Torrent reads (`r percent(readd(regioncounts)['is_057']/readd(demuxcounts)['is_057'], digits = 1, format = "f")`).
+Length and quality filtering resulted in `r readd(filtercounts_ITS2)['pb_500']` (`r percent(readd(filtercounts_ITS2)['pb_500']/readd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total) ITS2 sequences from the long amplicon PacBio reads, `r readd(filtercounts_ITS2)['pb_483']` (`r percent(readd(filtercounts_ITS2)['pb_483']/readd(demuxcounts)['pb_483'], digits = 1, format = "f")`) ITS2 sequences from the short amplicon PacBio reads, and `r readd(filtercounts_ITS2)['is_057']` (`r percent(readd(filtercounts_ITS2)['is_057']/readd(demuxcounts)['is_057'], digits = 1, format = "f")`) ITS2 sequences from the Ion torrent reads.
+The length distribution of different regions extracted from the long amplicon are shown in Supplementary Figure\ \@ref(fig:region-lengths).
+
+## Reproducibility of read counts from different technologies
+
+(ref:venn) Venn diagrams showing shared ITS2-based ASVs (*a*, *c*) and 97% OTUs (*b*, *d*) between different sequencing technologies from the same short amplicon library (*a*, *b*), and between long and short amplicon libraries (*c*, *d*).
 In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.
-
-```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
-readd(vennplot_ASV) %>% {
+For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
+
+```{r venn, fig.cap="(ref:venn)", fig.show="hold", out.width = "90%"}
+venn_A <- venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+  vennplot_data(ASVs) %>% {
     plot(
       select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-    legend = list(
-      labels = c("PacBio Long", "PacBio Short", "Ion Torrent", "Illumina")
-    ),
-    quantities = .$label,
-    main = "ASV"
-  )
-}
-
-readd(vennplot_OTU) %>% {
-  plot(
-    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
       legend = FALSE,
@@
       quantities = .$label,
-    main = "OTU"
-  )
-}
-```
-
-(ref:venn2) (*alternate to \@ref(fig:venn)*) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different sequencing technologies from the same short amplicon library.
-In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.
-
-```{r venn2, fig.cap = "(ref:venn2)", out.height = "30%", fig.show="hold"}
-venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
-  vennplot_data(ASVs) %>% {
-  plot(
-     select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-     legend = list(
-         labels = c("PacBio Short", "Illumina", "Ion Torrent")
-     ),
-     quantities = .$label,
-     main = "ASV",
+      main = "a",
       fills = c("grey85", "lightblue", "lightcoral"))
@@
 
-venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+venn_B <- venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
   vennplot_data(OTUs) %>% {
@@
       quantities = .$label,
-    main = "OTU",
+      main = "b",
       fills = c("grey85", "lightblue", "lightcoral")
@@
   }
-```
-
-(ref:venn3) (*alternate to \@ref(fig:venn)*) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different amplicons.
-In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each amplicon are shown below.
-Long amplicons were sequenced on PacBio RS II, short amplicons were sequenced on PacBio RS II, Illumina MiSeq, and Ion Torrent Ion S5.
-ASV/OTU counts for short amplicons reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
-
-```{r venn3, fig.cap = "(ref:venn2)", out.height = "30%", fig.show="hold"}
-readd(asv_table) %>%
+ggpubr::ggarrange(venn_A, venn_B, ncol = 2)
+
+grid::grid.legend(
+  labels =  c("PacBio Short", "Illumina", "Ion Torrent"),
+  nrow = 1,
+  pch = 21,
+  gp = grid::gpar(fill = c("grey85", "lightblue", "lightcoral")),
+  vp = grid::viewport(y = unit(0.1, "npc"))
+)
+
+grid::grid.newpage()
+
+venn_C <- readd(asv_table) %>%
   mutate_if(is.numeric, ~ . / sum(.)) %>%
@@
       select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-     legend = list(
-         labels = c("Long amplicon", "Short amplicon")
-     ),
+      legend = FALSE,
+      labels = FALSE,
       quantities = .$label,
-     main = "ASV",
+      main = "c",
       fills = c("white", "slateblue1"))
@@
 
-readd(otu_table) %>%
+venn_D <- readd(otu_table) %>%
   mutate_if(is.numeric, ~ . / sum(.)) %>%
@@
       quantities = .$label,
-    main = "OTU",
+      main = "d",
       fills = c("white", "slateblue1")
@@
   }
+# ggpubr::ggarrange(venn_C, venn_D, ncol = 2)
+# grid::grid.legend(
+#   labels =  c("Long amplicon", "Short amplicon"),
+#   nrow = 1,
+#   pch = 21,
+#   gp = grid::gpar(fill = c("white", "slateblue1")),
+#   vp = grid::viewport(y = unit(0.2, "npc"))
+# )
 ```
@@
 
-The number of ASVs and OTUs shared between the different sequencing strategies are shown in Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in figure \@ref(fig:venn).
+The number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in Figure\ \@ref(fig:venn).
 ASVs shared between all datasets represented `r text_list(percent(readd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
@@
 
-Figure \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
-In all cases, the read counts for shared ASVs/OTUs were correlated ($R^2 = `r compR2("SH-2257", "is_057", "OTU")` to $).
+Figure\ \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
+In all cases, the read counts for shared ASVs/OTUs were correlated , with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")`.
 Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
@@
 
-Almost all of the short amplicons from all three technologies were between 240 and 375 bp long (Supplementary Figure \@ref(fig:full-length)a).
-Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure \@ref(fig:full-length)b).
-In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of
+Almost all of the short amplicons from all three technologies were between 240 and 375 bp long (Supplementary Figure\ \@ref(fig:full-length)a).
+Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
+In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of `r filter(readd(reads_table), region == "long") %$% reldist::wtd.quantile(length, weight = reads)` bp.
 
@@
 
-The fraction of reads and ASVs assigned to each taxonomic rank is shown in Figure \@ref(fig:tax-chart).
+The fraction of reads and ASVs assigned to each taxonomic rank is shown in Figure\ \@ref(fig:tax-chart).
 For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs.
@@
 ```{r tax-chart, fig.cap = "(ref:tax-chart)"}
-readd(tax_chart)
+readd(tax_chart_plot)
 ```
@@
 
-set.seed(2)
 theme <- theme_update(panel.border = element_blank())
-readd(taxdata) %>%
-  metacoder::heat_tree(
-    layout = "davidson-harel",
-    # initial_layout = "reingold-tilford",
-    node_size = rowMeans(.$data$tax_read[,-1]),
-    node_color = rowMeans(.$data$tax_read[,-1]),
-    node_label = replace_na(taxon_names, "unknown"),
-    node_size_range = c(0.002, .035),
-    node_size_axis_label = "Read abundance",
-    node_color_axis_label = "Read abundance",
-    node_label_size_range = c(0.01, 0.03),
-    edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_trans = "linear",
-    edge_size_axis_label = "ASV count",
-    edge_size_range = c(0.001, 0.03),
-    edge_color = rowMeans(.$data$tax_asv[,-1]),
-    edge_color_axis_label = "ASV count",
-    # aspect_ratio = 3/2,
-    make_node_legend = FALSE,
-    make_edge_legend = FALSE,
-    output_file = here("temp/heattree.pdf")
-  )
+readd(heattree)
 theme_set(theme)
@@
 
-Taxonomic composition of the sequenced soil community is summarized in Figure \@ref(fig:taxa).
+Taxonomic composition of the sequenced soil community is summarized in Figure\ \@ref(fig:taxa).
 According to the PHYLOTAX assignments, Fungi represented `r percent(fungi_read["PacBio_Long"], format = "f", digits = 1)` of the reads and `r percent(fungi_asv["PacBio_Long"], format = "f", digits = 1)` of the ASVs in the long amplicon library, compared to `r paste(percent(range(fungi_read[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads and `r paste(percent(range(fungi_asv[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs in the short amplicon library.
 Fungi categorized as ECM made up `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads and `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs in the long amplicon library, and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs in the short amplicon library.
-Assignment to kingdom Fungi and ECM status for different amplicons, sequencing technologies, reference databases, and algorithms are shown in Supplementary Figure \@ref(ref:ecm).
+Assignment to kingdom Fungi and ECM status for different amplicons, sequencing technologies, reference databases, and algorithms are shown in Supplementary Figure\ \@ref(fig:ecm).
 
@@
 ## Spatial analysis
-
-### Autocorrelation
 
@@
 
-Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure \@ref(fig:correlog).
-Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 3m for the total fungal community, and 1–2m for the ECM community.
+Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure\ \@ref(fig:correlog).
+Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 3 m for the total fungal community, and 1–2 m for the ECM community.
 In both cases, the greatest correlation magnitudes and distances are associated with Illumina sequencing of the short amplicon library, followed by PacBio sequencing of the long amplicon library, with PacBio sequencing of the short amplicon library detecting the least spatial structure.
 
-In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1m 
+In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1 m 
 distance class showing a significant correlation at $p < 0.05$.
 
-The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1 year and 0m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
+The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1 year and 0 m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
 For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0m and 1 year was not significant, but there was a negative correlation at time lag of 1 year and a distance of 1m for both sequencing technologies.
-This negative correlation was significant in each short amplicon correlogram except for the ECM community sequenced by Illumina.
-
-Distance-decay curves relating community dissimilarities and spatial distances are shown in Figure \@ref(fig:spatio).
+This negative correlation was significant in all short amplicon correlograms.
+
+Distance-decay curves relating community dissimilarities and spatial distances are shown in Figure\ \@ref(fig:spatio).
 For the cases where the Mantel correlogram detected a significant positive correlation, the best-fit exponential curve is also displayed; for the long amplicon library, where there was also a significant positive correlation after a time lag of one year, the best fit curve for time lag is also included.
-The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in \@ref(tab:variofit).
+The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in Table\ \@ref(tab:variofit).
 Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0 m was included in the 95% confidence interval.
@@
 
-(ref:variofit) Parameters for variogram fits shown in Figure \@ref(fig:spatio-temporal). Range: range at which exponential function is at 95% of its maximum value.
+(ref:variofit) Parameters for variogram fits shown in Figure\ \@ref(fig:spatio). Range: range at which exponential function is at 95% of its maximum value.
 
@@
 
-## Comparison of technologies
-
-  - length bias not clearly observed between PacBio and Ion Torrent short reads
-  - long reads had more very short ITS2 sequences Supplementary Figure \@ref(fig:ITS2-length)) -- these are mostly Alveolata (Supplementary Figure \@ref(fig:short-its))
-  - lots of ASVs/OTUs detected in only one tech
-
-  - very few PacBio OTUs undetected by Ion torrent/Illumina from same library; only rare OTUS undetected by PacBio -- sequencing depth seems to be the biggest effect.
-
-  - sequencing technologies yield comparable results (given read depth) with sufficient clustering
-  - not such good match with ASVs -- suggests DADA2 doesn't perform equally on different datasets
-
-  - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.
-
-## Use of Phylogeny to improve taxonomic identification
-
-  - Strict consensus usually increases the number of assigned sequences relative to any single method, except for family and genus level identifications.
-  - Different methods and databases bring mostly complementary, non-contradictory information.
-  - Unite (using full ITS) identifies a similar fraction of reads and ASVs as RDP using LSU.
-
-## Differences in taxonomic bias of different amplicons
-  - Long amplicon library includes more non-fungi, particularly Alveolates and Metazoans.
-  - Also more of a few isolated fungal genera -- Tremellodendron, Marcelleina, Tulasnella.
-  - Short amplicon recovers more Sordariomycetes, Saccharomycetes, Glomeromycota -- the latter two maybe due to sequencing depth.
-
-## Bias in ecological results using different amplicons and sequencing technologies
-  - UNIFRAC does not detect spatial structure within this relatively homogeneous community -- largest fraction of weighted branch length is between Ascomycota and Basidiomycota.
-  - choice of PCR primer (and possibly PCR replicate?) is more important for repeatability that choice of sequencing technology.
+## Comparison of sequencing strategies
+
+The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
+Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(readd(bioinf_table)["ITS2","PacBio_Short_reads"]/readd(rawcounts)["pb_483"], format = "f", digits = 1)`) compared to Illumina (`r percent(readd(bioinf_table)["ITS2","Illumina_Short_reads"]/readd(rawcounts)["SH-2257"], format = "f", digits = 1)`) or Ion Torrent (`r percent(readd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/readd(rawcounts)["is_057"], format = "f", digits = 1)`),
+the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
+
+DADA2 denoising to form ASVs may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:pb-vs-it)).
+The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
+
+Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, no such amplicons were sequenced by any of the technologies, and Illumina showed a slightly higher fraction of longer sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Supplementary Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
+
+`r percent(readd(venn_ASV)["1000", "frac_pb_500"])` of long amplicon reads beloged to ASVs which were not found in any short amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])`.
+Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences which were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
+Taxonomic assignments reveal that many of these non-shared sequences fall outside kingdom Fungi (Supplementary Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Supplementary Figure\ \@ref(fig:short-its)).
+Within Fungi, the short amplicon datasets recovered more Sordariomycetes, possibly because this group was represented by diverse but rare taxa which were not detected in the PacBio datasets due to the lower sequencing depth.
+Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella*, *Tremellodendron*, and *Marcelleina* in the long amplicon dataset, and *Xylaria* and *Myerozyma* in the short amplicon datasets.
+These differences may be due to primer mismatches in these groups.
+It is of note that the three aforementioned groups which were more common in the long amplicon dataset are ECM.
+However, neither amplicon recovered any sequences identifiable to *Cantharellus*, an ECM genus which is known from the area from fruitbodies.
+
+## Reconstruction of long amplicons from denoised subregions
+
+ASV recovery for long amplicons was dramatically improved (`r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")`) by denoising homologous subregions independently.
+Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs, so this approach should increase ASV recovery from these platforms as well.
+It may also be adaptable to Oxford Nanopore sequencing, which has hitherto not been successfully applied to complex community metabarcoding [@loit2019].
+
+## Taxonomic identification
+
+The RDP fungal training set and Unite performed comparably at placing long amplicon sequences.
+The Warcup database placed notably fewer sequences at all taxonomic levels for all datasets, probably in part due to the fact that only fungal sequences are included.
+However, even with this taken into account, IDTAXA performed very poorly with the Warcup database, placing less than 25% of ASVs to kingdom in all datasets.
+IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores [@murali2018a].
+
+@gdanetz2017 showed that a majority-rule consensus of three assignment algorithms can improve the fraction of sequences assigned as well as decrease the false assignment rate.
+Here we find that strict consensus, which should provide more conservative taxonomic assignments, also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications.
+This suggests that different assignment algorithms and databases bring mostly complementary, non-contradictory information at higher taxonomic levels.
+However, contradictory assignments between different methods is more common at lower taxonomic levels, which can be problematic because accurate assignment at the family or genus level is generally required for ecological guild assignment using FUNGuild.
+
+For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
+For instance, `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()`.
+This led to a corresponding increase in the fraction of reads assigned to a functional guild (Supplementary Figure\ \@ref(fig:ecm)).
+
+## Turnover rate
+
+Weighted UNIFRAC did not reliably detect spatial structure within this relatively homogeneous community.
+Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1 m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0-5700 m.
+This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure\ \@ref(fig:taxa)), which are both well represented in the majority of samples.
+UNIFRAC would be more suited at larger spatial scales and/or larger ecological gradients.
+
+Mantel correlograms based on the Bray-Curtis dissimilarity (Figure\ \@ref(fig:correlog)) revealed spatial autocorrelation in the soil fungal community at distance classes $\le 3$ m for both Illumina and PacBio using long and short amplicons, and in the ECM fungal community at distance classes $\le 2$ m for Illumina and PacBio long amplicons, and $\le 1$ m for the PacBio short amplicons.
+These values are similar to the values reported for ECM root tips in temperate forests by @lilleskov2004 and @pickles2012a, but substantially smaller than those observed by @tedersoo2011.
+Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 17--19 m.
+The 95% confidence interval was substantially wider than this variation, generally covering a range of 9--35 m.
+All of these values are smaller than the 65 m reported by @Bahram2013, also from an ECM woodland habitat in Benin.
+The PacBio short amplicon dataset shows a longer spatial range, of 30 m for the total fungal community and 42--45 m for the ECM community, in both cases with wide confidence intervals spanning approximately 15--60 m.
+It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
+
+The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0 m and 1 year.
+The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.1 years for the total fungal comunity and 5.5 years for the ECM community, but with overlapping confidence intervals.
+Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0 m and 1 year, but a weak negative correlation at 1 m and 1 year.
+The general shape of the correlograms were similar for long and short amplicon datasets.
+We hypothesize that two different processes may be at work with differing spatiotemporal scales, whose superposition result in this pattern.
+
+# Conclusion
+
+The choice of amplicon and sequencing technology did not seem to affect the results of the spatial analysis, provided sufficient sequencing depth.
+However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments.
+DADA2 ASV yield was initially poor for long reads, but this was improved through extraction of subregions, seperate denoising, and then reconstruction of full-length sequences.
+This suggests a hybrid approach using long-read sequencing to aquire long amplicon sequences for the local species pool, and cost-effective short-read sequencing to provide high sampling depth and sample number.
 
@@
 
-```{r region-length}
+```{r region-limits}
 regions %>%
@@
 
-(ref:region-length) Length distribution for each region extracted from long amplicons.  Filtering limits are shown as dashed vertical lines.
-
-```{r region-lengths, fig.cap = "(ref:region-length)", fig.height = 6}
+(ref:region-lengths) Length distribution for each region extracted from long amplicons.  Filtering limits are shown as dashed vertical lines.
+
+```{r region-lengths, fig.cap = "(ref:region-lengths)", fig.height = 6}
 readd(region_length_fig)
@@
 )
+```
+
+(ref:venn-table-asv) Correspondences between ASVs found by different sequencing strategies.
+Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+
+```{r venn-table-asv}
+readd(venn_ASV) %>%
+venn_table(ASVs, "(ref:venn-table-asv)")
+```
+
+(ref:venn-table-otu) Correspondences between 97% OTUs found by different sequencing strategies.
+Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTU frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+
+```{r venn-table-otu}
+readd(venn_OTU) %>%
+venn_table(OTUs, "(ref:venn-table-otu)")
 ```
@@
 Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
-Refer to Figure \@ref(fig:taxa) for complete taxon labeling.
+Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.
 
 ```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare)"}
-taxdata2 <- readd(taxdata)
-# taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
-# taxdata2$data$read_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = read_cols)
-# taxdata2$data$tax_asv <- metacoder::calc_taxon_abund(taxdata2, "asv_table", cols = asv_cols)
-# taxdata2$data$tax_read <- metacoder::calc_taxon_abund(taxdata2, "read_table", cols = read_cols)
-taxdata2$data$diff_read <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_read",
-  cols = names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_.+_.+_.+_.+"),
-  groups = names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_.+_.+_.+_.+") %>%
-    str_replace("reads_.+_(.+)_.+_.+", "\\1"),
-  func = function(abund1, abund2) {
-    list(read_ratio = log10(mean(abund1) / mean(abund2)))
-  }
-  )
-taxdata2$data$diff_asv <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_asv",
-  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
-  groups = names(taxdata2$data$tax_asv) %>%
-    keep(startsWith, "ASVs_") %>%
-    str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
-  func = function(abund1, abund2) {
-    list(asv_ratio = log10(mean(abund1) / mean(abund2)))
-  }
-)
-
-set.seed(2)
 oldtheme <- theme_update(panel.border = element_blank())
-taxdata2 %>%
-  metacoder::heat_tree(
-    layout = "davidson-harel",
-    # initial_layout = "reingold-tilford",
-
-    node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.002, .035),
-    node_size_axis_label = "Read abundance",
-
-    node_color = read_ratio,
-    node_color_range = metacoder::diverging_palette(),
-    node_color_trans = "linear",
-    node_color_axis_label = "Read abundance ratio",
-    node_color_interval = c(-6, 6),
-
-    edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.001, 0.03),
-    edge_size_trans = "linear",
-    edge_size_axis_label = "ASV count",
-
-    edge_color = asv_ratio,
-    edge_color_range = metacoder::diverging_palette(),
-    edge_color_trans = "linear",
-    edge_color_axis_label = "ASV count ratio",
-    edge_color_interval = c(-3, 3),
-
-    node_label = ifelse(
-      pmax(abs(.$data$diff_read$read_ratio),
-           abs(.$data$diff_asv$asv_ratio)) > 0.5,
-      taxon_names,
-      ""
-    ),
-    # aspect_ratio = 3/2,
-    node_label_size_range = c(0.010, 0.03)
-    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2))
-  )
+readd(heattree_length_compare)
 theme_set(oldtheme)
@@
 
-```{r ECM, fig.cap = "(ref:ecm)"}
+```{r ecm, fig.cap = "(ref:ecm)"}
   readd(taxon_reads) %>%
@@
 
-```{r plate-loading}
+```{r plate-loading, include=FALSE}
 readd(proto_physeq) %>% {
@@
 
-```{r}
-physeq <- readd(proto_physeq) %>%
-  `sample_data<-`(inset2(
-    sample_data(.),
-    "qual",
-    value = fct_collapse(sample_data(.)$qual, "-" = c("", "*"))
-    )) %>%
-  subset_samples(!is.na(site)) %>%
-  merge_samples(group = glue::glue_data(
-    sample_data(.),
-    "{year}_{buffer}_{tech}_{amplicon}"
-  ))
-
-ranks <- c("root", "kingdom", "phylum", "class", "order", "family", "genus")
-physeq <- readd(taxon_reads) %>%
-  mutate(root = "Root") %>%
-  group_by(label) %>%
-  arrange(amplicon, reference, Algorithm, .by_group = TRUE) %>%
-  summarize_at(ranks, first) %>%
-  left_join(
-    tibble(label = taxa_names(physeq)),
-    .,
-    by = "label"
-  ) %>% {
-    split(select(., -label), .$label)
-  } %>%
-  lapply(unlist) %>%
-  build_tax_table() %>%
-  `tax_table<-`(physeq, .) %>%
-  tax_glom(taxrank = "family", NArm = FALSE)
-```
-
 (ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
@@
 ```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)"}
-taxmap <- metacoder::parse_phyloseq(physeq)
-taxmap$data$otu_table <- metacoder::calc_obs_props(taxmap, "otu_table")
-taxmap$data$read_count <- metacoder::calc_taxon_abund(taxmap, "otu_table")
-taxmap$data$read_diff <- metacoder::compare_groups(
-  taxmap,
-  "read_count",
-  cols = names(taxmap$data$read_count) %>%
-    keep(endsWith, "Long"),
-  groups = names(taxmap$data$read_count) %>%
-    keep(endsWith, "Long") %>%
-    str_match("(2015|2016)_(Xpedition|LifeGuard)") %>% {
-      paste(.[,3], .[,2])
-    },
-  func = function(abund1, abund2) {
-    list(read_ratio = log10(mean(abund1) / mean(abund2)) %>% ifelse(is.nan(.), 0, .))
-  },
-  combinations = list(
-    c("Xpedition 2016", "Xpedition 2015"),
-    c("Xpedition 2016", "LifeGuard 2016"),
-    c("Xpedition 2015", "LifeGuard 2016")
-  )
-)
 oldtheme <- theme_update(panel.border = element_blank())
-taxa::filter_taxa(
-  taxmap,
-  taxmap$data$read_count$`2015_Xpedition_PacBio_Long` > 0 |
-    taxmap$data$read_count$`2016_Xpedition_PacBio_Long` > 0 |
-    taxmap$data$read_count$`2016_LifeGuard_PacBio_Long` > 0,
-  drop_obs = TRUE,
-  reassign_obs = FALSE
-  ) %>%
-metacoder::heat_tree_matrix(
-  "read_diff",
-  node_label = taxon_names,
-  node_color = read_ratio,
-  node_size = rowMeans(.$data$read_count[-1]),
-  node_size_axis_label = "Read fraction",
-  node_color_range = metacoder::diverging_palette(),
-  node_color_axis_label = "Log10 Read Ratio",
-  node_color_interval = c(-3, 3),
-  node_color_trans = "linear",
-  edge_size = rowMeans(.$data$read_count[-1]),
-  edge_size_range = c(0, 0.03),
-  edge_size_trans = "linear",
-  edge_size_axis_label = "ASV fraction",
-  node_size_range = c(0, 0.03),
-  node_label_size_range = c(0.005, 0.03),
-  layout = "davidson-harel",
-  initial_layout = "reingold-tilford",
-  output_file = here("temp/compare_long.pdf")
-)
+readd(buffer_compare_long)
 theme_set(oldtheme)
-
-```
-
-(ref:buffer-compare-short) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
+```
+
+(ref:buffer-compare-short) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for short amplicon libraries.
 
 ```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
-taxmap$data$read_diff <- metacoder::compare_groups(
-  taxmap,
-  "read_count",
-  cols = names(taxmap$data$read_count) %>%
-    keep(endsWith, "Short"),
-  groups = names(taxmap$data$read_count) %>%
-    keep(endsWith, "Short") %>%
-    str_match("(2015|2016)_(Xpedition|LifeGuard)") %>% {
-      paste(.[,3], .[,2])
-    },
-  func = function(abund1, abund2) {
-    list(read_ratio = log10(mean(abund1) / mean(abund2)) %>% ifelse(is.nan(.), 0, .))
-  },
-  combinations = list(
-    c("Xpedition 2016", "Xpedition 2015"),
-    c("Xpedition 2016", "LifeGuard 2016"),
-    c("Xpedition 2015", "LifeGuard 2016")
-  )
-)
-taxa::filter_taxa(
-  taxmap,
-  taxmap$data$read_count %>% select(ends_with("Short")) %>% rowSums() > 0,
-  drop_obs = TRUE,
-  reassign_obs = FALSE
-  ) %>%
-metacoder::heat_tree_matrix(
-  "read_diff",
-  node_label = taxon_names,
-  node_color = read_ratio,
-  node_size = rowMeans(.$data$read_count[-1]),
-  node_size_axis_label = "Read fraction",
-  node_color_range = metacoder::diverging_palette(),
-  node_color_axis_label = "Log10 Read Ratio",
-  node_color_interval = c(-3, 3),
-  node_color_trans = "linear",
-  edge_size = rowMeans(.$data$read_count[-1]),
-  edge_size_range = c(0, 0.03),
-  edge_size_trans = "linear",
-  edge_size_axis_label = "ASV fraction",
-  node_size_range = c(0, 0.03),
-  node_label_size_range = c(0.005, 0.03),
-  layout = "davidson-harel",
-  initial_layout = "reingold-tilford",
-  output_file = here("temp/compare_short.pdf")
-)
-
-```
+readd(buffer_compare_short)
+theme_set(oldtheme)
+
+```

### 2020-02-13 scripts/drake_plan_light.R

@@ -610,62 +610,92 @@ plan2 <- drake_plan(
   rawcounts = readcounts %>%
     filter(is.na(step), is.na(well)) %>%
     group_by(seq_run) %>%
-    summarize(nreads =  prettyNum(sum(nreads), big.mark = " ")) %>%
+    summarize(
+      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      nreads = sum(nreads)
+    ) %>%
     deframe(),
   
   demuxcounts = readcounts %>%
     filter(is.na(step), !is.na(well)) %>%
     group_by(seq_run) %>%
-    summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
+    summarize(
+      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      nreads = sum(nreads)
+    ) %>%
     deframe(),
   
   filtercounts_full = readcounts %>%
     filter(step == "filter", is.na(region) | region %in% c("long", "short")) %>%
     group_by(seq_run, region) %>%
-    summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
+    summarize(
+      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      nreads = sum(nreads)
+    ) %>%
     select(seq_run, nreads) %>%
     deframe(), 
   
   filtercounts_ITS2 = readcounts %>%
     filter(step == "filter", region == "ITS2") %>%
     group_by(seq_run, region) %>%
-    summarize(nreads = prettyNum(sum(nreads), big.mark = " ")) %>%
+    summarize(
+      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      nreads = sum(nreads)
+    ) %>%
     select(seq_run, nreads) %>%
     deframe(),
     
   regioncounts = readcounts %>%
-    filter(step == "lsux") %>%
+    filter(step == "lsux", region == "5_8S") %>%
     group_by(seq_run) %>%
-    summarize(nreads = prettyNum(
-      sum(nreads) %/% if (seq_run[1] == "pb_500") 14 else 4,
-      big.mark = " ")
+    summarize(
+      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      nreads = sum(nreads)
     ) %>%
     deframe(),
+  
+  asvcounts =
+    pivot_longer(asv_table, -1, names_to = "seq_run", values_to = "reads") %>%
+    filter(reads > 0) %>%
[…]

### 2020-02-14 writing/transect_paper.Rmd
Move methods for Illumina denoising to end of section, clarify that this was different from PacBio/Ion Torrent and why; mention problems with DNA extracts from Ang 2015.

@@
 cache <- drake_cache(here(".light"))
-readd <- function(..., cache = drake_cache(here(".light"))) {
+rdd <- function(..., cache = drake_cache(here(".light"))) {
   force(cache)
@@
 }
-loadd <- function(..., cache = drake_cache(here(".light"))) {
+ldd <- function(..., cache = drake_cache(here(".light"))) {
   force(cache)
@@
 
-Forward and reverse Illumina reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
-Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using seperate error models and pseudo-pooling, and then forward and reverse reads were merged.
-
 We attempted to denoise both long and short PacBio amplicons using `DADA2` according to the steps outlined in  the supplementary information in @callahan2019.
-However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
+However, only `r rdd(region_table)$ASVs[rdd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
 We believe that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
@@
 Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = " ")`.
-Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r readd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
-In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
+For the combination of long reads (median $L=`r rdd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(rdd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - rdd(demuxqual)['pb_500'])^rdd(demuxlength)['pb_500'])), big.mark = " ")`.
+Given the high diversity relative to sequencing depth in this study (`r rdd(region_table)$ASVs[rdd(region_table)$seq_run == "pb_483" & rdd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r rdd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
+In comparison, the equivalent requirement for the short amplicon ($L=`r rdd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r rdd(demuxqual)['pb_483']`$) is only `r round(2/(1 - rdd(demuxqual)['pb_483'])^rdd(demuxlength)['pb_483'], 1)` reads.
 We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
@@
 
+Because the Illumina dataset consisted of paired-end reads, regions were not extracted prior to denoising.
+ASVs were instead generated according to a standard workflow for `DADA2`.
+Demultiplexed reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
+Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using seperate error models and pseudo-pooling, and then forward and reverse reads were merged.
+The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the other technologies.
+
 ### Taxonomy assignment
@@
 
+DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Supplementary Figure \@ref(fig:sample_depth).
+Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100 reads.
+Consequently Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
+
 The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table\ \@ref(tab:bioinfo).
-Long (short) amplicon sequencing with PacBio yielded `r readd(rawcounts)["pb_500"]` (`r readd(rawcounts)["pb_483"]`) circular consensus reads, of which `r readd(demuxcounts)["pb_500"]` (`r readd(demuxcounts)["pb_483"]`) were successfully trimmed and demultiplexed.
-Illumina sequencing yielded `r readd(rawcounts)["SH-2257"]` read pairs, of which `r readd(demuxcounts)["SH-2257"]` were successfully trimmed and demultiplexed, and `r readd(filtercounts_full)["SH-2257"]` passed quality filtering.
-Ion Torrent sequencing yielded `r readd(demuxcounts)["is_057"]` demultiplexed reads.
-`r readd(filtercounts_full)['pb_500']` (`r percent(readd(filtercounts_full)['pb_500']/readd(demuxcounts)['pb_500'], digits = 1, format = "f")`) of long amplicon PacBio reads passed length and quality filtering stages, compared with `r readd(filtercounts_full)['pb_483']` (`r percent(readd(filtercounts_full)['pb_483']/readd(demuxcounts)['pb_483'], digits = 1, format = "f")`) of short amplicon PacBio reads, `r readd(filtercounts_full)['SH-2257']` (`r percent(readd(filtercounts_full)['SH-2257']/readd(demuxcounts)['SH-2257'], digits = 1, format = "f")`) of Illumina read pairs, and `r readd(filtercounts_full)['is_057']` (`r percent(readd(filtercounts_full)['is_057']/readd(demuxcounts)['is_057'], digits = 1, format = "f")`) of Ion Torrent reads.
+Long (short) amplicon sequencing with PacBio yielded `r rdd(rawcounts)["pb_500"]` (`r rdd(rawcounts)["pb_483"]`) circular consensus reads, of which `r rdd(demuxcounts)["pb_500"]` (`r rdd(demuxcounts)["pb_483"]`) were successfully trimmed and demultiplexed.
+Illumina sequencing yielded `r rdd(rawcounts)["SH-2257"]` read pairs, of which `r rdd(demuxcounts)["SH-2257"]` were successfully trimmed and demultiplexed, and `r rdd(filtercounts_full)["SH-2257"]` passed quality filtering.
+Ion Torrent sequencing yielded `r rdd(demuxcounts)["is_057"]` demultiplexed reads.
+`r rdd(filtercounts_full)['pb_500']` (`r percent(rdd(filtercounts_full)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")`) of long amplicon PacBio reads passed length and quality filtering stages, compared with `r rdd(filtercounts_full)['pb_483']` (`r percent(rdd(filtercounts_full)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`) of short amplicon PacBio reads, `r rdd(filtercounts_full)['SH-2257']` (`r percent(rdd(filtercounts_full)['SH-2257']/rdd(demuxcounts)['SH-2257'], digits = 1, format = "f")`) of Illumina read pairs, and `r rdd(filtercounts_full)['is_057']` (`r percent(rdd(filtercounts_full)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`) of Ion Torrent reads.
 
@@
 #TODO asterisks for mapped ASVs
-# loadd(bioinf_table)
 options(knitr.kable.NA = '–')
 
-readd(bioinf_table) %>%
+rdd(bioinf_table) %>%
   set_names(gsub(".*_", "", names(.))) %>%
@@
       " " = 1,
-      readd(bioinf_header)$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
+      rdd(bioinf_header)$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
     )
@@
       " " = 1,
-      rle(readd(bioinf_header)$tech) %$% set_names(lengths, values)
+      rle(rdd(bioinf_header)$tech) %$% set_names(lengths, values)
     )
@@
 
-The 5.8S rDNA was located in `r readd(regioncounts)['pb_500']` long amplicon PacBio reads (`r percent(readd(regioncounts)['pb_500']/readd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total), `r readd(regioncounts)['pb_483']` short amplicon PacBio reads (`r percent(readd(regioncounts)['pb_483']/readd(demuxcounts)['pb_483'], digits = 1, format = "f")`), and `r readd(regioncounts)['is_057']` Ion Torrent reads (`r percent(readd(regioncounts)['is_057']/readd(demuxcounts)['is_057'], digits = 1, format = "f")`).
-Length and quality filtering resulted in `r readd(filtercounts_ITS2)['pb_500']` (`r percent(readd(filtercounts_ITS2)['pb_500']/readd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total) ITS2 sequences from the long amplicon PacBio reads, `r readd(filtercounts_ITS2)['pb_483']` (`r percent(readd(filtercounts_ITS2)['pb_483']/readd(demuxcounts)['pb_483'], digits = 1, format = "f")`) ITS2 sequences from the short amplicon PacBio reads, and `r readd(filtercounts_ITS2)['is_057']` (`r percent(readd(filtercounts_ITS2)['is_057']/readd(demuxcounts)['is_057'], digits = 1, format = "f")`) ITS2 sequences from the Ion torrent reads.
+The 5.8S rDNA was located in `r rdd(regioncounts)['pb_500']` long amplicon PacBio reads (`r percent(rdd(regioncounts)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total), `r rdd(regioncounts)['pb_483']` short amplicon PacBio reads (`r percent(rdd(regioncounts)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`), and `r rdd(regioncounts)['is_057']` Ion Torrent reads (`r percent(rdd(regioncounts)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`).
+Length and quality filtering resulted in `r rdd(filtercounts_ITS2)['pb_500']` (`r percent(rdd(filtercounts_ITS2)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total) ITS2 sequences from the long amplicon PacBio reads, `r rdd(filtercounts_ITS2)['pb_483']` (`r percent(rdd(filtercounts_ITS2)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`) ITS2 sequences from the short amplicon PacBio reads, and `r rdd(filtercounts_ITS2)['is_057']` (`r percent(rdd(filtercounts_ITS2)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`) ITS2 sequences from the Ion torrent reads.
 The length distribution of different regions extracted from the long amplicon are shown in Supplementary Figure\ \@ref(fig:region-lengths).
@@
 ```{r venn, fig.cap="(ref:venn)", fig.show="hold", out.width = "90%"}
-venn_A <- venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+venn_A <- venndata(rdd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
   vennplot_data(ASVs) %>% {
@@
 
-venn_B <- venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+venn_B <- venndata(rdd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
   vennplot_data(OTUs) %>% {
@@
 
-venn_C <- readd(asv_table) %>%
+venn_C <- rdd(asv_table) %>%
   mutate_if(is.numeric, ~ . / sum(.)) %>%
@@
 
-venn_D <- readd(otu_table) %>%
+venn_D <- rdd(otu_table) %>%
   mutate_if(is.numeric, ~ . / sum(.)) %>%
@@
   read_comparison(
-    readd(multi_table),
-    readd(comparisons),
+    rdd(multi_table),
+    rdd(comparisons),
     type = "ASV"
@@
   read_comparison(
-    readd(multi_table),
-    readd(comparisons),
+    rdd(multi_table),
+    rdd(comparisons),
     type = "OTU"
@@
 The number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in Figure\ \@ref(fig:venn).
-ASVs shared between all datasets represented `r text_list(percent(readd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
-These fractions increased to `r text_list(percent(readd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
-In particular, `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
-In contrast, `r percent(readd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
+ASVs shared between all datasets represented `r text_list(percent(rdd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
+These fractions increased to `r text_list(percent(rdd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+In particular, `r text_list(percent(colSums(mutate_all(rdd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
+In contrast, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
 
@@
 Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
-In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of `r filter(readd(reads_table), region == "long") %$% reldist::wtd.quantile(length, weight = reads)` bp.
+In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of `r filter(rdd(reads_table), region == "long") %$% reldist::wtd.quantile(length, weight = reads)` bp.
 
@@
 ```{r tax-chart, fig.cap = "(ref:tax-chart)"}
-readd(tax_chart_plot)
+rdd(tax_chart_plot)
 ```
@@
 ```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
-fungi_read <- readd(taxdata) %>%
+fungi_read <- rdd(taxdata) %>%
   taxa::filter_taxa(
@@
 
-fungi_asv <- readd(taxdata) %>%
+fungi_asv <- rdd(taxdata) %>%
   taxa::filter_taxa(
@@
 
-ecm_counts <- readd(taxon_reads) %>%
+ecm_counts <- rdd(taxon_reads) %>%
   filter(Algorithm == "PHYLOTAX") %>%
@@
 theme <- theme_update(panel.border = element_blank())
-readd(heattree)
+rdd(heattree)
 theme_set(theme)
@@
 ```{r heattree-db-compare, fig.cap = "(ref:heattree-db-compare)", include = FALSE}
-taxdata2 <- readd(taxdata)
+taxdata2 <- rdd(taxdata)
 # taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
@@
 *A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair.
-The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
+The abundance of the second ASV was `r rdd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
 > 0) %>% percent(format = "f", digits = 2) %>% paste(collapse = " and ")` that of the primary ASV in the two Illumina positive controls.
@@
 bisporus* ASV was the most abundant ASV in the long-amplicon PacBio dataset and the Ion Torrent dataset, and the fifth most abundant ASV in the short-amplicon PacBio dataset, and the third most abundant ASV in the Illumina dataset.
-This ASV was also found in `r readd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, and represented `r readd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
+This ASV was also found in `r rdd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, and represented `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
 
@@
   ) %>%
-  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
+  mutate_at("target", purrr::map, rdd, character_only = TRUE) %>%
   mutate_at("target", purrr::map, "mantel.res") %>%
@@
 
-readd(variog_empir) %>%
+rdd(variog_empir) %>%
   filter(algorithm == "PHYLO") %>%
@@
   # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
-  geom_line(data = filter(readd(variog_fit), timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
-  geom_line(data = filter(readd(variog_fit), timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
+  geom_line(data = filter(rdd(variog_fit), timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
+  geom_line(data = filter(rdd(variog_fit), timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
   scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
@@
 ```{r variofit, results="asis"}
-readd(variog_data) %>%
+rdd(variog_data) %>%
   select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
@@
 The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
-Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(readd(bioinf_table)["ITS2","PacBio_Short_reads"]/readd(rawcounts)["pb_483"], format = "f", digits = 1)`) compared to Illumina (`r percent(readd(bioinf_table)["ITS2","Illumina_Short_reads"]/readd(rawcounts)["SH-2257"], format = "f", digits = 1)`) or Ion Torrent (`r percent(readd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/readd(rawcounts)["is_057"], format = "f", digits = 1)`),
+Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 1)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 1)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 1)`),
 the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
@@
 
-`r percent(readd(venn_ASV)["1000", "frac_pb_500"])` of long amplicon reads beloged to ASVs which were not found in any short amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])`.
+`r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` of long amplicon reads beloged to ASVs which were not found in any short amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
 Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences which were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
@@
 
-ASV recovery for long amplicons was dramatically improved (`r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")`) by denoising homologous subregions independently.
+ASV recovery for long amplicons was dramatically improved (`r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` to `r percent((rdd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")`) by denoising homologous subregions independently.
 Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs, so this approach should increase ASV recovery from these platforms as well.
@@
 For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
-For instance, `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()`.
+For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()`.
 This led to a corresponding increase in the fraction of reads assigned to a functional guild (Supplementary Figure\ \@ref(fig:ecm)).
@@
 ```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
-readd(phylotax_fig)
+rdd(phylotax_fig)
 ```
@@
 ggpubr::ggarrange(
-  readd(full_dens_short),
-  readd(full_ecdf_short),
-  readd(full_dens_long),
-  readd(full_ecdf_long),
+  rdd(full_dens_short),
+  rdd(full_ecdf_short),
+  rdd(full_dens_long),
+  rdd(full_ecdf_long),
   ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
@@
 ```{r regions, fig.cap="(ref:region-asv)"}
-readd(region_table) %>%
+rdd(region_table) %>%
   dplyr::filter(seq_run == "pb_500") %>%
-  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", is.na(step), !is.na(well))$nreads)) %>%
+  mutate(reads = reads / sum(filter(rdd(readcounts), seq_run == "pb_500", is.na(step), !is.na(well))$nreads)) %>%
   dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
@@
 ```{r region-lengths, fig.cap = "(ref:region-lengths)", fig.height = 6}
-readd(region_length_fig)
+rdd(region_length_fig)
 ```
@@
 ggpubr::ggarrange(
-  readd(its2_dens),
-  readd(its2_ecdf),
+  rdd(its2_dens),
+  rdd(its2_ecdf),
   ncol = 1,
@@
 ```{r venn-table-asv}
-readd(venn_ASV) %>%
+rdd(venn_ASV) %>%
 venn_table(ASVs, "(ref:venn-table-asv)")
@@
 ```{r venn-table-otu}
-readd(venn_OTU) %>%
+rdd(venn_OTU) %>%
 venn_table(OTUs, "(ref:venn-table-otu)")
@@
 ```{r short-its, fig.cap = "(ref:short-its)"}
-readd(big_seq_table_ITS2) %>%
+rdd(big_seq_table_ITS2) %>%
   colSums() %>%
@@
   mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
+  semi_join(rdd(taxon_reads), ., by = c("label" = "ITS2")) %>%
   # group_by(ITS2) %>%
@@
 Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
-Values are log_10_ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
+Values are log~10~ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
 Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
@@
 oldtheme <- theme_update(panel.border = element_blank())
-readd(heattree_length_compare)
+rdd(heattree_length_compare)
 theme_set(oldtheme)
@@
 ```{r ecm, fig.cap = "(ref:ecm)"}
-  readd(taxon_reads) %>%
+  rdd(taxon_reads) %>%
   mutate_at("reference", replace_na, "All") %>%
@@
 Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
-Values are log_10_ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
+Values are log~10~ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
 
@@
 oldtheme <- theme_update(panel.border = element_blank())
-readd(ecm_heattree)
+rdd(ecm_heattree)
 theme_set(oldtheme)
@@
 Dotted horizontal line at 100 reads indicates cutoff for inclusion in community analysis.
-2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using LifeGuard.
+2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.
 
 ```{r sample-depth, fig.cap = "(ref:sample-depth)"}
-reads_plot <- readd(qstats_n) %>%
+reads_plot <- rdd(qstats_n) %>%
   mutate(file = str_replace(file, ".*/", "") %>%
@@
   left_join(
-    readd(proto_physeq) %>% {
+    rdd(proto_physeq) %>% {
       mutate(
@@
     type = factor(type, levels = c("raw_reads", "reads")),
+    buffer = factor(buffer, levels = c("Xpedition", "LifeGuard")),
+    tech = factor(tech, levels = c("PacBio", "Illumina")),
     "Sample group" = paste(year, buffer) %>%
@@
 
-conc_plot <- readd(proto_physeq) %>% 
+conc_plot <- rdd(proto_physeq) %>% 
         sample_data(.) %>%
@@
     amplicon = ifelse(step == "dna", "", amplicon),
+    buffer = factor(buffer, levels = c("Xpedition", "LifeGuard")),
     step = factor(step, levels = c("dna", "pcr"), labels = c("DNA", "PCR product")),
@@
 ```{r plate-loading, include=FALSE}
-readd(proto_physeq) %>% {
+rdd(proto_physeq) %>% {
   mutate(
@@
 oldtheme <- theme_update(panel.border = element_blank())
-readd(buffer_compare_long)
+rdd(buffer_compare_long)
 theme_set(oldtheme)
@@
 ```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
-readd(buffer_compare_short)
+rdd(buffer_compare_short)
 theme_set(oldtheme)

### 2020-02-16 writing/transect_paper.Rmd
raw Ion Torrent reads in in-line R code, table spacing

@@
 (ref:region-asv) Plot of total number of ASVs recovered vs.
-fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
+fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
 Some regions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU.
@@
   dplyr::filter(seq_run == "pb_500") %>%
-  mutate(reads = reads / sum(filter(rdd(readcounts), seq_run == "pb_500", is.na(step), !is.na(well))$nreads)) %>%
+  mutate(reads = reads / sum(filter(rdd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
   dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
@@
 Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+
+\setlength{\tabcolsep}{1pt}
 

### 2020-02-18 writing/transect_paper.Rmd
clean up before sending to Mohammad and Nourou

@@
 
+tinytex::tlmgr_install("pdfcrop")
+
 knitr::opts_chunk$set(
@@
   warning = FALSE,
-  external = TRUE
+  external = TRUE,
+  out.width = "70%",
+  fig.align = "center"
 )
@@
 Spatial turnover rate is one important characteristic of ecological communities, which is relevant for biodiversity assessment and sampling design.
-We investigated turnover rate in an ECM fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400 bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600 bp) with PacBio RS II.
+We investigated turnover rate in an ECM fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
 Resulting turnover rates depended more on statistical method than sequencing technology.
-We found significant autocorrelation in distance classes <3 m using a Mantel correlogram, and a best-fit distance-decay curve range of 17--19 m, both shorter than previously reported for similar habitats.
+We found significant autocorrelation in distance classes <3\ m using a Mantel correlogram, and a best-fit distance-decay curve range of 17--19\ m, both shorter than previously reported for similar habitats.
 The use of long-amplicon sequencing enables construction of a phylogenetic tree from metabarcoding reads, which facilitates taxonomic identification of sequences.
@@
 Several studies have investigated the turnover scale in ECM forests, but so far they have been based on analysis of ECM root tips, rather than bulk soil.
-@lilleskov2004 found autocorrelation only at ranges <2.6 m at most sites in temperate forests, based on Sanger sequencing.
-Similarly, @pickles2012a found autocorrelation at distances <3.4 m, also in temperate forests, based on T-RFLP.
+@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites in temperate forests, based on Sanger sequencing.
+Similarly, @pickles2012a found autocorrelation at distances <3.4\ m, also in temperate forests, based on T-RFLP.
 Two studies have included data from tropical African woodlands and wooded savannahs.
-@tedersoo2011 found autocorrelation at ranges <10 m in Miombo woodland based on Sanger sequencing.
-However, this study was based on irregular sampling with a minimum spacing of 8 m, so this as the smallest distance class.
-@Bahram2013 performed a meta-analysis of turnover scale in ECM communities, including data from Benin, where they found turnover range of 65 m based on distance-decay curves.
+@tedersoo2011 found autocorrelation at ranges <10\ m in Miombo woodland based on Sanger sequencing.
+However, this study was based on irregular sampling with a minimum spacing of 8\ m, so this as the smallest distance class.
+@Bahram2013 performed a meta-analysis of turnover scale in ECM communities, including data from Benin, where they found turnover range of 65\ m based on distance-decay curves.
 They concluded that the range of distance decay is greater in tropical forests than temperate forests.
@@
 
-Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30 km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
 Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25 soil samples were collected along a linear transect at intervals of 1 m in May 2015.
-One third of the sample locations (3 m spacing) were resampled one year later in June 2016.
+At each site, 25\ soil samples were collected along a linear transect at intervals of 1\ m in May 2015.
+One third of the sample locations (3\ m spacing) were resampled one year later in June 2016.
 For each sample, any coarse organic debris <!--Do you have any definition?
@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 50 mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0 mL microtube containing 750 mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
+Approximately 50\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
 <!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->
@@
 After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
-Samples preserved using LifeGuard were first centrifuged at 10000×g for 1 minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
+Samples preserved using LifeGuard were first centrifuged at 10000×g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
 DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
@@
 Two different fragments of the nuclear rDNA were amplified (Figure\ \@ref(fig:rDNA)).
-The short amplicon (approximately 300 bp) targetted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
-The long amplicon (approximately 1500 bp) targetted the full ITS region including the 5.8S rDNA and approximately 950 bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+The short amplicon (approximately 300\ bp) targetted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
+The long amplicon (approximately 1500\ bp) targetted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 
 The gITS7 primers for the short amplicon were indexed for multiplexing (**Supplementary File**).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1 U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7 ng purified DNA in Dream *Taq* buffer.
-The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
+The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
 Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity, which were pooled after amplification.
+
+(ref:rDNA-s) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions
 
@@
 <!--If limited by space, this could be in supp, and then just described in text. -MR-->
-```{tikz rDNA, fig.cap = "(ref:rDNA)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
+```{tikz rDNA, fig.cap = "(ref:rDNA)", fig.scap = "(ref:rDNA-s)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
 
@@
 Both primers for the long amplicon were indexed for multiplexing (**Supplementary File**).
-PCR was performed as for the short amplicons, but with 500 µM of each of the two primers.
-Reaction conditions were 10 min at 95°, 30 cycles of 45 s at 95°, 45 s at 59°, and 90 s at 72°, and finally 10 min at 72°.
+PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
+Reaction conditions were 10\ min at 95°, 30\ cycles of 45\ s at 95°, 45\ s at 59°, and 90\ s at 72°, and finally 10\ min at 72°.
 Each reaction was performed in three technical replicates as for short amplicons.
@@
 Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
-An aliquot of 100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
+An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
 Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
-for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752 37 UPPSALA, Sweden).
+for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752\ 37 UPPSALA, Sweden).
 Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
 
-Additionally the short amplicon libraries were combined and sequenced using an Ion S5  (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300 bp at the SNP&SEQ Technology Platform (Dept.
-of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751 44 UPPSALA, Sweden).
+Additionally the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
+of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden).
 The Illumina library was pooled with samples for another project, with one half of the reads from one lane devoted to this project.
@@
 First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
-No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
+No attempt was made to remove the approximately 12\ bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
 <!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
@@
 
-Unique LSU regions from long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10 iterations of progressive alignment and conserved RNA secondary structure calculation and 10 refinement iterations.
-<!-- The LSU tree ends up not looking as nice as the tree which includes ITS (I don't know why, except that maybe 5.8S is more helpful than ITS is harmful?) so I'm using that.
-Maybe I don't need to mention the LSU-only tree.-->
+Full length long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10\ iterations of progressive alignment and conserved RNA secondary structure calculation and 10\ refinement iterations.
 This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns
@@
 after this position, as also observed in several fungal species by @holst-jensen1999.
-Full length long amplicon ASVs (including ITS1, 5.8S, and ITS2 regions) were aligned and truncated in the same way.
 
@@
 Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
-Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m.
+Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12\ m.
 Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
@@
 
-DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Supplementary Figure \@ref(fig:sample_depth).
-Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100 reads.
+DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Supplementary Figure \@ref(fig:sample-depth).
+Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads.
 Consequently Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
@@
 
-(ref:readcounts) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages.
-*Raw*: raw reads as delivered by sequencing center;
+(ref:readcounts-s) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages
+
+(ref:readcounts) *Raw*: raw reads as delivered by sequencing center;
 *Trim*: reads with primers and demultiplexing barcodes removed;
@@
     booktabs = TRUE,
-    caption = "(ref:readcounts)",
+    caption = "(ref:readcounts-s). (ref:readcounts)",
+    caption.short = "(ref:readcounts-s)",
     align = "rrrrrr",
@@
 
+(ref:venn-s) Venn diagrams of shared ASVs and OTUs between different sequencing technologies
+
 (ref:venn) Venn diagrams showing shared ITS2-based ASVs (*a*, *c*) and 97% OTUs (*b*, *d*) between different sequencing technologies from the same short amplicon library (*a*, *b*), and between long and short amplicon libraries (*c*, *d*).
@@
 
-```{r venn, fig.cap="(ref:venn)", fig.show="hold", out.width = "90%"}
+```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", out.width = "70%", fig.height = 4, fig.align="center", results="asis"}
 venn_A <- venndata(rdd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
@@
   gp = grid::gpar(fill = c("grey85", "lightblue", "lightcoral")),
-  vp = grid::viewport(y = unit(0.1, "npc"))
+  vp = grid::viewport(y = unit(0.05, "npc"))
 )
@@
   }
-# ggpubr::ggarrange(venn_C, venn_D, ncol = 2)
-# grid::grid.legend(
-#   labels =  c("Long amplicon", "Short amplicon"),
-#   nrow = 1,
-#   pch = 21,
-#   gp = grid::gpar(fill = c("white", "slateblue1")),
-#   vp = grid::viewport(y = unit(0.2, "npc"))
-# )
-```
-
-(ref:readcap) Comparison between read numbers for different sequencing strategies, by ASV (top) and 97% OTU (bottom).
+ggpubr::ggarrange(venn_C, venn_D, ncol = 2)
+grid::grid.legend(
+  labels =  c("Long amplicon", "Short amplicon"),
+  nrow = 1,
+  pch = 21,
+  gp = grid::gpar(fill = c("white", "slateblue1")),
+  vp = grid::viewport(y = unit(0.15, "npc"))
+)
+```
+
+(ref:read-compare-s) Comparison between read numbers for different sequencing strategies
+
+(ref:read-compare) by ASV (top) and 97% OTU (bottom).
 ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
@@
 
-```{r pb-vs-it, fig.cap="(ref:readcap)", fig.height=8}
+```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8}
 ggpubr::ggarrange(
@@
 
-Figure\ \@ref(fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
+Figure\ \@ref(fig:read-compare) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
 In all cases, the read counts for shared ASVs/OTUs were correlated , with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")`.
@@
 
-Almost all of the short amplicons from all three technologies were between 240 and 375 bp long (Supplementary Figure\ \@ref(fig:full-length)a).
+Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Supplementary Figure\ \@ref(fig:full-length)a).
 Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
@@
 
-(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
+(ref:tax-chart-s) Summary of taxonomic assignments
+
+(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
@@
 
-```{r tax-chart, fig.cap = "(ref:tax-chart)"}
+```{r tax-chart, fig.cap = "(ref:tax-chart)", fig.scap = "(ref:tax-chart-s)"}
 rdd(tax_chart_plot)
 ```
-
-(ref:taxa) Heat tree summarizing taxonomic composition of soil community.
-Color and size of nodes represent the fractional read abundance of all ASVs belonging to that taxonomic group.
+(ref:taxa-s) Heat tree summarizing taxonomic composition of soil community
+
+(ref:taxa) Color and size of nodes represent the fractional read abundance of all ASVs belonging to that taxonomic group.
 Color and thickness of branches represent the fractional ASV richness of that taxonomic group.
-Both measures are a mean across sequencing runs, reference databases, and taxonomic assignment algorithms.
-Diverse but relatively rare groups are thus shown as small, light-colored nodes on relatively thick, darker-green-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as large nodes on relatively thin branches (e.g., Russulales).
+Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
 Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
 ASVs which could not be further assigned are labelled "?".
-
-```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
+ASV richness and read abundance are displayed as the mean across sequencing runs.
+Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
+The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
+
+```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8}
 fungi_read <- rdd(taxdata) %>%
@@
   extract2("tax_read") %>%
-  select(ends_with("PHYLOTAX")) %>%
   set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
@@
   extract2("tax_asv") %>%
-  select(ends_with("PHYLOTAX")) %>%
   set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
@@
 
-(ref:heattree-db-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
-Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
-Values are log_2_ of the ratio of median values across all datasets (sequencing technology, reference database, assignment algorithm) for each amplicon.
-
-```{r heattree-db-compare, fig.cap = "(ref:heattree-db-compare)", include = FALSE}
-taxdata2 <- rdd(taxdata)
-# taxdata2$data$asv_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = asv_cols)
-# taxdata2$data$read_table <- metacoder::calc_obs_props(taxdata2, "tax_data", cols = read_cols)
-# taxdata2$data$tax_asv <- metacoder::calc_taxon_abund(taxdata2, "asv_table", cols = asv_cols)
-# taxdata2$data$tax_read <- metacoder::calc_taxon_abund(taxdata2, "read_table", cols = read_cols)
-taxdata2$data$diff_read <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_read",
-  cols = names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_PacBio_Long_(Unite|RDP)"),
-  groups = names(taxdata2$data$tax_read) %>%
-    keep(str_detect, "reads_PacBio_Long_(Unite|RDP)") %>%
-    str_replace("reads_.+_.+_(.+)_.+", "\\1"),
-  func = function(abund1, abund2) {
-    list(read_ratio = log10(mean(abund1) / mean(abund2)))
-  }
-  )
-taxdata2$data$diff_asv <- metacoder::compare_groups(
-  taxdata2,
-  data = "tax_asv",
-  cols = names(taxdata2$data$tax_asv) %>% keep(startsWith, "ASVs_"),
-  groups = names(taxdata2$data$tax_asv) %>%
-    keep(startsWith, "ASVs_") %>%
-    str_replace("ASVs_.+_(.+)_.+_.+", "\\1"),
-  func = function(abund1, abund2) {
-    list(asv_ratio = log10(mean(abund1) / mean(abund2)))
-  }
-)
-
-taxdata2 <- taxdata2$filter_taxa(!is.nan(asv_ratio), !is.nan(read_ratio),
-                                 reassign_obs = FALSE)
-
-set.seed(2)
-theme <- theme_update(panel.border = element_blank())
-taxdata2 %>%
-  metacoder::heat_tree(
-    layout = "davidson-harel",
-    initial_layout = "reingold-tilford",
-
-    node_size = rowMeans(.$data$tax_read[,-1]),
-    node_size_range = c(0.005, .035),
-    node_size_axis_label = "Read abundance",
-
-    node_color = read_ratio,
-    node_color_range = metacoder::diverging_palette(),
-    node_color_trans = "linear",
-    node_color_axis_label = "Read abundance ratio",
-    node_color_interval = c(-4.5, 4.5),
-
-    edge_size = rowMeans(.$data$tax_asv[,-1]),
-    edge_size_range = c(0.0025, 0.03),
-    edge_size_trans = "linear",
-    edge_size_axis_label = "ASV count",
-
-    edge_color = asv_ratio,
-    edge_color_range = metacoder::diverging_palette(),
-    edge_color_trans = "linear",
-    edge_color_axis_label = "ASV count ratio",
-    edge_color_interval = c(-3, 3),
-
-    node_label = ifelse(
-      pmax(abs(.$data$diff_read$read_ratio),
-           abs(.$data$diff_asv$asv_ratio)) > 0.5,
-      taxon_names,
-      ""
-    ),
-    node_label_size_range = c(0.010, 0.03),
-    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
-    aspect_ratio = 3/2
-  )
-theme_set(theme)
-```
-
 ### Positive control
@@
 bisporus* ASV was the most abundant ASV in the long-amplicon PacBio dataset and the Ion Torrent dataset, and the fifth most abundant ASV in the short-amplicon PacBio dataset, and the third most abundant ASV in the Illumina dataset.
-This ASV was also found in `r rdd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, and represented `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
+This ASV was also found in `r rdd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, representing `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
 
@@
 
-(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).
+(ref:correlog-cap-s) Mantel correlograms for community dissimilarities and spatio-temporal distance
+
+(ref:correlog-cap) Community dissimilarities are Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).
 Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.
 
-```{r correlog, fig.cap="(ref:correlog-cap)"}
+```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)"}
 light_plan %>%
@@
 Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure\ \@ref(fig:correlog).
-Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 3 m for the total fungal community, and 1–2 m for the ECM community.
+Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 3\ m for the total fungal community, and 1–2\ m for the ECM community.
 In both cases, the greatest correlation magnitudes and distances are associated with Illumina sequencing of the short amplicon library, followed by PacBio sequencing of the long amplicon library, with PacBio sequencing of the short amplicon library detecting the least spatial structure.
 
-In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1 m 
+In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1\ m 
 distance class showing a significant correlation at $p < 0.05$.
 
-The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1 year and 0 m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
-For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0m and 1 year was not significant, but there was a negative correlation at time lag of 1 year and a distance of 1m for both sequencing technologies.
+The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1\ year and 0\ m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
+For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1m for both sequencing technologies.
 This negative correlation was significant in all short amplicon correlograms.
@@
 The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in Table\ \@ref(tab:variofit).
-Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0 m was included in the 95% confidence interval.
+Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m was included in the 95% confidence interval.
 Although the best fit spatial ranges vary between the different datasets by a factor of about 3, there is overlap of the 95% confidence intervals for all of the Bray-Curtis variograms across communities, amplicon libraries, and sequencing technologies.
 
-(ref:spatio) Relationship between spatiotemporal distance and community dissimilarities.
-Circles represent community data from short (top two rows) and long (bottom two rows) amplicon libraries, sequenced by Illumina MiSeq (top row) or PacBio RS II (bottom three rows).
+(ref:spatio-s) Distance-decay plot for community dissimilarities and spatio-temporal distance
+
+(ref:spatio) Circles represent community data from short (top two rows) and long (bottom two rows) amplicon libraries, sequenced by Illumina MiSeq (top row) or PacBio RS II (bottom three rows).
 Community dissimilarities are calculated using the Bray-Curtis dissimilarity for all datasets (top three rows) and using the weighted UNIFRAC dissimilarity for the long amplicon library, for which a phylogenetic tree could be constructed (bottom row).
 The left column represents the full fungal community, and the right column only sequences identified as ECM.
-The color of each circle represents the time lag between samples being compared (0 or 1 year), and the size represents the number of comparisons for that spatial distance and time lag.
+The color of each circle represents the time lag between samples being compared (0 or 1\ year), and the size represents the number of comparisons for that spatial distance and time lag.
 Lines are the best-fit lines for an exponential decay to max model.
-The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0 year timelag) or spatiotemporal (for the 1 year time lag) distance.
-
-```{r spatio, fig.cap = "(ref:spatio)"}
+The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0\ year timelag) or spatiotemporal (for the 1\ year time lag) distance.
+
+```{r spatio, fig.cap = "(ref:spatio-s). (ref:spatio)", fig.scap = "(ref:spatio-s)"}
 
@@
 
-DADA2 denoising to form ASVs may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:pb-vs-it)).
+DADA2 denoising to form ASVs may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
 The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
@@
 Within Fungi, the short amplicon datasets recovered more Sordariomycetes, possibly because this group was represented by diverse but rare taxa which were not detected in the PacBio datasets due to the lower sequencing depth.
-Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella*, *Tremellodendron*, and *Marcelleina* in the long amplicon dataset, and *Xylaria* and *Myerozyma* in the short amplicon datasets.
+Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella* in the long amplicon dataset (Supplementary Figure \@ref(fig:ecm-heattree)), and *Xylaria* and *Myerozyma* in the short amplicon datasets (Supplementary Figure\ \@ref(fig:heattree-length-compare)).
 These differences may be due to primer mismatches in these groups.
-It is of note that the three aforementioned groups which were more common in the long amplicon dataset are ECM.
 However, neither amplicon recovered any sequences identifiable to *Cantharellus*, an ECM genus which is known from the area from fruitbodies.
@@
 Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs, so this approach should increase ASV recovery from these platforms as well.
-It may also be adaptable to Oxford Nanopore sequencing, which has hitherto not been successfully applied to complex community metabarcoding [@loit2019].
+It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding [@loit2019].
 
@@
 Weighted UNIFRAC did not reliably detect spatial structure within this relatively homogeneous community.
-Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1 m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0-5700 m.
+Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0-5700\ m.
 This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure\ \@ref(fig:taxa)), which are both well represented in the majority of samples.
@@
 Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 17--19 m.
-The 95% confidence interval was substantially wider than this variation, generally covering a range of 9--35 m.
-All of these values are smaller than the 65 m reported by @Bahram2013, also from an ECM woodland habitat in Benin.
-The PacBio short amplicon dataset shows a longer spatial range, of 30 m for the total fungal community and 42--45 m for the ECM community, in both cases with wide confidence intervals spanning approximately 15--60 m.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 17--19\ m.
+The 95% confidence interval was substantially wider than this variation, generally covering a range of 9--35\ m.
+All of these values are smaller than the 65\ m reported by @Bahram2013, also from an ECM woodland habitat in Benin.
+The PacBio short amplicon dataset shows a longer spatial range, of 30\ m for the total fungal community and 42--45\ m for the ECM community, in both cases with wide confidence intervals spanning approximately 15--60\ m.
 It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
 
-The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0 m and 1 year.
-The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.1 years for the total fungal comunity and 5.5 years for the ECM community, but with overlapping confidence intervals.
-Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0 m and 1 year, but a weak negative correlation at 1 m and 1 year.
+The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0\ m and 1\ year.
+The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.1\ years for the total fungal comunity and 5.5\ years for the ECM community, but with overlapping confidence intervals.
+Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
 The general shape of the correlograms were similar for long and short amplicon datasets.
@@
 
-(ref:phylotax) Hypothetical example of phylogenetic refinement of taxonomic assignments.
-In the example, a clade includes  seven OTUs (A-F), which have been identified by two methods as belonging to taxa "Tax1" or "Tax2", or are unidentified ("unk"), as shown in the tip labels.
-No taxon is assigned at node 1, because one of the child branches (A) is completely unassigned.
-No taxon is assigned at node 2, because the assignments at C and F are inconsistent.
-Node 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
-Node 4 (and thus also node 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
-
-```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
+(ref:phylotax-s) Phylogenetic refinement of taxonomic assignments (PHYLOTAX)
+
+(ref:phylotax) In the example, a clade includes  seven OTUs (A-F), which have been identified by two methods as belonging to taxa "Tax1" or "Tax2", or are unidentified ("unk"), as shown in the tip labels.
+No taxon is assigned at node\ 1, because one of the child branches (A) is completely unassigned.
+No taxon is assigned at node\ 2, because the assignments at C and F are inconsistent.
+Node\ 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
+Node\ 4 (and thus also node\ 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
+
+```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)"}
 rdd(phylotax_fig)
@@
 
-(ref:full-length) Comparison of the length of denoised sequences from different length amplicons and sequencing technologies.
-Length distribution (*a*, *c*) and empirical cumulative distribution function (ECDF; *b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.
-
-```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
+(ref:full-length-s) Comparison of the length of denoised sequences from different length amplicons and sequencing technologies
+
+(ref:full-length) Length distribution (*a*, *c*) and empirical cumulative distribution function (ECDF; *b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.
+
+```{r full-length, fig.cap = "(ref:full-length-s). (ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6}
 ggpubr::ggarrange(
@@
 
-(ref:region-asv) Plot of total number of ASVs recovered vs.
-fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
-Some regions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU.
-For this figure, subregions were concatenated prior to generating ASVs with `DADA2`.
-
-```{r regions, fig.cap="(ref:region-asv)"}
+(ref:region-asv-s) Plot of total number of ASVs recovered vs. fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`
+
+(ref:region-asv) Some regions were formed by concatenation of extracted regions, as follows:
+LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
+ITS = ITS1--5.8S--ITS2;
+32S = 5.8S--ITS2--LSU;
+long = ITS--LSU.
+For this figure, subregions were concatenated prior to quality filtering and generating ASVs with `DADA2`.
+
+```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)"}
 rdd(region_table) %>%
@@
 
-(ref:region-lengths) Length distribution for each region extracted from long amplicons.  Filtering limits are shown as dashed vertical lines.
-
-```{r region-lengths, fig.cap = "(ref:region-lengths)", fig.height = 6}
+(ref:region-lengths-s) Length distribution for each region extracted from long amplicons
+
+(ref:region-lengths) Filtering limits are shown as dashed vertical lines.
+
+```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6}
 rdd(region_length_fig)
@@
 
-(ref:ITS2-length) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies.
-*a*) Length distribution.
-*b*) Empirical cumulative distribution function.
-
-```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
+(ref:ITS2-length-s) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies
+
+(ref:ITS2-length) *a*) Length distribution.  *b*) Empirical cumulative distribution function.
+
+```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6}
 ggpubr::ggarrange(
@@
 
-(ref:venn-table-asv) Correspondences between ASVs found by different sequencing strategies.
-Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+(ref:venn-table-asv-s) Correspondences between ASVs found by different sequencing strategies
+
+(ref:venn-table-asv) Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
 
@@
 rdd(venn_ASV) %>%
-venn_table(ASVs, "(ref:venn-table-asv)")
-```
-
-(ref:venn-table-otu) Correspondences between 97% OTUs found by different sequencing strategies.
-Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTU frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
+venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
+```
+
+(ref:venn-table-otu-s)  Correspondences between 97% OTUs found by different sequencing strategies
+
+(ref:venn-table-otu) Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTU frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
 
@@
 rdd(venn_OTU) %>%
-venn_table(OTUs, "(ref:venn-table-otu)")
-```
-
-(ref:short-its) Kingdom-level taxonomic composition of ITS2 ASVs <= 140 bp.
-Vertical axis shows fraction of total reads in each dataset.
-
-```{r short-its, fig.cap = "(ref:short-its)"}
+venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
+```
+
+(ref:short-its-s) Kingdom-level taxonomic composition of ITS2 ASVs $\le 140$ bp
+
+(ref:short-its) Vertical axis shows fraction of total reads in each dataset.
+
+```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)"}
 rdd(big_seq_table_ITS2) %>%
@@
 
-(ref:heattree-length-compare) Variation in read abundance and diversity for different amplicons across the taxonomic tree.
-Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
-Values are log~10~ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
+(ref:heattree-length-compare-s) Variation in read abundance and diversity for different amplicons across the taxonomic tree
+
+(ref:heattree-length-compare) Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
+Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
+Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
+ASVs which could not be further assigned are labelled "?".
 Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
@@
 
-```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare)"}
+```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s). (ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%"}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
-(ref:ecm) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments.
-Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
+(ref:ecm-s) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments
+
+(ref:ecm) Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
 "unidentified" denotes ASVs which were assigned to kingdom Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
@@
 
-```{r ecm, fig.cap = "(ref:ecm)"}
+```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)"}
   rdd(taxon_reads) %>%
+  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
   mutate_at("reference", replace_na, "All") %>%
@@
 
-(ref:ecm-heattree) Variation in read abundance and diversity for different amplicons, for ECM taxa only.
-Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
-Values are log~10~ of the ratio of mean values across all datasets (sequencing technology, assignment algorithm, reference database) for each amplicon.
-
-```{r ecm-heattree, fig.cap = "(ref:ecm-heattree)"}
+(ref:ecm-heattree-s) Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only
+
+(ref:ecm-heattree) Red edges/nodes represent increased read/ASV count for the short amplicon library, while teal edges/nodes represent increased read/ASV count for the long amplicon library.
+Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
+Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
+ASVs which could not be further assigned are labelled "?".
+Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
+The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
+
+```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%"}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
+(ref:sample-depth-s) DNA concentrations after extraction and PCR, and sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016
+
 (ref:sample-depth) DNA concentrations after extraction (leftmost column) and PCR (second and third columns), and sequencing depth (right three columns) along transects at the two sites Ang and Gan for the years 2015 and 2016.
 Sequencing reads which were filtered out during quality control are colored dark grey.
-Dotted horizontal line at 100 reads indicates cutoff for inclusion in community analysis.
+Dotted horizontal line at 100\ reads indicates cutoff for inclusion in community analysis.
 2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.
 
-```{r sample-depth, fig.cap = "(ref:sample-depth)"}
+```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)"}
 reads_plot <- rdd(qstats_n) %>%

### 2020-02-18 scripts/drake_plan_light.R

@@ -43,13 +43,13 @@ regions <- read_csv(config$regions, col_types = "cccciiiiic")
 #   group = c("fungi", "euk"),
 #   taxname = "short",
 #   reftree = c("fungi_tree_decipher_unconst_long", "tree_decipher_unconst_long"),
-#   conf_taxon = glue::glue("conf_taxon_short_{group}"),
+#   strict_taxon = glue::glue("strict_taxon_short_{group}"),
 #   phylotaxon = glue::glue("phylotaxon_unconst_long_{group}")
 # ) %>%
-#   mutate_at(c("reftree", "conf_taxon", "phylotaxon"), syms)
+#   mutate_at(c("reftree", "strict_taxon", "phylotaxon"), syms)
 
 guilds_meta <- tibble(
-  consensus_taxa = c("phylotaxon_decipher_unconst_long_fungi", "conf_taxon_short_fungi", "best_taxon_short_fungi"),
+  consensus_taxa = c("phylotaxon_decipher_unconst_long_fungi", "strict_taxon_short_fungi", "best_taxon_short_fungi"),
   amplicon = c("Long", "Short", "Short"),
   algorithm = c("PHYLOTAX", "Consensus", "PHYLOTAX+Cons")
 ) %>%
@@ -282,7 +282,7 @@ plan2 <- drake_plan(
     transform = map(phylotaxon, .tag_in = step, .id = c(treename, group))
   ),
   
-  conf_taxon_short_euk = target(
+  strict_taxon_short_euk = target(
     taxon_table %>%
       group_by(label, method, region, reference) %>%
       arrange(rank) %>%
@@ -302,25 +302,25 @@ plan2 <- drake_plan(
     transform = map(
       taxname = "short",
       group = "euk",
-      .tag_out = c(consensus_taxa, conf_taxon),
+      .tag_out = c(consensus_taxa, strict_taxon),
       .id = FALSE
     )
   ),
   
-  conf_taxon_short_fungi = target(
-    group_by(conf_taxon_short_euk$tip_taxa, label) %>%
+  strict_taxon_short_fungi = target(
+    group_by(strict_taxon_short_euk$tip_taxa, label) %>%
       filter("Fungi" %in% taxon | any(endsWith(taxon, "mycota"))) %>%
       list(tip_taxa = .),
     transform = map(
       taxname = "short",
       group = "fungi",
-      .tag_out = c(consensus_taxa, conf_taxon),
[…]

### 2020-02-18 scripts/output_functions.R

@@ -117,7 +117,7 @@ vennplot_data <- function(venndata, var) {
     )
 }
 
-venn_table <- function(venndata, var, caption) {
+venn_table <- function(venndata, var, caption, caption.short) {
   var <- enquo(var)
   
   linesep <- venndata %>%
@@ -155,6 +155,7 @@ venn_table <- function(venndata, var, caption) {
     kable(
       booktabs = TRUE,
       caption = caption,
+      caption.short = caption.short,
       linesep = linesep,
       align = ("rcrrcrrcrrcrr"),
       row.names = FALSE,

### 2020-02-27 writing/transect_paper.Rmd
Merge

@@
 The fungi provide nutrients and protection from pathogens in exchange for sugars from the plant [@Smith2010].
-Approximately 8% of described fungal species are believed to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
-Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
+Approximately 8% of described fungal species are thought to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
+Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
 Amherstieae [@brundrett2017a].
@@
 Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures [@horton2001].
-Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures.
-The advent of high-thoughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils [@Lindahl2013].
+Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures in the field.
+The advent of high-throughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils [@Lindahl2013].
 
 As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
-Traditionally, fungal metabarcoding studies have targetted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
+Traditionally, fungal metabarcoding studies have targeted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
@@
 Additionally the wide range of length variation of these regions may introduce bias in recovery of different taxa.
-Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].
+Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].
 The recent advent of long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) introduce the possibility of sequencing longer amplicons which include both the ITS regions and the flanking, more conserved SSU and/or LSU regions [@Tedersoo2018].
-This has the potential to improve taxonomic placement of sequences that lack close database matches, and to allow the alignment of metabarcoding reads.
-However, long-read technologies are currently more expensive per read, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
+This has the potential to improve taxonomic placement of sequences that lack close database matches, and to allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
+However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
 
@@
 The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
-Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are not as dataset specific [@callahan2017].
-
-Assignment of taxonomic identities using BLAST requires *a priori* choice of threshholds for different taxonomic ranks.
+Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are less dataset specific [@callahan2017].
+
+Assignment of taxonomic identities using BLAST requires *a priori* choice of thresholds for different taxonomic ranks.
 Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments [e.g.,@wang2007;@edgar2016a;@murali2018a].
@@
 
-We investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM Soudanian woodland in Benin.
+We investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM dominated Soudanian woodland in Benin.
 
@@
 It varies between different ecosystems and taxonomic groups.
-Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve community dissimilarity vs.
+Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
 distance [@legendre2012].
@@
 Several studies have investigated the turnover scale in ECM forests, but so far they have been based on analysis of ECM root tips, rather than bulk soil.
-@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites in temperate forests, based on Sanger sequencing.
-Similarly, @pickles2012a found autocorrelation at distances <3.4\ m, also in temperate forests, based on T-RFLP.
+@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites in temperate forests, using Sanger sequencing.
+Similarly, @pickles2012a found autocorrelation at distances <3.4\ m, also in temperate forests, based on T-RFLP analysis.
 Two studies have included data from tropical African woodlands and wooded savannahs.
@@
 @Bahram2013 performed a meta-analysis of turnover scale in ECM communities, including data from Benin, where they found turnover range of 65\ m based on distance-decay curves.
-They concluded that the range of distance decay is greater in tropical forests than temperate forests.
+Based on these results they concluded that the range of distance decay is greater in tropical forests than temperate forests.
 
@@
 Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-Both locations were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+Both sites were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
 At each site, 25\ soil samples were collected along a linear transect at intervals of 1\ m in May 2015.
@@
 Wider than 5 mm?
-MR--> was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a <!-- sterilized?
--MR--> knife blade.
+MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
@@
 The long amplicon (approximately 1500\ bp) targetted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
 
@@
 Each reaction was performed in three technical replicates as for short amplicons.
-
-Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
 
@@
 of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden).
-The Illumina library was pooled with samples for another project, with one half of the reads from one lane devoted to this project.
+The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.
 
@@
 The resulting sequences, as well as the paired-end Illumina sequences, were demultiplexed and sequencing primers were removed using `cutadapt` version 2.8 [@martin2011].
-Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible.
+Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible, and these reads were thus only analyzed as a pool.
 For Ion Torrent and PacBio, reads were discarded if they did not have the appropriate primers on both ends.
@@
 Read pairs were retained when PCR primers were detected at the 5' ends of both the forward and reverse read.
-Read pairs where the primers were found in reverse orientation were kept in seperate files, but were retained in their original orientation until after denoising.
+Read pairs where the primers were found in reverse orientation were kept in separate files, but were retained in their original orientation until after denoising.
 
@@
 However, only `r rdd(region_table)$ASVs[rdd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
-We believe that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
+We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
 The `DADA2` algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
@@
 Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median $L=`r rdd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(rdd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - rdd(demuxqual)['pb_500'])^rdd(demuxlength)['pb_500'])), big.mark = " ")`.
+For the combination of long reads (median $L=`r rdd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(rdd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - rdd(demuxqual)['pb_500'])^rdd(demuxlength)['pb_500'])), big.mark = ",")`.
 Given the high diversity relative to sequencing depth in this study (`r rdd(region_table)$ASVs[rdd(region_table)$seq_run == "pb_483" & rdd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r rdd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
@@
 For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
-Because ITS2 is the most variable of the amplified regions (Supplementary Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected have highly similar sequences in the other regions, unless the amplicon was chimeric.
+Because ITS2 is the most variable of the amplified regions (Supplementary Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
 The concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015].
@@
 Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
-The count and sample distribution of reads assigned to each ASV were calculated in order to form a sample × ASV community matrix.
+The count and sample distribution of reads assigned to each full-length ASV were calculated in order to form a sample × ASV community matrix.
 A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
@@
 Demultiplexed reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
-Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using seperate error models and pseudo-pooling, and then forward and reverse reads were merged.
+Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
 The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the other technologies.
@@
 
-The tree was rooted outside the Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
+The tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
 The kingdom Fungi was identified as the minimal clade containing all ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
-ASVs falling outside this clade were not included in the fungal community analysis.
+ASVs falling outside this clade were not included in downstream fungal community analysis.
 
 Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure\ \@ref(fig:phylotax)).
-A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
+A taxon at a particular rank was assigned to a node and all its descendants if that taxon was consistent with the reference-based taxonomic assignments for each of the descendants.
 A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
-The result of this process was to give a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade where which was consistently given an assignment, as well as clarify the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.
+The result of this process was twofold.
+First, it gives a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade which was consistently given an assignment.
+Second, it clarified the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.
+<!--I think this sentence is too long and it would be easier to read if it was like three sentences.-->
 This refinement algorithm is referred to as "PHYLOTAX".
@@
 Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
-For datasets where the Mantel corellogram indicated spatial correlation between samples taken in seperate years, the model was re-fit with an additional term to represent temporal correlation:
+For datasets where the Mantel correlogram indicated spatial correlation between samples taken in separate years, the model was re-fit with an additional term to represent temporal correlation:
 $$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
 where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity from replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill", and $a_t$ the equivalent temporal range.
-95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package in R.
+The 95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package in R.
 
@@
     align = "rrrrrr",
-    format.args = list(big.mark = " "),
+    format.args = list(big.mark = ","),
     linesep = c(rep("", 3), rep("\\addlinespace", 2), rep("", 4)),
@@
 
-## Region extraction
-
 The 5.8S rDNA was located in `r rdd(regioncounts)['pb_500']` long amplicon PacBio reads (`r percent(rdd(regioncounts)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total), `r rdd(regioncounts)['pb_483']` short amplicon PacBio reads (`r percent(rdd(regioncounts)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`), and `r rdd(regioncounts)['is_057']` Ion Torrent reads (`r percent(rdd(regioncounts)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`).
+<!--Not sure about , as thousand delimitator but not having delimitators makes it hard to read.
+You could also skip having the numbers in the text and just let them be in the table.
+Instead you could write that long read sequencing yield more than twice as many reads as short read etc… I think that would make the section more attractive to read.
+-AR-->
 Length and quality filtering resulted in `r rdd(filtercounts_ITS2)['pb_500']` (`r percent(rdd(filtercounts_ITS2)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total) ITS2 sequences from the long amplicon PacBio reads, `r rdd(filtercounts_ITS2)['pb_483']` (`r percent(rdd(filtercounts_ITS2)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`) ITS2 sequences from the short amplicon PacBio reads, and `r rdd(filtercounts_ITS2)['is_057']` (`r percent(rdd(filtercounts_ITS2)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`) ITS2 sequences from the Ion torrent reads.

### 2020-02-27 scripts/drake_plan_light.R

@@ -683,7 +683,7 @@ plan2 <- drake_plan(
     filter(is.na(step) | step == "raw", is.na(well)) %>%
     group_by(seq_run) %>%
     summarize(
-      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      # nreads =  prettyNum(sum(nreads), big.mark = ",")
       nreads = sum(nreads)
     ) %>%
     deframe(),
@@ -692,7 +692,7 @@ plan2 <- drake_plan(
     filter(is.na(step) | step == "demux", !is.na(well)) %>%
     group_by(seq_run) %>%
     summarize(
-      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      # nreads =  prettyNum(sum(nreads), big.mark = ",")
       nreads = sum(nreads)
     ) %>%
     deframe(),
@@ -701,7 +701,7 @@ plan2 <- drake_plan(
     filter(step == "filter", is.na(region) | region %in% c("long", "short")) %>%
     group_by(seq_run, region) %>%
     summarize(
-      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      # nreads =  prettyNum(sum(nreads), big.mark = ",")
       nreads = sum(nreads)
     ) %>%
     select(seq_run, nreads) %>%
@@ -711,7 +711,7 @@ plan2 <- drake_plan(
     filter(step == "filter", region == "ITS2") %>%
     group_by(seq_run, region) %>%
     summarize(
-      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      # nreads =  prettyNum(sum(nreads), big.mark = ",")
       nreads = sum(nreads)
     ) %>%
     select(seq_run, nreads) %>%
@@ -721,7 +721,7 @@ plan2 <- drake_plan(
     filter(step == "lsux", region == "5_8S") %>%
     group_by(seq_run) %>%
     summarize(
-      # nreads =  prettyNum(sum(nreads), big.mark = " ")
+      # nreads =  prettyNum(sum(nreads), big.mark = ",")
       nreads = sum(nreads)
     ) %>%
     deframe(),
@@ -935,7 +935,14 @@ plan2 <- drake_plan(
           pivot_wider(
             names_from = "rank",
             values_from = "taxon"
-          ),
+          ) %>%
+          mutate(
+            family = ifelse(is.na(genus), family, coalesce(family, "incertae_sedis")),
+            order = ifelse(is.na(family), order, coalesce(order, "incertae_sedis")),
[…]

### 2020-02-27 writing/transect_paper.Rmd
incorporate some coauthor changes

@@
 ---
-title: "Long- and Short-Read Metabarcoding Give Similar Results for Spatial Turnover Range in an Ectomycorrhizal Savanna Woodland"
+title: "Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities"
 author: 
   - Brendan Furneaux^1^,
-    Mohammad Bahram^1^,
-    Anna Rosling^3^,
-    Nourou S. Yorou^2^,
+    Mohammad Bahram^2,3^,
+    Anna Rosling^4^,
+    Nourou S. Yorou^5^,
     Martin Ryberg^1^
@@
     Uppsala, Sweden
-  - ^2^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
+  - ^2^Department of Ecology,
+    Swedish University of Agricultural Sciences,
+    Uppsala, Sweden
+  - ^3^Institute of Ecology and Earth Sciences,
+    University of Tartu,
+    Tartu, Estonia
+  - ^4^Program in Evolutionary Biology,
+    Department of Ecology and Genetics,
+    Uppsala University,
+    Uppsala, Sweden
+  - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
     LEB,
@@
     Parakou, Benin
-  - ^3^Program in Evolutionary Biology,
-    Department of Ecology and Genetics,
-    Uppsala University,
-    Uppsala, Sweden
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 
-Ectomycorrhizal (ECM) communities are characterized by mutualistic symbiosis between certain fungi and plant roots.
-Spatial turnover rate is one important characteristic of ecological communities, which is relevant for biodiversity assessment and sampling design.
-We investigated turnover rate in an ECM fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
-Resulting turnover rates depended more on statistical method than sequencing technology.
+Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in **xxx**.
+We investigated spatio-temporal structure of an ECM fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
+Resulting community structure and diversity depended more on statistical method than sequencing technology.
 We found significant autocorrelation in distance classes <3\ m using a Mantel correlogram, and a best-fit distance-decay curve range of 17--19\ m, both shorter than previously reported for similar habitats.
@@
 Ectomycorrhizas (ECM) are a symbiosis between fungi and plant root tips.
-The fungi provide nutrients and protection from pathogens in exchange for sugars from the plant [@Smith2010].
+The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant [@Smith2010].
 Approximately 8% of described fungal species are thought to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
@@
 
-Although many well-known mushrooms are produced by ECM fungi (e.g., *Amanita*, *Cantharellus*, *Boletus*), many ECM fungi instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
+Although many ECM fungi form many well-known mushrooms (e.g., *Amanita*, *Cantharellus*, *Boletus*), some instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
 Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures [@horton2001].
@@
 As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
-Traditionally, fungal metabarcoding studies have targeted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
+Fungal metabarcoding studies using short-read HTS technologies have targeted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
 If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013].
-However, this traditional approach comes with some potential weaknesses.
+However, this approach comes with some potential weaknesses.
 

### 2020-03-01 writing/transect_paper.Rmd
working on addressing comments

@@
   - "R.bib"
-csl: molecular-ecology.csl
+csl: new-phytologist.csl
 always_allow_html: true
@@
 
-Ectomycorrhizas (ECM) are a symbiosis between fungi and plant root tips.
+Ectomycorrhizas (ECM) form a symbiosis between fungi and plants.
 The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant [@Smith2010].
 Approximately 8% of described fungal species are thought to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
-Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr.
+Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, *Uapaca* (Phyllanthaceae) and Fabaceae tr.
 Amherstieae [@brundrett2017a].
@@
 
-While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the large taxonomic scales typically present in environmental samples [@Lindahl2013].
-Additionally the wide range of length variation of these regions may introduce bias in recovery of different taxa.
+While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
+Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa (**ref**).
 Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].
-The recent advent of long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) introduce the possibility of sequencing longer amplicons which include both the ITS regions and the flanking, more conserved SSU and/or LSU regions [@Tedersoo2018].
-This has the potential to improve taxonomic placement of sequences that lack close database matches, and to allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
+Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more conserved SSU and/or LSU regions [@Tedersoo2018].
+This can potentially improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
 However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
 
+In addition **XXX**.
 Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
@@
 
-Phylogenetic information has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
+Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
 Because OTU clustering may both "clump" different species into a single OTU, and "split" a single species into multiple OTUs [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
-In contrast, phylogenetic community distance measures like UNIFRAC [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
+In contrast, phylogenetic community distance measures [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
 Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
@@
 
-We investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM dominated Soudanian woodland in Benin.
+Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM-dominated Soudanian woodland in Benin.
 
@@
 Two studies have included data from tropical African woodlands and wooded savannahs.
-@tedersoo2011 found autocorrelation at ranges <10\ m in Miombo woodland based on Sanger sequencing.
+@tedersoo2011 found autocorrelation at ranges <10\ m in Miombo woodland using Sanger sequencing.
 However, this study was based on irregular sampling with a minimum spacing of 8\ m, so this as the smallest distance class.
@@
 
-Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Reservée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
 Both sites were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples were found not to be comparable with results preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-compare-long) and \@ref(fig:buffer-compare-short)), so these samples were not included in spatial analyses.
+Sequencing results for these samples differed significantly (p < 0.001 from permutational ANOVA) from samples preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-compare-long) and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
 However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
+
+```{r}
+buffer_asv_table <- rdd(proto_physeq) %>%
+  phyloseq::subset_samples(tech != "Ion Torrent") %>%
+  phyloseq::subset_samples(sample_type == "Sample") %>%
+  phyloseq::prune_taxa(taxa = rdd(fungi_PHYLOTAX.Cons)$label) %>%
+        phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 100)
+
+samples <- phyloseq::sample_data(buffer_asv_table) %>%
+  as_tibble(rownames = "sample") %>%
+  group_by(site, x, tech, amplicon) %>%
+  filter(n() >= 3) %>%
+  pull(sample) %>%
+  unique()
+
+buffer_asv_table <- phyloseq::prune_samples(buffer_asv_table, samples = samples)
+
+buffer_asv_bc_dist <- buffer_asv_table %>%
+  phyloseq::transform_sample_counts(function(x) x / sum(x)) %>%
+  phyloseq::distance("bray")
+vegan::adonis2(
+  buffer_asv_bc_dist ~ paste(site, x) + paste(tech, amplicon) + buffer + year,
+  data = as_tibble(phyloseq::sample_data(buffer_asv_table)),
+  by = "margin",
+  permutations = 9999
+)
+```
+
+```{r}
+buffer_asv_dbrda <- vegan::dbrda(
+  buffer_asv_bc_dist ~ 1,
+  data = as_tibble(phyloseq::sample_data(buffer_asv_table))
+)
+plot(buffer_asv_dbrda, type = "none")
+points(buffer_asv_dbrda, select = with(as_tibble(phyloseq::sample_data(buffer_asv_table)),
+                                     buffer == "Xpedition" & year == "2015" ),
+       col = "green", pch = 1)
+points(buffer_asv_dbrda, select = with(as_tibble(phyloseq::sample_data(buffer_asv_table)),
+                                     buffer == "Xpedition" & year == "2016"),
+       col = "blue", pch = 2)
+points(buffer_asv_dbrda, select = with(as_tibble(phyloseq::sample_data(buffer_asv_table)),
+                                     buffer == "LifeGuard" & year == "2016"),
+       col = "red", pch = 3)
+legend("bottomright", legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"), col = c("green", "blue", "red"), pch = 1:3)
+legend("topright", legend = with(as_tibble(phyloseq::sample_data(buffer_asv_table)), paste(tech, amplicon)) %>% unique() %>% sort(), lty = 1:3)
+vegan::ordiellipse(buffer_asv_dbrda, groups = with(as_tibble(phyloseq::sample_data(buffer_asv_table)), paste(tech, amplicon)),
+                   lty = 1:3)
+```
+
+```{r}
+tech_asv_table <- rdd(proto_physeq) %>%
+  phyloseq::subset_samples(tech != "Ion Torrent") %>%
+  phyloseq::subset_samples(buffer == "Xpedition") %>%
+  phyloseq::subset_samples(sample_type == "Sample") %>%
+  phyloseq::prune_taxa(taxa = rdd(fungi_PHYLOTAX.Cons)$label) %>%
+  phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 100)
+
+tech_class_table <- phyloseq::`tax_table<-`(
+  tech_asv_table,
+  phyloseq::tax_table(
+    rdd(fungi_PHYLOTAX.Cons) %>%
+      select(label, kingdom:genus) %>%
+      column_to_rownames("label") %>%
+      as.matrix()
+    )
+  ) %>%
+  phyloseq::tax_glom(taxrank = "class")
+
+phyloseq::taxa_names(tech_class_table) <- phyloseq::tax_table(tech_class_table)[,"class"]
+
+samples <- phyloseq::sample_data(tech_class_table) %>%
+  as_tibble(rownames = "sample") %>%
+  group_by(site, x, year) %>%
+  filter(n() >= 3) %>%
+  pull(sample) %>%
+  unique()
+
+tech_class_table <-
+  phyloseq::prune_samples(tech_class_table, samples = samples) %>%
+  phyloseq::transform_sample_counts(function(x) x / sum(x)) %>%
+  phyloseq::prune_taxa(taxa = (phyloseq::otu_table(tech_class_table) %>% t() %>% as.data.frame() %>% do.call(what = pmax)) > 0.01)
+
+tech_class_bc_dist <- tech_class_table %>%
+  phyloseq::distance("bray")
+vegan::adonis2(
+  tech_class_bc_dist ~ paste(site, x, year) + tech + amplicon,
+  data = as_tibble(phyloseq::sample_data(tech_class_table)),
+  by = "margin",
+  permutations = 9999
+)
+```
+
+```{r}
+tech_class_dbrda <- vegan::capscale(
+  phyloseq::otu_table(tech_class_table) ~ Condition(paste(site, x, year)),
+  data = as_tibble(phyloseq::sample_data(tech_class_table)),
+  distance = "bray"
+)
+tech_class_dbrda_scores <-
+  vegan::scores(tech_class_dbrda, display = "species") %>%
+  as_tibble(rownames = "class") %>%
+  mutate(L = sqrt(MDS1^2 + MDS2^2)) %>%
+  filter(L >= max(L) / 10) %>%
+  mutate(abbrev = substr(class, start = 1, stop = 2))
+```
+
+(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches.
+
+(ref:ppcoa) Only within-sample variation is displayed; variation between samples (site, position, year) have been partialed out.
+
+```{r ppcoa}
+plot(tech_class_dbrda, type = "none")
+text(
+  x = tech_class_dbrda_scores$MDS1,
+  y = tech_class_dbrda_scores$MDS2,
+  labels = tech_class_dbrda_scores$abbrev
+  )
+points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
+                                     tech == "PacBio" & amplicon == "Short" ),
+       col = "green", pch = 1)
+
+points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
+                                     tech == "PacBio" & amplicon == "Long"),
+       col = "blue", pch = 2)
+points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
+                                     tech == "Illumina" & amplicon == "Short"),
+       col = "red", pch = 3)
+legend("bottomleft", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
+```
+
+```{r}
+tech_class_nmds <- vegan::metaMDS(
+  phyloseq::otu_table(tech_class_table)
+)
+
+plot(tech_class_nmds, type = "n")
+vegan::ordispider(
+  tech_class_nmds,
+  groups = with(
+    as_tibble(phyloseq::sample_data(tech_class_table)),
+    paste(buffer, year, site, x)
+    ),
+  col = "gray80"
+  )
+text(
+  x = class_scores$MDS1,
+  y = class_scores$MDS2,
+  labels = class_scores$abbrev
+  )
+points(tech_class_nmds, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
+                                     tech == "PacBio" & amplicon == "Short" ),
+       col = "green", pch = 1)
+points(tech_class_nmds, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
+                                     tech == "PacBio" & amplicon == "Long"),
+       col = "blue", pch = 2)
+points(tech_class_nmds, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
+                                     tech == "Illumina" & amplicon == "Short"),
+       col = "red", pch = 3)
+legend("topright", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
+```
 
@@
 Two different fragments of the nuclear rDNA were amplified (Figure\ \@ref(fig:rDNA)).
-The short amplicon (approximately 300\ bp) targetted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
-The long amplicon (approximately 1500\ bp) targetted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
+The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
@@
 
-## Spatial analysis
+## Spatiotemporal analysis
 
@@
 Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure\ \@ref(fig:correlog).
-Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 3\ m for the total fungal community, and 1–2\ m for the ECM community.
+Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 2--3\ m for the total fungal community, and 1--2\ m for the ECM community.
 In both cases, the greatest correlation magnitudes and distances are associated with Illumina sequencing of the short amplicon library, followed by PacBio sequencing of the long amplicon library, with PacBio sequencing of the short amplicon library detecting the least spatial structure.
@@
 The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1\ year and 0\ m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
-For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1m for both sequencing technologies.
+For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
 This negative correlation was significant in all short amplicon correlograms.
@@
 Weighted UNIFRAC did not reliably detect spatial structure within this relatively homogeneous community.
-Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0-5700\ m.
+Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m.
 This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure\ \@ref(fig:taxa)), which are both well represented in the majority of samples.
@@
 Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 17--19\ m.
-The 95% confidence interval was substantially wider than this variation, generally covering a range of 9--35\ m.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 13--19\ m.
+The 95% confidence interval was substantially wider than this variation, generally covering a range of 6--35\ m.
 All of these values are smaller than the 65\ m reported by @Bahram2013, also from an ECM woodland habitat in Benin.
-The PacBio short amplicon dataset shows a longer spatial range, of 30\ m for the total fungal community and 42--45\ m for the ECM community, in both cases with wide confidence intervals spanning approximately 15--60\ m.
+The PacBio short amplicon dataset shows a longer spatial range, of 30\ m for the total fungal community and 42--43\ m for the ECM community, in both cases with wide confidence intervals spanning approximately 15--60\ m.
 It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
@@
 The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0\ m and 1\ year.
-The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.1\ years for the total fungal comunity and 5.5\ years for the ECM community, but with overlapping confidence intervals.
+The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.1\ years for the total fungal comunity and 4\ years for the ECM community, but with overlapping confidence intervals.
 Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.

### 2020-03-01 scripts/drake_plan_light.R

@@ -510,7 +510,7 @@ plan2 <- drake_plan(
         physeq <- phyloseq::prune_taxa(fungi$label, physeq) %>%
           phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 0)
       }
-      physeq
+      phyloseq::transform_sample_counts(physeq, function(x) x / sum(x))
     },
     transform = map(.data = !!physeq_meta, .tag_in = step, .id = c(guild, tech, amplicon, algorithm))
   ),

### 2020-03-02 writing/transect_paper.Rmd
PERMANOVA and PCoA

@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples differed significantly (p < 0.001 from permutational ANOVA) from samples preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-compare-long) and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
-However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
+Sequencing results for these samples differed significantly (p < `r rdd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")` from permutational MANOVA, $R^2 = `r rdd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$ ) from samples preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
+However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
 
 ```{r}
-buffer_asv_table <- rdd(proto_physeq) %>%
-  phyloseq::subset_samples(tech != "Ion Torrent") %>%
-  phyloseq::subset_samples(sample_type == "Sample") %>%
-  phyloseq::prune_taxa(taxa = rdd(fungi_PHYLOTAX.Cons)$label) %>%
-        phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 100)
-
-samples <- phyloseq::sample_data(buffer_asv_table) %>%
-  as_tibble(rownames = "sample") %>%
-  group_by(site, x, tech, amplicon) %>%
-  filter(n() >= 3) %>%
-  pull(sample) %>%
-  unique()
-
-buffer_asv_table <- phyloseq::prune_samples(buffer_asv_table, samples = samples)
-
-buffer_asv_bc_dist <- buffer_asv_table %>%
-  phyloseq::transform_sample_counts(function(x) x / sum(x)) %>%
-  phyloseq::distance("bray")
-vegan::adonis2(
-  buffer_asv_bc_dist ~ paste(site, x) + paste(tech, amplicon) + buffer + year,
-  data = as_tibble(phyloseq::sample_data(buffer_asv_table)),
-  by = "margin",
-  permutations = 9999
-)
-```
-
-```{r}
-buffer_asv_dbrda <- vegan::dbrda(
-  buffer_asv_bc_dist ~ 1,
-  data = as_tibble(phyloseq::sample_data(buffer_asv_table))
-)
-plot(buffer_asv_dbrda, type = "none")
-points(buffer_asv_dbrda, select = with(as_tibble(phyloseq::sample_data(buffer_asv_table)),
-                                     buffer == "Xpedition" & year == "2015" ),
-       col = "green", pch = 1)
-points(buffer_asv_dbrda, select = with(as_tibble(phyloseq::sample_data(buffer_asv_table)),
-                                     buffer == "Xpedition" & year == "2016"),
-       col = "blue", pch = 2)
-points(buffer_asv_dbrda, select = with(as_tibble(phyloseq::sample_data(buffer_asv_table)),
-                                     buffer == "LifeGuard" & year == "2016"),
-       col = "red", pch = 3)
-legend("bottomright", legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"), col = c("green", "blue", "red"), pch = 1:3)
-legend("topright", legend = with(as_tibble(phyloseq::sample_data(buffer_asv_table)), paste(tech, amplicon)) %>% unique() %>% sort(), lty = 1:3)
-vegan::ordiellipse(buffer_asv_dbrda, groups = with(as_tibble(phyloseq::sample_data(buffer_asv_table)), paste(tech, amplicon)),
-                   lty = 1:3)
-```
-
-```{r}
-tech_asv_table <- rdd(proto_physeq) %>%
-  phyloseq::subset_samples(tech != "Ion Torrent") %>%
-  phyloseq::subset_samples(buffer == "Xpedition") %>%
-  phyloseq::subset_samples(sample_type == "Sample") %>%
-  phyloseq::prune_taxa(taxa = rdd(fungi_PHYLOTAX.Cons)$label) %>%
-  phyloseq::prune_samples(samples = rowSums(phyloseq::otu_table(.)) > 100)
-
-tech_class_table <- phyloseq::`tax_table<-`(
-  tech_asv_table,
-  phyloseq::tax_table(
-    rdd(fungi_PHYLOTAX.Cons) %>%
-      select(label, kingdom:genus) %>%
-      column_to_rownames("label") %>%
-      as.matrix()
-    )
-  ) %>%
-  phyloseq::tax_glom(taxrank = "class")
-
-phyloseq::taxa_names(tech_class_table) <- phyloseq::tax_table(tech_class_table)[,"class"]
-
-samples <- phyloseq::sample_data(tech_class_table) %>%
-  as_tibble(rownames = "sample") %>%
-  group_by(site, x, year) %>%
-  filter(n() >= 3) %>%
-  pull(sample) %>%
-  unique()
-
-tech_class_table <-
-  phyloseq::prune_samples(tech_class_table, samples = samples) %>%
-  phyloseq::transform_sample_counts(function(x) x / sum(x)) %>%
-  phyloseq::prune_taxa(taxa = (phyloseq::otu_table(tech_class_table) %>% t() %>% as.data.frame() %>% do.call(what = pmax)) > 0.01)
-
-tech_class_bc_dist <- tech_class_table %>%
-  phyloseq::distance("bray")
-vegan::adonis2(
-  tech_class_bc_dist ~ paste(site, x, year) + tech + amplicon,
-  data = as_tibble(phyloseq::sample_data(tech_class_table)),
-  by = "margin",
-  permutations = 9999
-)
+
 ```
@@
 
-(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches.
-
-(ref:ppcoa) Only within-sample variation is displayed; variation between samples (site, position, year) have been partialed out.
-
-```{r ppcoa}
+(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches
+
+(ref:ppcoa) Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
+Only samples with at least 100 reads for all three approaches are included.
+ASVs are clustered by taxonomic class.
+Abbreviations represent the approximate direction of increase for the classes with the highest scores: `r glue::glue_data(tech_class_dbrda_scores, "**{abbrev}**: {class}") %>% glue::glue_collapse(sep = "; ", last = "; and ")`.
+
+```{r ppcoa, fig.cap = "(ref:ppcoa-short). (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
 plot(tech_class_dbrda, type = "none")
-text(
-  x = tech_class_dbrda_scores$MDS1,
-  y = tech_class_dbrda_scores$MDS2,
-  labels = tech_class_dbrda_scores$abbrev
-  )
 points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
@@
        col = "red", pch = 3)
+text(
+  x = tech_class_dbrda_scores$MDS1,
+  y = tech_class_dbrda_scores$MDS2,
+  labels = tech_class_dbrda_scores$abbrev
+  )
 legend("bottomleft", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
@@
 )
+
+tech_class_nmds_scores <-
+  vegan::scores(tech_class_nmds, display = "species") %>%
+  as_tibble(rownames = "class") %>%
+  mutate(L = sqrt(NMDS1^2 + NMDS2^2)) %>%
+  filter(!is.nan(L)) %>%
+  filter(L >= max(L) / 2) %>%
+  mutate(abbrev = substr(class, start = 1, stop = 2))
 
@@
 text(
-  x = class_scores$MDS1,
-  y = class_scores$MDS2,
-  labels = class_scores$abbrev
+  x = tech_class_nmds_scores$NMDS1,
+  y = tech_class_nmds_scores$NMDS2,
+  labels = tech_class_nmds_scores$abbrev
   )
@@
 
+(ref:buffer-pcoa-short) Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers
+
+(ref:buffer-pcoa) For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.
+
+```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)"}
+plot(rdd(buffer_asv_pcoa), type = "none")
+points(
+  rdd(buffer_asv_pcoa),
+  select = with(
+    as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))),
+    buffer == "Xpedition" & year == "2015" 
+  ),
+  col = "green",
+  pch = 1
+)
+points(
+  rdd(buffer_asv_pcoa),
+  select = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))),
+                buffer == "Xpedition" & year == "2016"
+  ),
+  col = "blue",
+  pch = 2
+)
+points(
+  rdd(buffer_asv_pcoa),
+  select = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))),
+                buffer == "LifeGuard" & year == "2016"
+  ),
+  col = "red",
+  pch = 3
+)
+legend(
+  "bottomright",
+  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
+  col = c("green", "blue", "red"),
+  pch = 1:3
+)
+# legend(
+#   "topright",
+#   legend = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))), paste(tech, amplicon)) %>%
+#     unique() %>%
+#     sort(),
+#   lty = 1:3
+# )
+# vegan::ordiellipse(
+#   rdd(buffer_asv_pcoa),
+#   groups = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))), paste(tech, amplicon)),
+#   lty = 1:3
+# )
+```
+
 (ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.

### 2020-03-02 writing/transect_paper.Rmd
coauthor changes up through beginning of results, plus added PCoA

@@
 }
-write_bib(c("base", "phyloseq", "gstat"), file = "R.bib", tweak = TRUE)
+write_bib(c("base", "phyloseq", "vegan"), file = "R.bib", tweak = TRUE)
 
@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples differed significantly (p < `r rdd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")` from permutational MANOVA, $R^2 = `r rdd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$ ) from samples preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
+Sequencing results for these samples differed significantly (p < `r rdd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")` from permutational MANOVA using 9999 permutations, $R^2 = `r rdd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
 However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
-
-```{r}
-
-```
-
-```{r}
-tech_class_dbrda <- vegan::capscale(
-  phyloseq::otu_table(tech_class_table) ~ Condition(paste(site, x, year)),
-  data = as_tibble(phyloseq::sample_data(tech_class_table)),
-  distance = "bray"
-)
-tech_class_dbrda_scores <-
-  vegan::scores(tech_class_dbrda, display = "species") %>%
-  as_tibble(rownames = "class") %>%
-  mutate(L = sqrt(MDS1^2 + MDS2^2)) %>%
-  filter(L >= max(L) / 10) %>%
-  mutate(abbrev = substr(class, start = 1, stop = 2))
-```
-
-(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches
-
-(ref:ppcoa) Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
-Only samples with at least 100 reads for all three approaches are included.
-ASVs are clustered by taxonomic class.
-Abbreviations represent the approximate direction of increase for the classes with the highest scores: `r glue::glue_data(tech_class_dbrda_scores, "**{abbrev}**: {class}") %>% glue::glue_collapse(sep = "; ", last = "; and ")`.
-
-```{r ppcoa, fig.cap = "(ref:ppcoa-short). (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
-plot(tech_class_dbrda, type = "none")
-points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
-                                     tech == "PacBio" & amplicon == "Short" ),
-       col = "green", pch = 1)
-
-points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
-                                     tech == "PacBio" & amplicon == "Long"),
-       col = "blue", pch = 2)
-points(tech_class_dbrda, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
-                                     tech == "Illumina" & amplicon == "Short"),
-       col = "red", pch = 3)
-text(
-  x = tech_class_dbrda_scores$MDS1,
-  y = tech_class_dbrda_scores$MDS2,
-  labels = tech_class_dbrda_scores$abbrev
-  )
-legend("bottomleft", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
-```
-
-```{r}
-tech_class_nmds <- vegan::metaMDS(
-  phyloseq::otu_table(tech_class_table)
-)
-
-tech_class_nmds_scores <-
-  vegan::scores(tech_class_nmds, display = "species") %>%
-  as_tibble(rownames = "class") %>%
-  mutate(L = sqrt(NMDS1^2 + NMDS2^2)) %>%
-  filter(!is.nan(L)) %>%
-  filter(L >= max(L) / 2) %>%
-  mutate(abbrev = substr(class, start = 1, stop = 2))
-
-plot(tech_class_nmds, type = "n")
-vegan::ordispider(
-  tech_class_nmds,
-  groups = with(
-    as_tibble(phyloseq::sample_data(tech_class_table)),
-    paste(buffer, year, site, x)
-    ),
-  col = "gray80"
-  )
-text(
-  x = tech_class_nmds_scores$NMDS1,
-  y = tech_class_nmds_scores$NMDS2,
-  labels = tech_class_nmds_scores$abbrev
-  )
-points(tech_class_nmds, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
-                                     tech == "PacBio" & amplicon == "Short" ),
-       col = "green", pch = 1)
-points(tech_class_nmds, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
-                                     tech == "PacBio" & amplicon == "Long"),
-       col = "blue", pch = 2)
-points(tech_class_nmds, select = with(as_tibble(phyloseq::sample_data(tech_class_table)),
-                                     tech == "Illumina" & amplicon == "Short"),
-       col = "red", pch = 3)
-legend("topright", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
-```
 
@@
 
-The gITS7 primers for the short amplicon were indexed for multiplexing (**Supplementary File**).
+The gITS7 primers for the short amplicon were indexed for multiplexing (**Supplementary File 1**).
 Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
@@
 
-Both primers for the long amplicon were indexed for multiplexing (**Supplementary File**).
+Both primers for the long amplicon were indexed for multiplexing (**Supplementary File 2**).
 PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
@@
 
-Additionally the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
+Additionally, the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
 of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden).
@@
 Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
-Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [ @adl2019], which utilizes hierarchical nameless ranks.
+Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
 FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).
 
-Taxonomic assignment was performed to genus level separately on the ITS and LSU regions using Unite/Warcup and RDP, respectively, as taxonomic references.
+Taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
 For each region/reference combination, taxonomy was assigned using three algorithms:
@@
 and IDTAXA [@murali2018].
-Each ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
+Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
+ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.
 
@@
 Full length long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10\ iterations of progressive alignment and conserved RNA secondary structure calculation and 10\ refinement iterations.
-This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns
-<!-- Just curious, what are Type 1 introns -AR--> <!-- It has to do with the enzyme that cuts them out, I think, but I don't actually know my introns are Type 1, just that this is a location where Type 1 introns are known to occur.
--BF-->
-after this position, as also observed in several fungal species by @holst-jensen1999.
+This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by @holst-jensen1999.
 
@@
 A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
-The result of this process was twofold.
-First, it gives a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade which was consistently given an assignment.
-Second, it clarified the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.
-<!--I think this sentence is too long and it would be easier to read if it was like three sentences.-->
+The result of this process is twofold.
+First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade which is consistently given an assignment.
+Second, it clarifies the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these is consistent with the assignments of other ASVs in the same clade.
 This refinement algorithm is referred to as "PHYLOTAX".
@@
 Additionally, a hybrid method, was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
+
+## Community comparison
+
+The fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina, PacBio Short, PacBio Long) were compared by permutational MANOVA.
+In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
+Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(rdd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(rdd(tech_class_table))` classes), were included.
+Permutational MANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs.
+short); and sequencing technology (Illumina MiSeq vs.
+PacBio RS II).
+The marginal significance of each term for explaining variation in the Bray-Curtis community dissimilarity matrix was performed using the `adonis2` function in the R package `vegan` [@R-vegan], with 9999 permutations.
+
+Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
+Spatiotemporal effects were partialled out in order to visualize effects due to sequencing technology and amplicon length.
 
@@
 Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads.
-Consequently Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
+Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
 
 The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table\ \@ref(tab:bioinfo).
-Long (short) amplicon sequencing with PacBio yielded `r rdd(rawcounts)["pb_500"]` (`r rdd(rawcounts)["pb_483"]`) circular consensus reads, of which `r rdd(demuxcounts)["pb_500"]` (`r rdd(demuxcounts)["pb_483"]`) were successfully trimmed and demultiplexed.
-Illumina sequencing yielded `r rdd(rawcounts)["SH-2257"]` read pairs, of which `r rdd(demuxcounts)["SH-2257"]` were successfully trimmed and demultiplexed, and `r rdd(filtercounts_full)["SH-2257"]` passed quality filtering.
-Ion Torrent sequencing yielded `r rdd(demuxcounts)["is_057"]` demultiplexed reads.
-`r rdd(filtercounts_full)['pb_500']` (`r percent(rdd(filtercounts_full)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")`) of long amplicon PacBio reads passed length and quality filtering stages, compared with `r rdd(filtercounts_full)['pb_483']` (`r percent(rdd(filtercounts_full)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`) of short amplicon PacBio reads, `r rdd(filtercounts_full)['SH-2257']` (`r percent(rdd(filtercounts_full)['SH-2257']/rdd(demuxcounts)['SH-2257'], digits = 1, format = "f")`) of Illumina read pairs, and `r rdd(filtercounts_full)['is_057']` (`r percent(rdd(filtercounts_full)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`) of Ion Torrent reads.
+Sequencing with PacBio RS II yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
+Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
+Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-rdd(filtercounts_full)/rdd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long reads, but only by `r (1-rdd(filtercounts_full)/rdd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short reads, resuling in a similar numbe rof filtered reads for the two strategies.
+Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-rdd(filtercounts_full)/rdd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
+In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-rdd(filtercounts_ITS2)['pb_500']/rdd(rawcounts)['pb_500'], digits = 1, format = "f")` of trimmed long amplicon PacBio reads, `r percent(1-rdd(filtercounts_ITS2)['pb_483']/rdd(rawcounts)['pb_483'], digits = 1, format = "f")` of trimmed short amplicon PacBio reads, and `r percent(1-rdd(filtercounts_ITS2)['is_057']/rdd(rawcounts)['is_057'], digits = 1, format = "f")` of trimmed Ion Torrent reads.
+This represented greater loss of PacBio short reads, but less loss of  PacBio long reads and Ion Torrent reads.
+
+The length distribution of the different regions extracted from the long amplicon are shown in Supplementary Figure\ \@ref(fig:region-lengths).
 
@@
 
-(ref:readcounts) *Raw*: raw reads as delivered by sequencing center;
-*Trim*: reads with primers and demultiplexing barcodes removed;
-*LSUx*: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
-*ITS2*: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
-*short*, *ITS*, *LSU*, *long*: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.
-Short amplicons were mapped to full ITS, LSU, and long amplicon ASVs when they had a perfect match in the ITS2 region.
+(ref:readcounts) **Raw**: raw reads as delivered by sequencing center (CCS reads for PacBio);
+**Trim**: reads with primers and demultiplexing barcodes removed;
+**Filter (full)**: full length amplicons with a maximum of three expected errors;
+**LSUx**: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
+**Filter (ITS2)**: extracted ITS2 regions with a maximum of three expected errors;
+**ITS2**: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
+**short**, **ITS**, **LSU**, **long**: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.
 
 ```{r bioinfo}
-#TODO add filtered reads (currently inflated numbers for non-Illumina)
-#TODO asterisks for mapped ASVs
 options(knitr.kable.NA = '–')
@@
     format.args = list(big.mark = ","),
-    linesep = c(rep("", 3), rep("\\addlinespace", 2), rep("", 4)),
+    linesep = c(
+      rep("", 2),
+      "\\addlinespace",
+      rep("", 2),
+      "\\addlinespace",
+      rep("", 4)
+    ),
     escape = FALSE
@@
 ```
-
-The 5.8S rDNA was located in `r rdd(regioncounts)['pb_500']` long amplicon PacBio reads (`r percent(rdd(regioncounts)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total), `r rdd(regioncounts)['pb_483']` short amplicon PacBio reads (`r percent(rdd(regioncounts)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`), and `r rdd(regioncounts)['is_057']` Ion Torrent reads (`r percent(rdd(regioncounts)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`).
-<!--Not sure about , as thousand delimitator but not having delimitators makes it hard to read.
-You could also skip having the numbers in the text and just let them be in the table.
-Instead you could write that long read sequencing yield more than twice as many reads as short read etc… I think that would make the section more attractive to read.
--AR-->
-Length and quality filtering resulted in `r rdd(filtercounts_ITS2)['pb_500']` (`r percent(rdd(filtercounts_ITS2)['pb_500']/rdd(demuxcounts)['pb_500'], digits = 1, format = "f")` of demultiplexed total) ITS2 sequences from the long amplicon PacBio reads, `r rdd(filtercounts_ITS2)['pb_483']` (`r percent(rdd(filtercounts_ITS2)['pb_483']/rdd(demuxcounts)['pb_483'], digits = 1, format = "f")`) ITS2 sequences from the short amplicon PacBio reads, and `r rdd(filtercounts_ITS2)['is_057']` (`r percent(rdd(filtercounts_ITS2)['is_057']/rdd(demuxcounts)['is_057'], digits = 1, format = "f")`) ITS2 sequences from the Ion torrent reads.
-The length distribution of different regions extracted from the long amplicon are shown in Supplementary Figure\ \@ref(fig:region-lengths).
 
@@
 Assignment to kingdom Fungi and ECM status for different amplicons, sequencing technologies, reference databases, and algorithms are shown in Supplementary Figure\ \@ref(fig:ecm).
+
+Figure\ \@ref(fig:ppcoa)
+
+(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches
+
+(ref:ppcoa) for all fungi (**a**) and ECM fungi (**b**).
+Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
+Only samples with at least 100 reads for all three approaches are included.
+ASVs are clustered by taxonomic class for all fungi, and by family for ECM fungi.
+Abbreviations represent the approximate direction of increase for the classes and families with the highest scores: `r glue::glue_data(rdd(tech_class_pcoa_scores), "**{abbrev}**: {class}") %>% glue::glue_collapse(sep = ", ", last = ", and ")`; `r glue::glue_data(rdd(tech_ecm_fam_pcoa_scores), "**{abbrev}**: {family}") %>% glue::glue_collapse(sep = ", ", last = ", and ")`.
+
+```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
+ggpubr::ggarrange(
+  rdd(tech_class_pcoa_plot),
+  rdd(tech_ecm_fam_pcoa_plot),
+  common.legend = TRUE,
+  legend = "bottom",
+  labels = "auto"
+)
+
+# plot(rdd(tech_class_pcoa), type = "none")
+# points(
+#   rdd(tech_class_pcoa),
+#   select = with(
+#     rdd(tech_class_sample_data),
+#     tech == "PacBio" & amplicon == "Short"
+#   ),
+#   col = "green",
+#   pch = 1
+# )
+# points(
+#   rdd(tech_class_pcoa),
+#   select = with(
+#     rdd(tech_class_sample_data),
+#     tech == "PacBio" & amplicon == "Long"
+#   ),
+#   col = "blue",
+#   pch = 2
+# )
+# points(
+#   rdd(tech_class_pcoa),
+#   select = with(
+#     rdd(tech_class_sample_data),
+#     tech == "Illumina" & amplicon == "Short"
+#   ),
+#   col = "red",
+#   pch = 3
+# )
+# with(
+#   rdd(tech_class_pcoa_scores),
+#   text(x = MDS1*2, y = MDS2*2, labels = abbrev)
+# )
+# legend(
+#   "bottomleft",
+#   legend = c("Illumina Short", "PacBio Short", "PacBio Long"),
+#   col = c("red", "green", "blue"),
+#   pch = c(3, 1, 2)
+# ) 
+# 
+# plot(rdd(tech_ecm_fam_pcoa), type = "none")
+# points(
+#   rdd(tech_ecm_fam_pcoa),
+#   select = with(
+#     rdd(tech_ecm_fam_sample_data),
+#     tech == "PacBio" & amplicon == "Short"
+#   ),
+#   col = "green",
+#   pch = 1
+# )
+# points(
+#   rdd(tech_ecm_fam_pcoa),
+#   select = with(
+#     rdd(tech_ecm_fam_sample_data),
+#     tech == "PacBio" & amplicon == "Long"
+#   ),
+#   col = "blue",
+#   pch = 2
+# )
+# points(
+#   rdd(tech_ecm_fam_pcoa),
+#   select = with(
+#     rdd(tech_ecm_fam_sample_data),
+#     tech == "Illumina" & amplicon == "Short"
+#   ),
+#   col = "red",
+#   pch = 3
+# )
+# with(
+#   rdd(tech_ecm_fam_pcoa_scores),
+#   text(x = MDS1*3, y = MDS2*3, labels = abbrev)
+# )
+# legend(
+#   "bottomleft",
+#   legend = c("Illumina Short", "PacBio Short", "PacBio Long"),
+#   col = c("red", "green", "blue"),
+#   pch = c(3, 1, 2)
+# )
+```
+
+```{r}
+tech_class_nmds <- vegan::metaMDS(
+  phyloseq::otu_table(rdd(tech_class_table))
+)
+
+tech_class_nmds_scores <-
+  vegan::scores(tech_class_nmds, display = "species") %>%
+  as_tibble(rownames = "class") %>%
+  mutate(L = sqrt(NMDS1^2 + NMDS2^2)) %>%
+  filter(!is.nan(L)) %>%
+  filter(L >= max(L) / 2) %>%
+  mutate(abbrev = substr(class, start = 1, stop = 2))
+
+plot(tech_class_nmds, type = "n")
+vegan::ordispider(
+  tech_class_nmds,
+  groups = with(
+    rdd(tech_class_sample_data),
+    paste(buffer, year, site, x)
+    ),
+  col = "gray80"
+  )
+text(
+  x = tech_class_nmds_scores$NMDS1,
+  y = tech_class_nmds_scores$NMDS2,
+  labels = tech_class_nmds_scores$abbrev
+  )
+points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
+                                     tech == "PacBio" & amplicon == "Short" ),
+       col = "green", pch = 1)
+points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
+                                     tech == "PacBio" & amplicon == "Long"),
+       col = "blue", pch = 2)
+points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
+                                     tech == "Illumina" & amplicon == "Short"),
+       col = "red", pch = 3)
+legend("topright", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
+```
 
@@
 
-(ref:buffer-pcoa-short) Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers
+(ref:buffer-pcoa-s) Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers
 

### 2020-03-02 writing/transect_paper.Rmd
Coauthor comments in tech comparison results

@@
 Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-rdd(filtercounts_full)/rdd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
-In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-rdd(filtercounts_ITS2)['pb_500']/rdd(rawcounts)['pb_500'], digits = 1, format = "f")` of trimmed long amplicon PacBio reads, `r percent(1-rdd(filtercounts_ITS2)['pb_483']/rdd(rawcounts)['pb_483'], digits = 1, format = "f")` of trimmed short amplicon PacBio reads, and `r percent(1-rdd(filtercounts_ITS2)['is_057']/rdd(rawcounts)['is_057'], digits = 1, format = "f")` of trimmed Ion Torrent reads.
+In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-rdd(filtercounts_ITS2)['pb_500']/rdd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-rdd(filtercounts_ITS2)['pb_483']/rdd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-rdd(filtercounts_ITS2)['is_057']/rdd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
 This represented greater loss of PacBio short reads, but less loss of PacBio long reads and Ion Torrent reads.
@@
 The length distribution of the different regions extracted from the long amplicon are shown in Supplementary Figure\ \@ref(fig:region-lengths).
+ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r rdd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r rdd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r rdd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r rdd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r rdd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
+Approximately `(rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Supplementary Figure\ \@ref(fig:region-lengths) due to rarity.
+Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.
 
@@
 
-## Reproducibility of read counts from different technologies
+## Reproducibility of sequence capture using different technologies
 
@@
 
-(ref:read-compare) by ASV (top) and 97% OTU (bottom).
+(ref:read-compare) by ASV (**a**) and 97% OTU (**b**).
 ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
@@
 
-The number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively, and graphically in Figure\ \@ref(fig:venn).
-ASVs shared between all datasets represented `r text_list(percent(rdd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
-These fractions increased to `r text_list(percent(rdd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))` when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Figure\ \@ref(fig:venn)).
+ASVs shared between all datasets represented `r paste(range(as.numeric(unlist(rdd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
+These fractions increased to `r paste(range(as.numeric(unlist(rdd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
 In particular, `r text_list(percent(colSums(mutate_all(rdd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
 In contrast, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
-
-Figure\ \@ref(fig:read-compare) shows the correspondence between the read count for different ASVs/OTUs in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted on the margins.
-In all cases, the read counts for shared ASVs/OTUs were correlated , with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")`.
+Complete tabulations of the number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively.
+
+Figure\ \@ref(fig:read-compare) shows the correspondence between the read count for different ASVs (\@ref(fig:read-compare)a) and OTUs(\@ref(fig:read-compare)b) in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted as lines along the axes.
+In all cases, the read counts for shared ASVs and OTUs were correlated, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")`.
 Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
@@
 `r compR2("SH-2257", "is_057", "ASV")` to `r compR2("SH-2257", "is_057", "OTU")`,
-for PacBio vs. Illumina, PacBio vs. Ion Torrent, and Illumina vs. Ion Torrent, respectively),
-but this was not, or only weakly, the case for correlations between the long amplicon library and short amplicon library
+for PacBio vs. Illumina, PacBio vs. Ion Torrent, and Illumina vs. Ion Torrent, respectively), but not between the long amplicon library and short amplicon library
 (`r compR2("pb_500", "pb_483", "ASV")` to `r compR2("pb_500", "pb_483", "OTU")`,
 `r compR2("pb_500", "SH-2257", "ASV")` to `r compR2("pb_500", "SH-2257", "OTU")`, and
-`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs PacBio, Illumina, and Ion Torrent short reads, respectively).
+`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short reads, respectively).
 
 Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Supplementary Figure\ \@ref(fig:full-length)a).
-Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
+Although the length profile of the three sequencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
+Although the difference in length distributions was statistically significant (Kruskal-Wallis H = )
 In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of `r filter(rdd(reads_table), region == "long") %$% reldist::wtd.quantile(length, weight = reads)` bp.

### 2020-03-02 writing/transect_paper.Rmd
analysis of length distributions (too much analysis!)

@@
 Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Supplementary Figure\ \@ref(fig:full-length)a).
-Although the length profile of the three sequencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
-Although the difference in length distributions was statistically significant (Kruskal-Wallis H = )
-In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of `r filter(rdd(reads_table), region == "long") %$% reldist::wtd.quantile(length, weight = reads)` bp.
+Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
+The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r rdd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(rdd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
+In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(rdd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
 

### 2020-03-03 writing/transect_paper.Rmd
a few more results for controls

@@
 We investigated turnover range in both the total fungal community and ECM community by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
-We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RPP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs.
+We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs.
 weighted UNIFRAC).
@@
 This represented greater loss of PacBio short reads, but less loss of PacBio long reads and Ion Torrent reads.
+
+Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Supplementary Figure\ \@ref(fig:full-length)a).
+Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
+The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r rdd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(rdd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
+In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(rdd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
 
@@
 
-Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Supplementary Figure\ \@ref(fig:full-length)a).
-Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
-The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r rdd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(rdd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
-In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(rdd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
-
 ## Taxonomic assignment
 
-The fraction of reads and ASVs assigned to each taxonomic rank is shown in Figure\ \@ref(fig:tax-chart).
-For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs.
+For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs (Figure\ \@ref(fig:tax-chart)).
 A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
-Across sequencing technologies, amplicons, and reference databases, the RDPC algorithm assigned the greatest fraction of reads and ASVs, followed by SINTAX, and then IDTAXA, with the exception of lower taxonomic ranks (order for the Long amplicon, family and genus for both amplicons) using the warcup reference database, where SINTAX assigned the greatest fraction of reads, but not ASVs.
+At most taxonomic ranks, the RDPC algorithm assigned the greatest fraction of reads and ASVs, followed by SINTAX, and then IDTAXA.
 
@@
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
-They are plotted in each column to compare with results for the individual databases.
+These two methods are plotted in each column to compare with results for the individual databases.
 
@@
 Taxonomic composition of the sequenced soil community is summarized in Figure\ \@ref(fig:taxa).
-According to the PHYLOTAX assignments, Fungi represented `r percent(fungi_read["PacBio_Long"], format = "f", digits = 1)` of the reads and `r percent(fungi_asv["PacBio_Long"], format = "f", digits = 1)` of the ASVs in the long amplicon library, compared to `r paste(percent(range(fungi_read[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads and `r paste(percent(range(fungi_asv[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs in the short amplicon library.
-Fungi categorized as ECM made up `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads and `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs in the long amplicon library, and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs in the short amplicon library.
+According to the PHYLOTAX assignments, Fungi represented `r percent(fungi_asv["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(fungi_read["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(fungi_asv[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(fungi_read[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
+Fungi categorized as ECM made up `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
 Assignment to kingdom Fungi and ECM status for different amplicons, sequencing technologies, reference databases, and algorithms are shown in Supplementary Figure\ \@ref(fig:ecm).
@@
 
-(ref:heattree-length-compare) Red nodes/edges represent increased read/ASV count for the short amplicon library, while teal nodes/edges represent increased read/ASV count for the long amplicon library.
+(ref:heattree-length-compare) Red nodes (edges) represent increased read (ASV) count for the short amplicon library, while blue-green nodes (edges) represent increased read (ASV) count for the long amplicon library.
 Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.

### 2020-03-03 scripts/drake_plan_light.R

@@ -483,7 +483,8 @@ plan2 <- drake_plan(
   
   physeq = target(
     {
-      physeq <- proto_physeq
+      physeq <- proto_physeq %>%
+        phyloseq::prune_taxa(taxa = !phyloseq::taxa_names(.) %in% pos_control)
       if (amplicon == "Long") {
         phyloseq::phy_tree(physeq) <- fungi_tree_decipher_unconst_long
       }
@@ -1151,6 +1152,11 @@ plan2 <- drake_plan(
         (seq_run == "is_057" & well == "E7")
     ),
   
+  neg_control_physeq = phyloseq::subset_samples(
+    proto_physeq,
+    sample_type == "Blank"
+  ),
+  
   nonpos_control_physeq = 
     phyloseq::subset_samples(
       proto_physeq,
@@ -1160,15 +1166,23 @@ plan2 <- drake_plan(
   
   pos_control =
     agaricus_reads %>%
-    as.matrix() %>%
-    colSums() %>%
-    which.max() %>%
-    names(),
+    as("matrix") %>%
+    as_tibble(rownames = "sample") %>%
+    pivot_longer(-1, names_to = "seq", values_to = "reads") %>%
+    filter(reads > 0) %>%
+    extract(
+      sample,
+      into = c("seq_run", "plate", "well"),
+      regex = "([a-zA-Z]{2}[-_]\\d{3,4})(00\\d)([A-H]1?\\d)"
+    ) %>%
+    filter(seq_run != "is_057") %$%
+    unique(seq),
   
   pos_control_reads =
     nonpos_control_physeq %>%
     phyloseq::prune_taxa(taxa = pos_control) %>%
     phyloseq::otu_table() %>%
+    as.matrix() %>%
     rowSums(),
   
   all_reads =
@@ -1180,13 +1194,44 @@ plan2 <- drake_plan(
   pos_control_data =
     nonpos_control_physeq %>%
     phyloseq::sample_data() %>%
-    as.data.frame() %>%
+    as("data.frame") %>%
     mutate(
       pc_reads = pos_control_reads,
       all_reads = all_reads,
       pc_frac = pc_reads/all_reads
     ) %>%
     select(seq_run, well, plate, pc_frac, pc_reads, all_reads),
+  
+  pc_pc_reads = 
+    pos_control_physeq %>%
+    phyloseq::prune_taxa(taxa = pos_control) %>%
[…]

### 2020-03-03 writing/transect_paper.Rmd
finished first pass through coauthor comments

@@
 
-We attempted to denoise both long and short PacBio amplicons using `DADA2` according to the steps outlined in the supplementary information in @callahan2019.
+We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in @callahan2019.
 However, only `r rdd(region_table)$ASVs[rdd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
 We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
-The `DADA2` algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
+The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
 If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
@@
 Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors.
-Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
-The error model for `DADA2` denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
+Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")` [R package `dada2`, @callahan2016; @callahan2019].
+The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
 Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
-For PacBio libraries, `DADA2` was run with complete pooling and a band size of 16.
-For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the `DADA2` FAQ (https://benjjneb.github.io/dada2/faq.html).
-Chimeras within each region were removed using `removeBimeraDenovoTable` from `DADA2`.
+For PacBio libraries, DADA2 was run with complete pooling and a band size of 16.
+For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the DADA2 FAQ (https://benjjneb.github.io/dada2/faq.html).
+Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
 
@@
 Because the Illumina dataset consisted of paired-end reads, regions were not extracted prior to denoising.
-ASVs were instead generated according to a standard workflow for `DADA2`.
+ASVs were instead generated according to a standard workflow for DADA2.
 Demultiplexed reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
-Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
+Forward and reverse reads were denoised using DADA2 version `r packageVersion("dada2")` [@callahan2016] using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
 The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the other technologies.
@@
 
-Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@wang2007] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
+Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
 In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
@@
 For each region/reference combination, taxonomy was assigned using three algorithms:
-the RDP Naïve Bayesian Classifier (RDPC) as implemented in DADA2;
+the RDP Naïve Bayesian Classifier [RDPC, @wang2007] as implemented in DADA2;
 SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
@@
 
+*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
+*A.
+bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair (`r Biostrings::readDNAStringSet(here::here("data/clusters/ITS2.fasta.gz")) %>% magrittr::extract(startsWith(names(.), "f2ab49e8")) %>% Biostrings::width() %>% unique() %>% divide_by(1, .) %>% subtract(1, .) %>% percent(digits = 1, format = "f")` similarity in ITS2).
+The abundance of the second ASV was `r rdd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
+> 0) %>% percent(format = "f", digits = 1) %>% paste(collapse = " and ")` that of the primary *A.
+bisporus* ASV in the two Illumina positive controls.
+The consistency of this ratio across replicate positive controls suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
+Despite higher total sequencing depth, this ASV was not identified from the Ion Torrent dataset.
+
+*A.
+bisporus* sequences represented `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively, giving similar estimates for the rate of tag-switching for all technologies.
+These reads were excluded from community analyses.
+
 ## Reproducibility of sequence capture using different technologies
@@
 )
-
-# plot(rdd(tech_class_pcoa), type = "none")
-# points(
-#   rdd(tech_class_pcoa),
-#   select = with(
-#     rdd(tech_class_sample_data),
-#     tech == "PacBio" & amplicon == "Short"
-#   ),
-#   col = "green",
-#   pch = 1
-# )
-# points(
-#   rdd(tech_class_pcoa),
-#   select = with(
-#     rdd(tech_class_sample_data),
-#     tech == "PacBio" & amplicon == "Long"
-#   ),
-#   col = "blue",
-#   pch = 2
-# )
-# points(
-#   rdd(tech_class_pcoa),
-#   select = with(
-#     rdd(tech_class_sample_data),
-#     tech == "Illumina" & amplicon == "Short"
-#   ),
-#   col = "red",
-#   pch = 3
-# )
-# with(
-#   rdd(tech_class_pcoa_scores),
-#   text(x = MDS1*2, y = MDS2*2, labels = abbrev)
-# )
-# legend(
-#   "bottomleft",
-#   legend = c("Illumina Short", "PacBio Short", "PacBio Long"),
-#   col = c("red", "green", "blue"),
-#   pch = c(3, 1, 2)
-# ) 
-# 
-# plot(rdd(tech_ecm_fam_pcoa), type = "none")
-# points(
-#   rdd(tech_ecm_fam_pcoa),
-#   select = with(
-#     rdd(tech_ecm_fam_sample_data),
-#     tech == "PacBio" & amplicon == "Short"
-#   ),
-#   col = "green",
-#   pch = 1
-# )
-# points(
-#   rdd(tech_ecm_fam_pcoa),
-#   select = with(
-#     rdd(tech_ecm_fam_sample_data),
-#     tech == "PacBio" & amplicon == "Long"
-#   ),
-#   col = "blue",
-#   pch = 2
-# )
-# points(
-#   rdd(tech_ecm_fam_pcoa),
-#   select = with(
-#     rdd(tech_ecm_fam_sample_data),
-#     tech == "Illumina" & amplicon == "Short"
-#   ),
-#   col = "red",
-#   pch = 3
-# )
-# with(
-#   rdd(tech_ecm_fam_pcoa_scores),
-#   text(x = MDS1*3, y = MDS2*3, labels = abbrev)
-# )
-# legend(
-#   "bottomleft",
-#   legend = c("Illumina Short", "PacBio Short", "PacBio Long"),
-#   col = c("red", "green", "blue"),
-#   pch = c(3, 1, 2)
-# )
-```
-
-```{r}
-tech_class_nmds <- vegan::metaMDS(
-  phyloseq::otu_table(rdd(tech_class_table))
-)
-
-tech_class_nmds_scores <-
-  vegan::scores(tech_class_nmds, display = "species") %>%
-  as_tibble(rownames = "class") %>%
-  mutate(L = sqrt(NMDS1^2 + NMDS2^2)) %>%
-  filter(!is.nan(L)) %>%
-  filter(L >= max(L) / 2) %>%
-  mutate(abbrev = substr(class, start = 1, stop = 2))
-
-plot(tech_class_nmds, type = "n")
-vegan::ordispider(
-  tech_class_nmds,
-  groups = with(
-    rdd(tech_class_sample_data),
-    paste(buffer, year, site, x)
-    ),
-  col = "gray80"
-  )
-text(
-  x = tech_class_nmds_scores$NMDS1,
-  y = tech_class_nmds_scores$NMDS2,
-  labels = tech_class_nmds_scores$abbrev
-  )
-points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
-                                     tech == "PacBio" & amplicon == "Short" ),
-       col = "green", pch = 1)
-points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
-                                     tech == "PacBio" & amplicon == "Long"),
-       col = "blue", pch = 2)
-points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
-                                     tech == "Illumina" & amplicon == "Short"),
-       col = "red", pch = 3)
-legend("topright", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
-```
-
-### Positive control
-
-*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
-Additional ASVs belonging to the genus *Agaricus* were also detected in the same wells as the positive control in the Ion Torrent dataset, but these were sufficiently different that they probably represent additional *Agaricus* species from the field samples, and were found in the same wells due to incomplete demultiplexing.
-*A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair.
-The abundance of the second ASV was `r rdd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
-> 0) %>% percent(format = "f", digits = 2) %>% paste(collapse = " and ")` that of the primary ASV in the two Illumina positive controls.
-The consistency of this ratio across PCR replicates suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
-
-The primary *A.
-bisporus* ASV was the most abundant ASV in the long-amplicon PacBio dataset and the Ion Torrent dataset, and the fifth most abundant ASV in the short-amplicon PacBio dataset, and the third most abundant ASV in the Illumina dataset.
-This ASV was also found in `r rdd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, representing `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.
+```
 
@@
 
-Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure\ \@ref(fig:correlog).
-Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 2--3\ m for the total fungal community, and 1--2\ m for the ECM community.
-In both cases, the greatest correlation magnitudes and distances are associated with Illumina sequencing of the short amplicon library, followed by PacBio sequencing of the long amplicon library, with PacBio sequencing of the short amplicon library detecting the least spatial structure.
-
-In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1\ m 
+Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at $p < 0.05$ for ranges of up to 2--3\ m for the total fungal community, and 1--2\ m for the ECM fungal community (Figure\ \@ref(fig:correlog)).
+In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
+The least spatial structure was detected with PacBio short amplicon sequencing.
+
+The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0\ m, time lag of 1\ year), for both the total fungal and ECM fungal communities in the long amplicon library.
+For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
+This puzzling negative correlation was significant in all correlograms based on short amplicon sequencing irrespective of technology.
+
+In contrast to the Bray-Curtis distance, the weighted UNIFRAC distance showed very little spatial structure, with only the total fungal community in the 1\ m 
 distance class showing a significant correlation at $p < 0.05$.
-
-The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1\ year and 0\ m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
-For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
-This negative correlation was significant in all short amplicon correlograms.
-
-Distance-decay curves relating community dissimilarities and spatial distances are shown in Figure\ \@ref(fig:spatio).
-For the cases where the Mantel correlogram detected a significant positive correlation, the best-fit exponential curve is also displayed; for the long amplicon library, where there was also a significant positive correlation after a time lag of one year, the best fit curve for time lag is also included.
-The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in Table\ \@ref(tab:variofit).
-Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m was included in the 95% confidence interval.
-Although the best fit spatial ranges vary between the different datasets by a factor of about 3, there is overlap of the 95% confidence intervals for all of the Bray-Curtis variograms across communities, amplicon libraries, and sequencing technologies.
+No temporal correlation was found for the weighted UNIFRAC distance.
+
+The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranged in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
+Although a distance-decay model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
 
@@
 ```{r variofit, results="asis"}
-rdd(variog_data) %>%
-  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
-  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
-  mutate(params = map(variofit, broom::tidy)) %>%
-  unnest(params) %>%
-  bind_cols(
-    pmap_dfr(
-      .,
-      function(mantel, variofit, term, ...) {
-        if (mantel) {
-          tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% inset2("term", term) %>% as.data.frame,
-                   error = function(e) tibble(term = term, conf.low = NA_real_, conf.high = NA_real_))
-        } else {
-          tibble(term = term, conf.low = NA_real_, conf.high = NA_real_)
-        }
-      }) %>%
-      select(-term)
-  ) %>%
-  filter(term %in% c("range", "timerange")) %>%
+ rdd(variofit_table) %>%
   mutate_at("conf.low", replace_na, 0) %>%
@@
 The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
-Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 1)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 1)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 1)`),
+Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 0)`),
 the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
 
-DADA2 denoising to form ASVs may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
+DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
 The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
 
-Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, no such amplicons were sequenced by any of the technologies, and Illumina showed a slightly higher fraction of longer sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Supplementary Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
-
-`r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` of long amplicon reads beloged to ASVs which were not found in any short amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
-Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences which were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
-Taxonomic assignments reveal that many of these non-shared sequences fall outside kingdom Fungi (Supplementary Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Supplementary Figure\ \@ref(fig:short-its)).
-Within Fungi, the short amplicon datasets recovered more Sordariomycetes, possibly because this group was represented by diverse but rare taxa which were not detected in the PacBio datasets due to the lower sequencing depth.
+Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any ITS2 fragments longer than those recovered by Illumina and Ion Torrent.
+Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
+Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Supplementary Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
+
+Of long amplicon reads, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
+Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
+Taxonomic assignments revealed that many of these non-shared sequences fall outside kingdom Fungi (Supplementary Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Supplementary Figure\ \@ref(fig:short-its)).
+Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figure \@ref(fig:pcoa)), possibly because this group was represented by diverse but rare taxa which remained undetected in the PacBio datasets due to the lower sequencing depth.
 Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella* in the long amplicon dataset (Supplementary Figure \@ref(fig:ecm-heattree)), and *Xylaria* and *Myerozyma* in the short amplicon datasets (Supplementary Figure\ \@ref(fig:heattree-length-compare)).
 These differences may be due to primer mismatches in these groups.
-However, neither amplicon recovered any sequences identifiable to *Cantharellus*, an ECM genus which is known from the area from fruitbodies.
 
@@
 
-ASV recovery for long amplicons was dramatically improved (`r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` to `r percent((rdd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")`) by denoising homologous subregions independently.
-Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs, so this approach should increase ASV recovery from these platforms as well.
+ASV recovery for long amplicons using DADA2 was dramatically improved (`r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` to `r percent((rdd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of reads) by denoising homologous subregions independently using the `lsux` and `tzara` packages.
+Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
+Thus, `tzara` should increase ASV recovery from these platforms as well.
 It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding [@loit2019].
@@
 
-The RDP fungal training set and Unite performed comparably at placing long amplicon sequences.
+The RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences.
 The Warcup database placed notably fewer sequences at all taxonomic levels for all datasets, probably in part due to the fact that only fungal sequences are included.
-However, even with this taken into account, IDTAXA performed very poorly with the Warcup database, placing less than 25% of ASVs to kingdom in all datasets.
+However, even with this considered, IDTAXA performed very poorly with the Warcup database, placing <25% of ASVs to kingdom in all datasets.
 IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores [@murali2018a].
@@
 For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
-For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()`.
+For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
 This led to a corresponding increase in the fraction of reads assigned to a functional guild (Supplementary Figure\ \@ref(fig:ecm)).
@@
 
-Weighted UNIFRAC did not reliably detect spatial structure within this relatively homogeneous community.
+Weighted UNIFRAC did not reliably detect spatial structure within this relatively ecologically homogeneous community.
 Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m.
@@
 Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 13--19\ m.
-The 95% confidence interval was substantially wider than this variation, generally covering a range of 6--35\ m.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from `r rdd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+The 95% confidence interval was substantially wider than this variation, generally covering a range of `r rdd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
 All of these values are smaller than the 65\ m reported by @Bahram2013, also from an ECM woodland habitat in Benin.
-The PacBio short amplicon dataset shows a longer spatial range, of 30\ m for the total fungal community and 42--43\ m for the ECM community, in both cases with wide confidence intervals spanning approximately 15--60\ m.
+The PacBio short amplicon dataset shows a longer spatial range, of `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the total fungal community and `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the ECM community, in both cases with wide confidence intervals spanning `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
 It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
@@
 The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0\ m and 1\ year.
-The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.1\ years for the total fungal comunity and 4\ years for the ECM community, but with overlapping confidence intervals.
+The spatiotemporal distance-decay fit estimated the temporal turnover range as `r rdd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r rdd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
 Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
@@
 However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments.
-DADA2 ASV yield was initially poor for long reads, but this was improved through extraction of subregions, seperate denoising, and then reconstruction of full-length sequences.
-This suggests a hybrid approach using long-read sequencing to aquire long amplicon sequences for the local species pool, and cost-effective short-read sequencing to provide high sampling depth and sample number.
+DADA2 ASV yield was initially poor for long reads, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
+Together these approaches provide a hybrid approach using long-read sequencing to acquire long amplicon sequences for the local species pool, and cost-effective short-read sequencing to provide high sampling depth and sample number.
+
+# Data Availability Statement
+
+- Raw data?
+- ASV sequences to Unite
+- alignment and trees to TreeBase
+- R packages on CRAN or Dryad (plus Github)
+- Analysis scripts on Github
+
+# Author contributions
+
+Sampling was planned and carried out by BF, NSY, and MR.
+Bioinformatics and data analysis were performed by BF with input from MB, AR, and MR.
+Scripts and R packages were written by BF.
+The manuscript was drafted by BF and MR.
+All authors contributed to and approved the final version of the manuscript.
 
@@
 Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden.
+
+This project was funded by Formas grant XXX
+
+Ylva for PCR and sequence prep.
 
@@
 long = ITS--LSU.
-For this figure, subregions were concatenated prior to quality filtering and generating ASVs with `DADA2`.
+For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.
 

### 2020-03-03 writing/transect_paper.Rmd
Add class taxonomy chart, switch back to Molecular Ecology

@@
   - Brendan Furneaux^1^,
-    Mohammad Bahram^2,3^,
-    Anna Rosling^4^,
-    Nourou S. Yorou^5^,
-    Martin Ryberg^1^
+  -  Mohammad Bahram^2,3^,
+  - Anna Rosling^4^,
+  - Nourou S. Yorou^5^,
+  - Martin Ryberg^1^
   - ^1^Program in Systematic Biology,
@@
     Uppsala University,
-    Uppsala, Sweden
+  - Uppsala, Sweden
   - ^2^Department of Ecology,
@@
     Uppsala University,
-    Uppsala, Sweden
+  - Uppsala, Sweden
   - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
     LEB,
-    University of Parakou,
+  - University of Parakou,
     Parakou, Benin
@@
   - "R.bib"
-csl: new-phytologist.csl
+csl: molecular-ecology.csl
 always_allow_html: true
@@
 
-# Methods
+# Materials and Methods
 
@@
 
-(ref:taxa) Color and size of nodes represent the fractional read abundance of all ASVs belonging to that taxonomic group.
-Color and thickness of branches represent the fractional ASV richness of that taxonomic group.
+(ref:taxa) Color and size of nodes represent the fractional ASV richness of that taxonomic group.
+Color and thickness of branches represent the fractional read abundance of all ASVs belonging to that taxonomic group.
 Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
@@
 theme_set(theme)
+```
+
+(ref:fungi-comm-s) Taxonomic composition of fungal community at the class level
+
+(ref:fungi-comm) Values represent the fraction of all ASVs, OTUs, or reads which were assigned to kingdom Fungi.
+Assignments based on PHYLOTAX.
+Classes which represented less than 2% of reads, OTUs, and ASVs in all datasets are grouped together as "other".
+
+```{r fungi-comm, fig.cap = "(ref:fungi-comm-s). (ref:fungi-comm)", fig.scap = "(ref:fungi-comm-s)"}
+rdd(taxon_reads) %>%
+  filter(
+    kingdom == "Fungi",
+    !(Algorithm == "Consensus" & region == "ITS"),
+    Algorithm == "PHYLOTAX"
+    ) %>%
+  group_by(tech, amplicon) %>%
+  mutate(reads = reads/sum(reads), ASVs = 1/n()) %>%
+  pivot_longer(cols = c("reads", "ASVs"), names_to = "type", values_to = "reads") %>%
+  bind_rows(
+    filter(., type == "ASVs", label == OTU) %>%
+      mutate(label = OTU) %>%
+      group_by(tech, amplicon) %>%
+      mutate(reads = 1/n(), type = "OTUs")
+  ) %>%
+  taxon_plot(
+    rank = class,
+    datasets = datasets,
+    cutoff = 0.02,
+    y = reads,
+    x = type,
+    facets = vars(tech, amplicon)
+  ) +
+  ylab("Fraction of Community") +
+  xlab(NULL)
 ```
@@
 
-# Conclusion
+## Conclusion
 
@@
 
-# Data Availability Statement
+# Acknowledgements {-}
+
+The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
+Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden.
+
+This project was funded by Formas grant XXX
+
+Ylva for PCR and sequence prep.
+
+# References {-}
+
+<div id="refs"></div>
+
+# Data Availability Statement {-}
 
@@
 
-# Author contributions
+# Author contributions {-}
 
@@
 All authors contributed to and approved the final version of the manuscript.
-
-# Acknowledgements
-
-The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
-Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden.
-
-This project was funded by Formas grant XXX
-
-Ylva for PCR and sequence prep.
-
-# Bibliography {-}
-
-<div id="refs"></div>
 
@@
 
-(ref:heattree-length-compare) Red nodes (edges) represent increased read (ASV) count for the short amplicon library, while blue-green nodes (edges) represent increased read (ASV) count for the long amplicon library.
+(ref:heattree-length-compare) Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
+In both cases red represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
 Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
@@
 
-(ref:ecm-heattree) Red edges/nodes represent increased read/ASV count for the short amplicon library, while teal edges/nodes represent increased read/ASV count for the long amplicon library.
+(ref:ecm-heattree) Red edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
 Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.

### 2020-03-04 writing/transect_paper.Rmd
finish (?) addressing coauthor comments

@@
 While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
-Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa (**ref**).
+Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa.
 Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].
-Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more conserved SSU and/or LSU regions [@Tedersoo2018].
-This can potentially improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
-However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
-
-In addition **XXX**.
+
 Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
-In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while leaving intra-species variation.
+In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while preserving intra-species variation.
 The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
@@
 
+Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions [@Tedersoo2018].
+This can potentially improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
 Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
@@
 However, long amplicon sequencing allows the inclusion of alignable regions for construction of more fully resolved phylogenetic trees directly from metabarcoding reads.
-
-Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatial community turnover rate along a transect in an ECM-dominated Soudanian woodland in Benin.
-
-Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
-Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
-It varies between different ecosystems and taxonomic groups.
-Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
-distance [@legendre2012].
-
-Several studies have investigated the turnover scale in ECM forests, but so far they have been based on analysis of ECM root tips, rather than bulk soil.
-@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites in temperate forests, using Sanger sequencing.
-Similarly, @pickles2012a found autocorrelation at distances <3.4\ m, also in temperate forests, based on T-RFLP analysis.
-Two studies have included data from tropical African woodlands and wooded savannahs.
-@tedersoo2011 found autocorrelation at ranges <10\ m in Miombo woodland using Sanger sequencing.
-However, this study was based on irregular sampling with a minimum spacing of 8\ m, so this as the smallest distance class.
-@Bahram2013 performed a meta-analysis of turnover scale in ECM communities, including data from Benin, where they found turnover range of 65\ m based on distance-decay curves.
-Based on these results they concluded that the range of distance decay is greater in tropical forests than temperate forests.
-
-We investigated turnover range in both the total fungal community and ECM community by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
+However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
+
+Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated Soudanian woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
 We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs.
@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples differed significantly (p < `r rdd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")` from permutational MANOVA using 9999 permutations, $R^2 = `r rdd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Supplementary Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
+Sequencing results for these samples differed significantly (p < `r rdd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")` from permutational MANOVA using 9999 permutations, $R^2 = `r rdd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
 However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
@@
 For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
-Because ITS2 is the most variable of the amplified regions (Supplementary Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
+Because ITS2 is the most variable of the amplified regions (Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
 The concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015].
@@
 
-Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure\ \@ref(fig:phylotax)).
+Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax)).
 A taxon at a particular rank was assigned to a node and all its descendants if that taxon was consistent with the reference-based taxonomic assignments for each of the descendants.
@@
 
+Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
+Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
+It varies between different ecosystems and taxonomic groups.
+Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
+distance [@legendre2012].
+
 Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
@@
 
-DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Supplementary Figure \@ref(fig:sample-depth).
+DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Figure \@ref(fig:sample-depth).
 Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads.
@@
 
-Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Supplementary Figure\ \@ref(fig:full-length)a).
-Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure\ \@ref(fig:full-length)b).
+Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Figure\ \@ref(fig:full-length)a).
+Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure\ \@ref(fig:full-length)b).
 The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r rdd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(rdd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
@@
 
-The length distribution of the different regions extracted from the long amplicon are shown in Supplementary Figure\ \@ref(fig:region-lengths).
+The length distribution of the different regions extracted from the long amplicon are shown in Figure\ \@ref(fig:region-lengths).
 ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r rdd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r rdd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r rdd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r rdd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r rdd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
-Approximately `(rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Supplementary Figure\ \@ref(fig:region-lengths) due to rarity.
+Approximately `(rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure\ \@ref(fig:region-lengths) due to rarity.
 Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.
@@
 According to the PHYLOTAX assignments, Fungi represented `r percent(fungi_asv["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(fungi_read["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(fungi_asv[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(fungi_read[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
-Fungi categorized as ECM made up `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
-Assignment to kingdom Fungi and ECM status for different amplicons, sequencing technologies, reference databases, and algorithms are shown in Supplementary Figure\ \@ref(fig:ecm).
+Fungi categorized as ECM made up `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library (Figure\ \@ref(fig:ecm)).
 
@@
 
-## Comparison of sequencing strategies
-
-The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
-Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 0)`),
-the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
-
-DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
-The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
-
-Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any ITS2 fragments longer than those recovered by Illumina and Ion Torrent.
-Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
-Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Supplementary Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
-
-Of long amplicon reads, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
-Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
-Taxonomic assignments revealed that many of these non-shared sequences fall outside kingdom Fungi (Supplementary Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Supplementary Figure\ \@ref(fig:short-its)).
-Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figure \@ref(fig:pcoa)), possibly because this group was represented by diverse but rare taxa which remained undetected in the PacBio datasets due to the lower sequencing depth.
-Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella* in the long amplicon dataset (Supplementary Figure \@ref(fig:ecm-heattree)), and *Xylaria* and *Myerozyma* in the short amplicon datasets (Supplementary Figure\ \@ref(fig:heattree-length-compare)).
-These differences may be due to primer mismatches in these groups.
-
 ## Reconstruction of long amplicons from denoised subregions
@@
 
+## Comparison of sequencing strategies
+
+The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
+Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 0)`),
+the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
+
+DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
+The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
+
+Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any ITS2 fragments longer than those recovered by Illumina and Ion Torrent.
+Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
+Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
+
+Of long amplicon reads, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
+Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
+Taxonomic assignments revealed that the majority of these non-shared sequences fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Figure\ \@ref(fig:short-its)).
+Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figure \@ref(fig:pcoa) and \@ref(fig:heattree-length-compare)).
+Additionally, several smaller groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
+These differences may be due to primer mismatches in these taxa.
+
 ## Taxonomic identification
@@
 @gdanetz2017 showed that a majority-rule consensus of three assignment algorithms can improve the fraction of sequences assigned as well as decrease the false assignment rate.
-Here we find that strict consensus, which should provide more conservative taxonomic assignments, also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications.
+Strict consensus rejects assignments whenever there is conflict between methods and should therefore provide more conservative taxonomic assignments than majority-rule consensus.
+Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications.
 This suggests that different assignment algorithms and databases bring mostly complementary, non-contradictory information at higher taxonomic levels.
@@
 For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
-This led to a corresponding increase in the fraction of reads assigned to a functional guild (Supplementary Figure\ \@ref(fig:ecm)).
+This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure\ \@ref(fig:ecm)).
 
@@
 Mantel correlograms based on the Bray-Curtis dissimilarity (Figure\ \@ref(fig:correlog)) revealed spatial autocorrelation in the soil fungal community at distance classes $\le 3$ m for both Illumina and PacBio using long and short amplicons, and in the ECM fungal community at distance classes $\le 2$ m for Illumina and PacBio long amplicons, and $\le 1$ m for the PacBio short amplicons.
-These values are similar to the values reported for ECM root tips in temperate forests by @lilleskov2004 and @pickles2012a, but substantially smaller than those observed by @tedersoo2011.
+These results are similar to autocorrelation ranges found in previous work based on ECM root tips in temperate forests [@lilleskov2004; @pickles2012a].
+@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites using Sanger sequencing.
+Similarly, @pickles2012a found autocorrelation at distances <3.4\ m based on T-RFLP analysis.
+However, previous work in Miombo woodland, a similar ecosystem to the Soudanian woodland in this study, found autocorrelation at ranges <10\ m using Sanger sequencing of ECM root tips [@tedersoo2011], which was their smallest distance class.
+
 Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
@@
 The 95% confidence interval was substantially wider than this variation, generally covering a range of `r rdd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
-All of these values are smaller than the 65\ m reported by @Bahram2013, also from an ECM woodland habitat in Benin.
+All of these values are smaller than the 65\ m reported by @Bahram2013, also based on distance-decay curves from an ECM woodland habitat in Benin.
+
 The PacBio short amplicon dataset shows a longer spatial range, of `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the total fungal community and `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the ECM community, in both cases with wide confidence intervals spanning `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
@@
 
+This project was funded by the Swedish research council FORMAS grant number 2014-01109.
+
+Laboratory work including PCR and library pooling was performed by Dr. Ylva Strid.
+
 The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
@@
 
-This project was funded by Formas grant XXX
-
-Ylva for PCR and sequence prep.
+Sequencing was performed by the SNP&SEQ Technology Platform in Uppsala.
+The facility is part of the National Genomics Infrastructure (NGI) Sweden and Science for Life Laboratory.
+The SNP&SEQ Platform is also supported by the Swedish Research Council and the Knut and Alice Wallenberg Foundation.
 
@@
 
-- Raw data?
+Sequence data, including raw reads and ASV sequences, are deposited at NCBI under BioProject XXXXXXX.
+Sequence alignment and phylogenetic tree are available at TreeBase under project S25885
 - ASV sequences to Unite

### 2020-03-05 writing/transect_paper.Rmd
data availability

@@
 Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in **xxx**.
-We investigated spatio-temporal structure of an ECM fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
+We investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
 Resulting community structure and diversity depended more on statistical method than sequencing technology.
@@
 Sequence data, including raw reads and ASV sequences, are deposited at NCBI under BioProject XXXXXXX.
-Sequence alignment and phylogenetic tree are available at TreeBase under project S25885
-- ASV sequences to Unite
-- alignment and trees to TreeBase
-- R packages on CRAN or Dryad (plus Github)
-- Analysis scripts on Github
+Sequence alignment and phylogenetic tree are available at TreeBase under project S25885.
+R packages `lsux`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/lsux, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
+Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.
 

### 2020-03-05 writing/transect_paper.Rmd
finish rewritten abstract

@@
 
-Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in **xxx**.
+Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in taxonomic and phylogenetic placement of fungi from environmental sequences.
 We investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
 Resulting community structure and diversity depended more on statistical method than sequencing technology.
-We found significant autocorrelation in distance classes <3\ m using a Mantel correlogram, and a best-fit distance-decay curve range of 17--19\ m, both shorter than previously reported for similar habitats.
 The use of long-amplicon sequencing enables construction of a phylogenetic tree from metabarcoding reads, which facilitates taxonomic identification of sequences.

### 2020-03-06 writing/transect_paper.Rmd
change cache locations to make knitting work better in light plan

@@
 
-# use the "light" plan cache instead of the default cache at .drake
-cache <- drake_cache(here(".light"))
-rdd <- function(..., cache = drake_cache(here(".light"))) {
-  force(cache)
-  drake::readd(..., cache = cache)
-}
-ldd <- function(..., cache = drake_cache(here(".light"))) {
-  force(cache)
-  drake::loadd(..., cache = cache, envir = parent.frame())
-}
-
 light_plan <- readRDS(here("data/plan/drake_light.rds"))
@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples differed significantly (p < `r rdd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")` from permutational MANOVA using 9999 permutations, $R^2 = `r rdd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
+Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
 However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
@@
 We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in @callahan2019.
-However, only `r rdd(region_table)$ASVs[rdd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
+However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
 We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
@@
 Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median $L=`r rdd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(rdd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - rdd(demuxqual)['pb_500'])^rdd(demuxlength)['pb_500'])), big.mark = ",")`.
-Given the high diversity relative to sequencing depth in this study (`r rdd(region_table)$ASVs[rdd(region_table)$seq_run == "pb_483" & rdd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r rdd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
-In comparison, the equivalent requirement for the short amplicon ($L=`r rdd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r rdd(demuxqual)['pb_483']`$) is only `r round(2/(1 - rdd(demuxqual)['pb_483'])^rdd(demuxlength)['pb_483'], 1)` reads.
+For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
+Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
+In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
 We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
@@
 In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
-Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(rdd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(rdd(tech_class_table))` classes), were included.
 Permutational MANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs.
@@
 PacBio RS II).
+Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
 The marginal significance of each term for explaining variation in the Bray-Curtis community dissimilarity matrix was performed using the `adonis2` function in the R package `vegan` [@R-vegan], with 9999 permutations.
@@
 Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
-Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-rdd(filtercounts_full)/rdd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long reads, but only by `r (1-rdd(filtercounts_full)/rdd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short reads, resuling in a similar numbe rof filtered reads for the two strategies.
-Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-rdd(filtercounts_full)/rdd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
-In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-rdd(filtercounts_ITS2)['pb_500']/rdd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-rdd(filtercounts_ITS2)['pb_483']/rdd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-rdd(filtercounts_ITS2)['is_057']/rdd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
+Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long reads, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short reads, resulting in a similar number of filtered reads for the two strategies.
+Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
+In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-readd(filtercounts_ITS2)['pb_483']/readd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-readd(filtercounts_ITS2)['is_057']/readd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
 This represented greater loss of PacBio short reads, but less loss of PacBio long reads and Ion Torrent reads.
@@
 Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure\ \@ref(fig:full-length)b).
-The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r rdd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(rdd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
-In contrast, the length of the long amplicon reads varied widely, from `r filter(rdd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(rdd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
+The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
+In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
 
 The length distribution of the different regions extracted from the long amplicon are shown in Figure\ \@ref(fig:region-lengths).
-ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r rdd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r rdd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r rdd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r rdd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r rdd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r rdd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
-Approximately `(rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/rdd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure\ \@ref(fig:region-lengths) due to rarity.
+ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
+Approximately `r (readd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/readd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure\ \@ref(fig:region-lengths) due to rarity.
 Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.
@@
 bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair (`r Biostrings::readDNAStringSet(here::here("data/clusters/ITS2.fasta.gz")) %>% magrittr::extract(startsWith(names(.), "f2ab49e8")) %>% Biostrings::width() %>% unique() %>% divide_by(1, .) %>% subtract(1, .) %>% percent(digits = 1, format = "f")` similarity in ITS2).
-The abundance of the second ASV was `r rdd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
+The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
 > 0) %>% percent(format = "f", digits = 1) %>% paste(collapse = " and ")` that of the primary *A.
@@
 *A.
-bisporus* sequences represented `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively, giving similar estimates for the rate of tag-switching for all technologies.
+bisporus* sequences represented `r readd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively, giving similar estimates for the rate of tag-switching for all technologies.
 These reads were excluded from community analyses.
@@
 ```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", out.width = "70%", fig.height = 4, fig.align="center", results="asis"}
-venn_A <- venndata(rdd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+venn_A <- venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
   vennplot_data(ASVs) %>% {
@@
 
-venn_B <- venndata(rdd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
+venn_B <- venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
   vennplot_data(OTUs) %>% {
@@
 
-venn_C <- rdd(asv_table) %>%
+venn_C <- readd(asv_table) %>%
   mutate_if(is.numeric, ~ . / sum(.)) %>%
@@
 
-venn_D <- rdd(otu_table) %>%
+venn_D <- readd(otu_table) %>%
   mutate_if(is.numeric, ~ . / sum(.)) %>%
@@
   read_comparison(
-    rdd(multi_table),
-    rdd(comparisons),
+    readd(multi_table),
+    readd(comparisons),
     type = "ASV"
@@
   read_comparison(
-    rdd(multi_table),
-    rdd(comparisons),
+    readd(multi_table),
+    readd(comparisons),
     type = "OTU"
@@
 The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Figure\ \@ref(fig:venn)).
-ASVs shared between all datasets represented `r paste(range(as.numeric(unlist(rdd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
-These fractions increased to `r paste(range(as.numeric(unlist(rdd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
-In particular, `r text_list(percent(colSums(mutate_all(rdd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
-In contrast, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
+ASVs shared between all datasets represented `r paste(range(as.numeric(unlist(readd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
+These fractions increased to `r paste(range(as.numeric(unlist(readd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+In particular, `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
+In contrast, `r percent(readd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
 Complete tabulations of the number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively.
@@
 ```{r tax-chart, fig.cap = "(ref:tax-chart)", fig.scap = "(ref:tax-chart-s)"}
-rdd(tax_chart_plot)
+readd(tax_chart_plot)
 ```
@@
 ```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8}
-fungi_read <- rdd(taxdata) %>%
+fungi_read <- readd(taxdata) %>%
   taxa::filter_taxa(
@@
 theme <- theme_update(panel.border = element_blank())
-rdd(heattree)
+readd(heattree)
 theme_set(theme)
@@
 ```{r fungi-comm, fig.cap = "(ref:fungi-comm-s). (ref:fungi-comm)", fig.scap = "(ref:fungi-comm-s)"}
-rdd(taxon_reads) %>%
+readd(taxon_reads) %>%
   filter(
@@
 
-Taxonomic composition of the sequenced soil community is summarized in Figure\ \@ref(fig:taxa).
-According to the PHYLOTAX assignments, Fungi represented `r percent(fungi_asv["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(fungi_read["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(fungi_asv[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(fungi_read[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
-Fungi categorized as ECM made up `r percent(ecm_asv["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(ecm_read["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(ecm_asv[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(ecm_read[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library (Figure\ \@ref(fig:ecm)).
+Taxonomic composition of the sequenced soil fungal community at the class level is is summarized in Figure\ \@ref(fig:fungi-comm) and as a heat tree [@foster2017] in Figure\ \@ref(fig:taxa).
+According to the PHYLOTAX assignments, Fungi represented `r percent(readd(fungi_asv)["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(readd(fungi_read)["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(readd(fungi_asv)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(fungi_read)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
+Measured fungal community composition at the class level varied significantly between amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
+The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm) and \@ref(fig:ppcoa)).
 
@@
 ggpubr::ggarrange(
-  rdd(tech_class_pcoa_plot),
-  rdd(tech_ecm_fam_pcoa_plot),
+  readd(tech_class_pcoa_plot),
+  readd(tech_ecm_fam_pcoa_plot),
   common.legend = TRUE,
@@
   ) %>%
-  mutate_at("target", purrr::map, rdd, character_only = TRUE) %>%
+  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
   mutate_at("target", purrr::map, "mantel.res") %>%
@@
 ```{r spatio, fig.cap = "(ref:spatio-s). (ref:spatio)", fig.scap = "(ref:spatio-s)"}
-
-rdd(variog_empir) %>%
+readd(variog_empir) %>%
   filter(algorithm == "PHYLO") %>%
@@
   # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
-  geom_line(data = filter(rdd(variog_fit), timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
-  geom_line(data = filter(rdd(variog_fit), timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
+  geom_line(data = filter(readd(variog_fit), timelag == "1", algorithm == "PHYLO"), size = 0.5, color = "darkcyan", alpha = 0.8) +
+  geom_line(data = filter(readd(variog_fit), timelag == "0", algorithm == "PHYLO"), size = 0.5, color = "red1", alpha = 0.8) +
   scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
@@
 ```{r variofit, results="asis"}
- rdd(variofit_table) %>%
+ readd(variofit_table) %>%
   mutate_at("conf.low", replace_na, 0) %>%
@@
 
-ASV recovery for long amplicons using DADA2 was dramatically improved (`r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` to `r percent((rdd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of reads) by denoising homologous subregions independently using the `lsux` and `tzara` packages.
+ASV recovery for long amplicons using DADA2 was dramatically improved (`r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of reads) by denoising homologous subregions independently using the `LSUx` and `tzara` packages.
 Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
@@
 The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
-Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 0)`),
+Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(readd(bioinf_table)["ITS2","PacBio_Short_reads"]/readd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(readd(bioinf_table)["ITS2","Illumina_Short_reads"]/readd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(readd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/readd(rawcounts)["is_057"], format = "f", digits = 0)`),
 the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
@@
 
-Of long amplicon reads, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
+Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])`.
 Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
@@
 Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from `r rdd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
-The 95% confidence interval was substantially wider than this variation, generally covering a range of `r rdd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+The 95% confidence interval was substantially wider than this variation, generally covering a range of `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
 All of these values are smaller than the 65\ m reported by @Bahram2013, also based on distance-decay curves from an ECM woodland habitat in Benin.
 
-The PacBio short amplicon dataset shows a longer spatial range, of `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the total fungal community and `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the ECM community, in both cases with wide confidence intervals spanning `r rdd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+The PacBio short amplicon dataset shows a longer spatial range, of `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the ECM community, in both cases with wide confidence intervals spanning `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
 It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
@@
 The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0\ m and 1\ year.
-The spatiotemporal distance-decay fit estimated the temporal turnover range as `r rdd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r rdd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
+The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
 Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
@@
 ```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)"}
-rdd(phylotax_fig)
+readd(phylotax_fig)
 ```
@@
 ggpubr::ggarrange(
-  rdd(full_dens_short),
-  rdd(full_ecdf_short),
-  rdd(full_dens_long),
-  rdd(full_ecdf_long),
+  readd(full_dens_short),
+  readd(full_ecdf_short),
+  readd(full_dens_long),
+  readd(full_ecdf_long),
   ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
@@
 ```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)"}
-rdd(region_table) %>%
+readd(region_table) %>%
   dplyr::filter(seq_run == "pb_500") %>%
-  mutate(reads = reads / sum(filter(rdd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
+  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
   dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
@@
 ```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6}
-rdd(region_length_fig)
+readd(region_length_fig)
 ```
@@
 ggpubr::ggarrange(
-  rdd(its2_dens),
-  rdd(its2_ecdf),
+  readd(its2_dens),
+  readd(its2_ecdf),
   ncol = 1,
@@
 ```{r venn-table-asv}
-rdd(venn_ASV) %>%
+readd(venn_ASV) %>%
 venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
@@
 ```{r venn-table-otu}
-rdd(venn_OTU) %>%
+readd(venn_OTU) %>%
 venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
@@
 ```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)"}
-rdd(big_seq_table_ITS2) %>%
+readd(big_seq_table_ITS2) %>%
   colSums() %>%
@@
   mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(rdd(taxon_reads), ., by = c("label" = "ITS2")) %>%
+  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
   # group_by(ITS2) %>%
@@
 oldtheme <- theme_update(panel.border = element_blank())
-rdd(heattree_length_compare)
+readd(heattree_length_compare)
 theme_set(oldtheme)
@@
 ```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)"}
-  rdd(taxon_reads) %>%
+  readd(taxon_reads) %>%
   filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
@@
 oldtheme <- theme_update(panel.border = element_blank())
-rdd(ecm_heattree)
+readd(ecm_heattree)
 theme_set(oldtheme)
@@
 ```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)"}
-reads_plot <- rdd(qstats_n) %>%
+reads_plot <- readd(qstats_n) %>%
   mutate(file = str_replace(file, ".*/", "") %>%
@@
   left_join(
-    rdd(proto_physeq) %>% {
+    readd(proto_physeq) %>% {
       mutate(
@@
 
-conc_plot <- rdd(proto_physeq) %>% 
+conc_plot <- readd(proto_physeq) %>% 
         sample_data(.) %>%
@@
 ```{r plate-loading, include=FALSE}
-rdd(proto_physeq) %>% {
+readd(proto_physeq) %>% {
   mutate(
@@
 ```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)"}
-plot(rdd(buffer_asv_pcoa), type = "none")
+plot(readd(buffer_asv_pcoa), type = "none")
 points(
-  rdd(buffer_asv_pcoa),
+  readd(buffer_asv_pcoa),
   select = with(
-    as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))),
+    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
     buffer == "Xpedition" & year == "2015" 
@@
 points(
-  rdd(buffer_asv_pcoa),
-  select = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))),
+  readd(buffer_asv_pcoa),
+  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                 buffer == "Xpedition" & year == "2016"
@@
 points(
-  rdd(buffer_asv_pcoa),
-  select = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))),
+  readd(buffer_asv_pcoa),
+  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                 buffer == "LifeGuard" & year == "2016"
@@
 #   "topright",
-#   legend = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))), paste(tech, amplicon)) %>%
+#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
 #     unique() %>%
@@
 # vegan::ordiellipse(
-#   rdd(buffer_asv_pcoa),
-#   groups = with(as_tibble(phyloseq::sample_data(rdd(buffer_asv_table))), paste(tech, amplicon)),
+#   readd(buffer_asv_pcoa),
+#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
 #   lty = 1:3
@@
 oldtheme <- theme_update(panel.border = element_blank())
-rdd(buffer_compare_long)
+readd(buffer_compare_long)
 theme_set(oldtheme)
@@
 ```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
-rdd(buffer_compare_short)
+readd(buffer_compare_short)
 theme_set(oldtheme)
 
-```
+```\processdelayedfloats

### 2020-03-06 scripts/drake-01-preITSx.R

@@ -36,7 +36,8 @@ if (length(preitsx_targets)) {
     cache_log_file = TRUE,
     targets = preitsx_targets,
     memory_strategy = "preclean",
-    console_log_file = get_logfile("preITSx")
+    console_log_file = get_logfile("preITSx"),
+    cache = cache
   )
   
   dod <- drake::outdated(dconfig)
@@ -48,7 +49,13 @@ if (length(preitsx_targets)) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+    )
+  )
   flog.info("Finished.")
 } else flog.info("All pre-itsx targets are up-to-date.")
 

### 2020-03-06 scripts/drake-02-ITSx.R

@@ -70,7 +70,8 @@ if (length(itsx_targets)) {
        caching = "worker",
        cache_log_file = TRUE,
        targets = itsx_targets,
-       console_log_file = get_logfile("ITSx")
+       console_log_file = get_logfile("ITSx"),
+       cache = cache
   )
   
   dod <- drake::outdated(dconfig)
@@ -80,7 +81,12 @@ if (length(itsx_targets)) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+  )
   flog.info("Finished.")
 } else flog.info("ITSx targets are up-to-date.")
 

### 2020-03-06 scripts/drake-03-preDADA.R

@@ -38,7 +38,8 @@ if (length(predada_targets)) {
        targets = predada_targets,
        caching = "worker",
        memory_strategy = "preclean",
-       console_log_file = get_logfile("predada")
+       console_log_file = get_logfile("predada"),
+       cache = cache
   )
   dod <- drake::outdated(dconfig)
   drake::make(config = dconfig)
@@ -47,7 +48,13 @@ if (length(predada_targets)) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+    )
+  )
   flog.info("Finished.")
 } else flog.info("Pre-DADA2 targets are up-to-date.")
 

### 2020-03-06 scripts/drake-04-errmodel.R

@@ -38,7 +38,8 @@ if (length(targets) > 0) {
     caching = "worker",
     cache_log_file = TRUE,
     targets = targets,
-    console_log_file = get_logfile("errmodel")
+    console_log_file = get_logfile("errmodel"),
+    cache = cache
   )
   dod <- drake::outdated(dconfig)
   drake::make(config = dconfig)
@@ -47,7 +48,13 @@ if (length(targets) > 0) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+    )
+  )
   flog.info("Finished.")
 } else flog.info("Error models are up-to-date.")
 

### 2020-03-06 scripts/drake-05-DADA.R

@@ -35,7 +35,8 @@ if (length(targets) > 0) {
        caching = "worker",
        cache_log_file = TRUE,
        targets = targets,
-       console_log_file = get_logfile("dada")
+       console_log_file = get_logfile("dada"),
+       cache = cache
   )
   dod <- drake::outdated(dconfig)
   drake::make(config = dconfig)
@@ -44,7 +45,13 @@ if (length(targets) > 0) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+    )
+  )
   flog.info("Finished.")
 } else flog.info("DADA targets are up-to-date.")
 

### 2020-03-06 scripts/drake-07-consensus.R

@@ -44,14 +44,21 @@ if (length(targets) > 0) {
     cache_log_file = TRUE,
     targets = targets,
     memory_strategy = "preclean",
-    console_log_file = get_logfile("consensus")
+    console_log_file = get_logfile("consensus"),
+    cache = cache
   )
   tictoc::toc()
   if (any(targets %in% drake::failed())) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+    )
+  )
   flog.info("Finished.")
 } else flog.info("Consensus targets are up-to-date.")
 

### 2020-03-06 scripts/drake-08-references.R

@@ -30,8 +30,8 @@ if (any(target %in% od) #|| !all(file.exists(outputs))
        elapsed = 3600, #1 hour
        keep_going = FALSE,
        cache_log_file = TRUE,
-       targets = target
-
+       targets = target,
+       cache = cache
   )
   dod <- drake::outdated(dconfig)
   drake::make(config = dconfig)

### 2020-03-06 scripts/drake-09-raxml.R

@@ -44,14 +44,21 @@ if (length(targets)) {
     cache_log_file = TRUE,
     targets = targets,
     console_log_file = get_logfile("raxml"),
-    prework = "options(expressions = 10000)"
+    prework = "options(expressions = 10000)",
+    cache = cache
   )
   tictoc::toc()
   if (any(targets %in% drake::failed())) {
     if (interactive()) stop() else quit(status = 1)
   }
   flog.info("Recalculating outdated targets...")
-  od <- drake::outdated(drake::drake_config(plan, jobs_preprocess = local_cpus()))
+  od <- drake::outdated(
+    drake::drake_config(
+      plan,
+      jobs_preprocess = local_cpus(),
+      cache = cache
+    )
+  )
   flog.info("Finished.")
 } else flog.info("RAxML targets are up-to-date.")
 

### 2020-03-06 scripts/drake-10-finish.R

@@ -44,7 +44,8 @@ if (length(targets)) {
        caching = "worker",
        cache_log_file = TRUE,
        console_log_file = get_logfile("finish"),
-       targets = targets
+       targets = targets,
+       cache = cache
   )
   dod <- drake::outdated(dconfig)
   drake::make(config = dconfig)
@@ -55,10 +56,9 @@ if (length(targets)) {
 }
 flog.info("All targets are up-to-date.")
 
-cache1 <- drake::drake_cache()
 
-cache2 <- drake::drake_cache(".light")
-if (is.null(cache2)) cache2 <- drake::new_cache(".light")
+cache2 <- drake::drake_cache(".drake")
+if (is.null(cache2)) cache2 <- drake::new_cache(".drake")
 exports <- intersect(
   c(
     "allseqs",
@@ -76,9 +76,9 @@ exports <- intersect(
     dplyr::filter(plan, step == "allchimeras")$target,
     "qstats"
   ),
-  cache1$list()
+  cache$list()
 )
 cache2$import(
-  from = cache1,
+  from = cache,
   list = exports
 )

### 2020-03-06 scripts/drake_plan_heavy.R

@@ -287,7 +287,7 @@ plan <- drake_plan(
   
   join_derep = target(
     tzara::combine_derep(
-      readd(derep1),
+      readd(derep1, cache = ignore(cache)),
       .data = dplyr::bind_rows(trace1)[ , c("seq_run", "plate", "well",
                                             "direction", "trim_file")]
     ),
@@ -428,7 +428,7 @@ plan <- drake_plan(
   err = target({
     err.fun <- if (tech == "PacBio" ) dada2::PacBioErrfun else
       dada2::loessErrfun
-    dereps <- readd(derep2) %>%
+    dereps <- readd(derep2, cache = ignore(cache)) %>%
       purrr::compact()
     dada2::learnErrors(
       fls = dereps,
@@ -451,7 +451,7 @@ plan <- drake_plan(
   ),
   
   err_illumina = target({
-    dereps <- readd(derep_illumina) %>%
+    dereps <- readd(derep_illumina, cache = ignore(cache)) %>%
       purrr::map(read) %>%
       purrr::compact()
     dada2::learnErrors(
@@ -476,9 +476,9 @@ plan <- drake_plan(
   
   # Run dada denoising algorithm (on different regions)
   dada = target(
-    readd(derep2) %>%
+    readd(derep2, cache = ignore(cache)) %>%
       set_names(
-        readd(position_map) %>%
+        readd(position_map, cache = ignore(cache)) %>%
           dplyr::bind_rows() %>%
           dplyr::mutate(region = region) %>%
           glue::glue_data("{seq_run}_{plate}_{well}_{region}")
@@ -500,10 +500,10 @@ plan <- drake_plan(
   ),
   
   dada_illumina = target(
-    readd(derep_illumina) %>%
+    readd(derep_illumina, cache = ignore(cache)) %>%
       purrr::map(read) %>%
       set_names(
-        dplyr::bind_rows(readd(illumina_id)) %>%
+        dplyr::bind_rows(readd(illumina_id, cache = ignore(cache))) %>%
           dplyr::mutate(read = read) %>%
           glue::glue_data("{seq_run}_{plate}_{well}_{read}_{direction}") %>%
           unique()
@@ -531,9 +531,9 @@ plan <- drake_plan(
   merge = target(
     dada2::mergePairs(
       dadaF = purrr::compact(dada_R1),
[…]

### 2020-03-06 writing/transect_paper.Rmd
Move supplementary info to its own file and knit in light plan

@@
 
-(ref:rDNA-s) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions
-
-(ref:rDNA) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).
-D1-3 (grey) represent the first three variable regions in LSU [@michot1984], while LSU1-4 (black) represent the conserved regions.
-Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
-
-<!--If limited by space, this could be in supp, and then just described in text. -MR-->
-```{tikz rDNA, fig.cap = "(ref:rDNA)", fig.scap = "(ref:rDNA-s)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
-
-```
-
 Both primers for the long amplicon were indexed for multiplexing (**Supplementary File 2**).
@@
 
-(ref:readcounts-s) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages
-
-(ref:readcounts) **Raw**: raw reads as delivered by sequencing center (CCS reads for PacBio);
-**Trim**: reads with primers and demultiplexing barcodes removed;
-**Filter (full)**: full length amplicons with a maximum of three expected errors;
-**LSUx**: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
-**Filter (ITS2)**: extracted ITS2 regions with a maximum of three expected errors;
-**ITS2**: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
-**short**, **ITS**, **LSU**, **long**: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.
-
-```{r bioinfo}
-options(knitr.kable.NA = '–')
-
-rdd(bioinf_table) %>%
-  set_names(gsub(".*_", "", names(.))) %>%
-  kable(
-    booktabs = TRUE,
-    caption = "(ref:readcounts-s). (ref:readcounts)",
-    caption.short = "(ref:readcounts-s)",
-    align = "rrrrrr",
-    format.args = list(big.mark = ","),
-    linesep = c(
-      rep("", 2),
-      "\\addlinespace",
-      rep("", 2),
-      "\\addlinespace",
-      rep("", 4)
-    ),
-    escape = FALSE
-    ) %>%
-  kableExtra::add_header_above(
-    header = c(
-      " " = 1,
-      rdd(bioinf_header)$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
-    )
-  ) %>%
-  kableExtra::add_header_above(
-    header = c(
-      " " = 1,
-      rle(rdd(bioinf_header)$tech) %$% set_names(lengths, values)
-    )
-  )
-```
-
 *Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
@@
 ```
-(ref:taxa-s) Heat tree summarizing taxonomic composition of soil community
-
-(ref:taxa) Color and size of nodes represent the fractional ASV richness of that taxonomic group.
-Color and thickness of branches represent the fractional read abundance of all ASVs belonging to that taxonomic group.
-Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
-Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
-ASVs which could not be further assigned are labelled "?".
-ASV richness and read abundance are displayed as the mean across sequencing runs.
-Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
-The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
-
-```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8}
-fungi_read <- readd(taxdata) %>%
-  taxa::filter_taxa(
-    taxon_names == "Fungi",
-    reassign_obs = FALSE,
-    reassign_taxa = FALSE
-    ) %>%
-  extract2("data") %>%
-  extract2("tax_read") %>%
-  set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
-  unlist()
-
-fungi_asv <- rdd(taxdata) %>%
-  taxa::filter_taxa(
-    taxon_names == "Fungi",
-    reassign_obs = FALSE,
-    reassign_taxa = FALSE
-    ) %>%
-  extract2("data") %>%
-  extract2("tax_asv") %>%
-  set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
-  unlist()
-
-ecm_counts <- rdd(taxon_reads) %>%
-  filter(Algorithm == "PHYLOTAX") %>%
-  group_by(seq_run) %>%
-  mutate(ASVs = 1/n()) %>%
-  filter(ECM %in% c("Possible ECM", "Probable ECM", "Highly Probable ECM")) %>%
-  summarize_at(c("ASVs", "reads"), sum)
-
-ecm_read <- select(ecm_counts, seq_run, reads) %>% deframe()
-ecm_asv <- select(ecm_counts, seq_run, ASVs) %>% deframe()
-
-theme <- theme_update(panel.border = element_blank())
-readd(heattree)
-theme_set(theme)
-```
 
@@
 
-```{r fungi-comm, fig.cap = "(ref:fungi-comm-s). (ref:fungi-comm)", fig.scap = "(ref:fungi-comm-s)"}
+```{r fungi-comm, fig.cap = "(ref:fungi-comm-s). (ref:fungi-comm)", fig.scap = "(ref:fungi-comm-s)", fig.width=7}
 readd(taxon_reads) %>%
@@
 
-Figure\ \@ref(fig:ppcoa)
-
-(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches
-
-(ref:ppcoa) for all fungi (**a**) and ECM fungi (**b**).
-Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
-Only samples with at least 100 reads for all three approaches are included.
-ASVs are clustered by taxonomic class for all fungi, and by family for ECM fungi.
-Abbreviations represent the approximate direction of increase for the classes and families with the highest scores: `r glue::glue_data(rdd(tech_class_pcoa_scores), "**{abbrev}**: {class}") %>% glue::glue_collapse(sep = ", ", last = ", and ")`; `r glue::glue_data(rdd(tech_ecm_fam_pcoa_scores), "**{abbrev}**: {family}") %>% glue::glue_collapse(sep = ", ", last = ", and ")`.
-
-```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
-ggpubr::ggarrange(
-  readd(tech_class_pcoa_plot),
-  readd(tech_ecm_fam_pcoa_plot),
-  common.legend = TRUE,
-  legend = "bottom",
-  labels = "auto"
-)
-```
+Fungi categorized as ECM made up `r percent(readd(ecm_asv)["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(readd(ecm_read)["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(readd(ecm_asv)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(ecm_read)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library (Figure\ \@ref(fig:ecm)).
+Although amplicon length had a significant effect on ECM community composition at the family level, the explained variation was very low (PERMANOVA with 9999 permutations, $p=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), and the majority of variation was again spatiotemporal ($p<`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$).
+Variation between sequencing technologies was not significant ($p=`r formatC(readd(tech_ecm_fam_adonis)["tech", "p.value"], format = "f", digits = 2)`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["tech", "R2"], format = "f", digits = 4)`$).
 
 ## Spatial analysis
-
-(ref:correlog-cap-s) Mantel correlograms for community dissimilarities and spatio-temporal distance
-
-(ref:correlog-cap) Community dissimilarities are Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).
-Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.
-
-```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)"}
-light_plan %>%
-  filter(startsWith(target, "correlog")) %>%
-  select(target) %>%
-  separate(
-    target,
-    into = c("step", "guild", "metric", "tech", "amplicon", "method", "tdist"),
-    sep = "_",
-    remove = FALSE
-  ) %>%
-  filter(startsWith(method, "PHYLOTAX")) %>%
-  mutate(
-    tdist = paste(tdist, "year"),
-    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
-    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
-    amplicon = factor(amplicon, levels = c("Short", "Long"))
-  ) %>%
-  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
-  mutate_at("target", purrr::map, "mantel.res") %>%
-  mutate_at("target", purrr::map, as.data.frame) %>%
-  unnest(target) %>%
-  filter(complete.cases(.)) %>%
-  mutate(Significant = `Pr(corrected)` < 0.05) %>%
-  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
-  geom_point(size = 3) +
-  geom_line(alpha = 0.3, size = 1.5) +
-  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
-  # geom_line(aes(y = exp(predict(bc_fit)))) +
-  scale_y_continuous(name = "Mantel correlation") +
-  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
-  scale_color_discrete(name = "Time lag") +
-  xlab("Distance (m)") +
-  ggnomics::facet_nested(amplicon + tech + metric ~ guild, scales = "fixed", nest_line = TRUE) +
-  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
-  theme(strip.background = element_blank())
-```
 
@@
   theme(strip.background = element_blank())
-```
-
-(ref:variofit) Parameters for variogram fits shown in Figure\ \@ref(fig:spatio). Range: range at which exponential function is at 95% of its maximum value.
-
-```{r variofit, results="asis"}
- readd(variofit_table) %>%
-  mutate_at("conf.low", replace_na, 0) %>%
-  mutate(value = ifelse(
-    is.na(estimate) | is.na(conf.high),
-    "-",
-    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
-  )) %>%
-  filter(timelag == "0" | term == "timerange") %>%
-  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
-  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
-  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
-  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
 ```
@@
 \processdelayedfloats
-
-# Supplementary Information {-}
-
-\beginsupplement
-
-(ref:phylotax-s) Phylogenetic refinement of taxonomic assignments (PHYLOTAX)
-
-(ref:phylotax) In the example, a clade includes  seven OTUs (A-F), which have been identified by two methods as belonging to taxa "Tax1" or "Tax2", or are unidentified ("unk"), as shown in the tip labels.
-No taxon is assigned at node\ 1, because one of the child branches (A) is completely unassigned.
-No taxon is assigned at node\ 2, because the assignments at C and F are inconsistent.
-Node\ 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
-Node\ 4 (and thus also node\ 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
-
-```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)"}
-readd(phylotax_fig)
-```
-
-(ref:full-length-s) Comparison of the length of denoised sequences from different length amplicons and sequencing technologies
-
-(ref:full-length) Length distribution (*a*, *c*) and empirical cumulative distribution function (ECDF; *b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.
-
-```{r full-length, fig.cap = "(ref:full-length-s). (ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6}
-ggpubr::ggarrange(
-  readd(full_dens_short),
-  readd(full_ecdf_short),
-  readd(full_dens_long),
-  readd(full_ecdf_long),
-  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
-)
-```
-
-```{r region-limits}
-regions %>%
-  select(Region = region, "Min. length" = min_length, "Max. length" = max_length) %>%
-  right_join(
-    tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4")),
-    by = "Region"
-    ) %>%
-  mutate_at("Region", str_replace_all, "_", ".") %>%
-  knitr::kable(
-    booktabs = TRUE,
-    caption = "Minimum and maximum allowed lengths for each extracted region."
-  )
-```
-
-(ref:region-asv-s) Plot of total number of ASVs recovered vs. fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`
-
-(ref:region-asv) Some regions were formed by concatenation of extracted regions, as follows:
-LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
-ITS = ITS1--5.8S--ITS2;
-32S = 5.8S--ITS2--LSU;
-long = ITS--LSU.
-For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.
-
-```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)"}
-readd(region_table) %>%
-  dplyr::filter(seq_run == "pb_500") %>%
-  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
-  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
-  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
-  geom_point() +
-  ggrepel::geom_text_repel() +
-  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
-  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
-```
-
-(ref:region-lengths-s) Length distribution for each region extracted from long amplicons
-
-(ref:region-lengths) Filtering limits are shown as dashed vertical lines.
-
-```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6}
-readd(region_length_fig)
-```
-
-(ref:ITS2-length-s) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies
-
-(ref:ITS2-length) *a*) Length distribution.  *b*) Empirical cumulative distribution function.
-
-```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6}
-ggpubr::ggarrange(
-  readd(its2_dens),
-  readd(its2_ecdf),
-  ncol = 1,
-  labels = "auto",
-  heights = c(4, 2)
-)
-```
-
-(ref:venn-table-asv-s) Correspondences between ASVs found by different sequencing strategies
-
-(ref:venn-table-asv) Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASV frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
-
-\setlength{\tabcolsep}{1pt}
-
-```{r venn-table-asv}
-readd(venn_ASV) %>%
-venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
-```
-
-(ref:venn-table-otu-s)  Correspondences between 97% OTUs found by different sequencing strategies
-
-(ref:venn-table-otu) Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTU frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*frac*).
-
-```{r venn-table-otu}
-readd(venn_OTU) %>%
-venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
-```
-
-(ref:short-its-s) Kingdom-level taxonomic composition of ITS2 ASVs $\le 140$ bp
-
-(ref:short-its) Vertical axis shows fraction of total reads in each dataset.
-
-```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)"}
-readd(big_seq_table_ITS2) %>%
-  colSums() %>%
-  enframe(name = "ITS2", value = "nreads") %>%
-  mutate(length = nchar(ITS2)) %>%
-  filter(length <= 140) %>%
-  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
-  # group_by(ITS2) %>%
-  # filter(!"Short" %in% amplicon) %>%
-  taxon_plot(kingdom, y = reads, datasets = datasets)
-```
-
-(ref:heattree-length-compare-s) Variation in read abundance and diversity for different amplicons across the taxonomic tree
-
-(ref:heattree-length-compare) Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
-In both cases red represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
-Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
-Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
-ASVs which could not be further assigned are labelled "?".
-Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
-Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.
-
-```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s). (ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%"}
-oldtheme <- theme_update(panel.border = element_blank())
-readd(heattree_length_compare)
-theme_set(oldtheme)
-```
-
-(ref:ecm-s) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments
-
-(ref:ecm) Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
-"unidentified" denotes ASVs which were assigned to kingdom Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
-Height of each bar represents the fraction of reads represented by that bar.
-Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; "All" denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.
-
-```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)"}
-  readd(taxon_reads) %>%
-  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
-  mutate_at("reference", replace_na, "All") %>%
-  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-(ref:ecm-heattree-s) Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only
-
-(ref:ecm-heattree) Red edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
-Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
-Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
-ASVs which could not be further assigned are labelled "?".
-Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
-The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
-
-```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%"}
-oldtheme <- theme_update(panel.border = element_blank())
-readd(ecm_heattree)
-theme_set(oldtheme)
-```
-
-(ref:sample-depth-s) DNA concentrations after extraction and PCR, and sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016
-
-(ref:sample-depth) DNA concentrations after extraction (leftmost column) and PCR (second and third columns), and sequencing depth (right three columns) along transects at the two sites Ang and Gan for the years 2015 and 2016.
-Sequencing reads which were filtered out during quality control are colored dark grey.
-Dotted horizontal line at 100\ reads indicates cutoff for inclusion in community analysis.
-2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.
-
-```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)"}
-reads_plot <- readd(qstats_n) %>%
-  mutate(file = str_replace(file, ".*/", "") %>%
-           str_replace("[.].*", "")) %>%
-  tidyr::extract(col = "file", into = c("seq_run", "plate", "row", "col", "read"),
-                 regex = "([[:alpha:]]{2}[-_]\\d+)_(\\d+)-([A-H])(\\d+)(_R[12])?[rf]?") %>%
-  filter(is.na(step), !is.na(seq_run), read != "_R2" | is.na(read)) %>%
-  group_by(seq_run, plate, row, col) %>%
-  summarize(raw_reads = sum(nreads)) %>%
-  ungroup() %>%
-  mutate_at("col", as.integer) %>%
-  left_join(
-    readd(proto_physeq) %>% {
-      mutate(
-        sample_data(.),
-        reads = otu_table(.) %>% rowSums()
-      )
-    } %>%
-      select(seq_run, plate, row, column, site, year, qual, sample_type, x, buffer, tech, amplicon, reads) %>%
-      mutate(row = LETTERS[row]) %>%
-      unique(),
-    by = c("seq_run", "plate", "row", "col" = "column")
-  ) %>%
-  filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x), !is.na(buffer)) %>%
-  complete(nesting(seq_run, tech, amplicon), site, x, nesting(year, buffer)) %>%
-  pivot_longer(cols = c("reads", "raw_reads"), names_to = "type", values_to = "reads") %>%
-  mutate(
-    type = factor(type, levels = c("raw_reads", "reads")),
-    buffer = factor(buffer, levels = c("Xpedition", "LifeGuard")),
-    tech = factor(tech, levels = c("PacBio", "Illumina")),
-    "Sample group" = paste(year, buffer) %>%
-      ifelse(type == "raw_reads", "Raw", .) %>%
-      fct_relevel("Raw")
-  ) %>%
-  arrange(`Sample group`) %>%
-  ggplot(aes(x = x, y = reads, group = `Sample group`, fill = `Sample group`)) +
-  geom_col(position = "identity") +
-  ggnomics::facet_nested(
-    cols = vars(tech, amplicon),
-    rows = vars(year, buffer, site),
-    nest_line = TRUE,
-    resect = unit(1, "mm")
-  ) +
-  scale_fill_manual(
-    values = c(RColorBrewer::brewer.pal(3, "Set2") %>% set_names(c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition")), Raw = "gray20"),
-    breaks = c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition", "Raw"),
-    labels = c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition", "Filtered")) +
-  scale_y_log10(breaks = c(10, 1e3, 1e5),
-                labels = c("10", "1k", "100k"), position = "right") +
-  xlab("x (m)") +
-  theme(
-    legend.position = "bottom",
-    strip.background = element_blank(),
-    panel.spacing = unit(1, "mm"),
-    plot.margin = margin(l = 0.5, unit = "mm"),
-    axis.text.x = element_text(angle = 90, vjust = 0.5)
-  ) +
-  geom_hline(yintercept = 100, linetype = 2, alpha = 0.5)
-
-conc_plot <- readd(proto_physeq) %>% 
-        sample_data(.) %>%
-      select(plate, row, column, site, year, sample_type, x, buffer, amplicon, dna_conc, pcr_conc) %>%
-  arrange(plate, row, column, amplicon) %>% 
-  unique() %>% 
-  mutate_at("dna_conc", as.double) %>%
-  pivot_longer(cols = c("dna_conc", "pcr_conc"), names_to = "step", values_to = "conc", names_pattern = "([[:alpha:]]+)_conc") %>%
-  filter(!is.na(year)) %>%
-  mutate(
-    amplicon = ifelse(step == "dna", "", amplicon),
-    buffer = factor(buffer, levels = c("Xpedition", "LifeGuard")),
-    step = factor(step, levels = c("dna", "pcr"), labels = c("DNA", "PCR product")),
-    conc = 1000 * conc,
-    "Sample group" = paste(year, buffer)
-  ) %>%
-  unique() %>%
-  complete(nesting(amplicon, step), site, x, nesting(year, buffer, `Sample group`)) %>%
-  arrange(`Sample group`) %>%
-  ggplot(aes(x = x, y = conc, group = `Sample group`, fill = `Sample group`)) +
-  geom_col() +
-  ggnomics::facet_nested(
-    cols = vars(step, amplicon),
-    rows = vars(year, buffer, site),
-    nest_line = TRUE,
-    resect = unit(1, "mm"),
-    switch = "y"
-  ) +
-  scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Set2") %>% set_names(c("2015 Xpedition", "2016 LifeGuard", "2016 Xpedition")), guide = "none") +
-  scale_y_log10(
-    breaks = c(10, 1e3, 1e5),
-    labels = c("0.1", "1", "100"),
-    name = "DNA concentration (ng/µL)"
-  ) +
-  xlab("x (m)") +
-  theme(
-    legend.position = "bottom",
-    strip.background = element_blank(),
-    strip.placement = "outside",
-    panel.spacing = unit(1, "mm"),
-    plot.margin = margin(r = 0.5, unit = "mm"),
-    axis.text.x = element_text(angle = 90, vjust = 0.5)
-  )
-
-ggpubr::ggarrange(conc_plot, reads_plot, ncol = 2, common.legend = TRUE) +
-  theme(panel.spacing = unit(0, "mm"))
-```
-
-```{r plate-loading, include=FALSE}
-readd(proto_physeq) %>% {
-  mutate(
-    sample_data(.),
-    reads = otu_table(.) %>% rowSums()
-  )
-} %>%
-  filter(tech != "Ion Torrent", amplicon == "Short", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
-  select(plate, well, tech, reads, year, site, buffer) %>%
-  pivot_wider(names_from = "tech", values_from = "reads") %>%
-  ggplot(aes(x = PacBio, y = Illumina, color = year)) +
-  geom_point()
-```
-
-(ref:buffer-pcoa-s) Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers
-
-(ref:buffer-pcoa) For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.
-
-```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)"}
-plot(readd(buffer_asv_pcoa), type = "none")
-points(
-  readd(buffer_asv_pcoa),
-  select = with(
-    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
-    buffer == "Xpedition" & year == "2015" 
-  ),
-  col = "green",
-  pch = 1
-)
-points(
-  readd(buffer_asv_pcoa),
-  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
-                buffer == "Xpedition" & year == "2016"
-  ),
-  col = "blue",
-  pch = 2
-)
-points(
-  readd(buffer_asv_pcoa),
-  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
-                buffer == "LifeGuard" & year == "2016"
-  ),
-  col = "red",
-  pch = 3
-)
-legend(
-  "bottomright",
-  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
-  col = c("green", "blue", "red"),
-  pch = 1:3
-)
-# legend(
-#   "topright",
-#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
-#     unique() %>%
-#     sort(),
-#   lty = 1:3
-# )
-# vegan::ordiellipse(
-#   readd(buffer_asv_pcoa),
-#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
-#   lty = 1:3
-# )
-```
-
-(ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
-
-```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)"}
-oldtheme <- theme_update(panel.border = element_blank())
-readd(buffer_compare_long)
-theme_set(oldtheme)
-```
-
-(ref:buffer-compare-short) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for short amplicon libraries.
-
-```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
-readd(buffer_compare_short)
-theme_set(oldtheme)
-
-```\processdelayedfloats

### 2020-03-06 writing/transect_supplement.Rmd
Move supplementary info to its own file and knit in light plan

---
title: Supplementary Information
always_allow_html: true
output:
  bookdown::pdf_document2:
    fig_crop: yes
    keep_tex: yes
    toc: false
    latex_engine: lualatex
    includes:
      in_header: preamble_supplement.tex
    citation_package: biblatex
    suppress-bibliography: true
    fontsize: 12pt
    geometry: left=2.5cm, right=2.5cm, top=3cm, bottom=3cm
---

```{r setup, include=FALSE}
library("knitr")
library("ggplot2")
library("magrittr")
library("tidyverse")
library("drake")
library("here")
library("phyloseq")
library("kableExtra")
tryCatch(
  tinytex::tinytex_root(),
  error = function(e) tinytex::install_tinytex()
  )

tinytex::tlmgr_install("pdfcrop")

knitr::opts_chunk$set(
  dev = "pdf",
  # dev.args = list(pointsize=32),
  echo = FALSE,
  # results = 'asis',
  # fig.show = 'asis',
  fig.width = 6,
  fig.height = 4.5,
  message = FALSE,
  warning = FALSE,
  external = TRUE,
  out.width = "70%",
  fig.align = "center"
)

options(knitr.table.format = "latex")

light_plan <- readRDS(here("data/plan/drake_light.rds"))
theme_set(theme_bw())

datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccicc")
regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
```

\listoffigures
\listoftables

(ref:buffer-pcoa-s) Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers

(ref:buffer-pcoa) For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.

```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)"}
plot(readd(buffer_asv_pcoa), type = "none")
points(
  readd(buffer_asv_pcoa),
  select = with(
    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
    buffer == "Xpedition" & year == "2015" 
  ),
  col = "green",
  pch = 1
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "Xpedition" & year == "2016"
  ),
  col = "blue",
  pch = 2
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "LifeGuard" & year == "2016"
  ),
  col = "red",
  pch = 3
)
legend(
  "bottomright",
  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
  col = c("green", "blue", "red"),
  pch = 1:3
)
# legend(
#   "topright",
#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
#     unique() %>%
#     sort(),
#   lty = 1:3
# )
# vegan::ordiellipse(
#   readd(buffer_asv_pcoa),
#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
#   lty = 1:3
# )
```

(ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.

```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)"}
oldtheme <- theme_update(panel.border = element_blank())
readd(buffer_compare_long)
theme_set(oldtheme)
```

(ref:buffer-compare-short) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for short amplicon libraries.

```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
readd(buffer_compare_short)
theme_set(oldtheme)

```

(ref:rDNA-s) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions

(ref:rDNA) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).  D1-3 (grey) represent the first three variable regions in LSU, while LSU1-4 (black) represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.

```{tikz rDNA, fig.cap = "(ref:rDNA)", fig.scap = "(ref:rDNA-s)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}

```

(ref:region-asv-s) Plot of total number of ASVs recovered vs. fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`

(ref:region-asv) Some regions were formed by concatenation of extracted regions, as follows:
LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
ITS = ITS1--5.8S--ITS2;
32S = 5.8S--ITS2--LSU;
long = ITS--LSU.
For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.

```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)"}
readd(region_table) %>%
  dplyr::filter(seq_run == "pb_500") %>%
  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
```

(ref:phylotax-s) Phylogenetic refinement of taxonomic assignments (PHYLOTAX)

(ref:phylotax) In the example, a clade includes  seven OTUs (A-F), which have been identified by two methods as belonging to taxa "Tax1" or "Tax2", or are unidentified ("unk"), as shown in the tip labels.
No taxon is assigned at node\ 1, because one of the child branches (A) is completely unassigned.
No taxon is assigned at node\ 2, because the assignments at C and F are inconsistent.
Node\ 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
Node\ 4 (and thus also node\ 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.

```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)"}
readd(phylotax_fig)
```

(ref:sample-depth-s) DNA concentrations after extraction and PCR, and sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016

(ref:sample-depth) DNA concentrations after extraction (leftmost column) and PCR (second and third columns), and sequencing depth (right three columns) along transects at the two sites Ang and Gan for the years 2015 and 2016.
Sequencing reads which were filtered out during quality control are colored dark grey.
Dotted horizontal line at 100\ reads indicates cutoff for inclusion in community analysis.
2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.

```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)"}
ggpubr::ggarrange(readd(conc_plot), readd(reads_plot), ncol = 2, common.legend = TRUE) +
  theme(panel.spacing = unit(0, "mm"))
```

(ref:full-length-s) Comparison of the length of denoised sequences from different length amplicons and sequencing technologies

(ref:full-length) Length distribution (*a*, *c*) and empirical cumulative distribution function (ECDF; *b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.

```{r full-length, fig.cap = "(ref:full-length-s). (ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6}
ggpubr::ggarrange(
  readd(full_dens_short),
  readd(full_ecdf_short),
  readd(full_dens_long),
  readd(full_ecdf_long),
  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
)
```

(ref:region-lengths-s) Length distribution for each region extracted from long amplicons

(ref:region-lengths) Filtering limits are shown as dashed vertical lines.

```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6}
readd(region_length_fig)
```

(ref:taxa-s) Heat tree summarizing taxonomic composition of soil community

(ref:taxa) Color and size of nodes represent the fractional ASV richness of that taxonomic group.
Color and thickness of branches represent the fractional read abundance of all ASVs belonging to that taxonomic group.
Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
ASV richness and read abundance are displayed as the mean across sequencing runs.
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.

```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8}
theme <- theme_update(panel.border = element_blank())
readd(heattree)
theme_set(theme)
```

(ref:ecm-s) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments

(ref:ecm) Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
"unidentified" denotes ASVs which were assigned to kingdom Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
Height of each bar represents the fraction of reads represented by that bar.
Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; "All" denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7}
  readd(taxon_reads) %>%
  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
  mutate_at("reference", replace_na, "All") %>%
  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches

(ref:ppcoa) for all fungi (**a**) and ECM fungi (**b**).
Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
Only samples with at least 100 reads for all three approaches are included.
ASVs are clustered by taxonomic class for all fungi, and by family for ECM fungi.
Abbreviations represent the approximate direction of increase for the classes and families with the highest scores: `r glue::glue_data(readd(tech_class_pcoa_scores), "**{abbrev}**: {class}") %>% glue::glue_collapse(sep = ", ", last = ", and ")`; `r glue::glue_data(readd(tech_ecm_fam_pcoa_scores), "**{abbrev}**: {family}") %>% glue::glue_collapse(sep = ", ", last = ", and ")`.

```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
ggpubr::ggarrange(
  readd(tech_class_pcoa_plot),
  readd(tech_ecm_fam_pcoa_plot),
  common.legend = TRUE,
  legend = "bottom",
  labels = "auto"
)
```

(ref:correlog-cap-s) Mantel correlograms for community dissimilarities and spatio-temporal distance

(ref:correlog-cap) Community dissimilarities are Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).  Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.

```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)"}
light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "method", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  filter(startsWith(method, "PHYLOTAX")) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
    amplicon = factor(amplicon, levels = c("Short", "Long"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  ggnomics::facet_nested(amplicon + tech + metric ~ guild, scales = "fixed", nest_line = TRUE) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
  theme(strip.background = element_blank())
```

(ref:ITS2-length-s) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies

(ref:ITS2-length) *a*) Length distribution.  *b*) Empirical cumulative distribution function.

```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6}
ggpubr::ggarrange(
  readd(its2_dens),
  readd(its2_ecdf),
  ncol = 1,
  labels = "auto",
  heights = c(4, 2)
)
```

(ref:heattree-length-compare-s) Variation in read abundance and diversity for different amplicons across the taxonomic tree

(ref:heattree-length-compare) Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
In both cases red represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.

```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s). (ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%"}
oldtheme <- theme_update(panel.border = element_blank())
readd(heattree_length_compare)
theme_set(oldtheme)
```

(ref:short-its-s) Kingdom-level taxonomic composition of ITS2 ASVs $\le 140$ bp

(ref:short-its) Vertical axis shows fraction of total reads $\le 140$ bp in each dataset.

```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7}
readd(big_seq_table_ITS2) %>%
  colSums() %>%
  enframe(name = "ITS2", value = "nreads") %>%
  mutate(length = nchar(ITS2)) %>%
  filter(length <= 140) %>%
  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
  group_by(tech, amplicon, Algorithm, reference) %>%
  mutate(reads = reads/sum(reads)) %>%
  ungroup() %>%
  # group_by(ITS2) %>%
  # filter(!"Short" %in% amplicon) %>%
  taxon_plot(kingdom, y = reads, datasets = datasets)
```

(ref:ecm-heattree-s) Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only

(ref:ecm-heattree) Red edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.

```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%"}
oldtheme <- theme_update(panel.border = element_blank())
readd(ecm_heattree)
theme_set(oldtheme)
```

```{r region-limits}
regions %>%
  select(Region = region, "Min. length" = min_length, "Max. length" = max_length) %>%
  right_join(
    tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4")),
    by = "Region"
    ) %>%
  mutate_at("Region", str_replace_all, "_", ".") %>%
  knitr::kable(
    booktabs = TRUE,
    caption = "Minimum and maximum allowed lengths for each extracted region."
  )
```

(ref:readcounts-s) Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages

(ref:readcounts) **Raw**: raw reads as delivered by sequencing center (CCS reads for PacBio);
**Trim**: reads with primers and demultiplexing barcodes removed;
**Filter (full)**: full length amplicons with a maximum of three expected errors;
**LSUx**: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
**Filter (ITS2)**: extracted ITS2 regions with a maximum of three expected errors;
**ITS2**: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
**short**, **ITS**, **LSU**, **long**: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.

```{r bioinfo}
options(knitr.kable.NA = '–')

readd(bioinf_table) %>%
  set_names(gsub(".*_", "", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:readcounts-s). (ref:readcounts)",
    caption.short = "(ref:readcounts-s)",
    align = "rrrrrr",
    format.args = list(big.mark = ","),
    linesep = c(
      rep("", 2),
      "\\addlinespace",
      rep("", 2),
      "\\addlinespace",
      rep("", 4)
    ),
    escape = FALSE
    ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      readd(bioinf_header)$amplicon[c(TRUE, FALSE)] %>% set_names(rep(2, length(.)), .)
    )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      rle(readd(bioinf_header)$tech) %$% set_names(lengths, values)
    )
  )
```

\setlength{\tabcolsep}{1ex}

(ref:venn-table-asv-s) Correspondences between ASVs found by different sequencing strategies

(ref:venn-table-asv) Each row shows the number of ASVs (*ASVs*) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (*ASVs frac*), the number of reads represented by those ASVs in each dataset (*reads*) and the fraction of total reads for the dataset (*reads frac*).

```{r venn-table-asv}
readd(venn_ASV) %>%
venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
```

(ref:venn-table-otu-s)  Correspondences between 97% OTUs found by different sequencing strategies

(ref:venn-table-otu) Each row shows the number of OTUs (*OTUs*) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (*OTUs frac*), the number of reads represented by those OTUs in each dataset (*reads*) and the fraction of total reads for the dataset (*reads frac*).

```{r venn-table-otu}
readd(venn_OTU) %>%
venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
```

(ref:variofit) Parameters for exponential distance-decay fits. Range: range at which exponential function is at 95% of its maximum value.

```{r variofit, results="asis"}
 readd(variofit_table) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

### 2020-03-06 scripts/drake_plan_light.R

@@ -2378,6 +2378,41 @@ plan2 <- drake_plan(
     mutate_at("seq_run", factor) %$%
     kruskal.test(length, g = seq_run),
   
+  fungi_read =
+    taxdata %>%
+    taxa::filter_taxa(
+      taxon_names == "Fungi",
+      reassign_obs = FALSE,
+      reassign_taxa = FALSE
+    ) %>%
+    extract2("data") %>%
+    extract2("tax_read") %>%
+    set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
+    unlist(),
+  
+  fungi_asv =
+    taxdata %>%
+    taxa::filter_taxa(
+      taxon_names == "Fungi",
+      reassign_obs = FALSE,
+      reassign_taxa = FALSE
+    ) %>%
+    extract2("data") %>%
+    extract2("tax_asv") %>%
+    set_names(names(.) %>% str_extract("(PacBio|Illumina|Ion.Torrent)_(Long|Short)")) %>%
+    unlist(),
+  
+  ecm_counts =
+    taxon_reads %>%
+    filter(Algorithm == "PHYLOTAX") %>%
+    group_by(seq_run) %>%
+    mutate(ASVs = 1/n()) %>%
+    filter(ECM %in% c("Possible ECM", "Probable ECM", "Highly Probable ECM")) %>%
+    summarize_at(c("ASVs", "reads"), sum),
+  
+  ecm_read = select(ecm_counts, seq_run, reads) %>% deframe(),
+  ecm_asv = select(ecm_counts, seq_run, ASVs) %>% deframe(),
+  
   short_otu_table = {
     big_fasta_short
     read_tsv(
@@ -2409,6 +2444,108 @@ plan2 <- drake_plan(
         by = "seq"
       )
   },
+  
+  # Plot of reads along transect
+  reads_plot = parsed_qstat %>%
+    filter(stat == "n") %>%
+    tidyr::extract(col = "well", into = c("row", "col"),
+                   regex = "([A-H])(\\d+)") %>%
+    filter(is.na(step), !is.na(seq_run), read != "_R2" | is.na(read)) %>%
+    group_by(seq_run, plate, row, col) %>%
+    summarize(raw_reads = sum(nreads)) %>%
+    ungroup() %>%
+    mutate_at("col", as.integer) %>%
+    left_join(
+      proto_physeq %>% {
+        mutate(
[…]

### 2020-03-06 writing/transect_paper.Rmd
Switch to biblatex for bibliography

@@
   - "R.bib"
-csl: molecular-ecology.csl
+biblio-style: apa
 always_allow_html: true
@@
     keep_tex: yes
+    toc: false
     latex_engine: lualatex
@@
       in_header: preamble.tex
+    citation_package: biblatex
+    suppress-bibliography: true
+    fontsize: 12pt
+    geometry: left=2.5cm, right=2.5cm, top=3cm, bottom=3cm
   bookdown::word_document2: 
@@
 }
-write_bib(c("base", "phyloseq", "vegan"), file = "R.bib", tweak = TRUE)
+write_bib(c("base", "phyloseq", "vegan", "metacoder"), file = "R.bib", tweak = TRUE)
 
@@
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
-
-source(here(file.path("scripts", "output_functions.R")))
 ```
@@
 
-# References {-}
+\singlespacing
 
-<div id="refs"></div>
+\printbibliography
+
+\doublespacing
 

### 2020-03-06 writing/transect_paper.Rmd
miscellaneous improvements before submitting

@@
 )
-options(
-  tikzDefaultEngine = "luatex",
-  tikzXelatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
-                          "\\usepackage[version=4]{mhchem}\n",
-                          "\\usepackage{siunitx}\n",
-                          "\\DeclareSIUnit\\year{a}\n",
-                          "\\usepackage{fontspec}"),
-  tikzMetricsDictionary = here("writing/transect_paper-tikzDictionary")
-)
 
 base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
-if (opts_chunk$get("dev") == "tikz") {
-  # styles$Expression <- styles$LaTeX.Expression
-  # styles$Name <- styles$LaTeX.Name
-  # styles$Short.Name <- styles$LaTeX.Short.Name
-  # styles$Unit <- paste0("\\si{", styles$LaTeX.Unit, "}")
-}
 write_bib(c("base", "phyloseq", "vegan", "metacoder"), file = "R.bib", tweak = TRUE)
@@
 
-# Abstract
+# Abstract {-}
 
 Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in taxonomic and phylogenetic placement of fungi from environmental sequences.
-We investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
+To address such challenges we investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
 Resulting community structure and diversity depended more on statistical method than sequencing technology.
@@
 
-Ectomycorrhizas (ECM) form a symbiosis between fungi and plants.
+Fungi are key drivers of nutrient cycling in terrestrial ecosystems.
+One important guild of fungi form ectomycorrhizas (ECM), a symbiosis between fungi and plants in which fungal hyphae enclose the plant's fine root tips.
 The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant [@Smith2010].
@@
 
-Although many ECM fungi form many well-known mushrooms (e.g., *Amanita*, *Cantharellus*, *Boletus*), some instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
+Although ECM fungi form many well-known mushrooms (e.g., *Amanita*, *Cantharellus*, *Boletus*), some instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
 Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures [@horton2001].
@@
 After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
-Samples preserved using LifeGuard were first centrifuged at 10000×g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
+Samples preserved using LifeGuard were first centrifuged at 10000\ g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
 DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
@@
 Reads were searched in both directions, and reads where the primers were found in the reverse direction were reverse complemented before further analysis.
-For Illumina sequences, primers were also searched for and removed on the 3' ends of the reads, in case of readthrough with short amplicons.
-Read pairs were retained when PCR primers were detected at the 5' ends of both the forward and reverse read.
+For Illumina sequences, read pairs were only retained when PCR primers were detected at the 5' ends of both the forward and reverse read.
+Primers were also searched for and removed on the 3' ends of the reads, in case of readthrough with short amplicons.
 Read pairs where the primers were found in reverse orientation were kept in separate files, but were retained in their original orientation until after denoising.
@@
 For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
-Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available on github (https://github.com/brendanf/LSUx).
+Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`.
 
 Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors.
-Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")` [R package `dada2`, @callahan2016; @callahan2019].
+Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
 The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
@@
 A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
-The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available on github (https://github.com/brendanf/tzara).
+The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R.
 
@@
 Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
-FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).
 
@@
 
-ML trees were produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
-The LSU tree was not constrained topologically, but the long amplicon tree was constrained by the result for the LSU tree.
-*Trees and alignments are available in Treebase. (not yet)*
-
+An ML treewas produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
 The tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
@@
 
-The fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina, PacBio Short, PacBio Long) were compared by permutational MANOVA.
+The fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina, PacBio Short, PacBio Long) were compared by PERMANOVA.
 In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
-Permutational MANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs.
-short); and sequencing technology (Illumina MiSeq vs.
-PacBio RS II).
 Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
+PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs. short); and sequencing technology (Illumina MiSeq vs. PacBio RS II).
 The marginal significance of each term for explaining variation in the Bray-Curtis community dissimilarity matrix was performed using the `adonis2` function in the R package `vegan` [@R-vegan], with 9999 permutations.
-
 Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
 Spatiotemporal effects were partialled out in order to visualize effects due to sequencing technology and amplicon length.
+
+A similar analysis was also applied to only fungi classified as ECM, clustered at the family level.
 
@@
 $$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
-where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity from replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill", and $a_t$ the equivalent temporal range.
+where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity from replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill"; and $a_t$, the equivalent temporal range.
 The 95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package in R.
@@
 The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Figure\ \@ref(fig:venn)).
-ASVs shared between all datasets represented `r paste(range(as.numeric(unlist(readd(venn_ASV)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
-These fractions increased to `r paste(range(as.numeric(unlist(readd(venn_OTU)["1111", c("frac_pb_500", "frac_pb_483", "frac_SH-2257", "frac_is_057")]))*100), collapse = "--")`% when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
-In particular, `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("frac_pb_483", "frac_SH-2257", "frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
-In contrast, `r percent(readd(venn_ASV)["1000", "frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])` after OTU clustering.
+ASVs shared between all datasets represented `r paste(range(as.numeric(unlist(readd(venn_ASV)["1111", c("reads_frac_pb_500", "reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")]))*100), collapse = "--")`% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
+These fractions increased to `r paste(range(as.numeric(unlist(readd(venn_OTU)["1111", c("reads_frac_pb_500", "reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")]))*100), collapse = "--")`% when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
+In particular, `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
+In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after OTU clustering.
 Complete tabulations of the number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively.
@@
 
-Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "frac_pb_500"])`.
+Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])`.
 Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
 Taxonomic assignments revealed that the majority of these non-shared sequences fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Figure\ \@ref(fig:short-its)).
-Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figure \@ref(fig:pcoa) and \@ref(fig:heattree-length-compare)).
+Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
 Additionally, several smaller groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
@@
 For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
-For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
+For instance, `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
 This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure\ \@ref(fig:ecm)).
@@
 Weighted UNIFRAC did not reliably detect spatial structure within this relatively ecologically homogeneous community.
-Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m) (Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship, and the functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m.
+Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m; Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship.
+The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m, indicating little evidence of spatial structure.
 This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure\ \@ref(fig:taxa)), which are both well represented in the majority of samples.
@@
 Sequence alignment and phylogenetic tree are available at TreeBase under project S25885.
-R packages `lsux`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/lsux, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
+R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
+FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
 Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.

### 2020-03-10 writing/transect_supplement.Rmd
add descriptions for supplementary files

@@
 ---
-title: Supplementary Information
+title: 
+  - "Supplementary Information for:"
+  - "Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities"
+author: 
+  - Brendan Furneaux^1^,
+  - Mohammad Bahram^2,3^,
+  - Anna Rosling^4^,
+  - Nourou S. Yorou^5^,
+  - Martin Ryberg^1^
+  - ^1^Program in Systematic Biology,
+    Department of Organismal Biology,
+    Uppsala University,
+  - Uppsala, Sweden
+  - ^2^Department of Ecology,
+    Swedish University of Agricultural Sciences,
+    Uppsala, Sweden
+  - ^3^Institute of Ecology and Earth Sciences,
+    University of Tartu,
+    Tartu, Estonia
+  - ^4^Program in Evolutionary Biology,
+    Department of Ecology and Genetics,
+    Uppsala University,
+  - Uppsala, Sweden
+  - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
+    LEB,
+  - University of Parakou,
+    Parakou, Benin
 always_allow_html: true
@@
 
-\listoffigures
-\listoftables
+**Supplementary file 1**:
 
@@
 
-```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)"}
+```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
 plot(readd(buffer_asv_pcoa), type = "none")

### 2020-03-10 writing/transect_supplement.Rmd
make a tree figure, add descriptions of supplementary files to the supplementary figures/tables file

@@
 
-**Supplementary file 1**:
+**Supplementary file 1**: Mapping between samples and indexed primers for short amplicons.
+
+**Supplementary file 2**: Mapping between samples and indexed primers for long amplicons.
+
+**Supplementary file 3**: ML tree of all long metabarcoding ASVs (5.8S--ITS2--LSU) assigned to kingdom Fungi.
+Thick branches represent rapid bootstrap values >90%.
+Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, and Unite, Warcup, and RDP reference databases. 
+Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
+Taxon names in red were reconstructed as polyphyletic.
+In some cases, this may be due to a poorly identified sequence "splitting" the group (e.g., the single *Aspergillus* sequence nested in *Monascus*, Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).
 

### 2020-03-10 scripts/drake_plan_light.R

@@ -2479,8 +2479,8 @@ plan2 <- drake_plan(
     left_join(
       proto_physeq %>% {
         mutate(
-          sample_data(.),
-          reads = otu_table(.) %>% rowSums()
+          phyloseq::sample_data(.),
+          reads = phyloseq::otu_table(.) %>% rowSums()
         )
       } %>%
         select(seq_run, plate, row, column, site, year, qual, sample_type, x, buffer, tech, amplicon, reads) %>%
@@ -2526,7 +2526,8 @@ plan2 <- drake_plan(
   
   conc_plot =
     proto_physeq %>% 
-    sample_data(.) %>%
+    phyloseq::sample_data(.) %>%
+    as("data.frame") %>%
     select(plate, row, column, site, year, sample_type, x, buffer, amplicon, dna_conc, pcr_conc) %>%
     arrange(plate, row, column, amplicon) %>% 
     unique() %>% 
@@ -2645,7 +2646,6 @@ plan2 <- drake_plan(
           "transect_supplement.Rmd",
           clean = FALSE
         )
-        file_out(!!file.path("writing", "transect_supplement.aux"))
         file_out(!!file.path("writing", "transect_supplement.tex"))
         readLines("transect_supplement.tex") %>%
           magrittr::extract(!grepl("\\\\printbibliography", .)) %>%
@@ -2663,7 +2663,6 @@ plan2 <- drake_plan(
         "transect_paper.Rmd",
         clean = FALSE
       )
-      file_out(!!file.path("writing", "transect_paper.aux"))
       file_out(!!file.path("writing", "transect_paper.tex"))
       
       readLines("transect_paper.tex") %>%
@@ -2676,8 +2675,7 @@ plan2 <- drake_plan(
     "writing",
     {
       file_in(!!file.path("writing", "transect_supplement.tex"))
-      file_in(!!file.path("writing", "transect_supplement.aux"))
-      file_in(!!file.path("writing", "transect_paper.aux"))
+      file_in(!!file.path("writing", "transect_paper.tex"))
       system2(
         command = "lualatex",
         args = c("--halt-on-error", "transect_supplement.tex")
@@ -2690,8 +2688,7 @@ plan2 <- drake_plan(
     "writing", 
     {
       file_in(!!file.path("writing", "transect_paper.tex"))
-      file_in(!!file.path("writing", "transect_paper.aux"))
[…]

### 2020-03-16 writing/transect_paper.Rmd
few more changes before submission

@@
 author: 
-  - Brendan Furneaux^1^,
+  - Brendan Furneaux^1,\*^,
   - Mohammad Bahram^2,3^,
@@
     Parakou, Benin
+  - ^\*^Corresponding author, `brendan.furneaux@ebc.uu.se`
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 write_bib(c("base", "phyloseq", "vegan", "metacoder"), file = "R.bib", tweak = TRUE)
+
+if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))
 
@@
 
-The gITS7 primers for the short amplicon were indexed for multiplexing (**Supplementary File 1**).
+The gITS7 primers for the short amplicon were indexed for multiplexing (Supplementary File 1).
 Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
@@
 
-Both primers for the long amplicon were indexed for multiplexing (**Supplementary File 2**).
+Both primers for the long amplicon were indexed for multiplexing (Supplementary File 2).
 PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
@@
 Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
-In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
+In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c as described in [@tedersoodata2017].
 Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
@@
 
-An ML treewas produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
+An ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
 The tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
@@
 Taxonomic composition of the sequenced soil fungal community at the class level is is summarized in Figure\ \@ref(fig:fungi-comm) and as a heat tree [@foster2017] in Figure\ \@ref(fig:taxa).
+The ML tree for fungal ASVs, along with taxonomic assignments, is shown in Supplementary File 3.
 According to the PHYLOTAX assignments, Fungi represented `r percent(readd(fungi_asv)["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(readd(fungi_read)["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(readd(fungi_asv)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(fungi_read)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
@@
 
-Sequence data, including raw reads and ASV sequences, are deposited at NCBI under BioProject XXXXXXX.
-Sequence alignment and phylogenetic tree are available at TreeBase under project S25885.
-R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
-FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
-Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.
+ - Sequence data, including raw reads and ASV sequences, will be deposited at the European Nucleotide Archive (ENA) prior to final publication.
+ - Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication [@furneauxdata2020].
+ - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
+These packages are currently being prepared for submission to Bioconductor.
+If they are not accepted at Bioconductor prior to final publication, they will be archived at Dryad.
+ - FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
+The versions used in this paper will be archived at Dryad prior to publication.
+ - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.
 

### 2020-03-16 scripts/output_functions.R

@@ -227,7 +227,7 @@ compR2 <- function(x, y, type) {
   )
   formatC(
     filter(
-      readd(comparisons, cache = cache),
+      readd(comparisons),
       x == x_var,
       xamp == x_amplicon,
       y == y_var,

### 2020-03-16 writing/transect_paper.Rmd
Submitted version

@@
   - Martin Ryberg^1^
-  - ^1^Program in Systematic Biology,
+date: |
+      | ^1^Program in Systematic Biology,
         Department of Organismal Biology,
         Uppsala University,
-  - Uppsala, Sweden
-  - ^2^Department of Ecology,
+        Uppsala, Sweden
+      | ^2^Department of Ecology,
         Swedish University of Agricultural Sciences,
         Uppsala, Sweden
-  - ^3^Institute of Ecology and Earth Sciences,
+      | ^3^Institute of Ecology and Earth Sciences,
         University of Tartu,
         Tartu, Estonia
-  - ^4^Program in Evolutionary Biology,
+      | ^4^Program in Evolutionary Biology,
         Department of Ecology and Genetics,
         Uppsala University,
-  - Uppsala, Sweden
-  - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
+        Uppsala, Sweden
+      | ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
         LEB,
-  - University of Parakou,
+        University of Parakou,
         Parakou, Benin
-  - ^\*^Corresponding author, `brendan.furneaux@ebc.uu.se`
-date: "`r format(Sys.time(), '%B %d, %Y')`"
+      | ^\*^Corresponding author, `brendan.furneaux@ebc.uu.se`
 bibliography:
@@
   bookdown::pdf_document2:
-    fig_crop: yes
-    keep_tex: yes
+    template: bare_template.tex
+    fig_crop: true
+    keep_tex: true
     toc: false
@@
     suppress-bibliography: true
-    fontsize: 12pt
-    geometry: left=2.5cm, right=2.5cm, top=3cm, bottom=3cm
   bookdown::word_document2: 
     keep_md: yes
+fontsize: 12pt
+geometry: ["left=2.5cm", "right=2.5cm", "top=3cm", "bottom=3cm"]
+graphics: yes
 ---
@@
 
-# Data Availability Statement {-}
+# Data Accessibility {-}
 
@@
 
-# Author contributions {-}
+# Author Contributions {-}
 

### 2020-03-19 writing/transect_paper.Rmd
detect missing tinytex

@@
 tryCatch(
-  tinytex::tinytex_root(),
+  stopifnot(dir.exists(tinytex::tinytex_root())),
   error = function(e) tinytex::install_tinytex()

### 2020-03-19 writing/transect_paper.Rmd
fix datasets format, correct soil sample size

@@
 
-datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccicc")
+datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccccccccicc")
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 50\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
+Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
 <!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->

### 2020-09-16 writing/transect_paper.Rmd
address most reviewer comments in intro

@@
 
+<!-- R1: This study seeks to determine how both sequencing target length and technology affect the spatial and temporal turnover of soil fungal communities in an African forest.
+The authors take an ambitious approach to these research questions by comparing multiple short-read technologies (Illumina, PacBio, and Ion Torrent) to a related long-read technology (PacBio), allowing them to make many valuable comparisons across datasets.
+Along with field sample replication, both spatially and temporally, the authors carefully consider a number of different bioinformatic steps for each sample, e.g.
+among taxonomic databases as well as whether to cluster ASV data or not.
+Importantly, after discovering that long read data has inherent properties that make it poorly suited for the popular DADA2 bioinformatic pipeline, the authors generate a series of new R-based tools to analyze long read data in ways that greater increase data retention.
+Further, they use a clever set of covariance models to improve the taxonomic assignments of ASVs that are not assigned at the bulk taxonomy step.
+This additional step is important in increasing the amount of sequence data that can be annotated to putative functional guilds, which in this case focuses on ECM fungi.
+
+Overall, I found the study to be a really excellent demonstration of how different sequencing technologies and target lengths can be compared for fungi.
+The overall conclusion that the ecological outcome is largely similar between the different datasets is heartening.
+I am quite sure that longer reads will continue to increase in popularity as costs decline and this study represents an important bridge in that transition – both for its careful benchmarking but also for its presentation of tools that allow for better incorporation of long read sequence data into existing bioinformatic pipelines.
+I hope that my comments below help the authors make some relatively minor improvements, particular in adding more biological interpretation in a few places to help more deeply understand the patterns they have identified.
+-->
+
+<!-- R2: The authors have collected an interesting data set but I feel strongly that the work requires significant improvement to the Introduction (particularly the justification) and Discussion (particularly some explanations, even speculative) sections, as detailed below in my specific comments.
+
+Regarding the justification and conclusions, a quick scan of the literature revealed a study that used long(er) reads (covering SSU, ITS, LSU; Heeger et al.
+2019 in this journal) and also separated these prior to clustering and assignment.
+I therefore see the contribution as stated here (line 614) as overstated except that denoising was applied prior to clustering.
+Phylogenetic reconstruction was also used here, although the conclusions are not justified (see my comment below).
+I see an important contribution of this work as developing a new method to separate fragments; however, although it should be clear (i.e.
+examined in the Introduction) why existing approaches like ITSx or others are not suitable.
+
+"Long-read" has become a vague descriptor and should be clearly defined early on for this manuscript (in the abstract).
+How many bp or markers?
+I have seen the term used in metabarcoding to mean anything from a few bp longer than a MiSeq PE 2x250, to nearly 5000 bp covering multiple rRNA genes.-->
+
+<!-- R3: This paper analyzes a transect of soils in Africa and reports how results differ from 3 differing sequencing technologies and variations to bioinformatics platforms.
+With sequencing technologies changing so rapidly, and so many methods to choose from, I’m sure this paper would be of interest to the Molecular Ecology Resources audience, as it provides useful data comparing sequencing technologies.
+I congratulate the author on herculean efforts to write so many R packages, for making his packages and scripts publicly available, and for making a pipeline that allegedly improves ASV recovery for long amplicons.
+I personally would find the paper more interesting if a treatment other than spatial turnover had been applied to these samples, like before and after a disturbance, or a different host tree, and one could examine if the treatment of interest gave similar inferences with different sequencing technologies.
+However, I believe their claim that sequencing technologies yield similar results.
+I do think the paper would be improved by adding in some alpha diversity analyses, which are likely to be less robust to vagaries of sequencing technologies than beta diversity analyses.
+I also did not see any discussion of normalization for variation in number of sequences per sample, which needs to be accounted for.
+Instead of figure 2 (which can be moved to the supplements), I would prefer instead to see a figure about correlations of alpha diversity among methods.
+I also think species accumulation curves would be useful to see.
+I have also suggested several ways the writing could be slightly altered to make it clearer and more engaging for the reader.
+The results section is quite dry, reporting the statistics rather than reporting the interesting results and using the statistics to support it.
+For example, “Microbial communities are spatially autocorrelated (R2 = 0.90, p <0.05)” is more interesting to read than the “Bray-curtis dissimilarity showed…” If some of the results section could be re-written to focus on the results rather than reporting statistics, it would be more engaging.
+I also think the introduction needs to be edited to be more cohesive with the rest of the document.
+For example, only abbreviations were used in the introduction but they were fully spelled out at numerous points in the methods, and this made the introduction unnecessarily difficult to follow.
+I also found that the discussion repeated the results a lot and thus could be truncated or re-focused to set the results more in the context of the current literature (for example some citations of similar papers was missing).
+Finally, I will freely admit that some portions of the methods section, like the discussion of LSUx, went over my head and were very difficult for me to understand and/or follow.
+I’m not sure if someone more bioinformatically inclined could do a better job reviewing some of the more technical portions or if they just need to be written more clearly.
+I actually spent 3x the amount of time I usually devote to reviews on this paper, and I made a lot of efforts to provide constructive criticisms of how this paper could be improved to be clearer for the reader, so I hope that they are helpful.
+See specific line comments below.-->
+
 # Abstract {-}
@@
 Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, *Uapaca* (Phyllanthaceae) and Fabaceae tr.
-Amherstieae [@brundrett2017a].
+Amherstieae [@brundrett2017a], which together represent approximately 60% of tree stems globally [@steidinger2019].
 
@@
 As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
-Fungal metabarcoding studies using short-read HTS technologies have targeted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
+Fungal metabarcoding studies using short-read HTS technologies such as 454, Illumina, and Ion Torrent
+have usually targeted the rDNA internal transcribed spacer regions ITS1 or ITS2,
+which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
+In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice [@opik2010], and variable regions of the rDNA large subunit (LSU) have also been used for barcoding [@house2016;@tedersoo2015].
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
-If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013].
-However, this approach comes with some potential weaknesses.
+If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013] with public databases such as Genbank [@benson2013] and Unite [@nilsson2019a].
+However, this approach comes with some potential weaknesses, discussed below.
 
-While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
+ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina [@schoch2012],
+but this variablity means that they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
+<!-- R2: This is true, but the fungi are a large and diverse group and thus it is no surprise that a highly variable rRNA marker can not be aligned.
+Perhaps this manuscript could discuss how meaningful a phylogenetic approach to analysis of complete fungal communities is anyway.
+Perhaps building a tree for each phylum is possible (and a realistic evolutionary model could be used).
+I think that this could be a contribution here - to explore how this variation could be dealt with using some knowledge of the clade "fungi".
+(Partly addressed?) -->
 Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa.
@@
 The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
-Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are less dataset specific [@callahan2017].
+Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and have been suggested to be less dataset specific [@callahan2017].
 
-Assignment of taxonomic identities using BLAST requires *a priori* choice of thresholds for different taxonomic ranks.
-Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments [e.g.,@wang2007;@edgar2016a;@murali2018a].
-In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017].
-
-Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions [@Tedersoo2018].
-This can potentially improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
-Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
-Because OTU clustering may both "clump" different species into a single OTU, and "split" a single species into multiple OTUs [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
+Because both OTU clustering and denoised ASVs may "clump" different species into a single unit and "split" a single species into multiple units [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
 In contrast, phylogenetic community distance measures [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
 Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
-Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018].
-However, long amplicon sequencing allows the inclusion of alignable regions for construction of more fully resolved phylogenetic trees directly from metabarcoding reads.
+Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018], or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases [@fouquier2016].
+
+Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
+Using BLAST requires *a priori* choice of thresholds for different taxonomic ranks.
+Although GenBank is often used as a first step for sequence identification, the database is ...
+Mention Unite, RDP, Warcup.
+<!-- R3: I would like to see a few more sentences of discussion of the issues with taxonomic assignments … I’m sure you know of them and are being brief for word limits…but I think it would be useful to mention more for young readers or people unfamiliar with fungi and issues you can come up against...like huge percentages of your sequences not getting names due to poor databases, multiple sequences getting the same names, multiple ITS copies within the same species getting different names, chimeric or incorrect taxonomic identifications in BLAST,  etc etc-->
+Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier [RDPC, @wang2007], SINTAX [@edgar2016a] and IDTAXA [@murali2018a] among others.
+In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017].
+<!-- R1: Palmer et al.(2018) use a similar ‘global’ approach to taxonomic assignment in the amptk pipeline that I think would be worth citing here.
+(They also discuss length variation in the ITS as a factor influencing taxa recovery, which is relevant to the points highlighted in the previous paragraph).
+-->
+
+Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
+This can improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis [@Tedersoo2018].
+Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
 However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
 
-Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated Soudanian woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
-We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs.
-weighted UNIFRAC).
+Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
+Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
+<!-- R2: This comes as a surprise - why use multiple sequencing strategies?
+There is no background given on the strategies used.
+Why use three different short-read platforms?
+Some discussion of how error rates differ and how the proprietary software (e.g., PacBio, Illumina) differs in its way of scoring is needed.
+The overall approach needs more justification in the Introduction section.
+I suggest either framing the question more in terms of using 1 vs 2 rRNA regions; although it must be placed in context based on the fact that other studies have used multiple rRNA regions in long-read studies); perhaps exploring in more detail how read depth affects the conclusions; however, read depth has not been mentioned in the Intro at all.-->
+We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
+and reference databases (Unite, Warcup, RDP),
+and with non-phylogenetic and phylogenetic community distance measures.
 We also present new algorithms for dividing the rDNA into regions, combining denoising results from multiple regions, and incorporating phylogenetic information into taxonomic assignments.
+<!-- R2: The need for new algorithms is not introduced or established in the Introduction. For example, what are the limitation of ITSx?-->
 
@@
 Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+<!-- R3: I entered those locations into google maps. The first one took me to Ghana, the second took me to the Pacific Ocean. Something is off – you need to fix this.-->
+Both sites were located in Soudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
 At each site, 25\ soil samples were collected along a linear transect at intervals of 1\ m in May 2015.
+<!-- R3: this is confusing. Do you mean that you had one 25m transect and 25 samples were taken along that one transect at 1m intervals?-->
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016.
@@
 <!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->
+<!-- R3: why would you lyse in the field? Aren’t you concerned about needing to refrigerate or freeze your sample? This is highly unorthodox-->
 
@@
 Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
+<!-- R3: which samples were excluded? The ones that were hand lysed or the ones that were preserved in lifeguard?-->
 However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
@@
 The gITS7 primers for the short amplicon were indexed for multiplexing (Supplementary File 1).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
+<!-- R3: so the gITS7 primer has a barcode and adaptor on it? I’m having a bit of trouble following your methods-->
+150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
 The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
@@
 Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
+<!-- R3: were technical replicates pooled prior to bead cleaning?
+Personally, I do not think technical replicates are necessary when using NGS (see Smith and Peay Plos One 2014) but I still need to be able to clearly follow your methods-->
 An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
@@
 The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.
+<!-- R3: not sure that this line is necessary-->
 
@@
 However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
+<!-- R3: Note: I have heard from many people that DADA2 performs poorly for ITS and that it was designed for 16S-->
 We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
@@
 
+<!-- R3: honestly this section is above my head and I can’t judge its accuracy. Hopefully another review is more bioinformatically inclined and can judge it better. -->
 Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
@@
 A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
+<!-- R3: you spell out RDP and cite it here, but not in your first mention of it at line 109. I would switch this around.-->
 A CM was generated from the alignment using `cmbuild` from Infernal.
@@
 Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`.
+<!-- R3: I tried to download LSUx but got an error. Can you do something about this?
+
+> install.packages("LSUx")
+And it threw the error
+“Warning message:
+package ‘LSUx’ is not available (for R version 4.0.2)”
+
+is this the github repo? https://rdrr.io/github/brendanf/LSUx/src/R/lsux.R can you put a link to this instead or in addition?
+
+If this is a methods paper, is the point to introduce LSUx?
+Can you provide a tutorial or better explanations of how one might actually implement LSUx?
+I have run analyses in QIIME2 and UPARSE many times but I don’t think I could figure out how to run your pipeline based on your paper here, and isn’t the point of this paper to make a new method available to your readers?-->
 
@@
 Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
+<!-- R3: similar to issue with RDP, you are stating amplicon sequencing variants (ASVs) here but introduced it for the first time on line 70. -->
 The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
@@
 
+<!-- R3: in general, I’m struggling to follow a lot of these methods and do not think that I could replicate your analyses -->
 For each ITS2 ASV from the long amplicon data set, the denoised sequences for the other regions corresponding to the same sequencing reads were concatenated to form a set of full-length reads.
@@
 Because ITS2 is the most variable of the amplified regions (Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
+<!-- R3: please explain how figure S5 shows me that ITS2 is the most variable of the amplified regions. I don’t get it.-->
 The concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015].
@@
 Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
+<!-- R3: I’m going to stop pointing this out – but state the full name (e.g.
+Ribosomal Data Project) the first time you introduce something, then in every other instance refer to it by its abbreviation (e.g.
+RDP).-->
 In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c as described in [@tedersoodata2017].
@@
 and IDTAXA [@murali2018].
+<!-- R3: why were these 3 algorithms chosen?-->
 Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
@@
 All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
+<!-- R3: why would you include at any level of confidence and not limit yourself to probable or highly probable?-->
 
@@
 For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
+<!-- R3: FYI, there is already a publication that compares OTU and ASV methods – Glassman and Martiny mSphere 2018 – and they result in the same inferences.
+Botnen et al Molecular Ecology Resources 2018 might also be worth citing.-->
 
@@
 The tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
+<!-- R3: But didn’t you use fungal specific primers? Wouldn’t the non-fungal kingdom reads just be PhiX?-->
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
@@
 Second, it clarifies the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these is consistent with the assignments of other ASVs in the same clade.
+<!-- R2: What is the justification for trusting the (fast, rough) tree here compared to the other algorithms?-->
 This refinement algorithm is referred to as "PHYLOTAX".
@@
 The fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina, PacBio Short, PacBio Long) were compared by PERMANOVA.
+<!-- R2: I am not a permanova expert but is it allowed when there is some problem of independence here because PacBio RS II includes both short and long reads?-->
+<!-- R3: I would say, the effect of the sequencing strategies and read length on the fungal community compositional dissimilarity, as assessed by either UNIFRAC or Bray-Curtis, was determined by PERMANOVA (cite Marti Anderson) using the Adonis function in vegan.
+I’m confused because you definitely mentioned in the intro UNIfrac but in this section only mention Bray-Curtis.
+Also you don’t mention any sort of normalization for differing sequence reads per sample, and that’s not really acceptable.
+I recommend avg.dist function in vegan.
+It lets you make the Bray-Curtis dissimilarity matrix subsampling for an even sequencing depth 100x and then take the median and perform a square root transformation all in one line of code.
+
+Also, what no alpha diversity analysis?
+Beta-diversity analyses are generally pretty robust.
+I would be way more interested to see differences in alpha diversity by sequencing technology.
+Can you please report this (normalized for differences in sequencing depth)?
+-->
 In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
@@
 Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
+<!-- Add citations for B-C and Unifrac-->
 Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12\ m.
+<!-- What is the justification for 0-12 m categories? Why not use 0-1 m? -->
 Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
@@
 
+<!-- R2: The Results section has many instances of the data "...are shown in Fig/Table..." This is difficult to read and I strongly suggest you summarize the findings in the text.
+In some cases it simply means deleting these sentences and adding the Fig/Table reference after the following sentence.
+It would be more helpful to talk about the overall trends here rather than listing all of the % values.
+-->
+
 DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Figure \@ref(fig:sample-depth).
@@
 Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
+<!-- R3: 20.7 M and 10.8M reads for only 25 samples is a lot of read depth! In the future you can go way lower than that to save money. I would multiplex in hundreds of samples for 20.8 M reads.-->
 Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long reads, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short reads, resulting in a similar number of filtered reads for the two strategies.
@@
 Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure\ \@ref(fig:full-length)b).
-The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic`, $p < 2.2\times10^{-16}$), but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
+The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic %>% formatC(format = "e", digits = 2)`, $p < 2.2\times10^{-16}$),
+<!-- S3: 2 decimal points is sufficient-->
+but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
 In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
@@
 These reads were excluded from community analyses.
+<!-- S3: so to clarify, everything was PCR amplified with a primer that had barcodes and adaptor attached to either forward or reverse primer?
+I was having trouble following this during you PCR description section.-->
 
@@
 For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
+<!-- S2: The values for Ion Torrent are 1-3 orders of magnitude higher than other methods. This to me is an important finding that is not mentioned(!) -->
+<!-- s3: I recommend using “VennDiagram” package in R.
+It allows you to make Venn diagrams with circle sizes that change based on how many samples it represents.
+That is way easier to interpret than one with circles that are the same size regardless.
+I also don’t really understand what the decimals represent – what do you mean by the fraction of reads for each sequencing technology?
+For example in figure C, is 601 sequences 79% of the long amplicon reads?
+Please clarify.
+-->
 
@@
 $R^2$ value displayed is for log-transformed non-zero read numbers.
+<!-- R2: Please provide axis labels e.g. panel (a) y-axis is "Count of ASVs" (?) -->
+<!-- R3: I’m not sure why this is useful, and I don’t think it warrants so much space in the document.
+It can be moved to the supplements.
+To me, a correlation of species richness by read technology might be interesting, or if you had some sort of treatment effect you were looking for and all of the methods gave you the same inferences of treatment effects, then that would be more interesting.
+I would prefer to see this replaced by correlation of alpha diversity.
+
+I also think species accumulation curves between sequencing technologies would be interesting to see. -->
 
@@
 In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after OTU clustering.
+<!-- R1: Is this due to the much greater length causing a different ASV than a shorter ASV, even if they have an identical sequence over the same region?
+That seems to be the case, as the clustering did not really lower the difference between the unique long and short read ASVs at all (going from 21% to 20%).
+More depth of explanation in the discussion of this issue, maybe with a couple of examples of what is driving this, would be helpful.-->
 Complete tabulations of the number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively.
@@
 For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs (Figure\ \@ref(fig:tax-chart)).
+<!-- R3: do you mean taxonomically identified? -->
 A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
@@
 These two methods are plotted in each column to compare with results for the individual databases.
+<!-- R2: Overall patterns are the same thus I suggest presenting only reads or ASVs but not both. Y-axis labels are needed as in Fig. 2. It is interesting that it is almost alinear decline-->
+<!-- R3: the barplots are not big enough to readily distinguish. Perhaps you can make the ASVs vs Reads vertical rather than horizonal and horizontally enlarge your image so it’s easier to read.-->
 
@@
 The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm) and \@ref(fig:ppcoa)).
+<!-- R1: Any thoughts on why this clade in particular was well represented?
+For example, do you think this a primer bias issue or is there some strange secondary structure in the Sordariomycetes rRNA operon that results in their removal during bioinformatic filtering?
+-->
 
@@
 For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
+<!-- R3: what does this mean?-->
 This puzzling negative correlation was significant in all correlograms based on short amplicon sequencing irrespective of technology.
@@
 The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranged in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
+<!-- R3: please re-write to better explain what you mean. Do you mean that communities generally experience full compositional turnover at 3m?-->
 Although a distance-decay model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
@@
 
+<!-- R2: The Discussion section is mostly descriptive, largely repeating the Results (e.g., lines 569-574; 578-581;589-593; 596-598; 600-606) with very little real discussion. -->
+
 ## Reconstruction of long amplicons from denoised subregions
@@
 ASV recovery for long amplicons using DADA2 was dramatically improved (`r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of reads) by denoising homologous subregions independently using the `LSUx` and `tzara` packages.
+<!-- Why does the difference range from 12-76% This is a wide range.
+I would like to know what the authors think is the reason for this.
+Under what circumstances might we expect 12% and what 76% and why?
+What is the contribution beyond that of other studies that separated long reads into constituent genes regions and examined these separately?
+-->
 Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
@@
 the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
+<!--R3: this could be visualized with a species accumulation curve-->
 
@@
 The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
+<!-- R2: I understand that dada2 may not work as well, but why not?-->
 
@@
 Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
+<!-- R3: what a bummer! -->
 Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
@@
 These differences may be due to primer mismatches in these taxa.
+<!-- R1: Can this be quantified?
+For example, do we know how well for example LR5 selects for fungi relative to alveolates based on priming site matches?
+Additionally, were the short reads more common on one side of the “long read” than the other?
+For example, if all the short reads are common on the LR5 end of the long read, that might suggest using a different LSU primer would minimize the inclusion of alveolates and other non-target taxa.-->
+<!-- R2: Primer mismatches are mentioned here, but only in relation to a few groups and as a final point in this section.
+But these are established primers whose biases are probably known and published.
+Do you results make sense?
+I would also need to see some discussion on how this bias might affect the study as a whole, because different platforms/reads used different primers.
+-->
+<!-- R3: what do you mean by smaller groups? Having fewer genera per class?-->
 
@@
 
+<!-- R2: This section is weak and could be removed. -->
 The RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences.
@@
 However, even with this considered, IDTAXA performed very poorly with the Warcup database, placing <25% of ASVs to kingdom in all datasets.
+<!-- R2: why is this the case?-->
 IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores [@murali2018a].
+<!-- "This is expected" is not a very interesting conclusion in this case.
+I do not see how any other results could occur.
+Less conflict at lower taxonomic levels would only be a reflection of different taxonomies among the databases, and would not be anything biological.
+-->
 
 @gdanetz2017 showed that a majority-rule consensus of three assignment algorithms can improve the fraction of sequences assigned as well as decrease the false assignment rate.
+<!-- R1: See my earlier note about Palmer et al. (2018) showing a similar result. -->
 Strict consensus rejects assignments whenever there is conflict between methods and should therefore provide more conservative taxonomic assignments than majority-rule consensus.
@@
 
+<!-- R2: This section is too descriptive, and would be greatly improved if reasons were given, even multiple possible reasons that lead to hypotheses.
+It appear that the only conclusion offered (UNIFRAC would be more suited to larger scales) could have been made without these data.
+-->
+<!-- R3: the Unifrac results should be reported after the Bray-Curtis results, which are more interesting. They could also be condensed.-->
+
 Weighted UNIFRAC did not reliably detect spatial structure within this relatively ecologically homogeneous community.
@@
 UNIFRAC would be more suited at larger spatial scales and/or larger ecological gradients.
+<!-- R1: A second way to try to understand this result is to see if the results differ based on weighting versus not weighting.
+Similarly, it would be nice to see is the distance-decay patterns are robust to abundance (read counts) versus incidence (pres/abs) based analyses.
+There are some in the fungal metabarcoding world who believe all high throughput sequencing (HTS) data should only be analyzed using presence absence data due to ITS copy number differences and other known issues (primer biases, etc).
+I don’t agree and I think the best way to push back on that critique is to show that the results of abundance- and incidence-based data classifications often yield similar patterns.
+The results of these suggested additional analyses could easily be incorporated as extra panels in Figure 5.
+-->
 
@@
 However, previous work in Miombo woodland, a similar ecosystem to the Soudanian woodland in this study, found autocorrelation at ranges <10\ m using Sanger sequencing of ECM root tips [@tedersoo2011], which was their smallest distance class.
+<!-- R3: I would not start this line with however, because it agrees with the sentences before if <10 m is the smallest distance class-->
 
@@
 All of these values are smaller than the 65\ m reported by @Bahram2013, also based on distance-decay curves from an ECM woodland habitat in Benin.
+<!-- R1: Is the habitat presented in Bahram et al.
+(2013) in Benin differ in some fundamentally important way from the other in this study (less disturbance, higher ECM tree basal area?It would be nice to try to better understand these differences, even if it requires a bit of speculation by the authors.-->
 
-The PacBio short amplicon dataset shows a longer spatial range, of `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m for the ECM community, in both cases with wide confidence intervals spanning `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+The PacBio short amplicon dataset shows a longer spatial range, of 
+`r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% format(digits = 0) %>% unique() %>% paste(collapse = "--")`\ m for the total fungal community and
+`r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% format(digits = 0) %>% unique() %>% paste(collapse = "--")`\ m for the ECM community,
+<!-- R1: 38-38m – is this a typo? FIXED-->
+<!-- R3: is 38-38m a typo?-->
+in both cases with wide confidence intervals spanning `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
 It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
+<!-- R1: Could this explanation be tested more formally by rarefying the non-ECM dataset down to the level of the ECM dataset and seeing if the confidence intervals and range estimates change?
+It would be nice to separate likely methodological artifact from potential differences in biology (as ECM communities likely have many more basids, so less asexual propagule production).
+-->
 
 The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0\ m and 1\ year.
+<!-- R3: This would be much more engaging to read, here and in other places, if you reword to report findings rather than stats “There was strong year to year correlation for both the total and ECM fungal communities (cite figure or stats)” -->
 The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
@@
 We hypothesize that two different processes may be at work with differing spatiotemporal scales, whose superposition result in this pattern.
+<!-- R2: This is one example of hypothesis formation, although these are very vague and could be more clearly presented - which scales and how would the different processes lead to this result? -->
 
@@
 However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments.
+<!-- R2: I don't see this as a valid conclusion, because this was not attempted with short reads. In principle it could be done. -->
 DADA2 ASV yield was initially poor for long reads, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
@@
 Laboratory work including PCR and library pooling was performed by Dr. Ylva Strid.
+<!-- R3: how come Dr Ylva Strid is not a co-author on the paper? Seems like she contributed substantially.-->
 

### 2020-09-16 writing/transect_supplement.Rmd
address most reviewer comments in intro

@@
 (ref:buffer-pcoa) For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.
+<!-- R3: your caption isn’t in line with your axes.
+Your axes say MDS1 and MDS2 – so it looks like a NMDS to me.
+But you caption says PCA.
+There is also insufficient information to jusdge this figure - No information on if reads were normalized or not, or what beta diversity metric this is based on, etc-->
 
@@
 (ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
+<!-- R3: I have no idea how to interpret these figures. More explanation is needed!-->
 
@@
 Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
+<!-- R3: Figure S4 is useful and should be in the main paper. -->
 
@@
 Node\ 4 (and thus also node\ 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
+<!-- Is there an error in the legend?
+It appears that node 3 is assigned to Taxa 1 based on box color coding and node 4 to Taxon 2 based on different box color coding.
+In the legend, it states that node 3 is assigned to Taxa 1, but there are three matches to Taxon 2 at that node and Taxon 2 is presented in Salmon color consistent with the box color for node 3.
+As such, I think the authors meant to designate node 3 as being assigned to Taxon 2.
+If this is wrong, then more details in the legend explaining why the greater presence of matches to Tax2 yet assignment to Tax1 is needed.
+-->
 
@@
 2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.
+<!-- R3: please enlarge. This is too small. -->
 

### 2020-09-16 writing/transect_paper.Rmd
Additional changes for MER and ENA submissions

--- 2020-03-19 writing/transect_paper.Rmd
@@
 bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair (`r Biostrings::readDNAStringSet(here::here("data/clusters/ITS2.fasta.gz")) %>% magrittr::extract(startsWith(names(.), "f2ab49e8")) %>% Biostrings::width() %>% unique() %>% divide_by(1, .) %>% subtract(1, .) %>% percent(digits = 1, format = "f")` similarity in ITS2).
-The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
+The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% filter(rowSums(.) > 0) %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~.
 > 0) %>% percent(format = "f", digits = 1) %>% paste(collapse = " and ")` that of the primary *A.
@@
 
- - Sequence data, including raw reads and ASV sequences, will be deposited at the European Nucleotide Archive (ENA) prior to final publication.
+ - Trimmed, demultiplexed sequencing reads have been deposited at the European Nucleotide Archive (ERA) under Project accession number PRJEB37385.
+Accession numbers are given in Supplementary Files 1 and 2.
+ - Consensus ASV sequences will also be deposited at ENA prior to publication.
  - Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication [@furneauxdata2020].
  - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
-These packages are currently being prepared for submission to Bioconductor.
-If they are not accepted at Bioconductor prior to final publication, they will be archived at Dryad.
+These packages are currently being prepared for submission to CRAN/Bioconductor.
+If they are not accepted prior to final publication, snapshots will be archived at Dryad.
  - FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.

### 2020-09-16 writing/transect_supplement.Rmd
Additional changes for MER and ENA submissions

--- 2020-03-19 writing/transect_supplement.Rmd
@@
 
-**Supplementary file 1**: Mapping between samples and indexed primers for short amplicons.
+**Supplementary file 1**: Mapping between samples, indexed primers, and read accession numbers for short amplicons.
 
-**Supplementary file 2**: Mapping between samples and indexed primers for long amplicons.
+**Supplementary file 2**: Mapping between samples, indexed primers, and read accession numbers for long amplicons.
 
@@
   mutate(value = ifelse(
-    is.na(estimate) | is.na(conf.high),
+    is.na(estimate) | is.na(conf.high) | mantel == FALSE,
     "-",
-    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
+    sprintf(
+      "%s (%s–%s)",
+      formatC(estimate, digits = 2, format = "fg"),
+      formatC(conf.low, digits = 2, format = "fg"),
+      formatC(conf.high, digits = 2, format = "fg")
+      )
   )) %>%
   filter(timelag == "0" | term == "timerange") %>%
-  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
+  pivot_wider(
+    names_from = "term",
+    values_from = "value",
+    id_cols = c("guild", "amplicon", "tech", "metric", "algorithm"),
+    values_fill = list(value = "-")
+  ) %>%
   select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%

### 2020-09-16 scripts/drake_plan_light.R

@@ -3,6 +3,7 @@ if (exists("snakemake")) {
 } else {
   load("data/plan/drake.Rdata")
 }
+if (file.exists("ENTREZ_KEY")) Sys.setenv(ENTREZ_KEY = readLines("ENTREZ_KEY"))
 suppressPackageStartupMessages({
   library(drake)
   library(magrittr)
@@ -14,7 +15,10 @@ suppressPackageStartupMessages({
   library(disk.frame)
   library(ape)
   library(ggplot2)
+  library(futile.logger)
 })
+
+setup_log("drake_light")
 
 source(file.path(config$rdir, "dada.R"))
 source(file.path(config$rdir, "mantel.R"))
@@ -29,15 +33,15 @@ regions <- read_csv(config$regions, col_types = "cccciiiiic")
 
 # Data frame for mapping between targets for guild assignment
 guilds_meta <- tibble(
-  consensus_taxa = c(
-    "phylotaxon_decipher_unconst_long_fungi",
-    "strict_taxon_short_fungi",
-    "best_taxon_short_fungi"
+  preguild_taxa = c(
+    "preguild_taxa_long_fungi_PHYLOTAX",
+    "preguild_taxa_short_fungi_Consensus",
+    "preguild_taxa_short_fungi_PHYLOTAX.Cons"
   ),
   amplicon = c("Long", "Short", "Short"),
   algorithm = c("PHYLOTAX", "Consensus", "PHYLOTAX+Cons")
 ) %>%
-  mutate_at("consensus_taxa", syms)
+  mutate_at("preguild_taxa", syms)
 
 # Data frame for mapping between targets for building phyloseq experiment objects
 physeq_meta <-
@@ -117,12 +121,6 @@ plan2 <- drake_plan(
   # Download the FUNGuild database
   funguild_db = FUNGuildR::get_funguild_db(),
   
-  # guilds_table ----
-  # Assign ecological guilds to the ASVs based on taxonomy.
-  # guilds_table = target(
-  #   FUNGuildR::funguild_assign(taxon_table, db = funguild_db),
-  #   format = "fst"),
-  
   # platemap ----
   # Read the map between plate locations and samples
   platemap = target(
@@ -137,8 +135,23 @@ plan2 <- drake_plan(
       read_rds(file_in(!!file.path("config", "tags", "long.rds"))),
       by = "well"
     ) %>%
-    select(year, site, buffer, x, sample_type, amplicon, everything()) %T>%
[…]

### 2020-09-16 scripts/mantel.R

@@ -49,7 +49,11 @@ assemble_physeq <- function(platemap, datasets, seqtable, tree = NULL, chimeras)
   asvs <- colnames(seqtable)
   if (!is.null(tree)) asvs <- intersect(asvs, tree$tip.label)
   asvs <- setdiff(asvs, chimeras)
-  asvtab <- seqtable[,asvs, drop = FALSE] %>%
+  miss_samples <- setdiff(row.names(samp), row.names(seqtable))
+  asvtab <- rbind(
+    seqtable[,asvs, drop = FALSE],
+    matrix(0, nrow = length(miss_samples), ncol = length(asvs), dimnames = list(miss_samples, asvs))
+  ) %>%
     phyloseq::otu_table(taxa_are_rows = FALSE)
   
   if (is.null(tree)) {

### 2020-09-16 scripts/variogram.R

@@ -129,7 +129,7 @@ variog_summary = function(model) {
     confint = purrr::map(
       term,
       ~ tryCatch(
-        broom::confint_tidy(model, term = .),
+        broom::confint_tidy(model, parm = .),
         error = function(e) tibble::tibble(conf.low = NA_real_, conf.high = NA_real_)
       )
     )

### 2020-09-16 writing/transect_supplement_original.tex
use tinytex to build PDFs; use latexdiff to compare original submission to new version

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfauthor={Brendan Furneaux1,; Mohammad Bahram2,3,; Anna Rosling4,; Nourou S. Yorou5,; Martin Ryberg1; 1Program in Systematic Biology, Department of Organismal Biology, Uppsala University,; Uppsala, Sweden; 2Department of Ecology, Swedish University of Agricultural Sciences, Uppsala, Sweden; 3Institute of Ecology and Earth Sciences, University of Tartu, Tartu, Estonia; 4Program in Evolutionary Biology, Department of Ecology and Genetics, Uppsala University,; Uppsala, Sweden; 5Research Unit in Tropical Mycology and Plant-Fungi Interactions, LEB,; University of Parakou, Parakou, Benin},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% table and figure caption width
\usepackage[width=0.9\textwidth]{caption}

% table formatting
\usepackage{colortbl}
\usepackage{makecell}

% lookup cross-references from the paper
\usepackage{xr}
\externaldocument{transect_paper}

% Figures and tables are prefixed with S for supplement
\renewcommand{\thetable}{S\arabic{table}} 
\renewcommand{\thefigure}{S\arabic{figure}}

% put all tables and figures on their own pages
\usepackage[nomarkers]{endfloat}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[]{biblatex}

\title{Supplementary Information for:Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities}
\author{Brendan Furneaux\textsuperscript{1}, \and Mohammad Bahram\textsuperscript{2,3}, \and Anna Rosling\textsuperscript{4}, \and Nourou S. Yorou\textsuperscript{5}, \and Martin Ryberg\textsuperscript{1} \and \textsuperscript{1}Program in Systematic Biology, Department of Organismal Biology, Uppsala University, \and Uppsala, Sweden \and \textsuperscript{2}Department of Ecology, Swedish University of Agricultural Sciences, Uppsala, Sweden \and \textsuperscript{3}Institute of Ecology and Earth Sciences, University of Tartu, Tartu, Estonia \and \textsuperscript{4}Program in Evolutionary Biology, Department of Ecology and Genetics, Uppsala University, \and Uppsala, Sweden \and \textsuperscript{5}Research Unit in Tropical Mycology and Plant-Fungi Interactions, LEB, \and University of Parakou, Parakou, Benin}
\date{}

\begin{document}
\maketitle

\textbf{Supplementary file 1}: Mapping between samples, indexed primers, and read accession numbers for short amplicons.

\textbf{Supplementary file 2}: Mapping between samples, indexed primers, and read accession numbers for long amplicons.

\textbf{Supplementary file 3}: ML tree of all long metabarcoding ASVs (5.8S--ITS2--LSU) assigned to kingdom Fungi.
Thick branches represent rapid bootstrap values \textgreater90\%.
Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, and Unite, Warcup, and RDP reference databases.
Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
Taxon names in red were reconstructed as polyphyletic.
In some cases, this may be due to a poorly identified sequence ``splitting'' the group (e.g., the single \emph{Aspergillus} sequence nested in \emph{Monascus}, Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/buffer-pcoa-1}

}

\caption[Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers]{Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers. For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.}\label{fig:buffer-pcoa}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/buffer-compare-long-1}

}

\caption{Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs.~LifeGuard) and sampling year, for long amplicon libraries.}\label{fig:buffer-compare-long}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/buffer-compare-short-1}

}

\caption{Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs.~LifeGuard) and sampling year, for short amplicon libraries.}\label{fig:buffer-compare-short}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/rDNA-1}

}

\caption[Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions]{Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines). D1-3 (grey) represent the first three variable regions in LSU, while LSU1-4 (black) represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.}\label{fig:rDNA}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/regions-1}

}

\caption[Plot of total number of ASVs recovered vs.~fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using \texttt{LSUx}]{Plot of total number of ASVs recovered vs.~fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using \texttt{LSUx}. Some regions were formed by concatenation of extracted regions, as follows:
LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
ITS = ITS1--5.8S--ITS2;
32S = 5.8S--ITS2--LSU;
long = ITS--LSU.
For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.}\label{fig:regions}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/phylotax-1}

}

\caption[Phylogenetic refinement of taxonomic assignments (PHYLOTAX)]{Phylogenetic refinement of taxonomic assignments (PHYLOTAX). In the example, a clade includes seven OTUs (A-F), which have been identified by two methods as belonging to taxa ``Tax1'' or ``Tax2'', or are unidentified (``unk''), as shown in the tip labels.
No taxon is assigned at node~1, because one of the child branches (A) is completely unassigned.
No taxon is assigned at node~2, because the assignments at C and F are inconsistent.
Node~3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
Node~4 (and thus also node~5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.}\label{fig:phylotax}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/sample-depth-1}

}

\caption[DNA concentrations after extraction and PCR, and sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016]{DNA concentrations after extraction (leftmost column) and PCR (second and third columns), and sequencing depth (right three columns) along transects at the two sites Ang and Gan for the years 2015 and 2016.
Sequencing reads which were filtered out during quality control are colored dark grey.
Dotted horizontal line at 100~reads indicates cutoff for inclusion in community analysis.
2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.}\label{fig:sample-depth}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/full-length-1}

}

\caption[Comparison of the length of denoised sequences from different length amplicons and sequencing technologies]{Comparison of the length of denoised sequences from different length amplicons and sequencing technologies. Length distribution (\emph{a}, \emph{c}) and empirical cumulative distribution function (ECDF; \emph{b}, \emph{d}) for short (\emph{a}, \emph{b}) and long (\emph{c}, \emph{d}) amplicons, respectively.}\label{fig:full-length}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/region-lengths-1}

}

\caption[Length distribution for each region extracted from long amplicons]{Length distribution for each region extracted from long amplicons. Filtering limits are shown as dashed vertical lines.}\label{fig:region-lengths}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/taxa-1}

}

\caption[Heat tree summarizing taxonomic composition of soil community]{Heat tree summarizing taxonomic composition of soil community. Color and size of nodes represent the fractional ASV richness of that taxonomic group.
Color and thickness of branches represent the fractional read abundance of all ASVs belonging to that taxonomic group.
Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
Groups which are not represented by at least 1\% of reads or 1\% of ASVs in any dataset are collapsed into nodes labeled "*``.
ASVs which could not be further assigned are labelled''?".
ASV richness and read abundance are displayed as the mean across sequencing runs.
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:taxa}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/ecm-1}

}

\caption[Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments]{Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments. Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
``unidentified'' denotes ASVs which were assigned to kingdom Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
Height of each bar represents the fraction of reads represented by that bar.
Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; ``All'' denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.}\label{fig:ecm}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/ppcoa-1}

}

\caption[Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches]{Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches, for all fungi (\textbf{a}) and ECM fungi (\textbf{b}).
Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
Only samples with at least 100 reads for all three approaches are included.
ASVs are clustered by taxonomic class for all fungi, and by family for ECM fungi.
Abbreviations represent the approximate direction of increase for the classes and families with the highest scores: \textbf{So}: Sordariomycetes, \textbf{Ag}: Agaricomycetes, \textbf{Pe}: Pezizomycetes, \textbf{Eu}: Eurotiomycetes, and \textbf{Do}: Dothideomycetes; \textbf{In}: Inocybaceae, \textbf{Am}: Amanitaceae, \textbf{Sc}: Sclerodermataceae, \textbf{Ru}: Russulaceae, \textbf{Pe}: Pezizaceae, and \textbf{Cl}: Clavulinaceae.}\label{fig:ppcoa}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/correlog-1}

}

\caption[Mantel correlograms for community dissimilarities and spatio-temporal distance]{Mantel correlograms for community dissimilarities and spatio-temporal distance. Community dissimilarities are Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right). Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.}\label{fig:correlog}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/ITS2-length-1}

}

\caption[Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies]{Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies. \emph{a}) Length distribution. \emph{b}) Empirical cumulative distribution function.}\label{fig:ITS2-length}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/heattree-length-compare-1}

}

\caption[Variation in read abundance and diversity for different amplicons across the taxonomic tree]{Variation in read abundance and diversity for different amplicons across the taxonomic tree. Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
In both cases red represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
Values are log\textsubscript{10} of the ratio of mean values across sequencing runs for each amplicon.
Groups which are not represented by at least 1\% of reads or 1\% of ASVs in any dataset are collapsed into nodes labeled "*``.
ASVs which could not be further assigned are labelled''?".
Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio \textgreater{} 3.2) are labeled.
Refer to Figure~\ref{fig:taxa} for complete taxon labeling.}\label{fig:heattree-length-compare}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_supplement_files/figure-latex/short-its-1}

}

\caption[Kingdom-level taxonomic composition of ITS2 ASVs \(\le 140\) bp]{Kingdom-level taxonomic composition of ITS2 ASVs \(\le 140\) bp. Vertical axis shows fraction of total reads \(\le 140\) bp in each dataset.}\label{fig:short-its}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/ecm-heattree-1}

}

\caption[Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only]{Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only. Red edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
Values are log\textsubscript{10} of the ratio of mean values across all sequencing technologies for each amplicon.
Groups which are not represented by at least 1\% of reads or 1\% of ASVs in any dataset are collapsed into nodes labeled "*``.
ASVs which could not be further assigned are labelled''?".
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:ecm-heattree}
\end{figure}

\begin{table}

\caption{\label{tab:region-limits}Minimum and maximum allowed lengths for each extracted region.}
\centering
\begin{tabular}[t]{lrr}
\toprule
Region & Min. length & Max. length\\
\midrule
ITS1 & 50 & 500\\
5.8S & 50 & 200\\
ITS2 & 50 & 500\\
LSU1 & 25 & 200\\
D1 & 20 & 500\\
\addlinespace
LSU2 & 50 & 500\\
D2 & 20 & 2999\\
LSU3 & 30 & 60\\
D3 & 50 & 500\\
LSU4 & 50 & 500\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption[Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages]{\label{tab:bioinfo}Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages. \textbf{Raw}: raw reads as delivered by sequencing center (CCS reads for PacBio);
\textbf{Trim}: reads with primers and demultiplexing barcodes removed;
\textbf{Filter (full)}: full length amplicons with a maximum of three expected errors;
\textbf{LSUx}: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
\textbf{Filter (ITS2)}: extracted ITS2 regions with a maximum of three expected errors;
\textbf{ITS2}: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
\textbf{short}, \textbf{ITS}, \textbf{LSU}, \textbf{long}: ITS2-based ASVs mapped to consensus ASVs for longer regions, where "short" and "long" denote the full-length short and long amplicons, respectively.}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{PacBio RS II} & \multicolumn{2}{c}{Ion Torrent Ion S5} & \multicolumn{2}{c}{Illumina MiSeq} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Long} & \multicolumn{2}{c}{Short} & \multicolumn{2}{c}{Short} & \multicolumn{2}{c}{Short} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
  & ASVs & reads & ASVs & reads & ASVs & reads & ASVs & reads\\
\midrule
Raw & – & 125,034 & – & 49,511 & – & 20,717,742 & – & 10,756,939\\
Trim & – & 104,305 & – & 41,096 & – & 15,208,677 & – & 9,513,433\\
Filter (full) & – & 44,190 & – & 38,742 & – & 12,177,705 & – & 7,674,712\\
\addlinespace
LSUx & – & 100,210 & – & 40,137 & – & 14,946,121 & – & –\\
Filter (ITS2) & – & 87,861 & – & 38,753 & – & 13,474,669 & – & –\\
ITS2 & 979 & 82,464 & 564 & 37,215 & 12,960 & 13,416,045 & 4,478 & 6,763,654\\
\addlinespace
short & – & – & 562 & 37,211 & 12,240 & 13,413,946 & 4,419 & 6,753,567\\
ITS & 805 & 81,217 & – & – & – & – & – & –\\
LSU & 777 & 81,087 & – & – & – & – & – & –\\
long & 708 & 80,287 & – & – & – & – & – & –\\
\bottomrule
\end{tabular}
\end{table}

\setlength{\tabcolsep}{1ex}

\begin{table}

\caption[Correspondences between ASVs found by different sequencing strategies]{\label{tab:venn-table-asv}Correspondences between ASVs found by different sequencing strategies. Each row shows the number of ASVs (\emph{ASVs}) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (\emph{ASVs frac}), the number of reads represented by those ASVs in each dataset (\emph{reads}) and the fraction of total reads for the dataset (\emph{reads frac}).}
\centering
\begin{tabular}[t]{rcrrcrrcrrcrrr}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{PacBio RS II} & \multicolumn{3}{c}{Illumina MiSeq} & \multicolumn{3}{c}{Ion Torrent Ion S5} \\
\cmidrule(l{3pt}r{3pt}){3-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Long} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\makecell[c]{\\} & ASVs & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac}\\
\midrule
 & 378 & 0.39 & 17k & 0.21 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 14 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.02 & 232 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 868 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.19 & 165k & 0.02 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 9373 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.72 & 1.8M & 0.14\\
\addlinespace
 & 28 & 0.03 & 682 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.01 & 12k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 41 & 0.04 & 886 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 3.2k & 0.00\\
 & 26 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.05 & 420 & 0.01 & 0.01 & 29k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 10 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.02 & 148 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 2.4k & 0.00\\
 & 2791 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.62 & 896k & 0.13 & 0.22 & 2.1M & 0.16\\
\addlinespace
 & 20 & 0.02 & 1.1k & 0.01 & 0.04 & 1.3k & 0.03 & 0.00 & 149k & 0.02 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 251 & 0.26 & 9k & 0.11 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.06 & 316k & 0.05 & 0.02 & 611k & 0.05\\
 & 233 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.41 & 5.5k & 0.15 & 0.05 & 703k & 0.10 & 0.02 & 1.4M & 0.10\\
\addlinespace
 & 261 & 0.27 & 54k & 0.65 & 0.46 & 30k & 0.80 & 0.06 & 4.5M & 0.66 & 0.02 & 7.5M & 0.56\\
\midrule
Total & 14294 & 1.00 & 82k & 1.00 & 1.00 & 37k & 1.00 & 1.00 & 6.8M & 1.00 & 1.00 & 13M & 1.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption[Correspondences between 97\% OTUs found by different sequencing strategies]{\label{tab:venn-table-otu}Correspondences between 97\% OTUs found by different sequencing strategies. Each row shows the number of OTUs (\emph{OTUs}) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (\emph{OTUs frac}), the number of reads represented by those OTUs in each dataset (\emph{reads}) and the fraction of total reads for the dataset (\emph{reads frac}).}
\centering
\begin{tabular}[t]{rcrrcrrcrrcrrr}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{PacBio RS II} & \multicolumn{3}{c}{Illumina MiSeq} & \multicolumn{3}{c}{Ion Torrent Ion S5} \\
\cmidrule(l{3pt}r{3pt}){3-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Long} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\makecell[c]{\\} & OTUs & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac}\\
\midrule
 & 327 & 0.36 & 16k & 0.20 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 1 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 2 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 133 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.05 & 8.1k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 4054 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.60 & 192k & 0.01\\
\addlinespace
 & 4 & 0.00 & 56 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 935 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 45 & 0.05 & 950 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.01 & 2.6k & 0.00\\
 & 2 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 18 & 0.00 & 0.00 & 3.1k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 4 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.01 & 18 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 712 & 0.00\\
 & 1866 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.67 & 332k & 0.05 & 0.28 & 919k & 0.07\\
\addlinespace
 & 242 & 0.26 & 5.1k & 0.06 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.09 & 124k & 0.02 & 0.04 & 350k & 0.03\\
 & 221 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.42 & 4.2k & 0.11 & 0.08 & 629k & 0.09 & 0.03 & 1.3M & 0.10\\
\addlinespace
 & 303 & 0.33 & 60k & 0.73 & 0.57 & 33k & 0.89 & 0.11 & 5.7M & 0.84 & 0.04 & 11M & 0.79\\
\midrule
Total & 7202 & 1.00 & 82k & 1.00 & 1.00 & 37k & 1.00 & 1.00 & 6.8M & 1.00 & 1.00 & 13M & 1.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:variofit}Parameters for exponential distance-decay fits. Range: range at which exponential function is at 95\% of its maximum value.}
\centering
\begin{tabular}[t]{lllllll}
\toprule
Guild & Amplicon & Tech & Algorithm & Metric & Space Range (m) & Time Range (a)\\
\midrule
All Fungi & Short & Illumina & Cons & Bray-Curtis & 13 (7.4– 23) & -\\
All Fungi & Short & Illumina & PHYLO & Bray-Curtis & 13 (7.4– 23) & -\\
All Fungi & Short & PacBio & Cons & Bray-Curtis & 25 ( 11– 59) & -\\
All Fungi & Short & PacBio & PHYLO & Bray-Curtis & 25 ( 11– 59) & -\\
All Fungi & Long & PacBio & PHYLO & Bray-Curtis & 18 (  9– 41) & 3.3 (1.9–6.9)\\
All Fungi & Long & PacBio & PHYLO & W-UNIFRAC & 478 (0–5583) & -\\
\addlinespace
ECM & Short & Illumina & Cons & Bray-Curtis & 13 (6.5– 26) & -\\
ECM & Short & Illumina & PHYLO & Bray-Curtis & 13 (6.7– 26) & -\\
ECM & Short & PacBio & Cons & Bray-Curtis & 38 ( 11– 66) & -\\
ECM & Short & PacBio & PHYLO & Bray-Curtis & 38 ( 11– 66) & -\\
ECM & Long & PacBio & PHYLO & Bray-Curtis & 14 (5.2– 30) & 4.1 (1.9– 18)\\
ECM & Long & PacBio & PHYLO & W-UNIFRAC & - & -\\
\bottomrule
\end{tabular}
\end{table}

\end{document}

### 2020-09-18 writing/transect_paper.Rmd
additional changes in response to reviewers

--- 2020-09-16 writing/transect_paper.Rmd
@@
 As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
-Fungal metabarcoding studies using short-read HTS technologies such as 454, Illumina, and Ion Torrent
+Fungal metabarcoding studies using short-read HTS technologies such as 454 Pyrosequencing, Illumina, and Ion Torrent
 have usually targeted the rDNA internal transcribed spacer regions ITS1 or ITS2,
@@
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
-If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013] with public databases such as Genbank [@benson2013] and Unite [@nilsson2019a].
+If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013] with public databases such as GenBank [@benson2013] and Unite [@nilsson2019a].
 However, this approach comes with some potential weaknesses, discussed below.
+
+Different sequencing technologies have different capabilities in terms of sequencing depth and read length, as well as differing quality profiles and potential biases [@yang2013].
+The rapid development of new HTS technologies, as well as subsequent iterative improvements in sequencing chemistry and read capacity, means that the technologies used in metabarcoding studies, along with any associated biases, change rapidly.
+As an example, the first study using HTS metabarcoding of soil fungi was published in 2009 [@buee2009] using 454 Pyrosequencing; production of 454 sequencers was subsequently discontinued in 2015, and sales of reagents stopped in 2016 [@hollmer2013].
+This brings into question the comparability of studies conducted only a few years apart.
 
@@
 (Partly addressed?) -->
-Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa.
+Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa [@ihrmark2012;@tedersoo2015;@palmer2018].
 Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].
@@
 Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and have been suggested to be less dataset specific [@callahan2017].
+Support for PacBio has recently been added to DADA2 [@callahan2019], but its application requires greater sequencing depth for longer reads, especially in high diversity samples.
 
@@
 Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
-Using BLAST requires *a priori* choice of thresholds for different taxonomic ranks.
-Although GenBank is often used as a first step for sequence identification, the database is ...
-Mention Unite, RDP, Warcup.
-<!-- R3: I would like to see a few more sentences of discussion of the issues with taxonomic assignments … I’m sure you know of them and are being brief for word limits…but I think it would be useful to mention more for young readers or people unfamiliar with fungi and issues you can come up against...like huge percentages of your sequences not getting names due to poor databases, multiple sequences getting the same names, multiple ITS copies within the same species getting different names, chimeric or incorrect taxonomic identifications in BLAST,  etc etc-->
-Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier [RDPC, @wang2007], SINTAX [@edgar2016a] and IDTAXA [@murali2018a] among others.
-In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017].
-<!-- R1: Palmer et al.(2018) use a similar ‘global’ approach to taxonomic assignment in the amptk pipeline that I think would be worth citing here.
-(They also discuss length variation in the ITS as a factor influencing taxa recovery, which is relevant to the points highlighted in the previous paragraph).
--->
+Although the public INSDC databases are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
+Consequently taxonomic assignments based on these databases may be incorrect or inconsistent.
+Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
+The Unite database is a more curated attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
+The Ribosomal Data Project [RDP, @cole2014] hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences.
+The Warcup ITS database contains 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set contains 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [@liu2012].
+Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.
+
+Assigning taxonomy to unknown sequences using BLAST requires *a priori* choice of similarity thresholds for different taxonomic ranks.
+Several algorithms specifically designed for taxonomic assignment have been published which instead use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier [RDPC, @wang2007], SINTAX [@edgar2016a] and IDTAXA [@murali2018a] among others.
+In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017;@palmer2018].
+However, all sequence-similarity based approaches are dependent on high taxonomic coverage in the reference database, making the placement of novel or undersampled groups problematic [@nilsson2016top;@Tedersoo2018].
 
@@
 Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
-<!-- R2: This comes as a surprise - why use multiple sequencing strategies?
-There is no background given on the strategies used.
-Why use three different short-read platforms?
-Some discussion of how error rates differ and how the proprietary software (e.g., PacBio, Illumina) differs in its way of scoring is needed.
-The overall approach needs more justification in the Introduction section.
-I suggest either framing the question more in terms of using 1 vs 2 rRNA regions; although it must be placed in context based on the fact that other studies have used multiple rRNA regions in long-read studies); perhaps exploring in more detail how read depth affects the conclusions; however, read depth has not been mentioned in the Intro at all.-->
 We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
@@
 and with non-phylogenetic and phylogenetic community distance measures.
-We also present new algorithms for dividing the rDNA into regions, combining denoising results from multiple regions, and incorporating phylogenetic information into taxonomic assignments.
-<!-- R2: The need for new algorithms is not introduced or established in the Introduction. For example, what are the limitation of ITSx?-->
+We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long reads in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
 
@@
 
-Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-<!-- R3: I entered those locations into google maps. The first one took me to Ghana, the second took me to the Pacific Ocean. Something is off – you need to fix this.-->
+Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
 Both sites were located in Soudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25\ soil samples were collected along a linear transect at intervals of 1\ m in May 2015.
-<!-- R3: this is confusing. Do you mean that you had one 25m transect and 25 samples were taken along that one transect at 1m intervals?-->
+At each site, 25\ soil samples were collected along a single 24 m linear transect at intervals of 1\ m in May 2015.
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016.
@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
-<!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->
-<!-- R3: why would you lyse in the field? Aren’t you concerned about needing to refrigerate or freeze your sample? This is highly unorthodox-->
+Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of field lysis/preservation buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
+Due to the inavailability of refrigeration, extraction was initiated in the field according to the manufacturers instructions by lysis using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
 
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
-<!-- R3: which samples were excluded? The ones that were hand lysed or the ones that were preserved in lifeguard?-->
+Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such the LifeGuard preserved samples were excluded from our spatial analyses.
 However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
-
-<!-- ## Soil Chemistry
-
-  - Need to get methods from Nourou
-  - or just not use it?
-  -->
 
@@
 
-The gITS7 primers for the short amplicon were indexed for multiplexing (Supplementary File 1).
+The forward primers for the short amplicon included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
 Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
-<!-- R3: so the gITS7 primer has a barcode and adaptor on it? I’m having a bit of trouble following your methods-->
 150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
 The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
-Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity, which were pooled after amplification.
+Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
 
-Both primers for the long amplicon were indexed for multiplexing (Supplementary File 2).
+Both forward and reverse primers for the long amplicon included indexes for combinatorical multiplexing (Supplementary File 2).
 PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
@@
 
-Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
-<!-- R3: were technical replicates pooled prior to bead cleaning?
-Personally, I do not think technical replicates are necessary when using NGS (see Smith and Peay Plos One 2014) but I still need to be able to clearly follow your methods-->
+After pooling of technical replicates, amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
 An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
@@
 Additionally, the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
-of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden).
+of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden) using one half of a lane.
 The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.
-<!-- R3: not sure that this line is necessary-->
 
@@
 
-### Denoising
+### Denoising and clustering
 
-We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in @callahan2019.
-However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
-<!-- R3: Note: I have heard from many people that DADA2 performs poorly for ITS and that it was designed for 16S-->
+All amplicons were denoised using DADA2 version `r packageVersion("dada2")`a ccording to the ITS pipeline workflow [@callahan2016;@callahan2020], with technology-specific modifications for Ion Torrent [@callahan2020b] and PacBio [@callahan2019].
+Although this was successful for the short reads of all technologies, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` ASVs were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
+
 We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
@@
 In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
-We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:
+
+We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into shorter homologous domains, independently denoising each segment, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
+We used this method, detailed below, for all of the PacBio and Ion Torrent datasets.
+The method is not applicable to Illumina paired-end reads.
 
@@
 <!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
-A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
-<!-- R3: you spell out RDP and cite it here, but not in your first mention of it at line 109. I would switch this around.-->
+A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
 A CM was generated from the alignment using `cmbuild` from Infernal.
 The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
-The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure\ \@ref(fig:rDNA),
-where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
+The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:rDNA),
+where LSU1-4 represent the conserved domains of LSU flanking the variable D1-3 domains [@michot1984].
 For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
-Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`.
-<!-- R3: I tried to download LSUx but got an error. Can you do something about this?
+Code to extract the segments, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available on Github at https://github.com/brendanf/LSUx.
 
-> install.packages("LSUx")
-And it threw the error
-“Warning message:
-package ‘LSUx’ is not available (for R version 4.0.2)”
-
-is this the github repo? https://rdrr.io/github/brendanf/LSUx/src/R/lsux.R can you put a link to this instead or in addition?
-
-If this is a methods paper, is the point to introduce LSUx?
-Can you provide a tutorial or better explanations of how one might actually implement LSUx?
-I have run analyses in QIIME2 and UPARSE many times but I don’t think I could figure out how to run your pipeline based on your paper here, and isn’t the point of this paper to make a new method available to your readers?-->
-
-Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors.
-Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
-<!-- R3: similar to issue with RDP, you are stating amplicon sequencing variants (ASVs) here but introduced it for the first time on line 70. -->
+Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors, and denoised using DADA2.
 The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
 Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
-For PacBio libraries, DADA2 was run with complete pooling and a band size of 16.
-For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the DADA2 FAQ (https://benjjneb.github.io/dada2/faq.html).
+DADA2 was run with complete pooling for PacBio libraries, and pseudo-pooling
+for Ion Torrent libraries.
 Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
 
-<!-- R3: in general, I’m struggling to follow a lot of these methods and do not think that I could replicate your analyses -->
-For each ITS2 ASV from the long amplicon data set, the denoised sequences for the other regions corresponding to the same sequencing reads were concatenated to form a set of full-length reads.
-For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
+For each raw read, the ASV sequences of the different domains were concatenated to form a set of full-length sequences.
+For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
 Because ITS2 is the most variable of the amplified regions (Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
 <!-- R3: please explain how figure S5 shows me that ITS2 is the most variable of the amplified regions. I don’t get it.-->
-The concatenated ASVs representing each long read were aligned in R using the `DECIPHER` package [@wright2015].
+Therefore, concatenated sequences with identical ITS2 were clustered, and each cluster was aligned in R using the `DECIPHER` package [@wright2015].
 Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the `odseq` package [@jehl2015], using the default threshold of 0.025.
-The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence.
+The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence for the entire cluster.
 Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
@@
 A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
-The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R.
+The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available at https://github.com/brendanf/tzara.
 
-Because the Illumina dataset consisted of paired-end reads, regions were not extracted prior to denoising.
-ASVs were instead generated according to a standard workflow for DADA2.
-Demultiplexed reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
-Forward and reverse reads were denoised using DADA2 version `r packageVersion("dada2")` [@callahan2016] using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
-The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the other technologies.
+Because the Illumina dataset consisted of paired-end reads, domains could not be extracted prior to denoising, and the ASVs generated according to the standard DADA2 workflow were used.
+The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.
 
-### Taxonomy assignment
+For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
+<!-- R3: FYI, there is already a publication that compares OTU and ASV methods – Glassman and Martiny mSphere 2018 – and they result in the same inferences.
+Botnen et al Molecular Ecology Resources 2018 might also be worth citing.-->
 
-Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
-<!-- R3: I’m going to stop pointing this out – but state the full name (e.g.
-Ribosomal Data Project) the first time you introduce something, then in every other instance refer to it by its abbreviation (e.g.
-RDP).-->
+### Phylogenetic inference and taxonomy assignment
+
+Full length long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10\ iterations of progressive alignment and conserved RNA secondary structure calculation and 10\ refinement iterations.
+This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by @holst-jensen1999.
+An ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
+
+Taxonomic annotations of the RDP LSU fungal training set version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
 In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c as described in [@tedersoodata2017].
@@
 
-Taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
-For each region/reference combination, taxonomy was assigned using three algorithms:
+Initial taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
+For each region/reference combination, taxonomy was assigned using three popular algorithms:
 the RDP Naïve Bayesian Classifier [RDPC, @wang2007] as implemented in DADA2;
@@
 
-Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
-All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
-<!-- R3: why would you include at any level of confidence and not limit yourself to probable or highly probable?-->
-
-### Clustering
-
-For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
-<!-- R3: FYI, there is already a publication that compares OTU and ASV methods – Glassman and Martiny mSphere 2018 – and they result in the same inferences.
-Botnen et al Molecular Ecology Resources 2018 might also be worth citing.-->
-
-### Alignment and phylogenetic inference
-
-Full length long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10\ iterations of progressive alignment and conserved RNA secondary structure calculation and 10\ refinement iterations.
-This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by @holst-jensen1999.
-
-An ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
-The tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
-<!-- R3: But didn’t you use fungal specific primers? Wouldn’t the non-fungal kingdom reads just be PhiX?-->
+The phylogenetic tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
@@
 <!-- R2: What is the justification for trusting the (fast, rough) tree here compared to the other algorithms?-->
-This refinement algorithm is referred to as "PHYLOTAX".
+This refinement algorithm is referred to as "PHYLOTAX", available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
 
@@
 A similar analysis was also applied to only fungi classified as ECM, clustered at the family level.
+Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
+All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
+<!-- R3: why would you include at any level of confidence and not limit yourself to probable or highly probable?-->
 

### 2020-09-18 writing/transect_paper_original.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities},
  pdfauthor={Brendan Furneaux1,*,; Mohammad Bahram2,3,; Anna Rosling4,; Nourou S. Yorou5,; Martin Ryberg1},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=2.5cm,right=2.5cm,top=3cm,bottom=3cm]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% line numbers -- only needed for MS
\usepackage{lineno}
\linenumbers

% put all tables and figures at the end -- only needed for MS
\usepackage[nomarkers]{endfloat}

% double spacing -- only needed for MS
\usepackage{setspace}
\doublespacing

% table and figure caption width
\usepackage[width=0.9\textwidth]{caption}

% crossreferences to supporting information -- transect_supplement.aux is required
\usepackage{xr}
\externaldocument{transect_supplement}
\usepackage[style=apa,]{biblatex}
% Produce [dataset] references
\NewBibliographyString{dataset}

\DefineBibliographyStrings{english}{
  dataset = {dataset},
}

\renewbibmacro{begentry}{%
  \ifkeyword{meta}{\textsuperscript{*}}{}%
  \ifentrytype{dataset}
    {\bibstring[\mkbibbrackets]{dataset}%
     \setunit{\addspace}}
    {}%
}
\addbibresource{all.bib}
\addbibresource{R.bib}

\title{Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities}
\author{Brendan Furneaux\textsuperscript{1,*}, \and Mohammad Bahram\textsuperscript{2,3}, \and Anna Rosling\textsuperscript{4}, \and Nourou S. Yorou\textsuperscript{5}, \and Martin Ryberg\textsuperscript{1}}
\date{\textsuperscript{1}Program in Systematic Biology, Department of Organismal Biology, Uppsala University, Uppsala, Sweden\\
\textsuperscript{2}Department of Ecology, Swedish University of Agricultural Sciences, Uppsala, Sweden\\
\textsuperscript{3}Institute of Ecology and Earth Sciences, University of Tartu, Tartu, Estonia\\
\textsuperscript{4}Program in Evolutionary Biology, Department of Ecology and Genetics, Uppsala University, Uppsala, Sweden\\
\textsuperscript{5}Research Unit in Tropical Mycology and Plant-Fungi Interactions, LEB, University of Parakou, Parakou, Benin\\
\textsuperscript{*}Corresponding author, \texttt{brendan.furneaux@ebc.uu.se}}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in taxonomic and phylogenetic placement of fungi from environmental sequences.
To address such challenges we investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400~bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600~bp) with PacBio RS II.
Resulting community structure and diversity depended more on statistical method than sequencing technology.
The use of long-amplicon sequencing enables construction of a phylogenetic tree from metabarcoding reads, which facilitates taxonomic identification of sequences.
However, long reads present issues for denoising algorithms in diverse communities.
We present a solution that splits the reads into shorter homologous regions prior to denoising, and then reconstructs the full denoised reads.
In the choice between short and long amplicons, we suggest a hybrid approach using short amplicons for sampling breadth and depth, and long amplicons to characterize the local species pool for improved identification and phylogenetic analyses.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Fungi are key drivers of nutrient cycling in terrestrial ecosystems.
One important guild of fungi form ectomycorrhizas (ECM), a symbiosis between fungi and plants in which fungal hyphae enclose the plant's fine root tips.
The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant \autocite{Smith2010}.
Approximately 8\% of described fungal species are thought to take part in ECM symbiosis \autocite{ainsworth2008ainsworth,Rinaldi2008}.
Although only about 2\% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, \emph{Uapaca} (Phyllanthaceae) and Fabaceae tr. Amherstieae \autocite{brundrett2017a}.

Although ECM fungi form many well-known mushrooms (e.g., \emph{Amanita}, \emph{Cantharellus}, \emph{Boletus}), some instead produce inconspicuous (e.g., \emph{Tomentella}) or no (e.g., \emph{Cenococcum}) fruit bodies.
Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures \autocite{horton2001}.
Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures in the field.
The advent of high-throughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils \autocite{Lindahl2013}.

As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
Fungal metabarcoding studies using short-read HTS technologies have targeted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS \autocite{schoch2012,Lindahl2013}.
The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis \autocite{Lindahl2013}.
If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST \autocite{altschul1990,Lindahl2013}.
However, this approach comes with some potential weaknesses.

While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the fungal kingdom \autocite{Lindahl2013,Tedersoo2018}.
Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa.
Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi \autocite{tedersoo2015}.

Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
In contrast, more recent denoising methods such as DADA2 \autocite{callahan2017}, Deblur \autocite{amir2017}, and UNOISE2 \autocite{edgar2016} utilize read quality information to control for sequencing error while preserving intra-species variation.
The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are less dataset specific \autocite{callahan2017}.

Assignment of taxonomic identities using BLAST requires \emph{a priori} choice of thresholds for different taxonomic ranks.
Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments \autocites[e.g.,][]{wang2007,edgar2016a,murali2018a}.
In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments \autocite{somervuo2016,gdanetz2017}.

Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions \autocite{Tedersoo2018}.
This can potentially improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
Because OTU clustering may both ``clump'' different species into a single OTU, and ``split'' a single species into multiple OTUs \autocite{ryberg2015}, diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
In contrast, phylogenetic community distance measures \autocite{wong2016} are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree \autocite{matsen2010,berger2011}, but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments \autocite{tedersoo2018}.
However, long amplicon sequencing allows the inclusion of alignable regions for construction of more fully resolved phylogenetic trees directly from metabarcoding reads.
However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number \autocite{Kennedy2018}.

Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated Soudanian woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs.~weighted UNIFRAC).
We also present new algorithms for dividing the rDNA into regions, combining denoising results from multiple regions, and incorporating phylogenetic information into taxonomic assignments.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{sampling}{%
\subsection{Sampling}\label{sampling}}

Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in woodlands dominated by the ECM host tree \emph{Isoberlinia doka} (Caesalpinioideae).
At each site, 25~soil samples were collected along a linear transect at intervals of 1~m in May 2015.
One third of the sample locations (3~m spacing) were resampled one year later in June 2016.
For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
Approximately 250~mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0~mL microtube containing 750~mL of lysis buffer and lysis beads (Xpedition\textsuperscript{TM} Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser\textsuperscript{TM}; Zymo Research Corporation).

An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard\textsuperscript{TM} Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, \(p < 0.0001\), \(R^2 = 0.06\)) from samples preserved using the Xpedition\textsuperscript{TM} lysis buffer (Figures \ref{fig:buffer-pcoa}, \ref{fig:buffer-compare-long}, and \ref{fig:buffer-compare-short}); as such these samples were excluded from our spatial analyses.
However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.

\hypertarget{dna-extraction-amplification-and-sequencing}{%
\subsection{DNA extraction, amplification, and sequencing}\label{dna-extraction-amplification-and-sequencing}}

After field lysis, DNA was extracted using the Xpedition\textsuperscript{TM} Soil/Fecal Prep kit (see above).
Samples preserved using LifeGuard were first centrifuged at 10000~g for 1~minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
DNA was quantified using fluorometrically using Quant-iT\textsuperscript{TM} PicoGreen\textsuperscript{TM} dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.

Two different fragments of the nuclear rDNA were amplified (Figure~\ref{fig:rDNA}).
The short amplicon (approximately 300~bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 \autocite{ihrmark2012} as the forward primer and a mix of ITS4 \autocite{white1990amplification} and ITS4a \autocite{urbina2016} as the reverse primer (hereafter, ITS4m).
The long amplicon (approximately 1500~bp) targeted the full ITS region including the 5.8S rDNA and approximately 950~bp at the 5' end of the LSU, including the first three variable regions (Figure~\ref{fig:rDNA}), using ITS1 \autocite{white1990amplification} as the forward primer and LR5 \autocite{vilgalys1990} as the reverse primer.
Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of \emph{Agaricus bisporus}.

The gITS7 primers for the short amplicon were indexed for multiplexing (Supplementary File 1).
Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200~µM dNTP mix, 250~µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl\textsubscript{2}, 0.1~U \emph{Taq} polymerase (Dream \emph{Taq}, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7~ng purified DNA in Dream \emph{Taq} buffer.
The reaction conditions were 10~min at 95°, followed by 35~cycles of 60~s at 95°, 45~s at 56°, and 50~s at 72°, and finally 3~min at 72°.
Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity, which were pooled after amplification.

Both primers for the long amplicon were indexed for multiplexing (Supplementary File 2).
PCR was performed as for the short amplicons, but with 500~µM of each of the two primers.
Reaction conditions were 10~min at 95°, 30~cycles of 45~s at 95°, 45~s at 59°, and 90~s at 72°, and finally 10~min at 72°.
Each reaction was performed in three technical replicates as for short amplicons.

Amplicons were purified using SPRI beads \autocite{vesterinen2016} and quantified fluorometrically as above.
An aliquot of 100~ng of DNA from each sample (or the total PCR product if less than 100~ng) was pooled into two libraries each for long and short amplicons.
Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752~37 UPPSALA, Sweden).
Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.

Additionally, the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300~bp at the SNP\&SEQ Technology Platform (Dept. of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751~44 UPPSALA, Sweden).
The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.

\hypertarget{bioinformatics}{%
\subsection{Bioinformatics}\label{bioinformatics}}

Circular consensus sequence (CCS) basecalls for PacBio sequences were made using \texttt{ccs} version 3.4 \autocite{pacificbiosciences2019} using the default settings.
The resulting sequences, as well as the paired-end Illumina sequences, were demultiplexed and sequencing primers were removed using \texttt{cutadapt} version 2.8 \autocite{martin2011}.
Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible, and these reads were thus only analyzed as a pool.
For Ion Torrent and PacBio, reads were discarded if they did not have the appropriate primers on both ends.
Reads were searched in both directions, and reads where the primers were found in the reverse direction were reverse complemented before further analysis.
For Illumina sequences, read pairs were only retained when PCR primers were detected at the 5' ends of both the forward and reverse read.
Primers were also searched for and removed on the 3' ends of the reads, in case of readthrough with short amplicons.
Read pairs where the primers were found in reverse orientation were kept in separate files, but were retained in their original orientation until after denoising.

\hypertarget{denoising}{%
\subsubsection{Denoising}\label{denoising}}

We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in \textcite{callahan2019}.
However, only 38 amplicon sequence variants (ASVs) were obtained for the long amplicons, representing 12\% of the trimmed reads.
We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads \autocite{callahan2016}.
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is \((1-\epsilon)^L\), where \(\epsilon\) is the sequencing error rate and \(L\) is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is \(2/(1-\epsilon)^L\).
For the combination of long reads (median \(L=1509 \text{ bp}\) after trimming) and moderate error rate (mean \(\epsilon= 0.0066\) based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is 43,720.
Given the high diversity relative to sequencing depth in this study (485 ASVs based on PacBio short amplicons, 104,305 trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon (\(L=265 \text{ bp}\), \(\epsilon=0.0022\)) is only 3.6 reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure \autocite{eddy1994}.
First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 \autocite{kalvari2018} using \texttt{cmsearch} from Infernal 1.1.2 \autocite{nawrocki2013}, and all bases before the 5.8S were assigned to ITS1.
No attempt was made to remove the approximately 12~bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx \autocite{bengtsson-palme2013}.
A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 \autocite{glockner2017,cole2014} by truncating after the LR5 primer site and using the reference line to annotate the variable regions \emph{sensu} \textcite{michot1984} and \textcite{raue1988}.
A CM was generated from the alignment using \texttt{cmbuild} from Infernal.
The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using \texttt{cmalign} from Infernal.
The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure~\ref{fig:rDNA},
where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions \autocite{michot1984}.
For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package \texttt{LSUx}.

Each of the extracted regions was independently filtered for length (Supplementary Table~\ref{tab:region-limits}) and a maximum of three expected errors.
Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version 1.12.1 \autocite{callahan2016,callahan2019}.
The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing run (i.e., long \emph{vs.} short amplicons, different sequencing technologies).
For PacBio libraries, DADA2 was run with complete pooling and a band size of 16.
For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the DADA2 FAQ (\url{https://benjjneb.github.io/dada2/faq.html}).
Chimeras within each region were removed using \texttt{removeBimeraDenovoTable} from DADA2.

For each ITS2 ASV from the long amplicon data set, the denoised sequences for the other regions corresponding to the same sequencing reads were concatenated to form a set of full-length reads.
For reads which were not assigned a denoised sequence for each region, the raw read for the region was used instead.
Because ITS2 is the most variable of the amplified regions (Figure~\ref{fig:regions}), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
The concatenated ASVs representing each long read were aligned in R using the \texttt{DECIPHER} package \autocite{wright2015}.
Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the \texttt{odseq} package \autocite{jehl2015}, using the default threshold of 0.025.
The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence.
Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide \textgreater50\% at a given position) were removed.
The count and sample distribution of reads assigned to each full-length ASV were calculated in order to form a sample × ASV community matrix.
A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
The process of assigning consensus full-length ASVs was carried out using the new \texttt{tzara} package for R.

Because the Illumina dataset consisted of paired-end reads, regions were not extracted prior to denoising.
ASVs were instead generated according to a standard workflow for DADA2.
Demultiplexed reads were truncated after the first base with quality score \textless= 10, and then reads with more than 3 expected errors in either read were discarded.
Forward and reverse reads were denoised using DADA2 version 1.12.1 \autocite{callahan2016} using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
The ITS2 region was extracted from the ASVs using \texttt{LSUx} for comparison to the other technologies.

\hypertarget{taxonomy-assignment}{%
\subsubsection{Taxonomy assignment}\label{taxonomy-assignment}}

Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 \autocite{cole2014} and Warcup ITS training set \autocite{deshpande2016} were mapped to the taxonomic classification system used in the Unite database version 8 \autocite{nilsson2019a}.
In particular, the classification for fungi was according to \textcite{tedersoo2018}, and for non-fungal eukaryotes was according to the proposed system of \textcite{tedersoo2017c} as described in \autocite{tedersoodata2017}.
Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists \autocite{adl2019}, which utilizes hierarchical nameless ranks.

Taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three algorithms:
the RDP Naïve Bayesian Classifier \autocite[RDPC,][]{wang2007} as implemented in DADA2;
SINTAX \autocite{edgar2016a} as implemented in VSEARCH v2.9.1 \autocite{rognes2016};
and IDTAXA \autocite{murali2018}.
Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references \(\times\) three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database \autocite{nguyen2016funguild} via the R package \texttt{FUNGuildR} (\url{https://github.com/brendanf/FUNGuildR}).
All taxa which included ``Ectomycorrhiza'' in the guild assignment at any level of confidence were included.

\hypertarget{clustering}{%
\subsubsection{Clustering}\label{clustering}}

For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97\% similarity using VSEARCH v2.9.1 \autocite{rognes2016}.

\hypertarget{alignment-and-phylogenetic-inference}{%
\subsubsection{Alignment and phylogenetic inference}\label{alignment-and-phylogenetic-inference}}

Full length long amplicon ASVs were aligned using DECIPHER \autocite{wright2015} with up to 10~iterations of progressive alignment and conserved RNA secondary structure calculation and 10~refinement iterations.
This alignment was truncated at a position after the D3 region corresponding to base 907 of the \emph{Saccharomyces cerevisiae} S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by \textcite{holst-jensen1999}.

An ML tree was produced using RAxML version 8.2.12 \autocite{stamatakis2014} using the GTR+GAMMA model and rapid bootstrapping with the MRE\_IGN stopping criterion.
The tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
The kingdom Fungi was identified as the minimal clade containing all ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
ASVs falling outside this clade were not included in downstream fungal community analysis.

Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Figure~\ref{fig:phylotax}).
A taxon at a particular rank was assigned to a node and all its descendants if that taxon was consistent with the reference-based taxonomic assignments for each of the descendants.
A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50\% confidence, or if no algorithm successfully classified the sequence at greater than 50\% confidence.
The result of this process is twofold.
First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade which is consistently given an assignment.
Second, it clarifies the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these is consistent with the assignments of other ASVs in the same clade.
This refinement algorithm is referred to as ``PHYLOTAX''.

ASVs from the short amplicon datasets were refined using only the final strict consensus step, i.e., an assignment at a given rank was accepted if there was no conflict between the different assignment algorithms at greater than 50\% confidence.
This refinement method is referred to as ``Consensus''.
Additionally, a hybrid method, was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.

\hypertarget{community-comparison}{%
\subsection{Community comparison}\label{community-comparison}}

The fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina, PacBio Short, PacBio Long) were compared by PERMANOVA.
In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
Only samples where all three strategies yielded at least 100~fungal reads (34 samples), and classes which represented at least 1\% of reads in at least one sample (14 classes), were included.
PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs.~short); and sequencing technology (Illumina MiSeq vs.~PacBio RS II).
The marginal significance of each term for explaining variation in the Bray-Curtis community dissimilarity matrix was performed using the \texttt{adonis2} function in the R package \texttt{vegan} \autocite{R-vegan}, with 9999 permutations.
Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the \texttt{capscale} function in \texttt{vegan} \autocite{R-vegan}.
Spatiotemporal effects were partialled out in order to visualize effects due to sequencing technology and amplicon length.

A similar analysis was also applied to only fungi classified as ECM, clustered at the family level.

\hypertarget{spatiotemporal-analysis}{%
\subsection{Spatiotemporal analysis}\label{spatiotemporal-analysis}}

Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
It varies between different ecosystems and taxonomic groups.
Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.~distance \autocite{legendre2012}.

Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in \texttt{phyloseq} version 1.26.0.
Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12~m.
Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.

Additionally, empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by \textcite{legendre2012} using the \texttt{nls} function in R.
Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
For datasets where the Mantel correlogram indicated spatial correlation between samples taken in separate years, the model was re-fit with an additional term to represent temporal correlation:
\[D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]\]
where \(D\), \(d\), and \(t\) represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are \(C_0\), the community dissimilarity from replicate samples (``nugget''); \(C_0 + C_1\), the community dissimilarity at long distances (``sill''); \(a_d\) the spatial range at which the community dissimilarity has moved 95\% of the way from ``nugget'' to ``sill''; and \(a_t\), the equivalent temporal range.
The 95\% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the \texttt{MASS} package in R.

\hypertarget{results}{%
\section{Results}\label{results}}

DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Figure \ref{fig:sample-depth}.
Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100~reads.
Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table~\ref{tab:bioinfo}.
Sequencing with PacBio RS II yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
Demultiplexing, primer trimming, and quality filtering reduced these totals by 64\% for PacBio long reads, but only by 21\% for PacBio short reads, resulting in a similar number of filtered reads for the two strategies.
Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with 41\% and 28\% loss, respectively.
In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of 29\% of trimmed long amplicon PacBio reads, 21\% of trimmed short amplicon PacBio reads, and 34\% of trimmed Ion Torrent reads.
This represented greater loss of PacBio short reads, but less loss of PacBio long reads and Ion Torrent reads.

Almost all of the short amplicons from all three technologies were between 240 and 375~bp long (Figure~\ref{fig:full-length}a).
Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure~\ref{fig:full-length}b).
The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = \ensuremath{8.5735947\times 10^{4}}, \(p < 2.2\times10^{-16}\)), but the difference between means was fairly small, with mean amplicon lengths of 276, 281, and 286 bp for PacBio, Ion Torrent, and Illumina, respectively.
In contrast, the length of the long amplicon reads varied widely, from 696 to 1638 bp, with a mean of 1431 bp.

The length distribution of the different regions extracted from the long amplicon are shown in Figure~\ref{fig:region-lengths}.
ITS1 showed the greatest length variability (mean \(\pm\) standard deviation: 193 \(\pm\) 55~bp), followed by ITS2 (184 \(\pm\) 41~bp) and the variable regions in LSU (D2: 227 \(\pm\) 36~bp; D3: 108 \(\pm\) 10~bp; D1: 159 \(\pm\) 5~bp).
Approximately 2\% of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure~\ref{fig:region-lengths} due to rarity.
Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations \textless{} 2 bp.

\emph{Agaricus bisporus}, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
\emph{A. bisporus} was represented by two ASVs in the Illumina dataset, which differed at one base pair (99.5\% similarity in ITS2).
The abundance of the second ASV was 1.1\% and 1.0\% that of the primary \emph{A. bisporus} ASV in the two Illumina positive controls.
The consistency of this ratio across replicate positive controls suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
Despite higher total sequencing depth, this ASV was not identified from the Ion Torrent dataset.

\emph{A. bisporus} sequences represented 0.01\%, 0.09\%, 0.09\%, and 0.09\% of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively, giving similar estimates for the rate of tag-switching for all technologies.
These reads were excluded from community analyses.

\hypertarget{reproducibility-of-sequence-capture-using-different-technologies}{%
\subsection{Reproducibility of sequence capture using different technologies}\label{reproducibility-of-sequence-capture-using-different-technologies}}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/venn-1} \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/venn-2}

}

\caption[Venn diagrams of shared ASVs and OTUs between different sequencing technologies]{Venn diagrams showing shared ITS2-based ASVs (\emph{a}, \emph{c}) and 97\% OTUs (\emph{b}, \emph{d}) between different sequencing technologies from the same short amplicon library (\emph{a}, \emph{b}), and between long and short amplicon libraries (\emph{c}, \emph{d}).
In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.
For short amplicons in \emph{c} and \emph{d}, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.}\label{fig:venn}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/read-compare-1}

}

\caption[Comparison between read numbers for different sequencing strategies]{Comparison between read numbers for different sequencing strategies, by ASV (\textbf{a}) and 97\% OTU (\textbf{b}). ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes. Dashed line represents a constant ratio of read numbers. The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading. \(R^2\) value displayed is for log-transformed non-zero read numbers.}\label{fig:read-compare}
\end{figure}

The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Figure~\ref{fig:venn}).
ASVs shared between all datasets represented 56--80\% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
These fractions increased to 73--89\% when differences at the intra-species scale were removed by clustering the ASVs into 97\% OTUs.
In particular, 100\%, 93\% , and 89\% of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
In contrast, 21\% of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to 20\% after OTU clustering.
Complete tabulations of the number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \ref{tab:venn-table-asv} and \ref{tab:venn-table-otu}, respectively.

Figure~\ref{fig:read-compare} shows the correspondence between the read count for different ASVs (\ref{fig:read-compare}a) and OTUs(\ref{fig:read-compare}b) in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted as lines along the axes.
In all cases, the read counts for shared ASVs and OTUs were correlated, with a minimum \(R^2\) value of 0.47.
Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
(0.69 to 0.72,
0.49 to 0.74, and
0.74 to 0.82,
for PacBio vs.~Illumina, PacBio vs.~Ion Torrent, and Illumina vs.~Ion Torrent, respectively), but not between the long amplicon library and short amplicon library
(0.65 to 0.62,
0.58 to 0.57, and
0.47 to 0.49, for PacBio long amplicon reads vs.~PacBio, Illumina, and Ion Torrent short reads, respectively).

\hypertarget{taxonomic-assignment}{%
\subsection{Taxonomic assignment}\label{taxonomic-assignment}}

For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs (Figure~\ref{fig:tax-chart}).
A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
At most taxonomic ranks, the RDPC algorithm assigned the greatest fraction of reads and ASVs, followed by SINTAX, and then IDTAXA.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/tax-chart-1}

}

\caption[Summary of taxonomic assignments]{Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
These two methods are plotted in each column to compare with results for the individual databases.}\label{fig:tax-chart}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/fungi-comm-1}

}

\caption[Taxonomic composition of fungal community at the class level]{Taxonomic composition of fungal community at the class level. Values represent the fraction of all ASVs, OTUs, or reads which were assigned to kingdom Fungi.
Assignments based on PHYLOTAX.
Classes which represented less than 2\% of reads, OTUs, and ASVs in all datasets are grouped together as ``other''.}\label{fig:fungi-comm}
\end{figure}

Taxonomic composition of the sequenced soil fungal community at the class level is is summarized in Figure~\ref{fig:fungi-comm} and as a heat tree \autocite{foster2017} in Figure~\ref{fig:taxa}.
The ML tree for fungal ASVs, along with taxonomic assignments, is shown in Supplementary File 3.
According to the PHYLOTAX assignments, Fungi represented 76\% of the ASVs and 90\% of the reads in the long amplicon library, compared to 89.8\%--94.9\% of the ASVs and 98.4\%--99.0\% of the reads in the short amplicon library.
Measured fungal community composition at the class level varied significantly between amplicons (PERMANOVA with 9999 permutations, \(p<0.0001\), \(R^2=0.047\)), but only marginally between sequencing technologies (\(p=0.0532\), \(R^2=0.002\)).
The majority of variation was spatiotemporal (i.e., between samples; \(p<0.0001\), \(R^2=0.90\)), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures~\ref{fig:fungi-comm} and \ref{fig:ppcoa}).

Fungi categorized as ECM made up 8.5\% of ASVs and 39.4\% of reads in the long amplicon library, and 6.3\%--13.8\% of the ASVs and 36.7\%--47.4\% of the reads in the short amplicon library (Figure~\ref{fig:ecm}).
Although amplicon length had a significant effect on ECM community composition at the family level, the explained variation was very low (PERMANOVA with 9999 permutations, \(p=0.0019\), \(R^2=0.002\)), and the majority of variation was again spatiotemporal (\(p<0.0001\), \(R^2=0.98\)).
Variation between sequencing technologies was not significant (\(p=0.76\), \(R^2=0.0002\)).

\hypertarget{spatial-analysis}{%
\subsection{Spatial analysis}\label{spatial-analysis}}

Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at \(p < 0.05\) for ranges of up to 2--3~m for the total fungal community, and 1--2~m for the ECM fungal community (Figure~\ref{fig:correlog}).
In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
The least spatial structure was detected with PacBio short amplicon sequencing.

The Bray-Curtis metric showed significant (\(p < 0.05\)) positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0~m, time lag of 1~year), for both the total fungal and ECM fungal communities in the long amplicon library.
For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0~m and 1~year was not significant, but there was a negative correlation at time lag of 1~year and a distance of 1~m for both sequencing technologies.
This puzzling negative correlation was significant in all correlograms based on short amplicon sequencing irrespective of technology.

In contrast to the Bray-Curtis distance, the weighted UNIFRAC distance showed very little spatial structure, with only the total fungal community in the 1~m
distance class showing a significant correlation at \(p < 0.05\).
No temporal correlation was found for the weighted UNIFRAC distance.

The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95\% confidence intervals for all of the Bray-Curtis spatial ranged in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}).
Although a distance-decay model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0~m, indicating no spatial structure, was included in the 95\% confidence interval.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/spatio-1}

}

\caption[Distance-decay plot for community dissimilarities and spatio-temporal distance]{Distance-decay plot for community dissimilarities and spatio-temporal distance. Circles represent community data from short (top two rows) and long (bottom two rows) amplicon libraries, sequenced by Illumina MiSeq (top row) or PacBio RS II (bottom three rows).
Community dissimilarities are calculated using the Bray-Curtis dissimilarity for all datasets (top three rows) and using the weighted UNIFRAC dissimilarity for the long amplicon library, for which a phylogenetic tree could be constructed (bottom row).
The left column represents the full fungal community, and the right column only sequences identified as ECM.
The color of each circle represents the time lag between samples being compared (0 or 1~year), and the size represents the number of comparisons for that spatial distance and time lag.
Lines are the best-fit lines for an exponential decay to max model.
The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0~year timelag) or spatiotemporal (for the 1~year time lag) distance.}\label{fig:spatio}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{reconstruction-of-long-amplicons-from-denoised-subregions}{%
\subsection{Reconstruction of long amplicons from denoised subregions}\label{reconstruction-of-long-amplicons-from-denoised-subregions}}

ASV recovery for long amplicons using DADA2 was dramatically improved (12\% to 76\% of reads) by denoising homologous subregions independently using the \texttt{LSUx} and \texttt{tzara} packages.
Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
Thus, \texttt{tzara} should increase ASV recovery from these platforms as well.
It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding \autocite{loit2019}.

\hypertarget{comparison-of-sequencing-strategies}{%
\subsection{Comparison of sequencing strategies}\label{comparison-of-sequencing-strategies}}

The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (75\%) compared to Illumina (63\%) or Ion Torrent (65\%),
the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.

DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97\% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure~\ref{fig:venn}) and the read counts for each sequence (Figure~\ref{fig:read-compare}).
The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed \autocite{callahan2016}.

Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any ITS2 fragments longer than those recovered by Illumina and Ion Torrent.
Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to \emph{Cantharellus}, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA \autocite{moncalvo2006} and longer ITS regions than other fungi \autocite{Feibelman_1994}, making it an especially difficult target for metabarcoding.
Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures~\ref{fig:full-length} and \ref{fig:ITS2-length}).

Of long amplicon reads, 21\% belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97\% similarity only reduced this fraction to 20\%.
Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \ref{fig:ITS2-length}).
Taxonomic assignments revealed that the majority of these non-shared sequences fall outside kingdom Fungi (Figure~\ref{fig:heattree-length-compare}), and that in particular the short ITS2 sequences are mostly Alveolates (Figure~\ref{fig:short-its}).
Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figures~\ref{fig:fungi-comm}, \ref{fig:ppcoa}, and \ref{fig:heattree-length-compare}).
Additionally, several smaller groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and \emph{Myerozyma} in the short amplicon datasets (Figures~\ref{fig:heattree-length-compare} and \ref{fig:ecm-heattree}).
These differences may be due to primer mismatches in these taxa.

\hypertarget{taxonomic-identification}{%
\subsection{Taxonomic identification}\label{taxonomic-identification}}

The RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences.
The Warcup database placed notably fewer sequences at all taxonomic levels for all datasets, probably in part due to the fact that only fungal sequences are included.
However, even with this considered, IDTAXA performed very poorly with the Warcup database, placing \textless25\% of ASVs to kingdom in all datasets.
IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores \autocite{murali2018a}.

\textcite{gdanetz2017} showed that a majority-rule consensus of three assignment algorithms can improve the fraction of sequences assigned as well as decrease the false assignment rate.
Strict consensus rejects assignments whenever there is conflict between methods and should therefore provide more conservative taxonomic assignments than majority-rule consensus.
Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications.
This suggests that different assignment algorithms and databases bring mostly complementary, non-contradictory information at higher taxonomic levels.
However, contradictory assignments between different methods is more common at lower taxonomic levels, which can be problematic because accurate assignment at the family or genus level is generally required for ecological guild assignment using FUNGuild.

For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
For instance, 56\% and 87\% of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to 75\% and 97\%.
This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure~\ref{fig:ecm}).

\hypertarget{turnover-rate}{%
\subsection{Turnover rate}\label{turnover-rate}}

Weighted UNIFRAC did not reliably detect spatial structure within this relatively ecologically homogeneous community.
Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1~m; Figure~\ref{fig:correlog}), the distance-decay plot in Figure~\ref{fig:spatio} does not show any clear relationship.
The functional fit showed poor convergence, with a 95\% confidence interval for spatial range of 0--5700~m, indicating little evidence of spatial structure.
This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure~\ref{fig:taxa}), which are both well represented in the majority of samples.
UNIFRAC would be more suited at larger spatial scales and/or larger ecological gradients.

Mantel correlograms based on the Bray-Curtis dissimilarity (Figure~\ref{fig:correlog}) revealed spatial autocorrelation in the soil fungal community at distance classes \(\le 3\) m for both Illumina and PacBio using long and short amplicons, and in the ECM fungal community at distance classes \(\le 2\) m for Illumina and PacBio long amplicons, and \(\le 1\) m for the PacBio short amplicons.
These results are similar to autocorrelation ranges found in previous work based on ECM root tips in temperate forests \autocite{lilleskov2004,pickles2012a}.
\textcite{lilleskov2004} found autocorrelation only at ranges \textless2.6~m at most sites using Sanger sequencing.
Similarly, \textcite{pickles2012a} found autocorrelation at distances \textless3.4~m based on T-RFLP analysis.
However, previous work in Miombo woodland, a similar ecosystem to the Soudanian woodland in this study, found autocorrelation at ranges \textless10~m using Sanger sequencing of ECM root tips \autocite{tedersoo2011}, which was their smallest distance class.

Distance-decay plots (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}) gave substantially longer autocorrelation distances.
There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 13--18~m.
The 95\% confidence interval was substantially wider than this variation, generally covering a range of 5--41~m.
All of these values are smaller than the 65~m reported by \textcite{Bahram2013}, also based on distance-decay curves from an ECM woodland habitat in Benin.

The PacBio short amplicon dataset shows a longer spatial range, of 25~m for the total fungal community and 38--38~m for the ECM community, in both cases with wide confidence intervals spanning 11--214~m.
It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.

The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0~m and 1~year.
The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.3~years for the total fungal community and 4.1~years for the ECM community, but with overlapping confidence intervals.
Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0~m and 1~year, but a weak negative correlation at 1~m and 1~year.
The general shape of the correlograms were similar for long and short amplicon datasets.
We hypothesize that two different processes may be at work with differing spatiotemporal scales, whose superposition result in this pattern.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

The choice of amplicon and sequencing technology did not seem to affect the results of the spatial analysis, provided sufficient sequencing depth.
However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments.
DADA2 ASV yield was initially poor for long reads, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
Together these approaches provide a hybrid approach using long-read sequencing to acquire long amplicon sequences for the local species pool, and cost-effective short-read sequencing to provide high sampling depth and sample number.

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

This project was funded by the Swedish research council FORMAS grant number 2014-01109.

Laboratory work including PCR and library pooling was performed by Dr.~Ylva Strid.

The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden.

Sequencing was performed by the SNP\&SEQ Technology Platform in Uppsala.
The facility is part of the National Genomics Infrastructure (NGI) Sweden and Science for Life Laboratory.
The SNP\&SEQ Platform is also supported by the Swedish Research Council and the Knut and Alice Wallenberg Foundation.

\singlespacing

\printbibliography

\doublespacing

\hypertarget{data-accessibility}{%
\section*{Data Accessibility}\label{data-accessibility}}
\addcontentsline{toc}{section}{Data Accessibility}

\begin{itemize}
\tightlist
\item
  Trimmed, demultiplexed sequencing reads have been deposited at the European Nucleotide Archive (ERA) under Project accession number PRJEB37385. Accession numbers are given in Supplementary Files 1 and 2.
\item
  Consensus ASV sequences will also be deposited at ENA prior to publication.
\item
  Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication \autocite{furneauxdata2020}.
\item
  R packages \texttt{LSUx}, \texttt{tzara}, \texttt{phylotax}, and \texttt{FUNGuildR} are available on Github at \url{https://github.com/brendanf/LSUx}, \url{https://github.com/brendanf/tzara}, \url{https://github.com/brendanf/phylotax}, and \url{https://github.com/brendanf/FUNGuildR}.
  These packages are currently being prepared for submission to CRAN/Bioconductor.
  If they are not accepted prior to final publication, snapshots will be archived at Dryad.
\item
  FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}. The versions used in this paper will be archived at Dryad prior to publication.
\item
  Bioinformatics pipeline and analysis scripts are available at \url{https://github.com/oueme-fungi/oueme-fungi-transect}.
\end{itemize}

\hypertarget{author-contributions}{%
\section*{Author Contributions}\label{author-contributions}}
\addcontentsline{toc}{section}{Author Contributions}

Sampling was planned and carried out by BF, NSY, and MR.
Bioinformatics and data analysis were performed by BF with input from MB, AR, and MR.
Scripts and R packages were written by BF.
The manuscript was drafted by BF and MR.
All authors contributed to and approved the final version of the manuscript.

\processdelayedfloats

\end{document}

### 2020-09-25 writing/transect_paper.Rmd
changes to address reviewer comments with Anna

@@
 Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in taxonomic and phylogenetic placement of fungi from environmental sequences.
-To address such challenges we investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
+To address such challenges we investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400\ bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, all from the same PCR library, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600\ bp) with PacBio RS II.
 Resulting community structure and diversity depended more on statistical method than sequencing technology.
@@
 and with non-phylogenetic and phylogenetic community distance measures.
-We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long reads in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
+We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
 
@@
 
-Two different fragments of the nuclear rDNA were amplified (Figure\ \@ref(fig:rDNA)).
+The DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:rDNA)) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
+Due to length restrictions for Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
 The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
@@
 
-Additionally, the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
+Additionally, the same short amplicon PCR libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
 of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden) using one half of a lane.
-The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.
 
@@
 All amplicons were denoised using DADA2 version `r packageVersion("dada2")`a ccording to the ITS pipeline workflow [@callahan2016;@callahan2020], with technology-specific modifications for Ion Torrent [@callahan2020b] and PacBio [@callahan2019].
-Although this was successful for the short reads of all technologies, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` ASVs were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
+Although this was successful for the short amplicons on all technologies, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` ASVs were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
 
-We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
+We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity.
 The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
@@
 
-Initial taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
+Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
 For each region/reference combination, taxonomy was assigned using three popular algorithms:
@@
 and IDTAXA [@murali2018].
-<!-- R3: why were these 3 algorithms chosen?-->
+A relatively lax confidence threshhold of 50% was used for all three algorithms.
 Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
 ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.
+Conflicts between the assignments made by the different algorithms were later resolved using a phylogenetic method and a consensus method, which are described below.
 
@@
 
-Taxonomic assignments of ASVs from the long amplicon dataset were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax)).
-A taxon at a particular rank was assigned to a node and all its descendants if that taxon was consistent with the reference-based taxonomic assignments for each of the descendants.
+For the ASVs from the long amplicon dataset, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax)).
+Taxa at each rank were assigned to a node and all its descendants if that taxon was consistent with the reference-based taxonomic assignments for each of the descendants.
 A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
 The result of this process is twofold.
-First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade which is consistently given an assignment.
-Second, it clarifies the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these is consistent with the assignments of other ASVs in the same clade.
-<!-- R2: What is the justification for trusting the (fast, rough) tree here compared to the other algorithms?-->
+First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade which is consistently assigned by the .
+Second, it resolves conflicts between the different reference-based algorithms, but only if one of the alternatives is consistent with the assignments of other ASVs in the same clade.
 This refinement algorithm is referred to as "PHYLOTAX", available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
@@
 
-## Community comparison
+## Effect of sequencing strategy on recovered community
 
-The fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina, PacBio Short, PacBio Long) were compared by PERMANOVA.
-<!-- R2: I am not a permanova expert but is it allowed when there is some problem of independence here because PacBio RS II includes both short and long reads?-->
+The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the R package `vegan` [@R-vegan].
+Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).
 <!-- R3: I would say, the effect of the sequencing strategies and read length on the fungal community compositional dissimilarity, as assessed by either UNIFRAC or Bray-Curtis, was determined by PERMANOVA (cite Marti Anderson) using the Adonis function in vegan.
@@
 Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
+The community matrix was normalized to proportional abundance by dividing the counts for each class in each sample by the total number of reads for that sample, but were not rarefied to even sampling depth [@mcmurdie2014].
 PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs. short); and sequencing technology (Illumina MiSeq vs. PacBio RS II).
-The marginal significance of each term for explaining variation in the Bray-Curtis community dissimilarity matrix was performed using the `adonis2` function in the R package `vegan` [@R-vegan], with 9999 permutations.
 Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
@@
 All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
-<!-- R3: why would you include at any level of confidence and not limit yourself to probable or highly probable?-->
 
@@
 
-Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
-<!-- Add citations for B-C and Unifrac-->
-Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12\ m.
-<!-- What is the justification for 0-12 m categories? Why not use 0-1 m? -->
+Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric [@bray1957, for both long and short amplicons] and the phylogenetically based weighted UniFrac metric [@lozupone2005;@lozupone2007, for only long amplicons] in `phyloseq` version `r packageVersion("phyloseq")`.
+Each of these distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
 Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
@@
 <!-- R3: 20.7 M and 10.8M reads for only 25 samples is a lot of read depth! In the future you can go way lower than that to save money. I would multiplex in hundreds of samples for 20.8 M reads.-->
-Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long reads, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short reads, resulting in a similar number of filtered reads for the two strategies.
+Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
 Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
 In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-readd(filtercounts_ITS2)['pb_483']/readd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-readd(filtercounts_ITS2)['is_057']/readd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
-This represented greater loss of PacBio short reads, but less loss of PacBio long reads and Ion Torrent reads.
+This represented greater loss of PacBio short amplicon reads, but less loss of PacBio long amplicon reads and Ion Torrent reads.
 
@@
 `r compR2("pb_500", "SH-2257", "ASV")` to `r compR2("pb_500", "SH-2257", "OTU")`, and
-`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short reads, respectively).
+`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short amplicon reads, respectively).
 
@@
 
-In contrast to the Bray-Curtis distance, the weighted UNIFRAC distance showed very little spatial structure, with only the total fungal community in the 1\ m 
+In contrast to the Bray-Curtis distance, the weighted UniFrac distance showed very little spatial structure, with only the total fungal community in the 1\ m
 distance class showing a significant correlation at $p < 0.05$.
-No temporal correlation was found for the weighted UNIFRAC distance.
+No temporal correlation was found for the weighted UniFrac distance.
 
@@
 <!-- R3: please re-write to better explain what you mean. Do you mean that communities generally experience full compositional turnover at 3m?-->
-Although a distance-decay model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
+Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
 
@@
 (ref:spatio) Circles represent community data from short (top two rows) and long (bottom two rows) amplicon libraries, sequenced by Illumina MiSeq (top row) or PacBio RS II (bottom three rows).
-Community dissimilarities are calculated using the Bray-Curtis dissimilarity for all datasets (top three rows) and using the weighted UNIFRAC dissimilarity for the long amplicon library, for which a phylogenetic tree could be constructed (bottom row).
+Community dissimilarities are calculated using the Bray-Curtis dissimilarity for all datasets (top three rows) and using the weighted UniFrac dissimilarity for the long amplicon library, for which a phylogenetic tree could be constructed (bottom row).
 The left column represents the full fungal community, and the right column only sequences identified as ECM.
@@
 
-Weighted UNIFRAC did not reliably detect spatial structure within this relatively ecologically homogeneous community.
+Weighted UniFrac did not reliably detect spatial structure within this relatively ecologically homogeneous community.
 Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m; Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship.
@@
 This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure\ \@ref(fig:taxa)), which are both well represented in the majority of samples.
-UNIFRAC would be more suited at larger spatial scales and/or larger ecological gradients.
+UniFrac would be more suited at larger spatial scales and/or larger ecological gradients.
 <!-- R1: A second way to try to understand this result is to see if the results differ based on weighting versus not weighting.
@@
 <!-- R2: I don't see this as a valid conclusion, because this was not attempted with short reads. In principle it could be done. -->
-DADA2 ASV yield was initially poor for long reads, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
+DADA2 ASV yield was initially poor for long amplicons, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
 Together these approaches provide a hybrid approach using long-read sequencing to acquire long amplicon sequences for the local species pool, and cost-effective short-read sequencing to provide high sampling depth and sample number.

### 2020-09-25 writing/transect_paper.Rmd
combine figs S4 and S5 and move to main document

@@
     suppress-bibliography: true
+    extra_dependencies: "subfig"
   bookdown::word_document2: 
@@
 
-The DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:rDNA)) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
+The DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
 Due to length restrictions for Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
 The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
-The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:rDNA)), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
+The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:regions)a), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
 Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
@@
 The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
-The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:rDNA),
+The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:regions),
 where LSU1-4 represent the conserved domains of LSU flanking the variable D1-3 domains [@michot1984].
@@
 For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
-Because ITS2 is the most variable of the amplified regions (Figure\ \@ref(fig:regions)), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
+Because ITS2 is the most variable of the amplified regions (Figure\ \@ref(fig:regions)b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
 <!-- R3: please explain how figure S5 shows me that ITS2 is the most variable of the amplified regions. I don’t get it.-->
@@
 The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available at https://github.com/brendanf/tzara.
+
+(ref:region-asv-s) rDNA regions
+
+(ref:region-asv) **(a)** Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).
+D1-3 (grey) represent the first three variable regions in LSU, while LSU1-4 (black) represent the conserved regions.
+Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
+**(b)** Total number of ASVs recovered vs.
+fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
+Some regions were formed by concatenation of extracted regions, as follows:
+LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
+ITS = ITS1--5.8S--ITS2;
+32S = 5.8S--ITS2--LSU;
+long = ITS--LSU.
+For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.
+
+```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)", fig.subcap=rep("", 2), fig.ncol = 1, fig.align = "center"}
+knitr::include_graphics("rDNA_amplicons", auto_pdf = TRUE)
+readd(region_table) %>%
+  dplyr::filter(seq_run == "pb_500") %>%
+  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
+  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
+  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
+  geom_point() +
+  ggrepel::geom_text_repel() +
+  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
+  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
+```
 

### 2020-09-25 writing/transect_supplement.Rmd
combine figs S4 and S5 and move to main document

@@
 
-```
-
-(ref:rDNA-s) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions
-
-(ref:rDNA) Portion of the rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).
-D1-3 (grey) represent the first three variable regions in LSU, while LSU1-4 (black) represent the conserved regions.
-Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
-<!-- R3: Figure S4 is useful and should be in the main paper. -->
-
-```{tikz rDNA, fig.cap = "(ref:rDNA)", fig.scap = "(ref:rDNA-s)", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}
-
-```
-
-(ref:region-asv-s) Plot of total number of ASVs recovered vs. fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`
-
-(ref:region-asv) Some regions were formed by concatenation of extracted regions, as follows:
-LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
-ITS = ITS1--5.8S--ITS2;
-32S = 5.8S--ITS2--LSU;
-long = ITS--LSU.
-For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.
-
-```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)"}
-readd(region_table) %>%
-  dplyr::filter(seq_run == "pb_500") %>%
-  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
-  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
-  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
-  geom_point() +
-  ggrepel::geom_text_repel() +
-  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
-  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
 ```

### 2020-09-25 writing/transect_supplement.Rmd
source output_functions if knitting from Rstudio

@@
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
+if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))
 ```

### 2020-09-25 writing/transect_paper.Rmd
improve Fig 1 caption and description in methods text

@@
 
-Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors, and denoised using DADA2.
-The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
-Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
+Each of the extracted domains was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors, and denoised using DADA2.
+To test the performance of DADA2 on intermediate-length amplicons, adjacent domains from each read were concatenated to form full ITS (ITS1 + 5.8S + ITS2), partial LSU (LSU1 + D1 + LSU2 + D2 + LSU3 + D3 + LSU4), and partial 32S (5.8S + ITS2 + LSU) datasets, and these were also denoised.
+The DADA2 error model was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
+Independent error models were fit for each sequencing strategy (i.e., long *vs.* short amplicons, different sequencing technologies).
 DADA2 was run with complete pooling for PacBio libraries, and pseudo-pooling
@@
 
-For each raw read, the ASV sequences of the different domains were concatenated to form a set of full-length sequences.
+For each raw read, the denoised sequences of the different domains were concatenated to form a set of full-length sequences.
 For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
-Because ITS2 is the most variable of the amplified regions (Figure\ \@ref(fig:regions)b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
-<!-- R3: please explain how figure S5 shows me that ITS2 is the most variable of the amplified regions. I don’t get it.-->
+Because ITS2 is the most variable of the amplified domains as indicated by the greatest number of unique ASVs (Figure\ \@ref(fig:regions)b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
 Therefore, concatenated sequences with identical ITS2 were clustered, and each cluster was aligned in R using the `DECIPHER` package [@wright2015].
@@
 Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
-**(b)** Total number of ASVs recovered vs.
-fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`.
-Some regions were formed by concatenation of extracted regions, as follows:
-LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4;
-ITS = ITS1--5.8S--ITS2;
-32S = 5.8S--ITS2--LSU;
-long = ITS--LSU.
-For this figure, subregions were concatenated prior to quality filtering and generating ASVs with DADA2.
+**(b)** Total number of DADA2 ASVs vs.
+fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using `LSUx`.
+Shorter and more conserved regions yield a greater fraction of successfully mapped reads<!-- This sentence is a result shown in the figure (although it is also theoretically expected)-->.
+At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs<!-- This sentence is how to interpret the figure -->.
 
@@
   scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
-  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
+  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1)) +
+  ylab("Unique ASVs")
 ```
@@
 Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
-<!-- R3: what a bummer! -->
 Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
@@
 Additionally, several smaller groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
+<!-- R3: what do you mean by smaller groups? Having fewer genera per class?-->
 These differences may be due to primer mismatches in these taxa.
@@
 -->
-<!-- R3: what do you mean by smaller groups? Having fewer genera per class?-->
 

### 2020-09-25 writing/qual-check.Rmd
update quality check so it runs again.

--- 2019-02-15 writing/qual-check.Rmd
@@
 title: "Read quality check"
-output: beamer_presentation
+output:
+  beamer_presentation:
+    latex_engine: xelatex
 ---
@@
 library(drake)
-knitr::opts_chunk$set(echo = FALSE)
+source(here::here("scripts", "qstats.R"))
+source(here::here("scripts", "output_functions.R"))
+knitr::opts_chunk$set(echo = FALSE, device = "tikz")
+theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
 ```
@@
 # pos <- readd(positions_pb500002A1f)
-reads <- readd(qstats_join)
+reads <- readd(parsed_qstat) %>%
+  dplyr::group_by(seq_run, region, step, stat, value, region, read) %>%
+  dplyr::summarize_at("nreads", sum) %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate_at(
+    "region",
+    factor,
+    levels = c(NA, "ITS2", "ITS1", "5_8S", "LSU1", "LSU2", "LSU3", "LSU4", "D1", "D2", "D3", "short", "ITS", "LSU", "32S", "long")
+  ) %>%
+  dplyr::left_join(
+    dplyr::select(datasets, dataset, seq_run, tech, machine, amplicon),
+    by = "seq_run"
+  ) %>%
+  dplyr::mutate_at(
+    "dataset", factor,
+    levels = c("short-ion", "short-illumina", "short-pacbio", "long-pacbio")
+  ) %>%
+  tidyr::replace_na(list(read = "")) %>%
+  dplyr::mutate(`Read type` = paste(tech, amplicon, substr(read, 2, 3)) %>%
+                  trimws())
 ```
@@
 
-## Minimum quality score
+## Minimum quality score -- strategies
+
 ```{r}
-ggplot(reads, aes(x = minq,
-                  linetype = Region,
-                  group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+dplyr::filter(reads, stat == "minq", step == "raw") %>%
+  dplyr::arrange(value) %>%
+  dplyr::group_by(`Read type`) %>%
+  dplyr::mutate(ecdf = 1 - cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = `Read type`,
+             group = `Read type`,
+             color = `Read type`)) +
+  geom_line() +
   scale_x_continuous(name = "Minimum quality score",
@@
                      oob = squish) +
-  scale_y_continuous(name = "Fraction passing")
+  scale_y_continuous(name = "Fraction passing") +
+  scale_color_read() +
+  scale_linetype_discrete()
 ```
 
-## Expected error rate
+## Minimum quality score -- regions
+
 ```{r}
+dplyr::filter(reads, stat == "minq", step == "lsux", amplicon == "Long",
+              region %in% c("ITS1", "ITS2", "5_8S", "ITS", "LSU", "long")) %>%
+  dplyr::arrange(value) %>%
+  dplyr::group_by(region) %>%
+  dplyr::mutate(ecdf = 1 - cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = region,
+             group = region,
+             color = region)) +
+  geom_line() +
+  scale_x_continuous(name = "Minimum quality score",
+                     # trans = reverselog_trans(10),
+                     # limits = c(NA, 1e-5),
+                     oob = squish) +
+  scale_y_continuous(name = "Fraction passing") +
+  scale_color_viridis_d(end = 0.95) +
+  scale_linetype_discrete()
+```
 
-ggplot(reads, aes(x = erate,
-                  linetype = Region,
-                  group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
-  scale_x_continuous(name = "Expected number of errors per base",
+## Expected error rate --strategies
+
+```{r}
+erate_plot <- dplyr::filter(reads, stat == "erate", step == "raw") %>%
+  dplyr::arrange(value) %>%
+  dplyr::group_by(`Read type`) %>%
+  dplyr::mutate(ecdf = cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = `Read type`,
+             group = `Read type`,
+             color = `Read type`)) +
+  geom_line() +
+  scale_x_continuous(name = "Expected error-rate",
                      trans = reverselog_trans(10),
@@
                      oob = squish) +
-  scale_y_continuous(name = "Fraction passing")
+  scale_y_continuous(name = "Fraction passing") +
+  scale_color_read(guide = guide_legend(title = NULL, ncol = 2)) +
+  scale_linetype_discrete(guide = guide_legend(title = NULL, ncol = 2)) +
+  theme(legend.position = "bottom")
+erate_plot
 ```
 
-## Expected number of errors
+## Expected number of errors --strategies
+
 ```{r}
-ggplot(reads, aes(x = eexp,
-                  linetype = Region,
-                  group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
+eexp_plot <- dplyr::filter(reads, stat == "eexp", step == "raw") %>%
+  dplyr::arrange(value) %>%
+  dplyr::group_by(`Read type`) %>%
+  dplyr::mutate(ecdf = cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = `Read type`,
+             group = `Read type`,
+             color = `Read type`)) +
+  geom_vline(xintercept = 3, linetype = "dashed", color = "gray50") +
+  geom_line() +
   scale_x_continuous(name = "Expected number of errors",
@@
                      oob = squish) +
-  scale_y_continuous(name = "Fraction passing")
+  scale_y_continuous(name = "Fraction passing") +
+  scale_color_read(guide = guide_legend(title = NULL, ncol = 2)) +
+  scale_linetype_discrete(guide = guide_legend(title = NULL, ncol = 2)) +
+  theme(legend.position = "bottom")
+eexp_plot
 ```
 
-## Error-free probability
+## Per-base and per-read error rate -- strategies
 
 ```{r}
-ggplot(reads, aes(x = p.noerr,
-                  linetype = Region,
-                  group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1 - ..y..), stat = "ecdf") +
+ggpubr::ggarrange(erate_plot, eexp_plot, ncol = 1, common.legend = TRUE,
+                  legend = "bottom", labels = "auto")
+```
+
+## Error-free probability --strategies
+
+```{r}
+dplyr::filter(reads, stat == "p.noerr", step == "raw") %>%
+  dplyr::arrange(dplyr::desc((value))) %>%
+  dplyr::group_by(`Read type`) %>%
+  dplyr::mutate(ecdf = cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = `Read type`,
+             group = `Read type`,
+             color = `Read type`)) +
+  geom_line() +
   scale_x_continuous(name = "Expectation of being error-free",
@@
                      oob = squish) +
-  scale_y_continuous(name = "Fraction passing")
+  scale_y_continuous(name = "Fraction passing") +
+  scale_color_read() +
+  scale_linetype_discrete()
 ```
 
-## Read length
+## Read length -- strategies
 
 ```{r}
-ggplot(reads, aes(x = length,
-                  linetype = Region,
-                  group = paste(Dataset, Region),
-                  color = Dataset)) +
-  geom_step(aes(y = 1- ..y..), stat = "ecdf") +
-  scale_x_log10(name = "Length (bp)") +
-  scale_y_continuous(name = "Fraction Greater than length")
+dplyr::filter(reads, stat == "length", step == "raw") %>%
+  dplyr::arrange(value) %>%
+  dplyr::group_by(`Read type`) %>%
+  dplyr::mutate(ecdf = 1 - cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = `Read type`,
+             group = `Read type`,
+             color = `Read type`)) +
+  geom_line() +
+  scale_x_log10(name = "Length (bp)",
+                limits = c(50, NA),
+                oob = squish) +
+  scale_y_continuous(name = "Fraction Greater than length") +
+  scale_color_viridis_d(option = "C", end = 0.9) +
+  scale_linetype_discrete()
+```
 
+## Read length density
+```{r}
+dplyr::filter(reads, stat == "erate", step == "raw") %>%
+  dplyr::arrange(value) %>%
+  dplyr::group_by(`Read type`) %>%
+  dplyr::mutate(ecdf = cumsum(nreads)/sum(nreads)) %>%
+  ggplot(aes(x = value,
+             y = ecdf,
+             linetype = `Read type`,
+             group = `Read type`,
+             color = `Read type`)) +
+  geom_line() +
+  scale_x_log10(name = "Length (bp)",
+                limits = c(50, NA),
+                oob = squish) +
+  scale_y_continuous(name = "count") +
+  scale_color_viridis_d(option = "C") +
+  scale_linetype_discrete()
 ```
@@
 reads %>%
-  group_by(Dataset, Plate, Region) %>%
-  summarize(`Expected Error-Free` = sum(p.noerr),
-            Total = n()) %>%
+  dplyr::filter(step %in% c("raw", "lsux"), stat == "p.noerr") %>%
+  group_by(dataset, read, region) %>%
+  summarize(`Expected Error-Free` = sum(value * nreads),
+            Total = sum(nreads)) %>%
   knitr::kable()
 ```
-
-```{r eval = FALSE}
-library(dada2)
-dada2::plotQualityProfile(unique(filter(reads, Tech == "PacBio")$files))
-```
-
-```{r eval = FALSE}
-counts <- readd(seq_counts) %>% 
-  tidyr::extract(col = "file",
-                     into = c("Seq.Run", "Plate", "Well","Region","Type"),
-                     regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)[fr]?-?([:alnum:]*)\\.([:alnum:]+).+") %>%
-  mutate(Type = ifelse(Type == "trim" & str_length(Region) > 0,
-                       "itsx",
-                       Type)) %>%
-  group_by(Seq.Run, Region, Type) %>%
-  summarize(reads = sum(reads)) %>%
-  spread(key = "Type", value = "reads") %>%
-  select(Seq.Run, trim, Region, itsx, qfilt) %>%
-  group_by(Seq.Run) %>%
-  mutate(trim = na.omit(trim)) %>%
-  filter(all(is.na(itsx)) | !is.na(itsx)) %>%
-  ungroup()
-
-demux.counts <- read_lines(file.path(data.dir, "demux.counts")) %>%
-  str_match("([^:]+): (\\d+) sequences ([^(]+)( \\(([FR]): (\\d+), ([FR]): (\\d+)\\))?\\.") %>%
-  magrittr::extract(,c(2,3,4,6,7,8, 9)) %>%
-  set_colnames(c("file", "reads", "type", "d1", "n1", "d2", "n2")) %>%
-  as_tibble() %>%
-  mutate_at(c("reads", "n1", "n2"), as.integer) %>%
-  mutate(fwd = case_when(!is.na(d1) & d1 == "F" ~ n1,
-                         !is.na(d2) & d2 == "F" ~ n2,
-                         TRUE ~ NA_integer_),
-         rev = case_when(!is.na(d1) & d1 == "R" ~ n1,
-                         !is.na(d2) & d2 == "R" ~ n2,
-                         TRUE ~ NA_integer_)) %>%
-  select(-(d1:n2)) %>%
-  mutate(dataset = map(file, str_extract, datasets$Dataset) %>%
-           map_chr(purrr::compose(dplyr::first, sort, unique)),
-         seq.run = map(file, str_extract, datasets$Seq.Run) %>%
-           map_chr(purrr::compose(dplyr::first, sort, unique)),
-         seq.plate = map(file, str_extract, paste0(datasets$Dataset, "-[0-9]{3}")) %>%
-           map_chr(purrr::compose(dplyr::first, sort, unique))) %>%
-  group_by(dataset, seq.run, seq.plate,type) %>%
-  summarize_at(c("reads", "fwd", "rev"), sum) %>%
-  ungroup() %>%
-  mutate(type = recode(type,
-                       "with at least one tag match" = "any.tag",
-                       "match at least 90% of tag length" = "tag.cover",
-                       "with at most 3 tag mismatches" = "tag.score",
-                       "with no more than one distinct tag in each direction" = "identifiable",
-                       "with tags in both directions" = "both.tags",
-                       "with tags appearing in correct order" = "tag.order")) %>%
-  gather(key = "dir", value = "reads", reads, fwd, rev) %>%
-  mutate_at("dir", str_replace, "^reads$", "") %>%
-  tidyr::unite("type", type, dir, sep = ".") %>%
-  spread(key = "type", value = "reads") %>%
-  set_names(str_replace(names(.), "\\.$", "")) %>%
-  keep(~!all(is.na(.))) %>%
-  select(dataset, starts_with("any.tag"), starts_with("tag.cover"),
-         starts_with("tag.score"), starts_with("identifiable"), both.tags, tag.order) %>%
-  left_join(select(counts, dataset, original, demux) %>%
-              filter(complete.cases(.)) %>%
-              unique()) %>%
-  select(dataset, original, any.tag:tag.order, demux) %>%
-  left_join(select(counts, dataset, region, ITSx, trim))
-
-dada.counts <-
-  demux.counts %>%
-  left_join(select(datasets, dataset = Dataset, Seq.Run)) %>%
-  mutate(file = glue("{data.dir}/{dataset}_{Seq.Run}.dada.seqtable.rds"),
-         file = ifelse(is.na(region),
-                       file,
-                       str_replace(file, "\\.dada", glue(".{region}.dada"))),
-         seq.tab = map(file, readRDS),
-         assigned = map_int(seq.tab, sum),
-         n_ASV = map_int(seq.tab, ncol),
-         file = str_replace(file, "\\.dada\\.seqtable\\.rds",
-                            ".dada.nochim.rds"),
-         nochim = map(file, readRDS),
-         nochim.reads = map_int(nochim, sum),
-         nochim.ASV = map_int(nochim, ncol)) %>%
-  select(-seq.tab, -file, -nochim)
-
-dada.counts
-```
-```{r eval = FALSE}
-s.pb.seqs <- readRDS(glue("{data.dir}/short-pacbio_pb_483.dada.nochim.rds")) %>%
-  colSums %>%
-  tibble(seq = names(.),
-         pb.reads = .)
-s.ion.seqs <- readRDS(glue("{data.dir}/short-ion_is_057.dada.nochim.rds")) %>%
-  colSums %>%
-  tibble(seq = names(.),
-         ion.reads = .)
-
-(crosstab <- full_join(s.pb.seqs, s.ion.seqs, by = "seq") %>%
-  mutate(len = str_length(seq)) %>%
-  replace_na(list(pb.reads = 0, ion.reads = 0))) %>%
-  ggplot(aes(x = ion.reads, y = pb.reads, color = len)) +
-  geom_point() +
-  scale_y_log10() +
-  scale_x_log10()
-```
-
-IonTorrent reads which match a PacBio ASV: ` filter(crosstab, pb.reads > 0) %$% sum(ion.reads)`
-IonTorrent reads which do NOT match PacBio: ` filter(crosstab, pb.reads == 0) %$% sum(ion.reads)`
-PacBio reads which match an IonTorrent ASV: ` filter(crosstab, ion.reads > 0) %$% sum(pb.reads)`
-PacBio reads which do NOT match IonTorrent: ` filter(crosstab, ion.reads == 0) %$% sum(pb.reads)`
-
-```{r eval = FALSE}
-crosstab %>% arrange(pb.reads, ion.reads) %>%
-  write.csv(file.path(data.dir, "pb_vs_ion_ASVs.csv"))
-```
-
-```{r eval = FALSE}
-demux.counts %>%
-  select(dataset, region) %>%
-  left_join(select(datasets, dataset = Dataset, Seq.Run)) %>%
-  mutate(file = glue("{data.dir}/{dataset}_{Seq.Run}.dada.taxonomy.rds"),
-         file = ifelse(is.na(region),
-                       file,
-                       str_replace(file, "\\.dada", glue(".{region}.dada")))) %>%
-  filter(file.exists(file)) %>%
-  mutate(tax = map(file, readRDS)) %>%
-  unnest %>%
-  filter(str_detect(Taxonomy, "Inocybe")) %>%
-  mutate(nreads = rowSums(select_if(., is.integer), na.rm = TRUE)) %>%
-  arrange(desc(nreads)) %>%
-  select(dataset, region, nreads, Taxonomy, seq) %>%
-  mutate(name = paste(dataset, region, substr(map_chr(seq, digest::digest, algo = "sha1"), 1, 8), sep = "_")) %$%
-  set_names(seq, name) %>%
-  DNAStringSet(use.names = TRUE) %>%
-  writeFasta(file.path(data.dir, "inocybe.fasta"))
-
-`short-ion_is_057.dada.nochim` %>%
-  colnames() %>%
-  set_names(paste0("seq", seq_along(.))) %>%
-  DNAStringSet(use.names = TRUE) %>%
-  writeFasta(file.path(data.dir, "short-ion.fasta"))
-```

### 2020-09-25 writing/transect_supplement.Rmd
add quality profile supplementary figure

@@
   theme(panel.spacing = unit(0, "mm"))
+```
+
+(ref:qual-check-s) Read quality profiles for different sequencing strategies
+
+(ref:qual-check) **(a)** per-base error rate; **(b)** expected number of errors per read.
+Statistics are based on quality scores assigned by technology-specific basecallers.
+**THIS NEEDS TO BE MENTIONED IN THE TEXT (addresses reviewer comment)**
+
+```{r qual-check, fig.cap = "(ref:qual-check-s). (ref:qual-check)", fig.scap = "(ref:qual-check-s)"}
+ggpubr::ggarrange(
+  readd(erate_plot),
+  readd(eexp_plot),
+  ncol = 1,
+  common.legend = TRUE,
+  legend = "bottom",
+  labels = "auto"
+)
 ```

### 2020-09-26 writing/transect_paper.Rmd
use ggplot instead of tikz to draw amplicon map, and put together subfigures using ggarrange instead of latex

@@
     suppress-bibliography: true
-    extra_dependencies: "subfig"
   bookdown::word_document2: 
@@
 
-(ref:region-asv) **(a)** Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).
-D1-3 (grey) represent the first three variable regions in LSU, while LSU1-4 (black) represent the conserved regions.
+(ref:region-asv) **(a)** Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions.
+D1-3 represent the first three variable regions in LSU, while LSU1-4 represent the conserved regions.
 Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
@@
 
-```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)", fig.subcap=rep("", 2), fig.ncol = 1, fig.align = "center"}
-knitr::include_graphics("rDNA_amplicons", auto_pdf = TRUE)
-readd(region_table) %>%
-  dplyr::filter(seq_run == "pb_500") %>%
-  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
-  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
-  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
-  geom_point() +
-  ggrepel::geom_text_repel() +
-  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
-  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1)) +
-  ylab("Unique ASVs")
+```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)", fig.align = "center"}
+ggpubr::ggarrange(
+  readd(amplicon_plot),
+  readd(length_diversity_plot),
+  ncol = 1,
+  labels = "auto",
+  heights = c(0.6, 1)
+)
 ```

### 2020-09-26 scripts/drake_plan_light.R

@@ -646,6 +646,32 @@ plan2 <- drake_plan(
   # for making the tables, figures, and in-line results in the paper.          #
   ##############################################################################
   
+  amplicon_plot =
+    readr::read_csv(file_in("writing/rDNA_amplicons.csv")) %>%
+    dplyr::mutate(xmid = (x + xend) / 2,
+                  label = tidyr::replace_na(label, "")) %>%
+    ggplot(aes(x = x, y = y, xend = xend, yend = yend, color = color,
+               linetype = linetype, size = size)) +
+    geom_segment(data = ~dplyr::filter(., !arrow)) +
+    geom_segment(
+      data = ~dplyr::filter(., arrow),
+      arrow = arrow(length = unit(0.1, "inches"))
+    ) +
+    geom_text(aes(x = xmid, y = y_label, label = label, color = labelcolor,
+                  size = size_label)) +
+    scale_linetype_identity() +
+    scale_color_identity() +
+    scale_size_identity() +
+    ylim(c(-19, 24)) +
+    coord_equal(ratio = 10) +
+    theme(axis.line = element_blank(),
+          panel.grid = element_blank(),
+          panel.border = element_blank(),
+          axis.text = element_blank(),
+          axis.ticks = element_blank(),
+          axis.title = element_blank()
+    ),
+  
   # list of all sequences by region and sequencing run, along with number of
   # reads and sequence length
   reads_table = target(
@@ -678,6 +704,28 @@ plan2 <- drake_plan(
       length_max = max(length),
       reads = sum(reads)
     ),
+  
+  # plot of ASV diversity vs. denoising success vs. length
+  length_diversity_plot = region_table %>%
+    dplyr::filter(seq_run == "pb_500") %>%
+    mutate_at(
+      "reads",
+      divide_by,
+      sum(filter(readcounts, seq_run == "pb_500", step == "demux")$nreads)
+    ) %>%
+    dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
+    ggplot(aes(reads, ASVs, color = length_med, label = region)) +
+    geom_point() +
+    ggrepel::geom_text_repel() +
+    scale_color_gradientn(
+      colors = c("blue", "red", "green1"),
+      limits = c(30, 2000),
+      breaks = c(50, 150, 400, 1000),
+      trans = "log10",
+      name = "median length"
+    ) +
[…]

### 2020-09-26 writing/transect_paper.Rmd
switch to cairo pdf in order to use unicode in figures

@@
 knitr::opts_chunk$set(
-  dev = "pdf",
+  dev = "cairo_pdf",
   # dev.args = list(pointsize=32),

### 2020-09-27 writing/transect_paper.Rmd
add species richness comparison figure and caption

@@
 
+(ref:alpha-compare-s) Comparison of alpha diversity between sequencing technologies
+
+(ref:alpha-compare) for ASVs **(a)** and OTUs **(b)**.
+Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
+All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.
+Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction.
+Blue line is a total least squares fit, with the given slope (and 95% confidence interval) and R^2^ value.
+Dashed diagonal line indicates a slope of 1.
+
+```{r alpha-compare, fig.scap = "(ref:alpha-compare-s)", fig.cap = "(ref:alpha-compare-s) (ref:alpha-compare)"}
+ggpubr::ggarrange(
+  remove_empty_facets(readd(alpha_plot_ASV)),
+  remove_empty_facets(readd(alpha_plot_OTU)),
+  ncol = 1,
+  labels = "auto"
+)
+```
+
 ## Taxonomic assignment

### 2020-09-27 writing/transect_paper.Rmd
clean up alpha diversity plot

@@
 
-```{r alpha-compare, fig.scap = "(ref:alpha-compare-s)", fig.cap = "(ref:alpha-compare-s) (ref:alpha-compare)"}
+```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s) (ref:alpha-compare)", fig.height=8}
 ggpubr::ggarrange(
@@
   ncol = 1,
-  labels = "auto"
+  labels = "auto",
+  label.x = 0.9
 )

### 2020-09-27 scripts/drake_plan_light.R

@@ -903,7 +903,7 @@ plan2 <- drake_plan(
       dplyr::mutate(
         slope_label = sprintf("slope==%.2f(%.2f-%.2f)",
                              value_x, `lower 0.95`, `upper 0.95`),
-        r2_label = sprintf("r^2==%.2f", r_squared)
+        r2_label = sprintf("R^2==%.2f", r_squared)
       ),
     transform = map(rarefy_data, .id = cluster)
   ),
@@ -914,27 +914,24 @@ plan2 <- drake_plan(
                   color = "gray50") +
       geom_point(shape = 1, alpha = 0.8) +
       geom_abline(
-        aes(slope = value_x, intercept = X.Intercept.),
+        aes(slope = value_x, intercept = `(Intercept)`),
         color = "blue",
-        alpha = 0.8,
         size = 1,
         data = demingfits
       ) +
       geom_text(
         aes(x = 0, y = 60, label = slope_label),
         hjust = 0,
-        size = 3,
+        size = 2.5,
         color = "blue",
-        alpha = 0.8,
         parse = TRUE,
         data = demingfits
       ) +
       geom_text(
         aes(x = 0, y = 52, label = r2_label),
         hjust = 0,
-        size = 3,
+        size = 2.5,
         color = "blue",
-        alpha = 0.8,
         parse = TRUE,
         data = demingfits
       ) +

### 2020-09-27 scripts/drake_plan_light.R
add accumulation plots

@@ -945,6 +945,34 @@ plan2 <- drake_plan(
     transform = map(demingfits, rarefy_data, cluster, .id = cluster)
   ),
 
+  accum_plot = target(
+    readd(sample_iNEXT) %>%
+      purrr::flatten() %>%
+      purrr::map_dfr(fortify, .id = "sample") %>%
+      tidyr::extract(
+        sample,
+        into = c("seq_run", "well"),
+        regex = "([ipS][sbH][-_][0-9]{3,4})_([A-H]1?[0-9])",
+        remove = FALSE
+      ) %>%
+      dplyr::left_join(datasets, by = "seq_run") %>%
+      dplyr::mutate(strategy = paste(tech, amplicon)) %>%
+      dplyr::filter(method != "extrapolated") %>%
+      ggplot(aes(x = x, y = y, color = strategy)) +
+      # lines for each sample
+      geom_line(aes(group = sample), alpha = 0.08) +
+      # small points to represent the actual observed read depth and richness
+      geom_point(data = ~dplyr::filter(., method == "observed"),
+                 alpha = 0.8, size = 1, shape = 1) +
+      xlab("reads") +
+      ylab(paste(cluster, "richness")) +
+      scale_color_strategy(name = NULL) +
+      scale_x_log10() +
+      scale_y_log10() +
+      theme_bw(),
+    transform = map(sample_iNEXT, cluster, .id = cluster)
+  ),
+
   ## Quality stats ----
 
   parsed_qstat = parse_qstat(qstats),

### 2020-09-30 writing/transect_paper.Rmd
add accumulation plots
work on reviewer comments

@@
 Different sequencing technologies have different capabilities in terms of sequencing depth and read length, as well as differing quality profiles and potential biases [@yang2013].
-The rapid development of new HTS technologies, as well as subsequent iterative improvements in sequencing chemistry and read capacity, means that the technologies used in metabarcoding studies, along with any associated biases, change rapidly.
+The rapid development of new HTS technologies, as well as subsequent iterative improvements in sequencing chemistry and read capacity, means that the technologies used in metabarcoding studies, along with any associated biases, change frequently.
 As an example, the first study using HTS metabarcoding of soil fungi was published in 2009 [@buee2009] using 454 Pyrosequencing; production of 454 sequencers was subsequently discontinued in 2015, and sales of reagents stopped in 2016 [@hollmer2013].
@@
 At each site, 25\ soil samples were collected along a single 24 m linear transect at intervals of 1\ m in May 2015.
-One third of the sample locations (3\ m spacing) were resampled one year later in June 2016.
+One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
 For each sample, any coarse organic debris <!--Do you have any definition?
@@
 An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such the LifeGuard preserved samples were excluded from our spatial analyses.
+Sequencing results for these 50 samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such the LifeGuard preserved samples were excluded from our spatial analyses.
 However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
@@
 An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
+
 Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
@@
 of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden) using one half of a lane.
+Platform-specific library preparation, including adapter ligation, was performed at the sequencing facilities according to their standard protocols.
 
@@
 The resulting sequences, as well as the paired-end Illumina sequences, were demultiplexed and sequencing primers were removed using `cutadapt` version 2.8 [@martin2011].
-Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible, and these reads were thus only analyzed as a pool.
+Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible, resulting in two samples sharing each tag.
+These reads were thus either analyzed as a pool, or comparisons were made to equivalently combined samples in the other datasets.
 For Ion Torrent and PacBio, reads were discarded if they did not have the appropriate primers on both ends.
@@
 
-<!-- R3: honestly this section is above my head and I can’t judge its accuracy. Hopefully another review is more bioinformatically inclined and can judge it better. -->
 Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
@@
 This refinement method is referred to as "Consensus".
-Additionally, a hybrid method, was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
+Additionally, a hybrid method was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
 
@@
 
+To assess the completeness of sequencing using each of the sequencing strategies, ASV and OTU accumulation curves were generated using the iNEXT package [citation].
+Additionally, average estimated ASV and OTU richness was calculated for each sample after rarefying to 100 sequences using 100 repetitions.
+ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies.
+For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the "deming" package [citation].
+For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.
+
 The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the R package `vegan` [@R-vegan].
 Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).
-<!-- R3: I would say, the effect of the sequencing strategies and read length on the fungal community compositional dissimilarity, as assessed by either UNIFRAC or Bray-Curtis, was determined by PERMANOVA (cite Marti Anderson) using the Adonis function in vegan.
-I’m confused because you definitely mentioned in the intro UNIfrac but in this section only mention Bray-Curtis.
-Also you don’t mention any sort of normalization for differing sequence reads per sample, and that’s not really acceptable.
-I recommend avg.dist function in vegan.
-It lets you make the Bray-Curtis dissimilarity matrix subsampling for an even sequencing depth 100x and then take the median and perform a square root transformation all in one line of code.
 
-Also, what no alpha diversity analysis?
-Beta-diversity analyses are generally pretty robust.
-I would be way more interested to see differences in alpha diversity by sequencing technology.
-Can you please report this (normalized for differences in sequencing depth)?
--->
 In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
@@
 
-DNA concentrations after extraction and PCR, as well as sequencing reads for PacBio and Illumina, are shown per sample in Figure \@ref(fig:sample-depth).
-Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads.
+Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads (Figure \@ref(fig:sample-depth)).
 Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
 
-The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table\ \@ref(tab:bioinfo).
+The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table\ \@ref(tab:bioinfo)).
 Sequencing with PacBio RS II yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
 Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
-<!-- R3: 20.7 M and 10.8M reads for only 25 samples is a lot of read depth! In the future you can go way lower than that to save money. I would multiplex in hundreds of samples for 20.8 M reads.-->
 Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
@@
 The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic %>% formatC(format = "e", digits = 2)`, $p < 2.2\times10^{-16}$),
-<!-- S3: 2 decimal points is sufficient-->
 but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
@@
 These reads were excluded from community analyses.
-<!-- S3: so to clarify, everything was PCR amplified with a primer that had barcodes and adaptor attached to either forward or reverse primer?
-I was having trouble following this during you PCR description section.-->
 
@@
 
-(ref:venn-s) Venn diagrams of shared ASVs and OTUs between different sequencing technologies
+(ref:venn-s) Shared richness and abundance of ASVs and OTUs between different sequencing technologies
 
-(ref:venn) Venn diagrams showing shared ITS2-based ASVs (*a*, *c*) and 97% OTUs (*b*, *d*) between different sequencing technologies from the same short amplicon library (*a*, *b*), and between long and short amplicon libraries (*c*, *d*).
-In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.
+(ref:venn) Shared richness and abundance of ITS2-based ASVs (*a*, *c*) and 97% OTUs (*b*, *d*) between different sequencing technologies from the same short amplicon library (*a*, *b*), and between long and short amplicon libraries (*c*, *d*).
+In each region, the ASV/OTU richness is given above, while the relative abundance of reads represented by these ASVs/OTUs in each sequencing strategy are shown below in the order PacBio/Illumina/Ion Torrent (*a*, *b*), or long/short (*c*, *d*).
 For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
-<!-- S2: The values for Ion Torrent are 1-3 orders of magnitude higher than other methods. This to me is an important finding that is not mentioned(!) -->
-<!-- s3: I recommend using “VennDiagram” package in R.
-It allows you to make Venn diagrams with circle sizes that change based on how many samples it represents.
-That is way easier to interpret than one with circles that are the same size regardless.
-I also don’t really understand what the decimals represent – what do you mean by the fraction of reads for each sequencing technology?
-For example in figure C, is 601 sequences 79% of the long amplicon reads?
-Please clarify.
--->
 
@@
 
-(ref:read-compare-s) Comparison between read numbers for different sequencing strategies
-
-(ref:read-compare) by ASV (**a**) and 97% OTU (**b**).
-ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
-Dashed line represents a constant ratio of read numbers.
-The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading.
-$R^2$ value displayed is for log-transformed non-zero read numbers.
-<!-- R2: Please provide axis labels e.g. panel (a) y-axis is "Count of ASVs" (?) -->
-<!-- R3: I’m not sure why this is useful, and I don’t think it warrants so much space in the document.
-It can be moved to the supplements.
-To me, a correlation of species richness by read technology might be interesting, or if you had some sort of treatment effect you were looking for and all of the methods gave you the same inferences of treatment effects, then that would be more interesting.
-I would prefer to see this replaced by correlation of alpha diversity.
-
-I also think species accumulation curves between sequencing technologies would be interesting to see. -->
-
-```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8}
-ggpubr::ggarrange(
-  read_comparison(
-    readd(multi_table),
-    readd(comparisons),
-    type = "ASV"
-  ),
-  read_comparison(
-    readd(multi_table),
-    readd(comparisons),
-    type = "OTU"
-  ),
-  ncol = 1,
-  labels = "auto",
-  label.x = 0.9
-)
-```
-
-The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Figure\ \@ref(fig:venn)).
-ASVs shared between all datasets represented `r paste(range(as.numeric(unlist(readd(venn_ASV)["1111", c("reads_frac_pb_500", "reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")]))*100), collapse = "--")`% of the reads for the long and short PacBio datasets, Illumina dataset, and Ion Torrent dataset, respectively.
-These fractions increased to `r paste(range(as.numeric(unlist(readd(venn_OTU)["1111", c("reads_frac_pb_500", "reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")]))*100), collapse = "--")`% when differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs.
-In particular, `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))` of reads in the PacBio, Illumina, and Ion Torrent short-amplicon datasets belonged to OTUs shared between all three datasets.
-In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs which were unique to that dataset, and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after OTU clustering.
-<!-- R1: Is this due to the much greater length causing a different ASV than a shorter ASV, even if they have an identical sequence over the same region?
-That seems to be the case, as the clustering did not really lower the difference between the unique long and short read ASVs at all (going from 21% to 20%).
-More depth of explanation in the discussion of this issue, maybe with a couple of examples of what is driving this, would be helpful.-->
-Complete tabulations of the number of ASVs and OTUs shared between the different sequencing strategies are shown in Supplementary Tables \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu), respectively.
+We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
+The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
+The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).
+When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% of reads, respectively (Figure\ \@ref(fig:venn)b).
+In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs, were found to be shared with other sequencing technologies upon OTU clustering.
+ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` of belonged to a unique OTU after clustering.
+In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).
 
@@
 
+```{r accum-ASV, fig.cap="ASV accumulation curves"}
+readd(accum_plot_ASV)
+```
+
+```{r accum-OTU, fig.cap="OTU accumulation curves"}
+readd(accum_plot_OTU)
+```
+
 ## Taxonomic assignment
 
-For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be identified than rare ASVs (Figure\ \@ref(fig:tax-chart)).
-<!-- R3: do you mean taxonomically identified? -->
+For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be taxonomically identified than rare ASVs (Figure\ \@ref(fig:tax-chart)).
 A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
@@
 These two methods are plotted in each column to compare with results for the individual databases.
-<!-- R2: Overall patterns are the same thus I suggest presenting only reads or ASVs but not both. Y-axis labels are needed as in Fig. 2. It is interesting that it is almost alinear decline-->
-<!-- R3: the barplots are not big enough to readily distinguish. Perhaps you can make the ASVs vs Reads vertical rather than horizonal and horizontally enlarge your image so it’s easier to read.-->
 
@@
 According to the PHYLOTAX assignments, Fungi represented `r percent(readd(fungi_asv)["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(readd(fungi_read)["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(readd(fungi_asv)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(fungi_read)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
-Measured fungal community composition at the class level varied significantly between amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
+Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
 The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm) and \@ref(fig:ppcoa)).
-<!-- R1: Any thoughts on why this clade in particular was well represented?
-For example, do you think this a primer bias issue or is there some strange secondary structure in the Sordariomycetes rRNA operon that results in their removal during bioinformatic filtering?
--->
 
@@
 
-The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0\ m, time lag of 1\ year), for both the total fungal and ECM fungal communities in the long amplicon library.
-For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
-<!-- R3: what does this mean?-->
-This puzzling negative correlation was significant in all correlograms based on short amplicon sequencing irrespective of technology.
+The Bray-Curtis metric showed positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0\ m, time lag of 1\ year), for both the total fungal and ECM fungal communities, although this result did not reach statistical significance for all sequencing strategies.
+This spatiotemporal correlation did not extend to a range of 1\ m, and in fact correlation was negative at a time lag of 1\ year and distance of 1\ m, indicating that samples collected 1\ m apart in different years were more different than randomly selected pairs of samples.
+This puzzling negative correlation reached statistical significance in the PacBio short amplicon dataset, and will be addressed further in the discussion.
 
@@
 
-The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranged in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
+The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
 <!-- R3: please re-write to better explain what you mean. Do you mean that communities generally experience full compositional turnover at 3m?-->
@@
 
-Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])`.
-Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
+Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% ITS2 similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])`, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
+ITS2 sequences extracted from the long amplicon dataset also included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
 Taxonomic assignments revealed that the majority of these non-shared sequences fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Figure\ \@ref(fig:short-its)).

### 2020-09-30 writing/transect_supplement.Rmd
work on reviewer comments

@@
 readd(region_length_fig)
+```
+
+(ref:read-compare-s) Comparison between read numbers for different sequencing strategies
+
+(ref:read-compare) by ASV (**a**) and 97% OTU (**b**).
+ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
+Dashed line represents a constant ratio of read numbers.
+The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading.
+$R^2$ value displayed is for log-transformed non-zero read numbers.
+
+```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8}
+ggpubr::ggarrange(
+  read_comparison(
+    readd(multi_table),
+    readd(comparisons),
+    type = "ASV"
+  ),
+  read_comparison(
+    readd(multi_table),
+    readd(comparisons),
+    type = "OTU"
+  ),
+  ncol = 1,
+  labels = "auto",
+  label.x = 0.9
+)
 ```

### 2020-10-07 writing/transect_paper.Rmd
modify taxonomy assignment chart to make it more readable
edits to respond to reviewer comments.  All addressed except supplementary figures.  diff latex is not rendering.

@@
 The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
-Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and have been suggested to be less dataset specific [@callahan2017].
+Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated [@glassman2018;@botnen2018].
+ASVs have been suggested to be less dataset specific that cluster-based OTUs [@callahan2017].
 Support for PacBio has recently been added to DADA2 [@callahan2019], but its application requires greater sequencing depth for longer reads, especially in high diversity samples.
@@
 Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
+Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
+Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
+It varies between different ecosystems and taxonomic groups.
+Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
+distance [@legendre2012].
+
 We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
@@
 
-For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
-<!-- R3: FYI, there is already a publication that compares OTU and ASV methods – Glassman and Martiny mSphere 2018 – and they result in the same inferences.
-Botnen et al Molecular Ecology Resources 2018 might also be worth citing.-->
+To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
 
@@
 
-Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
-Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
-It varies between different ecosystems and taxonomic groups.
-Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
-distance [@legendre2012].
-
-Ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric [@bray1957, for both long and short amplicons] and the phylogenetically based weighted UniFrac metric [@lozupone2005;@lozupone2007, for only long amplicons] in `phyloseq` version `r packageVersion("phyloseq")`.
+To estimate turnover scale, ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric [@bray1957, for both long and short amplicons] and the phylogenetically based weighted UniFrac metric [@lozupone2005;@lozupone2007, for only long amplicons] in `phyloseq` version `r packageVersion("phyloseq")`.
 Each of these distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
@@
 Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
-Demultiplexing, primer trimming, and quality filtering reduced these totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
+PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads (<= 1 expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
+Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \@ref(fig:qual-check)a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \@ref(fig:qual-check)b).
+Demultiplexing, primer trimming, and quality filtering reduced the read totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
 Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
-In contrast, extraction of only the ITS2 region before quality filtering resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-readd(filtercounts_ITS2)['pb_483']/readd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-readd(filtercounts_ITS2)['is_057']/readd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
+In contrast, extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-readd(filtercounts_ITS2)['pb_483']/readd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-readd(filtercounts_ITS2)['is_057']/readd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
 This represented greater loss of PacBio short amplicon reads, but less loss of PacBio long amplicon reads and Ion Torrent reads.
@@
 
-```{r tax-chart, fig.cap = "(ref:tax-chart)", fig.scap = "(ref:tax-chart-s)"}
+```{r tax-chart, fig.cap = "(ref:tax-chart)", fig.scap = "(ref:tax-chart-s)", out.width = "100%"}
 readd(tax_chart_plot)
@@
 
-The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
-<!-- R3: please re-write to better explain what you mean. Do you mean that communities generally experience full compositional turnover at 3m?-->
+The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from 13-32m for the total fungal community and 13-41m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit).
+However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies, so no strong conclusion of variability between methods can be drawn.
 Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
@@
 
-<!-- R2: The Discussion section is mostly descriptive, largely repeating the Results (e.g., lines 569-574; 578-581;589-593; 596-598; 600-606) with very little real discussion. -->
-
 ## Reconstruction of long amplicons from denoised subregions
 
-ASV recovery for long amplicons using DADA2 was dramatically improved (`r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of reads) by denoising homologous subregions independently using the `LSUx` and `tzara` packages.
-<!-- Why does the difference range from 12-76% This is a wide range.
-I would like to know what the authors think is the reason for this.
-Under what circumstances might we expect 12% and what 76% and why?
-What is the contribution beyond that of other studies that separated long reads into constituent genes regions and examined these separately?
--->
+Sequencing depth in the long amplicon PacBio dataset was not sufficient to successfully denoise using standard protocols, given the amplicon length and diversity of the samples.
+ASV recovery for long amplicons using DADA2 was dramatically improved from `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of reads by denoising homologous subregions independently using our new `LSUx` and `tzara` packages.
+LSUx was inspired by ITSx [@bengtsson-palme2013], which extracts the the ITS regions from the surrounding SSU, 5.8S, and LSU.
+However, we found that the resulting LSU segment was still too long for successful denoising (Figure \@ref(fig:regions)), necessitating more fine-grained extraction of domains.
 Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
@@
 Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(readd(bioinf_table)["ITS2","PacBio_Short_reads"]/readd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(readd(bioinf_table)["ITS2","Illumina_Short_reads"]/readd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(readd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/readd(rawcounts)["is_057"], format = "f", digits = 0)`),
-the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.
-<!--R3: this could be visualized with a species accumulation curve-->
+the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered, and were much closer to saturation of their respective species accumulation curves (Figure \ref(fig:accum-ASV)).
+OTU read counts were strongly correlated between technologies (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon == y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
+and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
+which we interpret as supporting the validity of abundance-based beta diversity measures in soil fungi (Figure \@ref(fig:read-compare)b).
 
-DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
-The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].
-<!-- R2: I understand that dada2 may not work as well, but why not?-->
+DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies  and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
+Ion Torrent generated more ASV richness than PacBio or Illumina from the common short amplicon library (Figure\ \@ref(fig:venn)), even after rarefication to equal sampling depth ().
+In contrast PacBio and Illumina generated very similar ASV diversity estimates after rarefication (Figure \ref(fig:alpha-compare)), even though the original sequencing depth for Illumina was two orders of magnitude higher than that for PacBio.
+This suggests that the ASV diversity is artificially inflated in the Ion Torrent dataset.
+This may be a result of the lower fraction of very high quality reads (Figure \ref(fig:qual-check)).
+Although we used options for DADA2 intended to improve performance on technologies with high rates of homopolymer indel errors [@callahan2020b], it is possible that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed [@callahan2016].
 
@@
 Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
-Additionally, several smaller groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
-<!-- R3: what do you mean by smaller groups? Having fewer genera per class?-->
-These differences may be due to primer mismatches in these taxa.
-<!-- R1: Can this be quantified?
-For example, do we know how well for example LR5 selects for fungi relative to alveolates based on priming site matches?
-Additionally, were the short reads more common on one side of the “long read” than the other?
-For example, if all the short reads are common on the LR5 end of the long read, that might suggest using a different LSU primer would minimize the inclusion of alveolates and other non-target taxa.-->
-<!-- R2: Primer mismatches are mentioned here, but only in relation to a few groups and as a final point in this section.
-But these are established primers whose biases are probably known and published.
-Do you results make sense?
-I would also need to see some discussion on how this bias might affect the study as a whole, because different platforms/reads used different primers.
--->
+Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
+These discrepancies are probably due to primer mismatches to the relevant groups.
+The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae (@tedersoo2015).
+ITS1 and LR5 match a much broader range of fungal and other eukaryote groups (@tedersoo2015).
+The alternate LR5-F primer (@tedersoo2008) would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
+We assert that more complete coverage of ECM fungal groups with high rDNA variability such as *Tulasnella* (and ideally other *Cantharellales*) is worth the read-depth spent on non-target groups.
 
@@
 
-<!-- R2: This section is weak and could be removed. -->
-The RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences.
-The Warcup database placed notably fewer sequences at all taxonomic levels for all datasets, probably in part due to the fact that only fungal sequences are included.
-However, even with this considered, IDTAXA performed very poorly with the Warcup database, placing <25% of ASVs to kingdom in all datasets.
-<!-- R2: why is this the case?-->
-IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores [@murali2018a].
-<!-- "This is expected" is not a very interesting conclusion in this case.
-I do not see how any other results could occur.
-Less conflict at lower taxonomic levels would only be a reflection of different taxonomies among the databases, and would not be anything biological.
--->
+Assignment of ecological function to environmental fungal sequences is dependent on accurate taxonomic identification, especially at the genus level or below (@nguyen2016funguild).
+However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
+Although the RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \@ref(fig:tax-chart)).
+This is probably due to two factors.
+First, the Warcup database does not include any non-fungi, so it cannot place any non-fungal sequences.
+Second, due to its low-density coverage of the fungal kingdom (18\ 000 vs.
+800\ 000 for Unite), it is likely that many ITS sequences, especially from uncultured tropical soil fungi, have no close match in the Warcup database, and so cannot be placed.
+The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
+Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, particularly the Warcup database, where it places <25% of ASVs even to kingdom.
+This is expected given its more conservative assignment of confidence scores [@murali2018a].
 
 @gdanetz2017 showed that a majority-rule consensus of three assignment algorithms can improve the fraction of sequences assigned as well as decrease the false assignment rate.
-<!-- R1: See my earlier note about Palmer et al. (2018) showing a similar result. -->
 Strict consensus rejects assignments whenever there is conflict between methods and should therefore provide more conservative taxonomic assignments than majority-rule consensus.
-Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications.
+AMPtk [@palmer2018] uses a strict consensus taxonomy between UTAX and SINTAX as an alternative when an initial BLAST search failed to give a hit with at least 97% sequence identity, but did not present results assessing the results of this approach.
+Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications (Figure \@ref(fig:tax-chart)).
 This suggests that different assignment algorithms and databases bring mostly complementary, non-contradictory information at higher taxonomic levels.
@@
 
-For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
+For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses phylogenetic relationships to resolve these disagreements in a principled manner.
 For instance, `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
@@
 
-<!-- R2: This section is too descriptive, and would be greatly improved if reasons were given, even multiple possible reasons that lead to hypotheses.
-It appear that the only conclusion offered (UNIFRAC would be more suited to larger scales) could have been made without these data.
--->
-<!-- R3: the Unifrac results should be reported after the Bray-Curtis results, which are more interesting. They could also be condensed.-->
+Mantel correlograms based on the Bray-Curtis dissimilarity (Figure\ \@ref(fig:correlog)) revealed spatial autocorrelation in the soil fungal community at distance classes $\le 3$ m for both Illumina and PacBio using long and short amplicons, and in the ECM fungal community at distance classes $\le 2$ m for Illumina and PacBio long amplicons, and $\le 1$ m for the PacBio short amplicons.
+These results are similar to autocorrelation ranges found in previous work based on ECM root tips in temperate forests [@lilleskov2004; @pickles2012a].
+@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites using Sanger sequencing.
+Similarly, @pickles2012a found autocorrelation at distances <3.4\ m based on T-RFLP analysis.
+Previous work in Miombo woodland, a similar ecosystem to the Soudanian woodland in this study, found autocorrelation at ranges <10\ m using Sanger sequencing of ECM root tips [@tedersoo2011], which was their smallest distance class.
+
+Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+The 95% confidence interval was substantially wider than this variation, generally covering a range of `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+All of these values are smaller than the 65\ m reported by @Bahram2013, also based on distance-decay curves from a similar ECM woodland habitat in Benin, but based on Sanger sequencing of ECM root tips rather than HTS metabarcoding of bulk soil.
+We speculate that this discrepancy is due to an increased ability to capture spatially variable rare species using HTS.
+For the short amplicon dataset, PacBio showed a spatial turnover range more than twice as long as showed by Illumina (Table \@ref(tab:variofit)) for both the total fungi and ECM communities, with wide confidence intervals.
+It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth in the PacBio short amplicon dataset.
+The long amplicon PacBio dataset, with more than twice the read depth of the short amplicon PacBio dataset, gave spatial turnover distance results much closer to those from Illumina.
+This is consistent with our speculation that the longer spatial turnover range found by @Bahram2013 is related to sequence sampling depth.
+
+Year-to-year correlation was found for both the total fungal and ECM communities in the long amplicon dataset (Figure \@ref(fig:correlog)).
+The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
+This corresponds to a space-for-time substitution rate of 6.15\ m/year for the total fungal community and 3.83\ m/year for the ECM community.
+In a recent study, @kivlin2020 reported a space-for-time substitution rate of `r format(1.5/6.8 * 365, digits = 2)`\ m/year (reported as 6.8\ d / 1.5\ m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
+However, comparison is obscured by different spatial and temporal sampling scales between the two studies.
+Year-to-year variation in ECM fungal communities, which we sampled, has been shown to be less than intra-annual variation [@bahram2015], as sampled by @kivlin2020.
+
+Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
+The general shape of the correlograms were similar for long and short amplicon datasets.
+We hypothesize that this pattern may result from a superposition of positively and negatively correlated temporal processes operating at different spatial scales.
+As an example, the positive year-to-year correlation at 0\ m distance could be due to one species or set of species which strongly persist between years, but are primarily confined to patch sizes $<1$\ m,
+while the negative year-to-year correlation could be due to a different species or set of species which deplete nutrients within a $>1$\ m patch over the course of a year, and are likely to be absent from the same patch in the following year.
 
@@
 The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m, indicating little evidence of spatial structure.
-This is probably due to the majority of weighted branch length in the community being between the Pezizomycotina and Agaricomycetes (Figure\ \@ref(fig:taxa)), which are both well represented in the majority of samples.
+This is probably because the majority of species turnover in this system, especially among ECM fungi, is between closely related species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
+This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac as compared to Bray-Curtis in Figure \@ref(fig:spatio)
 UniFrac would be more suited at larger spatial scales and/or larger ecological gradients.
-<!-- R1: A second way to try to understand this result is to see if the results differ based on weighting versus not weighting.
-Similarly, it would be nice to see is the distance-decay patterns are robust to abundance (read counts) versus incidence (pres/abs) based analyses.
-There are some in the fungal metabarcoding world who believe all high throughput sequencing (HTS) data should only be analyzed using presence absence data due to ITS copy number differences and other known issues (primer biases, etc).
-I don’t agree and I think the best way to push back on that critique is to show that the results of abundance- and incidence-based data classifications often yield similar patterns.
-The results of these suggested additional analyses could easily be incorporated as extra panels in Figure 5.
--->
-
-Mantel correlograms based on the Bray-Curtis dissimilarity (Figure\ \@ref(fig:correlog)) revealed spatial autocorrelation in the soil fungal community at distance classes $\le 3$ m for both Illumina and PacBio using long and short amplicons, and in the ECM fungal community at distance classes $\le 2$ m for Illumina and PacBio long amplicons, and $\le 1$ m for the PacBio short amplicons.
-These results are similar to autocorrelation ranges found in previous work based on ECM root tips in temperate forests [@lilleskov2004; @pickles2012a].
-@lilleskov2004 found autocorrelation only at ranges <2.6\ m at most sites using Sanger sequencing.
-Similarly, @pickles2012a found autocorrelation at distances <3.4\ m based on T-RFLP analysis.
-However, previous work in Miombo woodland, a similar ecosystem to the Soudanian woodland in this study, found autocorrelation at ranges <10\ m using Sanger sequencing of ECM root tips [@tedersoo2011], which was their smallest distance class.
-<!-- R3: I would not start this line with however, because it agrees with the sentences before if <10 m is the smallest distance class-->
-
-Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
-The 95% confidence interval was substantially wider than this variation, generally covering a range of `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
-All of these values are smaller than the 65\ m reported by @Bahram2013, also based on distance-decay curves from an ECM woodland habitat in Benin.
-<!-- R1: Is the habitat presented in Bahram et al.
-(2013) in Benin differ in some fundamentally important way from the other in this study (less disturbance, higher ECM tree basal area?It would be nice to try to better understand these differences, even if it requires a bit of speculation by the authors.-->
-
-The PacBio short amplicon dataset shows a longer spatial range, of 
-`r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "All Fungi") %$% range(estimate) %>% format(digits = 0) %>% unique() %>% paste(collapse = "--")`\ m for the total fungal community and
-`r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range", guild == "ECM") %$% range(estimate) %>% format(digits = 0) %>% unique() %>% paste(collapse = "--")`\ m for the ECM community,
-<!-- R1: 38-38m – is this a typo? FIXED-->
-<!-- R3: is 38-38m a typo?-->
-in both cases with wide confidence intervals spanning `r readd(variofit_table) %>% filter(amplicon == "Short", tech == "PacBio", metric == "Bray-Curtis", term == "range") %$% range(c(conf.high, conf.low), na.rm = TRUE) %>% unique() %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
-It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth.
-<!-- R1: Could this explanation be tested more formally by rarefying the non-ECM dataset down to the level of the ECM dataset and seeing if the confidence intervals and range estimates change?
-It would be nice to separate likely methodological artifact from potential differences in biology (as ECM communities likely have many more basids, so less asexual propagule production).
--->
-
-The Bray-Curtis Mantel correlogram for both the total fungal and ECM communities from the long amplicon dataset show a significant positive correlation at 0\ m and 1\ year.
-<!-- R3: This would be much more engaging to read, here and in other places, if you reword to report findings rather than stats “There was strong year to year correlation for both the total and ECM fungal communities (cite figure or stats)” -->
-The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
-Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
-The general shape of the correlograms were similar for long and short amplicon datasets.
-We hypothesize that two different processes may be at work with differing spatiotemporal scales, whose superposition result in this pattern.
-<!-- R2: This is one example of hypothesis formation, although these are very vague and could be more clearly presented - which scales and how would the different processes lead to this result? -->
 
@@
 The choice of amplicon and sequencing technology did not seem to affect the results of the spatial analysis, provided sufficient sequencing depth.
-However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments.
-<!-- R2: I don't see this as a valid conclusion, because this was not attempted with short reads. In principle it could be done. -->
+However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments using our new tool PHYLOTAX.
 DADA2 ASV yield was initially poor for long amplicons, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
-Together these approaches provide a hybrid approach using long-read sequencing to acquire long amplicon sequences for the local species pool, and cost-effective short-read sequencing to provide high sampling depth and sample number.
+Together these approaches provide a hybrid approach using long-read sequencing to acquire long amplicon sequences for the local species pool in order to improve taxonomic assignments, and cost-effective short-read sequencing to provide high sampling depth and sample number.
 
@@
 Laboratory work including PCR and library pooling was performed by Dr. Ylva Strid.
-<!-- R3: how come Dr Ylva Strid is not a co-author on the paper? Seems like she contributed substantially.-->
 

### 2020-10-07 writing/transect_supplement.Rmd
edits to respond to reviewer comments.  All addressed except supplementary figures.  diff latex is not rendering.

@@
 
-(ref:qual-check) **(a)** per-base error rate; **(b)** expected number of errors per read.
+(ref:qual-check) **(a)** per-base error rate; **(b)** expected number of errors per read, with threshold of 3 expected errors marked as a dashed vertical line.
 Statistics are based on quality scores assigned by technology-specific basecallers.
-**THIS NEEDS TO BE MENTIONED IN THE TEXT (addresses reviewer comment)**
 

### 2020-10-07 scripts/drake_plan_light.R

@@ -1158,6 +1158,7 @@ plan2 <- drake_plan(
 
   vennplot_OTU = vennplot_data(venn_OTU, OTUs),
 
+  ## read count comparisons ----
   big_table = {
     out <- bind_rows(
       mutate(asv_table, type = "ASV"),
@@ -1215,6 +1216,7 @@ plan2 <- drake_plan(
       label = paste("R^2 ==", formatC(r.squared, format = "f", digits = 2))
     ),
 
+  ## Taxonomy charts and heat trees ----
   taxon_reads = {
     out <-
       # Get the number of reads per sequencing run for each ASV
@@ -1448,6 +1450,7 @@ plan2 <- drake_plan(
           legend.position = "bottom",
           strip.background = element_blank()),
 
+  ## Stats for text about positive and negative controls ----
   agaricus =
     taxon_table %>%
     filter(rank == "genus", taxon == "Agaricus", n_diff == 1) %$%
@@ -1570,6 +1573,7 @@ plan2 <- drake_plan(
     apply(MARGIN = 1, order, decreasing = TRUE) %>%
     apply(MARGIN = 2, match, x = 1),
 
+  ## variogram plots ----
   variog_data = target({
     rlang::quo(list(correlog, variog, variogST, variofit2, variofitST2))
     ignore(plan2) %>%
@@ -3668,7 +3672,7 @@ make(
   cache = cache,
   parallelism = parallelism,
   jobs = njobs,
-  targets = c("article_pdf", "supplement_pdf"),
+  targets = c("article_pdf", "supplement_pdf", "article_diff_pdf", "supplement_diff_pdf"),
   lazy_load = FALSE,
   memory_strategy = "autoclean",
   garbage_collection = TRUE,

### 2020-10-08 writing/transect_supplement.Rmd
improve read quality plot

@@
 
-```{r qual-check, fig.cap = "(ref:qual-check-s). (ref:qual-check)", fig.scap = "(ref:qual-check-s)"}
+```{r qual-check, fig.cap = "(ref:qual-check-s). (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=6}
 ggpubr::ggarrange(
@@
   readd(eexp_plot),
+  readd(eexp_region_plot),
   ncol = 1,
-  common.legend = TRUE,
-  legend = "bottom",
+  # common.legend = TRUE,
+  # legend = "bottom",
   labels = "auto"

### 2020-10-08 writing/transect_paper.Rmd
switch to xelatex, as required by thesis printing.

@@
     toc: false
-    latex_engine: lualatex
+    latex_engine: xelatex
     includes:

### 2020-10-08 scripts/drake_plan_light.R

@@ -3207,7 +3207,7 @@ plan2 <- drake_plan(
       file_in(!!file.path("writing", "transect_supplement.tex"))
       file_in(!!file.path("writing", "transect_paper.tex"))
       file_out(!!file.path("writing", "transect_supplement.pdf"))
-      tinytex::lualatex("transect_supplement.tex", bib_engine = "biber")
+      tinytex::xelatex("transect_supplement.tex", bib_engine = "biber")
     }
   ),
 
@@ -3217,7 +3217,7 @@ plan2 <- drake_plan(
       file_in(!!file.path("writing", "transect_supplement_diff.tex"))
       file_in(!!file.path("writing", "transect_paper_diff.tex"))
       file_out(!!file.path("writing", "transect_supplement_diff.pdf"))
-      tinytex::lualatex("transect_supplement_diff.tex", bib_engine = "biber")
+      tinytex::xelatex("transect_supplement_diff.tex", bib_engine = "biber")
     }
   ),
 
@@ -3227,7 +3227,7 @@ plan2 <- drake_plan(
       file_in(!!file.path("writing", "transect_paper.tex"))
       file_in(!!file.path("writing", "transect_supplement.tex"))
       file_out(!!file.path("writing", "transect_paper.pdf"))
-      tinytex::lualatex("transect_paper.tex", bib_engine = "biber")
+      tinytex::xelatex("transect_paper.tex", bib_engine = "biber")
     }
   ),
 
@@ -3237,7 +3237,7 @@ plan2 <- drake_plan(
       file_in(!!file.path("writing", "transect_paper_diff.tex"))
       file_in(!!file.path("writing", "transect_supplement_diff.tex"))
       file_out(!!file.path("writing", "transect_paper_diff.pdf"))
-      tinytex::lualatex("transect_paper_diff.tex", bib_engine = "biber")
+      tinytex::xelatex("transect_paper_diff.tex", bib_engine = "biber")
     }
   ),
 

### 2020-10-08 writing/transect_supplement.Rmd
caption for quality check figure

@@
 
-(ref:qual-check-s) Read quality profiles for different sequencing strategies
+(ref:qual-check-s) Read quality profiles
 
-(ref:qual-check) **(a)** per-base error rate; **(b)** expected number of errors per read, with threshold of 3 expected errors marked as a dashed vertical line.
-Statistics are based on quality scores assigned by technology-specific basecallers.
+(ref:qual-check) based on quality scores assigned by technology-specific basecallers.
+**(a)** per-base error rate for different sequencing strategies; **(b)** expected number of errors per read for different sequencing strategies; **(c)** expected number of errors per read for regions extracted from PacBio long amplicon reads.
+In **(a)** and **(b)**, R1 and R2 denote the forward and reverse reads, respectively, from Illumina paired-end sequencing.
+PacBio short, Ion Torrent, and Illumina datasets were all sequenced from the same short amplicon library.
+In **(b)** and **(c)**, quality filtering threshold of 3 expected errors marked as a dashed vertical line.
 
-```{r qual-check, fig.cap = "(ref:qual-check-s). (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=6}
+```{r qual-check, fig.cap = "(ref:qual-check-s), (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=7}
 ggpubr::ggarrange(

### 2020-10-08 writing/transect_supplement.Rmd
fix caption for phylotax plot

@@
 No taxon is assigned at node\ 2, because the assignments at C and F are inconsistent.
-Node\ 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
-Node\ 4 (and thus also node\ 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
-<!-- Is there an error in the legend?
-It appears that node 3 is assigned to Taxa 1 based on box color coding and node 4 to Taxon 2 based on different box color coding.
-In the legend, it states that node 3 is assigned to Taxa 1, but there are three matches to Taxon 2 at that node and Taxon 2 is presented in Salmon color consistent with the box color for node 3.
-As such, I think the authors meant to designate node 3 as being assigned to Taxon 2.
-If this is wrong, then more details in the legend explaining why the greater presence of matches to Tax2 yet assignment to Tax1 is needed.
--->
+Node\ 3 is assigned to Tax2 because this is consistent with at least one of the assignments for both B and C.
+Node\ 4 (and thus also node\ 5) is assigned to Tax1 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
 

### 2020-10-08 writing/transect_supplement.Rmd
fix axis labels and captions for PCoA

@@
 
-(ref:buffer-pcoa-s) Unconstrained Principal Coordinate Analysis ordination of samples preserved with different buffers
+(ref:buffer-pcoa-s) Unconstrained Principal Coordinate Analysis (PCoA) ordination of samples preserved with different buffers
 
-(ref:buffer-pcoa) For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.
-<!-- R3: your caption isn’t in line with your axes.
-Your axes say MDS1 and MDS2 – so it looks like a NMDS to me.
-But you caption says PCA.
-There is also insufficient information to jusdge this figure - No information on if reads were normalized or not, or what beta diversity metric this is based on, etc-->
+(ref:buffer-pcoa) based on Bray-Curtis dissimilarities of relative read abundances.
+For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.
 
-```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s). (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
-plot(readd(buffer_asv_pcoa), type = "none")
+```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
+plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2")
 points(
@@
 
-(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding approaches
+(ref:ppcoa-short) Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding strategies
 
 (ref:ppcoa) for all fungi (**a**) and ECM fungi (**b**).
-Spatio-temporal variation (i.e., variation between different soil samples) has been partialled out.
-Only samples with at least 100 reads for all three approaches are included.
+Ordination based on the residual variation after spatio-temporal variation (i.e., variation between different soil samples) has been accounted for.
+Only samples with at least 100 reads for all three strategies are included.
 ASVs are clustered by taxonomic class for all fungi, and by family for ECM fungi.

### 2020-10-08 writing/transect_supplement.Rmd
add caching in supplement to speed up rendering

@@
 if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))
+
+gethash <- function(...) purrr::map_chr(list(...), drake_cache()$get_hash)
 ```
@@
 ```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
-plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2")
+plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2", cache = TRUE, extra.cache = gethash("buffer_asv_pcoa", "buffer_asv_table"))
 points(
@@
 
-```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)"}
+```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)", cache = TRUE, extra.cache = gethash("buffer_compare_long")}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
-```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
+```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)", cache = TRUE, extra.cache = gethash("buffer_compare_short"), more.cache=oldtheme}
 readd(buffer_compare_short)
@@
 
-```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)"}
+```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)", cache = TRUE, extra.cache = gethash("phylotax_fig")}
 readd(phylotax_fig)
@@
 
-```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)"}
+```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)", cache = TRUE, extra.cache = gethash("conc_plot", "reads_plot")}
 ggpubr::ggarrange(readd(conc_plot), readd(reads_plot), ncol = 2, common.legend = TRUE) +
@@
 
-```{r qual-check, fig.cap = "(ref:qual-check-s), (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=7}
+```{r qual-check, fig.cap = "(ref:qual-check-s), (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=7, cache = TRUE, extra.cache = gethash("erate_plot", "eexp_plot", "eexp_region_plot")}
 ggpubr::ggarrange(
@@
   ncol = 1,
-  # common.legend = TRUE,
-  # legend = "bottom",
   labels = "auto"
@@
 
-```{r full-length, fig.cap = "(ref:full-length-s). (ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6}
+```{r full-length, fig.cap = "(ref:full-length-s).
+(ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6, cache = TRUE, extra.cache = gethash("full_dens_short", "full_ecdf_short", "full_ecdf_short", "full_ecdf_long")}
 ggpubr::ggarrange(
@@
 
-```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6}
+```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6, cache = TRUE, extra.cache = gethash("region_length_fig")}
 readd(region_length_fig)
@@
 
-```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8}
+```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8, cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
 ggpubr::ggarrange(
@@
 
-```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8}
+```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8, cache=TRUE, extra.cache=gethash("heattree")}
 theme <- theme_update(panel.border = element_blank())
@@
 
-```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7}
+```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
   readd(taxon_reads) %>%
@@
 
-```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
+```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)", cache=TRUE, extra.cache=gethash("tech_class_pcoa_plot", "tech_ecm_fam_pcoa_plot")}
 ggpubr::ggarrange(
@@
 
-```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)"}
+```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)", cache=TRUE, extra.cache=gethash(keep(light_plan$target, startsWith, "correlog"))}
 light_plan %>%
@@
 
-```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6}
+```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6, cache=TRUE, extra.cache=gethash("its2_dens", "its2_ecdf")}
 ggpubr::ggarrange(
@@
 
-```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s). (ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%"}
+```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s).
+(ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%", cache=TRUE, extra.cache=gethash("heattree_length_compare")}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
-```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7}
+```{r short-its, fig.cap = "(ref:short-its-s).
+(ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
 readd(big_seq_table_ITS2) %>%
@@
 
-```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%"}
+```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%", , cache=TRUE, extra.cache=gethash("ecm_heattree")}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
-```{r bioinfo}
+```{r bioinfo, cache=TRUE, extra.cache=gethash("bioinf_table", "bioinf_header")}
 options(knitr.kable.NA = '–')
@@
 
-```{r venn-table-asv}
+```{r venn-table-asv, cache=TRUE, extra.cache=gethash("venn_ASV"), more.cache=venn_table}
 readd(venn_ASV) %>%
@@
 
-```{r venn-table-otu}
+```{r venn-table-otu, cache=TRUE, extra.cache=gethash("venn_OTU"), more.cache=venn_table}
 readd(venn_OTU) %>%
@@
 
-```{r variofit, results="asis"}
+```{r variofit, results="asis", cache=TRUE, extra.cache=gethash("variofit_table")}
  readd(variofit_table) %>%

### 2020-10-08 writing/transect_supplement.Rmd
enlarge and landscape some supplementary figures.

@@
 2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.
-<!-- R3: please enlarge. This is too small. -->
 
-```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)", cache = TRUE, extra.cache = gethash("conc_plot", "reads_plot")}
+```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)", fig.env = "sidewaysfigure", fig.width = 9, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("conc_plot", "reads_plot")}
 ggpubr::ggarrange(readd(conc_plot), readd(reads_plot), ncol = 2, common.legend = TRUE) +
@@
 
-```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8, cache=TRUE, extra.cache=gethash("heattree")}
+```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", fig.width = 9, fig.height = 5, out.width="100%", fig.env = "sidewaysfigure", cache=TRUE, extra.cache=gethash("heattree")}
 theme <- theme_update(panel.border = element_blank())

### 2020-10-08 writing/transect_supplement.Rmd
improve buffer comparison heat trees and their captions.

@@
 
-(ref:buffer-compare-long) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries.
-<!-- R3: I have no idea how to interpret these figures. More explanation is needed!-->
+(ref:buffer-compare-long-s) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for long amplicon libraries
 
-```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)", cache = TRUE, extra.cache = gethash("buffer_compare_long")}
+(ref:buffer-compare-long) Reference heat tree (bottom left) shows labeled taxonomic tree.
+Nodes and branches are scaled to represent mean relative read abundance across samples.
+Colored comparison trees (top and right) illustrate variation in relative read abundance between samples preserved by different techniques or from different years.
+For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with orange colors indicate greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicate greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
+The two Xpedition datasets (blue-green colors, top right and center right) show consistently greater relative abundance of Basidiomycota, especially Russulales and Sebacinales, as well as Glomeromycota, Morteriellomycota, Geoglossomycetes, and Pezizomycetes,
+while the LifeGuard samples (orange colors, top right and center right) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, and also Geminibasidiomycetes.
+Differences between the two Xpedition datasets (top center) are less pronounced.
+
+```{r buffer-compare-long, fig.scap = "(ref:buffer-compare-long-s)", fig.cap = "(ref:buffer-compare-long-s).
+(ref:buffer-compare-long)", fig.width = 5, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_long")}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
-(ref:buffer-compare-short) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for short amplicon libraries.
+(ref:buffer-compare-short-s) Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs. LifeGuard) and sampling year, for short amplicon libraries
 
-```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)", cache = TRUE, extra.cache = gethash("buffer_compare_short"), more.cache=oldtheme}
+(ref:buffer-compare-short) Reference heat tree (bottom left) shows labeled taxonomic tree.
+Nodes and branches are scaled to represent mean relative read abundance across samples.
+Colored comparison trees (top and right) illustrate variation in relative read abundance between samples preserved by different techniques or from different years.
+For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with orange colors indicate greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicate greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
+The two Xpedition datasets (blue-green colors, top right and center right) show consistently greater relative abundance of Basidiomycota, especially Sebacinales, as well as Glomeromycota, Geoglossomycetes, and Pezizomycetes,
+while the LifeGuard samples (orange colors, top right and center right) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, as well as Geminibasidiomycetes.
+Differences between the two Xpedition datasets (top center) are less pronounced.
+
+```{r buffer-compare-short, fig.scap="(ref:buffer-compare-short-s)", fig.cap = "(ref:buffer-compare-short-s).
+(ref:buffer-compare-short)", fig.width = 5, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_short")}
+oldtheme <- theme_update(panel.border = element_blank())
 readd(buffer_compare_short)

### 2020-10-08 scripts/drake_plan_light.R

@@ -2352,16 +2352,18 @@ plan2 <- drake_plan(
         node_color = read_ratio,
         node_size = rowMeans(.$data$read_count[-1]),
         node_size_range = c(0.0001, 0.03),
-        node_size_axis_label = "Read abundance",
-        node_label_size_range = c(0.005, 0.03),
+        node_size_axis_label = "Read abund.",
+        node_label_size_range = c(0.01, 0.05),
         node_color_range = metacoder::diverging_palette(),
-        node_color_axis_label = "Log abundance ratio",
+        node_color_axis_label = "Log abund. ratio",
         node_color_interval = c(-3, 3),
         node_color_trans = "linear",
         # edge_size = rowMeans(.$data$read_count[-1]),
         # edge_size_axis_label = "Read abundance",
         # edge_size_range = c(0.001, 0.03),
         # edge_size_trans = "linear",
+        aspect_ratio = 1,
+        key_size = 0.6,
         layout = "davidson-harel",
         initial_layout = "reingold-tilford",
         output_file = file_out("temp/compare_long.pdf")
@@ -2399,16 +2401,17 @@ plan2 <- drake_plan(
         node_color = read_ratio,
         node_size = rowMeans(.$data$read_count[-1]),
         node_size_range = c(0.0001, 0.03),
-        node_size_axis_label = "Read abundance",
-        node_label_size_range = c(0.005, 0.03),
+        node_size_axis_label = "Read abund.",
+        node_label_size_range = c(0.01, 0.05),
         node_color_range = metacoder::diverging_palette(),
-        node_color_axis_label = "Log abundance ratio",
+        node_color_axis_label = "Log abund. ratio",
         node_color_interval = c(-3, 3),
         node_color_trans = "linear",
         # edge_size = rowMeans(.$data$read_count[-1]),
         # edge_size_axis_label = "Read abundance",
         # edge_size_range = c(0.001, 0.03),
         # edge_size_trans = "linear",
+        aspect_ratio = 1,
         layout = "davidson-harel",
         initial_layout = "reingold-tilford",
         output_file = file_out("temp/compare_short.pdf")

### 2020-10-08 scripts/drake_plan_light.R
install listings package to remove latexdiff warning, and check before trying to install latex packages.

@@ -3167,7 +3167,7 @@ plan2 <- drake_plan(
       file_in(!!file.path("writing", "transect_supplement_original.tex"))
       file_in(!!file.path("writing", "transect_supplement.tex"))
       file_out(!!file.path("writing", "transect_supplement_diff.tex"))
-      tinytex::tlmgr_install("latexdiff")
+      tinytex::tlmgr_install(setdiff(c("latexdiff", "listings"), tinytex::tl_pkgs()))
       system2(
         command = "latexdiff",
         args = list(
@@ -3190,7 +3190,7 @@ plan2 <- drake_plan(
       file_in(!!file.path("writing", "transect_paper_original.tex"))
       file_in(!!file.path("writing", "transect_paper.tex"))
       file_out(!!file.path("writing", "transect_paper_diff.tex"))
-      tinytex::tlmgr_install("latexdiff")
+      tinytex::tlmgr_install(setdiff(c("latexdiff", "listings"), tinytex::tl_pkgs()))
       system2(
         command = "latexdiff",
         args = list(

### 2020-10-09 writing/transect_paper.Rmd
install listings package to remove latexdiff warning, and check before trying to install latex packages.
fix broken citations and cross-references

@@
 
-tinytex::tlmgr_install("pdfcrop")
+tinytex::tlmgr_install(setdiff("pdfcrop", tinytex::tl_pkgs()))
 
@@
 Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(readd(bioinf_table)["ITS2","PacBio_Short_reads"]/readd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(readd(bioinf_table)["ITS2","Illumina_Short_reads"]/readd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(readd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/readd(rawcounts)["is_057"], format = "f", digits = 0)`),
-the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered, and were much closer to saturation of their respective species accumulation curves (Figure \ref(fig:accum-ASV)).
+the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered, and were much closer to saturation of their respective species accumulation curves (Figure \@ref(fig:accum-ASV)).
 OTU read counts were strongly correlated between technologies (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon == y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
@@
 Ion Torrent generated more ASV richness than PacBio or Illumina from the common short amplicon library (Figure\ \@ref(fig:venn)), even after rarefication to equal sampling depth ().
-In contrast PacBio and Illumina generated very similar ASV diversity estimates after rarefication (Figure \ref(fig:alpha-compare)), even though the original sequencing depth for Illumina was two orders of magnitude higher than that for PacBio.
+In contrast PacBio and Illumina generated very similar ASV diversity estimates after rarefication (Figure \@ref(fig:alpha-compare)), even though the original sequencing depth for Illumina was two orders of magnitude higher than that for PacBio.
 This suggests that the ASV diversity is artificially inflated in the Ion Torrent dataset.
-This may be a result of the lower fraction of very high quality reads (Figure \ref(fig:qual-check)).
+This may be a result of the lower fraction of very high quality reads (Figure \@ref(fig:qual-check)).
 Although we used options for DADA2 intended to improve performance on technologies with high rates of homopolymer indel errors [@callahan2020b], it is possible that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed [@callahan2016].
@@
 These discrepancies are probably due to primer mismatches to the relevant groups.
-The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae (@tedersoo2015).
-ITS1 and LR5 match a much broader range of fungal and other eukaryote groups (@tedersoo2015).
-The alternate LR5-F primer (@tedersoo2008) would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
+The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae [@tedersoo2015].
+ITS1 and LR5 match a much broader range of fungal and other eukaryote groups [@tedersoo2015].
+The alternate LR5-F primer [@tedersoo2008] would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
 We assert that more complete coverage of ECM fungal groups with high rDNA variability such as *Tulasnella* (and ideally other *Cantharellales*) is worth the read-depth spent on non-target groups.
@@
 
-Assignment of ecological function to environmental fungal sequences is dependent on accurate taxonomic identification, especially at the genus level or below (@nguyen2016funguild).
+Assignment of ecological function to environmental fungal sequences is dependent on accurate taxonomic identification, especially at the genus level or below [@nguyen2016funguild].
 However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.

### 2020-10-09 writing/transect_supplement.Rmd
install listings package to remove latexdiff warning, and check before trying to install latex packages.
fix knitting errors in supplement

@@
 
-gethash <- function(...) purrr::map_chr(list(...), drake_cache()$get_hash)
+gethash <- function(...) purrr::map_chr(purrr::flatten(list(...)), drake_cache()$get_hash)
 ```
@@
 
-```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
-plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2", cache = TRUE, extra.cache = gethash("buffer_asv_pcoa", "buffer_asv_table"))
+```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)", cache = TRUE, extra.cache = gethash("buffer_asv_pcoa", "buffer_asv_table")}
+plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2")
 points(
@@
 
-```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)", cache=TRUE, extra.cache=gethash(keep(light_plan$target, startsWith, "correlog"))}
+```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)", cache=TRUE, extra.cache=gethash(as.list(keep(light_plan$target, startsWith, "correlog")))}
 light_plan %>%

### 2020-10-09 scripts/drake_plan_light.R
specify samples sizes explicitly to clean up low-sample values on accumulation curves, and move OTU curve to supplement.

@@ -852,7 +852,10 @@ plan2 <- drake_plan(
   ),
 
   sample_iNEXT = target(
-    lapply(pre_iNEXT, iNEXT::iNEXT, knots = 200),
+    lapply(pre_iNEXT, function(x) iNEXT:::iNEXT.Ind(
+      x,
+      m = unique(c(round(10^seq(0, log10(sum(x)), 0.1)), sum(x))),
+      q = 0)),
     transform = map(pre_iNEXT, .id = cluster),
     dynamic = map(pre_iNEXT)
   ),
@@ -954,7 +957,7 @@ plan2 <- drake_plan(
   accum_plot = target(
     readd(sample_iNEXT) %>%
       purrr::flatten() %>%
-      purrr::map_dfr(fortify, .id = "sample") %>%
+      dplyr::bind_rows(.id = "sample") %>%
       tidyr::extract(
         sample,
         into = c("seq_run", "well"),
@@ -964,13 +967,13 @@ plan2 <- drake_plan(
       dplyr::left_join(datasets, by = "seq_run") %>%
       dplyr::mutate(strategy = paste(tech, amplicon)) %>%
       dplyr::filter(method != "extrapolated") %>%
-      ggplot(aes(x = x, y = y, color = strategy)) +
+      ggplot(aes(x = m, y = qD, color = strategy)) +
       # lines for each sample
       geom_line(aes(group = sample), alpha = 0.08) +
       # small points to represent the actual observed read depth and richness
       geom_point(data = ~dplyr::filter(., method == "observed"),
                  alpha = 0.8, size = 1, shape = 1) +
-      xlab("reads") +
+      xlab("Number of rarefied reads") +
       ylab(paste(cluster, "richness")) +
       scale_color_strategy(name = NULL) +
       scale_x_log10() +

### 2020-10-09 writing/transect_paper.Rmd
specify samples sizes explicitly to clean up low-sample values on accumulation curves, and move OTU curve to supplement.
make read count comparison flow better in results

@@
 When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% of reads, respectively (Figure\ \@ref(fig:venn)b).
-In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs, were found to be shared with other sequencing technologies upon OTU clustering.
-ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` of belonged to a unique OTU after clustering.
+In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
+ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` belonged to a unique OTU after clustering.
 In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).
 
-Figure\ \@ref(fig:read-compare) shows the correspondence between the read count for different ASVs (\@ref(fig:read-compare)a) and OTUs(\@ref(fig:read-compare)b) in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted as lines along the axes.
-In all cases, the read counts for shared ASVs and OTUs were correlated, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")`.
+Read counts for shared ASVs and OTUs were highly correlated between strategies, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")` (Figure\ \@ref(fig:read-compare)).
 Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
@@
 `r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short amplicon reads, respectively).
+
+<!-- ADD HERE results about the alpha diversity comparison -->
 
@@
 readd(accum_plot_ASV)
-```
-
-```{r accum-OTU, fig.cap="OTU accumulation curves"}
-readd(accum_plot_OTU)
 ```

### 2020-10-09 writing/transect_paper.Rmd
improve alpha diversity comparison plots

@@
 
-```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s) (ref:alpha-compare)", fig.height=8}
+```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s) (ref:alpha-compare)", fig.height=8, fig.width = 4, out.width = "50%"}
 ggpubr::ggarrange(
@@
   ncol = 1,
-  labels = "auto",
-  label.x = 0.9
+  labels = "auto"
 )

### 2020-10-09 scripts/drake_plan_light.R

@@ -910,7 +910,7 @@ plan2 <- drake_plan(
       ) %>%
       tidyr::unnest(deming) %>%
       dplyr::mutate(
-        slope_label = sprintf("slope==%.2f(%.2f-%.2f)",
+        slope_label = sprintf("m==%.2f(%.2f-%.2f)",
                              value_x, `lower 0.95`, `upper 0.95`),
         r2_label = sprintf("R^2==%.2f", r_squared)
       ),
@@ -925,22 +925,23 @@ plan2 <- drake_plan(
       geom_abline(
         aes(slope = value_x, intercept = `(Intercept)`),
         color = "blue",
+        alpha = 0.7,
         size = 1,
         data = demingfits
       ) +
       geom_text(
         aes(x = 0, y = 60, label = slope_label),
         hjust = 0,
-        size = 2.5,
-        color = "blue",
+        # color = "blue",
+        size = 1.5,
         parse = TRUE,
         data = demingfits
       ) +
       geom_text(
         aes(x = 0, y = 52, label = r2_label),
         hjust = 0,
-        size = 2.5,
-        color = "blue",
+        # color = "blue",
+        size = 1.5,
         parse = TRUE,
         data = demingfits
       ) +

### 2020-10-09 writing/transect_supplement.Rmd
specify samples sizes explicitly to clean up low-sample values on accumulation curves, and move OTU curve to supplement.
reword distance decay table legend
Auto stash before merge of "master" and "origin/master"

@@
 ```{r buffer-compare-long, fig.scap = "(ref:buffer-compare-long-s)", fig.cap = "(ref:buffer-compare-long-s).
-(ref:buffer-compare-long)", fig.width = 5, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_long")}
+(ref:buffer-compare-long)", fig.width = 10, fig.height = 10, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_long"), dev = "pdf"}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 ```{r buffer-compare-short, fig.scap="(ref:buffer-compare-short-s)", fig.cap = "(ref:buffer-compare-short-s).
-(ref:buffer-compare-short)", fig.width = 5, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_short")}
+(ref:buffer-compare-short)", fig.width = 10, fig.height = 10, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_short"), dev = "pdf"}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
+```{r accum-OTU, fig.cap="OTU accumulation curves"}
+readd(accum_plot_OTU)
+```
+
 (ref:taxa-s) Heat tree summarizing taxonomic composition of soil community
@@
 
-```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)", cache=TRUE, extra.cache=gethash(as.list(keep(light_plan$target, startsWith, "correlog")))}
+```{r correlog, fig.cap="(ref:correlog-cap-s).
+(ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)", cache=TRUE, extra.cache=gethash(as.list(purrr::keep(light_plan$target, stringr::str_detect, "^correlog.+_PHYLOTAX_")))}
 light_plan %>%
@@
 
-(ref:variofit) Parameters for exponential distance-decay fits. Range: range at which exponential function is at 95% of its maximum value.
+(ref:variofit) Parameters for exponential distance-decay fits. Range: range at which exponential function is at 95% of its asymptotic value.
 

### 2020-10-09 scripts/scratch.R

@@ -1792,63 +1792,3 @@ ncbi_get_taxon_summary()%>%
   ) %>%
   paste("uncultured", .) %T>%
   writeLines("output/ENA/env_taxa.txt")
-
-drake_cache()$list() %>% purrr::keep(stringr::str_detect, "physeq")
-loadd(asv_table)
-pooled_inext <- asv_table %>%
-  dplyr::mutate_at("seq", tzara::seqhash) %>%
-  tibble::column_to_rownames("seq") %>%
-  as.matrix() %>%
-  subtract(1) %>%
-  pmax(0) %>%
-  iNEXT::iNEXT()
-
-sample_inext <-
-  readd(big_seq_table_ITS2) %>%
-  magrittr::extract(rowSums(.) > 100,) %>%
-  t() %>%
-  iNEXT::iNEXT(knots = 200)
-
-alpha100 <- readd(big_seq_table_ITS2) %>%
-  magrittr::extract(rowSums(.) > 100,) %>%
-  t() %>%
-  {
-    dplyr::bind_cols(
-      list(sample = rownames(.)),
-      iNEXT::estimateD(., level = 100) %>%
-        dplyr::filter(order == 0)
-    )
-  }
-
-
-  only_obs <- function(d) {
-    dplyr::filter(d$data, method == "observed")
-}
-
-library(ggplot2)
-sample_inext %>%
-  ggplot2::fortify() %>%
-  tidyr::extract(
-    site,
-    into = c("seq_run", "t", "plate", "well", "region"),
-    regex = "((is|pb|SH)[-_][0-9]{3,4})_([^_]+)_([A-H]1?[0-9])_(.+)",
-    remove = FALSE
-  ) %>%
-  dplyr::filter(method != "extrapolated", seq_run != "is_057") %>%
-  ggplot(aes(x = x, y = y, color = seq_run)) +
-  # lines for each sample
-  geom_line(aes(group = site), alpha = 0.08) +
-  # use the alpha values I provided as the literal alpha values
-  scale_alpha_identity() +
-  # small points to represent the actual observed read depth and richness
-  geom_point(data = ~dplyr::filter(., method == "observed"), alpha = 0.8, size = 1, shape = 1) +
-  # seperate facet for each forest type.
-  # They are side-by-side so we can compare them visually.
-  # facet_wrap(facets = ~seq_run, nrow = 1) +
-  xlab("reads") +
-  ylab("Species richness") +
-  scale_x_log10() +
[…]

### 2020-10-12 writing/transect_paper.Rmd
start addressing mohammad's comments, move some methods to the supplement.

@@
 base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
-write_bib(c("base", "phyloseq", "vegan", "metacoder"), file = "R.bib", tweak = TRUE)
+write_bib(c("base", "phyloseq", "vegan", "metacoder", "deming"), file = "R.bib", tweak = TRUE)
 
@@
 Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
-Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
-It varies between different ecosystems and taxonomic groups.
+Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
+Turnover scale varies between different ecosystems and taxonomic groups.
 Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
@@
 
-We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity.
-The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
-If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
-Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
-Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
-In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
-
-We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into shorter homologous domains, independently denoising each segment, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
-We used this method, detailed below, for all of the PacBio and Ion Torrent datasets.
-The method is not applicable to Illumina paired-end reads.
-
-Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
-First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
-No attempt was made to remove the approximately 12\ bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
-<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
-A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
-A CM was generated from the alignment using `cmbuild` from Infernal.
-The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
-The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:regions),
-where LSU1-4 represent the conserved domains of LSU flanking the variable D1-3 domains [@michot1984].
-For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
-Code to extract the segments, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available on Github at https://github.com/brendanf/LSUx.
-
-Each of the extracted domains was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors, and denoised using DADA2.
-To test the performance of DADA2 on intermediate-length amplicons, adjacent domains from each read were concatenated to form full ITS (ITS1 + 5.8S + ITS2), partial LSU (LSU1 + D1 + LSU2 + D2 + LSU3 + D3 + LSU4), and partial 32S (5.8S + ITS2 + LSU) datasets, and these were also denoised.
-The DADA2 error model was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
-Independent error models were fit for each sequencing strategy (i.e., long *vs.* short amplicons, different sequencing technologies).
-DADA2 was run with complete pooling for PacBio libraries, and pseudo-pooling
-for Ion Torrent libraries.
-Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
-
-For each raw read, the denoised sequences of the different domains were concatenated to form a set of full-length sequences.
-For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
-Because ITS2 is the most variable of the amplified domains as indicated by the greatest number of unique ASVs (Figure\ \@ref(fig:regions)b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
-Therefore, concatenated sequences with identical ITS2 were clustered, and each cluster was aligned in R using the `DECIPHER` package [@wright2015].
-Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the `odseq` package [@jehl2015], using the default threshold of 0.025.
-The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence for the entire cluster.
-Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
-The count and sample distribution of reads assigned to each full-length ASV were calculated in order to form a sample × ASV community matrix.
-A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
-The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available at https://github.com/brendanf/tzara.
+We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
+We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains (Figure\ \@ref(fig:regions)a), independently denoising each segment, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
+We used this method, implemented in the new R packages `LSUx` (https://github.com/brendanf/LSUx) and `tzara` (https://github.com/brendanf/tzara) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.
 
@@
 
-Because the Illumina dataset consisted of paired-end reads, domains could not be extracted prior to denoising, and the ASVs generated according to the standard DADA2 workflow were used.
+Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
 The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.
@@
 
-To assess the completeness of sequencing using each of the sequencing strategies, ASV and OTU accumulation curves were generated using the iNEXT package [citation].
+To assess the completeness of sequencing using each of the sequencing strategies, ASV and OTU accumulation curves were generated using the `iNEXT` package [@hsieh2016].
 Additionally, average estimated ASV and OTU richness was calculated for each sample after rarefying to 100 sequences using 100 repetitions.
 ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies.
-For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the "deming" package [citation].
+For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the `deming` package [@R-deming].
 For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.
 
-The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the R package `vegan` [@R-vegan].
+The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the `vegan` package [@R-vegan].
 Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).
 
-In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
+In order to detect potential biases at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
 Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
@@
 Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
-Spatiotemporal effects were partialled out in order to visualize effects due to sequencing technology and amplicon length.
+Spatiotemporal effects, i.e.
+differences between different samples, were partialed out using a `Condition()` term in the `capscale` model in order to visualize effects due to sequencing technology and amplicon length.
 
@@
 Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
-PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads (<= 1 expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
+PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads ($\le 1$ expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
 Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \@ref(fig:qual-check)a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \@ref(fig:qual-check)b).
@@
 Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
-In contrast, extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-readd(filtercounts_ITS2)['pb_483']/readd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-readd(filtercounts_ITS2)['is_057']/readd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
-This represented greater loss of PacBio short amplicon reads, but less loss of PacBio long amplicon reads and Ion Torrent reads.
+In contrast, extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads.
 
@@
 but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
-In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
+The length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
 
-The length distribution of the different regions extracted from the long amplicon are shown in Figure\ \@ref(fig:region-lengths).
+Among the different regions extracted from the long amplicon (Figure\ \@ref(fig:region-lengths),
 ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
@@
 Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.
+A larger fraction of ITS2 sequences extracted from the long amplicon dataset were shorter than 140 bp than from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
+The taxonomic identity of these sequences will be discussed below.
 
@@
 
-<!-- ADD HERE results about the alpha diversity comparison -->
+ASV and OTU richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
+The slope of the relationship between PacBio and Illumina richness estimates was not significantly different from 1, indicating that these two technologies give essentially identical rarefied richness estimates, dispite the approximately 200× difference in original sequencing depth.
+Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$value_x %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV (Figure\ \@ref(fig:accum-ASV)) and OTU (Figure\ \@ref(fig:accum-OTU)) accumulation curves.
+Richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
+Although the total least squares regression lines indicate that the long amplicon dataset resulted in richness estimates which were more similar to the short amplicon results from Ion Torrent than the other two technologies, the confidence intervals for all slope estimates overlap with each other and 1.
+
+```{r accum-ASV, fig.cap="ASV accumulation curves"}
+readd(accum_plot_ASV)
+```
 
@@
 )
-```
-
-```{r accum-ASV, fig.cap="ASV accumulation curves"}
-readd(accum_plot_ASV)
 ```
@@
 According to the PHYLOTAX assignments, Fungi represented `r percent(readd(fungi_asv)["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(readd(fungi_read)["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(readd(fungi_asv)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(fungi_read)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
+Many of the ASVs which were unique to the long-amplicon library thus fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)).
+In particular a large fraction of ITS2 sequences with length less than 140 (Figure\ \@ref(fig:ITS2-length)) were identified as Alveolates (Figure\ \@ref(fig:short-its)).
+
 Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
@@
 Fungi categorized as ECM made up `r percent(readd(ecm_asv)["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(readd(ecm_read)["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(readd(ecm_asv)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(ecm_read)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library (Figure\ \@ref(fig:ecm)).
-Although amplicon length had a significant effect on ECM community composition at the family level, the explained variation was very low (PERMANOVA with 9999 permutations, $p=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), and the majority of variation was again spatiotemporal ($p<`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$).
+Although amplicon length had a significant effect on ECM community composition at the family level (Figure\ \@ref(fig:ecm-heattree)), the explained variation was very low (PERMANOVA with 9999 permutations, $p=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), and the majority of variation was again spatiotemporal ($p<`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$).
 Variation between sequencing technologies was not significant ($p=`r formatC(readd(tech_ecm_fam_adonis)["tech", "p.value"], format = "f", digits = 2)`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["tech", "R2"], format = "f", digits = 4)`$).
@@
 
+<!-- 592 I’d start the discussion with a couple of introductory sentences pointing out the main finding and guiding the reader to the next sections.-->
+
 ## Reconstruction of long amplicons from denoised subregions
@@
 
+<!-- These are mostly  results, could you summarise them here (make sure all figures are cited in the results)? This would also address the reviewers comment. -->
 DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
-Ion Torrent generated more ASV richness than PacBio or Illumina from the common short amplicon library (Figure\ \@ref(fig:venn)), even after rarefication to equal sampling depth ().
-In contrast PacBio and Illumina generated very similar ASV diversity estimates after rarefication (Figure \@ref(fig:alpha-compare)), even though the original sequencing depth for Illumina was two orders of magnitude higher than that for PacBio.
-This suggests that the ASV diversity is artificially inflated in the Ion Torrent dataset.
+
+ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures\ \@ref(fig:venn) and \@ref(fig:alpha-compare)).
 This may be a result of the lower fraction of very high quality reads (Figure \@ref(fig:qual-check)).
-Although we used options for DADA2 intended to improve performance on technologies with high rates of homopolymer indel errors [@callahan2020b], it is possible that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed [@callahan2016].
+We used options for DADA2 intended to improve performance on technologies with higher rates of homopolymer indel errors [@callahan2020b], but our results suggest that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed [@callahan2016].
 
-Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any ITS2 fragments longer than those recovered by Illumina and Ion Torrent.
+Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any reads from the short amplicon library which were longer than those recovered by Illumina and Ion Torrent.
+<!-- Next sentence (57 words) could be cut... -->
 Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
@@
 
+<!-- These are mostly  results, could you summarise them here (make sure all figures are cited in the results)? This would also address the reviewers comment. -->
 Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% ITS2 similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])`, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
-ITS2 sequences extracted from the long amplicon dataset also included some sequences that were much shorter than those recovered from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
-Taxonomic assignments revealed that the majority of these non-shared sequences fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Figure\ \@ref(fig:short-its)).
+
 Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
 Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
+<!-- end range of previous comment -->
 These discrepancies are probably due to primer mismatches to the relevant groups.
@@
 
-Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
-The general shape of the correlograms were similar for long and short amplicon datasets.
-We hypothesize that this pattern may result from a superposition of positively and negatively correlated temporal processes operating at different spatial scales.
-As an example, the positive year-to-year correlation at 0\ m distance could be due to one species or set of species which strongly persist between years, but are primarily confined to patch sizes $<1$\ m,
-while the negative year-to-year correlation could be due to a different species or set of species which deplete nutrients within a $>1$\ m patch over the course of a year, and are likely to be absent from the same patch in the following year.
+Both datasets from the short amplicon library showed no temporal autocorrelation, but a weak negative spatiotemporal correlation at 1\ m and 1\ year, which was probably a statistical artifact.
+The correlograms for the long amplicon datasets showed similar trends.
 
@@
 The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m, indicating little evidence of spatial structure.
-This is probably because the majority of species turnover in this system, especially among ECM fungi, is between closely related species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
-This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac as compared to Bray-Curtis in Figure \@ref(fig:spatio)
+This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
+This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac as compared to Bray-Curtis in Figure \@ref(fig:spatio).
 UniFrac would be more suited at larger spatial scales and/or larger ecological gradients.
@@
 
-The choice of amplicon and sequencing technology did not seem to affect the results of the spatial analysis, provided sufficient sequencing depth.
+The choice of amplicon and sequencing technology did not affect the results of the spatial analysis, provided sufficient sequencing depth.
 However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments using our new tool PHYLOTAX.

### 2020-10-12 writing/transect_supplement.Rmd
start addressing mohammad's comments, move some methods to the supplement.

@@
     Parakou, Benin
+bibliography:
+  - "all.bib"
+  - "R.bib"
+biblio-style: apa
 always_allow_html: true
@@
     citation_package: biblatex
-    suppress-bibliography: true
+    suppress-bibliography: false
     fontsize: 12pt
@@
 
+# Supplementary methods
+
+The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
+If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
+Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
+For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
+Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
+In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
+
+In order to bypass this limitation, we implemented a workflow which divides the long rDNA amplicon into homologous domains (R package `LSUx`), denoises the domains individually using DADA2, and then recombines the denoised domains to form full-length denoised amplicons (R package `tzara`).
+The steps of this workflow are detailed below.
+
+## Splitting the rDNA into regions with covariance models: `LSUx`
+
+Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
+First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
+No attempt was made to remove the approximately 12\ bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
+<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
+A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
+A CM was generated from the alignment using `cmbuild` from Infernal.
+The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
+The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:regions),
+where LSU1-4 represent the conserved domains of LSU flanking the variable D1-3 domains [@michot1984].
+For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
+Code to extract the segments, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available and documented at https://github.com/brendanf/LSUx.
+
+## Denoising each domain independently with `DADA2`
+
+Each of the extracted domains was independently filtered for length (Table\ \@ref(tab:region-limits)) and a maximum of three expected errors, and denoised using DADA2.
+To test the performance of DADA2 on intermediate-length amplicons, adjacent domains from each read were concatenated to form full ITS (ITS1 + 5.8S + ITS2), partial LSU (LSU1 + D1 + LSU2 + D2 + LSU3 + D3 + LSU4), and partial 32S (5.8S + ITS2 + LSU) datasets, and these were also denoised.
+The DADA2 error model was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
+Independent error models were fit for each sequencing strategy (i.e., long *vs.* short amplicons, different sequencing technologies).
+DADA2 was run with complete pooling for PacBio libraries, and pseudo-pooling
+for Ion Torrent libraries.
+Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.
+
+## Generation of full-length denoised ASVs: `tzara`
+
+For each raw read, the denoised sequences of the different domains were concatenated to form a set of full-length sequences.
+For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
+Because ITS2 is the most variable of the amplified domains as indicated by the greatest number of unique ASVs (Figure\ \@ref(fig:regions)b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
+Therefore, concatenated sequences with identical ITS2 were clustered, and each cluster was aligned in R using the `DECIPHER` package [@wright2015].
+Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the `odseq` package [@jehl2015], using the default threshold of 0.025.
+The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence for the entire cluster.
+Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
+The count and sample distribution of reads assigned to each full-length ASV were calculated in order to form a sample × ASV community matrix.
+A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
+The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available and documented at https://github.com/brendanf/tzara.
+
+The two packages are modular; `LSUx` can be used to extract domains from LSU for any purpose, and `tzara` can also be used to recombine denoised sequences which were originally separated using other approaches, most notably using the popular and well-established ITSx [@bengtsson-palme2013].
+However, for our dataset, the single LSU region which would be extracted by ITSx was still too long to be effectively denoised by DADA2 (Figure\ \@ref(fig:regions)b in main text).
+
 (ref:buffer-pcoa-s) Unconstrained Principal Coordinate Analysis (PCoA) ordination of samples preserved with different buffers
@@
 Colored comparison trees (top and right) illustrate variation in relative read abundance between samples preserved by different techniques or from different years.
-For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with orange colors indicate greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicate greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
-The two Xpedition datasets (blue-green colors, top right and center right) show consistently greater relative abundance of Basidiomycota, especially Russulales and Sebacinales, as well as Glomeromycota, Morteriellomycota, Geoglossomycetes, and Pezizomycetes,
-while the LifeGuard samples (orange colors, top right and center right) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, and also Geminibasidiomycetes.
-Differences between the two Xpedition datasets (top center) are less pronounced.
+For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with red-orange colors indicating greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicating greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
+The two Xpedition datasets (blue-green colors in top right and center right trees) show consistently greater relative abundance of Basidiomycota, especially Russulales and Sebacinales, as well as Glomeromycota, Morteriellomycota, Geoglossomycetes, and Pezizomycetes,
+while the LifeGuard samples (red-orange colors in top right and center right trees) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, and also Geminibasidiomycetes.
+Differences between the two Xpedition datasets (top center tree) are less pronounced.
 
@@
 Colored comparison trees (top and right) illustrate variation in relative read abundance between samples preserved by different techniques or from different years.
-For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with orange colors indicate greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicate greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
-The two Xpedition datasets (blue-green colors, top right and center right) show consistently greater relative abundance of Basidiomycota, especially Sebacinales, as well as Glomeromycota, Geoglossomycetes, and Pezizomycetes,
-while the LifeGuard samples (orange colors, top right and center right) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, as well as Geminibasidiomycetes.
-Differences between the two Xpedition datasets (top center) are less pronounced.
+For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with red-orange colors indicating greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicating greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
+The two Xpedition datasets (blue-green colors in top right and center right trees) show consistently greater relative abundance of Basidiomycota, especially Sebacinales, as well as Glomeromycota, Geoglossomycetes, and Pezizomycetes,
+while the LifeGuard samples (red-orange colors in top right and center right trees) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, as well as Geminibasidiomycetes.
+Differences between the two Xpedition datasets (top center tree) are less pronounced.
 
@@
 
+(ref:ITS2-length-s) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies
+
+(ref:ITS2-length) *a*) Length distribution.  *b*) Empirical cumulative distribution function.
+
+```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6, cache=TRUE, extra.cache=gethash("its2_dens", "its2_ecdf")}
+ggpubr::ggarrange(
+  readd(its2_dens),
+  readd(its2_ecdf),
+  ncol = 1,
+  labels = "auto",
+  heights = c(4, 2)
+)
+```
+
 (ref:read-compare-s) Comparison between read numbers for different sequencing strategies
@@
 
+(ref:heattree-length-compare-s) Variation in read abundance and diversity for different amplicons across the taxonomic tree
+
+(ref:heattree-length-compare) Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
+In both cases red-orange represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
+Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
+Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
+ASVs which could not be further assigned are labelled "?".
+Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
+Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.
+
+```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s).
+(ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%", cache=TRUE, extra.cache=gethash("heattree_length_compare")}
+oldtheme <- theme_update(panel.border = element_blank())
+readd(heattree_length_compare)
+theme_set(oldtheme)
+```
+
+(ref:short-its-s) Kingdom-level taxonomic composition of ITS2 ASVs $\le 140$ bp
+
+(ref:short-its) Vertical axis shows fraction of total reads $\le 140$ bp in each dataset.
+
+```{r short-its, fig.cap = "(ref:short-its-s).
+(ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
+readd(big_seq_table_ITS2) %>%
+  colSums() %>%
+  enframe(name = "ITS2", value = "nreads") %>%
+  mutate(length = nchar(ITS2)) %>%
+  filter(length <= 140) %>%
+  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
+  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
+  group_by(tech, amplicon, Algorithm, reference) %>%
+  mutate(reads = reads/sum(reads)) %>%
+  ungroup() %>%
+  # group_by(ITS2) %>%
+  # filter(!"Short" %in% amplicon) %>%
+  taxon_plot(kingdom, y = reads, datasets = datasets)
+```
+
 (ref:ecm-s) Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments
@@
   theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+(ref:ecm-heattree-s) Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only
+
+(ref:ecm-heattree) Red-orange edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
+Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
+Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
+ASVs which could not be further assigned are labelled "?".
+Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
+The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
+
+```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%", , cache=TRUE, extra.cache=gethash("ecm_heattree")}
+oldtheme <- theme_update(panel.border = element_blank())
+readd(ecm_heattree)
+theme_set(oldtheme)
 ```
@@
   theme(strip.background = element_blank())
-```
-
-(ref:ITS2-length-s) Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies
-
-(ref:ITS2-length) *a*) Length distribution.  *b*) Empirical cumulative distribution function.
-
-```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6, cache=TRUE, extra.cache=gethash("its2_dens", "its2_ecdf")}
-ggpubr::ggarrange(
-  readd(its2_dens),
-  readd(its2_ecdf),
-  ncol = 1,
-  labels = "auto",
-  heights = c(4, 2)
-)
-```
-
-(ref:heattree-length-compare-s) Variation in read abundance and diversity for different amplicons across the taxonomic tree
-
-(ref:heattree-length-compare) Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
-In both cases red represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
-Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
-Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
-ASVs which could not be further assigned are labelled "?".
-Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
-Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.
-
-```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s).
-(ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%", cache=TRUE, extra.cache=gethash("heattree_length_compare")}
-oldtheme <- theme_update(panel.border = element_blank())
-readd(heattree_length_compare)
-theme_set(oldtheme)
-```
-
-(ref:short-its-s) Kingdom-level taxonomic composition of ITS2 ASVs $\le 140$ bp
-
-(ref:short-its) Vertical axis shows fraction of total reads $\le 140$ bp in each dataset.
-
-```{r short-its, fig.cap = "(ref:short-its-s).
-(ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
-readd(big_seq_table_ITS2) %>%
-  colSums() %>%
-  enframe(name = "ITS2", value = "nreads") %>%
-  mutate(length = nchar(ITS2)) %>%
-  filter(length <= 140) %>%
-  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
-  group_by(tech, amplicon, Algorithm, reference) %>%
-  mutate(reads = reads/sum(reads)) %>%
-  ungroup() %>%
-  # group_by(ITS2) %>%
-  # filter(!"Short" %in% amplicon) %>%
-  taxon_plot(kingdom, y = reads, datasets = datasets)
-```
-
-(ref:ecm-heattree-s) Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only
-
-(ref:ecm-heattree) Red edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
-Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
-Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
-ASVs which could not be further assigned are labelled "?".
-Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
-The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
-
-```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%", , cache=TRUE, extra.cache=gethash("ecm_heattree")}
-oldtheme <- theme_update(panel.border = element_blank())
-readd(ecm_heattree)
-theme_set(oldtheme)
 ```

### 2020-10-13 writing/transect_paper.Rmd
a little rewording

@@
 The result of this process is twofold.
-First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade which is consistently assigned by the .
+First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade with consistent preliminary assignments.
 Second, it resolves conflicts between the different reference-based algorithms, but only if one of the alternatives is consistent with the assignments of other ASVs in the same clade.
@@
 $$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
-where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity from replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill"; and $a_t$, the equivalent temporal range.
-The 95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package in R.
+where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity between replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill"; and $a_t$, the equivalent temporal range.
+The 95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package.
 
@@
 ASV recovery for long amplicons using DADA2 was dramatically improved from `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` to `r percent((readd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of reads by denoising homologous subregions independently using our new `LSUx` and `tzara` packages.
-LSUx was inspired by ITSx [@bengtsson-palme2013], which extracts the the ITS regions from the surrounding SSU, 5.8S, and LSU.
-However, we found that the resulting LSU segment was still too long for successful denoising (Figure \@ref(fig:regions)), necessitating more fine-grained extraction of domains.
 Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
-Thus, `tzara` should increase ASV recovery from these platforms as well.
+Thus, `tzara` should increase recovery of rare ASVs from these platforms as well.
 It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding [@loit2019].
@@
 This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
-This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac as compared to Bray-Curtis in Figure \@ref(fig:spatio).
+This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4-0.6) as compared to Bray-Curtis (0.8-1.0) in Figure \@ref(fig:spatio).
 UniFrac would be more suited at larger spatial scales and/or larger ecological gradients.

### 2020-10-13 writing/transect_paper.Rmd
move more text to supplementary methods

@@
 Wider than 5 mm?
-MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
-Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of field lysis/preservation buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
-Due to the inavailability of refrigeration, extraction was initiated in the field according to the manufacturers instructions by lysis using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
-
-An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
-Sequencing results for these 50 samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such the LifeGuard preserved samples were excluded from our spatial analyses.
-However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
+MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade and homogenized.
+A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (Supplementary methods).
 
@@
 
-After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
-Samples preserved using LifeGuard were first centrifuged at 10000\ g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
-DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
-
-The DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
+DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
 Due to length restrictions for Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
@@
 Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
-
-The forward primers for the short amplicon included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
-150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
-The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
-Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
-
-Both forward and reverse primers for the long amplicon included indexes for combinatorical multiplexing (Supplementary File 2).
-PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
-Reaction conditions were 10\ min at 95°, 30\ cycles of 45\ s at 95°, 45\ s at 59°, and 90\ s at 72°, and finally 10\ min at 72°.
-Each reaction was performed in three technical replicates as for short amplicons.
-
-After pooling of technical replicates, amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
-An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
+See Supplementary methods for PCR and library preparation protocols.
 
@@
 We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
-We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains (Figure\ \@ref(fig:regions)a), independently denoising each segment, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
+We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1-3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1-4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
 We used this method, implemented in the new R packages `LSUx` (https://github.com/brendanf/LSUx) and `tzara` (https://github.com/brendanf/tzara) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.
@@
 An ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
+The tree was rooted outside kingdom Fungi and non-Fungi were removed based on taxonomic assignments (see below and Supplementary methods).
 
-Taxonomic annotations of the RDP LSU fungal training set version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
-In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c as described in [@tedersoodata2017].
-Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
-Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
-
+Taxonomic annotations of the RDP LSU fungal training set version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods).
 Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
@@
 
-The phylogenetic tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
-Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
-The kingdom Fungi was identified as the minimal clade containing all ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
-ASVs falling outside this clade were not included in downstream fungal community analysis.
-
-For the ASVs from the long amplicon dataset, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax)).
-Taxa at each rank were assigned to a node and all its descendants if that taxon was consistent with the reference-based taxonomic assignments for each of the descendants.
-A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
-The result of this process is twofold.
-First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade with consistent preliminary assignments.
-Second, it resolves conflicts between the different reference-based algorithms, but only if one of the alternatives is consistent with the assignments of other ASVs in the same clade.
+For the ASVs from the long amplicon dataset, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax) and Supplementary Methods).
 This refinement algorithm is referred to as "PHYLOTAX", available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
@@
 
-To assess the completeness of sequencing using each of the sequencing strategies, ASV and OTU accumulation curves were generated using the `iNEXT` package [@hsieh2016].
-Additionally, average estimated ASV and OTU richness was calculated for each sample after rarefying to 100 sequences using 100 repetitions.
-ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies.
-For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the `deming` package [@R-deming].
-For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.
-
-The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the `vegan` package [@R-vegan].
-Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).
-
-In order to detect potential biases at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
-Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
-The community matrix was normalized to proportional abundance by dividing the counts for each class in each sample by the total number of reads for that sample, but were not rarefied to even sampling depth [@mcmurdie2014].
-PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs. short); and sequencing technology (Illumina MiSeq vs. PacBio RS II).
-Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
-Spatiotemporal effects, i.e.
-differences between different samples, were partialed out using a `Condition()` term in the `capscale` model in order to visualize effects due to sequencing technology and amplicon length.
-
-A similar analysis was also applied to only fungi classified as ECM, clustered at the family level.
-Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
-All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
+We compared alpha diversity estimates by the different sequencing strategies by calculating ASV and OTU accumulation curves, as well as comparing richness estimates after rarefaction for each sample (Supplementary methods).
+We also compared the effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, using PERMANOVA and heat tree visualizations (Supplementary methods).
 
@@
 
-Additionally, empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R.
-Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
-For datasets where the Mantel correlogram indicated spatial correlation between samples taken in separate years, the model was re-fit with an additional term to represent temporal correlation:
-$$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
-where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity between replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill"; and $a_t$, the equivalent temporal range.
-The 95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package.
+Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R (Supplementary methods).
 
@@
 The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table\ \@ref(tab:bioinfo)).
-Sequencing with PacBio RS II yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
-Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
+Sequencing with PacBio yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
+Ion Torrent and Illumina yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
 PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads ($\le 1$ expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
@@
 Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
-In contrast, extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads.
+Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
 
-Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Figure\ \@ref(fig:full-length)a).
-Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure\ \@ref(fig:full-length)b).
+Almost all of the short amplicon sequences from all three technologies were between 240 and 375\ bp long (Figure\ \@ref(fig:full-length)a).
+Although the length profile of the three sequencing runs were similar, Illumina had the largest fraction of reads near the top of the range, followed by Ion Torrent and PacBio (Figure\ \@ref(fig:full-length)b).
 The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic %>% formatC(format = "e", digits = 2)`, $p < 2.2\times10^{-16}$),
 but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
-The length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.
+The length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp (Figure\ \@ref(fig:full-length)c).
 
@@
 Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.
-A larger fraction of ITS2 sequences extracted from the long amplicon dataset were shorter than 140 bp than from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
-The taxonomic identity of these sequences will be discussed below.
+Around `r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %>% summarize(smallfrac = sum(reads[length <= 140]) / sum(reads)) %$% range(smallfrac) %>% percent(digits = 2) %>% unique() %>% paste(collapse = " to ")` of ITS2 sequences extracted from the long amplicon dataset were shorter than 140 bp, a much greater fraction than the `r readd(reads_table) %>% filter(region == "ITS2", seq_run != "pb_500") %>% group_by(seq_run) %>% summarize(smallfrac = sum(reads[length <= 140]) / sum(reads)) %$% range(smallfrac) %>% percent(digits = 2) %>% unique() %>% paste(collapse = " to ")` from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
+The taxonomic identity of these sequences is discussed below.
 
@@
 
-ASV and OTU richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
+ASV richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(readd(demingfits_ASV), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
 The slope of the relationship between PacBio and Illumina richness estimates was not significantly different from 1, indicating that these two technologies give essentially identical rarefied richness estimates, dispite the approximately 200× difference in original sequencing depth.
-Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$value_x %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV (Figure\ \@ref(fig:accum-ASV)) and OTU (Figure\ \@ref(fig:accum-OTU)) accumulation curves.
-Richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
+Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(readd(demingfits_ASV), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$value_x %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% unique() %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV accumulation curves (Figure\ \@ref(fig:accum-ASV)).
+ASV richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figures\ \@ref(fig:alpha-compare)).
 Although the total least squares regression lines indicate that the long amplicon dataset resulted in richness estimates which were more similar to the short amplicon results from Ion Torrent than the other two technologies, the confidence intervals for all slope estimates overlap with each other and 1.
+Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure\ \@ref(fig:venn)), sample-wise OTU accumulation curves (Figure\ \@ref(fig:accum-OTU)) and rarefied OTU richness relationships between sequencing strategies (Figure\ \@ref(fig:alpha-compare-otu)) were highly similar to those for ASVs.
 
@@
 
-(ref:alpha-compare-s) Comparison of alpha diversity between sequencing technologies
+(ref:alpha-compare-s) Comparison of ASV richness between sequencing technologies
 
-(ref:alpha-compare) for ASVs **(a)** and OTUs **(b)**.
-Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
+(ref:alpha-compare)  Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
 All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.
@@
 
-```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s) (ref:alpha-compare)", fig.height=8, fig.width = 4, out.width = "50%"}
-ggpubr::ggarrange(
-  remove_empty_facets(readd(alpha_plot_ASV)),
-  remove_empty_facets(readd(alpha_plot_OTU)),
-  ncol = 1,
-  labels = "auto"
-)
+```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s) (ref:alpha-compare)"}
+grid::grid.newpage()
+grid::grid.draw(remove_empty_facets(readd(alpha_plot_ASV)))
 ```
@@
 Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
-The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm) and \@ref(fig:ppcoa)).
+The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
+Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
 
@@
 
-<!-- These are mostly  results, could you summarise them here (make sure all figures are cited in the results)? This would also address the reviewers comment. -->
-Of long amplicon reads, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% ITS2 similarity only reduced this fraction to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])`, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
-
-Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
-Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
-<!-- end range of previous comment -->
-These discrepancies are probably due to primer mismatches to the relevant groups.
+The long amplicon dataset included `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` unique taxa, even after clustering at 97% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
 The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae [@tedersoo2015].
@@
 The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
+[@heeger2019] also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
+<!-- maybe the next 2 sentences could be dropped? 41 words-->
 Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, particularly the Warcup database, where it places <25% of ASVs even to kingdom.
@@
 Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications (Figure \@ref(fig:tax-chart)).
-This suggests that different assignment algorithms and databases bring mostly complementary, non-contradictory information at higher taxonomic levels.
-However, contradictory assignments between different methods is more common at lower taxonomic levels, which can be problematic because accurate assignment at the family or genus level is generally required for ecological guild assignment using FUNGuild.
+Inconsistent family and genus level assignments are problematic because accurate assignment at these ranks is generally required for ecological guild assignment using FUNGuild.
 
@@
 We speculate that this discrepancy is due to an increased ability to capture spatially variable rare species using HTS.
+
 For the short amplicon dataset, PacBio showed a spatial turnover range more than twice as long as showed by Illumina (Table \@ref(tab:variofit)) for both the total fungi and ECM communities, with wide confidence intervals.
@@
 
-Both datasets from the short amplicon library showed no temporal autocorrelation, but a weak negative spatiotemporal correlation at 1\ m and 1\ year, which was probably a statistical artifact.
+Neither datasets from the short amplicon library showed temporal autocorrelation, but a weak negative spatiotemporal correlation was detected at 1\ m and 1\ year, which was probably a statistical artifact.
 The correlograms for the long amplicon datasets showed similar trends.
@@
 This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4-0.6) as compared to Bray-Curtis (0.8-1.0) in Figure \@ref(fig:spatio).
-UniFrac would be more suited at larger spatial scales and/or larger ecological gradients.
+UniFrac analysis would be more suited at larger spatial scales and/or larger ecological gradients.
 

### 2020-10-13 writing/transect_supplement.Rmd
move more text to supplementary methods

@@
 
+## Sampling
+
+Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Both sites were located in Soudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+At each site, 25\ soil samples were collected along a single 24 m linear transect at intervals of 1\ m in May 2015.
+One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
+For each sample, any coarse organic debris <!--Do you have any definition?
+Wider than 5 mm?
+MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
+Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
+Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of field lysis/preservation buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
+Due to the inavailability of refrigeration, extraction was initiated in the field according to the manufacturers instructions by lysis using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
+
+An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
+Sequencing results for these 50 samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such the LifeGuard preserved samples were excluded from our spatial analyses.
+However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.
+
+## DNA extraction, amplification, and sequencing
+
+After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
+Samples preserved using LifeGuard were first centrifuged at 10000\ g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
+DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
+
+The forward primers for the short amplicon included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
+Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
+150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
+The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
+Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
+
+Both forward and reverse primers for the long amplicon included indexes for combinatorical multiplexing (Supplementary File 2).
+PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
+Reaction conditions were 10\ min at 95°, 30\ cycles of 45\ s at 95°, 45\ s at 59°, and 90\ s at 72°, and finally 10\ min at 72°.
+Each reaction was performed in three technical replicates as for short amplicons.
+
+After pooling of technical replicates, amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
+An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
+
+## Denoising and clustering
+
+### Denoising long-reads: the numbers are against us
+
 The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
@@
 
-## Splitting the rDNA into regions with covariance models: `LSUx`
+### Splitting the rDNA into regions with covariance models: `LSUx`
 
@@
 
-## Denoising each domain independently with `DADA2`
+### Denoising each domain independently with `DADA2`
 
@@
 
-## Generation of full-length denoised ASVs: `tzara`
+### Generation of full-length denoised ASVs: `tzara`
 
@@
 However, for our dataset, the single LSU region which would be extracted by ITSx was still too long to be effectively denoised by DADA2 (Figure\ \@ref(fig:regions)b in main text).
+
+## Phylogenetic inference and taxonomy assignment
+
+### Rooting the fungal tree
+
+The phylogenetic tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
+Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
+The kingdom Fungi was identified as the minimal clade containing all ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
+ASVs falling outside this clade were not included in downstream fungal community analysis.
+
+### Uniform taxonomic classification for Unite, Warcup, and RDP
+
+The taxonomic classification system is based on the Unite database version 8 [@nilsson2019a].
+In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c as described in [@tedersoodata2017].
+Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
+Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
+FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
+
+### Refining taxonomic assignments with a phylogenetic tree: PHYLOTAX
+
+PHYLOTAX refinements are based on (1) taxonomic assignments from other, primary assignment algorithms, and (2) a phylogenetic tree.
+Taxa at each rank are assigned to a node and all its descendants if that taxon is consistent with the reference-based taxonomic assignments for each of the descendants (Figure\ \@ref(fig:phylotax)).
+A taxon assignment is considered to be consistent if at least one primary algorithm assigned that taxon at greater than 50% confidence, or if no primary algorithm successfully classified the sequence at greater than 50% confidence.
+The result of this process is twofold.
+First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade with consistent preliminary assignments.
+Second, it resolves conflicts between the different reference-based algorithms, but only if one of the alternatives is consistent with the assignments of other ASVs in the same clade.
+PHYLOTAX is implemented in the new R package `phylotax` at https://github.com/brendanf/phylotax.
+
+## Effect of sequencing strategy on recovered community
+
+ASV and OTU accumulation curves were generated using the `iNEXT` package [@hsieh2016].
+Additionally, ASV and OTU richness was estimated for each sample by rarefying to 100 sequences per sample and taking the average of 100 repetitions.
+ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies.
+For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the `deming` package [@R-deming], with the intercept constrained to be 0.
+For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.
+
+The effect of sequencing technology and amplicon length on the recovered fungal community
+composition, as assessed by the Bray-Curtis dissimilarity was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the `vegan` package [@R-vegan].
+Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).
+
+In order to detect potential biases at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
+Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
+The community matrix was normalized to proportional abundance by dividing the counts for each class in each sample by the total number of reads for that sample, but were not rarefied to even sampling depth [@mcmurdie2014].
+PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs. short); and sequencing technology (Illumina vs. PacBio).
+Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
+Spatiotemporal effects, i.e.
+differences between different samples, were partialed out using a `Condition()` term in the `capscale` model in order to visualize effects due to sequencing technology and amplicon length.
+
+A similar analysis was also applied to only fungi classified as ECM, clustered at the family level.
+Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
+All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
+
+Overall community composition, and differences in composition between amplicon libraries, were visualized as heat trees using the `metacoder` package [@R-metacoder].
+
+## Spatiotemporal analysis
+
+Empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by Legendre and Legendre (2012) using the nls function in R.
+Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
+For datasets where the Mantel correlogram indicated spatial correlation between samples taken in separate years, the model was re-fit with an additional term to represent temporal correlation:
+$$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
+where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity between replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill"; and $a_t$, the equivalent temporal range.
+The 95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package.
+
+# Supplementary Figures and Tables
 
@@
 readd(accum_plot_OTU)
+```
+
+(ref:alpha-compare-otu-s) Comparison of OTU richness between sequencing technologies
+
+(ref:alpha-compare-OTU) Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
+All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.
+Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction.
+Blue line is a total least squares fit, with the given slope (and 95% confidence interval) and R^2^ value.
+Dashed diagonal line indicates a slope of 1.
+
+```{r alpha-compare-otu, fig.scap="(ref:alpha-compare-otu-s)", fig.cap="(ref:alpha-compare-otu-s) (ref:alpha-compare)"}
+grid::grid.newpage()
+grid::grid.draw(remove_empty_facets(readd(alpha_plot_OTU)))
 ```

### 2020-10-15 writing/transect_paper.Rmd
remove reviewer comments from manuscript

@@
 
-<!-- R1: This study seeks to determine how both sequencing target length and technology affect the spatial and temporal turnover of soil fungal communities in an African forest.
-The authors take an ambitious approach to these research questions by comparing multiple short-read technologies (Illumina, PacBio, and Ion Torrent) to a related long-read technology (PacBio), allowing them to make many valuable comparisons across datasets.
-Along with field sample replication, both spatially and temporally, the authors carefully consider a number of different bioinformatic steps for each sample, e.g.
-among taxonomic databases as well as whether to cluster ASV data or not.
-Importantly, after discovering that long read data has inherent properties that make it poorly suited for the popular DADA2 bioinformatic pipeline, the authors generate a series of new R-based tools to analyze long read data in ways that greater increase data retention.
-Further, they use a clever set of covariance models to improve the taxonomic assignments of ASVs that are not assigned at the bulk taxonomy step.
-This additional step is important in increasing the amount of sequence data that can be annotated to putative functional guilds, which in this case focuses on ECM fungi.
-
-Overall, I found the study to be a really excellent demonstration of how different sequencing technologies and target lengths can be compared for fungi.
-The overall conclusion that the ecological outcome is largely similar between the different datasets is heartening.
-I am quite sure that longer reads will continue to increase in popularity as costs decline and this study represents an important bridge in that transition – both for its careful benchmarking but also for its presentation of tools that allow for better incorporation of long read sequence data into existing bioinformatic pipelines.
-I hope that my comments below help the authors make some relatively minor improvements, particular in adding more biological interpretation in a few places to help more deeply understand the patterns they have identified.
--->
-
-<!-- R2: The authors have collected an interesting data set but I feel strongly that the work requires significant improvement to the Introduction (particularly the justification) and Discussion (particularly some explanations, even speculative) sections, as detailed below in my specific comments.
-
-Regarding the justification and conclusions, a quick scan of the literature revealed a study that used long(er) reads (covering SSU, ITS, LSU; Heeger et al.
-2019 in this journal) and also separated these prior to clustering and assignment.
-I therefore see the contribution as stated here (line 614) as overstated except that denoising was applied prior to clustering.
-Phylogenetic reconstruction was also used here, although the conclusions are not justified (see my comment below).
-I see an important contribution of this work as developing a new method to separate fragments; however, although it should be clear (i.e.
-examined in the Introduction) why existing approaches like ITSx or others are not suitable.
-
-"Long-read" has become a vague descriptor and should be clearly defined early on for this manuscript (in the abstract).
-How many bp or markers?
-I have seen the term used in metabarcoding to mean anything from a few bp longer than a MiSeq PE 2x250, to nearly 5000 bp covering multiple rRNA genes.-->
-
-<!-- R3: This paper analyzes a transect of soils in Africa and reports how results differ from 3 differing sequencing technologies and variations to bioinformatics platforms.
-With sequencing technologies changing so rapidly, and so many methods to choose from, I’m sure this paper would be of interest to the Molecular Ecology Resources audience, as it provides useful data comparing sequencing technologies.
-I congratulate the author on herculean efforts to write so many R packages, for making his packages and scripts publicly available, and for making a pipeline that allegedly improves ASV recovery for long amplicons.
-I personally would find the paper more interesting if a treatment other than spatial turnover had been applied to these samples, like before and after a disturbance, or a different host tree, and one could examine if the treatment of interest gave similar inferences with different sequencing technologies.
-However, I believe their claim that sequencing technologies yield similar results.
-I do think the paper would be improved by adding in some alpha diversity analyses, which are likely to be less robust to vagaries of sequencing technologies than beta diversity analyses.
-I also did not see any discussion of normalization for variation in number of sequences per sample, which needs to be accounted for.
-Instead of figure 2 (which can be moved to the supplements), I would prefer instead to see a figure about correlations of alpha diversity among methods.
-I also think species accumulation curves would be useful to see.
-I have also suggested several ways the writing could be slightly altered to make it clearer and more engaging for the reader.
-The results section is quite dry, reporting the statistics rather than reporting the interesting results and using the statistics to support it.
-For example, “Microbial communities are spatially autocorrelated (R2 = 0.90, p <0.05)” is more interesting to read than the “Bray-curtis dissimilarity showed…” If some of the results section could be re-written to focus on the results rather than reporting statistics, it would be more engaging.
-I also think the introduction needs to be edited to be more cohesive with the rest of the document.
-For example, only abbreviations were used in the introduction but they were fully spelled out at numerous points in the methods, and this made the introduction unnecessarily difficult to follow.
-I also found that the discussion repeated the results a lot and thus could be truncated or re-focused to set the results more in the context of the current literature (for example some citations of similar papers was missing).
-Finally, I will freely admit that some portions of the methods section, like the discussion of LSUx, went over my head and were very difficult for me to understand and/or follow.
-I’m not sure if someone more bioinformatically inclined could do a better job reviewing some of the more technical portions or if they just need to be written more clearly.
-I actually spent 3x the amount of time I usually devote to reviews on this paper, and I made a lot of efforts to provide constructive criticisms of how this paper could be improved to be clearer for the reader, so I hope that they are helpful.
-See specific line comments below.-->
-
 # Abstract {-}
@@
 but this variablity means that they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
-<!-- R2: This is true, but the fungi are a large and diverse group and thus it is no surprise that a highly variable rRNA marker can not be aligned.
-Perhaps this manuscript could discuss how meaningful a phylogenetic approach to analysis of complete fungal communities is anyway.
-Perhaps building a tree for each phylum is possible (and a realistic evolutionary model could be used).
-I think that this could be a contribution here - to explore how this variation could be dealt with using some knowledge of the clade "fungi".
-(Partly addressed?) -->
 Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa [@ihrmark2012;@tedersoo2015;@palmer2018].
@@
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, any coarse organic debris <!--Do you have any definition?
-Wider than 5 mm?
-MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade and homogenized.
+For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade and homogenized.
 A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (Supplementary methods).
@@
 # Results
-
-<!-- R2: The Results section has many instances of the data "...are shown in Fig/Table..." This is difficult to read and I strongly suggest you summarize the findings in the text.
-In some cases it simply means deleting these sentences and adding the Fig/Table reference after the following sentence.
-It would be more helpful to talk about the overall trends here rather than listing all of the % values.
--->
 
@@
 
-<!-- These are mostly  results, could you summarise them here (make sure all figures are cited in the results)? This would also address the reviewers comment. -->
 DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
-
 ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures\ \@ref(fig:venn) and \@ref(fig:alpha-compare)).
@@
 Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any reads from the short amplicon library which were longer than those recovered by Illumina and Ion Torrent.
-<!-- Next sentence (57 words) could be cut... -->
 Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
@@
 [@heeger2019] also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
-<!-- maybe the next 2 sentences could be dropped? 41 words-->
 Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, particularly the Warcup database, where it places <25% of ASVs even to kingdom.

### 2020-10-17 writing/transect_supplement.Rmd
Improve section on tree-rooting to match the current (improved) root.

@@
 
-### Rooting the fungal tree
-
-The phylogenetic tree was rooted outside the kingdom Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
-Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
-The kingdom Fungi was identified as the minimal clade containing all ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
-ASVs falling outside this clade were not included in downstream fungal community analysis.
-
 ### Uniform taxonomic classification for Unite, Warcup, and RDP
@@
 FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
+
+### Rooting the fungal tree
+
+The phylogenetic tree was rooted based on sequences which were confidently assigned to kingdom based on the primary taxonomic assignments.
+Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
+Assignments were considered as reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
+Two sequences belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.
+This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(read(sure_kingdoms), taxon) %>% glue::glue("{n} assigned to {taxon}, ", last = "and ")`.
+Inspection of the tree verified that these four kingdoms were reciprocally monophyletic, and the clade containing Alveolata and Viridiplantae was used as an outgroup to root the tree.
+The kingdom Fungi was then identified as the minimal clade containing all ASVs which were confidently identified  as kingdom Fungi.
+ASVs falling outside this clade, including the excluded *Tulasnella* sequences, were not included in downstream fungal community analysis.
 

### 2020-10-19 writing/transect_paper.Rmd

@@
 ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.
-Conflicts between the assignments made by the different algorithms were later resolved using a phylogenetic method and a consensus method, which are described below.
 
-For the ASVs from the long amplicon dataset, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax) and Supplementary Methods).
+For full-length long-amplicon ASVs, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax) and Supplementary Methods).
 This refinement algorithm is referred to as "PHYLOTAX", available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
 
-ASVs from the short amplicon datasets were refined using only the final strict consensus step, i.e., an assignment at a given rank was accepted if there was no conflict between the different assignment algorithms at greater than 50% confidence.
-This refinement method is referred to as "Consensus".
-Additionally, a hybrid method was applied to the short amplicon datasets, in which assignments from PHYLOTAX were used for ASVs which could be linked by an identical ITS2 region to a long amplicon, and assignments from Consensus were used for the remaining ASVs.
+ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the preliminary assignments.
+This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [LCAClassifier, @lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
+Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as "Consensus".
 

### 2020-10-22 writing/transect_paper.Rmd
radical reorganization of light plan and scripts directory.  everything works again.

@@
 library("phyloseq")
+library("glue")
 tryCatch(
@@
 
-if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))
+if (!exists("percent", mode = "function")) source(here::here("scripts", "functions_output.R"))
 
@@
 Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, *Uapaca* (Phyllanthaceae) and Fabaceae tr.
-Amherstieae [@brundrett2017a], which together represent approximately 60% of tree stems globally [@steidinger2019].
+Amherstieae [@brundrett2017a], together representing approximately 60% of tree stems globally [@steidinger2019].
 
 Although ECM fungi form many well-known mushrooms (e.g., *Amanita*, *Cantharellus*, *Boletus*), some instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
-Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by looking at vegetative structures [@horton2001].
+Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by sampling of vegetative structures [@horton2001].
 Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures in the field.
@@
 Due to length restrictions for Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
-The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer (hereafter, ITS4m).
+The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
 The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:regions)a), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
@@
   readd(amplicon_plot),
-  readd(length_diversity_plot),
+  readd(length_richness_plot),
   ncol = 1,
@@
 ```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", out.width = "70%", fig.height = 4, fig.align="center", results="asis"}
-venn_A <- venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
-  vennplot_data(ASVs) %>% {
-    plot(
-      select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-      legend = FALSE,
-      labels = FALSE,
-      quantities = .$label,
-      main = "a",
-      fills = c("grey85", "lightblue", "lightcoral"))
-  }
-
-venn_B <- venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
-  vennplot_data(OTUs) %>% {
-    plot(
-      select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-      legend = FALSE,
-      labels = FALSE,
-      quantities = .$label,
-      main = "b",
-      fills = c("grey85", "lightblue", "lightcoral")
-    )
-  }
-ggpubr::ggarrange(venn_A, venn_B, ncol = 2)
-
+ggpubr::ggarrange(readd(vennplot_tech_ASV), readd(vennplot_tech_OTU), ncol = 2)
 grid::grid.legend(
@@
 
-venn_C <- readd(asv_table) %>%
-  mutate_if(is.numeric, ~ . / sum(.)) %>%
-  mutate(long = pb_500,
-         short = rowMeans(.[,c("pb_483", "SH-2257", "is_057")])) %>%
-  select(seq, long, short) %>%
-  venndata(ASVs, cols = c("long", "short")) %>%
-  vennplot_data(ASVs) %>% {
-    plot(
-      select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-      legend = FALSE,
-      labels = FALSE,
-      quantities = .$label,
-      main = "c",
-      fills = c("white", "slateblue1"))
-  }
-
-venn_D <- readd(otu_table) %>%
-  mutate_if(is.numeric, ~ . / sum(.)) %>%
-  mutate(long = pb_500,
-         short = rowMeans(.[,c("pb_483", "SH-2257", "is_057")])) %>%
-  select(seq, long, short) %>%
-  venndata(OTUs, cols = c("long", "short")) %>%
-  vennplot_data(OTUs) %>% {
-    plot(
-      select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
-      legend = FALSE,
-      labels = FALSE,
-      quantities = .$label,
-      main = "d",
-      fills = c("white", "slateblue1")
-    )
-  }
-ggpubr::ggarrange(venn_C, venn_D, ncol = 2)
+ggpubr::ggarrange(readd(vennplot_amplicon_ASV), readd(vennplot_amplicon_OTU), ncol = 2)
 grid::grid.legend(
@@
 `r compR2("pb_500", "SH-2257", "ASV")` to `r compR2("pb_500", "SH-2257", "OTU")`, and
-`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short amplicon reads, respectively).
+`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs.
+PacBio, Illumina, and Ion Torrent short amplicon reads, respectively; Figure\ \@ref(fig:read-compare)).
 
 ASV richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(readd(demingfits_ASV), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
-The slope of the relationship between PacBio and Illumina richness estimates was not significantly different from 1, indicating that these two technologies give essentially identical rarefied richness estimates, dispite the approximately 200× difference in original sequencing depth.
-Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(readd(demingfits_ASV), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$value_x %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% unique() %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV accumulation curves (Figure\ \@ref(fig:accum-ASV)).
+The slope of the relationship between PacBio and Illumina richness estimates was not only slightly different from 1, indicating that these two technologies give highly comparable rarefied richness estimates, despite the approximately 200× difference in original sequencing depth.
+Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(readd(demingfits_ASV), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$slope %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% unique() %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV accumulation curves (Figure\ \@ref(fig:accum-ASV)).
 ASV richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figures\ \@ref(fig:alpha-compare)).
-Although the total least squares regression lines indicate that the long amplicon dataset resulted in richness estimates which were more similar to the short amplicon results from Ion Torrent than the other two technologies, the confidence intervals for all slope estimates overlap with each other and 1.
+Total least squares regression indicated that the long amplicon dataset resulted in richness estimates which were intermediate between the short amplicon results from Ion Torrent and the other two technologies.
 Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure\ \@ref(fig:venn)), sample-wise OTU accumulation curves (Figure\ \@ref(fig:accum-OTU)) and rarefied OTU richness relationships between sequencing strategies (Figure\ \@ref(fig:alpha-compare-otu)) were highly similar to those for ASVs.
@@
 Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction.
-Blue line is a total least squares fit, with the given slope (and 95% confidence interval) and R^2^ value.
-Dashed diagonal line indicates a slope of 1.
+Blue lines are total least squares fit with 95% confidence interval, with the given slope (and 95% confidence interval) and R^2^ value.
+Dashed diagonal line indicates ideal slope of 1.
 
@@
 
-For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be taxonomically identified than rare ASVs (Figure\ \@ref(fig:tax-chart)).
+For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be taxonomically identified than rare ASVs (Figure\ \@ref(fig:taxon-chart)).
 A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
@@
 
-(ref:tax-chart-s) Summary of taxonomic assignments
+(ref:taxon-chart-s) Summary of taxonomic assignments
 
-(ref:tax-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
+(ref:taxon-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
-These two methods are plotted in each column to compare with results for the individual databases.
 
-```{r tax-chart, fig.cap = "(ref:tax-chart)", fig.scap = "(ref:tax-chart-s)", out.width = "100%"}
-readd(tax_chart_plot)
+```{r taxon-chart, fig.cap = "(ref:taxon-chart)", fig.scap = "(ref:taxon-chart-s)", out.width = "100%"}
+readd(taxon_chart_plot)
 ```
@@
 ```{r fungi-comm, fig.cap = "(ref:fungi-comm-s). (ref:fungi-comm)", fig.scap = "(ref:fungi-comm-s)", fig.width=7}
-readd(taxon_reads) %>%
-  filter(
-    kingdom == "Fungi",
-    !(Algorithm == "Consensus" & region == "ITS"),
-    Algorithm == "PHYLOTAX"
-    ) %>%
-  group_by(tech, amplicon) %>%
-  mutate(reads = reads/sum(reads), ASVs = 1/n()) %>%
-  pivot_longer(cols = c("reads", "ASVs"), names_to = "type", values_to = "reads") %>%
-  bind_rows(
-    filter(., type == "ASVs", label == OTU) %>%
-      mutate(label = OTU) %>%
-      group_by(tech, amplicon) %>%
-      mutate(reads = 1/n(), type = "OTUs")
-  ) %>%
-  taxon_plot(
-    rank = class,
-    datasets = datasets,
-    cutoff = 0.02,
-    y = reads,
-    x = type,
-    facets = vars(tech, amplicon)
-  ) +
-  ylab("Fraction of Community") +
-  xlab(NULL)
+readd(taxon_plot_fungi_class)
 ```
@@
 However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
-Although the RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \@ref(fig:tax-chart)).
+Although the RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \@ref(fig:taxon-chart)).
 This is probably due to two factors.
@@
 AMPtk [@palmer2018] uses a strict consensus taxonomy between UTAX and SINTAX as an alternative when an initial BLAST search failed to give a hit with at least 97% sequence identity, but did not present results assessing the results of this approach.
-Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications (Figure \@ref(fig:tax-chart)).
+Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications (Figure \@ref(fig:taxon-chart)).
 Inconsistent family and genus level assignments are problematic because accurate assignment at these ranks is generally required for ecological guild assignment using FUNGuild.
@@
 For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses phylogenetic relationships to resolve these disagreements in a principled manner.
-For instance, `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
+The effect was most pronounced for the PacBio long amplicon dataset,
+where `r filter(readd(taxon_chart), tech == "PacBio", amplicon == "Long", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(taxon_chart), tech == "PacBio", amplicon == "Long", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
 This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure\ \@ref(fig:ecm)).
+For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs; this was the case for `r readd(taxon_reads_hybrid_phylotax) %>% filter(label %in% readd(phylotax_hybrid)$tree$tip.label, amplicon != "Long") %>% group_by(tech) %>% summarize_at("reads", sum) %$% paste(glue_collapse(percent(reads, format = "f", digits = 0), sep = ", ", last = ", and "), "of reads for the", glue_collapse(tech, sep = ", ", last = ", and "))` short amplicon datasets, respectively.
+Deeper long-amplicon seequencing would improve the
 

### 2020-10-22 writing/transect_supplement.Rmd
radical reorganization of light plan and scripts directory.  everything works again.

@@
 regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
-if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))
+if (!exists("percent", mode = "function")) source(here::here("scripts", "functions_output.R"))
 
@@
 
-**Supplementary file 3**: ML tree of all long metabarcoding ASVs (5.8S--ITS2--LSU) assigned to kingdom Fungi.
+**Supplementary file 3**: ML tree of all long metabarcoding ASVs (ITS1--5.8S--ITS2--LSU) assigned to kingdom Fungi.
 Thick branches represent rapid bootstrap values >90%.
@@
 
-## Sampling
+## Sampling {#Sampling-supplement}
 
@@
 
-## DNA extraction, amplification, and sequencing
+## DNA extraction, amplification, and sequencing {#DNA-supplement}
 
@@
 
-## Denoising and clustering
+## Denoising and clustering {#denoising-supplement}
 
@@
 
-## Phylogenetic inference and taxonomy assignment
+## Phylogenetic inference and taxonomy assignment {#phylogeny-supplement}
 
@@
 Two sequences belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.
-This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(read(sure_kingdoms), taxon) %>% glue::glue("{n} assigned to {taxon}, ", last = "and ")`.
+This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(readd(sure_kingdoms), taxon) %>% glue::glue_data("{n} assigned to {taxon}, ") %>% glue::glue_collapse(last = "and ")`.
 Inspection of the tree verified that these four kingdoms were reciprocally monophyletic, and the clade containing Alveolata and Viridiplantae was used as an outgroup to root the tree.
@@
 
-### Refining taxonomic assignments with a phylogenetic tree: PHYLOTAX
+### Refining taxonomic assignments with a phylogenetic tree: `PHYLOTAX`
 
@@
 
-## Effect of sequencing strategy on recovered community
+## Effect of sequencing strategy on recovered community {#sequencing-strategy-supplement}
 
@@
 
-## Spatiotemporal analysis
+## Spatiotemporal analysis {#spatiotemporal-supplement}
 
@@
   select = with(
-    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
+    as_tibble(phyloseq::sample_data(readd(buffer_asv_physeq))),
     buffer == "Xpedition" & year == "2015" 
@@
   readd(buffer_asv_pcoa),
-  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
+  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_physeq))),
                 buffer == "Xpedition" & year == "2016"
@@
   readd(buffer_asv_pcoa),
-  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
+  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_physeq))),
                 buffer == "LifeGuard" & year == "2016"
@@
 (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
-readd(big_seq_table_ITS2) %>%
-  colSums() %>%
-  enframe(name = "ITS2", value = "nreads") %>%
-  mutate(length = nchar(ITS2)) %>%
-  filter(length <= 140) %>%
-  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
-  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
-  group_by(tech, amplicon, Algorithm, reference) %>%
-  mutate(reads = reads/sum(reads)) %>%
-  ungroup() %>%
-  # group_by(ITS2) %>%
-  # filter(!"Short" %in% amplicon) %>%
-  taxon_plot(kingdom, y = reads, datasets = datasets)
+readd(taxon_plot_short_kingdom)
 ```
@@
 ```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
-  readd(taxon_reads) %>%
-  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
-  mutate_at("reference", replace_na, "All") %>%
-  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
-  theme(legend.position = "bottom", legend.title = element_blank())
+readd(taxon_plot_ecm)
 ```

### 2020-10-22 scripts/drake_plan_heavy.R

@@ -24,18 +24,7 @@ if (interactive()) {
 
 #### read scripts and configs ####
 # load various pipeline functions
-
-source(file.path(config$rdir, "dada.R"))
-source(file.path(config$rdir, "qstats.R"))
-source(file.path(config$rdir, "lsux.R"))
-source(file.path(config$rdir, "mantel.R"))
-source(file.path(config$rdir, "parallel_helpers.R"))
-source(file.path(config$rdir, "plate_check.R"))
-source(file.path(config$rdir, "raxml.R"))
-source(file.path(config$rdir, "taxonomy.R"))
-source(file.path(config$rdir, "utils.R"))
-source(file.path(config$rdir, "AlignSeqs.R"))
-source(file.path(config$rdir, "DECIPHER-utils.R"))
+for (f in list.files(config$rdir, "functions_.+.R", full.names = TRUE)) source(f)
 
 setup_log("drakeplan")
 

### 2020-10-22 scripts/functions_DECIPHERutils.R

@@ -1,0 +1,62 @@
+# fixes a bug in DECIPHER
+# contruibuted by Eric Scott Wright, author of DECIPHER
+
+.subset <- function(x, i) {
+	ans <- new(class(x))
+	g <- x@ranges@group[i]
+	u <- unique(g)
+	ans@pool <- x@pool[u]
+	ans@ranges@group <- match(g, u)
+	ans@ranges@start <- x@ranges@start[i]
+	ans@ranges@width <- x@ranges@width[i]
+	ans
+}
+
+.replace <- function(x, y, index) {
+	ans <- new(class(x))
+	ans@pool <- c(x@pool,
+		y@pool)
+	g <- .Call("getPools",
+		ans@pool@xp_list,
+		PACKAGE="DECIPHER")
+	w <- which(!duplicated(g))
+	ans@pool <- ans@pool[w]
+	m <- match(g, g[w])
+	ans@ranges@group <- x@ranges@group
+	ans@ranges@group[index] <- y@ranges@group + length(x@pool)
+	g <- m[ans@ranges@group]
+	u <- unique(g)
+	ans@pool <- ans@pool[u]
+	ans@ranges@group <- match(g, u)
+	ans@ranges@start <- x@ranges@start
+	ans@ranges@start[index] <- y@ranges@start
+	ans@ranges@width <- x@ranges@width
+	ans@ranges@width[index] <- y@ranges@width
+	ans
+}
+
+.append <- function(x, y) {
+	ans <- new(class(x))
+	ans@pool <- c(x@pool,
+		y@pool)
+	ans@ranges@group <- c(x@ranges@group,
+		y@ranges@group + length(x@pool))
+	ans@ranges@start <- c(x@ranges@start,
+		y@ranges@start)
+	ans@ranges@width <- c(x@ranges@width,
+		y@ranges@width)
+	ans
+}
+
+.switch <- function(x) {
+	if (class(x)=="DNAStringSet") {
+		ans <- new("RNAStringSet")
+	} else {
+		ans <- new("DNAStringSet")
+	}
+	ans@pool <- x@pool
+	ans@ranges@group <- x@ranges@group
+	ans@ranges@start <- x@ranges@start
+	ans@ranges@width <- x@ranges@width
+	ans
+}

### 2020-10-22 scripts/functions_heat_trees.R

@@ -1,0 +1,182 @@
+# functions for making heat trees
+# author Brendan Furneaux
+
+draw_full_heattree <- function(taxdata, outfile) {
+  set.seed(4)
+  theme_update(panel.border = element_blank())
+  taxdata %>%
+    metacoder::heat_tree(
+      layout = "davidson-harel",
+      # initial_layout = "reingold-tilford",
+      node_size = rowMeans(.$data$tax_asv[,-1]),
+      node_color = rowMeans(.$data$tax_asv[,-1]),
+      node_label = replace_na(taxon_names, "unknown"),
+      node_size_range = c(0.002, .035),
+      node_size_axis_label = "ASV count",
+      node_color_axis_label = "ASV count",
+      node_label_size_range = c(0.01, 0.03),
+      edge_size = rowMeans(.$data$tax_read[,-1]),
+      edge_size_trans = "linear",
+      edge_size_axis_label = "Read abundance",
+      edge_size_range = c(0.001, 0.03),
+      edge_color = rowMeans(.$data$tax_read[,-1]),
+      edge_color_axis_label = "Read abundance",
+      aspect_ratio = 9/5,
+      make_node_legend = FALSE,
+      make_edge_legend = FALSE,
+      output_file = outfile
+    )
+
+}
+
+log_read_ratio <- function(abund1, abund2) {
+  list(read_ratio = log10(mean(abund1) / mean(abund2)) %>% ifelse(is.nan(.), 0, .))
+}
+
+log_ASV_ratio <- function(abund1, abund2) {
+  list(ASV_ratio = log10(mean(abund1) / mean(abund2)) %>% ifelse(is.nan(.), 0, .))
+}
+
+draw_amplicons_heattree <- function(taxdata, outfile) {
+  taxdata2 <- taxdata
+  val_cols = names(taxdata2$data$tax_read) %>%
+    keep(str_detect, "reads_.+_.+_.+_Consensus")
+
+  taxdata2$filter_obs(
+    "tax_read",
+    rowSums(select_at(taxdata2$data$tax_read, val_cols)) > 0,
+    drop_taxa = TRUE,
+    drop_obs = TRUE,
+    supertaxa = TRUE,
+    reassign_obs = FALSE) %>%
+    invisible()
+
+  taxdata2$data$diff_read <- metacoder::compare_groups(
+    taxdata2,
+    data = "tax_read",
+    cols = val_cols,
+    groups = val_cols %>%
[…]

### 2020-10-22 scripts/functions_lsux.R

@@ -1,3 +1,7 @@
+# functions used in conjunction with LSUx
+# most of this has been moved to the LSUx package.
+# author Brendan Furneaux
+
 gather_regions <- function(pos) {
   starts <- dplyr::select(pos, -tidyselect::ends_with("_end"))
   starts <- tidyr::gather(starts, key = "region", value = "start", tidyselect::ends_with("_start"))

### 2020-10-22 scripts/functions_mantel.R

@@ -1,0 +1,21 @@
+# functions for mantel correlograms
+# author Brendan Furneaux
+
+correlog <- function(physeq, metric, timelag,
+                     break.pts = 0:13 - 0.5,
+                     cutoff = FALSE) {
+  dist_eco <- phyloseq::distance(physeq, method = metric)
+  dist_sp <- phyloseq::sample_data(physeq) %>%
+    with(x + 30000 * as.integer(site)) %>%
+    dist()
+  dist_t = phyloseq::sample_data(physeq) %>%
+    with(as.integer(year)) %>%
+    dist()
+  dist_spt = dist_sp + 100000 * (dist_t - timelag)
+  vegan::mantel.correlog(
+    dist_eco,
+    dist_spt,
+    break.pts = break.pts,
+    cutoff = cutoff
+  )
+}

### 2020-10-22 scripts/functions_parallel_helpers.R

@@ -1,4 +1,5 @@
 #### parallel setup ####
+# author Brendan Furneaux
 
 # are we running slurm?
 is_slurm <- function() nchar(Sys.which("sbatch")) > 0

### 2020-10-22 scripts/functions_phyloseq.R

@@ -1,0 +1,106 @@
+# functions to build phyloseq objects from subsets of other data
+# author Brendan Furneaux
+
+# rename the rows and columns of a sequence abundance table
+# convert rownames from a fileame to {Seq.Run}{Plate}{Well}
+# convert colnames from the full sequence to an 8-character hash.
+# also pools forward and reverse reads from same well.
+
+relabel_seqtable <- function(seqtable) {
+  seqtable %>%
+    # set the new column names
+    magrittr::set_colnames(tzara::seqhash(chartr("T", "U", colnames(.)))) %>%
+    # convert to a tibble for easier name column operations
+    tibble::as_tibble(rownames = "file") %>%
+    # parse the filename
+    tidyr::extract(
+      col = "file",
+      into = c("seq_run", "plate", "well", "direction", "region"),
+      regex = "([a-zA-Z]+[_-]\\d+)_(\\d+)_([A-H]1?\\d)([fr]?)_([a-zA-Z0-9]+).*") %>%
+    # create the ID
+    tidyr::unite("ID", seq_run, plate, well, sep = "") %>%
+    # remove unnecessary columns
+    dplyr::select(-direction, -region) %>%
+    # pool all reads from the same well
+    dplyr::group_by(ID) %>%
+    dplyr::summarise_all(sum) %>%
+    dplyr::ungroup() %>%
+    # convert back to a matrix
+    tibble::column_to_rownames("ID") %>%
+    as.matrix()
+}
+
+assemble_sample_data <- function(platemap, datasets) {
+  platemap %>%
+    dplyr::mutate_at("primer_pair", tolower) %>%
+    dplyr::left_join(
+      datasets %>%
+        dplyr::select(dataset, seq_run, tech, forward, reverse, amplicon) %>%
+        dplyr::mutate(
+          primer_pair = paste(
+            stringr::str_replace(forward, "_tag.*$", ""),
+            stringr::str_replace(reverse, "_tag.*$", ""),
+            sep = "_")
+        ),
+      by = "primer_pair") %>%
+    dplyr::mutate_at(c("year", "site", "qual", "plate", "well", "primer_pair"),
+                     factor) %>%
+    tidyr::unite("ID", seq_run, plate, well, sep = "", remove = FALSE) %>%
+    tibble::column_to_rownames("ID") %>%
[…]

### 2020-10-22 scripts/functions_platemap.R

@@ -1,3 +1,6 @@
+# function to read in, check, and format the platemap
+# author Brendan Furneaux
+
 read_platemap <- function(filename, sheet, skip = 2) {
   # read cells individually
   cells <- tidyxl::xlsx_cells(filename, sheet)
@@ -9,7 +12,7 @@ read_platemap <- function(filename, sheet, skip = 2) {
     {names(.)[which.max(.)]} %>%
     as.integer
   # read the data
-  raw <- readxl::read_excel(filename, sheet = sheet, 
+  raw <- readxl::read_excel(filename, sheet = sheet,
                             skip = header_line - 1,
                             col_names = TRUE, .name_repair = "minimal")
   # the data is not "tidy"; we have groups of columns for the different
@@ -59,7 +62,7 @@ read_platemap <- function(filename, sheet, skip = 2) {
                               TRUE ~ NA_character_)
     ) %>%
     # For some site/years, coordinates were measured from the edge of
-    # the plot rather than numbered from 1--25 
+    # the plot rather than numbered from 1--25
     dplyr::group_by(year, site) %>%
     dplyr::mutate(x = x - if (!any(is.na(x)) && max(x) > 25) 12 else 0) %>%
     dplyr::ungroup() %>%

### 2020-10-22 scripts/functions_raxml.R

@@ -1,5 +1,9 @@
+# function to interface with RAxML
+# author Brendan Furneaux, modified from raxml function in ips package version
+# 0.0.11 by C. Heibl
+
 #' Call raxml from R
-#' 
+#'
 #' @source modified from raxml function in ips package version 0.0.11 by C. Heibl
 #' #'@title Maximum Likelihood Tree Estimation with RAxML
 #'@description Provides an interface to the C program \bold{RAxML} (see
@@ -7,7 +11,7 @@
 #'  branch lengths, rapid and conventional non-parametric bootstrapping, mapping
 #'  splits onto individual topologies, and a lot more. See the RAxML manual for
 #'  details, especially if you are a new user of RAxML.
-#'@param RNAaln A matrix of RNA sequences of class 
+#'@param RNAaln A matrix of RNA sequences of class
 #'  \code{\link[Biostrings]{RNAMultipleStringSet}}.
 #'@param m A vector of mode \code{"character"} defining a model of molecular
 #'  evolution; currently only GTR model available.
@@ -92,21 +96,21 @@
 #'@seealso \code{\link{raxml.partitions}} to store partitioning information in a
 #'  data frame suitable for input as \code{partitions} argument in \code{raxml}.
 raxml_RNA <- function(RNAaln, m = "GTRCAT", f, N, p, b, x, k, S, A = "S16",
-                      weights, partitions, outgroup, backbone = NULL, 
+                      weights, partitions, outgroup, backbone = NULL,
                       file = paste0("fromR_", Sys.Date()), exec,
                       dir = tempdir(), threads){
-  
-  if (!inherits(RNAaln, "RNAMultipleAlignment")) stop("RNAaln is not of class 
+
+  if (!inherits(RNAaln, "RNAMultipleAlignment")) stop("RNAaln is not of class
                                                       'RNAMultipleAlignment'")
   if (!dir.exists(dir)) dir.create(dir, recursive = TRUE)
   olddir <- setwd(dir)
   on.exit(setwd(olddir))
-  
+
   # number of threads (PTHREADS only)
   # ---------------------------------
   if (!missing(threads))
     exec <- paste(exec, "-T", threads)
-  
+
   ## input file names
   ## ----------------
[…]

### 2020-10-22 scripts/functions_variogram.R

@@ -1,3 +1,7 @@
+# functions for spatiotemporal variograms
+# inspired by the gstat package, but ultimately totally reimplemented
+# author Brendan Furneaux
+
 variogram_table <- function(eco_dist, sp_dist, breaks) {
   tibble::tibble(
     dist = unclass(sp_dist),
@@ -134,4 +138,66 @@ variog_summary = function(model) {
     )
   ) %>%
     tidyr::unnest(confint)
+}
+
+fit_variogram <- function(variog) {
+    as_variogram(variog) %>%
+    stats::nls(
+      # this it reparameterized to give better convergence
+      # sill = 1 - (C0 + C1)
+      # nugget = C1 / (C0 + C1)
+      gamma ~ (1 - sill) - (1 - sill) * (1 - nugget) * exp(dist/range*log(0.05)),
+      data = .,
+      start = list(
+        nugget = min(.$gamma),
+        sill = 1 - max(.$gamma),
+        range = 30
+      ),
+      upper = list(nugget = 0.99, sill = 0.99, range = Inf),
+      lower = list(nugget = 0, sill = 0, range = 0.001),
+      algorithm = "port",
+      weights = pmin(.$np/.$dist, 1),
+      control = list(warnOnly = TRUE, maxiter = 1000, minFactor = 1/4096)
+    ) %>% {
+      list(
+        pars = as.list(.$m$getPars()),
+        summary = variog_summary(.),
+        predict = variog_predict(
+          .,
+          newdata = expand_grid(dist = seq(1, 25, 0.5), timelag = 0:1)
+        )
+      )
+    }
+}
+
+fit_variogramST <- function(variogST, variofit2) {
+  as_variogramST(variogST) %>%
+    stats::nls(
+      # this it reparameterized to give better convergence
+      # sill = 1 - (C0 + C1)
+      # nugget = C1 / (C0 + C1)
+      gamma ~ (1 - sill) - (1 - sill) * (1 - nugget) * exp((dist/range + timelag/timerange)*log(0.05)),
+      # add the parameters from the spatial-only fit.
+      data = .,
+      start = c(
+        list(timerange = 2),
+        variofit2$pars
+      ),
+      lower = c(
+        list(timerange = 0.001),
+        variofit2$pars
+      ),
+      algorithm = "port",
+      weights = pmin(.$np/.$dist, 1),
+      control = list(warnOnly = TRUE, maxiter = 1000, minFactor = 1/4096)
+    ) %>% {
[…]

### 2020-10-22 scripts/functions_venn.R

@@ -1,0 +1,147 @@
+# functions to make targets related to venn diagrams
+# author Brendan Furneaux
+
+# compile data to make a table with the information for a Venn diagram
+venndata <- function(data, vennvar, cols = names(data)[-1]) {
+  vennvar <- enquo(vennvar)
+  vennvarfrac <- paste0(as_label(vennvar), "_frac") %>% rlang::parse_quo(env = globalenv())
+  colpattern <- paste0("(", paste(cols, collapse = "|"), ")_(reads|", as_label(vennvar), ")")
+  data %>%
+    select_at(c("seq", cols)) %>%
+    mutate_at(cols, list(found = ~. > 0)) %>%
+    group_by_at(vars(ends_with("_found"))) %>%
+    summarize_at(cols, rlang::list2(!!vennvar := ~sum(. > 0), reads = sum)) %>%
+    ungroup() %>%
+    mutate_at(paste0(cols, "_found"), as.integer) %>%
+    full_join(
+      do.call(crossing, select_at(., paste0(cols, "_found"))),
+      by =  paste0(cols, "_found")
+    ) %>%
+    mutate_if(is.integer, replace_na, 0L) %>%
+    mutate_if(is.double, replace_na, 0L) %>%
+    tidyr::unite(col = "set", seq_along(cols), sep = "") %>%
+    filter(set != strrep(0, length(cols))) %>%
+    pivot_longer(
+      -1,
+      names_to = c("seq_run", "what"),
+      values_to = "value",
+      names_pattern = colpattern
+    ) %>%
+    pivot_wider(names_from = "what", values_from = "value") %>%
+    group_by(seq_run) %>%
+    mutate_if(is.numeric, list(frac = ~./sum(.))) %>%
+    bind_rows(
+      group_by(., seq_run) %>%
+        summarize_if(is.numeric, sum, na.rm = TRUE) %>%
+        mutate(set = "Total")
+    ) %>%
+    mutate_at(
+      vars(ends_with("_frac")),
+      na_if,
+      0
+    ) %>%
+    mutate_at(
+      vars(ends_with("_frac")),
+      formatC,
+      digits = 2,
+      format = "f",
+      drop0trailing = FALSE
+    ) %>%
+    mutate_at(
+      "reads",
+      k_or_M,
+      .sep = "",
+      .function = formatC,
+      format = "fg",
+      digits = 2
+    ) %>%
+    mutate(
[…]

### 2020-10-22 scripts/output_functions.R

@@ -1,580 +1,0 @@
-#### Output helper functions ####
-# These functions format in-line results or help to generate figures and tables
-# directly in the Rmarkdown for the paper, or in drake_plan_light.R
-
-# Format a number with "k" or "M" suffic for thousands or millions, as appropriate
-k_or_M <- function(x, ..., .sep = " ", .function = format) {
-  case_when(
-    abs(x) > 1e6 ~ paste(.function(x / 1e6, ...), "M", sep = .sep),
-    abs(x) > 1e3 ~ paste(.function(x / 1e3, ...), "k", sep = .sep),
-    is.na(x) ~ NA_character_,
-    TRUE ~ paste(.function(x, ...), " ", sep = .sep)
-  )
-}
-
-# Format a number as a percent
-percent <- function(x, ...) {
-  paste0(formatC(as.numeric(x) * 100, ...), "%")
-}
-
-# Format a vector of values to be used as in-line text
-text_list <- function(x, ...) {
-  x = format(x, ...)
-  glue::glue_collapse(x, sep = ", ", last = if (length(x) > 2) ", and " else " and ")
-}
-
-# compile data to make a table with the information for a Venn diagram
-venndata <- function(data, vennvar, cols = names(data)[-1]) {
-  vennvar <- enquo(vennvar)
-  vennvarfrac <- paste0(as_label(vennvar), "_frac") %>% rlang::parse_quo(env = globalenv())
-  colpattern <- paste0("(", paste(cols, collapse = "|"), ")_(reads|", as_label(vennvar), ")")
-  data %>%
-    select_at(c("seq", cols)) %>%
-    mutate_at(cols, list(found = ~. > 0)) %>%
-    group_by_at(vars(ends_with("_found"))) %>%
-    summarize_at(cols, rlang::list2(!!vennvar := ~sum(. > 0), reads = sum)) %>%
-    ungroup() %>%
-    mutate_at(paste0(cols, "_found"), as.integer) %>%
-    full_join(
-      do.call(crossing, select_at(., paste0(cols, "_found"))),
-      by =  paste0(cols, "_found")
-    ) %>%
-    mutate_if(is.integer, replace_na, 0L) %>%
-    mutate_if(is.double, replace_na, 0L) %>%
-    tidyr::unite(col = "set", seq_along(cols), sep = "") %>%
-    filter(set != strrep(0, length(cols))) %>%
-    pivot_longer(
-      -1,
[…]

### 2020-10-22 scripts/scratch.R

@@ -1426,6 +1426,76 @@ phylotax::relabel_tree(epa_decipher_full_graft, taxon_labels$label, paste0('"', 
   castor::write_tree("data/trees/labeled_epa_decipher.tree")
 
 
+
+max_cophenetic <- function(tree) {
+  max_coph <- max_depth <- max_length <- numeric(length(tree$tip.label) + tree$Nnode)
+  # tips have no depth, length, or cophenetic distance
+  max_coph[seq_along(tree$tip.label)] <- 0
+  max_depth[seq_along(tree$tip.label)] <- 0
+  max_length[seq_along(tree$tip.label)] <- 0
+
+  for (i in seq.int(length(max_coph), length(tree$tip.label) + 1, -1)) {
+    edges <- which(tree$edge[,1] == i)
+    if (length(edges) == 0) {
+      max_coph[i] <- max_length[i] <- max_depth[i] <- 0
+    } else {
+      max_length[i] <- max(max_length[tree$edge[edges,2]] + tree$edge.length[edges])
+      max_depth[i] <- max(max_depth[tree$edge[edges,2]] + 1)
+      if (length(edges) == 1) {
+        max_coph[i] <- max_coph[tree$edge[edges,2]]
+      } else {
+        max_coph[i] <- max(
+          max_coph[tree$edge[edges,2]],
+          sum(
+            sort(
+              max_length[tree$edge[edges,2]] + tree$edge.length[edges],
+              decreasing = TRUE
+            )[1:2]
+          )
+        )
+      }
+    }
+  }
+  tibble::tibble(
+    max_depth,
+    max_length,
+    max_coph
+  )
+}
+
+tip_depth <- function(tree) {
+  d <- integer(ape::Ntip(tree) + ape::Nnode(tree))
+  d[ape::Ntip(tree) + 1] <- 0
+  for (i in seq.int(ape::Ntip(tree) + 1, length(d))) {
+    edges <- which(tree$edge[,1] == i)
+    for (j in edges) {
+      d[tree$edge[j, 2]] <- d[i] + 1
+    }
+  }
+  d[1:ape::Ntip(tree)]
+}
+
+depth_order <- function(tree, reverse = FALSE) {
+  stopifnot(ape::is.binary(tree))
+  tree <- reorder(tree, "postorder")
+  mc <- max_cophenetic(tree)
+  kids <- lapply(
[…]

### 2020-10-22 scripts/taxonomy.R

@@ -1,125 +1,0 @@
-rank_factor <- function(r,
-                        ranks = c("rootrank", "domain", "kingdom", "phylum",
-                                  "class", "order", "family", "genus",
-                                  "species"),
-                        abbrev = FALSE) {
-  if (abbrev) {
-    factor(r, levels = substr(ranks, 1, 1), labels = ranks, ordered = TRUE)
-  } else {
-    factor(r, levels = ranks, ordered = TRUE)
-  }
-}
-
-# combine a named list of taxonomy tables
-combine_taxon_tables <- function(tables, allseqs) {
-  tibble::enframe(tables) %>%
-    tidyr::extract(
-      col = name,
-      into = c("region", "reference", "ref_region", "method"),
-      regex = "taxon_(32S|5_8S|ITS[12]?|LSU|long|short)_(unite|warcup|rdp_train)_(ITS[12]?|LSU)_(idtaxa|dada2|sintax)",
-      remove = FALSE
-    ) %>%
-    tidyr::unnest("value") %>%
-    dplyr::filter(
-      !is.na(rank),
-      !is.na(label),
-      taxon != "NA",
-      confidence >= 0.5,
-      !startsWith(taxon, "unidentified"),
-      !endsWith(taxon, "incertae_sedis"),
-      reference != "warcup" | rank != "kingdom"
-    ) %>%
-    dplyr::mutate_at("rank", rank_factor) %>%
-    dplyr::left_join(
-      dplyr::select(allseqs, label = "hash", n_reads = "nread") %>%
-        dplyr::group_by(label) %>%
-        dplyr::summarize(n_reads = sum(n_reads)) %>%
-        dplyr::ungroup(),
-      by = "label"
-    )
-}
-
-#### Find ASVs with consistent kingdom-level assignments
-# i.e., at least min_n assignments with greater than min_confidence
-# confidence, and also no conflicting assignments at any confidence level
-# they should also be present in the tree, and not one of the labels given
-# by "ignore"
-identify_taxa <- function(taxon_table, tree, rank, confidence, n,
-                          ignore = character()) {
-  taxon_table %>%
-    # look at kingdom assignments from the long amplicon dataset
-    # don't include tulasnella
-    dplyr::filter(
-      rank == !!rank,
-      label %in% tree$tip.label,
-      !label %in% ignore
-    ) %>%
-    dplyr::group_by(label) %>%
[…]

### 2020-10-22 writing/transect_paper.Rmd
minor fixes and changes

@@
 We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
-We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1-3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1-4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
+We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
 We used this method, implemented in the new R packages `LSUx` (https://github.com/brendanf/LSUx) and `tzara` (https://github.com/brendanf/tzara) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.
@@
 (ref:region-asv) **(a)** Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions.
-D1-3 represent the first three variable regions in LSU, while LSU1-4 represent the conserved regions.
+D1--3 represent the first three variable regions in LSU, while LSU1--4 represent the conserved regions.
 Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces.
@@
 Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
+Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of reads from the long-amplicon PacBio dataset.
+Mapping identical ITS2 ASVs from the short and long amplicon datasets allowed `r readd(bioinf_table)[c("long", "ITS2"), c("Ion Torrent_Short_reads", "Illumina_Short_reads", "PacBio_Short_reads")] %>% {glue_collapse(percent(.["long",]/.["ITS2",], format = "d", digits = 0), ", ", last = ", and ")}` of reads from the Ion Torrent, Illumina, and PacBio short amplicon datasets, respectively, to be assigned to a long amplicon ASV (Table \@ref(tab:bioinfo)).
 
@@
 
+; this was the case for `r readd(taxon_reads_hybrid_phylotax) %>% filter(label %in% readd(phylotax_hybrid)$tree$tip.label, amplicon != "Long") %>% group_by(tech) %>% summarize_at("reads", sum) %$% paste(glue_collapse(percent(reads, format = "f", digits = 0), sep = ", ", last = ", and "), "of reads for the", glue_collapse(tech, sep = ", ", last = ", and "))` short amplicon datasets, respectively.
+
 (ref:taxon-chart-s) Summary of taxonomic assignments
@@
 
-Taxonomic composition of the sequenced soil fungal community at the class level is is summarized in Figure\ \@ref(fig:fungi-comm) and as a heat tree [@foster2017] in Figure\ \@ref(fig:taxa).
+Taxonomic composition of the sequenced soil fungal community at the class level is summarized in Figure\ \@ref(fig:fungi-comm) and as a heat tree [@foster2017] in Figure\ \@ref(fig:taxa).
 The ML tree for fungal ASVs, along with taxonomic assignments, is shown in Supplementary File 3.
@@
 The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
-Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae and Pyronemataceae in the long amplicon dataset, and *Myerozyma* in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
+Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae (Agaricomycetes) and Pyronemataceae (Pezizomycetes) in the long amplicon dataset, and *Myerozyma* (Saccharomycetes) in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
 
@@
 
-The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from 13-32m for the total fungal community and 13-41m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit).
+The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "All Fungi") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the total fungal community and `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "ECM") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit).
 However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies, so no strong conclusion of variability between methods can be drawn.
@@
 The alternate LR5-F primer [@tedersoo2008] would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
-We assert that more complete coverage of ECM fungal groups with high rDNA variability such as *Tulasnella* (and ideally other *Cantharellales*) is worth the read-depth spent on non-target groups.
+We assert that, for studies targeting ECM fungi in particular, more complete capture of groups with high rDNA variability such as *Tulasnella* (and ideally other Cantharellales) is worth the read-depth spent on non-target groups.
 
@@
 This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure\ \@ref(fig:ecm)).
-For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs; this was the case for `r readd(taxon_reads_hybrid_phylotax) %>% filter(label %in% readd(phylotax_hybrid)$tree$tip.label, amplicon != "Long") %>% group_by(tech) %>% summarize_at("reads", sum) %$% paste(glue_collapse(percent(reads, format = "f", digits = 0), sep = ", ", last = ", and "), "of reads for the", glue_collapse(tech, sep = ", ", last = ", and "))` short amplicon datasets, respectively.
-Deeper long-amplicon seequencing would improve the
+For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs (last row of Table\ \@ref(tab:bioinfo)).
+Deeper long-amplicon sequencing would improve the coverage of long amplicons, allowing a greater fraction of short amplicon ASVs to also be placed phylogenetically.
 
@@
 The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
-This corresponds to a space-for-time substitution rate of 6.15\ m/year for the total fungal community and 3.83\ m/year for the ECM community.
+This corresponds to a space-for-time substitution rate of `r readd(variofit_table) %>% select(-variofit, -(std.error:conf.high)) %>% pivot_wider(names_from = term, values_from = estimate) %>% filter(timelag == 1, mantel == TRUE) %>% mutate(spacetime = range / timerange) %>% arrange(guild) %$% glue_collapse(format(spacetime, format = "f", digits = 2), " and ")`\ m/year for the total fungal community and ECM community, respecitvely.
 In a recent study, @kivlin2020 reported a space-for-time substitution rate of `r format(1.5/6.8 * 365, digits = 2)`\ m/year (reported as 6.8\ d / 1.5\ m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
@@
 
-Neither datasets from the short amplicon library showed temporal autocorrelation, but a weak negative spatiotemporal correlation was detected at 1\ m and 1\ year, which was probably a statistical artifact.
+Neither dataset from the short amplicon library showed temporal autocorrelation, but a weak negative spatiotemporal correlation was detected at 1\ m and 1\ year, which was probably a statistical artifact.
 The correlograms for the long amplicon datasets showed similar trends.
@@
 This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
-This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4-0.6) as compared to Bray-Curtis (0.8-1.0) in Figure \@ref(fig:spatio).
+This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4--0.6) as compared to Bray-Curtis (0.8--1.0) in Figure \@ref(fig:spatio).
 UniFrac analysis would be more suited at larger spatial scales and/or larger ecological gradients.

### 2020-10-22 writing/transect_supplement.Rmd
minor fixes and changes

@@
 Thick branches represent rapid bootstrap values >90%.
-Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, and Unite, Warcup, and RDP reference databases. 
+Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, with Unite, Warcup, and RDP reference databases. 
 Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
 Taxon names in red were reconstructed as polyphyletic.
-In some cases, this may be due to a poorly identified sequence "splitting" the group (e.g., the single *Aspergillus* sequence nested in *Monascus*, Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).
+In some cases, this may be due to a poorly identified sequence "splitting" the group (e.g., the single *Monascus* sequence nested in *Aspergillus*; Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).
 
@@
 Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in Soudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
 At each site, 25\ soil samples were collected along a single 24 m linear transect at intervals of 1\ m in May 2015.
@@
 Wider than 5 mm?
-MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
+MR-->was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with a sterilized knife blade.
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
@@
 The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:regions),
-where LSU1-4 represent the conserved domains of LSU flanking the variable D1-3 domains [@michot1984].
+where LSU1--4 represent the conserved domains of LSU flanking the variable D1--3 domains [@michot1984].
 For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
@@
 **short**, **ITS**, **LSU**, **long**: ITS2-based ASVs mapped to consensus ASVs for longer regions, where \"short\" and \"long\" denote the full-length short and long amplicons, respectively.
+*Italicized* numbers represent mapping via ITS2 ASVs to amplicons originally sequenced by another technology.
 
 ```{r bioinfo, cache=TRUE, extra.cache=gethash("bioinf_table", "bioinf_header")}
-options(knitr.kable.NA = '–')
-
 readd(bioinf_table) %>%
+  rownames_to_column("step") %>%
+  mutate_if(is.numeric, prettyNum, big.mark = ",") %>%
+  mutate_all(str_replace, "NA", "–") %>%
+  mutate_at(
+    vars(contains("Short")),
+    kableExtra::cell_spec,
+    italic = .$step %in% c("ITS", "LSU", "long"),
+    escape = FALSE
+  ) %>%
+  mutate_at(
+    vars(contains("Long")),
+    kableExtra::cell_spec,
+    italic = .$step %in% c("short"),
+    escape = FALSE
+  ) %>%
+  column_to_rownames("step") %>%
   set_names(gsub(".*_", "", names(.))) %>%
@@
     align = "rrrrrr",
-    format.args = list(big.mark = ","),
     linesep = c(
@@
     is.na(estimate) | is.na(conf.high) | mantel == FALSE,
-    "-",
+    "--",
     sprintf(
@@
     id_cols = c("guild", "amplicon", "tech", "metric", "algorithm"),
-    values_fill = list(value = "-")
+    values_fill = list(value = "--")
   ) %>%

### 2020-10-22 scripts/functions_ENA.R

@@ -178,16 +178,15 @@ compile_blank_samples <- function(sample_data) {
   )
 }
 
-write_ENA_samples <- function(soil_samples, template_file, output_file) {
+write_ena_samples <- function(samples, template_file, output_file) {
   template_names <- names(read_tsv(template_file, skip = 2))
   outdir <- dirname(output_file)
   if (!dir.exists(outdir)) {
     dir.create(outdir, recursive = TRUE)
   }
   file.copy(template_file, output_file, overwrite = TRUE)
-  soil_samples %>%
-    assertr::verify(template_names %in% names(.)) %>%
-    select_at(template_names) %>%
+  checkmate::assert_names(names(samples), must.include = template_names)
+  select_at(samples, template_names) %>%
     write_tsv(output_file, append = TRUE, col_names = FALSE)
 }
 

### 2020-10-22 scripts/functions_qstats.R

@@ -155,7 +155,7 @@ compile_bioinf_table <- function(rawcounts, demuxcounts, filtercounts_full,
       type = factor(type, c("ASVs", "reads"))
     ) %>%
     arrange(tech, amplicon, type) %>%
-    filter(ifelse(amplicon == "Short", !step  %in% c("ITS", "LSU", "long"), step != "short")) %>%
+    # filter(ifelse(amplicon == "Short", !step  %in% c("ITS", "LSU", "long"), step != "short")) %>%
     pivot_wider(
       id_cols = "step",
       names_from = c("tech", "amplicon", "type"),

### 2020-10-23 writing/transect_paper.Rmd
version as submitted

@@
 Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
-Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018], or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases [@fouquier2016].
+Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018], or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases [@fouquier2016], but these approaches are only applicable to sequences of known taxonomic affiliation.
 
 Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
-Although the public INSDC databases are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
+Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
 Consequently taxonomic assignments based on these databases may be incorrect or inconsistent.
@@
 The Unite database is a more curated attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
-The Ribosomal Data Project [RDP, @cole2014] hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences.
-The Warcup ITS database contains 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set contains 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [@liu2012].
+The Ribosomal Data Project [RDP, @cole2014] hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
+the Warcup ITS database, containing 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [@liu2012].
 Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.
@@
 
-Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
+Recent long-read HTS technologies such as Pacific Biosciences (PacBio) Single Molecule Real Time (SMRT) sequencing  enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
 This can improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis [@Tedersoo2018].
@@
 Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
-Turnover scale varies between different ecosystems and taxonomic groups.
-Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.
-distance [@legendre2012].
+Turnover scale varies between different ecosystems and taxonomic groups,
+and can be measured by the range at which a Mantel correlogram indicates significant autocorrelation,
+or by fitting a function to an empirical distance-decay curve of community dissimilarity vs. distance [@legendre2012].
 
 We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
-and reference databases (Unite, Warcup, RDP),
-and with non-phylogenetic and phylogenetic community distance measures.
+with three different reference databases (Unite, Warcup, RDP),
+and both non-phylogenetic and phylogenetic community distance measures.
 We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
@@
 
-Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in Soudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25\ soil samples were collected along a single 24 m linear transect at intervals of 1\ m in May 2015.
+Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
+Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade and homogenized.
-A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (Supplementary methods).
+For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with a sterilized knife blade and homogenized.
+A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary methods).
 
-## DNA extraction, amplification, and sequencing
+## DNA amplification and sequencing
 
 DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
-Due to length restrictions for Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
+Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
 The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
@@
 
-Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
+Each library was sequenced on a PacBio RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
 for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752\ 37 UPPSALA, Sweden).
 Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
-
 Additionally, the same short amplicon PCR libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept.
@@
 
-All amplicons were denoised using DADA2 version `r packageVersion("dada2")`a ccording to the ITS pipeline workflow [@callahan2016;@callahan2020], with technology-specific modifications for Ion Torrent [@callahan2020b] and PacBio [@callahan2019].
+All amplicons were denoised using DADA2 version `r packageVersion("dada2")` according to the ITS pipeline workflow [@callahan2016;@callahan2020], with technology-specific modifications for Ion Torrent [@callahan2020b] and PacBio [@callahan2019].
 Although this was successful for the short amplicons on all technologies, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` ASVs were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
@@
 
-Full length long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10\ iterations of progressive alignment and conserved RNA secondary structure calculation and 10\ refinement iterations.
+Full length long amplicon ASVs were aligned using DECIPHER [@wright2015] with up to 10\ iterations of alternating progressive alignment and conserved RNA secondary structure calculation, followed by 10\ refinement iterations.
 This alignment was truncated at a position after the D3 region corresponding to base 907 of the *Saccharomyces cerevisiae* S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by @holst-jensen1999.
@@
 Each of these distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
-Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
+Separate correlograms were drawn for samples taken during the same year and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
 
@@
 Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
-Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of reads from the long-amplicon PacBio dataset.
-Mapping identical ITS2 ASVs from the short and long amplicon datasets allowed `r readd(bioinf_table)[c("long", "ITS2"), c("Ion Torrent_Short_reads", "Illumina_Short_reads", "PacBio_Short_reads")] %>% {glue_collapse(percent(.["long",]/.["ITS2",], format = "d", digits = 0), ", ", last = ", and ")}` of reads from the Ion Torrent, Illumina, and PacBio short amplicon datasets, respectively, to be assigned to a long amplicon ASV (Table \@ref(tab:bioinfo)).
+Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of denoised ITS2 reads from the long-amplicon PacBio dataset.
+Mapping identical ITS2 ASVs from the short and long amplicon datasets allowed `r readd(bioinf_table)[c("long", "ITS2"), c("Ion Torrent_Short_reads", "Illumina_Short_reads", "PacBio_Short_reads")] %>% {glue_collapse(percent(.["long",]/.["ITS2",], format = "d", digits = 0), ", ", last = ", and ")}` of denoised reads from the Ion Torrent, Illumina, and PacBio short amplicon datasets, respectively, to be assigned to a long amplicon ASV (Table \@ref(tab:bioinfo)).
 
@@
 ASV richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(readd(demingfits_ASV), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
-The slope of the relationship between PacBio and Illumina richness estimates was not only slightly different from 1, indicating that these two technologies give highly comparable rarefied richness estimates, despite the approximately 200× difference in original sequencing depth.
+The slope of the relationship between PacBio and Illumina richness estimates was only slightly different from 1, indicating that these two technologies give highly comparable rarefied richness estimates, despite the approximately 200× difference in original sequencing depth.
 Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(readd(demingfits_ASV), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$slope %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% unique() %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV accumulation curves (Figure\ \@ref(fig:accum-ASV)).
-ASV richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figures\ \@ref(fig:alpha-compare)).
+ASV richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
 Total least squares regression indicated that the long amplicon dataset resulted in richness estimates which were intermediate between the short amplicon results from Ion Torrent and the other two technologies.
@@
 
-```{r accum-ASV, fig.cap="ASV accumulation curves"}
+(ref:accum-ASV-s) ASV accumulation curves
+
+(ref:accum-ASV) Each curve represents rarefaction of a single sample or pooled pair of samples.
+Points at the end of each curve represent the actual read depth and observed ASV richness.
+Note logarithmic axes.
+
+```{r accum-ASV, fig.cap="(ref:accum-ASV-s). (ref:accum-ASV)", fig.scap = "(ref:accum-ASV-s)"}
 readd(accum_plot_ASV)
@@
 
-For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads was assigned than of ASVs, indicating that common ASVs were more likely to be taxonomically identified than rare ASVs (Figure\ \@ref(fig:taxon-chart)).
+For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads were assigned than of ASVs, indicating that common ASVs were more likely to be taxonomically identified than rare ASVs (Figure\ \@ref(fig:taxon-chart)).
 A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
 At most taxonomic ranks, the RDPC algorithm assigned the greatest fraction of reads and ASVs, followed by SINTAX, and then IDTAXA.
-
-; this was the case for `r readd(taxon_reads_hybrid_phylotax) %>% filter(label %in% readd(phylotax_hybrid)$tree$tip.label, amplicon != "Long") %>% group_by(tech) %>% summarize_at("reads", sum) %$% paste(glue_collapse(percent(reads, format = "f", digits = 0), sep = ", ", last = ", and "), "of reads for the", glue_collapse(tech, sep = ", ", last = ", and "))` short amplicon datasets, respectively.
 
@@
 
-(ref:fungi-comm) Values represent the fraction of all ASVs, OTUs, or reads which were assigned to kingdom Fungi.
+(ref:fungi-comm) Values represent the fraction of all ASVs and reads which were assigned to kingdom Fungi.
 Assignments based on PHYLOTAX.
-Classes which represented less than 2% of reads, OTUs, and ASVs in all datasets are grouped together as "other".
+Classes which represented less than 2% of reads and ASVs in all datasets are grouped together as "other".
 
@@
 Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
-The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm), \@ref(fig:ppcoa), and \@ref(fig:heattree-length-compare)).
+The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm), \@ref(fig:heattree-length-compare), and \@ref(fig:ppcoa)).
 Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae (Agaricomycetes) and Pyronemataceae (Pezizomycetes) in the long amplicon dataset, and *Myerozyma* (Saccharomycetes) in the short amplicon datasets (Figures\ \@ref(fig:heattree-length-compare) and \@ref(fig:ecm-heattree)).
@@
 This spatiotemporal correlation did not extend to a range of 1\ m, and in fact correlation was negative at a time lag of 1\ year and distance of 1\ m, indicating that samples collected 1\ m apart in different years were more different than randomly selected pairs of samples.
-This puzzling negative correlation reached statistical significance in the PacBio short amplicon dataset, and will be addressed further in the discussion.
+This negative correlation, which reached marginal statistical significance in the PacBio short amplicon dataset, was probably a statistical artifact.
 
@@
 
-The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "All Fungi") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the total fungal community and `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "ECM") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit).
+The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "All Fungi") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the total fungal community and `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "ECM") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
 However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies, so no strong conclusion of variability between methods can be drawn.
@@
 OTU read counts were strongly correlated between technologies (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon == y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
-and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
-which we interpret as supporting the validity of abundance-based beta diversity measures in soil fungi (Figure \@ref(fig:read-compare)b).
+and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`, Figure \@ref(fig:read-compare)b),
+which we interpret as supporting the validity of abundance-based beta diversity measures in metabarcoding [but see @castano2020 and references therein].
 
-DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
+DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of OTUs recovered from the same library by different technologies and the read counts for each OTU (Figure\ \@ref(fig:read-compare)).
 ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures\ \@ref(fig:venn) and \@ref(fig:alpha-compare)).
-This may be a result of the lower fraction of very high quality reads (Figure \@ref(fig:qual-check)).
-We used options for DADA2 intended to improve performance on technologies with higher rates of homopolymer indel errors [@callahan2020b], but our results suggest that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed [@callahan2016].
+This may be a result of the lower fraction of very high quality reads in the Ion Torrent dataset (Figure \@ref(fig:qual-check)).
+We used options for DADA2 intended to improve performance on technologies, like Ion Torrent, with higher rates of homopolymer indel errors [@callahan2020b], but our results suggest that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed [@callahan2016].
 
-Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any reads from the short amplicon library which were longer than those recovered by Illumina and Ion Torrent.
-Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
-Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
+Although the longer read length capabilities of PacBio allow recovery of longer ITS2 sequences than the other two technologies, as has recently been demonstrated in mock communities [@castano2020], in our dataset from a natural community PacBio did not recover any reads from the short amplicon library which were longer than those recovered by Illumina and Ion Torrent.
+Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observations by BF and NSY), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
+Contrary to expectations, Illumina showed a slightly *higher* fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).
 
@@
 First, the Warcup database does not include any non-fungi, so it cannot place any non-fungal sequences.
-Second, due to its low-density coverage of the fungal kingdom (18\ 000 vs.
+Second, due to its low-density coverage of the fungal kingdom (18\ 000 sequences vs.
 800\ 000 for Unite), it is likely that many ITS sequences, especially from uncultured tropical soil fungi, have no close match in the Warcup database, and so cannot be placed.
 The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
-[@heeger2019] also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
-Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, particularly the Warcup database, where it places <25% of ASVs even to kingdom.
-This is expected given its more conservative assignment of confidence scores [@murali2018a].
+@heeger2019 also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
+Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, as expected given its more well-calibrated and conservative confidence scores [@murali2018a], but this was particularly dramatic when paired with the Warcup database, where IDTAXA placed <25% of ASVs even to phylum.
 
@@
 Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications (Figure \@ref(fig:taxon-chart)).
-Inconsistent family and genus level assignments are problematic because accurate assignment at these ranks is generally required for ecological guild assignment using FUNGuild.
+Inconsistent family and genus level assignments are particularly problematic because accurate assignment at these ranks is generally required for ecological guild assignment using FUNGuild.
 
@@
 Similarly, @pickles2012a found autocorrelation at distances <3.4\ m based on T-RFLP analysis.
-Previous work in Miombo woodland, a similar ecosystem to the Soudanian woodland in this study, found autocorrelation at ranges <10\ m using Sanger sequencing of ECM root tips [@tedersoo2011], which was their smallest distance class.
+Previous work in Miombo woodland, a similar ecosystem to the Sudanian woodland in this study, found autocorrelation at ranges <10\ m using Sanger sequencing of ECM root tips [@tedersoo2011], which was their smallest distance class.
 
@@
 Year-to-year variation in ECM fungal communities, which we sampled, has been shown to be less than intra-annual variation [@bahram2015], as sampled by @kivlin2020.
-
-Neither dataset from the short amplicon library showed temporal autocorrelation, but a weak negative spatiotemporal correlation was detected at 1\ m and 1\ year, which was probably a statistical artifact.
-The correlograms for the long amplicon datasets showed similar trends.
+Neither dataset from the short amplicon library showed significant temporal autocorrelatio.
 
@@
 Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m; Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship.
-The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m, indicating little evidence of spatial structure.
+The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--`r round(filter(readd(variofit_table), metric == "W-UniFrac", guild == "All Fungi", timelag == 0)$conf.high, -1)`\ m, indicating little evidence of spatial structure.
 This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
@@
 The choice of amplicon and sequencing technology did not affect the results of the spatial analysis, provided sufficient sequencing depth.
-However, the addition of long amplicon reads did allow the construction of a phylogenetic tree from the metabarcoding reads, which allowed refinement of taxonomic assignments using our new tool PHYLOTAX.
+Alpha diversity estimates were strongly correlated between methods, but somewhat inflated for Ion Torrent relative to the other technologies.
+However, the addition of long amplicon reads did allow the construction of a phylogenetic tree directly from the metabarcoding reads, which allowed refinement of taxonomic assignments using our new tool PHYLOTAX.
 DADA2 ASV yield was initially poor for long amplicons, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
@@
 
-The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure.
-Work performed at NGI / Uppsala Genome Center has been funded by RFI / VR and and Science for Life Laboratory, Sweden.
+The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure, funded by RFI / VR and and Science for Life Laboratory, Sweden.
 
-Sequencing was performed by the SNP&SEQ Technology Platform in Uppsala.
+Sequencing was also performed by the SNP&SEQ Technology Platform in Uppsala.
 The facility is part of the National Genomics Infrastructure (NGI) Sweden and Science for Life Laboratory.

### 2020-10-23 writing/transect_supplement.Rmd
version as submitted

@@
   external = TRUE,
-  out.width = "70%",
+  out.width = "100%",
   fig.align = "center"
@@
 Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25\ soil samples were collected along a single 24 m linear transect at intervals of 1\ m in May 2015.
+At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
@@
 After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
-Samples preserved using LifeGuard were first centrifuged at 10000\ g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
-DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
+Samples preserved using LifeGuard were first centrifuged at 10\ 000\ g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
+DNA was quantified fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
 
-The forward primers for the short amplicon included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
+For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
 Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
@@
 
-Both forward and reverse primers for the long amplicon included indexes for combinatorical multiplexing (Supplementary File 2).
+For the long amplicon, both forward and reverse primers included indexes for combinatorical multiplexing (Supplementary File 2).
 PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
@@
 
-### Denoising long-reads: the numbers are against us
+### Denoising long-reads: the numbers are against us {#long-numbers}
 
@@
 Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
-For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
-Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
+For the combination of long sequences (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = " ")`.
+Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = " ")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
 In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
 
-In order to bypass this limitation, we implemented a workflow which divides the long rDNA amplicon into homologous domains (R package `LSUx`), denoises the domains individually using DADA2, and then recombines the denoised domains to form full-length denoised amplicons (R package `tzara`).
+In order to bypass this limitation, we implemented a workflow which divides the long rDNA amplicon into homologous domains (new R package `LSUx`), denoises the domains individually using DADA2, and then recombines the denoised domains to form full-length denoised amplicons (new R package `tzara`).
 The steps of this workflow are detailed below.
@@
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
-Assignments were considered as reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
-Two sequences belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.
-This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(readd(sure_kingdoms), taxon) %>% glue::glue_data("{n} assigned to {taxon}, ") %>% glue::glue_collapse(last = "and ")`.
+Assignments were considered reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
+Two ASVs belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.
+This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(readd(sure_kingdoms), taxon) %>% glue::glue_data("{n} assigned to {taxon}") %>% glue::glue_collapse(", ", last = ", and ")`.
 Inspection of the tree verified that these four kingdoms were reciprocally monophyletic, and the clade containing Alveolata and Viridiplantae was used as an outgroup to root the tree.
@@
 
-# Supplementary Figures and Tables
+\newpage
+
+# Supplementary Figures
+
+\listoffigures
 
@@
 Node\ 3 is assigned to Tax2 because this is consistent with at least one of the assignments for both B and C.
-Node\ 4 (and thus also node\ 5) is assigned to Tax1 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.
+Node\ 4 (and thus also node\ 5) is assigned to Tax1 because this is consistent with at least one of the assignments for D and F, and because E, which is nested within the clade defined by D and F, is completely unassigned.
 
@@
 
-```{r accum-OTU, fig.cap="OTU accumulation curves"}
+(ref:accum-OTU-s) OTU accumulation curves
+
+(ref:accum-OTU) Each curve represents rarefaction of a single sample or pooled pair of samples.
+Points at the end of each curve represent the actual read depth and observed OTU richness.
+Note logarithmic axes.
+
+```{r accum-OTU, fig.cap="(ref:accum-OTU-s). (ref:accum-OTU)", fig.scap = "(ref:accum-OTU-s)"}
 readd(accum_plot_OTU)
@@
 
+\FloatBarrier
+
+# Supplementary tables
+
+\listoftables
+
+\newpage
+
+\FloatBarrier
+
 ```{r region-limits}
@@
 ```
+
+\FloatBarrier

### 2020-10-23 scripts/drake_plan_light.R

@@ -242,7 +242,7 @@ plan2 <- drake_plan(
       tree = rooted_tree,
       old = taxon_labels$old,
       new = taxon_labels$new,
-      chimeras = allchimeras_ITS2
+      drop = allchimeras_ITS2
     ) %T>%
     castor::write_tree(file_out(
       !!glue::glue("data/trees/labeled.tree", group = group)
@@ -1811,19 +1811,19 @@ plan2 <- drake_plan(
     }
   ),
 
-  supplement_diff_pdf = withr::with_dir(
-    "writing",
-    {
-      file_in(!!file.path("writing", "transect_supplement_diff.tex"))
-      file_in(!!file.path("writing", "transect_paper_diff.tex"))
-      file_out(!!file.path("writing", "transect_supplement_diff.pdf"))
-      tinytex::xelatex(
-        "transect_supplement_diff.tex",
-        bib_engine = "biber",
-        clean = FALSE
-      )
-    }
-  ),
+  # supplement_diff_pdf = withr::with_dir(
+  #   "writing",
+  #   {
+  #     file_in(!!file.path("writing", "transect_supplement_diff.tex"))
+  #     file_in(!!file.path("writing", "transect_paper_diff.tex"))
+  #     file_out(!!file.path("writing", "transect_supplement_diff.pdf"))
+  #     tinytex::xelatex(
+  #       "transect_supplement_diff.tex",
+  #       bib_engine = "biber",
+  #       clean = FALSE
+  #     )
+  #   }
+  # ),
 
   article_pdf = withr::with_dir(
     "writing",
@@ -1963,7 +1963,7 @@ plan2 <- drake_plan(
   taxid_all = target(
     {
       taxids <- bind_rows(readd(taxid))
-      write_csv(file_out(!!sprintf("output/%s_taxa.csv", taxon_type)))
+      write_csv(taxids, file_out(!!sprintf("output/%s_taxa.csv", taxon_type)))
       dplyr::select(taxids, label, targetTaxonomy, targetRank) %>%
         write_csv(file_out(!!sprintf("output/%s_taxa_target.csv", taxon_type)))
       dplyr::select(taxids, label, Taxonomy, rank) %>%
@@ -1987,11 +1987,11 @@ options(clustermq.scheduler = "multicore")
 cache_dir <- ".drake"
 cache <- drake_cache(cache_dir)
 dconfig <- drake_config(plan2, cache = cache)
-# vis_drake_graph(dconfig, targets_only = TRUE,
-#                 group = "step",
[…]

### 2020-10-25 writing/transect_paper.Rmd
adjust page formatting for thesis

@@
     keep_md: yes
-fontsize: 12pt
-geometry: ["left=2.5cm", "right=2.5cm", "top=3cm", "bottom=3cm"]
+fontsize: 10pt
 graphics: yes
@@
 
-\singlespacing
+<!-- \singlespacing -->
 
-\printbibliography
+<!-- \printbibliography -->
 
-\doublespacing
+<!-- \doublespacing -->
 
@@
 
-\processdelayedfloats
+<!-- \processdelayedfloats -->

### 2020-10-26 writing/transect_paper.Rmd
separate ASV and OTU read abundance into different figures

@@
 
-Read counts for shared ASVs and OTUs were highly correlated between strategies, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")` (Figure\ \@ref(fig:read-compare)).
+Read counts for shared ASVs were highly correlated between strategies, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")` (Figure\ \@ref(fig:read-compare-asv)).
 Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
@@
 `r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs.
-PacBio, Illumina, and Ion Torrent short amplicon reads, respectively; Figure\ \@ref(fig:read-compare)).
+PacBio, Illumina, and Ion Torrent short amplicon reads, respectively; Figure\ \@ref(fig:read-compare-otu)).
 
@@
 OTU read counts were strongly correlated between technologies (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon == y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
-and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`, Figure \@ref(fig:read-compare)b),
+and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`, Figure \@ref(fig:read-compare-otu)),
 which we interpret as supporting the validity of abundance-based beta diversity measures in metabarcoding [but see @castano2020 and references therein].
 
-DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of OTUs recovered from the same library by different technologies and the read counts for each OTU (Figure\ \@ref(fig:read-compare)).
+DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of OTUs recovered from the same library by different technologies and the read counts for each OTU (Figure\ \@ref(fig:read-compare-otu)).
 ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures\ \@ref(fig:venn) and \@ref(fig:alpha-compare)).

### 2020-10-26 writing/transect_supplement.Rmd
separate ASV and OTU read abundance into different figures

@@
 
-(ref:read-compare-s) Comparison between read numbers for different sequencing strategies
+(ref:read-compare-asv) Comparison between ASV read abundances for different sequencing strategies
 
@@
 
-```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8, cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
-ggpubr::ggarrange(
+```{r read-compare-asv, fig.cap="(ref:read-compare-asv). ASVs (ref:read-compare)", fig.scap="(ref:read-compare-asv)", cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
   read_comparison(
@@
     type = "ASV"
-  ),
+  )
+```
+
+(ref:read-compare-otu) Comparison between OTU read abundances for different sequencing strategies
+
+```{r read-compare-otu, fig.cap="(ref:read-compare-otu). OTUs (ref:read-compare)", fig.scap="(ref:read-compare-otu)", cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
   read_comparison(
@@
     type = "OTU"
-  ),
-  ncol = 1,
-  labels = "auto",
-  label.x = 0.9
   )

### 2020-11-08 writing/transect_supplement.Rmd
fix some figures in supplement

@@
 
-(ref:read-compare) by ASV (**a**) and 97% OTU (**b**).
-ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
+(ref:read-compare) which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
 Dashed line represents a constant ratio of read numbers.
@@
 ```{r read-compare-asv, fig.cap="(ref:read-compare-asv). ASVs (ref:read-compare)", fig.scap="(ref:read-compare-asv)", cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
+grid::grid.newpage()
+grid::grid.draw(
   read_comparison(
@@
     type = "ASV"
+  )
 )
@@
 ```{r read-compare-otu, fig.cap="(ref:read-compare-otu). OTUs (ref:read-compare)", fig.scap="(ref:read-compare-otu)", cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
+grid::grid.newpage()
+grid::grid.draw(
   read_comparison(
@@
     type = "OTU"
+  )
 )

### 2020-11-19 writing/transect_supplement.Rmd
format supplement for thesis

@@
     suppress-bibliography: false
-    fontsize: 12pt
-    geometry: left=2.5cm, right=2.5cm, top=3cm, bottom=3cm
+    fontsize: 10pt
 ---
@@
     ),
-    escape = FALSE
+    escape = FALSE,
+    table.envir = "sidewaystable"
     ) %>%
@@
 readd(venn_ASV) %>%
-venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
+venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)", table.envir = "sidewaystable")
 ```
@@
 readd(venn_OTU) %>%
-venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
+venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)", table.envir = "sidewaystable")
 ```
@@
   arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
-  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
+  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"), table.envir = "sidewaystable")
 ```

### 2020-11-19 scripts/functions_output.R

@@ -187,7 +187,7 @@ reverselog_trans <- function(base = exp(1)) {
 
 
 # Present the results of venndata() as a table
-venn_table <- function(venndata, var, caption, caption.short) {
+venn_table <- function(venndata, var, caption, caption.short, ...) {
   var <- enquo(var)
 
   linesep <- venndata %>%
@@ -229,7 +229,8 @@ venn_table <- function(venndata, var, caption, caption.short) {
       linesep = linesep,
       align = ("rcrrcrrcrrcrr"),
       row.names = FALSE,
-      escape = FALSE
+      escape = FALSE,
+      ...
     ) %>%
     kableExtra::add_header_above(
       header = c(

### 2021-02-03 writing/transect_paper.Rmd
final changes before thesis submission

@@
   - "R.bib"
-biblio-style: apa
+biblio-style: authoryear-comp
+biblatexoptions: [giveninits=true, maxbibnames=999, maxcitenames=2, mincitenames=1, uniquename=false, uniqueprimaryauthor=false, uniquelist=false, date=year, url=false, isbn=false, eprint=false, doi=true]
 always_allow_html: true
@@
 knitr::opts_chunk$set(
-  dev = "cairo_pdf",
+  dev = "pdf",
+  dpi = 600,
   # dev.args = list(pointsize=32),
@@
 
+\renewcommand{\bibname}{References}
+\renewcommand{\bibfont}{\normalfont\footnotesize}
+\renewbibmacro*{volume+number+eid}{%
+  \printfield{volume}%
+%  \setunit*{\adddot}% DELETED
+  \setunit*{\addnbthinspace}% NEW (optional); there's also \addnbthinspace
+  \printfield{number}%
+  \setunit{\addcomma\space}%
+  \printfield{eid}}
+\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
 <!-- \processdelayedfloats -->

### 2021-02-03 writing/transect_supplement.Rmd
final changes before thesis submission

@@
   - Martin Ryberg^1^
-  - ^1^Program in Systematic Biology,
+date: |
+      | ^1^Program in Systematic Biology,
         Department of Organismal Biology,
         Uppsala University,
-  - Uppsala, Sweden
-  - ^2^Department of Ecology,
+        Uppsala, Sweden
+      | ^2^Department of Ecology,
         Swedish University of Agricultural Sciences,
         Uppsala, Sweden
-  - ^3^Institute of Ecology and Earth Sciences,
+      | ^3^Institute of Ecology and Earth Sciences,
         University of Tartu,
         Tartu, Estonia
-  - ^4^Program in Evolutionary Biology,
+      | ^4^Program in Evolutionary Biology,
         Department of Ecology and Genetics,
         Uppsala University,
-  - Uppsala, Sweden
-  - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
+        Uppsala, Sweden
+      | ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
         LEB,
-  - University of Parakou,
+        University of Parakou,
         Parakou, Benin
@@
 knitr::opts_chunk$set(
-  dev = "cairo_pdf",
-  # dev.args = list(pointsize=32),
+  dev = "pdf",
+  dpi = 600,
   echo = FALSE,
@@
 ```{r buffer-compare-long, fig.scap = "(ref:buffer-compare-long-s)", fig.cap = "(ref:buffer-compare-long-s).
-(ref:buffer-compare-long)", fig.width = 10, fig.height = 10, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_long"), dev = "pdf"}
+(ref:buffer-compare-long)", fig.width = 10, fig.height = 10, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_long")}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 ```{r buffer-compare-short, fig.scap="(ref:buffer-compare-short-s)", fig.cap = "(ref:buffer-compare-short-s).
-(ref:buffer-compare-short)", fig.width = 10, fig.height = 10, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_short"), dev = "pdf"}
+(ref:buffer-compare-short)", fig.width = 10, fig.height = 10, out.width = "100%", cache = TRUE, extra.cache = gethash("buffer_compare_short")}
 oldtheme <- theme_update(panel.border = element_blank())
@@
 
-```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)", fig.env = "sidewaysfigure", fig.width = 9, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("conc_plot", "reads_plot")}
+\captionsetup{width=0.9\textheight}
+
+```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)", fig.env = "sidewaysfigure", fig.width = 8, fig.height = 5, out.width = "100%", cache = TRUE, extra.cache = gethash("conc_plot", "reads_plot"), results='asis'}
 ggpubr::ggarrange(readd(conc_plot), readd(reads_plot), ncol = 2, common.legend = TRUE) +
@@
 ```
+
+\captionsetup{width=0.9\textwidth}
 
@@
 
-```{r alpha-compare-otu, fig.scap="(ref:alpha-compare-otu-s)", fig.cap="(ref:alpha-compare-otu-s) (ref:alpha-compare)"}
+```{r alpha-compare-otu, fig.scap="(ref:alpha-compare-otu-s)", fig.cap="(ref:alpha-compare-otu-s) (ref:alpha-compare-OTU)"}
 grid::grid.newpage()
@@
 
-```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", fig.width = 9, fig.height = 5, out.width="100%", fig.env = "sidewaysfigure", cache=TRUE, extra.cache=gethash("heattree")}
+\captionsetup{width=0.9\textheight}
+
+```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", fig.width = 8, fig.height = 5, out.width="100%", fig.env = "sidewaysfigure", cache=TRUE, extra.cache=gethash("heattree")}
 theme <- theme_update(panel.border = element_blank())
@@
 
+\captionsetup{width=0.9\textheight}
+
 ```{r bioinfo, cache=TRUE, extra.cache=gethash("bioinf_table", "bioinf_header")}
@@
 \FloatBarrier
+
+\renewcommand{\bibname}{References}
+\renewcommand{\bibfont}{\normalfont\footnotesize}
+\renewbibmacro*{volume+number+eid}{%
+  \printfield{volume}%
+%  \setunit*{\adddot}% DELETED
+  \setunit*{\addnbthinspace}% NEW (optional); there's also \addnbthinspace
+  \printfield{number}%
+  \setunit{\addcomma\space}%
+  \printfield{eid}}
+\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}

### 2021-02-03 writing/transect_paper.Rmd
Comments from second round reviewers.

--- 2020-10-25 writing/transect_paper.Rmd
@@
     keep_md: yes
-fontsize: 10pt
+fontsize: 12pt
+geometry: ["left=2.5cm", "right=2.5cm", "top=3cm", "bottom=3cm"]
 graphics: yes
@@
 In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice [@opik2010], and variable regions of the rDNA large subunit (LSU) have also been used for barcoding [@house2016;@tedersoo2015].
+<!-- R4: consider citing Kurtzman and Robnett, 1998 Antonie van Leeuwenhoek 73: 331 (and other
+papers) who used the LSU region to ‘barcode’ yeasts -->
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
@@
 However, this approach comes with some potential weaknesses, discussed below.
+<!--R5: Awk -->
 
@@
 This brings into question the comparability of studies conducted only a few years apart.
+<!-- R5: Seems like this section below could use some references to studies that have tested communities with multiple methods.
+There are a bunch of citations like this but none featured here.
+For example:
+Smith, Dylan P., and Kabir G. Peay. "Sequence depth, not PCR replication, improves ecological inference from next generation DNA sequencing." PloS one 9.2 (2014): e90234. -->
 
@@
 but this variablity means that they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
+<!-- R5: citation formatting-->
 Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa [@ihrmark2012;@tedersoo2015;@palmer2018].
@@
 Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
+<!--R4: very true, great point. In fact, within-species ITS variation was also nicely discussed in
+Lindner and Banik, 2011 Mycologia 103: 731 in case you want another citation-->
 In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while preserving intra-species variation.
@@
 Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated [@glassman2018;@botnen2018].
-ASVs have been suggested to be less dataset specific that cluster-based OTUs [@callahan2017].
+ASVs have been suggested to be less dataset specific that cluster-based OTUs [@callahan2017]. <!-- R5: cross out that -->
 Support for PacBio has recently been added to DADA2 [@callahan2019], but its application requires greater sequencing depth for longer reads, especially in high diversity samples.
@@
 Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018], or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases [@fouquier2016], but these approaches are only applicable to sequences of known taxonomic affiliation.
+<!--R4: other literature you may want to consider citing: another phylogeny-based taxonomic
+assignment method is SAP (statistical assignment package, Munch et al., 2008a Syst Biol 57: 750
+and Munch et al., 2008b Phil Trans Roy Soc B 363: 3997), another lowest common ancestor
+approach MEGAN (metagenomic analysis) Huson et al., 2016 PLoS Comp Biol 12: e1004957 that
+can parse BLAST reports or RDP Classifier reports.Near L114 - More refs to consider: See Ashelford et al., 2005 Appl Env Microbiol 71: 7724,
+Bidartondo et al., 2008 Science 319:1616, but also LeRay et al., 2019 PNAS 116: 22651 for a
+more recent take-->
 
@@
 Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
-Consequently taxonomic assignments based on these databases may be incorrect or inconsistent.
+Consequently <!--R5: add comma--> taxonomic assignments based on these databases may be incorrect or inconsistent.
 Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
-The Unite database is a more curated attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
+<!--R4: see Nilsson et al., 2005 BMC Bioinformtics 6: 178-->
+The Unite database is a more curated <!--More curated?-->attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
 The Ribosomal Data Project [RDP, @cole2014] hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
@@
 However, all sequence-similarity based approaches are dependent on high taxonomic coverage in the reference database, making the placement of novel or undersampled groups problematic [@nilsson2016top;@Tedersoo2018].
+<!--R4: consider mentioning Nilsson et al., 2019 Nuc Acids Res 47: D259 on the naming of dark
+taxa in UNITE see: taxon hypotheses (THs)-->
 
@@
 Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
+<!--R5: Add citation for “turnover scale”-->
 Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
@@
 with three different reference databases (Unite, Warcup, RDP),
+<!--R4: RDP reference sets include UNITE ITS and Warcup ITS as well as an LSU reference set
+(Liu et al. 2012). I think it would be more clear to replace ...(Unite, Warcup, RDP) to (UNITE ITS,
+Warcup ITS, Liu LSU)? As ‘RDP’ generally refers to the better-known prokaryote SSU rRNA gene
+reference set.-->
 and both non-phylogenetic and phylogenetic community distance measures.
 We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
+
+<!--R5: I realize that this paper is strongly methods-based and is focused on figuring out which approaches might work best for understanding the fungi.
+However, I think there are likely some underlying questions and hypotheses.
+This is the place where I would spend a few sentences exploring the hypotheses.
+For example, what do you you expect to find with the new tree-based and distance-based approaches?
+Do you expect to find more diversity in the long read technology because the primers are more general, etc.?-->
 
@@
 Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
+<!-- R5: This does not match the taxonomic info in the intro-->
 At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
@@
 For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with a sterilized knife blade and homogenized.
+<!--R4: how was the knife sterilized, how were the samples homogenized?
+Preserved at what
+temperature? Please address in main text.-->
 A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary methods).
@@
 ## DNA amplification and sequencing
-
+<!--R5: -info on extraction techniques is missing here-->
 DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
@@
 The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
+<!--R4: “a mix of ITS4 and ITS4a” were equal mounts mixed? Please specify in text.-->
 The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:regions)a), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
-Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
+Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.<!--R5: In the future I strongly recommend using a positive control community of known taxa.
+this really helps to parameterize the bioinformatics to make sure you are capturing the approximate diversity that is expected based on this more complex control (rather than using one species, which is less helpful) -->
 See Supplementary methods for PCR and library preparation protocols.
+<!--I always recommend that full methods be in the main text, not in the supplement, it only
+takes a few sentences to describe PCR protocols at least.-->
 
@@
 We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
+<!--R4: I suggest that you refer to a specific supplementary figure or table, instead of the whole
+supplementary methods document here and elsewhere.-->
 We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
 We used this method, implemented in the new R packages `LSUx` (https://github.com/brendanf/LSUx) and `tzara` (https://github.com/brendanf/tzara) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.
+<!--R4: I suggest that the authors describe what the LSUx and tzara packages are used for if
+these are new and being introducing for the first time here.-->
 
@@
 At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs<!-- This sentence is how to interpret the figure -->.
+<!--R4: It would be nice if the colors used in part b corresponded with the same colors in part
+a. What does 32S refer to here? Is this data based on rarefied, normalized data?-->
+<!--R5: Not clear what 32S is here?-->
 
@@
 
+<!--R4: this looks like maybe a new paragraph? It should be connected to last paragraph.
+Regular indentations at the start of each paragraph would make this more clear.-->
 Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
@@
 
+<!--R4: this looks like maybe a new paragraph? It should be connected to last paragraph.-->
 To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
@@
 ### Phylogenetic inference and taxonomy assignment
+<!--R4: ‘Phylogenetic inference and TAXONOMIC assignment’-->
 
@@
 The tree was rooted outside kingdom Fungi and non-Fungi were removed based on taxonomic assignments (see below and Supplementary methods).
+<!--R5: I realize that is covered in the Supp. methods in more detail but I would at least give one sentence about how you chose some reference sequences to parameterize the tree here-->
 
@@
 Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
+<!--R4: ...LSU region using RDP. I suggest you refer to the RDP LSU reference set (Liu et al., 2012)
+instead of just to RDP which is better known for its prokaryote SSU rRNA gene classifier.-->
 For each region/reference combination, taxonomy was assigned using three popular algorithms:
@@
 A relatively lax confidence threshhold of 50% was used for all three algorithms.
+<!--R4: a relatively lax bootstrap support cutoff of 50% would indicate classification accuracy of
+what? 75% accuracy, more, less, does it vary across ITS and LSU? I suggest this be addressed in
+the main text, in https://rdp.cme.msu.edu/download/posters/fungalITSreport_062014.pdf see
+Fig 3. -->
+<!--R5: This seems very low.
+I think that some explanation for such a low value is needed to give a rationale.
+This is basically a coin toss so that is not exactly giving great confidence in the results-->
 Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
@@
 For full-length long-amplicon ASVs, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax) and Supplementary Methods).
+<!--R4: “refined using a phylogenetic tree” instead of ‘the’L275 - suggest you spend a few lines to describe what exactly what the phylotax package does if
+it’s new and you’re introducing it here for the first time-->
 This refinement algorithm is referred to as "PHYLOTAX", available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
@@
 ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the preliminary assignments.
+<!--R4: a good citation is for the LCA algorithm is MEGAN Huson et al. 2016, see comment above-->
 This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [LCAClassifier, @lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
@@
 Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
+<!--R5: So make a statement about how many samples are left in the analyses after these were excluded? Seems like it may be a very low number of samples here-->
 
@@
 Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
+<!--R5: But how does that impact quality? Seems like this would reduce quality significantly -->
 Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of denoised ITS2 reads from the long-amplicon PacBio dataset.
@@
 
-Among the different regions extracted from the long amplicon (Figure\ \@ref(fig:region-lengths),
+Among the different regions extracted from the long amplicon (Figure\ \@ref(fig:region-lengths)),
 ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
@@
 ## Reproducibility of sequence capture using different technologies
+<!--R4: ‘capture’ usually refers to the enrichment of metagenomic DNA fragments using bead
+based selection for example. Here, consider something like ‘Reproducibility of sequence
+detection using different sequencing technologies’-->
+<!--R5: This has a specific meaning for many people that is about sequences obtained via “capture” using hybridization- i would use a different heading here to make it more obvious what you are referring to-->
 
@@
 For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
+<!--R4: it would be helpful if you drew a line down the middle of this figure then labelled the
+left side ASVs and the right side OTUs. Instead of ‘abundance’ suggest you change to ‘read
+number or read abundance’. Is this figure based on rarefied, normalized data? Indicate in the
+Text.-->
+<!--R5: again, were these analyses conducted on the entire (i.e. fungal and non-fungal) dataset, or just the fungal community?-->
 
@@
 The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
-The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).
+The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).<!--R5: extra %-->
 When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% of reads, respectively (Figure\ \@ref(fig:venn)b).
-In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
+In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.<!--R5: extra %-->
 ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` belonged to a unique OTU after clustering.
 In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).
+<!--R5: Not clear what that means?-->
 
@@
 ASV richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(readd(demingfits_ASV), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
+<!--R4: here and onwards (L409, L418), I see that you have plotted rarefaction curves, but it’s
+not clear to me whether you compared samples at equal sampling depth, ex. smallest library
+size. If you did this, please add a line to say so in the main text. If you didn’t do this, how do
+you know that your community comparisons aren’t simply the effect of library size? Please
+address in main text.-->
 The slope of the relationship between PacBio and Illumina richness estimates was only slightly different from 1, indicating that these two technologies give highly comparable rarefied richness estimates, despite the approximately 200× difference in original sequencing depth.
@@
 Total least squares regression indicated that the long amplicon dataset resulted in richness estimates which were intermediate between the short amplicon results from Ion Torrent and the other two technologies.
-Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure\ \@ref(fig:venn)), sample-wise OTU accumulation curves (Figure\ \@ref(fig:accum-OTU)) and rarefied OTU richness relationships between sequencing strategies (Figure\ \@ref(fig:alpha-compare-otu)) were highly similar to those for ASVs.
+Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure\ \@ref(fig:venn)), sample-wise OTU accumulation curves
+<!--R5: Not clear what that means?-->
+(Figure\ \@ref(fig:accum-OTU)) and rarefied OTU richness relationships between sequencing strategies (Figure\ \@ref(fig:alpha-compare-otu)) were highly similar to those for ASVs.
 
@@
 Note logarithmic axes.
+<!--R4: If you normalized your dataset by rarefying your samples to the lowest library size, it
+would be helpful if you added a vertical line to the plot indicating ASV richness after
+Normalization.-->
+<!--R5: again, were these analyses conducted on the entire (i.e. fungal and non-fungal) dataset, or just the fungal community?-->
 
@@
 Dashed diagonal line indicates ideal slope of 1.
+<!--R4: Add a period after the figure title. change ‘ideal’ slope to ‘1:1 line’ or something
+similar. Is this figure based on rarefied, normalized data? Indicate in the text.-->
 
@@
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
+<!--R4: This figure needs a ‘short’ title. Suggest you change RDP to RDP LSU training set. I’d
+be interested in seeing a comparison of run times for each taxonomic assignment or a sentence
+added to the main text comparing run times. In the discussion you may want to indicate that
+phylogeny-based methods tend to be more computationally intensive compared to BLAST or
+RDP ex. shown for fungal ITS (Porter and Golding, 2011 New Phyt 192: 775 ) and fungal LSU
+(Porter and Golding, 2012 PLoS ONE 7: e35749).-->
 
@@
 Classes which represented less than 2% of reads and ASVs in all datasets are grouped together as "other".
+<!--R4: Is this comparison based on rarefied, normalized data? How do you know these
+patterns don’t reflect different library sizes? Please address in the main text. Change ‘which’ to
+‘That’.-->
 
@@
 Many of the ASVs which were unique to the long-amplicon library thus fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)).
+<!--R5: This needs to be mentioned above.
+This is the first mention of non-fungal reads.
+Need to address this in the Methods and also above near note on the ASVs in the long read vs.
+short read datasets-->
 In particular a large fraction of ITS2 sequences with length less than 140 (Figure\ \@ref(fig:ITS2-length)) were identified as Alveolates (Figure\ \@ref(fig:short-its)).
@@
 Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
+<!--R5: Here is not clear if you screened out the non-fungi from the various datasets before you created this part.
+I think you need to specifically address this in methods so the reader knows whether you are making comparisons between the full community or with the fungal communithy-->
 The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm), \@ref(fig:heattree-length-compare), and \@ref(fig:ppcoa)).
@@
 In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
+<!--R5: But if non-fungi were included then this may drive some of the pattern detection so needs to be clearer about whether non fungi are discarded for this or not?-->
 The least spatial structure was detected with PacBio short amplicon sequencing.
@@
 Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
+<!--R4: confidence intervals are not shown? If you’re doing to discuss CI then consider adding
+them to your figure.-->
 
@@
 which we interpret as supporting the validity of abundance-based beta diversity measures in metabarcoding [but see @castano2020 and references therein].
+<!--R4: ‘validity of abundance-based beta diversity measures’. Anyone can choose to treat their
+data as if read proportion/relative proportion is informative, but for fungi with different thallus
+sizes (yeast-like, filamentous), different nuclear states (haploid, diploid/dikaryons), abundant
+sporulators, etc. what does this really mean? This is a very contentious statement and I’m not
+sure that the data you have presented makes this case in any way. See Amend et al. 2010
+Molec Ecol 19: 5555, Amend et al. 2010 PNAS 107: 13748 supp mat, Elbrecht and Leese, 2015
+PLoS ONE 10: e0130324 for a non-fungal example, but interesting because like fungi,
+arthropods come in ranges of body sizes that differ by orders of magnitude and they are multi-
+cellular (unlike prokaryotes where relative abundance is often used, despite the fact that the
+rRNA genes exist in differently sized tandem arrays that differ across lineages). Some studies
+have even proposed ‘correction factors’ to make read counts more meaningful when assessing
+species abundance. I think this statement needs to be re-written to be more conservative, as
+your methods do not directly test whether read abundance reflects real species abundance.
+You may even want to discuss and/or cite some of the literature that has directly tackled this
+Issue.-->
+<!--R5: These are all short amplicons, and previous studies have indicated strong bias in short amplicon HTS, across different sequencing technologies.
+The correlation alone cannot be used as evidence for the validity of abundance-based diversity measures.
+Especially as these R2 values are on the borderline of what is interpreted as "strong".-->
 
@@
 ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures\ \@ref(fig:venn) and \@ref(fig:alpha-compare)).
+<!--R4: I’m uncertain how to interpret your rarefaction curves since both axes are on a log scale.
+Does adequate read depth, saturated sampling, plateau in the same way that you would expect
+curves to on regular axes as this is normally plotted? Please clarify in the main text.-->
 This may be a result of the lower fraction of very high quality reads in the Ion Torrent dataset (Figure \@ref(fig:qual-check)).
@@
 The long amplicon dataset included `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` unique taxa, even after clustering at 97% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
+<!--R5: If these 20% are mostly non-fungal reads then this needs to be a bigger part of the discussion!
+Furthermore, some of these primers have known biases against many non-Dikarya fungi.
+In which case, "total fungal community" in any resulting analyses has the very large caveat "biased towards Dikarya" attached to it.
+-->
 The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae [@tedersoo2015].
@@
 800\ 000 for Unite), it is likely that many ITS sequences, especially from uncultured tropical soil fungi, have no close match in the Warcup database, and so cannot be placed.
+<!-- R4: replace ‘RDP fungal training set’ with RDP LSU training set (Liu et al., 2012) as the RDP
+has 3 fungal training sets (UNITE ITS, Warcup ITS, LSU) in addition to the better-known
+prokaryote training set SSU rRNA.-->
 The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
@@
 Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, as expected given its more well-calibrated and conservative confidence scores [@murali2018a], but this was particularly dramatic when paired with the Warcup database, where IDTAXA placed <25% of ASVs even to phylum.
+<!--R5: This discussion of placement is good.
+However, it suffers from one serious weakness.
+It seems to assume that when things are placed that they are placed CORRECTLY.
+In my experience some things are obviously incorrectly placed.
+You should explicitly say whether you actually dealt with his problem or not.
+If not, at least give some caveats here!-->
 
@@
 For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses phylogenetic relationships to resolve these disagreements in a principled manner.
+<!--R4: in a ‘principled’ manner. Consider rephrasing to ‘uses phylogenetic relationships to
+resolve disagreements.’ Do you need to set a bootstrap support cutoff to get a good phylotax
+assignment? What cutoff did you use here? Does phylotax use neighbor joining or ML or
+Bayesian tree construction? These details need to be added to your method section in the main
+text. How does phylotax handle ambiguously aligned regions of ITS or LSU divergent domains?
+Are they removed, recoded, ignored?-->
+<!--R5: How were the results from this algorithm validated? On which taxonomic groups of fungi and/or non-fungi was it tested? How were the results of the other taxonomy methods checked for errors?-->
 The effect was most pronounced for the PacBio long amplicon dataset,
@@
 This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure\ \@ref(fig:ecm)).
+<!--R5: By how much?
+What proportion?
+This is the really important thing here because readers want to know how much extra information they would get for all this work and money to get the PacBio long read data!-->
 For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs (last row of Table\ \@ref(tab:bioinfo)).
@@
 In a recent study, @kivlin2020 reported a space-for-time substitution rate of `r format(1.5/6.8 * 365, digits = 2)`\ m/year (reported as 6.8\ d / 1.5\ m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
+<!--R4: I am not familiar with the term ‘space-for-time substitution rate’. Please explain this
+concept in the main text.-->
+
 However, comparison is obscured by different spatial and temporal sampling scales between the two studies.
 Year-to-year variation in ECM fungal communities, which we sampled, has been shown to be less than intra-annual variation [@bahram2015], as sampled by @kivlin2020.
-Neither dataset from the short amplicon library showed significant temporal autocorrelatio.
+Neither dataset from the short amplicon library showed significant temporal autocorrelatio.<!--R4: autocorrelation-->
 
@@
 
-<!-- \singlespacing -->
+\singlespacing
 
-<!-- \printbibliography -->
+\printbibliography
 
-<!-- \doublespacing -->
+\doublespacing
 
@@
 
-<!-- \processdelayedfloats -->
+\processdelayedfloats

### 2021-02-04 writing/transect_supplement.Rmd
side-by-side zoomed scales for accumulation curves instead of log-log plots

--- 2020-11-08 writing/transect_supplement.Rmd
@@
 
-(ref:read-compare-asv) Comparison between ASV read abundances for different sequencing strategies
+(ref:read-compare-s) Comparison between read numbers for different sequencing strategies
 
-(ref:read-compare) which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
+(ref:read-compare) by ASV (**a**) and 97% OTU (**b**).
+ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes.
 Dashed line represents a constant ratio of read numbers.
@@
 
-```{r read-compare-asv, fig.cap="(ref:read-compare-asv). ASVs (ref:read-compare)", fig.scap="(ref:read-compare-asv)", cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
-grid::grid.newpage()
-grid::grid.draw(
+```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8, cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
+ggpubr::ggarrange(
   read_comparison(
@@
     type = "ASV"
-  )
-)
-```
-
-(ref:read-compare-otu) Comparison between OTU read abundances for different sequencing strategies
-
-```{r read-compare-otu, fig.cap="(ref:read-compare-otu). OTUs (ref:read-compare)", fig.scap="(ref:read-compare-otu)", cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
-grid::grid.newpage()
-grid::grid.draw(
+  ),
   read_comparison(
@@
     type = "OTU"
-  )
+  ),
+  ncol = 1,
+  labels = "auto",
+  label.x = 0.9
 )
@@
 
-(ref:accum-OTU-s) OTU accumulation curves
+(ref:accum-OTU-s) 97% OTU accumulation curves
 
@@
 Points at the end of each curve represent the actual read depth and observed OTU richness.
-Note logarithmic axes.
+Panel at left has enlarged scale to show PacBio more clearly.
 

### 2021-02-04 scripts/functions_richness.R

@@ -156,14 +156,17 @@ make_accum_plot <- function(sample_iNEXT, cluster_type) {
     dplyr::filter(method != "extrapolated") %>%
     ggplot(aes(x = m, y = qD, color = strategy)) +
     # lines for each sample
-    geom_line(aes(group = sample), alpha = 0.08) +
+    geom_line(aes(group = sample), alpha = 0.15) +
     # small points to represent the actual observed read depth and richness
     geom_point(data = ~dplyr::filter(., method == "observed"),
                alpha = 0.8, size = 1, shape = 1) +
     xlab("Number of rarefied reads") +
     ylab(paste(cluster_type, "richness")) +
     scale_color_strategy(name = NULL) +
-    scale_x_log10() +
-    scale_y_log10() +
-    theme_bw()
+    ggforce::facet_zoom(
+      xy = startsWith(strategy, "PacBio"),
+      zoom.size = 1
+    ) +
+    theme_bw() +
+    theme(legend.position = "top")
 }

### 2021-02-04 writing/transect_paper.Rmd
side-by-side zoomed scales for accumulation curves instead of log-log plots
add line for rarefaction in accumulation plots

@@
 Points at the end of each curve represent the actual read depth and observed ASV richness.
-Note logarithmic axes.
-<!--R4: If you normalized your dataset by rarefying your samples to the lowest library size, it
-would be helpful if you added a vertical line to the plot indicating ASV richness after
-Normalization.-->
+Panel at left has enlarged scale to show PacBio more clearly.
+Vertical dashed line at 100 reads indicates rarefaction level for ASV richness comparisons.
 <!--R5: again, were these analyses conducted on the entire (i.e. fungal and non-fungal) dataset, or just the fungal community?-->
@@
 ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures\ \@ref(fig:venn) and \@ref(fig:alpha-compare)).
-<!--R4: I’m uncertain how to interpret your rarefaction curves since both axes are on a log scale.
-Does adequate read depth, saturated sampling, plateau in the same way that you would expect
-curves to on regular axes as this is normally plotted? Please clarify in the main text.-->
 This may be a result of the lower fraction of very high quality reads in the Ion Torrent dataset (Figure \@ref(fig:qual-check)).

### 2021-02-04 writing/transect_paper.Rmd
use MER output widths for figures in main text

@@
 
+onecolwidth = "80mm"
+twothirdwidth = "112mm"
+twocolwidth = "169mm"
+
 knitr::opts_chunk$set(
@@
   external = TRUE,
-  out.width = "70%",
+  out.width = twothirdwidth,
   fig.align = "center"
@@
 
-```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", out.width = "70%", fig.height = 4, fig.align="center", results="asis"}
+```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", fig.height = 4, fig.align="center", results="asis"}
 ggpubr::ggarrange(readd(vennplot_tech_ASV), readd(vennplot_tech_OTU), ncol = 2)
@@
 
-```{r taxon-chart, fig.cap = "(ref:taxon-chart)", fig.scap = "(ref:taxon-chart-s)", out.width = "100%"}
+```{r taxon-chart, fig.cap = "(ref:taxon-chart)", fig.scap = "(ref:taxon-chart-s)", out.width = twocolwidth}
 readd(taxon_chart_plot)

### 2021-02-04 writing/transect_paper.Rmd
Split combined comment

@@
 approach MEGAN (metagenomic analysis) Huson et al., 2016 PLoS Comp Biol 12: e1004957 that
-can parse BLAST reports or RDP Classifier reports.Near L114 - More refs to consider: See Ashelford et al., 2005 Appl Env Microbiol 71: 7724,
-Bidartondo et al., 2008 Science 319:1616, but also LeRay et al., 2019 PNAS 116: 22651 for a
-more recent take-->
+can parse BLAST reports or RDP Classifier reports.-->
 
@@
 Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
+<!--R4: Near L114 - More refs to consider: See Ashelford et al., 2005 Appl Env Microbiol 71: 7724,
+Bidartondo et al., 2008 Science 319:1616, but also LeRay et al., 2019 PNAS 116: 22651 for a
+more recent take-->
 Consequently <!--R5: add comma--> taxonomic assignments based on these databases may be incorrect or inconsistent.

### 2021-02-04 writing/transect_paper.Rmd
RDP -> RDP-LSU

@@
 The Ribosomal Data Project [RDP, @cole2014] hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
-the Warcup ITS database, containing 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [@liu2012].
+the Warcup ITS training set, containing 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [RDP-LSU, @liu2012].
 Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.
@@
 
-We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
-with three different reference databases (Unite, Warcup, RDP),
-<!--R4: RDP reference sets include UNITE ITS and Warcup ITS as well as an LSU reference set
-(Liu et al. 2012). I think it would be more clear to replace ...(Unite, Warcup, RDP) to (UNITE ITS,
-Warcup ITS, Liu LSU)? As ‘RDP’ generally refers to the better-known prokaryote SSU rRNA gene
-reference set.-->
+We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDPC, SINTAX, IDTAXA)
+with three different reference databases (Unite, Warcup, RDP-LSU),
 and both non-phylogenetic and phylogenetic community distance measures.
@@
 
-Taxonomic annotations of the RDP LSU fungal training set version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods).
-Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
-<!--R4: ...LSU region using RDP. I suggest you refer to the RDP LSU reference set (Liu et al., 2012)
-instead of just to RDP which is better known for its prokaryote SSU rRNA gene classifier.-->
+Taxonomic annotations of the RDP-LSU training set version 11.5 [@liu2012;@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods).
+Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
 For each region/reference combination, taxonomy was assigned using three popular algorithms:
-the RDP Naïve Bayesian Classifier [RDPC, @wang2007] as implemented in DADA2;
+the RDPC [@wang2007] as implemented in DADA2;
 SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
@@
 
-(ref:taxon-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
+(ref:taxon-chart) Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP-LSU), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
-<!--R4: This figure needs a ‘short’ title. Suggest you change RDP to RDP LSU training set. I’d
+<!--R4: This figure needs a ‘short’ title. I’d
 be interested in seeing a comparison of run times for each taxonomic assignment or a sentence
@@
 However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
-Although the RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \@ref(fig:taxon-chart)).
+Although RDP-LSU and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \@ref(fig:taxon-chart)).
 This is probably due to two factors.
@@
 800\ 000 for Unite), it is likely that many ITS sequences, especially from uncultured tropical soil fungi, have no close match in the Warcup database, and so cannot be placed.
-<!-- R4: replace ‘RDP fungal training set’ with RDP LSU training set (Liu et al., 2012) as the RDP
-has 3 fungal training sets (UNITE ITS, Warcup ITS, LSU) in addition to the better-known
-prokaryote training set SSU rRNA.-->
-The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
+RDP-LSU, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
 @heeger2019 also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
@@
 If they are not accepted prior to final publication, snapshots will be archived at Dryad.
- - FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
+ - FASTA-format files for the RDP fungal LSU training set, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
 The versions used in this paper will be archived at Dryad prior to publication.

### 2021-02-04 writing/transect_supplement.Rmd
add line for rarefaction in accumulation plots
RDP -> RDP-LSU

@@
 Thick branches represent rapid bootstrap values >90%.
-Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, with Unite, Warcup, and RDP reference databases. 
+Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, with Unite, Warcup, and RDP-LSU reference databases. 
 Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
@@
 
-### Uniform taxonomic classification for Unite, Warcup, and RDP
+### Uniform taxonomic classification for Unite, Warcup, and RDP-LSU
 
@@
 Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
-FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
+FASTA-format files for the RDP-LSU, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
 
@@
 Panel at left has enlarged scale to show PacBio more clearly.
+Vertical dashed line at 100 reads indicates rarefaction level for OTU richness comparisons.
 
@@
 Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
-The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
+The RDP-LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
 
@@
 Height of each bar represents the fraction of reads represented by that bar.
-Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; "All" denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.
+Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP-LSU; "All" denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.
 

### 2021-02-04 scripts/functions_taxonomy.R

@@ -208,7 +208,7 @@ format_taxon_reads <- function(taxon_reads, funguild_db) {
       "reference",
       factor,
       levels = c("All", "unite", "warcup", "rdp_train"),
-      labels = c("All", "Unite", "Warcup", "RDP")
+      labels = c("All", "Unite", "Warcup", "RDP-LSU")
     ) %>%
     mutate_at("reference", replace_na, "All") %>%
     mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
@@ -270,7 +270,7 @@ compile_taxon_chart <- function(taxon_reads) {
     ) %>%
     rename(Reads = reads) %>%
     mutate_at("reference", as.character) %>%
-    mutate_at("reference", factor, levels = c("All", "Unite", "Warcup", "RDP")) %>%
+    mutate_at("reference", factor, levels = c("All", "Unite", "Warcup", "RDP-LSU")) %>%
     pivot_longer(cols = c("Reads", "ASVs"), names_to = "type", values_to = "frac")
 }
 

### 2021-02-05 writing/transect_paper.Rmd
address more comments...

@@
 which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
-In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice [@opik2010], and variable regions of the rDNA large subunit (LSU) have also been used for barcoding [@house2016;@tedersoo2015].
-<!-- R4: consider citing Kurtzman and Robnett, 1998 Antonie van Leeuwenhoek 73: 331 (and other
-papers) who used the LSU region to ‘barcode’ yeasts -->
+In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice [@opik2010], and variable regions of the rDNA large subunit (LSU) have also been used for barcoding [e.g., @kurtzman1998;@tedersoo2015;@house2016].
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
@@
 
-Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
-<!--R4: very true, great point. In fact, within-species ITS variation was also nicely discussed in
-Lindner and Banik, 2011 Mycologia 103: 731 in case you want another citation-->
+Distance-based clustering conflates intra-species variation and sequencing error [@nilsson2008;@lindner2011], and results are dataset-specific.
 In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while preserving intra-species variation.
@@
 Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated [@glassman2018;@botnen2018].
-ASVs have been suggested to be less dataset specific that cluster-based OTUs [@callahan2017]. <!-- R5: cross out that -->
+ASVs have been suggested to be less dataset specific than cluster-based OTUs [@callahan2017].
 Support for PacBio has recently been added to DADA2 [@callahan2019], but its application requires greater sequencing depth for longer reads, especially in high diversity samples.
@@
 In contrast, phylogenetic community distance measures [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
-Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
+Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@munch2008a;@munch2008;@matsen2010;@berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
 Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018], or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases [@fouquier2016], but these approaches are only applicable to sequences of known taxonomic affiliation.
-<!--R4: other literature you may want to consider citing: another phylogeny-based taxonomic
-assignment method is SAP (statistical assignment package, Munch et al., 2008a Syst Biol 57: 750
-and Munch et al., 2008b Phil Trans Roy Soc B 363: 3997), another lowest common ancestor
-approach MEGAN (metagenomic analysis) Huson et al., 2016 PLoS Comp Biol 12: e1004957 that
-can parse BLAST reports or RDP Classifier reports.-->
 
@@
 more recent take-->
-Consequently <!--R5: add comma--> taxonomic assignments based on these databases may be incorrect or inconsistent.
+Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent.
 Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
 <!--R4: see Nilsson et al., 2005 BMC Bioinformtics 6: 178-->
-The Unite database is a more curated <!--More curated?-->attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
-The Ribosomal Data Project [RDP, @cole2014] hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
+The Unite database is a more curated <!--R5: More curated?-->attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
+The Ribosomal Data Project [RDP, @cole2014] hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
 the Warcup ITS training set, containing 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [RDP-LSU, @liu2012].
@@
 Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-<!-- R5: This does not match the taxonomic info in the intro-->
+Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Fabaceae tr. Amherstieae).
 At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with a sterilized knife blade and homogenized.
+For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade and homogenized by manually breaking up clumps and mixing inside a plastic zipper bag.
 <!--R4: how was the knife sterilized, how were the samples homogenized?
@@
 temperature? Please address in main text.-->
-A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary methods).
+A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).
 
 ## DNA amplification and sequencing
-<!--R5: -info on extraction techniques is missing here-->
 DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
 Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
-The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
-<!--R4: “a mix of ITS4 and ITS4a” were equal mounts mixed? Please specify in text.-->
+The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and an equimolar mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
 The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:regions)a), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
-Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.<!--R5: In the future I strongly recommend using a positive control community of known taxa.
-this really helps to parameterize the bioinformatics to make sure you are capturing the approximate diversity that is expected based on this more complex control (rather than using one species, which is less helpful) -->
-See Supplementary methods for PCR and library preparation protocols.
-<!--I always recommend that full methods be in the main text, not in the supplement, it only
+Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
+See Supplementary Methods \@ref(DNA-supplement) for PCR and library preparation protocols.
+<!--R4: I always recommend that full methods be in the main text, not in the supplement, it only
 takes a few sentences to describe PCR protocols at least.-->
@@
 
-We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
-<!--R4: I suggest that you refer to a specific supplementary figure or table, instead of the whole
-supplementary methods document here and elsewhere.-->
+We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods \@ref(long-numbers)).
 We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
-We used this method, implemented in the new R packages `LSUx` (https://github.com/brendanf/LSUx) and `tzara` (https://github.com/brendanf/tzara) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.
-<!--R4: I suggest that the authors describe what the LSUx and tzara packages are used for if
-these are new and being introducing for the first time here.-->
+We used this method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), for all of the PacBio and Ion Torrent datasets.
 
@@
 fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using `LSUx`.
-Shorter and more conserved regions yield a greater fraction of successfully mapped reads<!-- This sentence is a result shown in the figure (although it is also theoretically expected)-->.
+Data from all samples were analyzed as a single pool.
+*long*: entire long amplicon, including ITS1, 5.8S, ITS2, and partial LSU; *32S* partial 32S precursor to LSU, including 5.8S, ITS2, and partial LSU; *LSU*: section of LSU rDNA included in the long amplicon, from the 5' end to the LR5 primer site; *ITS*: full ITS region, including ITS1, 5.8S, and ITS2.
+Color indicates median region length.
+Shorter and more conserved regions yielded a greater fraction of successfully mapped reads<!-- This sentence is a result shown in the figure (although it is also theoretically expected)-->.
 At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs<!-- This sentence is how to interpret the figure -->.
-<!--R4: It would be nice if the colors used in part b corresponded with the same colors in part
-a. What does 32S refer to here? Is this data based on rarefied, normalized data?-->
-<!--R5: Not clear what 32S is here?-->
 
@@
 
-### Phylogenetic inference and taxonomy assignment
-<!--R4: ‘Phylogenetic inference and TAXONOMIC assignment’-->
+### Phylogenetic inference and taxonomic assignment
 
@@
 An ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
-The tree was rooted outside kingdom Fungi and non-Fungi were removed based on taxonomic assignments (see below and Supplementary methods).
-<!--R5: I realize that is covered in the Supp. methods in more detail but I would at least give one sentence about how you chose some reference sequences to parameterize the tree here-->
+Sequences confidently assigned outside kingdom Fungi by at least five of the primary taxonomic identification methods were used to root the tree, and sequences outside the clade defined by confidently identified Fungi were removed (see below and Supplementary Methods \@ref(rooting)).
 
-Taxonomic annotations of the RDP-LSU training set version 11.5 [@liu2012;@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods).
-Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
+Taxonomic annotations of the RDP-LSU training set version 11.5 [@liu2012;@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods \@ref(reannotate)).
+Primary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
 For each region/reference combination, taxonomy was assigned using three popular algorithms:
@@
 This is basically a coin toss so that is not exactly giving great confidence in the results-->
-Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
+Each full-length ASV was thus given up to nine primary taxonomic assignments (three references $\times$ three algorithms).
 ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.
 
-For full-length long-amplicon ASVs, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure\ \@ref(fig:phylotax) and Supplementary Methods).
-<!--R4: “refined using a phylogenetic tree” instead of ‘the’L275 - suggest you spend a few lines to describe what exactly what the phylotax package does if
-it’s new and you’re introducing it here for the first time-->
-This refinement algorithm is referred to as "PHYLOTAX", available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
+For full-length long-amplicon ASVs, the primary taxonomic assignments were refined with the new algorithm PHYLOTAX.
+The PHYLOTAX algorithm resolves conflicts and makes additional assignments based on an input tree (here the ML tree generated above) and taxonomic identifications from one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
+It is available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
 
-ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the preliminary assignments.
-<!--R4: a good citation is for the LCA algorithm is MEGAN Huson et al. 2016, see comment above-->
-This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [LCAClassifier, @lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
+ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the preliminary assignments [@huson2007].
+This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [e.g., MEGAN and LCAClassifier, @huson2016;@lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
 Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as "Consensus".
@@
 
-We compared alpha diversity estimates by the different sequencing strategies by calculating ASV and OTU accumulation curves, as well as comparing richness estimates after rarefaction for each sample (Supplementary methods).
-We also compared the effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, using PERMANOVA and heat tree visualizations (Supplementary methods).
+We compared alpha diversity estimates by the different sequencing strategies by calculating ASV and OTU accumulation curves, as well as comparing richness estimates after rarefaction for each sample (Supplementary Methods \@ref(sequencing-strategy-supplement)).
+We also compared the effect of sequencing technology and amplicon length on the recovered fungal community composition (i.e., after removal of non-fungi), as assessed by the Bray-Curtis dissimilarity, using PERMANOVA and heat tree visualizations (Supplementary Methods \@ref(sequencing-strategy-supplement)).
 
@@
 
-Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R (Supplementary methods).
+Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R (Supplementary Methods \@ref(spatiotemporal-supplement)).
 
@@
 
-## Reproducibility of sequence capture using different technologies
-<!--R4: ‘capture’ usually refers to the enrichment of metagenomic DNA fragments using bead
-based selection for example. Here, consider something like ‘Reproducibility of sequence
-detection using different sequencing technologies’-->
-<!--R5: This has a specific meaning for many people that is about sequences obtained via “capture” using hybridization- i would use a different heading here to make it more obvious what you are referring to-->
+## Reproducibility of sequence detection using different technologies
 
@@
 We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
-The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
-The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).<!--R5: extra %-->
-When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% of reads, respectively (Figure\ \@ref(fig:venn)b).
-In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.<!--R5: extra %-->
+The majority of abundant ASVs and OTUs were detected by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
+The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))` <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).
+When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))` of reads, respectively (Figure\ \@ref(fig:venn)b).
+In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
 ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` belonged to a unique OTU after clustering.
@@
 Blue lines are total least squares fit with 95% confidence interval, with the given slope (and 95% confidence interval) and R^2^ value.
-Dashed diagonal line indicates ideal slope of 1.
-<!--R4: Add a period after the figure title. change ‘ideal’ slope to ‘1:1 line’ or something
-similar. Is this figure based on rarefied, normalized data? Indicate in the text.-->
+Dashed diagonal line indicates 1:1 line.
 
-```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s) (ref:alpha-compare)"}
+```{r alpha-compare, fig.scap="(ref:alpha-compare-s)", fig.cap="(ref:alpha-compare-s). (ref:alpha-compare)"}
 grid::grid.newpage()
@@
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
-<!--R4: This figure needs a ‘short’ title. I’d
+<!--R4: I’d
 be interested in seeing a comparison of run times for each taxonomic assignment or a sentence
@@
 
-```{r taxon-chart, fig.cap = "(ref:taxon-chart)", fig.scap = "(ref:taxon-chart-s)", out.width = twocolwidth}
+```{r taxon-chart, fig.cap = "(ref:taxon-chart-s). (ref:taxon-chart)", fig.scap = "(ref:taxon-chart-s)", out.width = twocolwidth}
 readd(taxon_chart_plot)
@@
 The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "All Fungi") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the total fungal community and `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "ECM") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
-However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies, so no strong conclusion of variability between methods can be drawn.
-Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.
-<!--R4: confidence intervals are not shown? If you’re doing to discuss CI then consider adding
-them to your figure.-->
+However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Table\ \@ref(tab:variofit)), so no strong conclusion of variability between methods can be drawn.
+Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval (Table\ \@ref(tab:variofit)).
 
@@
 OTU read counts were strongly correlated between technologies (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon == y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`),
-and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`, Figure \@ref(fig:read-compare)b),
-which we interpret as supporting the validity of abundance-based beta diversity measures in metabarcoding [but see @castano2020 and references therein].
-<!--R4: ‘validity of abundance-based beta diversity measures’. Anyone can choose to treat their
-data as if read proportion/relative proportion is informative, but for fungi with different thallus
-sizes (yeast-like, filamentous), different nuclear states (haploid, diploid/dikaryons), abundant
-sporulators, etc. what does this really mean? This is a very contentious statement and I’m not
-sure that the data you have presented makes this case in any way. See Amend et al. 2010
-Molec Ecol 19: 5555, Amend et al. 2010 PNAS 107: 13748 supp mat, Elbrecht and Leese, 2015
-PLoS ONE 10: e0130324 for a non-fungal example, but interesting because like fungi,
-arthropods come in ranges of body sizes that differ by orders of magnitude and they are multi-
-cellular (unlike prokaryotes where relative abundance is often used, despite the fact that the
-rRNA genes exist in differently sized tandem arrays that differ across lineages). Some studies
-have even proposed ‘correction factors’ to make read counts more meaningful when assessing
-species abundance. I think this statement needs to be re-written to be more conservative, as
-your methods do not directly test whether read abundance reflects real species abundance.
-You may even want to discuss and/or cite some of the literature that has directly tackled this
-Issue.-->
-<!--R5: These are all short amplicons, and previous studies have indicated strong bias in short amplicon HTS, across different sequencing technologies.
-The correlation alone cannot be used as evidence for the validity of abundance-based diversity measures.
-Especially as these R2 values are on the borderline of what is interpreted as "strong".-->
+and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`, Figure \@ref(fig:read-compare)b).
+This lends some support for the technical repeatability of abundance-based beta diversity measures in metabarcoding, although bias at the amplification stage still presents issues [@castano2020 and references therein].
 
@@
 
-The long amplicon dataset included `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` unique taxa, even after clustering at 97% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
+The long amplicon dataset included `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` unique taxa, even after clustering at 97% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers amplify different parts of the community.
 <!--R5: If these 20% are mostly non-fungal reads then this needs to be a bigger part of the discussion!
@@
 The alternate LR5-F primer [@tedersoo2008] would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
-We assert that, for studies targeting ECM fungi in particular, more complete capture of groups with high rDNA variability such as *Tulasnella* (and ideally other Cantharellales) is worth the read-depth spent on non-target groups.
+We assert that, for studies targeting ECM fungi in particular, more complete detection of groups with high rDNA variability such as *Tulasnella* (and ideally other Cantharellales) is worth the read-depth spent on non-target groups.
 
@@
 Assignment of ecological function to environmental fungal sequences is dependent on accurate taxonomic identification, especially at the genus level or below [@nguyen2016funguild].
-However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
+However, different combinations of algorithms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
 Although RDP-LSU and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \@ref(fig:taxon-chart)).
@@
 Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, as expected given its more well-calibrated and conservative confidence scores [@murali2018a], but this was particularly dramatic when paired with the Warcup database, where IDTAXA placed <25% of ASVs even to phylum.
-<!--R5: This discussion of placement is good.
-However, it suffers from one serious weakness.
-It seems to assume that when things are placed that they are placed CORRECTLY.
-In my experience some things are obviously incorrectly placed.
-You should explicitly say whether you actually dealt with his problem or not.
-If not, at least give some caveats here!-->
 
@@
 
-For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses phylogenetic relationships to resolve these disagreements in a principled manner.
-<!--R4: in a ‘principled’ manner. Consider rephrasing to ‘uses phylogenetic relationships to
-resolve disagreements.’ Do you need to set a bootstrap support cutoff to get a good phylotax
-assignment? What cutoff did you use here? Does phylotax use neighbor joining or ML or
-Bayesian tree construction? These details need to be added to your method section in the main
-text. How does phylotax handle ambiguously aligned regions of ITS or LSU divergent domains?
-Are they removed, recoded, ignored?-->
-<!--R5: How were the results from this algorithm validated? On which taxonomic groups of fungi and/or non-fungi was it tested? How were the results of the other taxonomy methods checked for errors?-->
+For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses relationships from a provided phylogenetic tree to resolve these disagreements.
 The effect was most pronounced for the PacBio long amplicon dataset,
@@
 Deeper long-amplicon sequencing would improve the coverage of long amplicons, allowing a greater fraction of short amplicon ASVs to also be placed phylogenetically.
+
+Because our dataset was generated from environmental samples whose true taxonomic affinity is unknown, we were not able to assess the accuracy of taxonomic assignments by any of the methods used here.
+Accuracy has been assessed using leave-one-out validation for the primary assignment algorithms [e.g., @edgar2018;@murali2018] and other consensus methods [e.g., @gdanetz2017;@somervuo2016], and similar work could be carried out in the future for PHYLOTAX.
 
@@
 All of these values are smaller than the 65\ m reported by @Bahram2013, also based on distance-decay curves from a similar ECM woodland habitat in Benin, but based on Sanger sequencing of ECM root tips rather than HTS metabarcoding of bulk soil.
-We speculate that this discrepancy is due to an increased ability to capture spatially variable rare species using HTS.
+We speculate that this discrepancy is due to an increased ability to detect spatially variable rare species using HTS.
 
@@
 The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
-This corresponds to a space-for-time substitution rate of `r readd(variofit_table) %>% select(-variofit, -(std.error:conf.high)) %>% pivot_wider(names_from = term, values_from = estimate) %>% filter(timelag == 1, mantel == TRUE) %>% mutate(spacetime = range / timerange) %>% arrange(guild) %$% glue_collapse(format(spacetime, format = "f", digits = 2), " and ")`\ m/year for the total fungal community and ECM community, respecitvely.
+This corresponds to a space-for-time substitution rate (i.e., ratio of a spatial distance to a time delay which results in equivalent community dissimilarity) of `r readd(variofit_table) %>% select(-variofit, -(std.error:conf.high)) %>% pivot_wider(names_from = term, values_from = estimate) %>% filter(timelag == 1, mantel == TRUE) %>% mutate(spacetime = range / timerange) %>% arrange(guild) %$% glue_collapse(format(spacetime, format = "f", digits = 2), " and ")`\ m/year for the total fungal community and ECM community, respecitvely.
 In a recent study, @kivlin2020 reported a space-for-time substitution rate of `r format(1.5/6.8 * 365, digits = 2)`\ m/year (reported as 6.8\ d / 1.5\ m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
@@
 Year-to-year variation in ECM fungal communities, which we sampled, has been shown to be less than intra-annual variation [@bahram2015], as sampled by @kivlin2020.
-Neither dataset from the short amplicon library showed significant temporal autocorrelatio.<!--R4: autocorrelation-->
+Neither dataset from the short amplicon library showed significant temporal autocorrelation.
 

### 2021-02-05 writing/transect_supplement.Rmd
address more comments...

@@
 <!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
-A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
+A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal LSU RNA seed alignment from RDP release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
 A CM was generated from the alignment using `cmbuild` from Infernal.
@@
 
-### Uniform taxonomic classification for Unite, Warcup, and RDP-LSU
+### Uniform taxonomic classification for Unite, Warcup, and RDP-LSU {#reannotate}
 
@@
 
-### Rooting the fungal tree
+### Rooting the fungal tree {#rooting}
 
@@
 
-### Refining taxonomic assignments with a phylogenetic tree: `PHYLOTAX`
+### Refining taxonomic assignments with a phylogenetic tree: `PHYLOTAX` {#phylotax}
 

### 2021-02-16 writing/transect_paper_rev1.tex
add latex files for first revision (to make diffs)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities},
  pdfauthor={Brendan Furneaux1,*,; Mohammad Bahram2,3,; Anna Rosling4,; Nourou S. Yorou5,; Martin Ryberg1},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=2.5cm,right=2.5cm,top=3cm,bottom=3cm]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% line numbers -- only needed for MS
\usepackage{lineno}
\linenumbers

% put all tables and figures at the end -- only needed for MS
\usepackage[nomarkers]{endfloat}

% double spacing -- only needed for MS
\usepackage{setspace}
\doublespacing

% table and figure caption width
\usepackage[width=0.9\textwidth]{caption}

% crossreferences to supporting information -- transect_supplement.aux is required
\usepackage{xr}
\externaldocument{transect_supplement}
\usepackage[style=apa,]{biblatex}
% Produce [dataset] references
\NewBibliographyString{dataset}

\DefineBibliographyStrings{english}{
  dataset = {dataset},
}

\renewbibmacro{begentry}{%
  \ifkeyword{meta}{\textsuperscript{*}}{}%
  \ifentrytype{dataset}
    {\bibstring[\mkbibbrackets]{dataset}%
     \setunit{\addspace}}
    {}%
}
\addbibresource{all.bib}
\addbibresource{R.bib}

\title{Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities}
\author{Brendan Furneaux\textsuperscript{1,*}, \and Mohammad Bahram\textsuperscript{2,3}, \and Anna Rosling\textsuperscript{4}, \and Nourou S. Yorou\textsuperscript{5}, \and Martin Ryberg\textsuperscript{1}}
\date{\textsuperscript{1}Program in Systematic Biology, Department of Organismal Biology, Uppsala University, Uppsala, Sweden\\
\textsuperscript{2}Department of Ecology, Swedish University of Agricultural Sciences, Uppsala, Sweden\\
\textsuperscript{3}Institute of Ecology and Earth Sciences, University of Tartu, Tartu, Estonia\\
\textsuperscript{4}Program in Evolutionary Biology, Department of Ecology and Genetics, Uppsala University, Uppsala, Sweden\\
\textsuperscript{5}Research Unit in Tropical Mycology and Plant-Fungi Interactions, LEB, University of Parakou, Parakou, Benin\\
\textsuperscript{*}Corresponding author, \texttt{brendan.furneaux@ebc.uu.se}}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

Fungi form diverse communities and play essential roles in many terrestrial ecosystems, yet there are methodological challenges in taxonomic and phylogenetic placement of fungi from environmental sequences.
To address such challenges we investigated spatio-temporal structure of a fungal community using soil metabarcoding with four different sequencing strategies: short amplicon sequencing of the ITS2 region (300--400~bp) with Illumina MiSeq, Ion Torrent Ion S5, and PacBio RS II, all from the same PCR library, as well as long amplicon sequencing of the full ITS and partial LSU regions (1200--1600~bp) with PacBio RS II.
Resulting community structure and diversity depended more on statistical method than sequencing technology.
The use of long-amplicon sequencing enables construction of a phylogenetic tree from metabarcoding reads, which facilitates taxonomic identification of sequences.
However, long reads present issues for denoising algorithms in diverse communities.
We present a solution that splits the reads into shorter homologous regions prior to denoising, and then reconstructs the full denoised reads.
In the choice between short and long amplicons, we suggest a hybrid approach using short amplicons for sampling breadth and depth, and long amplicons to characterize the local species pool for improved identification and phylogenetic analyses.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Fungi are key drivers of nutrient cycling in terrestrial ecosystems.
One important guild of fungi form ectomycorrhizas (ECM), a symbiosis between fungi and plants in which fungal hyphae enclose the plant's fine root tips.
The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant \autocite{Smith2010}.
Approximately 8\% of described fungal species are thought to take part in ECM symbiosis \autocite{ainsworth2008ainsworth,Rinaldi2008}.
Although only about 2\% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, \emph{Uapaca} (Phyllanthaceae) and Fabaceae tr. Amherstieae \autocite{brundrett2017a}, together representing approximately 60\% of tree stems globally \autocite{steidinger2019}.

Although ECM fungi form many well-known mushrooms (e.g., \emph{Amanita}, \emph{Cantharellus}, \emph{Boletus}), some instead produce inconspicuous (e.g., \emph{Tomentella}) or no (e.g., \emph{Cenococcum}) fruit bodies.
Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by sampling of vegetative structures \autocite{horton2001}.
Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures in the field.
The advent of high-throughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils \autocite{Lindahl2013}.

As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
Fungal metabarcoding studies using short-read HTS technologies such as 454 Pyrosequencing, Illumina, and Ion Torrent
have usually targeted the rDNA internal transcribed spacer regions ITS1 or ITS2,
which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS \autocite{schoch2012,Lindahl2013}.
In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice \autocite{opik2010}, and variable regions of the rDNA large subunit (LSU) have also been used for barcoding \autocite{house2016,tedersoo2015}.
The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis \autocite{Lindahl2013}.
If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST \autocite{altschul1990,Lindahl2013} with public databases such as GenBank \autocite{benson2013} and Unite \autocite{nilsson2019a}.
However, this approach comes with some potential weaknesses, discussed below.

Different sequencing technologies have different capabilities in terms of sequencing depth and read length, as well as differing quality profiles and potential biases \autocite{yang2013}.
The rapid development of new HTS technologies, as well as subsequent iterative improvements in sequencing chemistry and read capacity, means that the technologies used in metabarcoding studies, along with any associated biases, change frequently.
As an example, the first study using HTS metabarcoding of soil fungi was published in 2009 \autocite{buee2009} using 454 Pyrosequencing; production of 454 sequencers was subsequently discontinued in 2015, and sales of reagents stopped in 2016 \autocite{hollmer2013}.
This brings into question the comparability of studies conducted only a few years apart.

ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina \autocite{schoch2012},
but this variablity means that they cannot be reliably aligned over the fungal kingdom \autocite{Lindahl2013,Tedersoo2018}.
Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa \autocite{ihrmark2012,tedersoo2015,palmer2018}.
Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi \autocite{tedersoo2015}.

Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
In contrast, more recent denoising methods such as DADA2 \autocite{callahan2017}, Deblur \autocite{amir2017}, and UNOISE2 \autocite{edgar2016} utilize read quality information to control for sequencing error while preserving intra-species variation.
The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated \autocite{glassman2018,botnen2018}.
ASVs have been suggested to be less dataset specific that cluster-based OTUs \autocite{callahan2017}.
Support for PacBio has recently been added to DADA2 \autocite{callahan2019}, but its application requires greater sequencing depth for longer reads, especially in high diversity samples.

Because both OTU clustering and denoised ASVs may ``clump'' different species into a single unit and ``split'' a single species into multiple units \autocite{ryberg2015}, diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
In contrast, phylogenetic community distance measures \autocite{wong2016} are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree \autocite{matsen2010,berger2011}, but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments \autocite{tedersoo2018}, or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases \autocite{fouquier2016}, but these approaches are only applicable to sequences of known taxonomic affiliation.

Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
Although the public INSDC databases \autocite{karsch-mizrachi2018} are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations \autocite{nilsson2006,steinegger2020} as well as sequences of poor technical quality \autocite{nilsson2012}.
Consequently taxonomic assignments based on these databases may be incorrect or inconsistent.
Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
The Unite database is a more curated attempt to include all publically available ITS sequences (800~000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes \autocite{nilsson2019a}, where efforts have been made to correct incorrect annotations and exclude low-quality sequences \autocite{abarenkov2018a}.
The Ribosomal Data Project \autocite[RDP,][]{cole2014} hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
the Warcup ITS database, containing 18~000 manually curated fungal ITS sequences \autocite{deshpande2016}, and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups \autocite{liu2012}.
Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.

Assigning taxonomy to unknown sequences using BLAST requires \emph{a priori} choice of similarity thresholds for different taxonomic ranks.
Several algorithms specifically designed for taxonomic assignment have been published which instead use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier \autocite[RDPC,][]{wang2007}, SINTAX \autocite{edgar2016a} and IDTAXA \autocite{murali2018a} among others.
In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments \autocite{somervuo2016,gdanetz2017,palmer2018}.
However, all sequence-similarity based approaches are dependent on high taxonomic coverage in the reference database, making the placement of novel or undersampled groups problematic \autocite{nilsson2016top,Tedersoo2018}.

Recent long-read HTS technologies such as Pacific Biosciences (PacBio) Single Molecule Real Time (SMRT) sequencing enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
This can improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis \autocite{Tedersoo2018}.
Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number \autocite{Kennedy2018}.

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
Turnover scale varies between different ecosystems and taxonomic groups,
and can be measured by the range at which a Mantel correlogram indicates significant autocorrelation,
or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.~distance \autocite{legendre2012}.

We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
with three different reference databases (Unite, Warcup, RDP),
and both non-phylogenetic and phylogenetic community distance measures.
We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{sampling}{%
\subsection{Sampling}\label{sampling}}

Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Caesalpinioideae).
At each site, 25~soil samples were collected along a single 24~m linear transect at intervals of 1~m in May 2015.
One third of the sample locations (3~m spacing) were resampled one year later in June 2016, for a total of 67 samples.
For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with a sterilized knife blade and homogenized.
A subsample of approximately 250~mg was preserved and returned to the laboratory for extraction (see Supplementary methods).

\hypertarget{dna-amplification-and-sequencing}{%
\subsection{DNA amplification and sequencing}\label{dna-amplification-and-sequencing}}

DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure~\ref{fig:regions}a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
The short amplicon (approximately 300~bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 \autocite{ihrmark2012} as the forward primer and a mix of ITS4 \autocite{white1990amplification} and ITS4a \autocite{urbina2016} as the reverse primer.
The long amplicon (approximately 1500~bp) targeted the full ITS region including the 5.8S rDNA and approximately 950~bp at the 5' end of the LSU, including the first three variable regions (Figure~\ref{fig:regions}a), using ITS1 \autocite{white1990amplification} as the forward primer and LR5 \autocite{vilgalys1990} as the reverse primer.
Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of \emph{Agaricus bisporus}.
See Supplementary methods for PCR and library preparation protocols.

Each library was sequenced on a PacBio RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752~37 UPPSALA, Sweden).
Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.
Additionally, the same short amplicon PCR libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300~bp at the SNP\&SEQ Technology Platform (Dept. of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751~44 UPPSALA, Sweden) using one half of a lane.
Platform-specific library preparation, including adapter ligation, was performed at the sequencing facilities according to their standard protocols.

\hypertarget{bioinformatics}{%
\subsection{Bioinformatics}\label{bioinformatics}}

Circular consensus sequence (CCS) basecalls for PacBio sequences were made using \texttt{ccs} version 3.4 \autocite{pacificbiosciences2019} using the default settings.
The resulting sequences, as well as the paired-end Illumina sequences, were demultiplexed and sequencing primers were removed using \texttt{cutadapt} version 2.8 \autocite{martin2011}.
Sequencing primers were similarly removed from the Ion Torrent sequences, but interference between the tagged gITS7 primers and the Ion XPress tags used in library prep made full demultiplexing of the Ion Torrent sequences impossible, resulting in two samples sharing each tag.
These reads were thus either analyzed as a pool, or comparisons were made to equivalently combined samples in the other datasets.
For Ion Torrent and PacBio, reads were discarded if they did not have the appropriate primers on both ends.
Reads were searched in both directions, and reads where the primers were found in the reverse direction were reverse complemented before further analysis.
For Illumina sequences, read pairs were only retained when PCR primers were detected at the 5' ends of both the forward and reverse read.
Primers were also searched for and removed on the 3' ends of the reads, in case of readthrough with short amplicons.
Read pairs where the primers were found in reverse orientation were kept in separate files, but were retained in their original orientation until after denoising.

\hypertarget{denoising-and-clustering}{%
\subsubsection{Denoising and clustering}\label{denoising-and-clustering}}

All amplicons were denoised using DADA2 version 1.12.1 according to the ITS pipeline workflow \autocite{callahan2016,callahan2020}, with technology-specific modifications for Ion Torrent \autocite{callahan2020b} and PacBio \autocite{callahan2019}.
Although this was successful for the short amplicons on all technologies, only 38 ASVs were obtained for the long amplicons, representing 12\% of the trimmed reads.

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU \autocite{michot1984}, and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure~\ref{fig:regions}a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
We used this method, implemented in the new R packages \texttt{LSUx} (\url{https://github.com/brendanf/LSUx}) and \texttt{tzara} (\url{https://github.com/brendanf/tzara}) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/regions-1}

}

\caption[rDNA regions]{rDNA regions. \textbf{(a)} Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions. D1--3 represent the first three variable regions in LSU, while LSU1--4 represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces. \textbf{(b)} Total number of DADA2 ASVs vs.~fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using \texttt{LSUx}. Shorter and more conserved regions yield a greater fraction of successfully mapped reads. At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs.}\label{fig:regions}
\end{figure}

Because the \texttt{LSUx} plus \texttt{tzara} method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using \texttt{LSUx} for comparison to the results from the other technologies.

To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97\% similarity using VSEARCH v2.9.1 \autocite{rognes2016}.

\hypertarget{phylogenetic-inference-and-taxonomy-assignment}{%
\subsubsection{Phylogenetic inference and taxonomy assignment}\label{phylogenetic-inference-and-taxonomy-assignment}}

Full length long amplicon ASVs were aligned using DECIPHER \autocite{wright2015} with up to 10~iterations of alternating progressive alignment and conserved RNA secondary structure calculation, followed by 10~refinement iterations.
This alignment was truncated at a position after the D3 region corresponding to base 907 of the \emph{Saccharomyces cerevisiae} S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by \textcite{holst-jensen1999}.
An ML tree was produced using RAxML version 8.2.12 \autocite{stamatakis2014} using the GTR+GAMMA model and rapid bootstrapping with the MRE\_IGN stopping criterion.
The tree was rooted outside kingdom Fungi and non-Fungi were removed based on taxonomic assignments (see below and Supplementary methods).

Taxonomic annotations of the RDP LSU fungal training set version 11.5 \autocite{cole2014} and Warcup ITS training set \autocite{deshpande2016} were mapped to a uniform taxonomic classification system (see Supplementary Methods).
Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three popular algorithms:
the RDP Naïve Bayesian Classifier \autocite[RDPC,][]{wang2007} as implemented in DADA2;
SINTAX \autocite{edgar2016a} as implemented in VSEARCH v2.9.1 \autocite{rognes2016};
and IDTAXA \autocite{murali2018}.
A relatively lax confidence threshhold of 50\% was used for all three algorithms.
Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references \(\times\) three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

For full-length long-amplicon ASVs, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure~\ref{fig:phylotax} and Supplementary Methods).
This refinement algorithm is referred to as ``PHYLOTAX'', available in the new R package \texttt{phylotax} at \url{https://github.com/brendanf/phylotax}.

ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the ``last common ancestor'' of the preliminary assignments.
This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits \autocite[LCAClassifier,][]{lanzen2012} or k-mer similarity scores \autocite[mothur's k-nearest neighbor method,][]{schloss2009}, but here is used to resolve conflicts between assignments from different algorithms and databases.
Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as ``Consensus''.

\hypertarget{effect-of-sequencing-strategy-on-recovered-community}{%
\subsection{Effect of sequencing strategy on recovered community}\label{effect-of-sequencing-strategy-on-recovered-community}}

We compared alpha diversity estimates by the different sequencing strategies by calculating ASV and OTU accumulation curves, as well as comparing richness estimates after rarefaction for each sample (Supplementary methods).
We also compared the effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, using PERMANOVA and heat tree visualizations (Supplementary methods).

\hypertarget{spatiotemporal-analysis}{%
\subsection{Spatiotemporal analysis}\label{spatiotemporal-analysis}}

To estimate turnover scale, ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric \autocite[for both long and short amplicons]{bray1957} and the phylogenetically based weighted UniFrac metric \autocites[for only long amplicons]{lozupone2005,lozupone2007} in \texttt{phyloseq} version 1.26.0.
Each of these distance matrices was used to calculate a Mantel correlogram with a 1~m bin size for distances in the range of 0--12~m, i.e., half the maximum separation present in the dataset.
Separate correlograms were drawn for samples taken during the same year and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.

Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by \textcite{legendre2012} using the \texttt{nls} function in R (Supplementary methods).

\hypertarget{results}{%
\section{Results}\label{results}}

Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100~reads (Figure \ref{fig:sample-depth}).
Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table~\ref{tab:bioinfo}).
Sequencing with PacBio yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
Ion Torrent and Illumina yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads (\(\le 1\) expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \ref{fig:qual-check}b).
Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \ref{fig:qual-check}a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \ref{fig:qual-check}b).
Demultiplexing, primer trimming, and quality filtering reduced the read totals by 64\% for PacBio long amplicons, but only by 21\% for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with 41\% and 28\% loss, respectively.
Extraction of only the ITS2 region before quality filtering (Figure \ref{fig:qual-check}c) reduced the loss of long amplicon PacBio reads to only 29\%, comparable to Illumina.
Application of \texttt{tzara} resulted in 708 reconstructed long-amplicon ASVs, representing 97\% of denoised ITS2 reads from the long-amplicon PacBio dataset.
Mapping identical ITS2 ASVs from the short and long amplicon datasets allowed 58\%, 71\%, and 81\% of denoised reads from the Ion Torrent, Illumina, and PacBio short amplicon datasets, respectively, to be assigned to a long amplicon ASV (Table \ref{tab:bioinfo}).

Almost all of the short amplicon sequences from all three technologies were between 240 and 375~bp long (Figure~\ref{fig:full-length}a).
Although the length profile of the three sequencing runs were similar, Illumina had the largest fraction of reads near the top of the range, followed by Ion Torrent and PacBio (Figure~\ref{fig:full-length}b).
The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = 8.57e+04, \(p < 2.2\times10^{-16}\)),
but the difference between means was fairly small, with mean amplicon lengths of 276, 281, and 286 bp for PacBio, Ion Torrent, and Illumina, respectively.
The length of the long amplicon reads varied widely, from 696 to 1638 bp, with a mean of 1431 bp (Figure~\ref{fig:full-length}c).

Among the different regions extracted from the long amplicon (Figure~\ref{fig:region-lengths},
ITS1 showed the greatest length variability (mean \(\pm\) standard deviation: 193 \(\pm\) 55~bp), followed by ITS2 (184 \(\pm\) 41~bp) and the variable regions in LSU (D2: 227 \(\pm\) 36~bp; D3: 108 \(\pm\) 10~bp; D1: 159 \(\pm\) 6~bp).
Approximately 2\% of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure~\ref{fig:region-lengths} due to rarity.
Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations \textless{} 2 bp.
Around 12\% of ITS2 sequences extracted from the long amplicon dataset were shorter than 140 bp, a much greater fraction than the 0.26\% to 0.44\% from the short amplicon datasets (Figure \ref{fig:ITS2-length}).
The taxonomic identity of these sequences is discussed below.

\emph{Agaricus bisporus}, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
\emph{A. bisporus} was represented by two ASVs in the Illumina dataset, which differed at one base pair (99.5\% similarity in ITS2).
The abundance of the second ASV was 1.1\% and 1.0\% that of the primary \emph{A. bisporus} ASV in the two Illumina positive controls.
The consistency of this ratio across replicate positive controls suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
Despite higher total sequencing depth, this ASV was not identified from the Ion Torrent dataset.

\emph{A. bisporus} sequences represented 0.01\%, 0.09\%, 0.09\%, and 0.09\% of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively, giving similar estimates for the rate of tag-switching for all technologies.
These reads were excluded from community analyses.

\hypertarget{reproducibility-of-sequence-capture-using-different-technologies}{%
\subsection{Reproducibility of sequence capture using different technologies}\label{reproducibility-of-sequence-capture-using-different-technologies}}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/venn-1} \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/venn-2}

}

\caption[Shared richness and abundance of ASVs and OTUs between different sequencing technologies]{Shared richness and abundance of ITS2-based ASVs (\emph{a}, \emph{c}) and 97\% OTUs (\emph{b}, \emph{d}) between different sequencing technologies from the same short amplicon library (\emph{a}, \emph{b}), and between long and short amplicon libraries (\emph{c}, \emph{d}).
In each region, the ASV/OTU richness is given above, while the relative abundance of reads represented by these ASVs/OTUs in each sequencing strategy are shown below in the order PacBio/Illumina/Ion Torrent (\emph{a}, \emph{b}), or long/short (\emph{c}, \emph{d}).
For short amplicons in \emph{c} and \emph{d}, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.}\label{fig:venn}
\end{figure}

We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables~\ref{tab:venn-table-asv} and \ref{tab:venn-table-otu}).
The short amplicon ASVs shared between all sequencing technologies represented 95\%, 76\%, and 66\%\% of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure~\ref{fig:venn}a).
When differences at the intra-species scale were removed by clustering the ASVs into 97\% OTUs, the number of OTUs shared between all three technologies increased to 524, representing 100\%, 93\% , and 89\% \% of reads, respectively (Figure~\ref{fig:venn}b).
In particular, the majority of the 9418 unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
ASVs unique to the Ion Torrent dataset made up 14\% of reads in that dataset, but only 1\% belonged to a unique OTU after clustering.
In contrast, 21\% of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure~\ref{fig:venn}c), and the fraction only reduced to 20\% after clustering the ITS2 regions into OTUs (Figure~\ref{fig:venn}d).

Read counts for shared ASVs and OTUs were highly correlated between strategies, with a minimum \(R^2\) value of 0.47 (Figure~\ref{fig:read-compare}).
Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
(0.69 to 0.72,
0.49 to 0.74, and
0.74 to 0.82,
for PacBio vs.~Illumina, PacBio vs.~Ion Torrent, and Illumina vs.~Ion Torrent, respectively), but not between the long amplicon library and short amplicon library
(0.65 to 0.62,
0.58 to 0.57, and
0.47 to 0.49, for PacBio long amplicon reads vs.~PacBio, Illumina, and Ion Torrent short amplicon reads, respectively; Figure~\ref{fig:read-compare}).

ASV richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library (\(R^2 =\) 0.91 to 0.94; Figure~\ref{fig:alpha-compare}).
The slope of the relationship between PacBio and Illumina richness estimates was only slightly different from 1, indicating that these two technologies give highly comparable rarefied richness estimates, despite the approximately 200× difference in original sequencing depth.
Ion Torrent resulted in rarefied richness estimates which were 24\% to 31\% greater than the other technologies, an effect which is also visible in ASV accumulation curves (Figure~\ref{fig:accum-ASV}).
ASV richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets (\(R^2 =\) 0.65 to 0.72; Figure~\ref{fig:alpha-compare}).
Total least squares regression indicated that the long amplicon dataset resulted in richness estimates which were intermediate between the short amplicon results from Ion Torrent and the other two technologies.
Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure~\ref{fig:venn}), sample-wise OTU accumulation curves (Figure~\ref{fig:accum-OTU}) and rarefied OTU richness relationships between sequencing strategies (Figure~\ref{fig:alpha-compare-otu}) were highly similar to those for ASVs.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/accum-ASV-1}

}

\caption[ASV accumulation curves]{ASV accumulation curves. Each curve represents rarefaction of a single sample or pooled pair of samples. Points at the end of each curve represent the actual read depth and observed ASV richness. Note logarithmic axes.}\label{fig:accum-ASV}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/alpha-compare-1}

}

\caption[Comparison of ASV richness between sequencing technologies]{Comparison of ASV richness between sequencing technologies Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies. All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads. Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction. Blue lines are total least squares fit with 95\% confidence interval, with the given slope (and 95\% confidence interval) and R\textsuperscript{2} value. Dashed diagonal line indicates ideal slope of 1.}\label{fig:alpha-compare}
\end{figure}

\hypertarget{taxonomic-assignment}{%
\subsection{Taxonomic assignment}\label{taxonomic-assignment}}

For all sequencing datasets and taxonomic assignment protocols, a higher proportion of reads were assigned than of ASVs, indicating that common ASVs were more likely to be taxonomically identified than rare ASVs (Figure~\ref{fig:taxon-chart}).
A greater fraction of ITS reads and ASVs were assigned using the Unite database than the Warcup database across sequencing technologies, amplicons, algorithms, and taxonomic ranks.
At most taxonomic ranks, the RDPC algorithm assigned the greatest fraction of reads and ASVs, followed by SINTAX, and then IDTAXA.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_paper_files/figure-latex/taxon-chart-1}

}

\caption[Summary of taxonomic assignments]{Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.}\label{fig:taxon-chart}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/fungi-comm-1}

}

\caption[Taxonomic composition of fungal community at the class level]{Taxonomic composition of fungal community at the class level. Values represent the fraction of all ASVs and reads which were assigned to kingdom Fungi.
Assignments based on PHYLOTAX.
Classes which represented less than 2\% of reads and ASVs in all datasets are grouped together as ``other''.}\label{fig:fungi-comm}
\end{figure}

Taxonomic composition of the sequenced soil fungal community at the class level is summarized in Figure~\ref{fig:fungi-comm} and as a heat tree \autocite{foster2017} in Figure~\ref{fig:taxa}.
The ML tree for fungal ASVs, along with taxonomic assignments, is shown in Supplementary File 3.
According to the PHYLOTAX assignments, Fungi represented 88\% of the ASVs and 81\% of the reads in the long amplicon library, compared to 92.4\%--96.5\% of the ASVs and 97.9\%--98.5\% of the reads in the short amplicon library.
Many of the ASVs which were unique to the long-amplicon library thus fall outside kingdom Fungi (Figure~\ref{fig:heattree-length-compare}).
In particular a large fraction of ITS2 sequences with length less than 140 (Figure~\ref{fig:ITS2-length}) were identified as Alveolates (Figure~\ref{fig:short-its}).

Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, \(p<0.0001\), \(R^2=0.046\)), but only marginally between sequencing technologies (\(p=0.0669\), \(R^2=0.002\)).
The majority of variation was spatiotemporal (i.e., between samples; \(p<0.0001\), \(R^2=0.90\)), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures~\ref{fig:fungi-comm}, \ref{fig:heattree-length-compare}, and \ref{fig:ppcoa}).
Additionally, several lower-rank taxonomic groups showed increased detection in either the long or short datasets, such as Tulasnellaceae (Agaricomycetes) and Pyronemataceae (Pezizomycetes) in the long amplicon dataset, and \emph{Myerozyma} (Saccharomycetes) in the short amplicon datasets (Figures~\ref{fig:heattree-length-compare} and \ref{fig:ecm-heattree}).

Fungi categorized as ECM made up 8.9\% of ASVs and 39.2\% of reads in the long amplicon library, and 5.4\%--13.4\% of the ASVs and 36.4\%--46.4\% of the reads in the short amplicon library (Figure~\ref{fig:ecm}).
Although amplicon length had a significant effect on ECM community composition at the family level (Figure~\ref{fig:ecm-heattree}), the explained variation was very low (PERMANOVA with 9999 permutations, \(p=0.0040\), \(R^2=0.002\)), and the majority of variation was again spatiotemporal (\(p<0.0001\), \(R^2=0.98\)).
Variation between sequencing technologies was not significant (\(p=0.76\), \(R^2=0.0002\)).

\hypertarget{spatial-analysis}{%
\subsection{Spatial analysis}\label{spatial-analysis}}

Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at \(p < 0.05\) for ranges of up to 2--3~m for the total fungal community, and 1--2~m for the ECM fungal community (Figure~\ref{fig:correlog}).
In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
The least spatial structure was detected with PacBio short amplicon sequencing.

The Bray-Curtis metric showed positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0~m, time lag of 1~year), for both the total fungal and ECM fungal communities, although this result did not reach statistical significance for all sequencing strategies.
This spatiotemporal correlation did not extend to a range of 1~m, and in fact correlation was negative at a time lag of 1~year and distance of 1~m, indicating that samples collected 1~m apart in different years were more different than randomly selected pairs of samples.
This negative correlation, which reached marginal statistical significance in the PacBio short amplicon dataset, was probably a statistical artifact.

In contrast to the Bray-Curtis distance, the weighted UniFrac distance showed very little spatial structure, with only the total fungal community in the 1~m
distance class showing a significant correlation at \(p < 0.05\).
No temporal correlation was found for the weighted UniFrac distance.

The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from 13--31~m for the total fungal community and 12--42~m for the ECM fungal community (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}).
However, there was overlap of the 95\% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies, so no strong conclusion of variability between methods can be drawn.
Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0~m, indicating no spatial structure, was included in the 95\% confidence interval.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/spatio-1}

}

\caption[Distance-decay plot for community dissimilarities and spatio-temporal distance]{Distance-decay plot for community dissimilarities and spatio-temporal distance. Circles represent community data from short (top two rows) and long (bottom two rows) amplicon libraries, sequenced by Illumina MiSeq (top row) or PacBio RS II (bottom three rows).
Community dissimilarities are calculated using the Bray-Curtis dissimilarity for all datasets (top three rows) and using the weighted UniFrac dissimilarity for the long amplicon library, for which a phylogenetic tree could be constructed (bottom row).
The left column represents the full fungal community, and the right column only sequences identified as ECM.
The color of each circle represents the time lag between samples being compared (0 or 1~year), and the size represents the number of comparisons for that spatial distance and time lag.
Lines are the best-fit lines for an exponential decay to max model.
The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0~year timelag) or spatiotemporal (for the 1~year time lag) distance.}\label{fig:spatio}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{reconstruction-of-long-amplicons-from-denoised-subregions}{%
\subsection{Reconstruction of long amplicons from denoised subregions}\label{reconstruction-of-long-amplicons-from-denoised-subregions}}

Sequencing depth in the long amplicon PacBio dataset was not sufficient to successfully denoise using standard protocols, given the amplicon length and diversity of the samples.
ASV recovery for long amplicons using DADA2 was dramatically improved from 12\% to 76\% of reads by denoising homologous subregions independently using our new \texttt{LSUx} and \texttt{tzara} packages.
Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
Thus, \texttt{tzara} should increase recovery of rare ASVs from these platforms as well.
It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding \autocite{loit2019}.

\hypertarget{comparison-of-sequencing-strategies}{%
\subsection{Comparison of sequencing strategies}\label{comparison-of-sequencing-strategies}}

The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (75\%) compared to Illumina (63\%) or Ion Torrent (65\%),
the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered, and were much closer to saturation of their respective species accumulation curves (Figure \ref{fig:accum-ASV}).
OTU read counts were strongly correlated between technologies (R\textsuperscript{2} = 0.72--0.82),
and even between primer pairs (R\textsuperscript{2} = 0.49--0.62, Figure \ref{fig:read-compare}b),
which we interpret as supporting the validity of abundance-based beta diversity measures in metabarcoding \autocite[but see][ and references therein]{castano2020}.

DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97\% led to substantially higher correspondence between both the set of OTUs recovered from the same library by different technologies and the read counts for each OTU (Figure~\ref{fig:read-compare}).
ASV diversity appears to be artificially inflated in the Ion Torrent dataset relative to the Illumina and PacBio datasets, which gave remarkably similar ASV richness after rarefaction, despite a difference of around 200x in unrarefied sequencing depth (Figures~\ref{fig:venn} and \ref{fig:alpha-compare}).
This may be a result of the lower fraction of very high quality reads in the Ion Torrent dataset (Figure \ref{fig:qual-check}).
We used options for DADA2 intended to improve performance on technologies, like Ion Torrent, with higher rates of homopolymer indel errors \autocite{callahan2020b}, but our results suggest that this still does not result in performance comparable to that which DADA2 achieves on Illumina sequences, for which it was developed \autocite{callahan2016}.

Although the longer read length capabilities of PacBio allow recovery of longer ITS2 sequences than the other two technologies, as has recently been demonstrated in mock communities \autocite{castano2020}, in our dataset from a natural community PacBio did not recover any reads from the short amplicon library which were longer than those recovered by Illumina and Ion Torrent.
Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to \emph{Cantharellus}, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observations by BF and NSY), but which is also known to have accelerated evolution in the rDNA \autocite{moncalvo2006} and longer ITS regions than other fungi \autocite{Feibelman_1994}, making it an especially difficult target for metabarcoding.
Contrary to expectations, Illumina showed a slightly \emph{higher} fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Figures~\ref{fig:full-length} and \ref{fig:ITS2-length}).

The long amplicon dataset included 20\% unique taxa, even after clustering at 97\% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae \autocite{tedersoo2015}.
ITS1 and LR5 match a much broader range of fungal and other eukaryote groups \autocite{tedersoo2015}. The alternate LR5-F primer \autocite{tedersoo2008} would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
We assert that, for studies targeting ECM fungi in particular, more complete capture of groups with high rDNA variability such as \emph{Tulasnella} (and ideally other Cantharellales) is worth the read-depth spent on non-target groups.

\hypertarget{taxonomic-identification}{%
\subsection{Taxonomic identification}\label{taxonomic-identification}}

Assignment of ecological function to environmental fungal sequences is dependent on accurate taxonomic identification, especially at the genus level or below \autocite{nguyen2016funguild}.
However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
Although the RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \ref{fig:taxon-chart}).
This is probably due to two factors.
First, the Warcup database does not include any non-fungi, so it cannot place any non-fungal sequences.
Second, due to its low-density coverage of the fungal kingdom (18~000 sequences vs.~800~000 for Unite), it is likely that many ITS sequences, especially from uncultured tropical soil fungi, have no close match in the Warcup database, and so cannot be placed.
The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
\textcite{heeger2019} also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
Of the three algorithms tested, IDTAXA placed fewer sequences than RDPC or SINTAX with all databases, as expected given its more well-calibrated and conservative confidence scores \autocite{murali2018a}, but this was particularly dramatic when paired with the Warcup database, where IDTAXA placed \textless25\% of ASVs even to phylum.

\textcite{gdanetz2017} showed that a majority-rule consensus of three assignment algorithms can improve the fraction of sequences assigned as well as decrease the false assignment rate.
Strict consensus rejects assignments whenever there is conflict between methods and should therefore provide more conservative taxonomic assignments than majority-rule consensus.
AMPtk \autocite{palmer2018} uses a strict consensus taxonomy between UTAX and SINTAX as an alternative when an initial BLAST search failed to give a hit with at least 97\% sequence identity, but did not present results assessing the results of this approach.
Here, we found that strict consensus also usually increases the number of assigned sequences relative to any single method, except at family and genus level identifications (Figure \ref{fig:taxon-chart}).
Inconsistent family and genus level assignments are particularly problematic because accurate assignment at these ranks is generally required for ecological guild assignment using FUNGuild.

For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses phylogenetic relationships to resolve these disagreements in a principled manner.
The effect was most pronounced for the PacBio long amplicon dataset,
where 46\% and 62\% of reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to 73\% and 81\%.
This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure~\ref{fig:ecm}).
For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs (last row of Table~\ref{tab:bioinfo}).
Deeper long-amplicon sequencing would improve the coverage of long amplicons, allowing a greater fraction of short amplicon ASVs to also be placed phylogenetically.

\hypertarget{turnover-rate}{%
\subsection{Turnover rate}\label{turnover-rate}}

Mantel correlograms based on the Bray-Curtis dissimilarity (Figure~\ref{fig:correlog}) revealed spatial autocorrelation in the soil fungal community at distance classes \(\le 3\) m for both Illumina and PacBio using long and short amplicons, and in the ECM fungal community at distance classes \(\le 2\) m for Illumina and PacBio long amplicons, and \(\le 1\) m for the PacBio short amplicons.
These results are similar to autocorrelation ranges found in previous work based on ECM root tips in temperate forests \autocite{lilleskov2004,pickles2012a}.
\textcite{lilleskov2004} found autocorrelation only at ranges \textless2.6~m at most sites using Sanger sequencing.
Similarly, \textcite{pickles2012a} found autocorrelation at distances \textless3.4~m based on T-RFLP analysis.
Previous work in Miombo woodland, a similar ecosystem to the Sudanian woodland in this study, found autocorrelation at ranges \textless10~m using Sanger sequencing of ECM root tips \autocite{tedersoo2011}, which was their smallest distance class.

Distance-decay plots (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}) gave substantially longer autocorrelation distances.
There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 12--18~m.
The 95\% confidence interval was substantially wider than this variation, generally covering a range of 5--41~m.
All of these values are smaller than the 65~m reported by \textcite{Bahram2013}, also based on distance-decay curves from a similar ECM woodland habitat in Benin, but based on Sanger sequencing of ECM root tips rather than HTS metabarcoding of bulk soil.
We speculate that this discrepancy is due to an increased ability to capture spatially variable rare species using HTS.

For the short amplicon dataset, PacBio showed a spatial turnover range more than twice as long as showed by Illumina (Table \ref{tab:variofit}) for both the total fungi and ECM communities, with wide confidence intervals.
It is possible that the weaker fit for this dataset, which also showed weaker autocorrelation in the Mantel correlogram, is due to low sequencing depth in the PacBio short amplicon dataset.
The long amplicon PacBio dataset, with more than twice the read depth of the short amplicon PacBio dataset, gave spatial turnover distance results much closer to those from Illumina.
This is consistent with our speculation that the longer spatial turnover range found by \textcite{Bahram2013} is related to sequence sampling depth.

Year-to-year correlation was found for both the total fungal and ECM communities in the long amplicon dataset (Figure \ref{fig:correlog}).
The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.3~years for the total fungal community and 4.2~years for the ECM community, but with overlapping confidence intervals.
This corresponds to a space-for-time substitution rate of 5.4 and 3.3~m/year for the total fungal community and ECM community, respecitvely.
In a recent study, \textcite{kivlin2020} reported a space-for-time substitution rate of 81~m/year (reported as 6.8~d / 1.5~m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
However, comparison is obscured by different spatial and temporal sampling scales between the two studies.
Year-to-year variation in ECM fungal communities, which we sampled, has been shown to be less than intra-annual variation \autocite{bahram2015}, as sampled by \textcite{kivlin2020}.
Neither dataset from the short amplicon library showed significant temporal autocorrelatio.

Weighted UniFrac did not reliably detect spatial structure within this relatively ecologically homogeneous community.
Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1~m; Figure~\ref{fig:correlog}), the distance-decay plot in Figure~\ref{fig:spatio} does not show any clear relationship.
The functional fit showed poor convergence, with a 95\% confidence interval for spatial range of 0--5470~m, indicating little evidence of spatial structure.
This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages \emph{sensu} \textcite{Tedersoo2010}) are more spatially constant.
This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4--0.6) as compared to Bray-Curtis (0.8--1.0) in Figure \ref{fig:spatio}.
UniFrac analysis would be more suited at larger spatial scales and/or larger ecological gradients.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

The choice of amplicon and sequencing technology did not affect the results of the spatial analysis, provided sufficient sequencing depth.
Alpha diversity estimates were strongly correlated between methods, but somewhat inflated for Ion Torrent relative to the other technologies.
However, the addition of long amplicon reads did allow the construction of a phylogenetic tree directly from the metabarcoding reads, which allowed refinement of taxonomic assignments using our new tool PHYLOTAX.
DADA2 ASV yield was initially poor for long amplicons, but this was improved by developing a workflow for extraction of subregions, separate denoising, and then reconstruction of full-length unique sequences.
Together these approaches provide a hybrid approach using long-read sequencing to acquire long amplicon sequences for the local species pool in order to improve taxonomic assignments, and cost-effective short-read sequencing to provide high sampling depth and sample number.

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

This project was funded by the Swedish research council FORMAS grant number 2014-01109.

Laboratory work including PCR and library pooling was performed by Dr.~Ylva Strid.

The authors would like to acknowledge support of the National Genomics Infrastructure (NGI) / Uppsala Genome Center and UPPMAX for providing assistance in massive parallel sequencing and computational infrastructure, funded by RFI / VR and and Science for Life Laboratory, Sweden.

Sequencing was also performed by the SNP\&SEQ Technology Platform in Uppsala.
The facility is part of the National Genomics Infrastructure (NGI) Sweden and Science for Life Laboratory.
The SNP\&SEQ Platform is also supported by the Swedish Research Council and the Knut and Alice Wallenberg Foundation.

\singlespacing

\printbibliography

\doublespacing

\hypertarget{data-accessibility}{%
\section*{Data Accessibility}\label{data-accessibility}}
\addcontentsline{toc}{section}{Data Accessibility}

\begin{itemize}
\tightlist
\item
  Trimmed, demultiplexed sequencing reads have been deposited at the European Nucleotide Archive (ERA) under Project accession number PRJEB37385. Accession numbers are given in Supplementary Files 1 and 2.
\item
  Consensus ASV sequences will also be deposited at ENA prior to publication.
\item
  Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication \autocite{furneauxdata2020}.
\item
  R packages \texttt{LSUx}, \texttt{tzara}, \texttt{phylotax}, and \texttt{FUNGuildR} are available on Github at \url{https://github.com/brendanf/LSUx}, \url{https://github.com/brendanf/tzara}, \url{https://github.com/brendanf/phylotax}, and \url{https://github.com/brendanf/FUNGuildR}.
  These packages are currently being prepared for submission to CRAN/Bioconductor.
  If they are not accepted prior to final publication, snapshots will be archived at Dryad.
\item
  FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}. The versions used in this paper will be archived at Dryad prior to publication.
\item
  Bioinformatics pipeline and analysis scripts are available at \url{https://github.com/oueme-fungi/oueme-fungi-transect}.
\end{itemize}

\hypertarget{author-contributions}{%
\section*{Author Contributions}\label{author-contributions}}
\addcontentsline{toc}{section}{Author Contributions}

Sampling was planned and carried out by BF, NSY, and MR.
Bioinformatics and data analysis were performed by BF with input from MB, AR, and MR.
Scripts and R packages were written by BF.
The manuscript was drafted by BF and MR.
All authors contributed to and approved the final version of the manuscript.

\processdelayedfloats

\end{document}

### 2021-02-16 writing/transect_supplement_rev1.tex
add latex files for first revision (to make diffs)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfauthor={Brendan Furneaux1,; Mohammad Bahram2,3,; Anna Rosling4,; Nourou S. Yorou5,; Martin Ryberg1; 1Program in Systematic Biology, Department of Organismal Biology, Uppsala University,; Uppsala, Sweden; 2Department of Ecology, Swedish University of Agricultural Sciences, Uppsala, Sweden; 3Institute of Ecology and Earth Sciences, University of Tartu, Tartu, Estonia; 4Program in Evolutionary Biology, Department of Ecology and Genetics, Uppsala University,; Uppsala, Sweden; 5Research Unit in Tropical Mycology and Plant-Fungi Interactions, LEB,; University of Parakou, Parakou, Benin},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% table and figure caption width
\usepackage[width=0.9\textwidth]{caption}

% table formatting
\usepackage{colortbl}
\usepackage{makecell}

% lookup cross-references from the paper
\usepackage{xr}
\externaldocument{transect_paper}

% Figures and tables are prefixed with S for supplement
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

% put all tables and figures on their own pages
% \usepackage[nomarkers]{endfloat}
\usepackage{placeins}

% full page landscape figures
\usepackage{rotating}
% \DeclareDelayedFloatFlavour{sidewaysfigure}{figure}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[style=apa,]{biblatex}
\addbibresource{all.bib}
\addbibresource{R.bib}

\title{Supplementary Information for:Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities}
\author{Brendan Furneaux\textsuperscript{1}, \and Mohammad Bahram\textsuperscript{2,3}, \and Anna Rosling\textsuperscript{4}, \and Nourou S. Yorou\textsuperscript{5}, \and Martin Ryberg\textsuperscript{1} \and \textsuperscript{1}Program in Systematic Biology, Department of Organismal Biology, Uppsala University, \and Uppsala, Sweden \and \textsuperscript{2}Department of Ecology, Swedish University of Agricultural Sciences, Uppsala, Sweden \and \textsuperscript{3}Institute of Ecology and Earth Sciences, University of Tartu, Tartu, Estonia \and \textsuperscript{4}Program in Evolutionary Biology, Department of Ecology and Genetics, Uppsala University, \and Uppsala, Sweden \and \textsuperscript{5}Research Unit in Tropical Mycology and Plant-Fungi Interactions, LEB, \and University of Parakou, Parakou, Benin}
\date{}

\begin{document}
\maketitle

\textbf{Supplementary file 1}: Mapping between samples, indexed primers, and read accession numbers for short amplicons.

\textbf{Supplementary file 2}: Mapping between samples, indexed primers, and read accession numbers for long amplicons.

\textbf{Supplementary file 3}: ML tree of all long metabarcoding ASVs (ITS1--5.8S--ITS2--LSU) assigned to kingdom Fungi.
Thick branches represent rapid bootstrap values \textgreater90\%.
Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, with Unite, Warcup, and RDP reference databases.
Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
Taxon names in red were reconstructed as polyphyletic.
In some cases, this may be due to a poorly identified sequence ``splitting'' the group (e.g., the single \emph{Monascus} sequence nested in \emph{Aspergillus}; Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).

\hypertarget{supplementary-methods}{%
\section{Supplementary methods}\label{supplementary-methods}}

\hypertarget{Sampling-supplement}{%
\subsection{Sampling}\label{Sampling-supplement}}

Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Caesalpinioideae).
At each site, 25~soil samples were collected along a single 24~m linear transect at intervals of 1~m in May 2015.
One third of the sample locations (3~m spacing) were resampled one year later in June 2016, for a total of 67 samples.
For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with a sterilized knife blade.
Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
Approximately 250~mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0~mL microtube containing 750~mL of field lysis/preservation buffer and lysis beads (Xpedition\textsuperscript{TM} Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
Due to the inavailability of refrigeration, extraction was initiated in the field according to the manufacturers instructions by lysis using a handheld bead-beater (TerraLyser\textsuperscript{TM}; Zymo Research Corporation).

An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard\textsuperscript{TM} Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
Sequencing results for these 50 samples differed significantly (PERMANOVA with 9999 permutations, \(p < 0.0001\), \(R^2 = 0.06\)) from samples preserved using the Xpedition\textsuperscript{TM} lysis buffer (Figures \ref{fig:buffer-pcoa}, \ref{fig:buffer-compare-long}, and \ref{fig:buffer-compare-short}); as such the LifeGuard preserved samples were excluded from our spatial analyses.
However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.

\hypertarget{DNA-supplement}{%
\subsection{DNA extraction, amplification, and sequencing}\label{DNA-supplement}}

After field lysis, DNA was extracted using the Xpedition\textsuperscript{TM} Soil/Fecal Prep kit (see above).
Samples preserved using LifeGuard were first centrifuged at 10~000~g for 1~minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
DNA was quantified fluorometrically using Quant-iT\textsuperscript{TM} PicoGreen\textsuperscript{TM} dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.

For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200~µM dNTP mix, 250~µM indexed gITS7 primer,
150µM ITS4m, 2mM MgCl\textsubscript{2}, 0.1~U \emph{Taq} polymerase (Dream \emph{Taq}, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7~ng purified DNA in Dream \emph{Taq} buffer.
The reaction conditions were 10~min at 95°, followed by 35~cycles of 60~s at 95°, 45~s at 56°, and 50~s at 72°, and finally 3~min at 72°.
Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.

For the long amplicon, both forward and reverse primers included indexes for combinatorical multiplexing (Supplementary File 2).
PCR was performed as for the short amplicons, but with 500~µM of each of the two primers.
Reaction conditions were 10~min at 95°, 30~cycles of 45~s at 95°, 45~s at 59°, and 90~s at 72°, and finally 10~min at 72°.
Each reaction was performed in three technical replicates as for short amplicons.

After pooling of technical replicates, amplicons were purified using SPRI beads \autocite{vesterinen2016} and quantified fluorometrically as above.
An aliquot of 100~ng of DNA from each sample (or the total PCR product if less than 100~ng) was pooled into two libraries each for long and short amplicons.

\hypertarget{denoising-supplement}{%
\subsection{Denoising and clustering}\label{denoising-supplement}}

\hypertarget{long-numbers}{%
\subsubsection{Denoising long-reads: the numbers are against us}\label{long-numbers}}

The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads \autocite{callahan2016}.
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is \((1-\epsilon)^L\), where \(\epsilon\) is the sequencing error rate and \(L\) is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is \(2/(1-\epsilon)^L\).
For the combination of long sequences (median \(L=1509 \text{ bp}\) after trimming) and moderate error rate (mean \(\epsilon= 0.0066\) based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is 43 720.
Given the high diversity relative to sequencing depth in this study (485 ASVs based on PacBio short amplicons, 104 305 trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon (\(L=265 \text{ bp}\), \(\epsilon=0.0022\)) is only 3.6 reads.

In order to bypass this limitation, we implemented a workflow which divides the long rDNA amplicon into homologous domains (new R package \texttt{LSUx}), denoises the domains individually using DADA2, and then recombines the denoised domains to form full-length denoised amplicons (new R package \texttt{tzara}).
The steps of this workflow are detailed below.

\hypertarget{splitting-the-rdna-into-regions-with-covariance-models-lsux}{%
\subsubsection{\texorpdfstring{Splitting the rDNA into regions with covariance models: \texttt{LSUx}}{Splitting the rDNA into regions with covariance models: LSUx}}\label{splitting-the-rdna-into-regions-with-covariance-models-lsux}}

Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure \autocite{eddy1994}.
First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 \autocite{kalvari2018} using \texttt{cmsearch} from Infernal 1.1.2 \autocite{nawrocki2013}, and all bases before the 5.8S were assigned to ITS1.
No attempt was made to remove the approximately 12~bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx \autocite{bengtsson-palme2013}.
A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 \autocite{glockner2017,cole2014} by truncating after the LR5 primer site and using the reference line to annotate the variable regions \emph{sensu} \textcite{michot1984} and \textcite{raue1988}.
A CM was generated from the alignment using \texttt{cmbuild} from Infernal.
The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using \texttt{cmalign} from Infernal.
The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure~\ref{fig:regions},
where LSU1--4 represent the conserved domains of LSU flanking the variable D1--3 domains \autocite{michot1984}.
For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
Code to extract the segments, including annotated seed alignments and CMs, is available in the new R package \texttt{LSUx}, available and documented at \url{https://github.com/brendanf/LSUx}.

\hypertarget{denoising-each-domain-independently-with-dada2}{%
\subsubsection{\texorpdfstring{Denoising each domain independently with \texttt{DADA2}}{Denoising each domain independently with DADA2}}\label{denoising-each-domain-independently-with-dada2}}

Each of the extracted domains was independently filtered for length (Table~\ref{tab:region-limits}) and a maximum of three expected errors, and denoised using DADA2.
To test the performance of DADA2 on intermediate-length amplicons, adjacent domains from each read were concatenated to form full ITS (ITS1 + 5.8S + ITS2), partial LSU (LSU1 + D1 + LSU2 + D2 + LSU3 + D3 + LSU4), and partial 32S (5.8S + ITS2 + LSU) datasets, and these were also denoised.
The DADA2 error model was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing strategy (i.e., long \emph{vs.} short amplicons, different sequencing technologies).
DADA2 was run with complete pooling for PacBio libraries, and pseudo-pooling
for Ion Torrent libraries.
Chimeras within each region were removed using \texttt{removeBimeraDenovoTable} from DADA2.

\hypertarget{generation-of-full-length-denoised-asvs-tzara}{%
\subsubsection{\texorpdfstring{Generation of full-length denoised ASVs: \texttt{tzara}}{Generation of full-length denoised ASVs: tzara}}\label{generation-of-full-length-denoised-asvs-tzara}}

For each raw read, the denoised sequences of the different domains were concatenated to form a set of full-length sequences.
For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
Because ITS2 is the most variable of the amplified domains as indicated by the greatest number of unique ASVs (Figure~\ref{fig:regions}b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
Therefore, concatenated sequences with identical ITS2 were clustered, and each cluster was aligned in R using the \texttt{DECIPHER} package \autocite{wright2015}.
Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the \texttt{odseq} package \autocite{jehl2015}, using the default threshold of 0.025.
The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence for the entire cluster.
Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide \textgreater50\% at a given position) were removed.
The count and sample distribution of reads assigned to each full-length ASV were calculated in order to form a sample × ASV community matrix.
A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
The process of assigning consensus full-length ASVs was carried out using the new \texttt{tzara} package for R, available and documented at \url{https://github.com/brendanf/tzara}.

The two packages are modular; \texttt{LSUx} can be used to extract domains from LSU for any purpose, and \texttt{tzara} can also be used to recombine denoised sequences which were originally separated using other approaches, most notably using the popular and well-established ITSx \autocite{bengtsson-palme2013}.
However, for our dataset, the single LSU region which would be extracted by ITSx was still too long to be effectively denoised by DADA2 (Figure~\ref{fig:regions}b in main text).

\hypertarget{phylogeny-supplement}{%
\subsection{Phylogenetic inference and taxonomy assignment}\label{phylogeny-supplement}}

\hypertarget{uniform-taxonomic-classification-for-unite-warcup-and-rdp}{%
\subsubsection{Uniform taxonomic classification for Unite, Warcup, and RDP}\label{uniform-taxonomic-classification-for-unite-warcup-and-rdp}}

The taxonomic classification system is based on the Unite database version 8 \autocite{nilsson2019a}.
In particular, the classification for fungi was according to \textcite{tedersoo2018}, and for non-fungal eukaryotes was according to the proposed system of \textcite{tedersoo2017c} as described in \autocite{tedersoodata2017}.
Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists \autocite{adl2019}, which utilizes hierarchical nameless ranks.
FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}.

\hypertarget{rooting-the-fungal-tree}{%
\subsubsection{Rooting the fungal tree}\label{rooting-the-fungal-tree}}

The phylogenetic tree was rooted based on sequences which were confidently assigned to kingdom based on the primary taxonomic assignments.
Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
Assignments were considered reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
Two ASVs belonging to \emph{Tulasnella}, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies \autocite{moncalvo2006}, were excluded at this step.
This resulted in 516 confident kingdom assignments, with 41 assigned to Alveolata, 453 assigned to Fungi, 1 assigned to Metazoa, and 21 assigned to Viridiplantae.
Inspection of the tree verified that these four kingdoms were reciprocally monophyletic, and the clade containing Alveolata and Viridiplantae was used as an outgroup to root the tree.
The kingdom Fungi was then identified as the minimal clade containing all ASVs which were confidently identified as kingdom Fungi.
ASVs falling outside this clade, including the excluded \emph{Tulasnella} sequences, were not included in downstream fungal community analysis.

\hypertarget{refining-taxonomic-assignments-with-a-phylogenetic-tree-phylotax}{%
\subsubsection{\texorpdfstring{Refining taxonomic assignments with a phylogenetic tree: \texttt{PHYLOTAX}}{Refining taxonomic assignments with a phylogenetic tree: PHYLOTAX}}\label{refining-taxonomic-assignments-with-a-phylogenetic-tree-phylotax}}

PHYLOTAX refinements are based on (1) taxonomic assignments from other, primary assignment algorithms, and (2) a phylogenetic tree.
Taxa at each rank are assigned to a node and all its descendants if that taxon is consistent with the reference-based taxonomic assignments for each of the descendants (Figure~\ref{fig:phylotax}).
A taxon assignment is considered to be consistent if at least one primary algorithm assigned that taxon at greater than 50\% confidence, or if no primary algorithm successfully classified the sequence at greater than 50\% confidence.
The result of this process is twofold.
First, it gives a taxonomic assignment to ASVs which were previously unassigned if they are nested within a clade with consistent preliminary assignments.
Second, it resolves conflicts between the different reference-based algorithms, but only if one of the alternatives is consistent with the assignments of other ASVs in the same clade.
PHYLOTAX is implemented in the new R package \texttt{phylotax} at \url{https://github.com/brendanf/phylotax}.

\hypertarget{sequencing-strategy-supplement}{%
\subsection{Effect of sequencing strategy on recovered community}\label{sequencing-strategy-supplement}}

ASV and OTU accumulation curves were generated using the \texttt{iNEXT} package \autocite{hsieh2016}.
Additionally, ASV and OTU richness was estimated for each sample by rarefying to 100 sequences per sample and taking the average of 100 repetitions.
ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies.
For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the \texttt{deming} package \autocite{R-deming}, with the intercept constrained to be 0.
For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.

The effect of sequencing technology and amplicon length on the recovered fungal community
composition, as assessed by the Bray-Curtis dissimilarity was determined by PERMANOVA \autocite{anderson2003} using the \texttt{adonis2} function in the \texttt{vegan} package \autocite{R-vegan}.
Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).

In order to detect potential biases at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
Only samples where all three strategies yielded at least 100~fungal reads ( samples), and classes which represented at least 1\% of reads in at least one sample ( classes), were included.
The community matrix was normalized to proportional abundance by dividing the counts for each class in each sample by the total number of reads for that sample, but were not rarefied to even sampling depth \autocite{mcmurdie2014}.
PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs.~short); and sequencing technology (Illumina vs.~PacBio).
Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the \texttt{capscale} function in \texttt{vegan} \autocite{R-vegan}.
Spatiotemporal effects, i.e.~differences between different samples, were partialed out using a \texttt{Condition()} term in the \texttt{capscale} model in order to visualize effects due to sequencing technology and amplicon length.

A similar analysis was also applied to only fungi classified as ECM, clustered at the family level.
Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database \autocite{nguyen2016funguild} via the R package \texttt{FUNGuildR} (\url{https://github.com/brendanf/FUNGuildR}).
All taxa which included ``Ectomycorrhiza'' in the guild assignment at any level of confidence were included.

Overall community composition, and differences in composition between amplicon libraries, were visualized as heat trees using the \texttt{metacoder} package \autocite{R-metacoder}.

\hypertarget{spatiotemporal-supplement}{%
\subsection{Spatiotemporal analysis}\label{spatiotemporal-supplement}}

Empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by Legendre and Legendre (2012) using the nls function in R.
Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
For datasets where the Mantel correlogram indicated spatial correlation between samples taken in separate years, the model was re-fit with an additional term to represent temporal correlation:
\[D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]\]
where \(D\), \(d\), and \(t\) represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are \(C_0\), the community dissimilarity between replicate samples (``nugget''); \(C_0 + C_1\), the community dissimilarity at long distances (``sill''); \(a_d\) the spatial range at which the community dissimilarity has moved 95\% of the way from ``nugget'' to ``sill''; and \(a_t\), the equivalent temporal range.
The 95\% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the \texttt{MASS} package.

\newpage

\hypertarget{supplementary-figures}{%
\section{Supplementary Figures}\label{supplementary-figures}}

\listoffigures

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/buffer-pcoa-1}

}

\caption[Unconstrained Principal Coordinate Analysis (PCoA) ordination of samples preserved with different buffers]{Unconstrained Principal Coordinate Analysis (PCoA) ordination of samples preserved with different buffers, based on Bray-Curtis dissimilarities of relative read abundances.
For each sequencing strategy, only those sampling locations with more than 100 reads each for all three year/preservation buffer combinations were included.}\label{fig:buffer-pcoa}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/buffer-compare-long-1}

}

\caption[Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs.~LifeGuard) and sampling year, for long amplicon libraries]{Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs.~LifeGuard) and sampling year, for long amplicon libraries. Reference heat tree (bottom left) shows labeled taxonomic tree.
Nodes and branches are scaled to represent mean relative read abundance across samples.
Colored comparison trees (top and right) illustrate variation in relative read abundance between samples preserved by different techniques or from different years.
For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with red-orange colors indicating greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicating greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
The two Xpedition datasets (blue-green colors in top right and center right trees) show consistently greater relative abundance of Basidiomycota, especially Russulales and Sebacinales, as well as Glomeromycota, Morteriellomycota, Geoglossomycetes, and Pezizomycetes,
while the LifeGuard samples (red-orange colors in top right and center right trees) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, and also Geminibasidiomycetes.
Differences between the two Xpedition datasets (top center tree) are less pronounced.}\label{fig:buffer-compare-long}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/buffer-compare-short-1}

}

\caption[Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs.~LifeGuard) and sampling year, for short amplicon libraries]{Comparison of read counts by taxonomic groups for different preservation buffers (Xpedition vs.~LifeGuard) and sampling year, for short amplicon libraries. Reference heat tree (bottom left) shows labeled taxonomic tree.
Nodes and branches are scaled to represent mean relative read abundance across samples.
Colored comparison trees (top and right) illustrate variation in relative read abundance between samples preserved by different techniques or from different years.
For each comparison tree, colors indicate the base 10 log relative abundance ratio for each taxon between two datasets, with red-orange colors indicating greater relative abundance in the dataset indicated above the tree (Xpedition 2015, LifeGuard 2016), and blue-green colors indicating greater relative abundance in the dataset indicated to the right of the tree (Xpedition 2016, Xpedition 2015).
The two Xpedition datasets (blue-green colors in top right and center right trees) show consistently greater relative abundance of Basidiomycota, especially Sebacinales, as well as Glomeromycota, Geoglossomycetes, and Pezizomycetes,
while the LifeGuard samples (red-orange colors in top right and center right trees) showed greater relative abundance of Ascomycota in general, especially Aspergillaceae, as well as Geminibasidiomycetes.
Differences between the two Xpedition datasets (top center tree) are less pronounced.}\label{fig:buffer-compare-short}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/phylotax-1}

}

\caption[Phylogenetic refinement of taxonomic assignments (PHYLOTAX)]{Phylogenetic refinement of taxonomic assignments (PHYLOTAX). In the example, a clade includes seven OTUs (A-F), which have been identified by two methods as belonging to taxa ``Tax1'' or ``Tax2'', or are unidentified (``unk''), as shown in the tip labels.
No taxon is assigned at node~1, because one of the child branches (A) is completely unassigned.
No taxon is assigned at node~2, because the assignments at C and F are inconsistent.
Node~3 is assigned to Tax2 because this is consistent with at least one of the assignments for both B and C.
Node~4 (and thus also node~5) is assigned to Tax1 because this is consistent with at least one of the assignments for D and F, and because E, which is nested within the clade defined by D and F, is completely unassigned.}\label{fig:phylotax}
\end{figure}

\begin{sidewaysfigure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/sample-depth-1}

}

\caption[DNA concentrations after extraction and PCR, and sequencing depth along transects at the two sites Ang and Gan for the years 2015 and 2016]{DNA concentrations after extraction (leftmost column) and PCR (second and third columns), and sequencing depth (right three columns) along transects at the two sites Ang and Gan for the years 2015 and 2016.
Sequencing reads which were filtered out during quality control are colored dark grey.
Dotted horizontal line at 100~reads indicates cutoff for inclusion in community analysis.
2016 samples were preserved using two different methods (LifeGuard, Xpedition), while 2015 samples were preserved only using Xpedition.}\label{fig:sample-depth}
\end{sidewaysfigure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/qual-check-1}

}

\caption[Read quality profiles]{Read quality profiles, based on quality scores assigned by technology-specific basecallers. \textbf{(a)} per-base error rate for different sequencing strategies; \textbf{(b)} expected number of errors per read for different sequencing strategies; \textbf{(c)} expected number of errors per read for regions extracted from PacBio long amplicon reads. In \textbf{(a)} and \textbf{(b)}, R1 and R2 denote the forward and reverse reads, respectively, from Illumina paired-end sequencing. PacBio short, Ion Torrent, and Illumina datasets were all sequenced from the same short amplicon library. In \textbf{(b)} and \textbf{(c)}, quality filtering threshold of 3 expected errors marked as a dashed vertical line.}\label{fig:qual-check}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/full-length-1}

}

\caption[Comparison of the length of denoised sequences from different length amplicons and sequencing technologies]{Comparison of the length of denoised sequences from different length amplicons and sequencing technologies. Length distribution (\emph{a}, \emph{c}) and empirical cumulative distribution function (ECDF; \emph{b}, \emph{d}) for short (\emph{a}, \emph{b}) and long (\emph{c}, \emph{d}) amplicons, respectively.}\label{fig:full-length}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/region-lengths-1}

}

\caption[Length distribution for each region extracted from long amplicons]{Length distribution for each region extracted from long amplicons. Filtering limits are shown as dashed vertical lines.}\label{fig:region-lengths}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/ITS2-length-1}

}

\caption[Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies]{Comparison of the length of denoised ITS2 sequences extracted from different length amplicons and sequencing technologies. \emph{a}) Length distribution. \emph{b}) Empirical cumulative distribution function.}\label{fig:ITS2-length}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/read-compare-1}

}

\caption[Comparison between read numbers for different sequencing strategies]{Comparison between read numbers for different sequencing strategies, by ASV (\textbf{a}) and 97\% OTU (\textbf{b}). ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes. Dashed line represents a constant ratio of read numbers. The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading. \(R^2\) value displayed is for log-transformed non-zero read numbers.}\label{fig:read-compare}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/accum-OTU-1}

}

\caption[OTU accumulation curves]{OTU accumulation curves. Each curve represents rarefaction of a single sample or pooled pair of samples. Points at the end of each curve represent the actual read depth and observed OTU richness. Note logarithmic axes.}\label{fig:accum-OTU}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/alpha-compare-otu-1}

}

\caption[Comparison of OTU richness between sequencing technologies]{Comparison of OTU richness between sequencing technologies (ref:alpha-compare)}\label{fig:alpha-compare-otu}
\end{figure}

\begin{sidewaysfigure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/taxa-1}

}

\caption[Heat tree summarizing taxonomic composition of soil community]{Heat tree summarizing taxonomic composition of soil community. Color and size of nodes represent the fractional ASV richness of that taxonomic group.
Color and thickness of branches represent the fractional read abundance of all ASVs belonging to that taxonomic group.
Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
Groups which are not represented by at least 1\% of reads or 1\% of ASVs in any dataset are collapsed into nodes labeled "*``.
ASVs which could not be further assigned are labelled''?".
ASV richness and read abundance are displayed as the mean across sequencing runs.
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:taxa}
\end{sidewaysfigure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/heattree-length-compare-1}

}

\caption[Variation in read abundance and diversity for different amplicons across the taxonomic tree]{Variation in read abundance and diversity for different amplicons across the taxonomic tree. Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
In both cases red-orange represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
Values are log\textsubscript{10} of the ratio of mean values across sequencing runs for each amplicon.
Groups which are not represented by at least 1\% of reads or 1\% of ASVs in any dataset are collapsed into nodes labeled "*``.
ASVs which could not be further assigned are labelled''?".
Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio \textgreater{} 3.2) are labeled.
Refer to Figure~\ref{fig:taxa} for complete taxon labeling.}\label{fig:heattree-length-compare}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/short-its-1}

}

\caption[Kingdom-level taxonomic composition of ITS2 ASVs \(\le 140\) bp]{Kingdom-level taxonomic composition of ITS2 ASVs \(\le 140\) bp. Vertical axis shows fraction of total reads \(\le 140\) bp in each dataset.}\label{fig:short-its}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/ecm-1}

}

\caption[Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments]{Assignment of ECM status to fungal ASVs using FUNGuild database and taxonomic assignments. Reads which were not identified to kingdom, or which were identified as a non-Fungi kingdom, are not included.
``unidentified'' denotes ASVs which were assigned to kingdom Fungi, but could not be identified to family level or could not be assigned ECM status based on the available identification.
Height of each bar represents the fraction of reads represented by that bar.
Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; ``All'' denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.}\label{fig:ecm}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/ecm-heattree-1}

}

\caption[Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only]{Heat tree showing variation in read abundance and diversity for different amplicons, for ECM taxa only. Red-orange edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
Values are log\textsubscript{10} of the ratio of mean values across all sequencing technologies for each amplicon.
Groups which are not represented by at least 1\% of reads or 1\% of ASVs in any dataset are collapsed into nodes labeled "*``.
ASVs which could not be further assigned are labelled''?".
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:ecm-heattree}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/ppcoa-1}

}

\caption[Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding strategies]{Partial Principal Coordinates Analysis (PPCoA) of fungal community using different metabarcoding strategies, for all fungi (\textbf{a}) and ECM fungi (\textbf{b}).
Ordination based on the residual variation after spatio-temporal variation (i.e., variation between different soil samples) has been accounted for.
Only samples with at least 100 reads for all three strategies are included.
ASVs are clustered by taxonomic class for all fungi, and by family for ECM fungi.
Abbreviations represent the approximate direction of increase for the classes and families with the highest scores: \textbf{So}: Sordariomycetes, \textbf{Ag}: Agaricomycetes, \textbf{Pe}: Pezizomycetes, \textbf{Eu}: Eurotiomycetes, and \textbf{Do}: Dothideomycetes; \textbf{In}: Inocybaceae, \textbf{Am}: Amanitaceae, \textbf{Sc}: Sclerodermataceae, \textbf{Ru}: Russulaceae, \textbf{Pe}: Pezizaceae, and \textbf{Cl}: Clavulinaceae.}\label{fig:ppcoa}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{transect_supplement_files/figure-latex/correlog-1}

}

\caption[Mantel correlograms for community dissimilarities and spatio-temporal distance]{Mantel correlograms for community dissimilarities and spatio-temporal distance. Community dissimilarities are Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right). Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.}\label{fig:correlog}
\end{figure}

\FloatBarrier

\hypertarget{supplementary-tables}{%
\section{Supplementary tables}\label{supplementary-tables}}

\listoftables

\newpage

\FloatBarrier

\begin{table}

\caption{\label{tab:region-limits}Minimum and maximum allowed lengths for each extracted region.}
\centering
\begin{tabular}[t]{lrr}
\toprule
Region & Min. length & Max. length\\
\midrule
ITS1 & 50 & 500\\
5.8S & 50 & 200\\
ITS2 & 50 & 500\\
LSU1 & 25 & 200\\
D1 & 20 & 500\\
\addlinespace
LSU2 & 50 & 500\\
D2 & 20 & 2999\\
LSU3 & 30 & 60\\
D3 & 50 & 500\\
LSU4 & 50 & 500\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption[Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages]{\label{tab:bioinfo}Number of Amplicon Sequence Variants (ASV) and reads at different pipeline stages. \textbf{Raw}: raw reads as delivered by sequencing center (CCS reads for PacBio);
\textbf{Trim}: reads with primers and demultiplexing barcodes removed;
\textbf{Filter (full)}: full length amplicons with a maximum of three expected errors;
\textbf{LSUx}: reads with a positive CM hit for 5.8S, allowing regions to be extracted;
\textbf{Filter (ITS2)}: extracted ITS2 regions with a maximum of three expected errors;
\textbf{ITS2}: ASVs obtained for the ITS2 region, and reads successfully mapped to an ITS2 ASV;
\textbf{short}, \textbf{ITS}, \textbf{LSU}, \textbf{long}: ITS2-based ASVs mapped to consensus ASVs for longer regions, where "short" and "long" denote the full-length short and long amplicons, respectively.
\emph{Italicized} numbers represent mapping via ITS2 ASVs to amplicons originally sequenced by another technology.}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{PacBio RS II} & \multicolumn{2}{c}{Ion Torrent Ion S5} & \multicolumn{2}{c}{Illumina MiSeq} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Long} & \multicolumn{2}{c}{Short} & \multicolumn{2}{c}{Short} & \multicolumn{2}{c}{Short} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
  & ASVs & reads & ASVs & reads & ASVs & reads & ASVs & reads\\
\midrule
Raw & – & 125,034 & – & 49,511 & – & 20,717,742 & – & 10,756,939\\
Trim & – & 104,305 & – & 41,096 & – & 15,208,677 & – & 9,513,433\\
Filter (full) & – & 44,190 & – & 38,742 & – & 12,177,705 & – & 7,674,712\\
\addlinespace
LSUx & – & 100,235 & – & 40,137 & – & 14,946,121 & – & –\\
Filter (ITS2) & – & 87,860 & – & 38,756 & – & 13,474,693 & – & –\\
ITS2 & 978 & 82,460 & 564 & 37,218 & 12,964 & 13,416,070 & 4,478 & 6,763,654\\
\addlinespace
short & \em{598} & \em{65,191} & 562 & 37,214 & 12,246 & 13,413,990 & 4,419 & 6,753,567\\
ITS & 806 & 81,234 & \em{268} & \em{30,582} & \em{470} & \em{7,938,390} & \em{491} & \em{4,848,591}\\
LSU & 779 & 81,096 & \em{268} & \em{30,741} & \em{479} & \em{8,074,310} & \em{498} & \em{4,930,083}\\
long & 708 & 80,290 & \em{257} & \em{30,481} & \em{434} & \em{7,895,829} & \em{457} & \em{4,817,606}\\
\bottomrule
\end{tabular}
\end{table}

\setlength{\tabcolsep}{1ex}

\begin{table}

\caption[Correspondences between ASVs found by different sequencing strategies]{\label{tab:venn-table-asv}Correspondences between ASVs found by different sequencing strategies. Each row shows the number of ASVs (\emph{ASVs}) shared uniquely by one or more datasets, including the fraction of total ASVs for each dataset (\emph{ASVs frac}), the number of reads represented by those ASVs in each dataset (\emph{reads}) and the fraction of total reads for the dataset (\emph{reads frac}).}
\centering
\begin{tabular}[t]{rcrrcrrcrrcrrr}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{PacBio RS II} & \multicolumn{3}{c}{Illumina MiSeq} & \multicolumn{3}{c}{Ion Torrent Ion S5} \\
\cmidrule(l{3pt}r{3pt}){3-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Long} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\makecell[c]{\\} & ASVs & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{ASVs\\frac} & reads & \makecell[c]{reads\\frac}\\
\midrule
 & 378 & 0.39 & 17k & 0.21 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 14 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.02 & 232 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 868 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.19 & 165k & 0.02 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 9378 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.72 & 1.8M & 0.14\\
\addlinespace
 & 28 & 0.03 & 682 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.01 & 12k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 40 & 0.04 & 879 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 3.2k & 0.00\\
 & 26 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.05 & 420 & 0.01 & 0.01 & 29k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 10 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.02 & 148 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 2.4k & 0.00\\
 & 2791 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.62 & 896k & 0.13 & 0.22 & 2.1M & 0.16\\
\addlinespace
 & 20 & 0.02 & 1.1k & 0.01 & 0.04 & 1.3k & 0.03 & 0.00 & 149k & 0.02 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 251 & 0.26 & 9k & 0.11 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.06 & 316k & 0.05 & 0.02 & 611k & 0.05\\
 & 233 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.41 & 5.5k & 0.15 & 0.05 & 703k & 0.10 & 0.02 & 1.4M & 0.10\\
\addlinespace
 & 261 & 0.27 & 54k & 0.65 & 0.46 & 30k & 0.80 & 0.06 & 4.5M & 0.66 & 0.02 & 7.5M & 0.56\\
\midrule
Total & 14298 & 1.00 & 82k & 1.00 & 1.00 & 37k & 1.00 & 1.00 & 6.8M & 1.00 & 1.00 & 13M & 1.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption[Correspondences between 97\% OTUs found by different sequencing strategies]{\label{tab:venn-table-otu}Correspondences between 97\% OTUs found by different sequencing strategies. Each row shows the number of OTUs (\emph{OTUs}) shared uniquely by one or more datasets, including the fraction of total OTUs for each dataset (\emph{OTUs frac}), the number of reads represented by those OTUs in each dataset (\emph{reads}) and the fraction of total reads for the dataset (\emph{reads frac}).}
\centering
\begin{tabular}[t]{rcrrcrrcrrcrrr}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{PacBio RS II} & \multicolumn{3}{c}{Illumina MiSeq} & \multicolumn{3}{c}{Ion Torrent Ion S5} \\
\cmidrule(l{3pt}r{3pt}){3-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Long} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} & \multicolumn{3}{c}{Short} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11} \cmidrule(l{3pt}r{3pt}){12-14}
\makecell[c]{\\} & OTUs & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac} & \makecell[c]{OTUs\\frac} & reads & \makecell[c]{reads\\frac}\\
\midrule
 & 326 & 0.35 & 16k & 0.20 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 1 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 2 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 133 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.05 & 8.1k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 4054 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.60 & 192k & 0.01\\
\addlinespace
 & 4 & 0.00 & 56 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 935 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 45 & 0.05 & 950 & 0.01 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.01 & 2.6k & 0.00\\
 & 2 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 18 & 0.00 & 0.00 & 3.1k & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  }\\
 & 4 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.01 & 18 & 0.00 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.00 & 712 & 0.00\\
 & 1866 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.67 & 332k & 0.05 & 0.28 & 919k & 0.07\\
\addlinespace
 & 242 & 0.26 & 5.1k & 0.06 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.09 & 124k & 0.02 & 0.04 & 350k & 0.03\\
 & 221 & \cellcolor{gray}{  } & \cellcolor{gray}{  } & \cellcolor{gray}{  } & 0.42 & 4.2k & 0.11 & 0.08 & 629k & 0.09 & 0.03 & 1.3M & 0.10\\
\addlinespace
 & 303 & 0.33 & 60k & 0.73 & 0.57 & 33k & 0.89 & 0.11 & 5.7M & 0.84 & 0.04 & 11M & 0.79\\
\midrule
Total & 7201 & 1.00 & 82k & 1.00 & 1.00 & 37k & 1.00 & 1.00 & 6.8M & 1.00 & 1.00 & 13M & 1.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:variofit}Parameters for exponential distance-decay fits. Range: range at which exponential function is at 95\% of its asymptotic value.}
\centering
\begin{tabular}[t]{lllllll}
\toprule
Guild & Amplicon & Tech & Algorithm & Metric & Space Range (m) & Time Range (a)\\
\midrule
All Fungi & Short & Illumina & Cons & Bray-Curtis & 13 (7.4– 23) & --\\
All Fungi & Short & Illumina & PHYLO & Bray-Curtis & 13 (7.4– 23) & --\\
All Fungi & Short & PacBio & Cons & Bray-Curtis & 31 ( 15– 57) & --\\
All Fungi & Short & PacBio & PHYLO & Bray-Curtis & 31 ( 15– 57) & --\\
All Fungi & Long & PacBio & PHYLO & Bray-Curtis & 18 (  9– 41) & 3.3 (1.9–6.9)\\
All Fungi & Long & PacBio & PHYLO & W-UniFrac & 466 (0–5471) & --\\
\addlinespace
ECM & Short & Illumina & Cons & Bray-Curtis & 12 (6.4– 26) & --\\
ECM & Short & Illumina & PHYLO & Bray-Curtis & 13 (6.7– 26) & --\\
ECM & Short & PacBio & Cons & Bray-Curtis & 42 ( 16– 60) & --\\
ECM & Short & PacBio & PHYLO & Bray-Curtis & 42 ( 16– 61) & --\\
ECM & Long & PacBio & PHYLO & Bray-Curtis & 14 (5.2– 30) & 4.2 (  2– 19)\\
ECM & Long & PacBio & PHYLO & W-UniFrac & -- & --\\
\bottomrule
\end{tabular}
\end{table}

\FloatBarrier

\end{document}

### 2021-02-16 writing/transect_paper.Rmd
more revisions

@@
 Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
-Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
-<!--R4: Near L114 - More refs to consider: See Ashelford et al., 2005 Appl Env Microbiol 71: 7724,
-Bidartondo et al., 2008 Science 319:1616, but also LeRay et al., 2019 PNAS 116: 22651 for a
-more recent take-->
-Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent.
+Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@bidartondo2008;@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@ashelford2006;@nilsson2012].
+Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent [@nilsson2005]. 
 Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
-<!--R4: see Nilsson et al., 2005 BMC Bioinformtics 6: 178-->
-The Unite database is a more curated <!--R5: More curated?-->attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
+The Unite database is an attempt to include all publically available high-quality ITS sequences (800\ 000 as of release 8.0), originally limited to fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
 The Ribosomal Data Project [RDP, @cole2014] hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
@@
 However, all sequence-similarity based approaches are dependent on high taxonomic coverage in the reference database, making the placement of novel or undersampled groups problematic [@nilsson2016top;@Tedersoo2018].
-<!--R4: consider mentioning Nilsson et al., 2019 Nuc Acids Res 47: D259 on the naming of dark
-taxa in UNITE see: taxon hypotheses (THs)-->
 
@@
 Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
-Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
-<!--R5: Add citation for “turnover scale”-->
+Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool [@legendre2012].
 Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
@@
 For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade and homogenized by manually breaking up clumps and mixing inside a plastic zipper bag.
-<!--R4: how was the knife sterilized, how were the samples homogenized?
-Preserved at what
-temperature? Please address in main text.-->
 A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).
@@
 See Supplementary Methods \@ref(DNA-supplement) for PCR and library preparation protocols.
-<!--R4: I always recommend that full methods be in the main text, not in the supplement, it only
-takes a few sentences to describe PCR protocols at least.-->
 
@@
 We used this method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), for all of the PacBio and Ion Torrent datasets.
+Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
+The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.
+To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
 
@@
 
-<!--R4: this looks like maybe a new paragraph? It should be connected to last paragraph.
-Regular indentations at the start of each paragraph would make this more clear.-->
-Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
-The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.
-
-<!--R4: this looks like maybe a new paragraph? It should be connected to last paragraph.-->
-To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].
-
 ### Phylogenetic inference and taxonomic assignment
@@
 and IDTAXA [@murali2018].
-A relatively lax confidence threshhold of 50% was used for all three algorithms.
-<!--R4: a relatively lax bootstrap support cutoff of 50% would indicate classification accuracy of
-what? 75% accuracy, more, less, does it vary across ITS and LSU? I suggest this be addressed in
-the main text, in https://rdp.cme.msu.edu/download/posters/fungalITSreport_062014.pdf see
-Fig 3. -->
-<!--R5: This seems very low.
-I think that some explanation for such a low value is needed to give a rationale.
-This is basically a coin toss so that is not exactly giving great confidence in the results-->
+A relatively lax confidence threshold of 50% was used for all three algorithms, in order to increase the amount of input for  consensus algorithms.
 Each full-length ASV was thus given up to nine primary taxonomic assignments (three references $\times$ three algorithms).
@@
 For full-length long-amplicon ASVs, the primary taxonomic assignments were refined with the new algorithm PHYLOTAX.
-The PHYLOTAX algorithm resolves conflicts and makes additional assignments based on an input tree (here the ML tree generated above) and taxonomic identifications from one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
+The PHYLOTAX algorithm resolves conflictsm one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
 It is available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
 
-ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the preliminary assignments [@huson2007].
+ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different primary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the primary assignments [@huson2007].
 This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [e.g., MEGAN and LCAClassifier, @huson2016;@lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
@@
 To estimate turnover scale, ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric [@bray1957, for both long and short amplicons] and the phylogenetically based weighted UniFrac metric [@lozupone2005;@lozupone2007, for only long amplicons] in `phyloseq` version `r packageVersion("phyloseq")`.
-Each of these distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
+Dissimilarities were based on relative read abundance within each sample.
+Samples were not rarefied to a standard sequencing depth within datasets, as both the Bray-Curtis dissimilarity and the UniFrac metric are robust to unequal sampling depths [@mcmurdie2014].
+In addition, we did not standardize sequencing depth between datasets, because this would remove one of the potential benefits of Illumina and Ion Torrent relative to PacBio.
+
+Each of the distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
 Separate correlograms were drawn for samples taken during the same year and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
-
 Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R (Supplementary Methods \@ref(spatiotemporal-supplement)).
+Spatiotemporal analyses were performed on the full recovered fungal community after removal of non-fungal sequences, and on the ECM community.
 
@@
 Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads (Figure \@ref(fig:sample-depth)).
-Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
-<!--R5: So make a statement about how many samples are left in the analyses after these were excluded? Seems like it may be a very low number of samples here-->
+Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, taxonomic assignment, and all non-spatial analyses.
+Spatial analyses were based on the remaining 34 samples for Illumina and 30 samples each for the PacBio Long and Short amplicons.
 
@@
 Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
-<!--R5: But how does that impact quality? Seems like this would reduce quality significantly -->
 Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of denoised ITS2 reads from the long-amplicon PacBio dataset.
@@
 For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
-<!--R4: it would be helpful if you drew a line down the middle of this figure then labelled the
-left side ASVs and the right side OTUs. Instead of ‘abundance’ suggest you change to ‘read
-number or read abundance’. Is this figure based on rarefied, normalized data? Indicate in the
-Text.-->
-<!--R5: again, were these analyses conducted on the entire (i.e. fungal and non-fungal) dataset, or just the fungal community?-->
+Analyses performed on pooled ASVs/OTUs from all samples.
 
 ```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", fig.height = 4, fig.align="center", results="asis"}
-ggpubr::ggarrange(readd(vennplot_tech_ASV), readd(vennplot_tech_OTU), ncol = 2)
+line = ggplot() +
+  geom_vline(xintercept = 0) +
+  theme_pubclean() +
+  theme(axis.title = element_text(color = "white"),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        plot.background = element_blank(),
+        panel.background = element_blank()
+        )
+ggpubr::ggarrange(
+  readd(vennplot_tech_ASV),
+  line,
+  readd(vennplot_tech_OTU),
+  ncol = 3,
+  widths = c(1,0.02,1),
+  labels = c("ASVs", "", "OTUs"),
+  label.x = 0.5,
+  hjust = 0.5,
+  font.label = list(size = 16)
+)
 grid::grid.legend(
@@
 
-ggpubr::ggarrange(readd(vennplot_amplicon_ASV), readd(vennplot_amplicon_OTU), ncol = 2)
+ggpubr::ggarrange(
+  readd(vennplot_amplicon_ASV),
+  line,
+  readd(vennplot_amplicon_OTU),
+  ncol = 3,
+  widths = c(1,0.02,1)
+)
 grid::grid.legend(
@@
 In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).
-<!--R5: Not clear what that means?-->
+The taxonomic identity of these ASVs is discussed below.
 
@@
 ASV richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(readd(demingfits_ASV), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% unique() %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
-<!--R4: here and onwards (L409, L418), I see that you have plotted rarefaction curves, but it’s
-not clear to me whether you compared samples at equal sampling depth, ex. smallest library
-size. If you did this, please add a line to say so in the main text. If you didn’t do this, how do
-you know that your community comparisons aren’t simply the effect of library size? Please
-address in main text.-->
 The slope of the relationship between PacBio and Illumina richness estimates was only slightly different from 1, indicating that these two technologies give highly comparable rarefied richness estimates, despite the approximately 200× difference in original sequencing depth.
@@
 Total least squares regression indicated that the long amplicon dataset resulted in richness estimates which were intermediate between the short amplicon results from Ion Torrent and the other two technologies.
-Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure\ \@ref(fig:venn)), sample-wise OTU accumulation curves
-<!--R5: Not clear what that means?-->
+Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure\ \@ref(fig:venn)), OTU accumulation curves for each sample
 (Figure\ \@ref(fig:accum-OTU)) and rarefied OTU richness relationships between sequencing strategies (Figure\ \@ref(fig:alpha-compare-otu)) were highly similar to those for ASVs.
@@
 Vertical dashed line at 100 reads indicates rarefaction level for ASV richness comparisons.
-<!--R5: again, were these analyses conducted on the entire (i.e. fungal and non-fungal) dataset, or just the fungal community?-->
 
@@
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.
-<!--R4: I’d
-be interested in seeing a comparison of run times for each taxonomic assignment or a sentence
-added to the main text comparing run times. In the discussion you may want to indicate that
-phylogeny-based methods tend to be more computationally intensive compared to BLAST or
-RDP ex. shown for fungal ITS (Porter and Golding, 2011 New Phyt 192: 775 ) and fungal LSU
-(Porter and Golding, 2012 PLoS ONE 7: e35749).-->
 
@@
 Assignments based on PHYLOTAX.
-Classes which represented less than 2% of reads and ASVs in all datasets are grouped together as "other".
-<!--R4: Is this comparison based on rarefied, normalized data? How do you know these
-patterns don’t reflect different library sizes? Please address in the main text. Change ‘which’ to
-‘That’.-->
+Classes that represented less than 2% of reads and ASVs in all datasets are grouped together as "other".
 
@@
 Many of the ASVs which were unique to the long-amplicon library thus fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)).
-<!--R5: This needs to be mentioned above.
-This is the first mention of non-fungal reads.
-Need to address this in the Methods and also above near note on the ASVs in the long read vs.
-short read datasets-->
 In particular a large fraction of ITS2 sequences with length less than 140 (Figure\ \@ref(fig:ITS2-length)) were identified as Alveolates (Figure\ \@ref(fig:short-its)).
@@
 Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
-<!--R5: Here is not clear if you screened out the non-fungi from the various datasets before you created this part.
-I think you need to specifically address this in methods so the reader knows whether you are making comparisons between the full community or with the fungal communithy-->
 The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm), \@ref(fig:heattree-length-compare), and \@ref(fig:ppcoa)).
@@
 
-Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at $p < 0.05$ for ranges of up to 2--3\ m for the total fungal community, and 1--2\ m for the ECM fungal community (Figure\ \@ref(fig:correlog)).
+Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at $p < 0.05$ for ranges of up to 2--3\ m for the total recovered fungal community, and 1--2\ m for the ECM fungal community (Figure\ \@ref(fig:correlog)).
 In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
-<!--R5: But if non-fungi were included then this may drive some of the pattern detection so needs to be clearer about whether non fungi are discarded for this or not?-->
 The least spatial structure was detected with PacBio short amplicon sequencing.
 
-The Bray-Curtis metric showed positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0\ m, time lag of 1\ year), for both the total fungal and ECM fungal communities, although this result did not reach statistical significance for all sequencing strategies.
+The Bray-Curtis metric showed positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0\ m, time lag of 1\ year), for both the total recovered fungal and ECM fungal communities, although this result did not reach statistical significance for all sequencing strategies.
 This spatiotemporal correlation did not extend to a range of 1\ m, and in fact correlation was negative at a time lag of 1\ year and distance of 1\ m, indicating that samples collected 1\ m apart in different years were more different than randomly selected pairs of samples.
@@
 
-In contrast to the Bray-Curtis distance, the weighted UniFrac distance showed very little spatial structure, with only the total fungal community in the 1\ m
+In contrast to the Bray-Curtis distance, the weighted UniFrac distance showed very little spatial structure, with only the total recovered fungal community in the 1\ m
 distance class showing a significant correlation at $p < 0.05$.
@@
 
-The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "All Fungi") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the total fungal community and `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "ECM") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
-However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Table\ \@ref(tab:variofit)), so no strong conclusion of variability between methods can be drawn.
-Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval (Table\ \@ref(tab:variofit)).
+The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "All Fungi") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the total recovered fungal community and `r filter(readd(variofit_table), metric == "Bray-Curtis", timelag == 0, guild == "ECM") %$% glue_collapse(format(range(estimate), format = "f", digits = 0), "--")`\ m for the ECM fungal community (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
+However, there was overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranges in both the total recovered fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Table\ \@ref(tab:variofit)), so no strong conclusion of variability between methods can be drawn.
+Although a distance-decay model was fit for the weighted UniFrac distance applied to the total recovered fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval (Table\ \@ref(tab:variofit)).
 
@@
 
-<!-- 592 I’d start the discussion with a couple of introductory sentences pointing out the main finding and guiding the reader to the next sections.-->
-
 ## Reconstruction of long amplicons from denoised subregions
@@
 and even between primer pairs (R^2^ = `r readd(comparisons) %>% filter(type == "OTU") %>% filter(x_amplicon != y_amplicon) %$% paste(format(range(r.squared), format = "f", digits = 2), collapse = "--")`, Figure \@ref(fig:read-compare)b).
-This lends some support for the technical repeatability of abundance-based beta diversity measures in metabarcoding, although bias at the amplification stage still presents issues [@castano2020 and references therein].
+This lends some support for the technical repeatability of abundance-based beta diversity measures in metabarcoding, although bias at the amplification stage still presents issues [@polz1998;@Kanagawa_2003;@bellemain2010;@castano2020].
 
@@
 The long amplicon dataset included `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` unique taxa, even after clustering at 97% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers amplify different parts of the community.
-<!--R5: If these 20% are mostly non-fungal reads then this needs to be a bigger part of the discussion!
-Furthermore, some of these primers have known biases against many non-Dikarya fungi.
-In which case, "total fungal community" in any resulting analyses has the very large caveat "biased towards Dikarya" attached to it.
--->
 The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae [@tedersoo2015].
@@
 where `r filter(readd(taxon_chart), tech == "PacBio", amplicon == "Long", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(readd(taxon_chart), tech == "PacBio", amplicon == "Long", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
-This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure\ \@ref(fig:ecm)).
-<!--R5: By how much?
-What proportion?
-This is the really important thing here because readers want to know how much extra information they would get for all this work and money to get the PacBio long read data!-->
+This led to a corresponding increase in the fraction of fungal reads assigned to a functional guild from `r filter(readd(taxon_reads), ifelse(Algorithm == "Consensus", region == "hybrid", TRUE), amplicon == "Long", kingdom == "Fungi", Algorithm %in% c("Consensus", "PHYLOTAX")) %>% group_by(Algorithm) %>% summarize(1 - sum(is.na(as.character(ECM)) * reads)/sum(reads)) %>% tibble::deframe() %>% magrittr::extract(c("Consensus", "PHYLOTAX")) %>% percent(format = "f", digits = 0) %>% paste(collapse = " to ")` (Figure\ \@ref(fig:ecm)).
 For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs (last row of Table\ \@ref(tab:bioinfo)).
@@
 Distance-decay plots (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total recovered fungal community and the ECM community, with best fit estimates ranging from `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(estimate) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
 The 95% confidence interval was substantially wider than this variation, generally covering a range of `r readd(variofit_table) %>% filter((amplicon == "Long") | (tech == "Illumina"), metric == "Bray-Curtis", term == "range") %$% range(c(conf.low, conf.high), na.rm = TRUE) %>% format(digits = 0) %>% paste(collapse = "--")`\ m.
@@
 
-Year-to-year correlation was found for both the total fungal and ECM communities in the long amplicon dataset (Figure \@ref(fig:correlog)).
-The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
-This corresponds to a space-for-time substitution rate (i.e., ratio of a spatial distance to a time delay which results in equivalent community dissimilarity) of `r readd(variofit_table) %>% select(-variofit, -(std.error:conf.high)) %>% pivot_wider(names_from = term, values_from = estimate) %>% filter(timelag == 1, mantel == TRUE) %>% mutate(spacetime = range / timerange) %>% arrange(guild) %$% glue_collapse(format(spacetime, format = "f", digits = 2), " and ")`\ m/year for the total fungal community and ECM community, respecitvely.
+Year-to-year correlation was found for both the total recovered fungal and ECM communities in the long amplicon dataset (Figure \@ref(fig:correlog)).
+The spatiotemporal distance-decay fit estimated the temporal turnover range as `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "All Fungi") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the total recovered fungal community and `r readd(variofit_table) %>% filter(amplicon == "Long", tech == "PacBio", metric == "Bray-Curtis", term == "timerange", guild == "ECM") %$% range(estimate, na.rm = TRUE) %>% unique() %>% formatC(format = "f", digits = 1) %>% paste(collapse = "--")`\ years for the ECM community, but with overlapping confidence intervals.
+This corresponds to a space-for-time substitution rate (i.e., ratio of a spatial distance to a time delay which results in equivalent community dissimilarity) of `r readd(variofit_table) %>% select(-variofit, -(std.error:conf.high)) %>% pivot_wider(names_from = term, values_from = estimate) %>% filter(timelag == 1, mantel == TRUE) %>% mutate(spacetime = range / timerange) %>% arrange(guild) %$% glue_collapse(format(spacetime, format = "f", digits = 2), " and ")`\ m/year for the total recovered fungal community and ECM community, respecitvely.
 In a recent study, @kivlin2020 reported a space-for-time substitution rate of `r format(1.5/6.8 * 365, digits = 2)`\ m/year (reported as 6.8\ d / 1.5\ m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
-<!--R4: I am not familiar with the term ‘space-for-time substitution rate’. Please explain this
-concept in the main text.-->
-
 However, comparison is obscured by different spatial and temporal sampling scales between the two studies.

### 2021-02-16 writing/transect_supplement.Rmd
more revisions

@@
 
-```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
+```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_plot_ecm"), more.cache = datasets}
 readd(taxon_plot_ecm)

### 2021-02-18 writing/transect_paper.Rmd
Additional changes from Martin

@@
 For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade and homogenized by manually breaking up clumps and mixing inside a plastic zipper bag.
-A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).
+A subsample of approximately 250\ mg was stored in DNA-preservation buffer and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).
 
@@
 
-We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods \@ref(long-numbers)).
-We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
+We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity, which lead to most biological variants being represented only by a cluster of reads differing by a small number of unique sequencing errors (for calculations see Supplementary Methods \@ref(long-numbers)).
+We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a).
+We then independently denoised reads from each domain and concatenating the denoised domains for each read.
+Finally denoised reads were clustered based on 100% ITS2 identity, and a consensus ASV was calculated for each cluster.
 We used this method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), for all of the PacBio and Ion Torrent datasets.
@@
 An ML tree was produced using RAxML version 8.2.12 [@stamatakis2014] using the GTR+GAMMA model and rapid bootstrapping with the MRE_IGN stopping criterion.
-Sequences confidently assigned outside kingdom Fungi by at least five of the primary taxonomic identification methods were used to root the tree, and sequences outside the clade defined by confidently identified Fungi were removed (see below and Supplementary Methods \@ref(rooting)).
+Sequences confidently (i.e., by at least five of the primary taxonomic identification methods) assigned outside kingdom Fungi  were used to root the tree, and sequences outside the clade defined by confidently identified Fungi were removed (see below and Supplementary Methods \@ref(rooting)).
 
@@
 For full-length long-amplicon ASVs, the primary taxonomic assignments were refined with the new algorithm PHYLOTAX.
-The PHYLOTAX algorithm resolves conflictsm one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
+The PHYLOTAX algorithm resolves conflicts among one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
 It is available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
@@
 Spatiotemporal analyses were performed on the full recovered fungal community after removal of non-fungal sequences, and on the ECM community.
+Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
+All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
 
@@
   geom_vline(xintercept = 0) +
-  theme_pubclean() +
+  ggpubr::theme_pubclean() +
   theme(axis.title = element_text(color = "white"),

### 2021-02-18 writing/transect_supplement.Rmd
Additional changes from Martin

@@
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
-Assignments were considered reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
+Assignments were considered confident if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
 Two ASVs belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.

### 2021-02-18 writing/transect_paper.Rmd
move sampling and PCR methods back to main text

@@
 One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade and homogenized by manually breaking up clumps and mixing inside a plastic zipper bag.
-A subsample of approximately 250\ mg was stored in DNA-preservation buffer and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).
+For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade.
+Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
+A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in DNA-preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement) for preservation and extraction methods).
 
 ## DNA amplification and sequencing
+
 DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
@@
 Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
-See Supplementary Methods \@ref(DNA-supplement) for PCR and library preparation protocols.
+
+For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
+PCR amplification was performed in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
+150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
+The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
+Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
+
+For the long amplicon, both forward and reverse primers included indexes for combinatorial multiplexing (Supplementary File 2).
+PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
+Reaction conditions were 10\ min at 95°, 30\ cycles of 45\ s at 95°, 45\ s at 59°, and 90\ s at 72°, and finally 10\ min at 72°.
+Each reaction was performed in three technical replicates as for short amplicons.
+
+After pooling of technical replicates, amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
+An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
 

### 2021-02-18 writing/transect_supplement.Rmd
move sampling and PCR methods back to main text

@@
 
-## Sampling {#Sampling-supplement}
+## Sample preservation and extraction {#Sampling-supplement}
 
-Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
-At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
-One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, any coarse organic debris <!--Do you have any definition?
-Wider than 5 mm?
-MR-->was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with a sterilized knife blade.
-Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of field lysis/preservation buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
+Soil samples described in the main text were placed into separate 2.0\ mL microtubes containing 750\ mL of field lysis/preservation buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
 Due to the inavailability of refrigeration, extraction was initiated in the field according to the manufacturers instructions by lysis using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
@@
 
-## DNA extraction, amplification, and sequencing {#DNA-supplement}
-
-After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
+After returning to the laboratory, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
 Samples preserved using LifeGuard were first centrifuged at 10\ 000\ g for 1\ minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
 DNA was quantified fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
-
-For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
-150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
-The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
-Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
-
-For the long amplicon, both forward and reverse primers included indexes for combinatorical multiplexing (Supplementary File 2).
-PCR was performed as for the short amplicons, but with 500\ µM of each of the two primers.
-Reaction conditions were 10\ min at 95°, 30\ cycles of 45\ s at 95°, 45\ s at 59°, and 90\ s at 72°, and finally 10\ min at 72°.
-Each reaction was performed in three technical replicates as for short amplicons.
-
-After pooling of technical replicates, amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
-An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
 

### 2021-03-02 writing/transect_paper.Rmd
version submitted for second revision

@@
 
-Although ECM fungi form many well-known mushrooms (e.g., *Amanita*, *Cantharellus*, *Boletus*), some instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruit bodies.
+Although ECM fungi form many well-known mushrooms (e.g., *Amanita*, *Cantharellus*, *Boletus*), some instead produce inconspicuous (e.g., *Tomentella*) or no (e.g., *Cenococcum*) fruitbodies.
 Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by sampling of vegetative structures [@horton2001].
@@
 If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013] with public databases such as GenBank [@benson2013] and Unite [@nilsson2019a].
-However, this approach comes with some potential weaknesses, discussed below.
-<!--R5: Awk -->
+However, there is potential to improve this approach at several stages, including sequencing technology, amplicon choice, clustering, and taxonomic assignment.
 
@@
 This brings into question the comparability of studies conducted only a few years apart.
-<!-- R5: Seems like this section below could use some references to studies that have tested communities with multiple methods.
-There are a bunch of citations like this but none featured here.
-For example:
-Smith, Dylan P., and Kabir G. Peay. "Sequence depth, not PCR replication, improves ecological inference from next generation DNA sequencing." PloS one 9.2 (2014): e90234. -->
+Existing studies which sequenced the same environmental samples using different HTS technologies [e.g., @claesson2010;@smith2014;@divoll2018;@Kennedy2018;@speranskaya2018;@Tedersoo2018] have found that most differences in results seem to be attributable to differences in sequencing depth or different primer biases, rather than differences in the technologies themselves.
+Only a few of these studies have controlled for primer biases by using the same primer pairs in each technology [@claesson2010;@divoll2018;@speranskaya2018], and to our knowledge none have sequenced the same PCR products using multiple HTS technologies.
 
 ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina [@schoch2012],
-but this variablity means that they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
+but this variability means that they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
 <!-- R5: citation formatting-->
@@
 Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
-The Unite database is an attempt to include all publically available high-quality ITS sequences (800\ 000 as of release 8.0), originally limited to fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
+The Unite database is an attempt to include all publicly available high-quality ITS sequences (800\ 000 as of release 8.0), originally limited to fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
 The Ribosomal Data Project [RDP, @cole2014] hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
@@
 Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
+PacBio sequencing has also been shown to recover longer variants of variable-length regions such as ITS1 and ITS2, which are excluded by other technologies [@castano2020].
 However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].
@@
 and both non-phylogenetic and phylogenetic community distance measures.
-We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
-
-<!--R5: I realize that this paper is strongly methods-based and is focused on figuring out which approaches might work best for understanding the fungi.
-However, I think there are likely some underlying questions and hypotheses.
-This is the place where I would spend a few sentences exploring the hypotheses.
-For example, what do you you expect to find with the new tree-based and distance-based approaches?
-Do you expect to find more diversity in the long read technology because the primers are more general, etc.?-->
+We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse communities, and incorporating phylogenetic information into taxonomic assignments.
+We hypothesize that (a) PacBio sequencing of the short amplicon gives less bias against longer ITS2 amplicons than Illumina and Ion Torrent, both qualitatively (recovering amplicons missed by the others) and quantitatively (greater fraction of reads in longer amplicons);
+(b) our long amplicons (ITS1-LR5) recover a more complete view of the fungal community than our short amplicons (gITS7-ITS4), due to reduced length and primer biases;
+(c) these differences lead to differing results for ecological metrics, specifically OTU/ASV richness and turnover distance;
+and (d) incorporating LSU in the long amplicon allows for better taxonomic assignment.
 
@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in DNA-preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement) for preservation and extraction methods).
+A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement) for preservation and extraction methods).
 
@@
 We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a).
-We then independently denoised reads from each domain and concatenating the denoised domains for each read.
-Finally denoised reads were clustered based on 100% ITS2 identity, and a consensus ASV was calculated for each cluster.
-We used this method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), for all of the PacBio and Ion Torrent datasets.
+We then independently denoised reads from each domain and concatenated the denoised domains for each read.
+Finally denoised reads were clustered based on 100% ITS2 identity, and a full-length consensus ASV was calculated for each cluster.
+This method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), was used for all of the PacBio and Ion Torrent datasets.
 Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
@@
 
-For full-length long-amplicon ASVs, the primary taxonomic assignments were refined with the new algorithm PHYLOTAX.
-The PHYLOTAX algorithm resolves conflicts among one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
+For full-length long-amplicon ASVs, the primary taxonomic assignments were refined based on the ML phylogenetic tree generated above using the new algorithm PHYLOTAX.
+The PHYLOTAX algorithm resolves conflicts among one or more primary assignment methods using a supplied phylogenetic tree (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
 It is available in the new R package `phylotax` at https://github.com/brendanf/phylotax.
@@
 Spatiotemporal analyses were performed on the full recovered fungal community after removal of non-fungal sequences, and on the ECM community.
-Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
+Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database [as of 20 Feb, 2020\;@nguyen2016funguild] via the R package `FUNGuildR` (https://github.com/brendanf/FUNGuildR).
 All taxa which included "Ectomycorrhiza" in the guild assignment at any level of confidence were included.
@@
 Lines are the best-fit lines for an exponential decay to max model.
-The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0\ year timelag) or spatiotemporal (for the 1\ year time lag) distance.
+The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0\ year time lag) or spatiotemporal (for the 1\ year time lag) distance.
 
@@
 The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--`r round(filter(readd(variofit_table), metric == "W-UniFrac", guild == "All Fungi", timelag == 0)$conf.high, -1)`\ m, indicating little evidence of spatial structure.
-This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
+This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades [e.g., ECM lineages \textit{sensu} @Tedersoo2010] are more spatially constant.
 This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4--0.6) as compared to Bray-Curtis (0.8--1.0) in Figure \@ref(fig:spatio).
@@
 
-The choice of amplicon and sequencing technology did not affect the results of the spatial analysis, provided sufficient sequencing depth.
+Contrary to our hypothesis, we found that Illumina and Ion Torrent sequencing of real environmental samples resulted in neither qualitative nor quantitative bias against longer ITS2 amplicons, relative to PacBio.
+Furthermore, although we did find an increased ability to detect certain fungal groups using the more universal ITS1-LR5 primer pair, the choice of amplicon and sequencing technology did not affect the results of the spatial analysis, provided sufficient sequencing depth.
 Alpha diversity estimates were strongly correlated between methods, but somewhat inflated for Ion Torrent relative to the other technologies.

### 2021-03-09 writing/transect_paper.Rmd
rearrange figures so they appear in text ordering
reference S1, 2, 3 in main text

@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement) for preservation and extraction methods).
+A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement), as well as Figures \@ref(buffer-pcoa), \@ref(buffer-compare-long), and \@ref(buffer-compare-short), for preservation and extraction methods).
 
@@
 
-(ref:accum-ASV-s) ASV accumulation curves
-
-(ref:accum-ASV) Each curve represents rarefaction of a single sample or pooled pair of samples.
-Points at the end of each curve represent the actual read depth and observed ASV richness.
-Panel at left has enlarged scale to show PacBio more clearly.
-Vertical dashed line at 100 reads indicates rarefaction level for ASV richness comparisons.
-
-```{r accum-ASV, fig.cap="(ref:accum-ASV-s). (ref:accum-ASV)", fig.scap = "(ref:accum-ASV-s)"}
-readd(accum_plot_ASV)
-```
-
 (ref:alpha-compare-s) Comparison of ASV richness between sequencing technologies
@@
 grid::grid.draw(remove_empty_facets(readd(alpha_plot_ASV)))
+```
+
+(ref:accum-ASV-s) ASV accumulation curves
+
+(ref:accum-ASV) Each curve represents rarefaction of a single sample or pooled pair of samples.
+Points at the end of each curve represent the actual read depth and observed ASV richness.
+Panel at left has enlarged scale to show PacBio more clearly.
+Vertical dashed line at 100 reads indicates rarefaction level for ASV richness comparisons.
+
+```{r accum-ASV, fig.cap="(ref:accum-ASV-s). (ref:accum-ASV)", fig.scap = "(ref:accum-ASV-s)"}
+readd(accum_plot_ASV)
 ```

### 2021-03-12 writing/transect_paper.Rmd
text updates for final revision; still don't have Dryad accession number

--- 2021-03-02 writing/transect_paper.Rmd
@@
 Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement) for preservation and extraction methods).
+A subsample of approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \@ref(Sampling-supplement), as well as Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short), for preservation and extraction methods).
 
@@
 
- - Trimmed, demultiplexed sequencing reads have been deposited at the European Nucleotide Archive (ERA) under Project accession number PRJEB37385.
-Accession numbers are given in Supplementary Files 1 and 2.
- - Consensus ASV sequences will also be deposited at ENA prior to publication.
+ - Trimmed, demultiplexed sequencing reads are deposited at the European Nucleotide Archive (ENA) under Project accession number PRJEB37385. Accession numbers are given in Supplementary Files 1 and 2.
+ - Consensus ASV sequences from the PacBio and Illumina datasets are deposited at ENA under Project accession number PRJEN37385, accession numbers HG995461-HG996435 (ASVs detected in the long amplicon dataset) and FR984199-FR988025 (ASVs only detected in the short amplicon dataset).
+Because of concern that the Ion Torrent dataset may contain many ASVs with uncorrected sequencing errors, we elected not to submit ASVs which were only detected in the Ion Torrent dataset.
+However, these sequences, along with all other consensus ASVs, are available in Dryad repository XXXXXXXX [@furneauxdata2020].
  - Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication [@furneauxdata2020].
  - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
-These packages are currently being prepared for submission to CRAN/Bioconductor.
-If they are not accepted prior to final publication, snapshots will be archived at Dryad.
+Snapshots of the versions used in this paper are archived in Zenodo repository XXXXXX
  - FASTA-format files for the RDP fungal LSU training set, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
-The versions used in this paper will be archived at Dryad prior to publication.
- - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.
+The versions used in this paper are archived in Dryad repository XXXXXXX [@furneauxdata2020].
+ - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect, and archived in Zenodo repository XXXXXXX.
 

### 2021-03-12 writing/transect_paper.Rmd
Add Dryad doi

@@
 Because of concern that the Ion Torrent dataset may contain many ASVs with uncorrected sequencing errors, we elected not to submit ASVs which were only detected in the Ion Torrent dataset.
-However, these sequences, along with all other consensus ASVs, are available in Dryad repository XXXXXXXX [@furneauxdata2020].
- - Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication [@furneauxdata2020].
+However, these sequences, along with all other consensus ASVs, are archived in Dryad [@furneauxdata2021].
+ - Nucleotide alignment and ML tree are archived at Dryad [@furneauxdata2021].
  - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
-Snapshots of the versions used in this paper are archived in Zenodo repository XXXXXX
+Snapshots of the versions used in this paper are archived in Dryad [@furneauxdata2021]
  - FASTA-format files for the RDP fungal LSU training set, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
-The versions used in this paper are archived in Dryad repository XXXXXXX [@furneauxdata2020].
- - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect, and archived in Zenodo repository XXXXXXX.
+The versions used in this paper are archived in Dryad [@furneauxdata2021].
+ - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect, and archived in Dryad [@furneauxdata2021].
 

### 2021-03-16 writing/transect_paper_rev2.tex
output diff from second revision (to make sure there are no regressions from accepted version)

--- 2021-02-16 writing/transect_paper_rev1.tex
@@
 
-Although ECM fungi form many well-known mushrooms (e.g., \emph{Amanita}, \emph{Cantharellus}, \emph{Boletus}), some instead produce inconspicuous (e.g., \emph{Tomentella}) or no (e.g., \emph{Cenococcum}) fruit bodies.
+Although ECM fungi form many well-known mushrooms (e.g., \emph{Amanita}, \emph{Cantharellus}, \emph{Boletus}), some instead produce inconspicuous (e.g., \emph{Tomentella}) or no (e.g., \emph{Cenococcum}) fruitbodies.
 Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by sampling of vegetative structures \autocite{horton2001}.
@@
 which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS \autocite{schoch2012,Lindahl2013}.
-In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice \autocite{opik2010}, and variable regions of the rDNA large subunit (LSU) have also been used for barcoding \autocite{house2016,tedersoo2015}.
+In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice \autocite{opik2010}, and variable regions of the rDNA large subunit (LSU) have also been used for barcoding \autocites[e.g.,][]{kurtzman1998,tedersoo2015,house2016}.
 The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis \autocite{Lindahl2013}.
 If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST \autocite{altschul1990,Lindahl2013} with public databases such as GenBank \autocite{benson2013} and Unite \autocite{nilsson2019a}.
-However, this approach comes with some potential weaknesses, discussed below.
+However, there is potential to improve this approach at several stages, including sequencing technology, amplicon choice, clustering, and taxonomic assignment.
 
@@
 This brings into question the comparability of studies conducted only a few years apart.
+Existing studies which sequenced the same environmental samples using different HTS technologies \autocites[e.g.,][]{claesson2010,smith2014,divoll2018,Kennedy2018,speranskaya2018,Tedersoo2018} have found that most differences in results seem to be attributable to differences in sequencing depth or different primer biases, rather than differences in the technologies themselves.
+Only a few of these studies have controlled for primer biases by using the same primer pairs in each technology \autocite{claesson2010,divoll2018,speranskaya2018}, and to our knowledge none have sequenced the same PCR products using multiple HTS technologies.
 
 ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina \autocite{schoch2012},
-but this variablity means that they cannot be reliably aligned over the fungal kingdom \autocite{Lindahl2013,Tedersoo2018}.
+but this variability means that they cannot be reliably aligned over the fungal kingdom \autocite{Lindahl2013,Tedersoo2018}.
 Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa \autocite{ihrmark2012,tedersoo2015,palmer2018}.
@@
 
-Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
+Distance-based clustering conflates intra-species variation and sequencing error \autocite{nilsson2008,lindner2011}, and results are dataset-specific.
 In contrast, more recent denoising methods such as DADA2 \autocite{callahan2017}, Deblur \autocite{amir2017}, and UNOISE2 \autocite{edgar2016} utilize read quality information to control for sequencing error while preserving intra-species variation.
@@
 Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated \autocite{glassman2018,botnen2018}.
-ASVs have been suggested to be less dataset specific that cluster-based OTUs \autocite{callahan2017}.
+ASVs have been suggested to be less dataset specific than cluster-based OTUs \autocite{callahan2017}.
 Support for PacBio has recently been added to DADA2 \autocite{callahan2019}, but its application requires greater sequencing depth for longer reads, especially in high diversity samples.
@@
 In contrast, phylogenetic community distance measures \autocite{wong2016} are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
-Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree \autocite{matsen2010,berger2011}, but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
+Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree \autocite{munch2008a,munch2008,matsen2010,berger2011}, but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
 Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments \autocite{tedersoo2018}, or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases \autocite{fouquier2016}, but these approaches are only applicable to sequences of known taxonomic affiliation.
@@
 Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
-Although the public INSDC databases \autocite{karsch-mizrachi2018} are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations \autocite{nilsson2006,steinegger2020} as well as sequences of poor technical quality \autocite{nilsson2012}.
-Consequently taxonomic assignments based on these databases may be incorrect or inconsistent.
+Although the public INSDC databases \autocite{karsch-mizrachi2018} are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations \autocite{bidartondo2008,nilsson2006,steinegger2020} as well as sequences of poor technical quality \autocite{ashelford2006,nilsson2012}.
+Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent \autocite{nilsson2005}.
 Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
-The Unite database is a more curated attempt to include all publically available ITS sequences (800~000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes \autocite{nilsson2019a}, where efforts have been made to correct incorrect annotations and exclude low-quality sequences \autocite{abarenkov2018a}.
-The Ribosomal Data Project \autocite[RDP,][]{cole2014} hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
-the Warcup ITS database, containing 18~000 manually curated fungal ITS sequences \autocite{deshpande2016}, and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups \autocite{liu2012}.
+The Unite database is an attempt to include all publicly available high-quality ITS sequences (800~000 as of release 8.0), originally limited to fungi but now expanded to include all eukaryotes \autocite{nilsson2019a}, where efforts have been made to correct incorrect annotations and exclude low-quality sequences \autocite{abarenkov2018a}.
+The Ribosomal Data Project \autocite[RDP,][]{cole2014} hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
+the Warcup ITS training set, containing 18~000 manually curated fungal ITS sequences \autocite{deshpande2016}, and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups \autocite[RDP-LSU,][]{liu2012}.
 Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.
@@
 Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
+PacBio sequencing has also been shown to recover longer variants of variable-length regions such as ITS1 and ITS2, which are excluded by other technologies \autocite{castano2020}.
 However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number \autocite{Kennedy2018}.
@@
 Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
-Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
+Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool \autocite{legendre2012}.
 Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
@@
 
-We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
-with three different reference databases (Unite, Warcup, RDP),
+We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDPC, SINTAX, IDTAXA)
+with three different reference databases (Unite, Warcup, RDP-LSU),
 and both non-phylogenetic and phylogenetic community distance measures.
-We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.
+We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse communities, and incorporating phylogenetic information into taxonomic assignments.
+We hypothesize that (a) PacBio sequencing of the short amplicon gives less bias against longer ITS2 amplicons than Illumina and Ion Torrent, both qualitatively (recovering amplicons missed by the others) and quantitatively (greater fraction of reads in longer amplicons);
+(b) our long amplicons (ITS1-LR5) recover a more complete view of the fungal community than our short amplicons (gITS7-ITS4), due to reduced length and primer biases;
+(c) these differences lead to differing results for ecological metrics, specifically OTU/ASV richness and turnover distance;
+and (d) incorporating LSU in the long amplicon allows for better taxonomic assignment.
 
@@
 Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Caesalpinioideae).
+Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Fabaceae tr. Amherstieae).
 At each site, 25~soil samples were collected along a single 24~m linear transect at intervals of 1~m in May 2015.
 One third of the sample locations (3~m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with a sterilized knife blade and homogenized.
-A subsample of approximately 250~mg was preserved and returned to the laboratory for extraction (see Supplementary methods).
+For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with an ethanol sterilized knife blade.
+Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
+A subsample of approximately 250~mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \ref{Sampling-supplement} for preservation and extraction methods).
 
@@
 Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
-The short amplicon (approximately 300~bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 \autocite{ihrmark2012} as the forward primer and a mix of ITS4 \autocite{white1990amplification} and ITS4a \autocite{urbina2016} as the reverse primer.
+The short amplicon (approximately 300~bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 \autocite{ihrmark2012} as the forward primer and an equimolar mix of ITS4 \autocite{white1990amplification} and ITS4a \autocite{urbina2016} as the reverse primer.
 The long amplicon (approximately 1500~bp) targeted the full ITS region including the 5.8S rDNA and approximately 950~bp at the 5' end of the LSU, including the first three variable regions (Figure~\ref{fig:regions}a), using ITS1 \autocite{white1990amplification} as the forward primer and LR5 \autocite{vilgalys1990} as the reverse primer.
 Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of \emph{Agaricus bisporus}.
-See Supplementary methods for PCR and library preparation protocols.
+
+For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
+PCR amplification was performed in 20µl reactions containing 200~µM dNTP mix, 250~µM indexed gITS7 primer,
+150µM ITS4m, 2mM MgCl\textsubscript{2}, 0.1~U \emph{Taq} polymerase (Dream \emph{Taq}, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7~ng purified DNA in Dream \emph{Taq} buffer.
+The reaction conditions were 10~min at 95°, followed by 35~cycles of 60~s at 95°, 45~s at 56°, and 50~s at 72°, and finally 3~min at 72°.
+Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
+
+For the long amplicon, both forward and reverse primers included indexes for combinatorial multiplexing (Supplementary File 2).
+PCR was performed as for the short amplicons, but with 500~µM of each of the two primers.
+Reaction conditions were 10~min at 95°, 30~cycles of 45~s at 95°, 45~s at 59°, and 90~s at 72°, and finally 10~min at 72°.
+Each reaction was performed in three technical replicates as for short amplicons.
+
+After pooling of technical replicates, amplicons were purified using SPRI beads \autocite{vesterinen2016} and quantified fluorometrically using Quant-iT\textsuperscript{TM} PicoGreen\textsuperscript{TM} dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
+An aliquot of 100~ng of DNA from each sample (or the total PCR product if less than 100~ng) was pooled into two libraries each for long and short amplicons.
 
@@
 
-We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
-We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU \autocite{michot1984}, and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure~\ref{fig:regions}a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
-We used this method, implemented in the new R packages \texttt{LSUx} (\url{https://github.com/brendanf/LSUx}) and \texttt{tzara} (\url{https://github.com/brendanf/tzara}) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.
+We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity, which lead to most biological variants being represented only by a cluster of reads differing by a small number of unique sequencing errors (for calculations see Supplementary Methods \ref{long-numbers}).
+We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU \autocite{michot1984}, and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure~\ref{fig:regions}a).
+We then independently denoised reads from each domain and concatenated the denoised domains for each read.
+Finally denoised reads were clustered based on 100\% ITS2 identity, and a full-length consensus ASV was calculated for each cluster.
+This method, implemented in the new R packages \texttt{LSUx} (splitting reads into homologous regions; \url{https://github.com/brendanf/LSUx}) and \texttt{tzara} (reassembling regions and generating full-length consensus ASVs; \url{https://github.com/brendanf/tzara}) and detailed in Supplementary Methods \ref{denoising-supplement}, was used for all of the PacBio and Ion Torrent datasets.
+Because the \texttt{LSUx} plus \texttt{tzara} method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
+The ITS2 region was extracted from the ASVs using \texttt{LSUx} for comparison to the results from the other technologies.
+To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97\% similarity using VSEARCH v2.9.1 \autocite{rognes2016}.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/regions-1}
+{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/regions-1}
 
@@
 \textbf{(b)} Total number of DADA2 ASVs vs.~fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using \texttt{LSUx}.
-Shorter and more conserved regions yield a greater fraction of successfully mapped reads.
+Data from all samples were analyzed as a single pool.
+\emph{long}: entire long amplicon, including ITS1, 5.8S, ITS2, and partial LSU; \emph{32S} partial 32S precursor to LSU, including 5.8S, ITS2, and partial LSU; \emph{LSU}: section of LSU rDNA included in the long amplicon, from the 5' end to the LR5 primer site; \emph{ITS}: full ITS region, including ITS1, 5.8S, and ITS2.
+Color indicates median region length.
+Shorter and more conserved regions yielded a greater fraction of successfully mapped reads.
 At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs.}\label{fig:regions}
@@
 
-Because the \texttt{LSUx} plus \texttt{tzara} method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
-The ITS2 region was extracted from the ASVs using \texttt{LSUx} for comparison to the results from the other technologies.
-
-To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97\% similarity using VSEARCH v2.9.1 \autocite{rognes2016}.
-
-\hypertarget{phylogenetic-inference-and-taxonomy-assignment}{%
-\subsubsection{Phylogenetic inference and taxonomy assignment}\label{phylogenetic-inference-and-taxonomy-assignment}}
+\hypertarget{phylogenetic-inference-and-taxonomic-assignment}{%
+\subsubsection{Phylogenetic inference and taxonomic assignment}\label{phylogenetic-inference-and-taxonomic-assignment}}
 
@@
 An ML tree was produced using RAxML version 8.2.12 \autocite{stamatakis2014} using the GTR+GAMMA model and rapid bootstrapping with the MRE\_IGN stopping criterion.
-The tree was rooted outside kingdom Fungi and non-Fungi were removed based on taxonomic assignments (see below and Supplementary methods).
+Sequences confidently (i.e., by at least five of the primary taxonomic identification methods) assigned outside kingdom Fungi were used to root the tree, and sequences outside the clade defined by confidently identified Fungi were removed (see below and Supplementary Methods \ref{rooting}).
 
-Taxonomic annotations of the RDP LSU fungal training set version 11.5 \autocite{cole2014} and Warcup ITS training set \autocite{deshpande2016} were mapped to a uniform taxonomic classification system (see Supplementary Methods).
-Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
+Taxonomic annotations of the RDP-LSU training set version 11.5 \autocite{liu2012,cole2014} and Warcup ITS training set \autocite{deshpande2016} were mapped to a uniform taxonomic classification system (see Supplementary Methods \ref{reannotate}).
+Primary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
 For each region/reference combination, taxonomy was assigned using three popular algorithms:
-the RDP Naïve Bayesian Classifier \autocite[RDPC,][]{wang2007} as implemented in DADA2;
+the RDPC \autocite{wang2007} as implemented in DADA2;
 SINTAX \autocite{edgar2016a} as implemented in VSEARCH v2.9.1 \autocite{rognes2016};
 and IDTAXA \autocite{murali2018}.
-A relatively lax confidence threshhold of 50\% was used for all three algorithms.
-Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references \(\times\) three algorithms).
+A relatively lax confidence threshold of 50\% was used for all three algorithms, in order to increase the amount of input for consensus algorithms.
+Each full-length ASV was thus given up to nine primary taxonomic assignments (three references \(\times\) three algorithms).
 ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.
 
-For full-length long-amplicon ASVs, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure~\ref{fig:phylotax} and Supplementary Methods).
-This refinement algorithm is referred to as ``PHYLOTAX'', available in the new R package \texttt{phylotax} at \url{https://github.com/brendanf/phylotax}.
+For full-length long-amplicon ASVs, the primary taxonomic assignments were refined based on the ML phylogenetic tree generated above using the new algorithm PHYLOTAX.
+The PHYLOTAX algorithm resolves conflicts among one or more primary assignment methods using a supplied phylogenetic tree (see Figure~\ref{fig:phylotax} and Supplementary Methods \ref{phylotax}).
+It is available in the new R package \texttt{phylotax} at \url{https://github.com/brendanf/phylotax}.
 
-ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the ``last common ancestor'' of the preliminary assignments.
-This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits \autocite[LCAClassifier,][]{lanzen2012} or k-mer similarity scores \autocite[mothur's k-nearest neighbor method,][]{schloss2009}, but here is used to resolve conflicts between assignments from different algorithms and databases.
+ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different primary assignments at each rank, resulting in a consensus assignment equivalent to the ``last common ancestor'' of the primary assignments \autocite{huson2007}.
+This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits \autocites[e.g., MEGAN and LCAClassifier,][]{huson2016,lanzen2012} or k-mer similarity scores \autocite[mothur's k-nearest neighbor method,][]{schloss2009}, but here is used to resolve conflicts between assignments from different algorithms and databases.
 Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as ``Consensus''.
@@
 
-We compared alpha diversity estimates by the different sequencing strategies by calculating ASV and OTU accumulation curves, as well as comparing richness estimates after rarefaction for each sample (Supplementary methods).
-We also compared the effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, using PERMANOVA and heat tree visualizations (Supplementary methods).
+We compared alpha diversity estimates by the different sequencing strategies by calculating ASV and OTU accumulation curves, as well as comparing richness estimates after rarefaction for each sample (Supplementary Methods \ref{sequencing-strategy-supplement}).
+We also compared the effect of sequencing technology and amplicon length on the recovered fungal community composition (i.e., after removal of non-fungi), as assessed by the Bray-Curtis dissimilarity, using PERMANOVA and heat tree visualizations (Supplementary Methods \ref{sequencing-strategy-supplement}).
 
@@
 To estimate turnover scale, ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric \autocite[for both long and short amplicons]{bray1957} and the phylogenetically based weighted UniFrac metric \autocites[for only long amplicons]{lozupone2005,lozupone2007} in \texttt{phyloseq} version 1.26.0.
-Each of these distance matrices was used to calculate a Mantel correlogram with a 1~m bin size for distances in the range of 0--12~m, i.e., half the maximum separation present in the dataset.
+Dissimilarities were based on relative read abundance within each sample.
+Samples were not rarefied to a standard sequencing depth within datasets, as both the Bray-Curtis dissimilarity and the UniFrac metric are robust to unequal sampling depths \autocite{mcmurdie2014}.
+In addition, we did not standardize sequencing depth between datasets, because this would remove one of the potential benefits of Illumina and Ion Torrent relative to PacBio.
+
+Each of the distance matrices was used to calculate a Mantel correlogram with a 1~m bin size for distances in the range of 0--12~m, i.e., half the maximum separation present in the dataset.
 Separate correlograms were drawn for samples taken during the same year and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
-
-Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by \textcite{legendre2012} using the \texttt{nls} function in R (Supplementary methods).
+Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by \textcite{legendre2012} using the \texttt{nls} function in R (Supplementary Methods \ref{spatiotemporal-supplement}).
+Spatiotemporal analyses were performed on the full recovered fungal community after removal of non-fungal sequences, and on the ECM community.
+Sequences were assigned as ECM based on taxonomic assignments using the FUNGuild database \autocite[as of 20 Feb, 2020;][]{nguyen2016funguild} via the R package \texttt{FUNGuildR} (\url{https://github.com/brendanf/FUNGuildR}).
+All taxa which included ``Ectomycorrhiza'' in the guild assignment at any level of confidence were included.
 
@@
 Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100~reads (Figure \ref{fig:sample-depth}).
-Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
+Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, taxonomic assignment, and all non-spatial analyses.
+Spatial analyses were based on the remaining 34 samples for Illumina and 30 samples each for the PacBio Long and Short amplicons.
 
@@
 
-Among the different regions extracted from the long amplicon (Figure~\ref{fig:region-lengths},
+Among the different regions extracted from the long amplicon (Figure~\ref{fig:region-lengths}),
 ITS1 showed the greatest length variability (mean \(\pm\) standard deviation: 193 \(\pm\) 55~bp), followed by ITS2 (184 \(\pm\) 41~bp) and the variable regions in LSU (D2: 227 \(\pm\) 36~bp; D3: 108 \(\pm\) 10~bp; D1: 159 \(\pm\) 6~bp).
@@
 
-\hypertarget{reproducibility-of-sequence-capture-using-different-technologies}{%
-\subsection{Reproducibility of sequence capture using different technologies}\label{reproducibility-of-sequence-capture-using-different-technologies}}
+\hypertarget{reproducibility-of-sequence-detection-using-different-technologies}{%
+\subsection{Reproducibility of sequence detection using different technologies}\label{reproducibility-of-sequence-detection-using-different-technologies}}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/venn-1} \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/venn-2}
+{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/venn-1} \includegraphics[width=112mm]{transect_paper_files/figure-latex/venn-2}
 
@@
 In each region, the ASV/OTU richness is given above, while the relative abundance of reads represented by these ASVs/OTUs in each sequencing strategy are shown below in the order PacBio/Illumina/Ion Torrent (\emph{a}, \emph{b}), or long/short (\emph{c}, \emph{d}).
-For short amplicons in \emph{c} and \emph{d}, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.}\label{fig:venn}
+For short amplicons in \emph{c} and \emph{d}, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
+Analyses performed on pooled ASVs/OTUs from all samples.}\label{fig:venn}
 \end{figure}
@@
 We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
-The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables~\ref{tab:venn-table-asv} and \ref{tab:venn-table-otu}).
-The short amplicon ASVs shared between all sequencing technologies represented 95\%, 76\%, and 66\%\% of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure~\ref{fig:venn}a).
-When differences at the intra-species scale were removed by clustering the ASVs into 97\% OTUs, the number of OTUs shared between all three technologies increased to 524, representing 100\%, 93\% , and 89\% \% of reads, respectively (Figure~\ref{fig:venn}b).
+The majority of abundant ASVs and OTUs were detected by all sequencing strategies used (Tables~\ref{tab:venn-table-asv} and \ref{tab:venn-table-otu}).
+The short amplicon ASVs shared between all sequencing technologies represented 95\%, 76\%, and 66\% of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure~\ref{fig:venn}a).
+When differences at the intra-species scale were removed by clustering the ASVs into 97\% OTUs, the number of OTUs shared between all three technologies increased to 524, representing 100\%, 93\% , and 89\% of reads, respectively (Figure~\ref{fig:venn}b).
 In particular, the majority of the 9418 unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
@@
 In contrast, 21\% of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure~\ref{fig:venn}c), and the fraction only reduced to 20\% after clustering the ITS2 regions into OTUs (Figure~\ref{fig:venn}d).
+The taxonomic identity of these ASVs is discussed below.
 
@@
 Total least squares regression indicated that the long amplicon dataset resulted in richness estimates which were intermediate between the short amplicon results from Ion Torrent and the other two technologies.
-Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure~\ref{fig:venn}), sample-wise OTU accumulation curves (Figure~\ref{fig:accum-OTU}) and rarefied OTU richness relationships between sequencing strategies (Figure~\ref{fig:alpha-compare-otu}) were highly similar to those for ASVs.
+Despite the fact that experiment-wide OTU richness was lower than ASV richness (Figure~\ref{fig:venn}), OTU accumulation curves for each sample
+(Figure~\ref{fig:accum-OTU}) and rarefied OTU richness relationships between sequencing strategies (Figure~\ref{fig:alpha-compare-otu}) were highly similar to those for ASVs.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/accum-ASV-1}
+{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/accum-ASV-1}
 
@@
 
-\caption[ASV accumulation curves]{ASV accumulation curves.
-Each curve represents rarefaction of a single sample or pooled pair of samples.
+\caption[ASV accumulation curves]{ASV accumulation curves. Each curve represents rarefaction of a single sample or pooled pair of samples.
 Points at the end of each curve represent the actual read depth and observed ASV richness.
-Note logarithmic axes.}\label{fig:accum-ASV}
+Panel at left has enlarged scale to show PacBio more clearly.
+Vertical dashed line at 100 reads indicates rarefaction level for ASV richness comparisons.}\label{fig:accum-ASV}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/alpha-compare-1}
+{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/alpha-compare-1}
 
@@
 
-\caption[Comparison of ASV richness between sequencing technologies]{Comparison of ASV richness between sequencing technologies Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
+\caption[Comparison of ASV richness between sequencing technologies]{Comparison of ASV richness between sequencing technologies.
+Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
 All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.
@@
 Blue lines are total least squares fit with 95\% confidence interval, with the given slope (and 95\% confidence interval) and R\textsuperscript{2} value.
-Dashed diagonal line indicates ideal slope of 1.}\label{fig:alpha-compare}
+Dashed diagonal line indicates 1:1 line.}\label{fig:alpha-compare}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{transect_paper_files/figure-latex/taxon-chart-1}
+{\centering \includegraphics[width=169mm]{transect_paper_files/figure-latex/taxon-chart-1}
 
@@
 
-\caption[Summary of taxonomic assignments]{Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
+\caption[Summary of taxonomic assignments]{Summary of taxonomic assignments.
+Fraction of ASVs (left) and reads (right) assigned to each taxonomic rank, for different sequencing technologies (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), amplicons (Long, Short), reference databases (Unite, Warcup, RDP-LSU), and assignment algorithms (PHYLOTAX, Consensus, RDPC, SINTAX, IDTAXA).
 Consensus and PHYLOTAX assignments are based on the consensus of RDPC, SINTAX, and IDTAXA, using all available databases and, in the case of PHYLOTAX, phylogenetic information.}\label{fig:taxon-chart}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/fungi-comm-1}
+{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/fungi-comm-1}
 
@@
 Assignments based on PHYLOTAX.
-Classes which represented less than 2\% of reads and ASVs in all datasets are grouped together as ``other''.}\label{fig:fungi-comm}
+Classes that represented less than 2\% of reads and ASVs in all datasets are grouped together as ``other''.}\label{fig:fungi-comm}
 \end{figure}
@@
 
-Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at \(p < 0.05\) for ranges of up to 2--3~m for the total fungal community, and 1--2~m for the ECM fungal community (Figure~\ref{fig:correlog}).
+Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at \(p < 0.05\) for ranges of up to 2--3~m for the total recovered fungal community, and 1--2~m for the ECM fungal community (Figure~\ref{fig:correlog}).
 In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
@@
 
-The Bray-Curtis metric showed positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0~m, time lag of 1~year), for both the total fungal and ECM fungal communities, although this result did not reach statistical significance for all sequencing strategies.
+The Bray-Curtis metric showed positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0~m, time lag of 1~year), for both the total recovered fungal and ECM fungal communities, although this result did not reach statistical significance for all sequencing strategies.
 This spatiotemporal correlation did not extend to a range of 1~m, and in fact correlation was negative at a time lag of 1~year and distance of 1~m, indicating that samples collected 1~m apart in different years were more different than randomly selected pairs of samples.
@@
 
-In contrast to the Bray-Curtis distance, the weighted UniFrac distance showed very little spatial structure, with only the total fungal community in the 1~m
+In contrast to the Bray-Curtis distance, the weighted UniFrac distance showed very little spatial structure, with only the total recovered fungal community in the 1~m
 distance class showing a significant correlation at \(p < 0.05\).
@@
 
-The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from 13--31~m for the total fungal community and 12--42~m for the ECM fungal community (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}).
-However, there was overlap of the 95\% confidence intervals for all of the Bray-Curtis spatial ranges in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies, so no strong conclusion of variability between methods can be drawn.
-Although a distance-decay model was fit for the weighted UniFrac distance applied to the total fungal community, the result was very poorly constrained, and a range of 0~m, indicating no spatial structure, was included in the 95\% confidence interval.
+The best fit spatial turnover ranges based on Bray-Curtis distance-decay curves calculated from different sequencing strategies range widely from 13--31~m for the total recovered fungal community and 12--42~m for the ECM fungal community (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}).
+However, there was overlap of the 95\% confidence intervals for all of the Bray-Curtis spatial ranges in both the total recovered fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Table~\ref{tab:variofit}), so no strong conclusion of variability between methods can be drawn.
+Although a distance-decay model was fit for the weighted UniFrac distance applied to the total recovered fungal community, the result was very poorly constrained, and a range of 0~m, indicating no spatial structure, was included in the 95\% confidence interval (Table~\ref{tab:variofit}).
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/spatio-1}
+{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/spatio-1}
 
@@
 Lines are the best-fit lines for an exponential decay to max model.
-The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0~year timelag) or spatiotemporal (for the 1~year time lag) distance.}\label{fig:spatio}
+The model was only fit for datasets where the Mantel test indicated a significant relationship between community dissimilarity and spatial (for the 0~year time lag) or spatiotemporal (for the 1~year time lag) distance.}\label{fig:spatio}
 \end{figure}
@@
 OTU read counts were strongly correlated between technologies (R\textsuperscript{2} = 0.72--0.82),
-and even between primer pairs (R\textsuperscript{2} = 0.49--0.62, Figure \ref{fig:read-compare}b),
-which we interpret as supporting the validity of abundance-based beta diversity measures in metabarcoding \autocite[but see][ and references therein]{castano2020}.
+and even between primer pairs (R\textsuperscript{2} = 0.49--0.62, Figure \ref{fig:read-compare}b).
+This lends some support for the technical repeatability of abundance-based beta diversity measures in metabarcoding, although bias at the amplification stage still presents issues \autocite{polz1998,Kanagawa_2003,bellemain2010,castano2020}.
 
@@
 
-The long amplicon dataset included 20\% unique taxa, even after clustering at 97\% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers capture different parts of the community.
+The long amplicon dataset included 20\% unique taxa, even after clustering at 97\% ITS2 similarity, indicating that the differences in the communities recovered are not due to small sequencing errors, but rather that the different primers amplify different parts of the community.
 The ITS4 primer used in the short amplicon dataset has known mismatches to Tulasnellaceae and Alveolata, while gITS7 also has mismatches for Tulasnellaceae \autocite{tedersoo2015}.
@@
 The alternate LR5-F primer \autocite{tedersoo2008} would select against the non-target Alveolata, at the expense of also having mismatches for the Tulasnellaceae.
-We assert that, for studies targeting ECM fungi in particular, more complete capture of groups with high rDNA variability such as \emph{Tulasnella} (and ideally other Cantharellales) is worth the read-depth spent on non-target groups.
+We assert that, for studies targeting ECM fungi in particular, more complete detection of groups with high rDNA variability such as \emph{Tulasnella} (and ideally other Cantharellales) is worth the read-depth spent on non-target groups.
 
@@
 Assignment of ecological function to environmental fungal sequences is dependent on accurate taxonomic identification, especially at the genus level or below \autocite{nguyen2016funguild}.
-However, different combinations of algortihms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
-Although the RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \ref{fig:taxon-chart}).
+However, different combinations of algorithms and reference datasets vary in their performance at confidently assigning taxonomy to sequences.
+Although RDP-LSU and Unite performed comparably at taxonomic placement of long amplicon sequences, the Warcup database placed notably fewer sequences at all taxonomic levels for all datasets (Figure \ref{fig:taxon-chart}).
 This is probably due to two factors.
@@
 Second, due to its low-density coverage of the fungal kingdom (18~000 sequences vs.~800~000 for Unite), it is likely that many ITS sequences, especially from uncultured tropical soil fungi, have no close match in the Warcup database, and so cannot be placed.
-The RDP fungal training set, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
+RDP-LSU, which has even fewer sequences (8000 fungi plus 3000 other eukaryotes), is probably more successful due to higher sequence conservation in LSU.
 \textcite{heeger2019} also found that a more conserved region, in their case 5.8S, outperformed ITS at placing sequences without close database matches.
@@
 
-For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses phylogenetic relationships to resolve these disagreements in a principled manner.
+For ASVs where a long amplicon sequence is available, our novel PHYLOTAX algorithm uses relationships from a provided phylogenetic tree to resolve these disagreements.
 The effect was most pronounced for the PacBio long amplicon dataset,
 where 46\% and 62\% of reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to 73\% and 81\%.
-This led to a corresponding increase in the fraction of reads assigned to a functional guild (Figure~\ref{fig:ecm}).
+This led to a corresponding increase in the fraction of fungal reads assigned to a functional guild from 71\% to 91\% (Figure~\ref{fig:ecm}).
 For short amplicon sequencing strategies, the improvement was more modest, because PHYLOTAX could only be applied for ITS2 ASVs with a match to one of the long amplicon ASVs (last row of Table~\ref{tab:bioinfo}).
 Deeper long-amplicon sequencing would improve the coverage of long amplicons, allowing a greater fraction of short amplicon ASVs to also be placed phylogenetically.
+
+Because our dataset was generated from environmental samples whose true taxonomic affinity is unknown, we were not able to assess the accuracy of taxonomic assignments by any of the methods used here.
+Accuracy has been assessed using leave-one-out validation for the primary assignment algorithms \autocites[e.g.,][]{edgar2018,murali2018} and other consensus methods \autocites[e.g.,][]{gdanetz2017,somervuo2016}, and similar work could be carried out in the future for PHYLOTAX.
 
@@
 Distance-decay plots (Figure~\ref{fig:spatio}, Table~\ref{tab:variofit}) gave substantially longer autocorrelation distances.
-There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total fungal community and the ECM community, with best fit estimates ranging from 12--18~m.
+There was little variation in the results between the Illumina and long-amplicon PacBio datasets for both the total recovered fungal community and the ECM community, with best fit estimates ranging from 12--18~m.
 The 95\% confidence interval was substantially wider than this variation, generally covering a range of 5--41~m.
 All of these values are smaller than the 65~m reported by \textcite{Bahram2013}, also based on distance-decay curves from a similar ECM woodland habitat in Benin, but based on Sanger sequencing of ECM root tips rather than HTS metabarcoding of bulk soil.
-We speculate that this discrepancy is due to an increased ability to capture spatially variable rare species using HTS.
+We speculate that this discrepancy is due to an increased ability to detect spatially variable rare species using HTS.
 
@@
 
-Year-to-year correlation was found for both the total fungal and ECM communities in the long amplicon dataset (Figure \ref{fig:correlog}).
-The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.3~years for the total fungal community and 4.2~years for the ECM community, but with overlapping confidence intervals.
-This corresponds to a space-for-time substitution rate of 5.4 and 3.3~m/year for the total fungal community and ECM community, respecitvely.
+Year-to-year correlation was found for both the total recovered fungal and ECM communities in the long amplicon dataset (Figure \ref{fig:correlog}).
+The spatiotemporal distance-decay fit estimated the temporal turnover range as 3.3~years for the total recovered fungal community and 4.2~years for the ECM community, but with overlapping confidence intervals.
+This corresponds to a space-for-time substitution rate (i.e., ratio of a spatial distance to a time delay which results in equivalent community dissimilarity) of 5.4 and 3.3~m/year for the total recovered fungal community and ECM community, respecitvely.
 In a recent study, \textcite{kivlin2020} reported a space-for-time substitution rate of 81~m/year (reported as 6.8~d / 1.5~m) in the soil fungal community of a nonseasonal tropical forest in Costa Rica.
@@
 Year-to-year variation in ECM fungal communities, which we sampled, has been shown to be less than intra-annual variation \autocite{bahram2015}, as sampled by \textcite{kivlin2020}.
-Neither dataset from the short amplicon library showed significant temporal autocorrelatio.
+Neither dataset from the short amplicon library showed significant temporal autocorrelation.
 
@@
 The functional fit showed poor convergence, with a 95\% confidence interval for spatial range of 0--5470~m, indicating little evidence of spatial structure.
-This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages \emph{sensu} \textcite{Tedersoo2010}) are more spatially constant.
+This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades \autocite[e.g., ECM lineages \textit{sensu}][]{Tedersoo2010} are more spatially constant.
 This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac (0.4--0.6) as compared to Bray-Curtis (0.8--1.0) in Figure \ref{fig:spatio}.
@@
 \item
-  FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}.
+  FASTA-format files for the RDP fungal LSU training set, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}.
 The versions used in this paper will be archived at Dryad prior to publication.

### 2021-03-16 writing/transect_supplement_rev2.tex
output diff from second revision (to make sure there are no regressions from accepted version)

--- 2021-02-16 writing/transect_supplement_rev1.tex
@@
 % \DeclareDelayedFloatFlavour{sidewaysfigure}{figure}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \ifluatex
@@
 Thick branches represent rapid bootstrap values \textgreater90\%.
-Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, with Unite, Warcup, and RDP reference databases.
+Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, with Unite, Warcup, and RDP-LSU reference databases.
 Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
@@
 \hypertarget{Sampling-supplement}{%
-\subsection{Sampling}\label{Sampling-supplement}}
+\subsection{Sample preservation and extraction}\label{Sampling-supplement}}
 
-Sampling was conducted at two sites (Ang: N 9.75456° E 2.14064°; Gan: N 9.75678° E 2.31058°) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
-Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Caesalpinioideae).
-At each site, 25~soil samples were collected along a single 24~m linear transect at intervals of 1~m in May 2015.
-One third of the sample locations (3~m spacing) were resampled one year later in June 2016, for a total of 67 samples.
-For each sample, any coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with a sterilized knife blade.
-Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
-Approximately 250~mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0~mL microtube containing 750~mL of field lysis/preservation buffer and lysis beads (Xpedition\textsuperscript{TM} Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
+Soil samples described in the main text were placed into separate 2.0~mL microtubes containing 750~mL of field lysis/preservation buffer and lysis beads (Xpedition\textsuperscript{TM} Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA).
 Due to the inavailability of refrigeration, extraction was initiated in the field according to the manufacturers instructions by lysis using a handheld bead-beater (TerraLyser\textsuperscript{TM}; Zymo Research Corporation).
@@
 
-\hypertarget{DNA-supplement}{%
-\subsection{DNA extraction, amplification, and sequencing}\label{DNA-supplement}}
-
-After field lysis, DNA was extracted using the Xpedition\textsuperscript{TM} Soil/Fecal Prep kit (see above).
+After returning to the laboratory, DNA was extracted using the Xpedition\textsuperscript{TM} Soil/Fecal Prep kit (see above).
 Samples preserved using LifeGuard were first centrifuged at 10~000~g for 1~minute, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
 DNA was quantified fluorometrically using Quant-iT\textsuperscript{TM} PicoGreen\textsuperscript{TM} dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
-
-For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
-Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200~µM dNTP mix, 250~µM indexed gITS7 primer,
-150µM ITS4m, 2mM MgCl\textsubscript{2}, 0.1~U \emph{Taq} polymerase (Dream \emph{Taq}, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7~ng purified DNA in Dream \emph{Taq} buffer.
-The reaction conditions were 10~min at 95°, followed by 35~cycles of 60~s at 95°, 45~s at 56°, and 50~s at 72°, and finally 3~min at 72°.
-Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.
-
-For the long amplicon, both forward and reverse primers included indexes for combinatorical multiplexing (Supplementary File 2).
-PCR was performed as for the short amplicons, but with 500~µM of each of the two primers.
-Reaction conditions were 10~min at 95°, 30~cycles of 45~s at 95°, 45~s at 59°, and 90~s at 72°, and finally 10~min at 72°.
-Each reaction was performed in three technical replicates as for short amplicons.
-
-After pooling of technical replicates, amplicons were purified using SPRI beads \autocite{vesterinen2016} and quantified fluorometrically as above.
-An aliquot of 100~ng of DNA from each sample (or the total PCR product if less than 100~ng) was pooled into two libraries each for long and short amplicons.
 
@@
 No attempt was made to remove the approximately 12~bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx \autocite{bengtsson-palme2013}.
-A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 \autocite{glockner2017,cole2014} by truncating after the LR5 primer site and using the reference line to annotate the variable regions \emph{sensu} \textcite{michot1984} and \textcite{raue1988}.
+A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal LSU RNA seed alignment from RDP release 11.5 \autocite{glockner2017,cole2014} by truncating after the LR5 primer site and using the reference line to annotate the variable regions \emph{sensu} \textcite{michot1984} and \textcite{raue1988}.
 A CM was generated from the alignment using \texttt{cmbuild} from Infernal.
@@
 
-\hypertarget{uniform-taxonomic-classification-for-unite-warcup-and-rdp}{%
-\subsubsection{Uniform taxonomic classification for Unite, Warcup, and RDP}\label{uniform-taxonomic-classification-for-unite-warcup-and-rdp}}
+\hypertarget{reannotate}{%
+\subsubsection{Uniform taxonomic classification for Unite, Warcup, and RDP-LSU}\label{reannotate}}
 
@@
 Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists \autocite{adl2019}, which utilizes hierarchical nameless ranks.
-FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}.
+FASTA-format files for the RDP-LSU, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at \url{https://github.com/brendanf/reannotate}.
 
-\hypertarget{rooting-the-fungal-tree}{%
-\subsubsection{Rooting the fungal tree}\label{rooting-the-fungal-tree}}
+\hypertarget{rooting}{%
+\subsubsection{Rooting the fungal tree}\label{rooting}}
 
@@
 Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
-Assignments were considered reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
+Assignments were considered confident if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
 Two ASVs belonging to \emph{Tulasnella}, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies \autocite{moncalvo2006}, were excluded at this step.
@@
 
-\hypertarget{refining-taxonomic-assignments-with-a-phylogenetic-tree-phylotax}{%
-\subsubsection{\texorpdfstring{Refining taxonomic assignments with a phylogenetic tree: \texttt{PHYLOTAX}}{Refining taxonomic assignments with a phylogenetic tree: PHYLOTAX}}\label{refining-taxonomic-assignments-with-a-phylogenetic-tree-phylotax}}
+\hypertarget{phylotax}{%
+\subsubsection{\texorpdfstring{Refining taxonomic assignments with a phylogenetic tree: \texttt{PHYLOTAX}}{Refining taxonomic assignments with a phylogenetic tree: PHYLOTAX}}\label{phylotax}}
 
@@
 
-\caption[OTU accumulation curves]{OTU accumulation curves.
-Each curve represents rarefaction of a single sample or pooled pair of samples.
+\caption[97\% OTU accumulation curves]{97\% OTU accumulation curves. Each curve represents rarefaction of a single sample or pooled pair of samples.
 Points at the end of each curve represent the actual read depth and observed OTU richness.
-Note logarithmic axes.}\label{fig:accum-OTU}
+Panel at left has enlarged scale to show PacBio more clearly.
+Vertical dashed line at 100 reads indicates rarefaction level for OTU richness comparisons.}\label{fig:accum-OTU}
 \end{figure}
@@
 
-\caption[Comparison of OTU richness between sequencing technologies]{Comparison of OTU richness between sequencing technologies (ref:alpha-compare)}\label{fig:alpha-compare-otu}
+\caption[Comparison of OTU richness between sequencing technologies]{Comparison of OTU richness between sequencing technologies Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies.
+All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.
+Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction.
+Blue lines are total least squares fit with 95\% confidence interval, with the given slope (and 95\% confidence interval) and R\textsuperscript{2} value.
+Dashed diagonal line indicates ideal slope of 1.}\label{fig:alpha-compare-otu}
 \end{figure}
@@
 Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
-The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:taxa}
+The RDP-LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:taxa}
 \end{sidewaysfigure}
@@
 Height of each bar represents the fraction of reads represented by that bar.
-Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP; ``All'' denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.}\label{fig:ecm}
+Results are shown for different sequencing technologies, amplicons (Long, Short), reference databases (Unite, Warcup, RDP-LSU; ``All'' denotes algorithms which combine results from multiple databases), and taxonomic assignment algorithms.}\label{fig:ecm}
 \end{figure}
@@
 Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
-The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:ecm-heattree}
+The RDP-LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.}\label{fig:ecm-heattree}
 \end{figure}

### 2021-03-16 scripts/drake_plan_light.R

@@ -1777,7 +1777,7 @@ plan2 <- drake_plan(
     "writing",
     {
       file_in(!!file.path("writing", "transect_supplement_rev1.tex"))
-      file_in(!!file.path("writing", "transect_supplement.tex"))
+      file_in(!!file.path("writing", "transect_supplement_rev2.tex"))
       file_out(!!file.path("writing", "transect_supplement_diff2.tex"))
       tinytex::tlmgr_install(setdiff(c("latexdiff", "listings"), tinytex::tl_pkgs()))
       system2(
@@ -1788,7 +1788,7 @@ plan2 <- drake_plan(
           '"FLOATENV=sidewaysfigure"',
           "--graphics-markup=2",
           "transect_supplement_rev1.tex",
-          "transect_supplement.tex"
+          "transect_supplement_rev2.tex"
         ),
         stdout = TRUE
       ) %>%
@@ -1797,6 +1797,33 @@ plan2 <- drake_plan(
           "transect_paper_diff2"
         ) %>%
         writeLines("transect_supplement_diff2.tex")
+    }
+  ),
+
+  supplement_diff3_tex = withr::with_dir(
+    "writing",
+    {
+      file_in(!!file.path("writing", "transect_supplement_rev2.tex"))
+      file_in(!!file.path("writing", "transect_supplement.tex"))
+      file_out(!!file.path("writing", "transect_supplement_diff3.tex"))
+      tinytex::tlmgr_install(setdiff(c("latexdiff", "listings"), tinytex::tl_pkgs()))
+      system2(
+        command = "latexdiff",
+        args = list(
+          "--type=CFONT",
+          "--add-to-config",
+          '"FLOATENV=sidewaysfigure"',
+          "--graphics-markup=2",
+          "transect_supplement_rev2.tex",
+          "transect_supplement.tex"
+        ),
+        stdout = TRUE
+      ) %>%
+        stringr::str_replace_all(
+          "transect_paper",
+          "transect_paper_diff3"
+        ) %>%
+        writeLines("transect_supplement_diff3.tex")
     }
   ),
 
@@ -1828,7 +1855,7 @@ plan2 <- drake_plan(
     "writing",
     {
       file_in(!!file.path("writing", "transect_paper_rev1.tex"))
-      file_in(!!file.path("writing", "transect_paper.tex"))
+      file_in(!!file.path("writing", "transect_paper_rev2.tex"))
[…]

### 2021-03-17 writing/transect_paper.Rmd
add Dryad links

@@
 Finally denoised reads were clustered based on 100% ITS2 identity, and a full-length consensus ASV was calculated for each cluster.
-This method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), was used for all of the PacBio and Ion Torrent datasets.
+This method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement) and Table \@ref(tab:region-limits), was used for all of the PacBio and Ion Torrent datasets.
 Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
@@
 Because of concern that the Ion Torrent dataset may contain many ASVs with uncorrected sequencing errors, we elected not to submit ASVs which were only detected in the Ion Torrent dataset.
-However, these sequences, along with all other consensus ASVs, are archived in Dryad [@furneauxdata2021].
- - Nucleotide alignment and ML tree are archived at Dryad [@furneauxdata2021].
+However, these sequences, along with all other consensus ASVs, are archived in Dryad [@furneauxdata2021\; https://doi.org/10.5061/dryad.6wwpzgmvf].
+ - Nucleotide alignment and ML tree are archived at Dryad [@furneauxdata2021\; https://doi.org/10.5061/dryad.6wwpzgmvf].
  - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
-Snapshots of the versions used in this paper are archived in Dryad [@furneauxdata2021]
+Snapshots of the versions used in this paper are archived in Dryad [@furneauxdata2021\; https://doi.org/10.5061/dryad.6wwpzgmvf].
  - FASTA-format files for the RDP fungal LSU training set, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate.
-The versions used in this paper are archived in Dryad [@furneauxdata2021].
- - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect, and archived in Dryad [@furneauxdata2021].
+The versions used in this paper are archived in Dryad [@furneauxdata2021\; https://doi.org/10.5061/dryad.6wwpzgmvf].
+ - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect, and archived in Dryad [@furneauxdata2021\; https://doi.org/10.5061/dryad.6wwpzgmvf].
 
